### 2012-04-26 writedown/linear.response/main.tex
add writedown

\documentclass[aps,pre,preprint,unsortedaddress]{revtex4}
% \documentclass[aps,pre,twocolumn]{revtex4-1}
% \documentclass[aps,jcp,groupedaddress,twocolumn,unsortedaddress]{revtex4}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[dvips]{graphicx}
\usepackage{color}
\usepackage{tabularx}
\usepackage{algorithm}
\usepackage{algorithmic}

\newcommand{\recheck}[1]{{\color{red} #1}}
\newcommand{\redc}[1]{{\color{red} #1}}
\newcommand{\bluec}[1]{{\color{blue} #1}}
\renewcommand{\v}[1]{\textbf{\textit{#1}}}
\renewcommand{\d}[1]{\textsf{#1}}

\begin{document}

\title{Onsager regression method via Adaptive Resolution Simulation}

\begin{abstract}
We combine the Onsager regression method with the adaptive resolution simulation scheme.
\end{abstract}

\maketitle

\section{Theoretical Principles}
Let us suppose we have a system which is in equilibrium with a reservoir of energy and particles (i.e. Grand Canonical ensemble), at a certain time we apply a perturbation to the system. Such a perturbation will induce a response from the system. Let us identify $\hat{O}({\bf r,p})$ as the physical observable which expresses the response to the perturbation. If the system was characterized by a constant number of particle $\hat{N}$, then the average response of the system at time $t$ would be expressed by the equation:
\begin{equation}
O(t)=\int d{\bf r}d{\bf p} \hat{O}({\bf r,p})\rho_{\hat{N}}({\bf r,p};t)
\label{o1}
\end{equation}
or equivalently:
\begin{equation}
\int d{\bf r}d{\bf p} \hat{O}({\bf r,p})\rho_{\hat{N}}({\bf r,p};t)=\int d{\bf r}d{\bf p} \hat{O}({\bf r}(t),{\bf p}(t))\rho_{\hat{N}}({\bf r,p};0)
\label{o2}
\end{equation}
with:
\begin{equation}
\rho_{\hat{N}}({\bf r,p};0)=\rho_{\hat{N}}({\bf r,p}) lim_{T \to\infty}\frac{1}{T}\int_{0}^{T} d\tau \hat{O}\left({\bf r}(t; {\bf r}_{0}(\tau), {\bf p}_{0}(\tau)),{\bf p}(t;{\bf r}_{0}(\tau), {\bf p}_{0}(\tau)\right)
\label{rhon}
\end{equation} 
with $\rho_{\hat{N}}({\bf r,p})$ the standard canonical distribution at fixed $\hat{N}$. However in the present case, already before applying a perturbation we have a system whose total number of particles $N$ fluctuates, moreover when we apply the pertubation to such system the average response needs to be sampled over all possible realization (corresponding to each value of $N$), thus the formulas above are likely to be modified as:
\begin{equation}
O(t)=\left<\left<\hat{O}_{N}(t)|N\right>\right>_{N}=\sum_{N=0}^{\infty}P(N)\int d{\bf r}d{\bf p}\rho_{N}({\bf r,p})\hat{O}({\bf r}(t),{\bf p}(t))
\label{newvers}
\end{equation}
where:
\begin{equation}
P(N)=\frac{e^{\beta\mu N}Q_{N}}{Q}
\end{equation}
and
\begin{equation}
Q=\sum_{N}e^{\beta\mu N}Q_{N}
\end{equation}
with
\begin{equation}
Q_{N}=\int d{\bf r}d{\bf p} \frac{e^{-\beta H_{N}}}{N!h^{3N}}
\end{equation}
and
\begin{equation}
\rho_{N}({\bf r,p})=\frac{e^{-\beta H_{N}}}{N}.
\end{equation}
The essential meaning of Eq.\ref{newvers} is that we let the system evolve and calculate the average, counting the contribution of each $\hat{O}$ at each realization of $N$, weighted by the corresponding distribution for this realization $N$. So far this is the standard average in a Grand Canonical ensemble. The formulation above is nothing else than the extension of the Onsager-Kubo equation to the Grand Canonical case.
Let us discuss this in detail. Since our central quantities are $\rho_{N}$ , and in general all operators are of the form $X_{N}$ ($X$ generic operator), let us apply the procedure behind the Onsager-Kubo equation to quantities that are $N$ labeled, meaning that they depend on $N$. Liouville equation can then be written as:
\begin{equation}
\frac{\partial\rho_{N}}{\partial t}=i L_{N}\rho_{N}=iL_{N}^{0}\rho_{N}+iL_{N}^{P}\rho_{N}=\{H_{N}^{0},\rho_{N}\}+\{H^{P}_{N},\rho_{N}\}
\label{liouville}
\end{equation}
where $0$ and $p$ are indicating the unperturbed $L$ or $H$ operator and the perturbed one respectively.
For the distribution of the system we have:
\begin{equation}
\rho_{N}({\bf r},{\bf p},t)=S^{\dagger}_{N}(t)\rho_{N}({\bf r},{\bf p},0)
\end{equation}
% \bluec{ Noted by Han: This may be problematic, because I am not sure
%   if the probabilty density of an OPEN subsystem can be evolved by the
%   Liouville operator for a subsystem with FIXED number of particles
%   $N$.  Actually this equation assumes the number of particle in the
%   system is a constant, which may not be the case under perturbation.}
An observable $\hat{O}$ of the system evolves as:
\begin{equation}
\hat{O}_{N}(t)=\hat{O}({\bf r}(t),{\bf p}(t),N)=S_{N}(t)\hat{O}(0)
\end{equation}
The non equilibrium average of a given observable, $N$-labeled, i.e. $N$-dependent is obtained via:
\begin{equation}
\frac{d \hat{J}_{N}}{dt}=iL_{N}\hat{J}_{N};\hat{J}_{N}(t)=S_{N}(t)\hat{J}_{N}
\end{equation}
this could be calculated via an "equilibrium average as (Onsager-Kubo equation):
\begin{equation}
\hat{J}_{N}(t)=<\hat{J}>_{N}
\end{equation}
that is
\begin{equation}
<\hat{J}>_{N}=\int d\Gamma_{N}\hat{J}_{N}\rho_{N}(t)=(\hat{J}_{N},S^{\dagger}_{N}(t)\rho_{N}^{0})
\end{equation}
where $\Gamma_{N}$ is the phase space of the system, and finally:
\begin{equation}
 (\hat{J}_{N},S^{\dagger}_{N}(t)\rho_{N}^{0})=(S_{N}(t)\hat{J}_{N}, \rho_{N}^{0}).
\end{equation}
Now its average over the realizations $N$ follows as:
\begin{equation}
\left<\left<\hat{J}(t|N)\right>\right>_{N}=\sum_{N}p(N)\hat{J}_{N}(t)=\sum_{N}\int d{\bf r}d{\bf p}\hat{J}_{N}(t)\rho_{N}({\bf r},{\bf p}, t=0)
\end{equation}
where it must be noticed that $\rho_{N}({\bf r},{\bf p}, t=0)$ is the $\rho_{equil}({\bf r},{\bf p})$.
\bluec{Noted by Han:
  It is unclear for me that maybe $p(N)$ is also a function of time
  under perturbation. How to deal with it?}
This obvious extension is physically correct.
Let us see how to implement it. We have taken an initial condition with a given number of particles. Then, starting from this initial condition we allowed the system to freely evolve (with no constraints on $N$) and let the system go to the Grand Canonical equilibrium. Then we picked from time to time at regular intervals of time, initial conditions for the perturbed dynamics each with its $N$ kept fixed. After certain time we will have picked up a statistically significant realizations with different possible $N$ from the distribution $P(N)$.\\
The difference between the Grand Canonical approach and the Canonical approach is that:\\
(a)We need long non equilibrium trajectories to be sure that we sample enough segments with given $N$ along the non equilibrium trajectories\\
(b)A number of initial conditions with different $N$ large enough to be meaningful, that is, a dense branching along the Grand Canonical equilibrium trajectory.\\
In any case the advantage of this approach compared to an all atom simulation is the drastic reduction of computational cost of each simulation.

\section{An alternative consideration for the explicit (atomistic) reference system}
If we assume that the Observable we are interested can be expend by:
\begin{align}
  \hat O(\v r, \v p) =
  \sum_i\hat O^{(1)}(\v r_i, \v p_i) +
  \sum_{i\neq j}\hat O^{(2)}(\v r_i, \v p_i; \v r_j, \v p_j),
\end{align}
namely, we can explicitly writen down the single and pairwise
component of the observable (this is the case for most of the
properties, such as energy, pressure, density, RDF, etc.). We further
assume $\hat O^{(1)}$ and $\hat O^{(2)}$ are smooth up to the infinite
order (this may not be the case, but it is possible to polish these
observables and consider the limit of the parameter in the end. So the
smoothness of the observables is not essential in this argument).

Now assume that the grand-canonical system as a subsystem of an
``infinitely'' large canonical system. We denote the domain of the
subsystem by $\Omega$. $\v r_1, \cdots, \v r_N, \v p_1, \cdots, \v
p_N$ are all degrees of freedoms in the large canonical system. In the fully atomistic system, the
probability density is govened by a Liouville equation.  The
observable of the interested sub grand canonical can be written as a
function of the degrees of freedoms of the large canonical system,
revised by some proper screening function that shields the unconcerned
degrees of freedoms outside of the subsystem:
\begin{align}
  \hat O_{\textrm{sub-gc}}(\v r, \v p) =
  \sum_i
  \chi_\Omega(\v r_i)\,\hat O^{(1)}(\v r_i, \v p_i) +
  \sum_{i\neq j}
  \frac12 \,[\,
  \chi_\Omega(\v r_i) + \chi_\Omega(\v r_j)
  \,]\,
  \hat O^{(2)}(\v r_i, \v p_i; \v r_j, \v p_j),  
\end{align}
where $\chi_\Omega$ is defined by
\begin{align}
  \chi_\Omega(\v r)=
  \left\{
    \begin{array}{lcl}
      1 &\qquad\qquad& \v r \in \Omega; \\
      0 && \textrm{otherwise}.
    \end{array}
  \right.
\end{align}
Again, though the screening funciton $\chi_\Omega$ is not smooth, it
can be polished by some infinitly smooth function and consider the
limit of the polishing parameter in the end.  Following the argument
for the nonequilibrium canonical system:
% , we have the similar argument
% for the grand canonical subsystem:
\begin{align}\nonumber
  O_{\textrm{sub-gc}}(t)
  & = \langle \hat O_{\textrm{sub-gc}}(\v r, \v p)\rangle_t\\\nonumber
  & =
  \int
  \hat O_{\textrm{sub-gc}}(\v r, \v p) \rho(\v r, \v p, t)\,
  \d d\v r\d d\v p\\\nonumber
  & =
  \int
  \hat O_{\textrm{sub-gc}}(\v r, \v p) S(t)\rho(\v r, \v p)\,
  \d d\v r\d d\v p\\\nonumber
  & =
  \int
  S^{\dagger}(t)\hat O_{\textrm{sub-gc}}(\v r, \v p) \rho(\v r, \v p)\,
  \d d\v r\d d\v p\\
  & =
  \int
  \hat O_{\textrm{sub-gc}}(\v r(t), \v p(t)) \rho(\v r, \v p)\,
  \d d\v r\d d\v p
\end{align}
The $\rho$ is the standard canonical density distribution, which
includes all information of the grand canonical subsystem (especially
$p(N)$). The point is that we are able to study the observable of the
grand canonical subsystem (as a function of time), by the
non-equilibrium time evolution of the large canonical system, whose
governing function is well defined (do not need to consider the
exchange of particles, which is considered by $\rho(\v r, \v p, t)$
automatically).

\bibliography{ref}{}
\bibliographystyle{unsrt}

\end{document}

### 2012-04-26 writedown/linear.response/main.tex
init text

@@
-\documentclass[aps,pre,preprint,unsortedaddress]{revtex4}
+% \documentclass[aip,jcp,preprint,unsortedaddress,a4paper,onecolum]{revtex4-1}
+\documentclass[aip,jcp,a4paper,reprint,onecolumn]{revtex4-1}
 % \documentclass[aps,pre,twocolumn]{revtex4-1}
@@
 
-\usepackage{amsmath}
+\usepackage[fleqn]{amsmath}
 \usepackage{amssymb}
@@
 
+\makeatletter
+\makeatother
+
 \newcommand{\recheck}[1]{{\color{red} #1}}
@@
 \newcommand{\bluec}[1]{{\color{blue} #1}}
-\renewcommand{\v}[1]{\textbf{\textit{#1}}}
-\renewcommand{\d}[1]{\textsf{#1}}
+\newcommand{\vect}[1]{\textbf{\textit{#1}}}
+\newcommand{\dd}{\textsf{d}}
+
+\newcommand{\mh}{\mathcal H}
+\newcommand{\ml}{\mathcal L}
+\newcommand{\mt}{\mathcal T}
+\newcommand{\EX}{{\textrm{EX}}}
+\newcommand{\CG}{{\textrm{CG}}}
+\newcommand{\HY}{{\Delta}}
 
@@
 
-\title{Onsager regression method via Adaptive Resolution Simulation}
+\title{Linear response theory for core set identification}
+\author{Han Wang}
+% \affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Berlin, Germany}
+\author{Christof Sch\"utte}
+\affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Berlin, Germany}
+% \affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Berlin, Germany}
 
 \begin{abstract}
-We combine the Onsager regression method with the adaptive resolution simulation scheme.
+  % In this draft, we are trying to study the core set based Markov
+  % State Model (MSM) under non-equilirbium conditions.
 \end{abstract}
@@
 
-\section{Theoretical Principles}
-Let us suppose we have a system which is in equilibrium with a reservoir of energy and particles (i.e.
-Grand Canonical ensemble), at a certain time we apply a perturbation to the system.
-Such a perturbation will induce a response from the system.
-Let us identify $\hat{O}({\bf r,p})$ as the physical observable which expresses the response to the perturbation.
-If the system was characterized by a constant number of particle $\hat{N}$, then the average response of the system at time $t$ would be expressed by the equation:
-\begin{equation}
-O(t)=\int d{\bf r}d{\bf p} \hat{O}({\bf r,p})\rho_{\hat{N}}({\bf r,p};t)
-\label{o1}
-\end{equation}
+\section{Theoretical considerations}
+\subsection{The linear response theory}
+In this sub-section, we give a brief desciption of the linear response
+theory, for details see Ref.~\cite{tuckeman2010statistical} for
+example. Here we assume ths system is goverened by equation:
+\begin{align}\label{eqn:gov-1}
+  \dot{\vect q}_i & = \frac{\partial\mh}{\partial \vect p_i}
+  + \vect C_i(\vect q,\vect p) F_e(t)\\\label{eqn:gov-2}
+  \dot{\vect p}_i & =- \frac{\partial\mh}{\partial \vect q_i}
+  + \vect D_i(\vect q,\vect p) F_e(t)
+\end{align}
+Where $\vect C$ and $\vect D$ are perturbations to the system, and
+$F_e(t)$ is the strength of the perturbation as a function of time.
+The phase space imcompressibility condition is assumed:
+\begin{align}
+  \sum_{i=1}^{3N}
+  \bigg[
+  \frac{\partial \vect C_i}{\partial \vect q_i} +
+  \frac{\partial \vect D_i}{\partial \vect p_i}
+  \bigg] = 0
+\end{align}
+Then the phase space distribution $f(\vect x, t)$ ($\vect x = \{\vect
+q_i, \vect p_i\}$) is subject to the Liouville equation:
+\begin{align}
+  \frac{\partial}{\partial t} f(\vect x, t) + i\ml f(\vect x, t) = 0
+\end{align}
+where $\ml$ is the Liouville operator given by $i\ml = \dot{\vect
+  x}\cdot\nabla_{\vect x}$.  We assume that the perturbation to the
+system is so small that both the phase space distribution and the
+Liouville operator can be viewed as perturbation from the quilibirum
+distribution (invariant measure) and the unperturbed operator,
+respectively. Namely
+\begin{align}
+  f(\vect x, t) &= f_0(\vect x) + \Delta  f(\vect x, t)
+\end{align}
+and since the Eqn.~\eqref{eqn:gov-1} and \eqref{eqn:gov-2} imply that
+$\dot{\vect x} = \dot{\vect x}_0 + \Delta \dot{\vect x}(t)$
+\begin{align}
+  i\ml =
+  (\dot{\vect x}_0 + \Delta \dot{\vect x}(t))\cdot\nabla_{\vect x}
+  = i\ml_0 + i\Delta\ml(t)
+\end{align}
+Where the invariant measure satisfies $i\ml_0f_0(\vect x) = 0$. Also
+$f_0(\vect x)$ has the form $f_0(\mh(\vect x))$ (?). The Liouville equation
+for the perturbed system is therefore:
+\begin{align}
+  \frac{\partial}{\partial t}
+  [f_0(\mh(\vect x)) + \Delta  f(\vect x, t)]
+  +
+  [i\ml_0 + i\Delta\ml(t)]
+  [f_0(\mh(\vect x)) + \Delta  f(\vect x, t)] = 0
+\end{align}
+The basic assumption for the ``linear response'' is that the high order
+term $i\Delta\ml(t)  \Delta  f(\vect x, t)$ in the Liouville equation
+can be neglected. Noticing $i\ml_0 f_0 = 0$, we have the Liouville equation
+for the leading order of perturbation:
+\begin{align}
+  \bigg[
+  \frac{\partial}{\partial t}
+  + i\ml_0
+  \bigg]
+  \Delta  f(\vect x, t)
+  =
+  - i\Delta\ml(t) f_0(\mh(\vect x))
+\end{align}
+The operator on l.h.s. is the same as the unperturbed Liouville equation.
+Formally solving this equation with initial condition
+$\Delta f(\vect x, 0) = 0$ gives
+\begin{align}
+  \Delta  f(\vect x, t) =
+  - \int_0^t \dd s \:
+  e^{-i\ml_0(t-s)}
+  i\Delta\ml(s)f_0(\mh(\vect x))
+\end{align}
+By defining the \emph{dissipative flux},
+\begin{align}
+  j(\vect x) =
+  -\sum_{i=1}^{3N}
+  \bigg[
+  \vect D_i(\vect x) \frac{\partial\mh}{\partial \vect p_i} +
+  \vect C_i(\vect x) \frac{\partial\mh}{\partial \vect q_i}   
+  \bigg]
+\end{align}
+it has been shown that~\cite{tuckeman2010statistical}
+\begin{align}
+  i\Delta\ml(s)f_0(\mh(\vect x)) =
+  -\frac{f_0(\mh)}{\partial\mh} j(\vect x)\,F_e(s)
+\end{align}
+For the canonical invariant measure, this is equivalent to
+\begin{align}
+  i\Delta\ml(s)f_0(\mh(\vect x)) =
+  \beta f_0(\mh(\vect x)) j(\vect x)F_e(s)
+\end{align}
+For any observation:
+\begin{align}\nonumber
+  \mathcal A(t)
+  &=
+  \int\dd \vect x A(\vect x)f(\vect x, t)  \\\nonumber
+  &=
+  \int\dd \vect x\, A(\vect x)\,
+  [f_0(\mh(\vect x)) + \Delta f(\vect x, t)] \\\nonumber
+  &=
+  \langle A\rangle_0
+  -
+  \beta
+  \int \dd \vect x
+  A(\vect x)
+  \int_0^t\dd s\,
+  e^{-i\ml_0(t-s)}
+  f_0(\mh(\vect x)) j(\vect x)F_e(s)  \\\nonumber
+  &=
+  \langle A\rangle_0
+  -
+  \beta
+  \int_0^t\dd s\,
+  F_e(s)
+  \int \dd \vect x\,
+  A(\vect x)
+  e^{-i\ml_0(t-s)}
+  f_0(\mh(\vect x)) j(\vect x)  \\
+  &=
+  \langle A\rangle_0
+  -
+  \beta
+  \int_0^t\dd s\,
+  F_e(s)
+  \int \dd \vect x\,
+  f_0(\mh(\vect x))
+  A(\vect x)
+  e^{-i\ml_0(t-s)}
+  j(\vect x)  
+\end{align}
+Notice that $e^{i\ml_0(t-s)}A(\vect x) = A(\phi_{t-s}(\vect x))$,
+where $\phi_t(\vect x)$ is the flow function of the \emph{unperturbed}
+Hamiltonian system. We have
+\begin{align}
+  \mathcal A(t)
+  =
+  \langle A\rangle_0
+  -
+  \beta
+  \int_0^t\dd s\,
+  F_e(s)
+  \int \dd \vect x\,
+  f_0(\mh(\vect x))
+  A(\phi_{t-s}(\vect x))
+  j(\vect x)  
+\end{align}
 or equivalently:
-\begin{equation}
-\int d{\bf r}d{\bf p} \hat{O}({\bf r,p})\rho_{\hat{N}}({\bf r,p};t)=\int d{\bf r}d{\bf p} \hat{O}({\bf r}(t),{\bf p}(t))\rho_{\hat{N}}({\bf r,p};0)
-\label{o2}
-\end{equation}
-with:
-\begin{equation}
-\rho_{\hat{N}}({\bf r,p};0)=\rho_{\hat{N}}({\bf r,p}) lim_{T \to\infty}\frac{1}{T}\int_{0}^{T} d\tau \hat{O}\left({\bf r}(t; {\bf r}_{0}(\tau), {\bf p}_{0}(\tau)),{\bf p}(t;{\bf r}_{0}(\tau), {\bf p}_{0}(\tau)\right)
-\label{rhon}
-\end{equation} 
-with $\rho_{\hat{N}}({\bf r,p})$ the standard canonical distribution at fixed $\hat{N}$.
-However in the present case, already before applying a perturbation we have a system whose total number of particles $N$ fluctuates, moreover when we apply the pertubation to such system the average response needs to be sampled over all possible realization (corresponding to each value of $N$), thus the formulas above are likely to be modified as:
-\begin{equation}
-O(t)=\left<\left<\hat{O}_{N}(t)|N\right>\right>_{N}=\sum_{N=0}^{\infty}P(N)\int d{\bf r}d{\bf p}\rho_{N}({\bf r,p})\hat{O}({\bf r}(t),{\bf p}(t))
-\label{newvers}
-\end{equation}
-where:
-\begin{equation}
-P(N)=\frac{e^{\beta\mu N}Q_{N}}{Q}
-\end{equation}
-and
-\begin{equation}
-Q=\sum_{N}e^{\beta\mu N}Q_{N}
-\end{equation}
-with
-\begin{equation}
-Q_{N}=\int d{\bf r}d{\bf p} \frac{e^{-\beta H_{N}}}{N!h^{3N}}
-\end{equation}
-and
-\begin{equation}
-\rho_{N}({\bf r,p})=\frac{e^{-\beta H_{N}}}{N}.
-\end{equation}
-The essential meaning of Eq.\ref{newvers} is that we let the system evolve and calculate the average, counting the contribution of each $\hat{O}$ at each realization of $N$, weighted by the corresponding distribution for this realization $N$.
-So far this is the standard average in a Grand Canonical ensemble.
-The formulation above is nothing else than the extension of the Onsager-Kubo equation to the Grand Canonical case.
-Let us discuss this in detail.
-Since our central quantities are $\rho_{N}$ , and in general all operators are of the form $X_{N}$ ($X$ generic operator), let us apply the procedure behind the Onsager-Kubo equation to quantities that are $N$ labeled, meaning that they depend on $N$.
-Liouville equation can then be written as:
-\begin{equation}
-\frac{\partial\rho_{N}}{\partial t}=i L_{N}\rho_{N}=iL_{N}^{0}\rho_{N}+iL_{N}^{P}\rho_{N}=\{H_{N}^{0},\rho_{N}\}+\{H^{P}_{N},\rho_{N}\}
-\label{liouville}
-\end{equation}
-where $0$ and $p$ are indicating the unperturbed $L$ or $H$ operator and the perturbed one respectively.
-For the distribution of the system we have:
-\begin{equation}
-\rho_{N}({\bf r},{\bf p},t)=S^{\dagger}_{N}(t)\rho_{N}({\bf r},{\bf p},0)
-\end{equation}
-% \bluec{ Noted by Han: This may be problematic, because I am not sure
-%   if the probabilty density of an OPEN subsystem can be evolved by the
-%   Liouville operator for a subsystem with FIXED number of particles
-%   $N$.  Actually this equation assumes the number of particle in the
-%   system is a constant, which may not be the case under perturbation.}
-An observable $\hat{O}$ of the system evolves as:
-\begin{equation}
-\hat{O}_{N}(t)=\hat{O}({\bf r}(t),{\bf p}(t),N)=S_{N}(t)\hat{O}(0)
-\end{equation}
-The non equilibrium average of a given observable, $N$-labeled, i.e. $N$-dependent is obtained via:
-\begin{equation}
-\frac{d \hat{J}_{N}}{dt}=iL_{N}\hat{J}_{N};\hat{J}_{N}(t)=S_{N}(t)\hat{J}_{N}
-\end{equation}
-this could be calculated via an "equilibrium average as (Onsager-Kubo equation):
-\begin{equation}
-\hat{J}_{N}(t)=<\hat{J}>_{N}
-\end{equation}
-that is
-\begin{equation}
-<\hat{J}>_{N}=\int d\Gamma_{N}\hat{J}_{N}\rho_{N}(t)=(\hat{J}_{N},S^{\dagger}_{N}(t)\rho_{N}^{0})
-\end{equation}
-where $\Gamma_{N}$ is the phase space of the system, and finally:
-\begin{equation}
- (\hat{J}_{N},S^{\dagger}_{N}(t)\rho_{N}^{0})=(S_{N}(t)\hat{J}_{N}, \rho_{N}^{0}).
-\end{equation}
-Now its average over the realizations $N$ follows as:
-\begin{equation}
-\left<\left<\hat{J}(t|N)\right>\right>_{N}=\sum_{N}p(N)\hat{J}_{N}(t)=\sum_{N}\int d{\bf r}d{\bf p}\hat{J}_{N}(t)\rho_{N}({\bf r},{\bf p}, t=0)
-\end{equation}
-where it must be noticed that $\rho_{N}({\bf r},{\bf p}, t=0)$ is the $\rho_{equil}({\bf r},{\bf p})$.
-\bluec{Noted by Han:
-  It is unclear for me that maybe $p(N)$ is also a function of time
-  under perturbation. How to deal with it?}
-This obvious extension is physically correct.
-Let us see how to implement it.
-We have taken an initial condition with a given number of particles.
-Then, starting from this initial condition we allowed the system to freely evolve (with no constraints on $N$) and let the system go to the Grand Canonical equilibrium.
-Then we picked from time to time at regular intervals of time, initial conditions for the perturbed dynamics each with its $N$ kept fixed.
-After certain time we will have picked up a statistically significant realizations with different possible $N$ from the distribution $P(N)$.\\
-The difference between the Grand Canonical approach and the Canonical approach is that:\\
-(a)We need long non equilibrium trajectories to be sure that we sample enough segments with given $N$ along the non equilibrium trajectories\\
-(b)A number of initial conditions with different $N$ large enough to be meaningful, that is, a dense branching along the Grand Canonical equilibrium trajectory.\\
-In any case the advantage of this approach compared to an all atom simulation is the drastic reduction of computational cost of each simulation.
+\begin{align}
+  \mathcal A(t)
+  =
+  \langle A\rangle_0
+  -
+  \beta
+  \int_0^t\dd s\,
+  F_e(s)
+  \langle
+  A(t-s)\,
+  j(0)\,
+  \rangle_0
+\end{align}
+where $\langle A(t-s)\,j(0)\,\rangle_0$ is the \emph{equilibrium time
+  correlation} between the observation $A$ at time $t-s$ and the
+dissipative flux $j$ at time 0.
 
-\section{An alternative consideration for the explicit (atomistic) reference system}
-If we assume that the Observable we are interested can be expend by:
+\subsection{The Core set as an observation of the system}
+
+The core set plays a very important role in the MSM theory. One way to
+identify core set is to consider the ``stability'' of the invariant
+measure $f_0$ under some mapping $\mt_\alpha$ of time step $\alpha$:
 \begin{align}
-  \hat O(\v r, \v p) =
-  \sum_i\hat O^{(1)}(\v r_i, \v p_i) +
-  \sum_{i\neq j}\hat O^{(2)}(\v r_i, \v p_i; \v r_j, \v p_j),
+  C_\alpha = \big\{
+  \vect x \,\vert\, \mt_\alpha f_0(\vect x) > f_0(\vect x)
+  \big\}
 \end{align}
-namely, we can explicitly writen down the single and pairwise
-component of the observable (this is the case for most of the
-properties, such as energy, pressure, density, RDF, etc.). We further
-assume $\hat O^{(1)}$ and $\hat O^{(2)}$ are smooth up to the infinite
-order (this may not be the case, but it is possible to polish these
-observables and consider the limit of the parameter in the end. So the
-smoothness of the observables is not essential in this argument).
+Generally, it is impossible and not necessary to consider the core set
+in the $3N$ dimensional configuration space. For example, a
+macromolecule solving in water, the DOFs of water is not of interest
+at all. Therefore, we consider some phase space compressing function
+$\vect y = \vect y(\vect x)$, with $\vect y \in \Omega \subset \mathbb
+R^M,\ M\ll 3N$. We call $\Omega$ the observation space.  The projected
+invariant measure on the observation space is therefore
+\begin{align}
+  \mathcal P f_0 (\vect y)
+  =
+  \int
+  \dd \vect x\,
+  f_0(\vect x) \,
+  \delta (\vect y - \vect y(\vect x)) 
+\end{align}
+The core set on the observation space is
+\begin{align}
+  \widetilde
+  C_\alpha = \big\{
+  \vect y \,\vert\,
+  \mathcal P\mt_\alpha f_0(\vect y) > \mathcal P f_0(\vect y)
+  \big\}
+\end{align}
+% with $\widetilde \mt_\alpha$ defined by
+% $\widetilde\mt_\alpha = \mathcal P\circ\mt_\alpha$, namely
+% \begin{align}
+%   \widetilde\mt_\alpha\, p_0(\vect y)
+%   =
+%   \mathcal P\circ\mt_\alpha f_0 (\vect y)
+%   =
+%   \int
+%   \dd \vect x\,
+%   [\mt_\alpha f_0](\vect x) \,
+%   \delta (\vect y - \vect y(\vect x)) 
+% \end{align}
+For convenience, we define $\Delta \mt_\alpha =
+\mt_\alpha - \mathcal I$, where $\mathcal I$ is the
+identity. If $[\mathcal P\circ\Delta \mt_\alpha f_0](\vect y) > 0$, then
+$\vect y$ is in the core set $\widetilde C_\alpha $.
+\begin{align}\nonumber
+  [\mathcal P\circ\Delta \mt_\alpha f_0](\vect y)
+  &=
+  \int
+  \dd \vect x\,
+  \Delta \mt_\alpha f_0(\vect x) \,
+  \delta (\vect y - \vect y(\vect x))  \\\nonumber
+  &=
+  \lim_{\sigma \rightarrow 0}
+  \int
+  \dd \vect x\,
+  \Delta \mt_\alpha f_0(\vect x) \,
+  K_{\sigma} (\vect y - \vect y(\vect x))  \\\nonumber
+  &=
+  \lim_{\sigma \rightarrow 0}
+  \int
+  \dd \vect x\,
+  \Delta \mt_\alpha f_0(\vect x) \,
+  K_{\sigma, \vect y} (\vect x)  \\\nonumber
+  &=
+  \lim_{\sigma \rightarrow 0}
+  \int
+  \dd \vect x\,
+  f_0(\vect x) \,
+  [\Delta \mt_\alpha]^\dagger K_{\sigma, \vect y} (\vect x)\\
+  &=
+  \lim_{\sigma \rightarrow 0}\,
+  \langle
+  [\Delta \mt_\alpha]^\dagger K_{\sigma, \vect y}
+  \rangle_0
+\end{align}
+where $[\Delta \mt_\alpha]^\dagger$ is the adjoint operator of $\Delta
+\mt_\alpha$, which is actually $\Delta \mt_{-\alpha}$. The delta
+function is replaced by the limit of a mollifier $K_\sigma(\vect y)$,
+which is inifinely smooth, compact supported, bounded and $\int
+K_\sigma(\vect y)\dd \vect y = 1$.  The physical meaning is that we
+can consider the expectation value of the stability of the molifier
+$K_{\sigma, \vect y}$ instead of the stability of the invariant measure.
+This provides the covenience of using the linear response theory showed
+in the previous section:
+\begin{align}\nonumber
+  [\mathcal P\circ\Delta \mt_\alpha f_t](\vect y)
+  &=
+  \lim_{\sigma \rightarrow 0}\,
+  \langle
+  [\Delta \mt_\alpha]^\dagger K_{\sigma, \vect y}
+  \rangle_t \\
+  &=
+  \lim_{\sigma \rightarrow 0}\,
+  \Big\{
+  \langle
+  [\Delta \mt_\alpha]^\dagger K_{\sigma, \vect y}
+  \rangle_0  -
+  \beta
+  \int_0^t\dd s\,
+  F_e(s)
+  \langle
+  [\Delta \mt_\alpha]^\dagger K_{\sigma, \vect y}(t-s)\,
+  j(0)\,
+  \rangle_0
+  \Big\}
+\end{align}
 
-Now assume that the grand-canonical system as a subsystem of an
-``infinitely'' large canonical system. We denote the domain of the
-subsystem by $\Omega$. $\v r_1, \cdots, \v r_N, \v p_1, \cdots, \v
-p_N$ are all degrees of freedoms in the large canonical system. In the fully atomistic system, the
-probability density is govened by a Liouville equation.  The
-observable of the interested sub grand canonical can be written as a
-function of the degrees of freedoms of the large canonical system,
-revised by some proper screening function that shields the unconcerned
-degrees of freedoms outside of the subsystem:
-\begin{align}
-  \hat O_{\textrm{sub-gc}}(\v r, \v p) =
-  \sum_i
-  \chi_\Omega(\v r_i)\,\hat O^{(1)}(\v r_i, \v p_i) +
-  \sum_{i\neq j}
-  \frac12 \,[\,
-  \chi_\Omega(\v r_i) + \chi_\Omega(\v r_j)
-  \,]\,
-  \hat O^{(2)}(\v r_i, \v p_i; \v r_j, \v p_j),  
-\end{align}
-where $\chi_\Omega$ is defined by
-\begin{align}
-  \chi_\Omega(\v r)=
-  \left\{
-    \begin{array}{lcl}
-      1 &\qquad\qquad& \v r \in \Omega; \\
-      0 && \textrm{otherwise}.
-    \end{array}
-  \right.
-\end{align}
-Again, though the screening funciton $\chi_\Omega$ is not smooth, it
-can be polished by some infinitly smooth function and consider the
-limit of the polishing parameter in the end.  Following the argument
-for the nonequilibrium canonical system:
-% , we have the similar argument
-% for the grand canonical subsystem:
-\begin{align}\nonumber
-  O_{\textrm{sub-gc}}(t)
-  & = \langle \hat O_{\textrm{sub-gc}}(\v r, \v p)\rangle_t\\\nonumber
-  & =
-  \int
-  \hat O_{\textrm{sub-gc}}(\v r, \v p) \rho(\v r, \v p, t)\,
-  \d d\v r\d d\v p\\\nonumber
-  & =
-  \int
-  \hat O_{\textrm{sub-gc}}(\v r, \v p) S(t)\rho(\v r, \v p)\,
-  \d d\v r\d d\v p\\\nonumber
-  & =
-  \int
-  S^{\dagger}(t)\hat O_{\textrm{sub-gc}}(\v r, \v p) \rho(\v r, \v p)\,
-  \d d\v r\d d\v p\\
-  & =
-  \int
-  \hat O_{\textrm{sub-gc}}(\v r(t), \v p(t)) \rho(\v r, \v p)\,
-  \d d\v r\d d\v p
-\end{align}
-The $\rho$ is the standard canonical density distribution, which
-includes all information of the grand canonical subsystem (especially
-$p(N)$). The point is that we are able to study the observable of the
-grand canonical subsystem (as a function of time), by the
-non-equilibrium time evolution of the large canonical system, whose
-governing function is well defined (do not need to consider the
-exchange of particles, which is considered by $\rho(\v r, \v p, t)$
-automatically).
-
+\section*{References}
 \bibliography{ref}{}

### 2012-04-26 writedown/linear.response/main.tex
fix bugs

@@
 \subsection{The linear response theory}
-In this sub-section, we give a brief desciption of the linear response
+In this sub-section, we give a brief description of the linear response
 theory, for details see Ref.~\cite{tuckeman2010statistical} for
-example. Here we assume ths system is goverened by equation:
+example. Here we assume the system is governed by equation:
 \begin{align}\label{eqn:gov-1}
@@
 system is so small that both the phase space distribution and the
-Liouville operator can be viewed as perturbation from the quilibirum
+Liouville operator can be viewed as perturbation from the equilibrium
 distribution (invariant measure) and the unperturbed operator,
@@
 function is replaced by the limit of a mollifier $K_\sigma(\vect y)$,
-which is inifinely smooth, compact supported, bounded and $\int
+which is infinitely smooth, compact supported, bounded and $\int
 K_\sigma(\vect y)\dd \vect y = 1$.  The physical meaning is that we
-can consider the expectation value of the stability of the molifier
+can consider the expectation value of the stability of the mollifier
 $K_{\sigma, \vect y}$ instead of the stability of the invariant measure.
-This provides the covenience of using the linear response theory showed
+This provides the convenience of using the linear response theory showed
 in the previous section:

### 2012-05-15 writedown/linear.response/main.tex
improve

@@
 \author{Han Wang}
-% \affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Berlin, Germany}
+% \affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}
 \author{Christof Sch\"utte}
-\affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Berlin, Germany}
-% \affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Berlin, Germany}
+\affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}
+% \affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}
+
+\date{\today}
 
@@
   i\Delta\ml(s)f_0(\mh(\vect x)) =
-  -\frac{f_0(\mh)}{\partial\mh} j(\vect x)\,F_e(s)
+  -\frac{\partial f_0(\mh)}{\partial\mh} j(\vect x)\,F_e(s)
 \end{align}
@@
 Notice that $e^{i\ml_0(t-s)}A(\vect x) = A(\phi_{t-s}(\vect x))$,
-where $\phi_t(\vect x)$ is the flow function of the \emph{unperturbed}
+where $\phi_t(\vect x)$ is the flow mapping of the \emph{unperturbed}
 Hamiltonian system. We have
@@
 or equivalently:
-\begin{align}
+\begin{align}\label{eqn:lr}
   \mathcal A(t)
@@
 function is replaced by the limit of a mollifier $K_\sigma(\vect y)$,
-which is infinitely smooth, compact supported, bounded and $\int
+which is infinitely smooth, compact supported of size $\sigma$,
+bounded and $\int
 K_\sigma(\vect y)\dd \vect y = 1$.  The physical meaning is that we
@@
 This provides the convenience of using the linear response theory showed
-in the previous section:
+in the previous section. By Eqn.~\eqref{eqn:lr},
 \begin{align}\nonumber
@@
 \end{align}
+Therefore, the core set at time $t$ is defined by
+\begin{align}
+  \widetilde C_\alpha^t =
+  \Big\{
+  \vect y
+  \,\Big\vert\,
+  [\mathcal P\circ\Delta \mt_\alpha f_t](\vect y) > 0
+  \Big\}
+\end{align}
 

### 2012-05-15 writedown/linear.response/langevin.tex
add file for langevin dynamics

--- 2012-05-15 writedown/linear.response/main.tex
@@
 \newcommand{\mt}{\mathcal T}
+\newcommand{\fwg}{{\mathcal A}}
+\newcommand{\bwg}{{\mathcal B}}
 \newcommand{\EX}{{\textrm{EX}}}
@@
 theory, for details see Ref.~\cite{tuckeman2010statistical} for
-example. Here we assume the system is governed by equation:
+example. Here we assume the system is governed by the
+perturbed Langevin equation:
 \begin{align}\label{eqn:gov-1}
-  \dot{\vect q}_i & = \frac{\partial\mh}{\partial \vect p_i}
-  + \vect C_i(\vect q,\vect p) F_e(t)\\\label{eqn:gov-2}
-  \dot{\vect p}_i & =- \frac{\partial\mh}{\partial \vect q_i}
-  + \vect D_i(\vect q,\vect p) F_e(t)
+  \dot{\vect q} & = \frac{\partial\mh}{\partial \vect p}
+  + \vect C(\vect q,\vect p) F_e(t)\\\label{eqn:gov-2}
+  \dot{\vect p} & =- \frac{\partial\mh}{\partial \vect q}
+  + \vect D(\vect q,\vect p) F_e(t)
+  - \gamma\vect p
+  + \sigma\dot{\vect W}
 \end{align}
@@
 \begin{align}
-  \sum_{i=1}^{3N}
-  \bigg[
-  \frac{\partial \vect C_i}{\partial \vect q_i} +
-  \frac{\partial \vect D_i}{\partial \vect p_i}
-  \bigg] = 0
+  \nabla_{\vect q}\cdot\vect C + \nabla_{\vect p}\cdot\vect D = 0
 \end{align}
 Then the phase space distribution $f(\vect x, t)$ ($\vect x = \{\vect
-q_i, \vect p_i\}$) is subject to the Liouville equation:
+q, \vect p\}$) is subject to the Fokker-Planck equation:
 \begin{align}
-  \frac{\partial}{\partial t} f(\vect x, t) + i\ml f(\vect x, t) = 0
+  \frac{\partial}{\partial t} f(\vect x, t) - \fwg f(\vect x, t) = 0
 \end{align}
-where $\ml$ is the Liouville operator given by $i\ml = \dot{\vect
-  x}\cdot\nabla_{\vect x}$.  We assume that the perturbation to the
+where $\fwg$ is the infiniesimal generator gieven by
+\begin{align}
+  \fwg =
+  \frac{\sigma^2}2\Delta_{\vect p}
+  - \Big(
+  \frac{\partial \mh}{\partial\vect p} + \vect C F_e(t)
+  \Big)\cdot\nabla_{\vect q}
+  - \Big(
+  -\frac{\partial \mh}{\partial\vect q} +
+  \vect D F_e(t) - \gamma\vect p
+  \Big)\cdot\nabla_{\vect p}
+  + 3N\gamma
+\end{align}
+The equilibrium and unperturbed system
+is governed by
+the standard Langevin equation, i.e. with
+vanishing input perturbation $F_e(t) = 0$ in
+Eqn.~\eqref{eqn:gov-1} and \eqref{eqn:gov-2}.
+We assume that the perturbation to the
 system is so small that both the phase space distribution and the
-Liouville operator can be viewed as perturbation from the equilibrium
+infiniesimal generator can be viewed as perturbation from the equilibrium
 distribution (invariant measure) and the unperturbed operator,
-respectively. Namely
+respectively. We have
 \begin{align}
@@
 \end{align}
-and since the Eqn.~\eqref{eqn:gov-1} and \eqref{eqn:gov-2} imply that
-$\dot{\vect x} = \dot{\vect x}_0 + \Delta \dot{\vect x}(t)$
+and 
 \begin{align}
-  i\ml =
-  (\dot{\vect x}_0 + \Delta \dot{\vect x}(t))\cdot\nabla_{\vect x}
-  = i\ml_0 + i\Delta\ml(t)
+  \fwg = \fwg_0 + \Delta\fwg(t)
 \end{align}
-Where the invariant measure satisfies $i\ml_0f_0(\vect x) = 0$. Also
-$f_0(\vect x)$ has the form $f_0(\mh(\vect x))$ (?). The Liouville equation
-for the perturbed system is therefore:
+Where the invariant measure satisfies $\fwg_0f_0(\vect x) = 0$,
+which is nothing but the Boltzmann distribution
+\begin{align}
+  f_0(\vect x) \propto e^{-\beta\mh(\vect x)}
+\end{align}
+The Fokker-Planck equation for the perturbation is therefore
 \begin{align}
   \frac{\partial}{\partial t}
-  [f_0(\mh(\vect x)) + \Delta  f(\vect x, t)]
-  +
-  [i\ml_0 + i\Delta\ml(t)]
-  [f_0(\mh(\vect x)) + \Delta  f(\vect x, t)] = 0
+  [f_0(\vect x) + \Delta  f(\vect x, t)]
+  -
+  [\,\fwg_0 + \Delta\fwg(t)\,]
+  [\,f_0(\vect x) + \Delta  f(\vect x, t)\,] = 0
 \end{align}
 The basic assumption for the ``linear response'' is that the high order
-term $i\Delta\ml(t)  \Delta  f(\vect x, t)$ in the Liouville equation
-can be neglected. Noticing $i\ml_0 f_0 = 0$, we have the Liouville equation
-for the leading order of perturbation:
+term $\Delta\fwg(t)  \Delta  f(\vect x, t)$ in the Fokker-Planck equation
+can be neglected. Noticing $\fwg_0 f_0(\vect x) = 0$, we have the
+Fokker-Planck
+for the leading order of the perturbation:
 \begin{align}
@@
   \frac{\partial}{\partial t}
-  + i\ml_0
+  - \fwg_0
   \bigg]
@@
   =
-  - i\Delta\ml(t) f_0(\mh(\vect x))
+  \Delta\fwg(t) f_0(\vect x)
 \end{align}
-The operator on l.h.s. is the same as the unperturbed Liouville equation.
+The operator on l.h.s. is the same as the unperturbed Fokker-Planck equation.
 Formally solving this equation with initial condition
@@
   \Delta  f(\vect x, t) =
-  - \int_0^t \dd s \:
-  e^{-i\ml_0(t-s)}
-  i\Delta\ml(s)f_0(\mh(\vect x))
+  \int_0^t \dd s \:
+  e^{(t-s)\fwg_0}
+  \Delta\fwg(s)f_0(\vect x)
 \end{align}
@@
   j(\vect x) =
-  -\sum_{i=1}^{3N}
-  \bigg[
-  \vect D_i(\vect x) \frac{\partial\mh}{\partial \vect p_i} +
-  \vect C_i(\vect x) \frac{\partial\mh}{\partial \vect q_i}   
-  \bigg]
+  -\vect C\cdot\nabla_{\vect q}\mh 
+  -\vect D\cdot\nabla_{\vect p}\mh
 \end{align}
@@
 \begin{align}
-  i\Delta\ml(s)f_0(\mh(\vect x)) =
-  -\frac{\partial f_0(\mh)}{\partial\mh} j(\vect x)\,F_e(s)
-\end{align}
-For the canonical invariant measure, this is equivalent to
-\begin{align}
-  i\Delta\ml(s)f_0(\mh(\vect x)) =
-  \beta f_0(\mh(\vect x)) j(\vect x)F_e(s)
+  \Delta\fwg(s)f_0(\vect x) =
+  -\beta f_0(\vect x)\, j(\vect x) F_e(s)
 \end{align}
@@
 \begin{align}\nonumber
-  \mathcal A(t)
+  \mathcal O(t)
   &=
-  \int\dd \vect x A(\vect x)f(\vect x, t)  \\\nonumber
+  \int\dd \vect x \:O(\vect x)f(\vect x, t)  \\\nonumber
   &=
-  \int\dd \vect x\, A(\vect x)\,
-  [f_0(\mh(\vect x)) + \Delta f(\vect x, t)] \\\nonumber
+  \int\dd \vect x\, O(\vect x)\,
+  [f_0(\vect x) + \Delta f(\vect x, t)] \\\nonumber
   &=
-  \langle A\rangle_0
+  \langle O\rangle_0
   -
   \beta
-  \int \dd \vect x
-  A(\vect x)
+  \int \dd \vect x\:
+  O(\vect x)
   \int_0^t\dd s\,
-  e^{-i\ml_0(t-s)}
-  f_0(\mh(\vect x)) j(\vect x)F_e(s)  \\\nonumber
+  e^{(t-s)\fwg_0}
+  f_0(\vect x)\, j(\vect x)F_e(s)  \\\nonumber
   &=
-  \langle A\rangle_0
+  \langle O\rangle_0
   -
@@
   \int \dd \vect x\,
-  A(\vect x)
-  e^{-i\ml_0(t-s)}
-  f_0(\mh(\vect x)) j(\vect x)  \\
+  O(\vect x)\,
+  e^{(t-s)\fwg_0}
+  f_0(\vect x) j(\vect x)  \\
   &=
-  \langle A\rangle_0
+  \langle O\rangle_0
   -
@@
   \int \dd \vect x\,
-  f_0(\mh(\vect x))
-  A(\vect x)
-  e^{-i\ml_0(t-s)}
-  j(\vect x)  
-\end{align}
-Notice that $e^{i\ml_0(t-s)}A(\vect x) = A(\phi_{t-s}(\vect x))$,
-where $\phi_t(\vect x)$ is the flow mapping of the \emph{unperturbed}
-Hamiltonian system. We have
-\begin{align}
-  \mathcal A(t)
-  =
-  \langle A\rangle_0
+  f_0(\vect x) j(\vect x)\,
+  e^{(t-s)\bwg_0}
+  O(\vect x)\\
+  &=
+  \langle O\rangle_0
   -
@@
   \int_0^t\dd s\,
-  F_e(s)
+  F_e(t-s)
   \int \dd \vect x\,
-  f_0(\mh(\vect x))
-  A(\phi_{t-s}(\vect x))
-  j(\vect x)  
+  f_0(\vect x) j(\vect x)\,
+  e^{s\bwg_0}
+  O(\vect x)
+\end{align}
+Where $\bwg_0$ is the infiniesimal generator of the \emph{unperturbed} backward
+Kolmogorov equation, which is the adjoint operator of $\fwg_0$.
+We have
+\begin{align}
+  e^{s\bwg_0} O(\vect x) = \mathbb E_{\vect x} [O(\vect X_s)]
+\end{align}
+where $\vect X_t$ is the trajectory of the unperturbed Langevin equation
+starting at $\vect x$.
+Therefore, we have
+% Notice that $e^{i\ml_0(t-s)}O(\vect x) = O(\phi_{t-s}(\vect x))$,
+% where $\phi_t(\vect x)$ is the flow mapping of the \emph{unperturbed}
+% Hamiltonian system. We have
+\begin{align}
+  \mathcal O(t)
+  =
+  \langle O\rangle_0
+  -
+  \beta
+  \int_0^t\dd s\,
+  F_e(t-s)
+  \int \dd \vect x\,
+  f_0(\vect x)
+  j(\vect x)\,
+  \mathbb E_{\vect x} [O(\vect X_s)]
 \end{align}
@@
 \begin{align}\label{eqn:lr}
-  \mathcal A(t)
+  \mathcal O(t)
   =
-  \langle A\rangle_0
+  \langle O\rangle_0
   -
@@
   \int_0^t\dd s\,
-  F_e(s)
+  F_e(t-s)
   \langle
-  A(t-s)\,
   j(0)\,
+  \mathbb E_{\vect x} [O(\vect X_s)]
   \rangle_0
 \end{align}
-where $\langle A(t-s)\,j(0)\,\rangle_0$ is the \emph{equilibrium time
-  correlation} between the observation $A$ at time $t-s$ and the
+where $  \langle
+  j(0)\,
+  \mathbb E_{\vect x} [O(\vect X_s)]
+  \rangle_0$ is the \emph{equilibrium time
+  correlation} between the observation $O$ at time $s$ and the
 dissipative flux $j$ at time 0.
@@
   \int_0^t\dd s\,
-  F_e(s)
+  F_e(t-s)
   \langle
-  [\Delta \mt_\alpha]^\dagger K_{\sigma, \vect y}(t-s)\,
   j(0)\,
+  \mathbb E_{\vect x}\{
+  [\Delta \mt_\alpha]^\dagger K_{\sigma, \vect y}(s)\,
+  \}
   \rangle_0

### 2012-05-21 writedown/linear.response/langevin.tex
small revise
fix bugs in text

@@
 Then the phase space distribution $f(\vect x, t)$ ($\vect x = \{\vect
-q, \vect p\}$) is subject to the Fokker-Planck equation:
+q, \vect p\}$) is subject to the Kolmogorov forward equation:
 \begin{align}
@@
 \end{align}
-where $\fwg$ is the infiniesimal generator gieven by
+where $\fwg$ is the forward infiniesimal generator gieven by
 \begin{align}
@@
 infiniesimal generator can be viewed as perturbation from the equilibrium
-distribution (invariant measure) and the unperturbed operator,
+distribution (invariant measure) and the unperturbed operator
+($\vect C = \vect D = 0$),
 respectively. We have
@@
 \end{align}
-The Fokker-Planck equation for the perturbation is therefore
+The Kolmogorov forward equation for the perturbation is therefore
 \begin{align}
@@
 The basic assumption for the ``linear response'' is that the high order
-term $\Delta\fwg(t)  \Delta  f(\vect x, t)$ in the Fokker-Planck equation
+term $\Delta\fwg(t)  \Delta  f(\vect x, t)$ 
 can be neglected. Noticing $\fwg_0 f_0(\vect x) = 0$, we have the
-Fokker-Planck
+Kolmogorov forward equation
 for the leading order of the perturbation:
@@
 \end{align}
-The operator on l.h.s. is the same as the unperturbed Fokker-Planck equation.
+The operator on l.h.s. is the same as the unperturbed equation.
 Formally solving this equation with initial condition
@@
 \end{align}
-it has been shown that~\cite{tuckeman2010statistical}
+it has been shown that %~\cite{tuckeman2010statistical}
 \begin{align}
@@
   e^{(t-s)\fwg_0}
-  f_0(\vect x) j(\vect x)  \\
+  f_0(\vect x) j(\vect x)  \\\nonumber
   &=
@@
 \end{align}
-Where $\bwg_0$ is the infiniesimal generator of the \emph{unperturbed} backward
-Kolmogorov equation, which is the adjoint operator of $\fwg_0$.
+Where $\bwg_0$ is the infiniesimal generator of the \emph{unperturbed} 
+Kolmogorov backward equation, which is the adjoint operator of $\fwg_0$.
 We have
@@
 
-\subsection{The Core set as an observation of the system}
+\subsection{The core set as an observation of the system}
 

### 2012-05-21 one.test/template/tools/simulator/app/main.cpp
implement Euler-Maruyama integrator

@@ -1,0 +1,21 @@
+#include <algorithm>
+#include <fstream>
+#include <iomanip>
+#include <iostream>
+#include <iterator>
+#include <list>
+#include <map>
+#include <set>
+#include <sstream>
+#include <string>
+#include <queue>
+#include <vector>
+#include <cmath>
+
+#include "Force.h"
+#include "Integrator.h"
+
+int main(int argc, char * argv[])
+{
+  
+}

### 2012-05-21 one.test/template/tools/simulator/include/Defines.h

@@ -1,0 +1,37 @@
+#ifndef __Defines_h_NONEQMSM_wanghan__
+#define __Defines_h_NONEQMSM_wanghan__
+
+#include <algorithm>
+#include <fstream>
+#include <iomanip>
+#include <iostream>
+#include <iterator>
+#include <list>
+#include <map>
+#include <set>
+#include <sstream>
+#include <string>
+#include <queue>
+#include <vector>
+#include <cmath>
+
+#define NUMDOFS	1
+
+using namespace std;
+
+struct Dofs
+{
+  vector<double > xx;
+  vector<double > vv;
+  Dofs ();
+}
+    ;
+
+inline Dofs::
+Dofs ()
+    : xx(NUMDOFS), vv(NUMDOFS)
+{
+}
+
+
+#endif

### 2012-05-21 one.test/template/tools/simulator/include/Force.h

@@ -1,0 +1,53 @@
+#ifndef __Force_h_NONEQMSM_wanghan__
+#define __Force_h_NONEQMSM_wanghan__
+
+#include "Defines.h"
+
+class Perturbation 
+{
+public:
+  virtual void operator () (const Dofs & dofs,
+			    const double & time,
+			    Dofs & pvalue) const = 0;
+}
+    ;
+
+class Force 
+{
+public:
+  virtual double potential (const Dofs & dofs,
+			    const double & time) const = 0;
+  virtual void operator () (const Dofs & dofs,
+			    const double & time,
+			    vector<double> & fvalue) const = 0;
+}
+    ;
+
+class PertConstTilt : public Perturbation
+{
+  double strength;
+public:
+  PertConstTilt (const double & s);
+  virtual void operator () (const Dofs & dofs,
+			    const double & time,
+			    Dofs & pvalue) const;
+}
+    ;
+
+class DoubleWell : public Force
+{
+  double kk;
+  double aa;
+public:
+  DoubleWell (const double & k,
+	      const double & a);
+  virtual double potential (const Dofs & dofs,
+			    const double & time) const;
+  virtual void operator () (const Dofs & dofs,
+			    const double & time,
+			    vector<double> & fvalue) const ;
+}
+    ;
+
+
+#endif

### 2012-05-21 one.test/template/tools/simulator/include/Integrator.h

@@ -1,0 +1,42 @@
+#ifndef __Integrator_NONEQMSM_h_wanghan__
+#define __Integrator_NONEQMSM_h_wanghan__
+
+#include "Defines.h"
+#include "Force.h"
+
+class Integrator
+{
+public:
+  virtual void step (Dofs & dofs,
+		     const double time) const = 0;
+}
+    ;
+
+class EulerMaruyama : public Integrator
+{
+  double dt;
+  double sqrtdt;
+  double gamma;
+  double sigma;
+  const Perturbation * pert;
+  const Force * force;
+public:
+  EulerMaruyama ();
+  EulerMaruyama (const double & gamma,
+		 const double & kT,
+		 const double & dt,
+		 const Perturbation * p = NULL,
+		 const Force * f = NULL);
+public:
+  void reinit (const double & gamma,
+	       const double & kT,
+	       const double & dt,
+	       const Perturbation * p = NULL,
+	       const Force * f = NULL);
+  virtual void step (Dofs & dofs,
+		     const double time) const;
+}
+    ;
+
+
+#endif

### 2012-05-21 one.test/template/tools/simulator/include/RandomGenerator.h

@@ -1,0 +1,23 @@
+#ifndef __wanghan_ToolBox_h__
+#define __wanghan_ToolBox_h__
+#include<vector>
+
+typedef double value_type;
+
+namespace RandomGenerator_MT19937{
+  void init_by_array(unsigned long init_key[], int key_length);
+  void init_genrand(unsigned long s);
+  unsigned long genrand_int32(void);
+  long genrand_int31(void);
+  double genrand_real1(void); // in [0,1]
+  double genrand_real2(void); // in [0,1)
+  double genrand_real3(void); // in (0,1)
+  double genrand_res53(void);
+  double gaussian ();		// Box Muller
+  void sphere (double & x, double & y, double & z);
+}
+
+
+
+
+#endif

### 2012-05-21 one.test/template/tools/simulator/src/Force.cpp

@@ -1,0 +1,54 @@
+#include "Force.h"
+
+PertConstTilt::
+PertConstTilt (const double & s)
+    : strength(s)
+{
+}
+
+void PertConstTilt::
+operator () (const Dofs & dofs,
+	     const double & time,
+	     Dofs & pvalue) const
+{
+  pvalue.vv[0] = strength;
+}
+
+DoubleWell::
+DoubleWell (const double & k,
+	    const double & a)
+    : kk(k), aa(a)
+{
+}
+
+void DoubleWell::
+operator () (const Dofs & dofs,
+	     const double & time,
+	     vector<double> & fvalue) const
+{
+  double tmp = 0;
+  for (unsigned dd = 0; dd < NUMDOFS; ++dd){
+    tmp += dofs.xx[dd] * dofs.xx[dd];
+  }
+  tmp -= aa * aa;
+
+  for (unsigned dd = 0; dd < NUMDOFS; ++dd){
+    fvalue[dd] = -2. * kk * tmp * dofs.xx[dd];
+  }
+}
+
+double DoubleWell::
+potential (const Dofs & dofs,
+	   const double & time) const
+{
+  double tmp = 0;
+  for (unsigned dd = 0; dd < NUMDOFS; ++dd){
+    tmp += dofs.xx[dd] * dofs.xx[dd];
+  }
+  tmp -= aa * aa;
+  
+  return 0.5 * kk * tmp * tmp;
+}
+
+
+

### 2012-05-21 one.test/template/tools/simulator/src/Integrator.cpp

@@ -1,0 +1,69 @@
+#include "Integrator.h"
+#include "RandomGenerator.h"
+
+void EulerMaruyama::
+step (Dofs & dofs,
+      const double time) const
+{
+  Dofs pvalue;
+  if (pert != NULL){
+    (*pert) (dofs, time, pvalue);
+  }
+  else {
+    for (unsigned dd = 0; dd < NUMDOFS; ++dd){
+      pvalue.xx[dd] = 0.;
+      pvalue.vv[dd] = 0.;
+    }
+  }
+  vector<double > fvalue (NUMDOFS);
+  if (pert != NULL){
+    (*force) (dofs, time, fvalue);
+  }
+  else {
+    for (unsigned dd = 0; dd < NUMDOFS; ++dd){
+      fvalue[dd] = 0.;
+    }
+  }
+
+  Dofs oldDofs(dofs);
+  for (unsigned dd = 0; dd < NUMDOFS; ++dd){
+    dofs.xx[dd] = oldDofs.xx[dd] + dt * oldDofs.vv[dd] + dt * pvalue.xx[dd];
+    dofs.vv[dd] = oldDofs.vv[dd] + dt * fvalue[dd] - dt * gamma * oldDofs.vv[dd] + dt * pvalue.vv[dd]
+	+ sqrtdt * sigma * RandomGenerator_MT19937::gaussian();
+  }
+}
+
+
+EulerMaruyama::
+EulerMaruyama ()
+    : dt(0.), sqrtdt(0.), gamma(0.), sigma(0.),
+      pert(NULL), force(NULL)
+{
+}
+
+EulerMaruyama::
+EulerMaruyama (const double & gamma,
+	       const double & kT,
+	       const double & d,
+	       const Perturbation * p,
+	       const Force * f)
+{
+  reinit (gamma, kT, d, p, f);
+}
+
+void EulerMaruyama::
+reinit (const double & g,
+	const double & kT,
+	const double & d,
+	const Perturbation * p,
+	const Force * f)
+{
+  dt = d;
+  sqrtdt = sqrt(dt);
+  gamma = g;
+  sigma = sqrt(2. * kT * gamma);
+  pert = p;
+  force = f;
+}
+
+

### 2012-05-21 one.test/template/tools/simulator/src/RandomGenerator_MT19937.cpp

@@ -1,0 +1,207 @@
+#include "RandomGenerator.h"
+#include <cmath>
+
+/* 
+   A C-program for MT19937, with initialization improved 2002/1/26.
+   Coded by Takuji Nishimura and Makoto Matsumoto.
+
+   Before using, initialize the state by using init_genrand(seed)  
+   or init_by_array(init_key, key_length).
+
+   Copyright (C) 1997 - 2002, Makoto Matsumoto and Takuji Nishimura,
+   All rights reserved.                          
+
+   Redistribution and use in source and binary forms, with or without
+   modification, are permitted provided that the following conditions
+   are met:
+
+     1. Redistributions of source code must retain the above copyright
+        notice, this list of conditions and the following disclaimer.
+
+     2. Redistributions in binary form must reproduce the above copyright
+        notice, this list of conditions and the following disclaimer in the
+        documentation and/or other materials provided with the distribution.
+
+     3. The names of its contributors may not be used to endorse or promote 
+        products derived from this software without specific prior written 
+        permission.
+
+   THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
+   "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
+   LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR
+   A PARTICULAR PURPOSE ARE DISCLAIMED.  IN NO EVENT SHALL THE COPYRIGHT OWNER OR
+   CONTRIBUTORS BE LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL,
+   EXEMPLARY, OR CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO,
+   PROCUREMENT OF SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR
+   PROFITS; OR BUSINESS INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF
+   LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING
[…]

### 2012-05-24 writedown/linear.response/langevin.tex
improve

@@
 identify core set is to consider the ``stability'' of the invariant
-measure $f_0$ under some mapping $\mt_\alpha$ of time step $\alpha$:
+measure $f_0$ under a propagator $\mt_\alpha$
+of time step $\alpha$, which is associated with a
+stochasitc process denoted by $\vect Y_\alpha$:
 \begin{align}
@@
 where $[\Delta \mt_\alpha]^\dagger$ is the adjoint operator of $\Delta
-\mt_\alpha$, which is actually $\Delta \mt_{-\alpha}$. The delta
+\mt_\alpha$. The delta
 function is replaced by the limit of a mollifier $K_\sigma(\vect y)$,
@@
   [\Delta \mt_\alpha]^\dagger K_{\sigma, \vect y}
-  \rangle_t \\
+  \rangle_t \\\nonumber
   &=
@@
   \mathbb E_{\vect x}\{
-  [\Delta \mt_\alpha]^\dagger K_{\sigma, \vect y}(s)\,
+  [\Delta \mt_\alpha]^\dagger K_{\sigma, \vect y}(\vect X_s)\,
   \}
   \rangle_0
+  \Big\} \\
+  &=
+  \lim_{\sigma \rightarrow 0}\,
+  \Big\{
+  \langle
+  [\Delta \mt_\alpha]^\dagger K_{\sigma, \vect y}
+  \rangle_0  -
+  \beta
+  \int_0^t\dd s\,
+  F_e(t-s)
+  \Big\langle
+  j(0)\,
+  \mathbb E_{\vect x}
+  \big[
+  \mathbb E_{\vect X_s}
+  (K_{\sigma, \vect y}(\vect Y_\alpha))
+  -
+  K_{\sigma, \vect y}(\vect X_s)
+  \big]
+  \Big\rangle_0
   \Big\}
+\end{align}
+with initial condistions:
+\begin{align}
+  \vect X_0 = \vect x; \quad \vect Y_0 = \vect X_s.
 \end{align}

### 2012-06-01 writedown/linear.response/langevin.tex
add results to langevin.tex

@@
 $F_e(t)$ is the strength of the perturbation as a function of time.
-The phase space imcompressibility condition is assumed:
+The phase space incompressibility condition is assumed:
 \begin{align}
@@
 \end{align}
-where $\fwg$ is the forward infiniesimal generator gieven by
+where $\fwg$ is the forward infiniesimal generator given by
 \begin{align}
@@
 of time step $\alpha$, which is associated with a
-stochasitc process denoted by $\vect Y_\alpha$:
+stochastic process denoted by $\vect Y_\alpha$:
 \begin{align}
@@
   \rangle_0
-  \Big\} \\
+  \Big\} \\ \label{eqn:main}
   &=
@@
 \end{align}
-with initial condistions:
+with initial conditions:
 \begin{align}
@@
 
+\begin{figure}
+  \centering
+  \includegraphics[width=0.35\textwidth]{figs/fig-pot.eps}
+  \caption{The potential with and without perturbation.}
+  \label{fig:tmp1}
+\end{figure}
+
+\section{Numerical test}
+
+We test the idea by a one-dimensional model system: one particle in a
+double-well potential. For convenience, we let the mass of the
+particle to be 1 \textsf{amu}. The Hamiltonian of the system is given
+by:
+\begin{align}
+  \mh (\vect p, \vect q) = \frac 12 \vect p^2 + U(\vect q) 
+\end{align}
+with
+\begin{align}
+  U(\vect q) = \frac12 k (\vect q^2 - a^2)^2
+\end{align}
+Here $k = 8$ $\textsf{kJ} / (\textsf{mol nm}^4)$, and $ a = 1 \textsf{nm}$.
+Notice at room temperature $300 \textsf{K}$, $k_BT = 2.48$ \textsf{kJ/mol}.
+See the red line in Fig.~\ref{fig:tmp1}.
+%$\textsf{kJ} / (\textsf{mol nm}^4)}$
+The perturbation is given by
+\begin{align}
+  \vect C(\vect p, \vect q) = 0; \qquad
+  \vect D(\vect p, \vect q) = -\nabla_{\vect q} V(\vect q) = 1
+\end{align}
+Here $V(\vect q) = -\vect q$, which effectively tilt the original
+potential $U(\vect q)$. The strength of the perturbation $F_e(t)$ is given
+by the following function:
+\begin{align}
+  F_e(t) = 
+  \begin{cases}
+    F_e^{\textrm{max}}\times (t / t_c) & t < t_c \\
+    F_e^{\textrm{max}} & t \geq t_c
+  \end{cases}
+\end{align}
+$t_c = 100$ \textsf{ps} so that the perturbation increases slow
+enough: the typical decaying time scale of the correlation function
+in~\eqref{eqn:main} is roughly 3 \textsf{ps}.  We choose $\mt_\alpha$
+to be the propagator of the Langevin dynamics at temperature $150$
+\textsf{K} with $\alpha = 1$ \textsf{ps}.  Since the dynamics is in
+1-d, the projection is identity.  We consider the stability of the
+distribution $f_t$ at time $t$: $-\Delta\mt_\alpha f_t$. The smaller
+this value, the more stable the core sets are.
+
+See Fig. \ref{fig:tmp2} and \ref{fig:tmp3} for numerical result.  Both
+the direct non-equilirbium simulaiton and the linear response
+approximation are presented. Time slice $t = 1$, 30, 60 and 100
+\textsf{ps} are shown.  When the perturbation is small,
+i.e. $F_e^{\textrm{max}} = 2$, the direct non-equilibrium simulation
+and the linear response results consistent very well. As the
+perturbation grows stronger, the left well becomes less stable, while
+the right well becomes more stable.  When the perturbation is big,
+i.e. $F_e^{\textrm{max}} = 6$, the linear response results are no
+longer precise. When $t\geq 60$ \textsf{ps}, the left well basically
+disappears, however, the linear response calculation present a well
+that is nearly as stable as the right well.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.95\textwidth]
+  {figs/warm100.str2.0.nst1e09.smallgrid/fig-2d.eps}
+  \caption{The stability of distribution $f_t$ under perturbation of
+    $F_e^{\textrm{max}} = 2$.
+    Upper row: the direct non-equilibrium simulation. Lower row: the
+    results of linear response formula~\eqref{eqn:main}.
+  }
+  \label{fig:tmp2}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.95\textwidth]
+  {figs/warm100.str6.0.nst1e09.smallgrid/fig-2d.eps}
+  \caption{The stability of distribution $f_t$ under perturbation of
+    $F_e^{\textrm{max}} = 6$.  Upper row: the direct non-equilibrium
+    simulation. Lower row: the results of linear response
+    formula~\eqref{eqn:main}.  }
+  \label{fig:tmp3}
+\end{figure}
+
 \section*{References}

### 2012-06-07 writedown/linear.response/langevin.tex
fix bug
add noneq linear response

@@
 \newcommand{\HY}{{\Delta}}
+\newcommand{\bsigma}{\boldsymbol\sigma}
 
@@
 
-\section{Theoretical considerations}
+\section{Perturbation from equilibrium state}
 \subsection{The linear response theory}
@@
 perturbed Langevin equation:
-\begin{align}\label{eqn:gov-1}
+\begin{align}\label{eqn:langevin-1}
   \dot{\vect q} & = \frac{\partial\mh}{\partial \vect p}
-  + \vect C(\vect q,\vect p) F_e(t)\\\label{eqn:gov-2}
+  + \vect C(\vect q,\vect p) F_e(t)\\\label{eqn:langevin-2}
   \dot{\vect p} & =- \frac{\partial\mh}{\partial \vect q}
@@
 vanishing input perturbation $F_e(t) = 0$ in
-Eqn.~\eqref{eqn:gov-1} and \eqref{eqn:gov-2}.
+Eqn.~\eqref{eqn:langevin-1} and \eqref{eqn:langevin-2}.
 We assume that the perturbation to the
@@
   e^{(t-s)\bwg_0}
-  O(\vect x)\\
+  O(\vect x)\\ \label{eqn:eqi-pert-1}
   &=
@@
 
-\section{Numerical test}
+\subsection{Numerical test}
 
@@
 See Fig. \ref{fig:tmp2} and \ref{fig:tmp3} for numerical result.  Both
-the direct non-equilirbium simulaiton and the linear response
-approximation are presented. Time slice $t = 1$, 30, 60 and 100
+the direct non-equilibrium simulation and the linear response
+approximation are presented. Time slices $t = 0$, 30, 60 and 100
 \textsf{ps} are shown.  When the perturbation is small,
@@
 
+\section{Non-equilibrium linear response}
+
+\subsection{Path integral}
+The path integral is very useful to develop the non-equilibrium version of
+linear response theory. 
+Comparing with the idea of ensemble, which is a collection of state points
+in the configuration space, the path integral consider the ``ensemble'' of
+trajectories.
+(In the following, the develop of formula are formal. The strict mathematical
+theory is still unclear for me.)
+First, we take the standard 1-dimensional Wiener process for example.
+Consider the time interval $[0,T]$, the probability density of one trajectory
+is given by
+\begin{align}
+  p[W_t] = \frac 1Z e^{-I[W_t]}
+\end{align}
+where $I$ is the action functional of Wiener process given by
+\begin{align}
+  I[W_t] = \int_0^T\frac12 \dot{W}_t^2\dd t
+\end{align}
+the partition function $Z$ is
+\begin{align}
+  Z = \int  e^{-I[W_t]} \prod_{0\leq t\leq T}\dd W_t
+\end{align}
+Consider the following two stochastic processes:
+\begin{align}\label{eqn:st-orig}
+  \dd \vect X_t &= \vect b(\vect X_t, t) \dd t + \bsigma(\vect X_t, t)\,\dd\vect W_t \\\label{eqn:st-pert}
+  \dd \vect Y_t &= \vect b(\vect Y_t, t) \dd t + \vect c(\vect Y_t, t) \dd t + \bsigma(\vect Y_t, t)\,\dd\vect W_t
+\end{align}
+Where $\vect W_t$ is the standard Wiener process. Obviously
+\eqref{eqn:st-pert} is the perturbed version of
+\eqref{eqn:st-orig}. It has been shown that, starting from the same
+measure saying $\mu$ (which is not necessarily equilibrium or
+stationary), the ratio of trajectory probability density between the perturbed
+and unperturbed process is~\cite{li2010lecture} (\recheck{I think here
+  the incompressibility of phase space should be assumed. Otherwise the
+  ratio should be more complicated.})
+\begin{align}\label{eqn:girsanov}
+  \frac{\dd\mathcal P_{\vect Y}}{\dd\mathcal P_{\vect X}}
+  =
+  \exp
+  \Big\{
+  -\frac12 \int_0^T\vert\bsigma^{-1}\cdot\vect c(\vect X_t,t)\vert^2\,\dd t
+  +\int_0^T[\dot{\vect X}_t - \vect b(\vect X_t,t)]^T
+  \cdot\bsigma^{-T}\cdot\bsigma^{-1}\cdot\vect c(\vect X_t,t)\,\dd t 
+  \Big\}
+\end{align}
+where $\mathcal P_{\vect Y}$ and $\mathcal P_{\vect Y}$ are measures
+of the trajectories generated by \eqref{eqn:st-pert} and
+\eqref{eqn:st-orig}, respectively.  This result (Girsanov formula?) is
+(formally) precise without any approximation.
+
+\subsection{Non-equilibrium perturbation with linear approximation}
+
+To be more general, we study the two perturbed Langevin dynamics
+(i.e. Eqn.~\eqref{eqn:langevin-1} and \eqref{eqn:langevin-2}) with
+different perturbations that are given by $F_e^1(t)$ and $F_e^2(t)$,
+respectively. We consider the difference between the observations of
+these dynamics:
+\begin{align}\label{eqn:noneq-pert-1}
+  \langle Q(t)\rangle_{\mu}^{2}  -\langle Q(t)\rangle_{\mu}^{1}
+  &=
+  \int Q(\vect X_t) \,\dd\mathcal P^2[\vect X_t]
+  -\int Q(\vect X_t) \,\dd\mathcal P^1[\vect X_t]
+  =
+  \int Q(\vect X_t)
+  \Big\{
+  \frac{\dd\mathcal P^2[\vect X_t]}{\dd\mathcal P^1[\vect X_t]}
+  -1
+  \Big\}
+  \,\dd\mathcal P^1[\vect X_t]
+\end{align}
+\recheck{Here we further assume that $\vect C(\vect q, \vect p) = 0$.
+  (otherwise the Girsanov formula~\eqref{eqn:girsanov} may be unstable...
+  By letting the $\epsilon$ in the following formula to 0, if
+  $\vect C\neq 0$, the limit may not be exist. This
+  is intuitionally not likely... I do not know why.)}
+By writing Eqn.~\eqref{eqn:langevin-1} and \eqref{eqn:langevin-2} as
+\begin{align}
+  \frac{\dd }{\dd t}
+  \left[
+    \begin{array}{l}
+      \vect q\\
+      \vect p
+    \end{array}
+  \right]
+  =
+  \left[
+    \begin{array}{l}
+      \nabla_{\vect p}\mh\\
+      -\nabla_{\vect q}\mh - \gamma\vect p + F_e\vect D
+    \end{array}
+  \right]
+  \dd t
+  +
+  \left[
+    \begin{array}{ll}
+      \epsilon \vect I & 0\\
+      0 & \sigma\vect I
+    \end{array}
+  \right]
+  \cdot
+  \left[
+    \begin{array}{l}
+      \dd\vect W_t\\
+      \dd\vect W_t
+    \end{array}
+  \right]
+\end{align}
+Using Girsanov's formula, to the leading order of $\Delta F_e = F_e^2 - F_e^1$,
+we have
+\begin{align}\nonumber
+  \frac{\dd\mathcal P^2[\vect X_t]}{\dd\mathcal P^1[\vect X_t]}
+  -1
+  &=
+  \exp
+  \Big\{
+  \frac 1 {\sigma^2}
+  \int_0^t
+  [\dot{\vect p}_s +
+  \gamma\vect p_s +
+  \nabla_{\vect q}\mh(\vect q_s,\vect p_s)
+  ]\cdot
+  \vect D(\vect q_s,\vect p_s)
+  \Delta F_e(s)\,
+  \dd s
+  + \mathcal O(\Delta F_e^2)
+  \Big\} - 1 \\\label{eqn:girsanov-langevin-1}
+  &\approx
+  \frac 1 {\sigma^2}
+  \int_0^t
+  [\dot{\vect p}_s +
+  \gamma\vect p_s +
+  \nabla_{\vect q}\mh(\vect q_s,\vect p_s)
+  ]\cdot
+  \vect D(\vect q_s,\vect p_s)
+  \Delta F_e(s)\,
+  \dd s  
+\end{align}
+This is where the \emph{linear approximation} applies.
+By using the relation $\beta = 2\gamma / \sigma^2$ and split the
+Eqn.~\eqref{eqn:girsanov-langevin-1} into time symmetric and
+antisymmetric parts~\cite{baiesi2010nonequilibrium2}, we have:
+\begin{align}\nonumber
+  \frac{\dd\mathcal P^2[\vect X_t]}{\dd\mathcal P^1[\vect X_t]}
+  -1
+  \approx& \:
+  \frac\beta 2
+  \int_0^t
+  \Delta F_e(s)\,
+  \vect p_s\cdot \vect D(\vect q_s,\vect p_s)\,
+  \dd s\\  \label{eqn:girsanov-langevin-2}
+  &\,+
+  \frac 1{\sigma^2}
+  \int_0^t
+  \Delta F_e(s)\,
+  \nabla_q\mh(\vect q_s,\vect p_s)\cdot \vect D(\vect q_s,\vect p_s)\,
+  \dd s  
+  +
+  \frac 1{\sigma^2}
+  \int_0^t
+  \Delta F_e(s) \vect D(\vect q_s,\vect p_s)\cdot \dd \vect p_s
+\end{align}
+The first line of the r.h.s. of Eqn.~\eqref{eqn:girsanov-langevin-2}
+is the time integrated excess entropy flux, which is
+time-antisymmetric, while the second line is the excess dynamic
+activity that is time-symmetric.  Inserting
+Eqn.~\eqref{eqn:girsanov-langevin-2} into \eqref{eqn:noneq-pert-1}, we
+have
+\begin{align}\nonumber
+  \langle Q(t)\rangle_{\mu}^{2}  -\langle Q(t)\rangle_{\mu}^{1}
+  =&\,
+  \frac\beta 2
+  \int_0^t\dd s\,
+  \Delta F_e(s)\,
+  \big\langle
+  \vect p_s\cdot\vect D(\vect X_s)\,Q(\vect X_t)
+  \big\rangle_\mu^1 \\  \label{eqn:noneq-pert-2}
+  &
+  +
+  \frac1{\sigma^2}
+  \int_0^t\dd s\,
+  \Delta F_e(s)\,
+  \big\langle
+  \nabla_{\vect q}\mh(\vect X_s)\cdot\vect D(\vect X_s)\,Q(\vect X_t)
+  \big\rangle_\mu^1
+  +
+  \frac1{\sigma^2}
+  \int_0^t
+  \Delta F_e(s)\,
+  \big\langle
+  \dd \vect p_s\cdot\,
+  \vect D(\vect X_s)\,Q(\vect X_t)
+  \big\rangle_\mu^1  
+\end{align}
+If $\vect D(\vect x)$ is independent of momentum and is conservative,
+saying $\vect D(\vect x) = \vect D(\vect q) = -\nabla_{\vect q}
+V(\vect q)$, then the last term of \eqref{eqn:noneq-pert-2} can be
+further simplified, because
+\begin{align}
+  -\dd\, [\nabla_{\vect q}V(\vect q_s)\cdot \vect p_s]
+  =
+  -\nabla_{\vect q}V(\vect q_s)\cdot\dd \vect p_s
+  - \vect p_s\cdot\nabla^2_{\vect q}V(\vect q_s)\cdot\vect p_s\,\dd s,
+\end{align}
+then \eqref{eqn:noneq-pert-2} changes to
+\begin{align}\nonumber
+  \langle Q(t)\rangle_{\mu}^{2}  -\langle Q(t)\rangle_{\mu}^{1}
+  =&\,
+  \frac\beta 2
+  \int_0^t\dd s\,
+  \Delta F_e(s)\,
+  \big\langle
+  \vect p_s\cdot\vect D(\vect X_s)\,Q(\vect X_t)
+  \big\rangle_\mu^1 \\\nonumber
+  & -
+  \frac1{\sigma^2}
+  \int_0^t\dd s\,
+  \Delta F_e(s)\,
+  \big\langle
+  \nabla_{\vect q}\mh(\vect X_s)
+  \cdot\nabla_{\vect q} V(\vect X_s)
+  \,Q(\vect X_t)
+  \big\rangle_\mu^1 \\\nonumber
+  & -
+  \frac1{\sigma^2}
+  \int_0^t
+  \dd s\,
+  \Delta F_e(s)\,
+  \frac{\dd}{\dd s}
+  \big\langle
+  \vect p_s\cdot\,
+  \nabla_{\vect q} V(\vect X_s)
+  \,Q(\vect X_t)
+  \big\rangle_\mu^1    \\\label{eqn:noneq-pert-3}
+  & +
+  \frac1{\sigma^2}
+  \int_0^t
+  \dd s\,
+  \Delta F_e(s)\,
+  \big\langle
+  \vect p_s\vect p_s:
+  \nabla^2_{\vect q} V(\vect X_s)
+  \,Q(\vect X_t)
+  \big\rangle_\mu^1      
+\end{align}
+
+\subsection{Remarks and possibilities}
+\bluec{
+\begin{itemize}
+\item It can be shown that 
+  Eqn.~\eqref{eqn:noneq-pert-3} simply reduces to
+  \eqref{eqn:eqi-pert-1}
+  in the equilibrium case~\cite{baiesi2010nonequilibrium2}.
+\item By Eqn.~\eqref{eqn:noneq-pert-3}, the trajectory average of
+  $F_e^2$ is calculated from the trajectory average of $F_e^1$. We
+  assume: (1) they start from the same initial distribution $\mu$, and
+  (2) $F_e^2$ is not very far from $F_e^1$ so that the linear
+  approximation is reasonably valid.  The condition (1) can be made
+  weaker, for example $F_e^2$ starts from $\mu_2$ and $F_e^1$ starts
+  from $\mu_1$, however, $\mu_2$ should be similar to $\mu_1$,
+  otherwise the sampling is of poor accuracy.
+\item Eqn.~\eqref{eqn:noneq-pert-3}, may be used in this way:
+  assume we want to study the dynamics of $F_e^2$, which is very slow
+  and difficult to handle directly.
+  We can instead study the dynamics of $F_e^1$, which is
+  similar to $F_e^2$, but is much faster, so that it can be handle by MD.
+  And then use \eqref{eqn:noneq-pert-3} to calculate the averages of
+  the target dynamics $F_e^2$.
+  A possible choice of  $F_e^1$ may be  $F_e^1 = \alpha F_e^2,\ \alpha > 1$.
+\item The perturbation $F_e^1$ can be kept a constant so that the
+  system reaches a stationary state.
+  $F_e^2$ is perturbation from that stationary state.
+  The averages in
+  \eqref{eqn:noneq-pert-3} can be calculated easier, because
+  what we sample is the stationary state, rather the trajectories.
+\item It may also be possible to study multiple reference dynamics,
+  and use interpolations to increase the accuracy of the linear
+  approximation.
+\end{itemize}
+}
+
 \section*{References}

### 2012-06-07 writedown/linear.response/langevin.tex
small revises

@@
   in the equilibrium case~\cite{baiesi2010nonequilibrium2}.
-\item By Eqn.~\eqref{eqn:noneq-pert-3}, the trajectory average of
+\item  Eqn.~\eqref{eqn:noneq-pert-3} shows that the trajectory average of
   $F_e^2$ is calculated from the trajectory average of $F_e^1$. We
@@
   what we sample is the stationary state, rather the trajectories.
+  When the system is far from the stationary state, the
+  accuracy is low.
 \item It may also be possible to study multiple reference dynamics,

### 2012-06-12 writedown/linear.response/langevin.tex
fix bug, more considerations

@@
   \left[
-    \begin{array}{ll}
+    \begin{array}{cc}
       \epsilon \vect I & 0\\
@@
   \dd s\\  \label{eqn:girsanov-langevin-2}
-  &\,+
+  &\,-
   \frac 1{\sigma^2}
@@
   \Delta F_e(s)\,
-  \nabla_q\mh(\vect q_s,\vect p_s)\cdot \vect D(\vect q_s,\vect p_s)\,
+  \vect F^1(\vect q_s,\vect p_s)\cdot \vect D(\vect q_s,\vect p_s)\,
   \dd s  
@@
   \int_0^t
-  \Delta F_e(s) \vect D(\vect q_s,\vect p_s)\cdot \dd \vect p_s
+  \Delta F_e(s) \vect D(\vect q_s,\vect p_s)\cdot \dd \vect p_s,
 \end{align}
-The first line of the r.h.s. of Eqn.~\eqref{eqn:girsanov-langevin-2}
-is the time integrated excess entropy flux, which is
-time-antisymmetric, while the second line is the excess dynamic
-activity that is time-symmetric.  Inserting
-Eqn.~\eqref{eqn:girsanov-langevin-2} into \eqref{eqn:noneq-pert-1}, we
-have
+where $\vect F^1(\vect q_s,\vect p_s) = -\nabla_{\vect q}\mh(\vect
+q_s,\vect p_s) + F_e^1(t) \vect D(\vect q_s,\vect p_s)$.  The first
+line of the r.h.s. of Eqn.~\eqref{eqn:girsanov-langevin-2} is the time
+integrated excess entropy flux, which is time-antisymmetric, while the
+second line is the excess dynamic activity that is time-symmetric.
+Inserting Eqn.~\eqref{eqn:girsanov-langevin-2} into
+\eqref{eqn:noneq-pert-1}, we have
 \begin{align}\nonumber
-  \langle Q(t)\rangle_{\mu}^{2}  -\langle Q(t)\rangle_{\mu}^{1}
+  \langle O(t)\rangle_{\mu}^{2}  -\langle O(t)\rangle_{\mu}^{1}
   =&\,
@@
   \big\langle
-  \vect p_s\cdot\vect D(\vect X_s)\,Q(\vect X_t)
+  \vect p_s\cdot\vect D(\vect X_s)\,O(\vect X_t)
   \big\rangle_\mu^1 \\  \label{eqn:noneq-pert-2}
   &
-  +
+  -
   \frac1{\sigma^2}
@@
   \big\langle
-  \nabla_{\vect q}\mh(\vect X_s)\cdot\vect D(\vect X_s)\,Q(\vect X_t)
+  \vect F^1(\vect X_s)\cdot\vect D(\vect X_s)\,O(\vect X_t)
   \big\rangle_\mu^1
@@
   \dd \vect p_s\cdot\,
-  \vect D(\vect X_s)\,Q(\vect X_t)
+  \vect D(\vect X_s)\,O(\vect X_t)
   \big\rangle_\mu^1  
@@
 \begin{align}\nonumber
-  \langle Q(t)\rangle_{\mu}^{2}  -\langle Q(t)\rangle_{\mu}^{1}
+  \langle O(t)\rangle_{\mu}^{2}  -\langle O(t)\rangle_{\mu}^{1}
   =&\,
-  \frac\beta 2
+  -\frac\beta 2
   \int_0^t\dd s\,
@@
   \big\langle
-  \vect p_s\cdot\vect D(\vect X_s)\,Q(\vect X_t)
+  \vect p_s\cdot\nabla_{\vect q} V(\vect X_s)\,O(\vect X_t)
   \big\rangle_\mu^1 \\\nonumber
-  & -
+  & +
   \frac1{\sigma^2}
@@
   \big\langle
-  \nabla_{\vect q}\mh(\vect X_s)
+  \vect F^1(\vect X_s)
   \cdot\nabla_{\vect q} V(\vect X_s)
-  \,Q(\vect X_t)
+  \,O(\vect X_t)
   \big\rangle_\mu^1 \\\nonumber
@@
   \nabla_{\vect q} V(\vect X_s)
-  \,Q(\vect X_t)
+  \,O(\vect X_t)
   \big\rangle_\mu^1    \\\label{eqn:noneq-pert-3}
@@
   \nabla^2_{\vect q} V(\vect X_s)
-  \,Q(\vect X_t)
+  \,O(\vect X_t)
   \big\rangle_\mu^1      
@@
   otherwise the sampling is of poor accuracy.
+  (2) means all possible trajectories of perturbation $F_e^2$ should be
+  effciently sampled by perturbation $F_e^1$.
 \item Eqn.~\eqref{eqn:noneq-pert-3}, may be used in this way:
@@
   A possible choice of  $F_e^1$ may be  $F_e^1 = \alpha F_e^2,\ \alpha > 1$.
+  On the other hand, perturbation $F_e^2$ should be not very far
+  from $F_e^1$, that means $F_e^2$ can not arbitrarily accelerate
+  $F_e^1$. Fortunately we have the freedom to design $F_e^2$, therefore,
+  an error estimate may help, and a path optimization of $F_e^2$
+  may be developed.
 \item The perturbation $F_e^1$ can be kept a constant so that the
@@
   approximation.
+\item Extension to nonlinear response?
 \end{itemize}

### 2012-07-19 writedown/linear.response/langevin.tex
fix bug
important improvement of the theory

@@
 \newcommand{\dd}{\textsf{d}}
+\newcommand{\inv}{\textrm{inv}}
 
@@
 \newcommand{\mt}{\mathcal T}
+\newcommand{\mo}{\mathcal O}
+\newcommand{\mi}{\mathcal I}
+\newcommand{\mc}{\mathcal C}
 \newcommand{\fwg}{{\mathcal A}}
 \newcommand{\bwg}{{\mathcal B}}
-\newcommand{\EX}{{\textrm{EX}}}
-\newcommand{\CG}{{\textrm{CG}}}
-\newcommand{\HY}{{\Delta}}
 \newcommand{\bsigma}{\boldsymbol\sigma}
@@
 \begin{align}\label{eqn:langevin-1}
-  \dot{\vect q} & = \frac{\partial\mh}{\partial \vect p}
+  \dot{\vect q} & = \nabla_{\vect p}\mh(\vect q,\vect p)
   + \vect C(\vect q,\vect p) F_e(t)\\\label{eqn:langevin-2}
-  \dot{\vect p} & =- \frac{\partial\mh}{\partial \vect q}
+  \dot{\vect p} & =- \nabla_{\vect q}\mh(\vect q,\vect p)
   + \vect D(\vect q,\vect p) F_e(t)
@@
 Where $\vect C$ and $\vect D$ are perturbations to the system, and
-$F_e(t)$ is the strength of the perturbation as a function of time.
+$F_e(t)$ is the strength of the perturbation as a function of time,
+and is of order $\mo (\varepsilon)$, where $\varepsilon$ is assumed to be small.
 The phase space incompressibility condition is assumed:
@@
 q, \vect p\}$) is subject to the Kolmogorov forward equation:
-\begin{align}
-  \frac{\partial}{\partial t} f(\vect x, t) - \fwg f(\vect x, t) = 0
-\end{align}
-where $\fwg$ is the forward infiniesimal generator given by
-\begin{align}
-  \fwg =
+\begin{align}\label{eqn:orig-forward}
+  \frac{\partial}{\partial t} f(\vect x, t) - \fwg^\beta(t) f(\vect x, t) = 0
+\end{align}
+where $\fwg^\beta(t)$ is the forward infiniesimal generator given by
+\begin{align}
+  \fwg^\beta(t) =
   \frac{\sigma^2}2\Delta_{\vect p}
   - \Big(
-  \frac{\partial \mh}{\partial\vect p} + \vect C F_e(t)
+  \nabla_{\vect p}\mh + \vect C F_e(t)
   \Big)\cdot\nabla_{\vect q}
   - \Big(
-  -\frac{\partial \mh}{\partial\vect q} +
+  -\nabla_{\vect q}\mh +
   \vect D F_e(t) - \gamma\vect p
   \Big)\cdot\nabla_{\vect p}
-  + 3N\gamma
-\end{align}
+  + 3N\gamma,
+\end{align}
+and $\beta$ is the inverse temperature.
+The fluctuation-dissipation relation is always true:
+$\beta = 2\gamma / \sigma^2$.
 The equilibrium and unperturbed system
@@
 system is so small that both the phase space distribution and the
-infiniesimal generator can be viewed as perturbation from the equilibrium
-distribution (invariant measure) and the unperturbed operator
+infiniesimal generator can be viewed as perturbation from the initial
+distribution and the unperturbed operator
 ($\vect C = \vect D = 0$),
-respectively. We have
-\begin{align}
-  f(\vect x, t) &= f_0(\vect x) + \Delta  f(\vect x, t)
+respectively. We the expansion with respect to small
+variable $\varepsilon$:
+\begin{align}\label{eqn:f-expan}
+  f(\vect x, t) &= f_0(\vect x, t) + \varepsilon f_1(\vect x, t)
+  +\varepsilon^2 f_2(\vect x, t) + \mo (\varepsilon^3),
 \end{align}
 and 
-\begin{align}
-  \fwg = \fwg_0 + \Delta\fwg(t)
-\end{align}
-Where the invariant measure satisfies $\fwg_0f_0(\vect x) = 0$,
+\begin{align}\label{eqn:A-expan}
+  \fwg^\beta(t) = \fwg^\beta_0 + \varepsilon\fwg_1(t),
+\end{align}
+where
+\begin{align}
+  \fwg_0^\beta =&
+  \frac{\sigma^2}2\Delta_{\vect p}
+  -
+  \nabla_{\vect p}\mh\cdot\nabla_{\vect q}
+  - \big(
+  -\nabla_{\vect q}\mh - \gamma\vect p
+  \big)\cdot\nabla_{\vect p}
+  + 3N\gamma,\\
+  \fwg_1(t) =&
+  - \frac{F_e(t)}{\varepsilon} \big(
+  \vect C \cdot\nabla_{\vect q}
+  +
+  \vect D \cdot\nabla_{\vect p}
+  \big).
+\end{align}
+Notice $F_e(t) \sim \mo(\varepsilon)$, so $\Vert\fwg_1(t)\Vert\sim\mo(1)$.
+The invariant measure of $\fwg^\beta_0$ satisfies $\fwg^\beta_0f^\beta_{\inv}(\vect x) = 0$,
 which is nothing but the Boltzmann distribution
 \begin{align}
-  f_0(\vect x) \propto e^{-\beta\mh(\vect x)}
-\end{align}
-The Kolmogorov forward equation for the perturbation is therefore
+  f^\beta_{\inv}(\vect x) \propto e^{-\beta\mh(\vect x)}
+\end{align}
+Inserting \eqref{eqn:f-expan} and \eqref{eqn:A-expan} into
+\eqref{eqn:orig-forward}, 
+the Kolmogorov forward equation becomes:
 \begin{align}
   \frac{\partial}{\partial t}
-  [f_0(\vect x) + \Delta  f(\vect x, t)]
+  [f_0(\vect x, t) + \varepsilon  f_1(\vect x, t) + \varepsilon^2f_2(\vect x, t)]
   -
-  [\,\fwg_0 + \Delta\fwg(t)\,]
-  [\,f_0(\vect x) + \Delta  f(\vect x, t)\,] = 0
-\end{align}
-The basic assumption for the ``linear response'' is that the high order
-term $\Delta\fwg(t)  \Delta  f(\vect x, t)$ 
-can be neglected. Noticing $\fwg_0 f_0(\vect x) = 0$, we have the
-Kolmogorov forward equation
-for the leading order of the perturbation:
-\begin{align}
+  [\,\fwg^\beta_0 + \varepsilon\fwg_1(t)\,]
+  [f_0(\vect x, t) + \varepsilon  f_1(\vect x, t) + \varepsilon^2f_2(\vect x, t)]
+  = 0
+\end{align}
+Matching this equation at different orders of $\varepsilon$, the zeroth
+order gives:
+\begin{align}\label{eqn:e-o0}
   \bigg[
   \frac{\partial}{\partial t}
-  - \fwg_0
+  - \fwg_0^\beta
   \bigg]
-  \Delta  f(\vect x, t)
-  =
-  \Delta\fwg(t) f_0(\vect x)
-\end{align}
-The operator on l.h.s. is the same as the unperturbed equation.
-Formally solving this equation with initial condition
-$\Delta f(\vect x, 0) = 0$ gives
-\begin{align}
-  \Delta  f(\vect x, t) =
-  \int_0^t \dd s \:
-  e^{(t-s)\fwg_0}
-  \Delta\fwg(s)f_0(\vect x)
-\end{align}
-By defining the \emph{dissipative flux},
+  f_0(\vect x, t)
+  = 0
+\end{align}
+which is the unperturbed equation. Matching the first and the second order gives:
+\begin{align}\label{eqn:e-o1}
+  \bigg[
+  \frac{\partial}{\partial t}
+  - \fwg_0^\beta
+  \bigg]
+  f_1(\vect x, t)
+  =&
+  \fwg_1(t) f_0(\vect x, t)\\\label{eqn:e-o2}
+  \bigg[
+  \frac{\partial}{\partial t}
+  - \fwg_0^\beta
+  \bigg]
+  f_2(\vect x, t)
+  =&
+  \fwg_1(t) f_1(\vect x, t)
+\end{align}
+Formally solving \eqref{eqn:e-o0} -- \eqref{eqn:e-o2} with
+inital conditions $f_1(\vect x, 0) = 0$ and $f_2(\vect x, 0) = 0$
+gives
+\begin{align}\label{eqn:f0-0}
+  f_0(\vect x, t)
+  =&\,
+  e^{t\fwg^\beta_0}f_0(\vect x, 0) \\\label{eqn:f1-0}
+  f_1(\vect x, t)
+  =&\,
+  \int_0^t\dd s\,
+  e^{(t-s)\fwg^\beta_0}\circ
+  \fwg_1(s) f_0(\vect x,s)
+  =
+  \int_0^t\dd s\,
+  e^{(t-s)\fwg^\beta_0}\circ
+  \fwg_1(s)\circ
+  e^{s\fwg^\beta_0}f_0(\vect x, 0) \\\nonumber
+  f_2(\vect x, t)
+  =&\,
+  \int_0^t\dd s\,
+  e^{(t-s)\fwg^\beta_0}\circ
+  \fwg_1(s) f_1(\vect x,s) \\\label{eqn:f2-0}
+  &=
+  \int_0^t\dd s
+  \int_0^s\dd u\,\,
+  e^{(t-s)\fwg^\beta_0}\circ
+  \fwg_1(s)\circ
+  e^{(s-u)\fwg^\beta_0}\circ
+  \fwg_1(u)\circ
+  e^{u\fwg^\beta_0}
+  f_0(\vect x, 0) 
+\end{align}
+If the initial distribution is the invariant distribution, i.e.
+$f_0(\vect x, 0) = f_{\inv}^\beta(\vect x)$,
+\eqref{eqn:f0-0} becomes:
+\begin{align}
+  f_0(\vect x, t) = f_{\inv}^\beta(\vect x)
+\end{align}
+It can be shown that
+\begin{align}
+  \fwg_1(s) f_{\inv}^\beta(\vect x)
+  =
+  -\frac{F_e(s)}{\varepsilon}\beta
+  j(\vect x)
+  f_{\inv}^\beta(\vect x),
+\end{align}
+where the \emph{dissipative flux} $j(\vect x)$ is defined by,
 \begin{align}
@@
 \end{align}
-it has been shown that %~\cite{tuckeman2010statistical}
-\begin{align}
-  \Delta\fwg(s)f_0(\vect x) =
-  -\beta f_0(\vect x)\, j(\vect x) F_e(s)
-\end{align}
-For any observation:
+Then $f_1(\vect x, t)$ becomes:
+\begin{align}
+  f_1(\vect x, t)
+  =
+  -\frac\beta\varepsilon
+  \int_0^t\dd s\,
+  F_e(s)\,
+  e^{(t-s)\fwg^\beta_0}
+  \big[
+  j(\vect x)
+  f_{\inv}^\beta(\vect x)
+  \big]
+\end{align}
+Therefore, the linear order approximation to
+the solution of~\eqref{eqn:orig-forward} is 
+\begin{align}\label{eqn:solv-orig-linear}
+  f(\vect x, t) =
+  f_{\inv}^\beta(\vect x)
+  - \beta
+  \int_0^t\dd s\,
+  F_e(s)\,
+  e^{(t-s)\fwg^\beta_0}
+  \big[
+  j(\vect x)
+  f_{\inv}^\beta(\vect x)
+  \big]
+  + \mo(\varepsilon^2)
+\end{align}
+For any time-dependent observation, the linear response approximation
+would be:
 \begin{align}\nonumber
@@
   \int\dd \vect x\, O(\vect x)\,
-  [f_0(\vect x) + \Delta f(\vect x, t)] \\\nonumber
-  &=
-  \langle O\rangle_0
+  [\,f_{\inv}^\beta(\vect x) + \varepsilon f_1(\vect x, t)] \\\nonumber
+  &=
+  \langle O\rangle
   -
@@
   O(\vect x)
-  \int_0^t\dd s\,
-  e^{(t-s)\fwg_0}
-  f_0(\vect x)\, j(\vect x)F_e(s)  \\\nonumber
-  &=
-  \langle O\rangle_0
+  \int_0^t\dd s\,F_e(s)
+  e^{(t-s)\fwg_0^\beta}
+  \big[j(\vect x)f^\beta_{\inv}(\vect x)\big]
+  \\\nonumber
+  &=
+  \langle O\rangle
   -
@@
   O(\vect x)\,
-  e^{(t-s)\fwg_0}
-  f_0(\vect x) j(\vect x)  \\\nonumber
-  &=
-  \langle O\rangle_0
+  e^{(t-s)\fwg_0^\beta}
+  \big[j(\vect x)f^\beta_{\inv}(\vect x)\big]
+  \\\nonumber
+  &=
+  \langle O\rangle
   -
@@
   \int \dd \vect x\,
-  f_0(\vect x) j(\vect x)\,
-  e^{(t-s)\bwg_0}
+  f^\beta_{\inv}(\vect x) j(\vect x)\,
+  e^{(t-s)\bwg^\beta_0}
   O(\vect x)\\ \label{eqn:eqi-pert-1}
   &=
-  \langle O\rangle_0
+  \langle O\rangle
   -
@@
   \int \dd \vect x\,
-  f_0(\vect x) j(\vect x)\,
-  e^{s\bwg_0}
+  f^\beta_{\inv}(\vect x) j(\vect x)\,
+  e^{s\bwg_0^\beta}
   O(\vect x)
 \end{align}
-Where $\bwg_0$ is the infiniesimal generator of the \emph{unperturbed} 
-Kolmogorov backward equation, which is the adjoint operator of $\fwg_0$.
+Where $\bwg^\beta_0$ is the infiniesimal generator of the \emph{unperturbed} 
+Kolmogorov backward equation, which is the adjoint operator of $\fwg^\beta_0$.
 We have
 \begin{align}
-  e^{s\bwg_0} O(\vect x) = \mathbb E_{\vect x} [O(\vect X_s)]
-\end{align}
-where $\vect X_t$ is the trajectory of the unperturbed Langevin equation
+  e^{s\bwg^\beta_0} O(\vect x) = \mathbb E_{\vect x} [O(\vect X^\beta_s)]
+\end{align}
+where $\vect X^\beta_t$ is the trajectory of the unperturbed Langevin equation
 starting at $\vect x$.
@@
   =
-  \langle O\rangle_0
+  \langle O\rangle
   -
@@
   \int \dd \vect x\,
-  f_0(\vect x)
+  f^\beta_{\inv}(\vect x)
   j(\vect x)\,
-  \mathbb E_{\vect x} [O(\vect X_s)]
+  \mathbb E_{\vect x} [O(\vect X^\beta_s)]
 \end{align}
@@
   F_e(t-s)
-  \langle
+  \big\langle
   j(0)\,
-  \mathbb E_{\vect x} [O(\vect X_s)]
-  \rangle_0
+  \mathbb E_{\vect x} [O(\vect X^\beta_s)]
+  \big\rangle
 \end{align}
@@
   j(0)\,
-  \mathbb E_{\vect x} [O(\vect X_s)]
-  \rangle_0$ is the \emph{equilibrium time
+  \mathbb E_{\vect x} [O(\vect X^\beta_s)]
+  \rangle$ is the \emph{equilibrium time
   correlation} between the observation $O$ at time $s$ and the
-dissipative flux $j$ at time 0.
+dissipative flux $j$.
 
@@
 identify core set is to consider the ``stability'' of the invariant
-measure $f_0$ under a propagator $\mt_\alpha$
-of time step $\alpha$, which is associated with a
-stochastic process denoted by $\vect Y_\alpha$:
-\begin{align}
-  C_\alpha = \big\{
-  \vect x \,\vert\, \mt_\alpha f_0(\vect x) > f_0(\vect x)
+measure $f^\beta_{\inv}$ under a propagator $\mt_\tau$
+of time step $\tau$.
+For example, the propagator is chosen $\mt_\tau = e^{\tau\fwg_t^{2\beta}}$,
+i.e. integrating the Langevin equation~\eqref{eqn:langevin-1}
+and \eqref{eqn:langevin-2} at temperature $1/2\beta$ for time $\tau$.
+Notice $\tau$ is an artificial time that is ``orthogonal'' to the real
+time $t$.
+% , which is associated with a
+% stochastic process denoted by $\vect Y_\alpha$:
+The core set is identified by:
+\begin{align}
+  C_\tau(t) = \big\{
+  \vect x \,\vert\, \mt_\tau f(\vect x, t) > f(\vect x, t)
   \big\}
@@
 \begin{align}
-  \mathcal P f_0 (\vect y)
+  \mathcal P f (\vect y, t)
   =
@@
   \dd \vect x\,
-  f_0(\vect x) \,
+  f(\vect x, t) \,
   \delta (\vect y - \vect y(\vect x)) 
@@
   \widetilde
-  C_\alpha = \big\{
+  C_\tau(t) = \big\{
   \vect y \,\vert\,
-  \mathcal P\mt_\alpha f_0(\vect y) > \mathcal P f_0(\vect y)
+  \mathcal P\mt_\tau f(\vect y, t) > \mathcal P f(\vect y, t)
   \big\}
@@
 % \end{align}
-For convenience, we define $\Delta \mt_\alpha =
-\mt_\alpha - \mathcal I$, where $\mathcal I$ is the
-identity. If $[\mathcal P\circ\Delta \mt_\alpha f_0](\vect y) > 0$, then
-$\vect y$ is in the core set $\widetilde C_\alpha $.
+For convenience, we define $\Delta \mt_\tau =
+\mt_\tau - \mathcal I$, where $\mathcal I$ is the
+identity. If $[\mathcal P\circ\Delta \mt_\tau f](\vect y, t) > 0$, then
+$\vect y$ is in the core set $\widetilde C_\tau(t) $.
 \begin{align}\nonumber
-  [\mathcal P\circ\Delta \mt_\alpha f_0](\vect y)
+  [\mathcal P\circ\Delta \mt_\tau f](\vect y, t)
   &=
@@
   \dd \vect x\,
-  \Delta \mt_\alpha f_0(\vect x) \,
+  \Delta \mt_\tau f(\vect x, t) \,
   \delta (\vect y - \vect y(\vect x))  \\\nonumber
@@
   \dd \vect x\,
-  \Delta \mt_\alpha f_0(\vect x) \,
-  K_{\sigma} (\vect y - \vect y(\vect x))  \\\nonumber
+  \Delta \mt_\tau f(\vect x, t) \,
+  K_{\sigma} (\vect y - \vect y(\vect x))  \\\label{eqn:core-identify}
   &=
@@
   \dd \vect x\,
-  \Delta \mt_\alpha f_0(\vect x) \,
-  K_{\sigma, \vect y} (\vect x)  \\\nonumber
-  &=
+  \Delta \mt_\tau f(\vect x, t) \,
+  K_{\sigma, \vect y} (\vect x)
+  % \\\nonumber
+  % &=
+  % \lim_{\sigma \rightarrow 0}
+  % \int
+  % \dd \vect x\,
+  % f_0(\vect x) \,
+  % [\Delta \mt_\alpha]^\dagger K_{\sigma, \vect y} (\vect x)\\
+  % &=
+  % \lim_{\sigma \rightarrow 0}\,
+  % \langle
+  % [\Delta \mt_\alpha]^\dagger K_{\sigma, \vect y}
+  % \rangle_0
+\end{align}
+For convenience, we denote $f_t(\vect x) = f(\vect x, t)$.
+Notice that $\mt_\tau f_t(\vect x) = e^{\tau \fwg_t^{2\beta}} f_t(\vect x)$
+is the solution of
+\begin{align}
+  \frac{\partial}{\partial \tau} f_t(\vect x, \tau)
+  -
+  \fwg^{2\beta}(t)  f_t(\vect x, \tau)
+  =0
+\end{align}
+with inital condition $f_t(\vect x, \tau = 0) = f_t(\vect x)$. We do the same
+expansion as before, and then
+matching different orders of $\varepsilon$ gives:
+\begin{align}
+  &\bigg[
+  \frac{\partial}{\partial \tau}
+  - \fwg_0^{2\beta}
+  \bigg]
+  f_{t,0}(\vect x, \tau)
+  = 0 \\
+  &\bigg[
+  \frac{\partial}{\partial \tau}
+  - \fwg_0^{2\beta}
+  \bigg]
+  f_{t,1}(\vect x, \tau)
+  =
+  \fwg_1(t) f_{t,0}(\vect x, \tau)
+\end{align}
+Formally solving these equations with
+inital conditions $f_{t,0}(\vect x, 0) = f_t(\vect x) = f(\vect x, t)$
+and  $f_{t,1}(\vect x, 0) = 0$ 
+gives
+\begin{align}\label{eqn:f0-1}
+  f_{t,0}(\vect x, \tau)
+  =&\,
+  e^{\tau\fwg^{2\beta}_0}f_{t,0}(\vect x, 0) \\\label{eqn:f1-1}
+  f_{t,1}(\vect x, \tau)
+  =&\,
+  \int_0^\tau\dd \eta\,
+  e^{(\tau-\eta)\fwg^{2\beta}_0}\circ
+  \fwg_1(t) f_{t,0}(\vect x,\eta)
+  =
+  \int_0^\tau\dd \eta\,
+  e^{(\tau-\eta)\fwg^{2\beta}_0}\circ
+  \fwg_1(t)\circ
+  e^{\eta\fwg^{2\beta}_0}f_{t,0}(\vect x, 0) 
+\end{align}
+By inserting the linear order approximated solution~\eqref{eqn:solv-orig-linear}
+in to the above equations, we have:
+\begin{align}\nonumber
+  \mt_\tau f_t(\vect x)=
+  f_{t}(\vect x, \tau)
+  =&\,
+  e^{\tau\fwg^{2\beta}_0} f_{\inv}^\beta(\vect x) -
+  \beta
+  \int_0^t\dd s\,
+  F_e(s)\,
+  e^{\tau\fwg^{2\beta}_0} \circ
+  e^{(t-s)\fwg^\beta_0}
+  \big[
+  j(\vect x)
+  f_{\inv}^\beta(\vect x)
+  \big]
+  \\\label{eqn:solve-tau-linear}
+  &\,+ \varepsilon
+  \int_0^\tau\dd \eta\,
+  e^{(\tau-\eta)\fwg^{2\beta}_0}\circ
+  \fwg_1(t)\circ
+  e^{\eta\fwg^{2\beta}_0}f_{\inv}^\beta(\vect x)  + \mo(\varepsilon^2)
+\end{align}
+By inserting~\eqref{eqn:solv-orig-linear} and~\eqref{eqn:solve-tau-linear}
+into~\eqref{eqn:core-identify}, we have, up to the first order
+approximation,
+\begin{align}\nonumber
+  [\mathcal P\circ\Delta\mt_\tau f](\vect y, t)
+  \approx
   \lim_{\sigma \rightarrow 0}
-  \int
-  \dd \vect x\,
-  f_0(\vect x) \,
-  [\Delta \mt_\alpha]^\dagger K_{\sigma, \vect y} (\vect x)\\
-  &=
-  \lim_{\sigma \rightarrow 0}\,
-  \langle
-  [\Delta \mt_\alpha]^\dagger K_{\sigma, \vect y}
-  \rangle_0
-\end{align}
-where $[\Delta \mt_\alpha]^\dagger$ is the adjoint operator of $\Delta
-\mt_\alpha$. The delta
-function is replaced by the limit of a mollifier $K_\sigma(\vect y)$,
-which is infinitely smooth, compact supported of size $\sigma$,
-bounded and $\int
-K_\sigma(\vect y)\dd \vect y = 1$.  The physical meaning is that we
-can consider the expectation value of the stability of the mollifier
-$K_{\sigma, \vect y}$ instead of the stability of the invariant measure.
-This provides the convenience of using the linear response theory showed
-in the previous section. By Eqn.~\eqref{eqn:lr},
-\begin{align}\nonumber
-  [\mathcal P\circ\Delta \mt_\alpha f_t](\vect y)
-  &=
-  \lim_{\sigma \rightarrow 0}\,
-  \langle
-  [\Delta \mt_\alpha]^\dagger K_{\sigma, \vect y}
-  \rangle_t \\\nonumber
-  &=
-  \lim_{\sigma \rightarrow 0}\,
-  \Big\{
-  \langle
-  [\Delta \mt_\alpha]^\dagger K_{\sigma, \vect y}
-  \rangle_0  -
-  \beta
-  \int_0^t\dd s\,
-  F_e(t-s)
-  \langle
-  j(0)\,
-  \mathbb E_{\vect x}\{
-  [\Delta \mt_\alpha]^\dagger K_{\sigma, \vect y}(\vect X_s)\,
-  \}
-  \rangle_0
-  \Big\} \\ \label{eqn:main}
-  &=
-  \lim_{\sigma \rightarrow 0}\,
-  \Big\{
-  \langle
-  [\Delta \mt_\alpha]^\dagger K_{\sigma, \vect y}
-  \rangle_0  -
-  \beta
-  \int_0^t\dd s\,
-  F_e(t-s)
+  \bigg\{
+    &\,
+    \int \dd \vect x\,
+    \big(e^{\tau\fwg_0^{2\beta}}-\mi\big)
+    f_{\inv}^\beta(\vect x)\,
+    K_{\sigma,\vect y}(\vect x)\\\nonumber
+    &
+    - \int_0^t\dd s\,
+    F_e(s)
+    \int\dd \vect x\,
+    \big(e^{\tau\fwg^{2\beta}_0} - \mi\big)\circ
+    e^{(t-s)\fwg^\beta_0}
+    \big[
+    j(\vect x)
+    f_{\inv}^\beta(\vect x)
+    \big]\, K_{\sigma,\vect y}(\vect x) \\\nonumber
+    &
+    +
+    \varepsilon
+    \int_0^\tau\dd \eta
+    \int\dd \vect x\,
+    e^{(\tau-\eta)\fwg^{2\beta}_0}\circ
+    \fwg_1(t)\circ
+    e^{\eta\fwg^{2\beta}_0}f_{\inv}^\beta(\vect x)\,K_{\sigma,\vect y}(\vect x)
+    \bigg\}     \\\nonumber
+  \approx
+  \lim_{\sigma \rightarrow 0}
+  \bigg\{
+  &\,
     \Big\langle
-  j(0)\,
-  \mathbb E_{\vect x}
-  \big[
-  \mathbb E_{\vect X_s}
-  (K_{\sigma, \vect y}(\vect Y_\alpha))
+    \mathbb E_{\vect x}[ K_{\sigma,\vect y}(\vect X^{2\beta}_\tau) ]
+    - K_{\sigma,\vect y}(\vect x)
+    \Big\rangle\\\nonumber
+    &
+    - \int_0^t\dd s\,
+    F_e(s)
+    \Big\langle
+    j(\vect x)\,
+    \mathbb E_{\vect x}\big\{
+    \mathbb E_{\vect X_{t-s}^\beta}[
+    K_{\sigma,\vect y}(\vect X_\tau^{2\beta})
+    ]
     -
-  K_{\sigma, \vect y}(\vect X_s)
-  \big]
-  \Big\rangle_0
-  \Big\}
-\end{align}
-with initial conditions:
-\begin{align}
-  \vect X_0 = \vect x; \quad \vect Y_0 = \vect X_s.
-\end{align}
-Therefore, the core set at time $t$ is defined by
-\begin{align}
-  \widetilde C_\alpha^t =
-  \Big\{
-  \vect y
-  \,\Big\vert\,
-  [\mathcal P\circ\Delta \mt_\alpha f_t](\vect y) > 0
-  \Big\}
-\end{align}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.35\textwidth]{figs/fig-pot.eps}
-  \caption{The potential with and without perturbation.}
-  \label{fig:tmp1}
-\end{figure}
-
-\subsection{Numerical test}
-
-We test the idea by a one-dimensional model system: one particle in a
-double-well potential. For convenience, we let the mass of the
-particle to be 1 \textsf{amu}. The Hamiltonian of the system is given
-by:
-\begin{align}
-  \mh (\vect p, \vect q) = \frac 12 \vect p^2 + U(\vect q) 
-\end{align}
-with
-\begin{align}
-  U(\vect q) = \frac12 k (\vect q^2 - a^2)^2
-\end{align}
-Here $k = 8$ $\textsf{kJ} / (\textsf{mol nm}^4)$, and $ a = 1 \textsf{nm}$.
-Notice at room temperature $300 \textsf{K}$, $k_BT = 2.48$ \textsf{kJ/mol}.
-See the red line in Fig.~\ref{fig:tmp1}.
-%$\textsf{kJ} / (\textsf{mol nm}^4)}$
-The perturbation is given by
-\begin{align}
-  \vect C(\vect p, \vect q) = 0; \qquad
-  \vect D(\vect p, \vect q) = -\nabla_{\vect q} V(\vect q) = 1
-\end{align}
-Here $V(\vect q) = -\vect q$, which effectively tilt the original
-potential $U(\vect q)$. The strength of the perturbation $F_e(t)$ is given
-by the following function:
-\begin{align}
-  F_e(t) = 
-  \begin{cases}
-    F_e^{\textrm{max}}\times (t / t_c) & t < t_c \\
-    F_e^{\textrm{max}} & t \geq t_c
-  \end{cases}
-\end{align}
-$t_c = 100$ \textsf{ps} so that the perturbation increases slow
-enough: the typical decaying time scale of the correlation function
-in~\eqref{eqn:main} is roughly 3 \textsf{ps}.  We choose $\mt_\alpha$
-to be the propagator of the Langevin dynamics at temperature $150$
-\textsf{K} with $\alpha = 1$ \textsf{ps}.  Since the dynamics is in
-1-d, the projection is identity.  We consider the stability of the
-distribution $f_t$ at time $t$: $-\Delta\mt_\alpha f_t$. The smaller
-this value, the more stable the core sets are.
-
-See Fig. \ref{fig:tmp2} and \ref{fig:tmp3} for numerical result.  Both
-the direct non-equilibrium simulation and the linear response
-approximation are presented. Time slices $t = 0$, 30, 60 and 100
-\textsf{ps} are shown.  When the perturbation is small,
-i.e. $F_e^{\textrm{max}} = 2$, the direct non-equilibrium simulation
-and the linear response results consistent very well. As the
-perturbation grows stronger, the left well becomes less stable, while
-the right well becomes more stable.  When the perturbation is big,
-i.e. $F_e^{\textrm{max}} = 6$, the linear response results are no
-longer precise. When $t\geq 60$ \textsf{ps}, the left well basically
-disappears, however, the linear response calculation present a well
-that is nearly as stable as the right well.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.95\textwidth]
-  {figs/warm100.str2.0.nst1e09.smallgrid/fig-2d.eps}
-  \caption{The stability of distribution $f_t$ under perturbation of
-    $F_e^{\textrm{max}} = 2$.
-    Upper row: the direct non-equilibrium simulation. Lower row: the
-    results of linear response formula~\eqref{eqn:main}.
-  }
-  \label{fig:tmp2}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.95\textwidth]
-  {figs/warm100.str6.0.nst1e09.smallgrid/fig-2d.eps}
-  \caption{The stability of distribution $f_t$ under perturbation of
-    $F_e^{\textrm{max}} = 6$.  Upper row: the direct non-equilibrium
-    simulation. Lower row: the results of linear response
-    formula~\eqref{eqn:main}.  }
-  \label{fig:tmp3}
-\end{figure}
-
-\section{Non-equilibrium linear response}
+    K_{\sigma,\vect y}(\vect X_{t-s}^\beta)
+    \big\}
+    \Big\rangle\\\label{eqn:core-identify-approx-1}
+    &
+    +
+    \varepsilon
+    \int_0^\tau\dd \eta
+    \int\dd \vect x\,
+    e^{(\tau-\eta)\fwg^{2\beta}_0}\circ
+    \fwg_1(t)\circ
+    e^{\eta\fwg^{2\beta}_0}f_{\inv}^\beta(\vect x)\,K_{\sigma,\vect y}(\vect x)
+    \bigg\}     
+\end{align}
+Unfortunately the last term on the r.h.s. of~\eqref{eqn:core-identify-approx-1}
+is not easy to calculate. We will deal with it with
+the path integral method introduced later.
 
@@
 
-\subsection{Non-equilibrium perturbation with linear approximation}
-
 To be more general, we study the two perturbed Langevin dynamics
@@
 different perturbations that are given by $F_e^1(t)$ and $F_e^2(t)$,
-respectively. We consider the difference between the observations of
-these dynamics:
-\begin{align}\label{eqn:noneq-pert-1}
-  \langle Q(t)\rangle_{\mu}^{2}  -\langle Q(t)\rangle_{\mu}^{1}
-  &=
-  \int Q(\vect X_t) \,\dd\mathcal P^2[\vect X_t]
-  -\int Q(\vect X_t) \,\dd\mathcal P^1[\vect X_t]
-  =
-  \int Q(\vect X_t)
-  \Big\{
-  \frac{\dd\mathcal P^2[\vect X_t]}{\dd\mathcal P^1[\vect X_t]}
-  -1
-  \Big\}
-  \,\dd\mathcal P^1[\vect X_t]
-\end{align}
-\recheck{Here we further assume that $\vect C(\vect q, \vect p) = 0$.
-  (otherwise the Girsanov formula~\eqref{eqn:girsanov} may be unstable...
-  By letting the $\epsilon$ in the following formula to 0, if
-  $\vect C\neq 0$, the limit may not be exist. This
-  is intuitionally not likely... I do not know why.)}
+respectively. 
 By writing Eqn.~\eqref{eqn:langevin-1} and \eqref{eqn:langevin-2} as
@@
 we have
-\begin{align}\nonumber
+\begin{align}\label{eqn:girsanov-1}
   \frac{\dd\mathcal P^2[\vect X_t]}{\dd\mathcal P^1[\vect X_t]}
-  -1
   &=
@@
   \nabla_{\vect q}\mh(\vect q_s,\vect p_s)
+  - F_e^1(s) \vect D(\vect q_s,\vect p_s)
+  ]\cdot
+  \vect D(\vect q_s,\vect p_s)
+  \Delta F_e(s)\,
+  \dd s
+  + \mathcal O(\Delta F_e^2)
+  \Big\} 
+\end{align}
+
+\subsection{Linear response theory of the core set identification,
+  revisited by the path integral method}
+For some observable $O[\vect X_\eta]$, which is a functional of the
+stochastic process $\vect X_\eta$,
+the trajectory ensemble average yields:
+\begin{align}
+  \mo(t)
+  =
+  \int \dd \mu_t(\vect x)
+  \int_{\mc\{\vect x,0;\tau\}}
+  O[\vect X_\eta]\,
+  \dd \mathcal P^{2\beta}_t[\vect X_\eta],
+\end{align}
+where $\mu_t(\vect x)$
+is the measure of the initial position, and obviously, we have
+$\dd \mu_t(\vect x) = f_t(\vect x)\,\dd \vect x$.
+The trajectory $\vect X_\eta \in {\mc\{\vect x,0;\tau\}}$, where
+${\mc\{\vect x,0;\tau\}}$ denotes the set of all continuous but not
+differentiable trajectories starting at $\vect x$ at artificial
+time 0, and ending
+at artificial time $\tau$.
+$\mathcal P^{2\beta}_t$ denotes the measure of trajectories
+developped under temperature $1/2\beta$
+at real time $t$.
+Since the perturbation is a function of time, this
+measure also depends on time. The reference artificial dynamics is chosen
+at $t=0$.
+In this case, \eqref{eqn:girsanov-1} rewrites:
+\begin{align}\nonumber
+  \frac{\dd \mathcal P^{2\beta}_t[\vect X_\eta]}
+  {\dd \mathcal P^{2\beta}_0[\vect X_\eta]}\,
+  &=
+  \exp
+  \Big\{
+  \frac 1 {\sigma^2}
+  \int_0^\tau
+  [\dot{\vect p}_\eta +
+  \gamma\vect p_\eta +
+  \nabla_{\vect q}\mh(\vect q_\eta,\vect p_\eta)
+  ]\cdot
+  \vect D(\vect q_\eta,\vect p_\eta)
+  F_e(t)\,
+  \dd \eta
+  + \mathcal O(\varepsilon^2)
+  \Big\}\\
+  &=
+  1+
+  F_e(t)\,
+  \frac 1 {\sigma^2}
+  \int_0^\tau
+  [\dot{\vect p}_\eta +
+  \gamma\vect p_\eta +
+  \nabla_{\vect q}\mh(\vect q_\eta,\vect p_\eta)
+  ]\cdot
+  \vect D(\vect q_\eta,\vect p_\eta)\,
+  \dd \eta
+  + \mathcal O(\varepsilon^2)  
+\end{align}
+Notice that $F_e(t)\sim \mo(\varepsilon)$.
+For convenience, we denote
+\begin{align}
+  G[\vect X_\eta] =
+    \frac 1 {\sigma^2}
+  \int_0^\tau
+  [\dot{\vect p}_\eta +
+  \gamma\vect p_\eta +
+  \nabla_{\vect q}\mh(\vect q_\eta,\vect p_\eta)
+  ]\cdot
+  \vect D(\vect q_\eta,\vect p_\eta)\,
+  \dd \eta
+\end{align}
+Therefore, we have
+\begin{align}\nonumber
+  \mo(t)
+  =&\,
+  \int \dd \vect x
+  f(\vect x, t)
+  \int_{\mc\{\vect x,0;\tau\}}
+  O[\vect X_\eta]\,
+  \frac{\dd \mathcal P^{2\beta}_t[\vect X_\eta]}
+  {\dd \mathcal P^{2\beta}_0[\vect X_\eta]}\,
+  \dd \mathcal P^{2\beta}_0[\vect X_\eta]  \\\nonumber
+  =&\,
+  \int \dd \vect x
+  \big[
+  f_{0}(\vect x, t)
+  + \varepsilon f_{1}(\vect x, t)
+  + \mo(\varepsilon^2)
+  \big]
+  \int_{\mc\{\vect x,0;\tau\}}
+  O[\vect X_\eta]\,
+  \big[
+  1 
+  + F_e(t) G[\vect X_\eta]
+  + \mo(\varepsilon^2)
+  \big]
+  \dd \mathcal P^{2\beta}_0[\vect X_\eta]
+  \\\nonumber
+  =&\,
+  \int \dd \vect x
+  f_{0}(\vect x, t)
+  \int_{\mc\{\vect x,0;\tau\}}
+  O[\vect X_\eta]\,
+  \dd \mathcal P^{2\beta}_0[\vect X_\eta]
+  +
+  \varepsilon
+  \int \dd \vect x\,
+  f_{1}(\vect x, t)
+  \int_{\mc\{\vect x,0;\tau\}}
+  O[\vect X_\eta]\,
+  \dd \mathcal P^{2\beta}_0[\vect X_\eta]\\\label{eqn:solve-tau-linear-2-0}
+  &+
+  F_e(t)
+  \int \dd \vect x\,
+  f_{0}(\vect x, t)
+  \int_{\mc\{\vect x,0;\tau\}}
+  O[\vect X_\eta]\,
+  G[\vect X_\eta]\,
+  \dd \mathcal P^{2\beta}_0[\vect X_\eta]
+  + \mo(\varepsilon^2) 
+\end{align}
+In the case of core set identification,
+\begin{align}\nonumber
+  [\mathcal P\circ \mt_\tau] f(\vect y,t)
+  =&\,
+  \int\dd \vect x\,
+  \mt_\tau f(\vect x,t) K_{\sigma,\vect y}(\vect x)\\\nonumber
+  =&\,
+  \int\dd \vect x\,
+  e^{\tau\fwg_t^{2\beta}} f(\vect x,t) \,K_{\sigma,\vect y}(\vect x)\\\nonumber
+  =&\,
+  \int\dd \vect x\,
+  f(\vect x,t) \,e^{\tau\bwg_t^{2\beta}}K_{\sigma,\vect y}(\vect x)  \\\nonumber
+  =&\,
+  \int\dd \vect x\,
+  f(\vect x,t) \,\mathbb E_{\vect x}
+  \big[
+  K_{\sigma,\vect y}(\vect X^{2\beta}_{t,\tau})
+  \big]  
+  \\
+  =&\,
+  \int\dd \vect x\,
+  f(\vect x,t) 
+  \int_{\mc\{\vect x,0;\tau\}} K_{\sigma,\vect y}(\vect X_\eta)\,
+  \dd\mathcal P^{2\beta}_t[\vect X_\eta]
+\end{align}
+So the observable $O[\vect X_\eta]$
+is:
+\begin{align}
+  O[\vect X_\eta] = 
+  K_{\sigma,\vect y}(\vect X_\tau),
+  % -
+  % K_{\sigma,\vect y}(\vect x)
+\end{align}
+which is the end point of the artificial
+trajectory measured by $K_{\sigma,\vect y}$.
+Inserting everything we known in~\eqref{eqn:solve-tau-linear-2-0},
+and notice $G[\vect X_\eta] = 0$ when $\tau = 0$,
+\begin{align}\nonumber
+  [\mathcal P\circ\Delta\mt_\tau f](\vect y, t)
+  =
+  \lim_{\sigma \rightarrow 0}
+  \big\{
+  &\,
+  [\mathcal P\circ\mt_\tau f - \mathcal P\circ\mt_0 f](\vect y, t)
+  \big\}\\\nonumber
+  \approx
+  \lim_{\sigma \rightarrow 0}
+  \bigg\{
+  &\,
+  \int \dd \vect x\,
+  f_\inv^\beta(\vect x)
+  \int_{\mc\{\vect x,0;\tau\}}
+  [K_{\sigma,\vect y}(\vect X_\tau)
+  -K_{\sigma,\vect y}(\vect x)]  \,
+  \dd \mathcal P^{2\beta}_0[\vect X_\eta]\\\nonumber
+  &\,-
+  \beta
+  \int_0^t\dd s\, F_e(s)
+  \int \dd \vect x\,
+  e^{(t-s)\fwg_0^\beta}
+  \big[
+  j(\vect x)
+  f_\inv^\beta(\vect x)
+  \big]
+  \int_{\mc\{\vect x,0;\tau\}}
+  [K_{\sigma,\vect y}(\vect X_\tau)
+  -K_{\sigma,\vect y}(\vect x)]  \,
+  \dd \mathcal P^{2\beta}_0[\vect X_\eta]\\\nonumber
+  &\,+
+  F_e(t)
+  \int \dd \vect x\,
+  f_\inv^\beta(\vect x)
+  \int_{\mc\{\vect x,0;\tau\}}
+  K_{\sigma,\vect y}(\vect X_\tau)
+  % \Big\{
+  % \int_0^\tau\dd \eta\,
+  % [\dot{\vect p}_\eta + \gamma\vect p_\eta + \nabla_{\vect q}\mh(\vect X_\eta)]
+  % \cdot\vect D(\vect X_\eta)
+  % \Big\}\,
+  \,G[\vect X_\eta]\,
+  \dd \mathcal P^{2\beta}_0[\vect X_\eta]
+  \bigg\}\\\nonumber
+  \approx
+  \lim_{\sigma \rightarrow 0}
+  \bigg\{
+  &\,
+  \Big\langle
+  \mathbb E_{\vect x}[K_{\sigma,\vect y}(\vect X_\tau^{2\beta})]
+  -
+  K_{\sigma,\vect y}(\vect x)
+  \Big\rangle\\\nonumber
+  &\,
+  - \beta
+  \int_0^t\dd s\, F_e(s)
+  \Big\langle
+  j(\vect x)
+  \mathbb E_{\vect x} \big\{
+  \mathbb E_{\vect X^\beta_{t-s}}\big[
+  K_{\sigma, \vect y}(\vect X^{2\beta}_\tau)
+  \big]
+  -
+  K_{\sigma, y}(\vect X^\beta_{t-s})
+  \big\}
+  \Big\rangle\\\label{eqn:core-identify-approx-2}
+  &\,+
+  F_e(t)
+  \int \dd \vect x\,
+  f_\inv^\beta(\vect x)
+  \int_{\mc\{\vect x,0;\tau\}}
+  K_{\sigma,\vect y}(\vect X_\tau)
+  \,G[\vect X_\eta]\,
+  \dd \mathcal P^{2\beta}_0[\vect X_\eta]
+  \bigg\}
+\end{align}
+Comparing~\eqref{eqn:core-identify-approx-2}
+with~\eqref{eqn:core-identify-approx-1}, the first
+two terms on the r.h.s. are identical. 
+The last term is now computable.
+
+\subsection{Numerical test}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.4\textwidth]{figs/fig-pot.eps}
+  \caption{The potential with and without perturbation.}
+  \label{fig:tmp1}
+\end{figure}
+
+We test the idea by a one-dimensional model system: one particle in a
+double-well potential. For convenience, we let the mass of the
+particle to be 1 \textsf{amu}. The Hamiltonian of the system is given
+by:
+\begin{align}
+  \mh (\vect p, \vect q) = \frac 12 \vect p^2 + U(\vect q) 
+\end{align}
+with
+\begin{align}
+  U(\vect q) = \frac12 k (\vect q^2 - a^2)^2
+\end{align}
+Here $k = 8$ $\textsf{kJ} / (\textsf{mol nm}^4)$, and $ a = 1 \textsf{nm}$.
+Notice at room temperature $300 \textsf{K}$, $k_BT = 2.48$ \textsf{kJ/mol}.
+See the red line in Fig.~\ref{fig:tmp1}.
+%$\textsf{kJ} / (\textsf{mol nm}^4)}$
+The perturbation is given by
+\begin{align}
+  \vect C(\vect p, \vect q) = 0; \qquad
+  \vect D(\vect p, \vect q) = -\nabla_{\vect q} V(\vect q) = 1
+\end{align}
+Here $V(\vect q) = -\vect q$, which effectively tilt the original
+potential $U(\vect q)$. The strength of the perturbation $F_e(t)$ is given
+by the following function:
+\begin{align}
+  F_e(t) = 
+  \begin{cases}
+    F_e^{\textrm{max}}\times (t / t_c) & t < t_c \\
+    F_e^{\textrm{max}} & t \geq t_c
+  \end{cases}
+\end{align}
+$t_c = 100$ \textsf{ps} so that the perturbation increases slow
+enough: the typical decaying time scale of the correlation function
+in~\eqref{eqn:main} is roughly 3 \textsf{ps}.  We choose $\mt_\alpha$
+to be the propagator of the Langevin dynamics at temperature $150$
+\textsf{K} with $\alpha = 1$ \textsf{ps}.  Since the dynamics is in
+1-d, the projection is identity.  We consider the stability of the
+distribution $f_t$ at time $t$: $-\Delta\mt_\alpha f_t$. The smaller
+this value, the more stable the core sets are.
+
+See Fig. \ref{fig:tmp2} and \ref{fig:tmp3} for numerical result.  Both
+the direct non-equilibrium simulation and the linear response
+approximation are presented. Time slices $t = 0$, 30, 60 and 100
+\textsf{ps} are shown.  When the perturbation is small,
+i.e. $F_e^{\textrm{max}} = 2$, the direct non-equilibrium simulation
+and the linear response results consistent very well. As the
+perturbation grows stronger, the left well becomes less stable, while
+the right well becomes more stable.  When the perturbation is big,
+i.e. $F_e^{\textrm{max}} = 6$, the linear response results are no
+longer precise. When $t\geq 60$ \textsf{ps}, the left well basically
+disappears, however, the linear response calculation present a well
+that is nearly as stable as the right well.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.95\textwidth]
+  {figs/warm100.str2.0.nst1e09.smallgrid/fig-2d.eps}
+  \caption{The stability of distribution $f_t$ under perturbation of
+    $F_e^{\textrm{max}} = 2$.
+    Upper row: the direct non-equilibrium simulation. Lower row: the
+    results of linear response formula~\eqref{eqn:main}.
+  }
+  \label{fig:tmp2}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.95\textwidth]
+  {figs/warm100.str6.0.nst1e09.smallgrid/fig-2d.eps}
+  \caption{The stability of distribution $f_t$ under perturbation of
+    $F_e^{\textrm{max}} = 6$.  Upper row: the direct non-equilibrium
+    simulation. Lower row: the results of linear response
+    formula~\eqref{eqn:main}.  }
+  \label{fig:tmp3}
+\end{figure}
+
+\section{Non-equilibrium linear response}
+
+% \subsection{Path integral}
+% The path integral is very useful to develop the non-equilibrium version of
+% linear response theory. 
+% Comparing with the idea of ensemble, which is a collection of state points
+% in the configuration space, the path integral consider the ``ensemble'' of
+% trajectories.
+% (In the following, the develop of formula are formal. The strict mathematical
+% theory is still unclear for me.)
+% First, we take the standard 1-dimensional Wiener process for example.
+% Consider the time interval $[0,T]$, the probability density of one trajectory
+% is given by
+% \begin{align}
+%   p[W_t] = \frac 1Z e^{-I[W_t]}
+% \end{align}
+% where $I$ is the action functional of Wiener process given by
+% \begin{align}
+%   I[W_t] = \int_0^T\frac12 \dot{W}_t^2\dd t
+% \end{align}
+% the partition function $Z$ is
+% \begin{align}
+%   Z = \int  e^{-I[W_t]} \prod_{0\leq t\leq T}\dd W_t
+% \end{align}
+% Consider the following two stochastic processes:
+% \begin{align}\label{eqn:st-orig}
+%   \dd \vect X_t &= \vect b(\vect X_t, t) \dd t + \bsigma(\vect X_t, t)\,\dd\vect W_t \\\label{eqn:st-pert}
+%   \dd \vect Y_t &= \vect b(\vect Y_t, t) \dd t + \vect c(\vect Y_t, t) \dd t + \bsigma(\vect Y_t, t)\,\dd\vect W_t
+% \end{align}
+% Where $\vect W_t$ is the standard Wiener process. Obviously
+% \eqref{eqn:st-pert} is the perturbed version of
+% \eqref{eqn:st-orig}. It has been shown that, starting from the same
+% measure saying $\mu$ (which is not necessarily equilibrium or
+% stationary), the ratio of trajectory probability density between the perturbed
+% and unperturbed process is~\cite{li2010lecture} (\recheck{I think here
+%   the incompressibility of phase space should be assumed. Otherwise the
+%   ratio should be more complicated.})
+% \begin{align}\label{eqn:girsanov}
+%   \frac{\dd\mathcal P_{\vect Y}}{\dd\mathcal P_{\vect X}}
+%   =
+%   \exp
+%   \Big\{
+%   -\frac12 \int_0^T\vert\bsigma^{-1}\cdot\vect c(\vect X_t,t)\vert^2\,\dd t
+%   +\int_0^T[\dot{\vect X}_t - \vect b(\vect X_t,t)]^T
+%   \cdot\bsigma^{-T}\cdot\bsigma^{-1}\cdot\vect c(\vect X_t,t)\,\dd t 
+%   \Big\}
+% \end{align}
+% where $\mathcal P_{\vect Y}$ and $\mathcal P_{\vect Y}$ are measures
+% of the trajectories generated by \eqref{eqn:st-pert} and
+% \eqref{eqn:st-orig}, respectively.  This result (Girsanov formula?) is
+% (formally) precise without any approximation.
+
+\subsection{Non-equilibrium perturbation with linear approximation}
+
+To be more general, we study the two perturbed Langevin dynamics
+(i.e. Eqn.~\eqref{eqn:langevin-1} and \eqref{eqn:langevin-2}) with
+different perturbations that are given by $F_e^1(t)$ and $F_e^2(t)$,
+respectively. We consider the difference between the observations of
+these dynamics:
+\begin{align}\label{eqn:noneq-pert-1}
+  \langle Q(t)\rangle_{\mu}^{2}  -\langle Q(t)\rangle_{\mu}^{1}
+  &=
+  \int Q(\vect X_t) \,\dd\mathcal P^2[\vect X_t]
+  -\int Q(\vect X_t) \,\dd\mathcal P^1[\vect X_t]
+  =
+  \int Q(\vect X_t)
+  \Big\{
+  \frac{\dd\mathcal P^2[\vect X_t]}{\dd\mathcal P^1[\vect X_t]}
+  -1
+  \Big\}
+  \,\dd\mathcal P^1[\vect X_t]
+\end{align}
+\recheck{Here we further assume that $\vect C(\vect q, \vect p) = 0$.
+  (otherwise the Girsanov formula~\eqref{eqn:girsanov} may be unstable...
+  By letting the $\epsilon$ in the following formula to 0, if
+  $\vect C\neq 0$, the limit may not be exist. This
+  is intuitionally not likely... I do not know why.)}
+By writing Eqn.~\eqref{eqn:langevin-1} and \eqref{eqn:langevin-2} as
+\begin{align}
+  \frac{\dd }{\dd t}
+  \left[
+    \begin{array}{l}
+      \vect q\\
+      \vect p
+    \end{array}
+  \right]
+  =
+  \left[
+    \begin{array}{l}
+      \nabla_{\vect p}\mh\\
+      -\nabla_{\vect q}\mh - \gamma\vect p + F_e\vect D
+    \end{array}
+  \right]
+  \dd t
+  +
+  \left[
+    \begin{array}{cc}
+      \epsilon \vect I & 0\\
+      0 & \sigma\vect I
+    \end{array}
+  \right]
+  \cdot
+  \left[
+    \begin{array}{l}
+      \dd\vect W_t\\
+      \dd\vect W_t
+    \end{array}
+  \right]
+\end{align}
+Using Girsanov's formula, to the leading order of $\Delta F_e = F_e^2 - F_e^1$,
+we have
+\begin{align}\nonumber
+  \frac{\dd\mathcal P^2[\vect X_t]}{\dd\mathcal P^1[\vect X_t]}
+  -1
+  &=
+  \exp
+  \Big\{
+  \frac 1 {\sigma^2}
+  \int_0^t
+  [\dot{\vect p}_s +
+  \gamma\vect p_s +
+  \nabla_{\vect q}\mh(\vect q_s,\vect p_s)
+  - F_e^1(s) \vect D(\vect q_s,\vect p_s)
   ]\cdot
@@
   \nabla_{\vect q}\mh(\vect q_s,\vect p_s)
+  - F_e^1(s) \vect D(\vect q_s,\vect p_s)
   ]\cdot

### 2012-07-22 writedown/linear.response/langevin.tex
new figs

@@
 
-\subsection{Numerical test}
+\subsection{Numerical tests}
+
+\subsubsection{The tilting double-well potential}
 
@@
   \centering
-  \includegraphics[width=0.4\textwidth]{figs/fig-pot.eps}
-  \caption{The potential with and without perturbation.}
+  \includegraphics[width=0.4\textwidth]{figs/fig-tilt-pot.eps}
+  \caption{The double-well potential with tilting perturbation.}
   \label{fig:tmp1}
@@
 
-We test the idea by a one-dimensional model system: one particle in a
-double-well potential. For convenience, we let the mass of the
-particle to be 1 \textsf{amu}. The Hamiltonian of the system is given
-by:
+We test the idea firstly by a one-dimensional model system: one particle in a
+tilting double-well potential. For convenience, we let the mass of the
+particle to be 1 \textsf{amu}. The unperturbed
+Hamiltonian of the system is given by:
 \begin{align}
@@
 \end{align}
-with
+with potential
 \begin{align}
@@
 \end{align}
-Here $k = 8$ $\textsf{kJ} / (\textsf{mol nm}^4)$, and $ a = 1 \textsf{nm}$.
-Notice at room temperature $300 \textsf{K}$, $k_BT = 2.48$ \textsf{kJ/mol}.
+Here $k = 8$~$\textsf{kJ} / (\textsf{mol nm}^4)$, and $ a = 1\ \textsf{nm}$.
+Notice at room temperature $300\ \textsf{K}$, $k_BT = 2.48$~\textsf{kJ/mol}.
 See the red line in Fig.~\ref{fig:tmp1}.
@@
 \end{align}
-Here $V(\vect q) = -\vect q$, which effectively tilt the original
+Here $V(\vect q) = -\vect q$ is  effectively tilt the original
 potential $U(\vect q)$. The strength of the perturbation $F_e(t)$ is given
@@
 \end{align}
-$t_c = 100$ \textsf{ps} so that the perturbation increases slow
+$t_c = 20$~\textsf{ps} so that the perturbation increases slow
 enough: the typical decaying time scale of the correlation function
-in~\eqref{eqn:main} is roughly 3 \textsf{ps}.  We choose $\mt_\alpha$
-to be the propagator of the Langevin dynamics at temperature $150$
-\textsf{K} with $\alpha = 1$ \textsf{ps}.  Since the dynamics is in
-1-d, the projection is identity.  We consider the stability of the
-distribution $f_t$ at time $t$: $-\Delta\mt_\alpha f_t$. The smaller
+in~\eqref{eqn:core-identify-approx-2} is roughly 3~\textsf{ps}.
+We choose $\tau = 1$~\textsf{ps}.
+% We choose $\mt_\alpha$
+% to be the propagator of the Langevin dynamics at temperature $150$
+% \textsf{K} with $\alpha = 1$ \textsf{ps}.
+Since the dynamics is in
+1-d, the projection $\mathcal P$ is identity.  We consider the stability of the
+distribution $f(\vect x, t)$ at time $t$: $\Delta\mt_\tau f(\vect x, t)$.
+The larger
 this value, the more stable the core sets are.
@@
 the direct non-equilibrium simulation and the linear response
-approximation are presented. Time slices $t = 0$, 30, 60 and 100
+approximation are presented. Time slices $t = 0$, 10, 15 and 20
 \textsf{ps} are shown.  When the perturbation is small,
 i.e. $F_e^{\textrm{max}} = 2$, the direct non-equilibrium simulation
-and the linear response results consistent very well. As the
+and the linear response results are accurately consistent. As the
 perturbation grows stronger, the left well becomes less stable, while
 the right well becomes more stable.  When the perturbation is big,
-i.e. $F_e^{\textrm{max}} = 6$, the linear response results are no
-longer precise. When $t\geq 60$ \textsf{ps}, the left well basically
-disappears, however, the linear response calculation present a well
-that is nearly as stable as the right well.
+i.e. $F_e^{\textrm{max}} = 4$, the linear response results are no
+longer precise. When $t\geq 10$ \textsf{ps}, the left well basically
+disappears, however, the linear response calculation presents
+artifical occilation at the right well.
 
@@
   \includegraphics[width=0.95\textwidth]
-  {figs/warm100.str2.0.nst1e09.smallgrid/fig-2d.eps}
-  \caption{The stability of distribution $f_t$ under perturbation of
+  {figs/tilt.warm020.str2.0.nst1e09.smallgrid/fig-2d.eps}
+  \caption{The plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
     $F_e^{\textrm{max}} = 2$.
     Upper row: the direct non-equilibrium simulation. Lower row: the
-    results of linear response formula~\eqref{eqn:main}.
+    results of linear response formula~\eqref{eqn:core-identify-approx-2}.
   }
@@
   \includegraphics[width=0.95\textwidth]
-  {figs/warm100.str6.0.nst1e09.smallgrid/fig-2d.eps}
-  \caption{The stability of distribution $f_t$ under perturbation of
-    $F_e^{\textrm{max}} = 6$.  Upper row: the direct non-equilibrium
+  {figs/tilt.warm020.str4.0.nst1e09.smallgrid/fig-2d.eps}
+  \caption{The plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
+    $F_e^{\textrm{max}} = 4$.  Upper row: the direct non-equilibrium
     simulation. Lower row: the results of linear response
-    formula~\eqref{eqn:main}.  }
+    formula~\eqref{eqn:core-identify-approx-2}.  }
   \label{fig:tmp3}
+\end{figure}
+
+\subsubsection{Splitting single-well potential}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.4\textwidth]{figs/fig-split-pot.eps}
+  \caption{The single-well potential with splitting perturbation.}
+  \label{fig:tmp4}
+\end{figure}
+
+In this splitting single-well potential example, the unperturbed
+potential is simply the harmonic potential given by:
+\begin{align}
+  U(\vect q) = \frac12\,k\,\vect q^2 
+\end{align}
+Here $k = 8$~$\textsf{kJ} / (\textsf{mol nm}^2)$.
+Notice at room temperature $300\ \textsf{K}$, $k_BT = 2.48$~\textsf{kJ/mol}.
+See the red line in Fig.~\ref{fig:tmp4}.
+%$\textsf{kJ} / (\textsf{mol nm}^4)}$
+The perturbation is again given by
+\begin{align}
+  \vect C(\vect p, \vect q) = 0; \qquad
+  \vect D(\vect p, \vect q) = -\nabla_{\vect q} V(\vect q) 
+\end{align}
+where $V(\vect q)$ is a Gaussian type potential:
+\begin{align}
+  V(\vect q) = \frac{1}{\sqrt{2\pi \sigma^2}}
+  \exp\Big\{-\frac{\vect q^2}{2\sigma^2}\Big\}
+\end{align}
+we use $\sigma = 0.16$~\textsf{nm}. Please see Fig~\ref{fig:tmp4}.
+For the results, please see Fig.~\ref{fig:tmp5} -- Fig.~\ref{fig:tmp6}.
+\begin{figure}
+  \centering
+  \includegraphics[width=0.95\textwidth]
+  {figs/split.warm020.str1.0.nst1e09.smallgrid.more/fig-2d.eps}
+  \caption{The plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
+    $F_e^{\textrm{max}} = 1.0$.
+    Upper row: the direct non-equilibrium simulation. Lower row: the
+    results of linear response formula~\eqref{eqn:core-identify-approx-2}.
+  }
+  \label{fig:tmp5}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.95\textwidth]
+  {figs/split.warm020.str2.0.nst1e09.smallgrid.more/fig-2d.eps}
+  \caption{The plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
+    $F_e^{\textrm{max}} = 2.0$.  Upper row: the direct non-equilibrium
+    simulation. Lower row: the results of linear response
+    formula~\eqref{eqn:core-identify-approx-2}.  }
+  \label{fig:tmp6}
 \end{figure}

### 2012-07-24 writedown/linear.response/langevin.tex
change notation for the projection operator

@@
 \newcommand{\mc}{\mathcal C}
+\newcommand{\proj}{\mathit\Pi}
 \newcommand{\fwg}{{\mathcal A}}
@@
 
-\subsection{The core set as an observation of the system}
+\subsection{The linear response theory for the core set identification}
 
@@
 \begin{align}
-  \mathcal P f (\vect y, t)
+  \proj f (\vect y, t)
   =
@@
   \vect y \,\vert\,
-  \mathcal P\mt_\tau f(\vect y, t) > \mathcal P f(\vect y, t)
+  \proj\mt_\tau f(\vect y, t) > \proj f(\vect y, t)
   \big\}
@@
 \mt_\tau - \mathcal I$, where $\mathcal I$ is the
-identity. If $[\mathcal P\circ\Delta \mt_\tau f](\vect y, t) > 0$, then
+identity. If $[\proj\circ\Delta \mt_\tau f](\vect y, t) > 0$, then
 $\vect y$ is in the core set $\widetilde C_\tau(t) $.
 \begin{align}\nonumber
-  [\mathcal P\circ\Delta \mt_\tau f](\vect y, t)
+  [\proj\circ\Delta \mt_\tau f](\vect y, t)
   &=
@@
 \begin{align}\nonumber
-  [\mathcal P\circ\Delta\mt_\tau f](\vect y, t)
+  [\proj\circ\Delta\mt_\tau f](\vect y, t)
   \approx
@@
 \begin{align}
-  \frac{\dd }{\dd t}
   \left[
     \begin{array}{l}
-      \vect q\\
-      \vect p
+      \dd\vect q\\
+      \dd\vect p
     \end{array}
@@
 
-\subsection{Linear response theory of the core set identification,
-  revisited by the path integral method}
+\subsection{The linear response theory for the core set identification,
+  revisited by the path integral}
 For some observable $O[\vect X_\eta]$, which is a functional of the
@@
 \begin{align}
-  G[\vect X_\eta] =
+  G[\vect X_\eta] &=
     \frac 1 {\sigma^2}
@@
   \vect D(\vect q_\eta,\vect p_\eta)\,
-  \dd \eta
-\end{align}
+  \dd \eta =
+  \frac1\sigma\int_0^\tau
+  \vect D(\vect q_\eta,\vect p_\eta)
+  \cdot\dd \vect W_\eta
+\end{align}
+where $\vect W_\eta$ is the Wiener process that generates
+$\vect X_\eta$.
 Therefore, we have
@@
 \begin{align}\nonumber
-  [\mathcal P\circ \mt_\tau] f(\vect y,t)
+  [\proj\circ \mt_\tau] f(\vect y,t)
   =&\,
@@
   f(\vect x,t) 
-  \int_{\mc\{\vect x,0;\tau\}} K_{\sigma,\vect y}(\vect X_\eta)\,
+  \int_{\mc\{\vect x,0;\tau\}} K_{\sigma,\vect y}(\vect X_\tau)\,
   \dd\mathcal P^{2\beta}_t[\vect X_\eta]
@@
 \begin{align}\nonumber
-  [\mathcal P\circ\Delta\mt_\tau f](\vect y, t)
+  [\proj\circ\Delta\mt_\tau f](\vect y, t)
   =
@@
   &\,
-  [\mathcal P\circ\mt_\tau f - \mathcal P\circ\mt_0 f](\vect y, t)
+  [\proj\circ\mt_\tau f - \proj\circ\mt_0 f](\vect y, t)
   \big\}\\\nonumber

### 2012-07-24 writedown/linear.response/langevin.tex
add non linear response theory

@@
 
-\subsection{Numerical tests}
-
-\subsubsection{The tilting double-well potential}
+\subsection{General response theory for core set identification}
+
+Here, we assume that the system undergoes a series of perturbations,
+which are denoted by $F^0_e(t),\ F^1_e(t),\ \cdots,\ F^i_e(t),\ \cdots$.
+Here, $F^0_e(t) = 0$.
+The measure for the trajectory space
+$\mc\{\vect x, 0; t\}$
+is given by
+$\mathcal P_0^\beta[\vect X_s],\ \mathcal P_1^\beta[\vect X_s],\ \cdots,\
+\mathcal P_i^\beta[\vect X_s], \cdots$, respectively.
+The measure for the quenching trajectory are denoted by
+$\mathcal P_{0,t}^{2\beta}[\vect X_\eta],\
+\mathcal P_{1,t}^{2\beta}[\vect X_\eta],\ \cdots,\
+\mathcal P_{i,t}^{2\beta}[\vect X_\eta], \cdots$, respectively.
+The notation $t$ on the subscript means the trajectory is branched from
+the main trajectory at time $t$.
+Now the phase space distribution $f_i(\vect x, t)$ for the
+$i$-th perturbation can be studied in terms
+of path integral:
+\begin{align}
+  f_i(\vect x, t) =
+  \int\dd\vect u\,
+  f_i(\vect u, 0)
+  \int_{\mc\{\vect u,0; \vect x,t\}}
+  \dd\mathcal P_i^\beta[\vect X_s]
+\end{align}
+Applied by operator $\mt_\tau$, we have
+\begin{align}\nonumber
+  \mt_\tau f_i(\vect x, t)
+  =&\,
+  \int\dd\vect u f_i(\vect u, t)
+  \int_{\mc\{\vect u,0; \vect x,\tau\}}
+  \dd\mathcal P_{i,t}^{2\beta}[\vect X_\eta]\\
+  =&\,
+  \int\dd\vect v\,
+  f_i(\vect v, 0)
+  \int\dd\vect u
+  \int_{\mc\{\vect v,0; \vect u,t\}}
+  \dd\mathcal P_i^\beta[\vect X_s]  
+  \int_{\mc\{\vect u,0; \vect x,\tau\}}
+  \dd\mathcal P_{i,t}^{2\beta}[\vect X_\eta]
+\end{align}
+For the core set identification,
+\begin{align}
+  [\proj\circ\mt_\tau f](\vect y, t)
+  &=
+  \int\dd\vect v\,
+  f_i(\vect v, 0)
+  \int\dd\vect u
+  \int_{\mc\{\vect v,0; \vect u,t\}}
+  \dd\mathcal P_i^\beta[\vect X_s]  
+  \int_{\mc\{\vect u,0; \tau\}}
+  K_{\sigma,\vect y}(\vect X_\tau)\,
+  \dd\mathcal P_{i,t}^{2\beta}[\vect X_\eta]  \\
+  [\proj\circ\mt_0 f](\vect y, t)
+  &=
+  \int\dd\vect v\,
+  f_i(\vect v, 0)
+  \int_{\mc\{\vect v,0; t\}}
+  K_{\sigma,\vect y}(\vect X_t)\,
+  \dd\mathcal P_i^\beta[\vect X_s]  
+  % \int_{\mc\{\vect u,0; \tau\}}
+  % K_{\sigma,\vect y}(\vect X_\tau)\,
+  % \dd\mathcal P_{i,t}^{2\beta}[\vect X_\eta]  
+\end{align}
+Here we assume that the system always starts from invariant measure,
+i.e. $f_i(\vect x, 0) = f_{\inv}^\beta(\vect x)$, and that we have a
+good sampling of the $(i-1)$-th perturbation, then
+\begin{align}
+  [\proj\circ\mt_\tau f](\vect y, t)
+  &=
+  \int\dd\vect v\,
+  f_\inv^\beta(\vect v)
+  \int\dd\vect u
+  \int_{\mc\{\vect v,0; \vect u,t\}}
+  \dd\mathcal P_{i-1}^\beta[\vect X_s]  
+  \int_{\mc\{\vect u,0; \tau\}}
+  \dd\mathcal P_{i-1,t}^{2\beta}[\vect X_\eta]\cdot
+  K_{\sigma,\vect y}(\vect X_\tau)\,
+  \frac
+  {\dd\mathcal P_{i}^\beta[\vect X_s]}
+  {\dd\mathcal P_{i-1}^\beta[\vect X_s]}
+  \frac
+  {\dd\mathcal P_{i,t}^{2\beta}[\vect X_\eta]}
+  {\dd\mathcal P_{i-1,t}^{2\beta}[\vect X_\eta]}
+  \\
+  [\proj\circ\mt_0 f](\vect y, t)
+  &=
+  \int\dd\vect v\,
+  f_{\inv}^\beta(\vect v)
+  \int_{\mc\{\vect v,0; t\}}
+  \dd\mathcal P_{i-1}^\beta[\vect X_s]
+  \cdot
+  K_{\sigma,\vect y}(\vect X_t)\,
+  \frac
+  {\dd\mathcal P_{i}^\beta[\vect X_s]}
+  {\dd\mathcal P_{i-1}^\beta[\vect X_s]}  
+  % \int_{\mc\{\vect u,0; \tau\}}
+  % K_{\sigma,\vect y}(\vect X_\tau)\,
+  % \dd\mathcal P_{i,t}^{2\beta}[\vect X_\eta]  
+\end{align}
+Here
+\begin{align}\label{eqn:exact-ratio-1}
+  \frac
+  {\dd\mathcal P_{i}^\beta[\vect X_s]}
+  {\dd\mathcal P_{i-1}^\beta[\vect X_s]}  
+  =&\,
+  \exp
+  \bigg\{
+  \Delta F_e^{i-1}(s)\,
+  \frac{1}\sigma\int_0^t
+  \vect D(\vect q_s,\vect p_s)\cdot\dd\vect W_s
+  -
+  [\Delta F_e^{i-1}(s)]^2\,
+  \frac 1{2\sigma}\int_0^t
+  \vect D^2(\vect q_s,\vect p_s)\,\dd s
+  \bigg\}\\\label{eqn:exact-ratio-2}
+  \frac
+  {\dd\mathcal P_{i,t}^{2\beta}[\vect X_\eta]}
+  {\dd\mathcal P_{i-1,t}^{2\beta}[\vect X_\eta]}
+  =&\,
+  \exp
+  \bigg\{
+  \Delta F_e^{i-1}(t)\,
+  \frac{1}\sigma\int_0^\tau
+  \vect D(\vect q_\eta,\vect p_\eta)\cdot\dd\vect W_\eta
+  -
+  [\Delta F_e^{i-1}(t)]^2\,
+  \frac 1{2\sigma}\int_0^\tau
+  \vect D^2(\vect q_\eta,\vect p_\eta)\,\dd \eta
+  \bigg\}
+\end{align}
+where $\Delta F_e^{i-1}(t) = F_e^{i}(t) - F_e^{i-1}(t) $.
+If the trajectory at perturbation $ F_e^{i-1}(t) $ is well sampled, then
+\eqref{eqn:exact-ratio-1} and \eqref{eqn:exact-ratio-2} can be measured
+without linear approximation.
+
+\section{Numerical tests}
+
+\subsection{The tilting double-well potential}
 
@@
 
-\subsubsection{Splitting single-well potential}
+\subsection{Splitting single-well potential}
 

### 2012-07-26 writedown/linear.response/langevin.tex
fix bugs
add non-linear response

@@
 
-\subsection{General response theory for core set identification}
+\subsection{Higher orders response theory for core set identification}
 
@@
 good sampling of the $(i-1)$-th perturbation, then
-\begin{align}
+\begin{align}\label{eqn:pert-exact-1}
   [\proj\circ\mt_\tau f](\vect y, t)
@@
   {\dd\mathcal P_{i-1,t}^{2\beta}[\vect X_\eta]}
-  \\
+  \\\label{eqn:pert-exact-2}
   [\proj\circ\mt_0 f](\vect y, t)
@@
   \bigg\{
-  \Delta F_e^{i-1}(s)\,
   \frac{1}\sigma\int_0^t
-  \vect D(\vect q_s,\vect p_s)\cdot\dd\vect W_s
-  -
-  [\Delta F_e^{i-1}(s)]^2\,
+  \Delta F_e^{i-1}(s)\,\vect D(\vect q_s,\vect p_s)\cdot\dd\vect W_s
+  -
   \frac 1{2\sigma}\int_0^t
-  \vect D^2(\vect q_s,\vect p_s)\,\dd s
+  [\Delta F_e^{i-1}(s)]^2\,\vect D^2(\vect q_s,\vect p_s)\,\dd s
   \bigg\}\\\label{eqn:exact-ratio-2}
@@
   \bigg\{
-  \Delta F_e^{i-1}(t)\,
   \frac{1}\sigma\int_0^\tau
-  \vect D(\vect q_\eta,\vect p_\eta)\cdot\dd\vect W_\eta
-  -
-  [\Delta F_e^{i-1}(t)]^2\,
+  \Delta F_e^{i-1}(t)\,\vect D(\vect q_\eta,\vect p_\eta)\cdot\dd\vect W_\eta
+  -
   \frac 1{2\sigma}\int_0^\tau
-  \vect D^2(\vect q_\eta,\vect p_\eta)\,\dd \eta
+  [\Delta F_e^{i-1}(t)]^2\,\vect D^2(\vect q_\eta,\vect p_\eta)\,\dd \eta
   \bigg\}
@@
 without linear approximation.
+This is the exact result, from where we can develop
+the response (also non-linear) theroy.
+Before doing so, we should assume the perturbation can be expanded,
+and well approximated in a finite dimensional space:
+\begin{align}
+  F^i_e(t)
+  =
+  \sum_{j = 1}^\infty
+  s^i_j M_j(t)
+\end{align}
+with this expansion, we have
+\begin{align}\label{eqn:exact-ratio-11}
+  \frac
+  {\dd\mathcal P_{i}^\beta[\vect X_s]}
+  {\dd\mathcal P_{i-1}^\beta[\vect X_s]}  
+  =&\,
+  \exp
+  \bigg\{
+  \sum_{j=1}^\infty
+  \Delta s_j^{i-1}G_j[\vect X_s]
+  -
+  \frac12 \sum_{j,k=1}^\infty
+  \Delta s_j^{i-1} \Delta s_k^{i-1}H_{jk}[\vect X_s]
+  \bigg\}\\\label{eqn:exact-ratio-21}
+  \frac
+  {\dd\mathcal P_{i,t}^{2\beta}[\vect X_\eta]}
+  {\dd\mathcal P_{i-1,t}^{2\beta}[\vect X_\eta]}
+  =&\,
+  \exp
+  \bigg\{
+  \Delta F_e^{i-1}(t)\,
+  G_0[\vect X_\eta]
+  -
+  \frac12[\Delta F_e^{i-1}(t)]^2
+  H_{00}[\vect X_\eta]
+  \bigg\}
+\end{align}
+where
+\begin{align}\label{eqn:Gj}
+  G_j[\vect X_s]
+  =&\,
+  \frac 1\sigma
+  \int_0^t
+  M_j(s)\,\vect D(\vect X_s)\cdot\dd\vect W_s\\\label{eqn:Hjk}
+  H_{jk}[\vect X_s]
+  =&\,
+  \frac 1{\sigma}
+  \int_0^t
+  M_j(s)\,M_k(s)\,\vect D^2(\vect X_s)\cdot\dd s
+\end{align}
+and 
+\begin{align}\label{eqn:G0}
+  G_0[\vect X_\eta]
+  =&\,
+  \frac 1\sigma
+  \int_0^\tau
+  M_j(\eta)\,\vect D(\vect X_\eta)\cdot\dd\vect W_\eta\\\label{eqn:H00}
+  H_{00}[\vect X_\eta]
+  =&\,
+  \frac 1{\sigma}
+  \int_0^\tau
+  M_j(\eta)\,M_k(\eta)\,\vect D^2(\vect X_\eta)\cdot\dd \eta
+\end{align}
+Notice that Eqn.~\eqref{eqn:Gj}--\eqref{eqn:H00} are perturbation
+independent, therefore can be measured with the dynamics $i-1$.
+Expanding \eqref{eqn:exact-ratio-11} and \eqref{eqn:exact-ratio-21}
+up to $\mathcal O(\varepsilon^3)$ yields
+\begin{align}\label{eqn:expan-ratio-1}
+  \frac
+  {\dd\mathcal P_{i}^\beta[\vect X_s]}
+  {\dd\mathcal P_{i-1}^\beta[\vect X_s]}
+  =&\,
+  1
+  +
+  \sum_{j=1}^\infty
+  \Delta s_j^{i-1}G_j[\vect X_s]
+  -
+  \frac12
+  \sum_{j,k=1}^\infty
+  \Delta s_j^{i-1} \Delta s_k^{i-1}H_{jk}[\vect X_s]
+  +
+  \frac12  
+  \sum_{j,k=1}^\infty
+  \Delta s_j^{i-1}\Delta s_k^{i-1}G_j[\vect X_s]G_k[\vect X_s]
+  +
+  \mathcal O(\varepsilon^3)
+  \\\label{eqn:expan-ratio-2}
+  \frac
+  {\dd\mathcal P_{i,t}^{2\beta}[\vect X_\eta]}
+  {\dd\mathcal P_{i-1,t}^{2\beta}[\vect X_\eta]}
+  =&\,
+  1 +
+  \Delta F_e^{i-1}(t)\,
+  G_0[\vect X_\eta]
+  -
+  \frac12[\Delta F_e^{i-1}(t)]^2
+  H_{00}[\vect X_\eta]
+  +
+  \frac12[\Delta F_e^{i-1}(t)]^2
+  \big\{
+  G_0[\vect X_\eta]
+  \big\}^2
+  +
+  \mathcal O(\varepsilon^3)
+\end{align}
+Therefore, by the expansion formula \eqref{eqn:expan-ratio-1} and
+\eqref{eqn:expan-ratio-2}, the response theory for the
+\eqref{eqn:pert-exact-1} and \eqref{eqn:pert-exact-2} can be
+established.
 

### 2012-07-27 writedown/linear.response/langevin.tex
add non-linear response code. Fix bug in text

@@
   \int_0^\tau
-  M_j(\eta)\,\vect D(\vect X_\eta)\cdot\dd\vect W_\eta\\\label{eqn:H00}
+  \vect D(\vect X_\eta)\cdot\dd\vect W_\eta\\\label{eqn:H00}
   H_{00}[\vect X_\eta]
@@
   \int_0^\tau
-  M_j(\eta)\,M_k(\eta)\,\vect D^2(\vect X_\eta)\cdot\dd \eta
+  \vect D^2(\vect X_\eta)\cdot\dd \eta
 \end{align}
@@
 established.
+For convenience, we denote
+\begin{align}
+  \big\langle
+  O[\vect X_s]\tilde O[\vect X_\eta]
+  \big\rangle_{0,t,\tau}^{i-1}
+  =&\,
+  \int\dd\vect v\,
+  f_\inv^\beta(\vect v)
+  \int\dd\vect u
+  \int_{\mc\{\vect v, 0; \vect u, t\}}
+  \dd\mathcal P^\beta_{i-1}[\vect X_s]
+  \,
+  O[\vect X_s]
+  \int_{\mc\{\vect u, 0; \tau\}}
+  \dd\mathcal P^{2\beta}_{i-1,t}[\vect X_\eta]
+  \,
+  \tilde O[\vect X_\eta]
+  \\
+  \big\langle
+  O[\vect X_s]
+  \big\rangle_{0,t,\tau}^{i-1}
+  = &\,
+  \int\dd\vect v\,
+  f_\inv^\beta(\vect v)
+  \int_{\mc\{\vect v, 0; t\}}
+  \dd\mathcal P^\beta_{i-1}[\vect X_s]
+  \,
+  O[\vect X_s]  
+\end{align}
+Then the approximated \eqref{eqn:pert-exact-1} is
+\begin{align}\nonumber
+  [\proj\circ\mt_\tau f](\vect y, t)
+  =&\,
+  \big\langle
+  K_{\sigma,\vect y}(\vect X_\tau)
+  \big\rangle_{0,t,\tau}^{i-1}
+  \\ \nonumber
+  &\, +
+  \Delta F_e^{i-1}(t) 
+  \big\langle
+  K_{\sigma,\vect y}(\vect X_\tau)\,
+  G_0[\vect X_\eta]
+  \big\rangle_{0,t,\tau}^{i-1}
+  +
+  \sum_{j=1}^\infty
+  \Delta s_j^{i-1}
+  \big\langle
+  K_{\sigma,\vect y}(\vect X_\tau)\,
+  G_j[\vect X_s]
+  \big\rangle_{0,t,\tau}^{i-1}
+  \\\nonumber
+  &\, 
+  -
+  \frac12[\Delta F_e^{i-1}(t)]^2
+  \big\langle
+  K_{\sigma,\vect y}(\vect X_\tau)\,
+  H_{00}[\vect X_\eta]
+  \big\rangle_{0,t,\tau}^{i-1}
+  -
+  \frac 12
+  \sum_{j,k=1}^\infty
+  \Delta s_j^{i-1}\Delta s_k^{i-1}
+  \big\langle
+  K_{\sigma,\vect y}(\vect X_\tau)\,
+  H_{jk}[\vect X_s]
+  \big\rangle_{0,t,\tau}^{i-1}
+  \\\label{eqn:pert-approx-1}
+  &\,+
+  \sum_{j=1}^\infty
+  \Delta F_e^{i-1}(t) \Delta s_j^{i-1}
+  \big\langle
+  K_{\sigma,\vect y}(\vect X_\tau)\,
+  G_0[\vect X_\eta]\,
+  G_j[\vect X_s]
+  \big\rangle_{0,t,\tau}^{i-1}
+\end{align}
+The first line on the r.h.s. of \eqref{eqn:pert-approx-1}
+is the zeroth order approximation,  the second line is the
+first order approximation, and the last two lines are the
+second order approximation.
+And the approximated \eqref{eqn:pert-exact-2} is
+\begin{align}
+  \nonumber
+  [\proj\circ\mt_0 f](\vect y, t)
+  =&\,
+  \big\langle
+  K_{\sigma,\vect y}(\vect X_t)
+  \big\rangle_{0,t}^{i-1}
+  \\ \nonumber
+  &\,+
+  \sum_{j=1}^\infty
+  \Delta s_j^{i-1}
+  \big\langle
+  K_{\sigma,\vect y}(\vect X_t)\,
+  G_j[\vect X_s]
+  \big\rangle_{0,t}^{i-1}
+  \\\label{eqn:pert-approx-2}
+  &\,
+  -
+  \frac 12
+  \sum_{j,k=1}^\infty
+  \Delta s_j^{i-1}\Delta s_k^{i-1}
+  \big\langle
+  K_{\sigma,\vect y}(\vect X_t)\,
+  H_{jk}[\vect X_s]
+  \big\rangle_{0,t}^{i-1}  
+\end{align}
+The first, second and third line on the r.h.s. of \eqref{eqn:pert-approx-2}
+are the zeroth order, first order and second order approximations, respectively.
 

### 2012-07-27 one.test/template/tools/simulator/app/splitSingleWell.noneqResp.cpp

@@ -1,0 +1,269 @@
+#include <algorithm>
+#include <fstream>
+#include <iomanip>
+#include <iostream>
+#include <iterator>
+#include <list>
+#include <map>
+#include <set>
+#include <sstream>
+#include <string>
+#include <queue>
+#include <vector>
+#include <cmath>
+
+#include "Force.h"
+#include "Perturbation.h"
+#include "Integrator.h"
+#include "Distribution.h"
+#include "NoneqResponseInfo.h"
+
+#include <boost/program_options.hpp>
+#include "RandomGenerator.h"
+
+namespace po = boost::program_options;
+
+int main(int argc, char * argv[])
+{
+  double gamma;
+  double kT;
+  double T;
+  double dt;
+  double pertSt0, pertSt1;
+  double sigma;
+  double kk;
+  double nst;
+  unsigned nstprint = 1;
+  double warmTime;
+  
+  double noneqTime;
+  double noneqCheckFeq;
+  double branchFeq;
+  double quenchTime;
+  double quenchTemperture;
+  double quenchkT;
+  unsigned noneqCheckNumFeq;
+  unsigned branchNumFeq;
+  double x0, x1, v0, v1;
+  unsigned nx, nv;
+      
+  unsigned long seed;
+  
+  po::options_description desc ("Allow options");
+  desc.add_options()
+      ("help,h", "print this message")
+      ("dt,d", po::value<double > (&dt)->default_value(0.0001), "time step [ps]")
+      ("nst,n", po::value<double > (&nst)->default_value(10000), "number of time step")
+      ("print-feq,p", po::value<unsigned > (&nstprint)->default_value(10000), "print frequency")
+      ("gamma,g", po::value<double > (&gamma)->default_value(1.), "gamma [ps^-1]")
+      ("temperature,t", po::value<double > (&T)->default_value(300.), "temperature [K]")
+      ("single-well-k,k",po::value<double > (&kk)->default_value(8.0), "k parameter of the double single potential [kJ/(mol nm^2)]")
+      ("gaussian-sigma,a", po::value<double > (&sigma)->default_value(0.16), "sigma of Gaussian perturbation [nm]")
[…]

### 2012-07-27 one.test/template/tools/simulator/include/NoneqResponseInfo.h

@@ -1,0 +1,82 @@
+#ifndef __NoneqResponseInfo_h_NONEQMSM_wanghan__
+#define __NoneqResponseInfo_h_NONEQMSM_wanghan__
+
+#include <vector>
+#include "Defines.h"
+#include "Distribution.h"
+#include "Perturbation.h"
+
+using namespace std;
+
+class NoneqResponseInfo 
+{
+  double x0, x1, v0, v1;
+  unsigned nx, nv;
+  
+  double dt;
+  double noneqTime;
+  double noneqCheckFeq;
+  int numCheck;
+  int checkNumFeq;
+  double quenchTime;
+  int quenchNumStep;
+
+  int countNoneq;
+  int countNoneqSeg;
+  int countQuench;
+
+  const Perturbation * ppert;
+  int numMode;
+  
+  vector<Distribution_1d >			order0;		// time
+  vector<vector<Distribution_1d > >		order1;		// time x vec
+  vector<vector<vector<Distribution_1d > > >	order2;		// time x mat  
+  
+  vector<Distribution_1d >			quench_order0;		// time
+  vector<Distribution_1d >			quench_order1_term1;	// time
+  vector<vector<Distribution_1d > >		quench_order1_term2;	// time x vec
+  vector<Distribution_1d >			quench_order2_term1;	// time
+  vector<vector<vector<Distribution_1d > > >	quench_order2_term2;	// time x mat
+  vector<vector<Distribution_1d > >		quench_order2_term3;	// time x vec
+  
+  vector<double >				Gj;		// vec 
+  vector<vector<double > >			Hjk;		// mat 
+  double					G0;		// 1
+  double					H00;		// 1
+
+  void newQuenchTraj ();
+public:
+  NoneqResponseInfo ();
+  // ~NoneqResponseInfo ();
+public:
+  void reinit (const double & x0,
+	       const double & x1,
+	       const unsigned & nx,
+	       const double & v0,
+	       const double & v1,
+	       const unsigned & nv,
+	       const double & dt,
+	       const double & noneqTime,
+	       const double & noneqCheckFeq,
+	       const double & quenchTime,
+	       const Perturbation & pert);
+  void newTraj ();
+  void depositMainTraj (const Dofs & oldx,
+			const Dofs & newx,
+			const double & sigma,
+			const Dofs & dw);
+  void depositQuenchTraj (const Dofs & oldx,
+			  const Dofs & newx,
+			  const double & sigma,
[…]

### 2012-07-27 one.test/template/tools/simulator/include/Perturbation.h

@@ -3,13 +3,21 @@
 
 #include "Defines.h"
 #include "Force.h"
+#include <vector>
+
+using namespace std;
 
 class Perturbation 
 {
 public:
+  virtual double Fe (const double & time) const = 0;
+  virtual int numMode () const;
+  virtual void FeMode     (const double & time,
+			   vector<double > & modes) const;
+  virtual void FeModePref (vector<double > & pref ) const;
+public:
   virtual void operator () (const Dofs & dofs,
   			    Dofs & pvalue) const = 0;
-  virtual double Fe (const double & time) const = 0;
   virtual void operator () (const Dofs & dofs,
 			    const double & time,
 			    Dofs & pvalue) const = 0;
@@ -37,9 +45,15 @@ class PertConstTilt : public Perturbation
 public:
   PertConstTilt (const double & s,
 		 const double warmTime = 0.);
+public:
+  virtual double Fe (const double & time) const;
+  virtual int numMode () const;
+  virtual void FeMode (const double & time,
+		       vector<double > & modes) const;
+  virtual void FeModePref (vector<double > & pref ) const;
+public:
   virtual void operator () (const Dofs & dofs,
   			    Dofs & pvalue) const;
-  virtual double Fe (const double & time) const;
   virtual void operator () (const Dofs & dofs,
 			    const double & time,
 			    Dofs & pvalue) const;
@@ -61,9 +75,15 @@ public:
 		const double m = 0.,
 		const double sig = 1.,
 		const double warmTime = 0.);
+public:
+  virtual double Fe (const double & time) const;
+  virtual int numMode () const;
+  virtual void FeMode (const double & time,
+		       vector<double > & modes) const;
+  virtual void FeModePref (vector<double > & pref ) const;
+public:
   virtual void operator () (const Dofs & dofs,
   			    Dofs & pvalue) const;
-  virtual double Fe (const double & time) const;
   virtual void operator () (const Dofs & dofs,
 			    const double & time,
 			    Dofs & pvalue) const;

### 2012-07-27 one.test/template/tools/simulator/src/NoneqResponseInfo.cpp

@@ -1,0 +1,284 @@
+#include "NoneqResponseInfo.h"
+
+NoneqResponseInfo::
+NoneqResponseInfo ()
+    : ppert (NULL)
+{
+}
+
+
+void NoneqResponseInfo::
+reinit (const double & x0_,
+	const double & x1_,
+	const unsigned & nx_,
+	const double & v0_,
+	const double & v1_,
+	const unsigned & nv_,
+	const double & dt_,
+	const double & noneqTime_,
+	const double & noneqCheckFeq_,
+	const double & quenchTime_,
+	const Perturbation & pert)
+{
+  x0 = x0_;
+  x1 = x1_;
+  nx = nx_;
+  v0 = v0_;
+  v1 = v1_;
+  nv = nv_;
+  dt = dt_;
+  noneqTime = noneqTime_;
+  noneqCheckFeq = noneqCheckFeq_;
+  quenchTime = quenchTime_;
+  ppert = &pert;
+
+  numMode = ppert->numMode();  
+  numCheck = int((noneqTime + 0.5 * noneqCheckFeq) / noneqCheckFeq) + 1;
+  checkNumFeq = int((noneqCheckFeq + 0.5 * dt) / dt);
+  quenchNumStep = int((quenchTime + 0.5 * dt) / dt);
+  
+  order0.resize(numCheck);
+  order1.resize(numCheck);
+  order2.resize(numCheck);
+  
+  quench_order0.resize(numCheck);
+  quench_order1_term1.resize(numCheck);
+  quench_order1_term2.resize(numCheck);
+  quench_order2_term1.resize(numCheck);
+  quench_order2_term2.resize(numCheck);
+  quench_order2_term3.resize(numCheck);
+
+  for (int jj = 0; jj < numCheck; ++jj){
+    order1[jj].resize (numMode);
+    order2[jj].resize (numMode);
+    quench_order1_term2[jj].resize (numMode);
+    quench_order2_term2[jj].resize (numMode);
+    quench_order2_term3[jj].resize (numMode);
+    for (int kk = 0; kk < numMode; ++kk){
+      order2[jj][kk].resize (numMode);
+      quench_order2_term2[jj][kk].resize (numMode);
+    }
+  }
+
+  for (int tt = 0; tt < numCheck; ++tt){
+    order0[tt]				.reinit (x0, x1, nx, v0, v1, nv);
[…]

### 2012-07-27 one.test/template/tools/simulator/src/Perturbation.cpp

@@ -19,6 +19,27 @@ operator () (const Dofs & dof) const
   return tmp;
 }
 
+int Perturbation::
+numMode () const
+{
+  return 1;
+}
+
+void Perturbation::
+FeMode (const double & time,
+	vector<double > & modes) const
+{
+  modes.resize(1);
+  modes[0] = Fe(time);
+}
+
+void Perturbation::
+FeModePref (vector<double > & perf) const
+{
+  perf.resize(1);
+  perf[0] = 1;
+}
+
 
 PertConstTilt::
 PertConstTilt (const double & s,
@@ -53,6 +74,31 @@ Fe (const double & time) const
   return tmp;
 }
 
+int PertConstTilt::
+numMode () const
+{
+  return 1;
+}
+
+void PertConstTilt::
+FeMode (const double & time,
+	vector<double > & modes) const
+{
+  modes.resize(1);
+  modes[0] = 1.;
+  if (warmTime > 0.0 && time < warmTime){
+    modes[0] *= time / warmTime;
+  }
+}
+
+void PertConstTilt::
+FeModePref (vector<double > & pref ) const
+{
+  pref.resize(1);
+  pref[0] = strength;
+}
+
+
 
 PertGaussian::
 PertGaussian (const double s,
@@ -74,6 +120,31 @@ Fe (const double & time) const
   }
   return tmp;
 }
+
+int PertGaussian::
+numMode () const
+{
+  return 1;
+}
+
+void PertGaussian::
+FeMode (const double & time,
+	vector<double > & modes) const
+{
+  modes.resize(1);
+  modes[0] = 1.;
+  if (warmTime > 0.0 && time < warmTime){
+    modes[0] *= time / warmTime;
+  }
+}
+
+void PertGaussian::
+FeModePref (vector<double > & pref ) const
+{
+  pref.resize(1);
+  pref[0] = strength;
+}
+
 
 void PertGaussian::
 operator () (const Dofs & dofs,

### 2012-07-27 one.test/template/tools/simulator/app/splitSingleWell.noneqResp.cpp
fig bugs

@@ -171,9 +171,6 @@ int main(int argc, char * argv[])
   
   int count = 0;
   int countBranch = 0;
-  double ttNoneq;
-  double ttQuench;
-  unsigned iiCheck = 0;
   double time = 0;
 
   NoneqResponseInfo resInfo;
@@ -194,33 +191,43 @@ int main(int argc, char * argv[])
       countBranch = 0;
       Dofs branchXX (xx);
       Dofs branchXX_old (xx);
-      iiCheck = 0;
       resInfo.newTraj ();
-      // branching
-      int countNoneqCheck = noneqCheckNumFeq;
-      for (ttNoneq = 0.; ttNoneq < noneqTime-0.5*dt; ttNoneq += dt){
-	// printf ("%f %d %d\n", ttNoneq, countNoneqCheck, noneqCheckNumFeq);
- 	if (countNoneqCheck == int(noneqCheckNumFeq)){
-	  // printf ("check non eq at time %f \n", ttNoneq);
-	  countNoneqCheck = 0;
-	  // dists[iiCheck].deposite (branchXX);
-	  // quenching
-	  Dofs quenchXX (branchXX);
-	  Dofs quenchXX_old (branchXX);
-	  for (ttQuench = 0.; ttQuench < quenchTime-0.5*dt; ttQuench += dt){
-	    quenchXX_old = quenchXX;
-	    quenchInte.step(quenchXX, ttNoneq);
-	    Dofs quenchDw = quenchInte.getDw();
-	    resInfo.depositQuenchTraj (quenchXX_old, quenchXX, quenchSigma, quenchDw);
-	  }
-	  // distsQuench[iiCheck].deposite (quenchXX);
-	  iiCheck ++;
+      // at time 0, quench any way
+      {
+	// quenching
+	resInfo.newQuenchTraj ();
+	Dofs quenchXX (branchXX);
+	Dofs quenchXX_old (branchXX);
+	for (double ttQuench = 0.; ttQuench < quenchTime-0.5*dt; ttQuench += dt){
+	  quenchXX_old = quenchXX;
+	  quenchInte.step(quenchXX, 0.);
+	  Dofs quenchDw = quenchInte.getDw();
+	  resInfo.depositQuenchTraj (quenchXX_old, quenchXX, quenchSigma, quenchDw);
 	}
+      }      
+      int countNoneqCheck = 0;
+      // branching
[…]

### 2012-07-27 one.test/template/tools/simulator/include/NoneqResponseInfo.h

@@ -44,7 +44,6 @@ class NoneqResponseInfo
   double					G0;		// 1
   double					H00;		// 1
 
-  void newQuenchTraj ();
 public:
   NoneqResponseInfo ();
   // ~NoneqResponseInfo ();
@@ -61,6 +60,7 @@ public:
 	       const double & quenchTime,
 	       const Perturbation & pert);
   void newTraj ();
+  void newQuenchTraj ();
   void depositMainTraj (const Dofs & oldx,
 			const Dofs & newx,
 			const double & sigma,

### 2012-07-27 one.test/template/tools/simulator/src/NoneqResponseInfo.cpp

@@ -96,8 +96,6 @@ newTraj ()
 
   countNoneq = 0;
   countNoneqSeg = 0;
-
-  newQuenchTraj ();
 }
 
 void NoneqResponseInfo::
@@ -118,6 +116,8 @@ depositMainTraj (const Dofs & oldx,
   // if (countNoneq - countNoneqSeg * checkNumFeq == 0){
   if (countNoneq == 0){
     order0[0].deposite (oldx);
+    countNoneq ++;
+    return;
   }
     
   double nowTime = countNoneq * dt;
@@ -140,7 +140,10 @@ depositMainTraj (const Dofs & oldx,
   }
 
   countNoneq ++;
-  if (countNoneq - countNoneqSeg * checkNumFeq == 0){
+  // std::cout << "noneq count is " << countNoneq
+  // 	    << "  countNoneqSeg is " << countNoneqSeg 
+  // 	    << std::endl;
+  if (countNoneq == (countNoneqSeg + 1) * checkNumFeq){
     countNoneqSeg ++;
     if (countNoneqSeg >= numCheck){
       std::cerr << "# the non eq traj is too long, problematic!" << std::endl;
@@ -153,7 +156,6 @@ depositMainTraj (const Dofs & oldx,
 	order2[countNoneqSeg][jj][kk].deposite (newx, Hjk[jj][kk]);
       }
     }
-    newQuenchTraj ();
   }
 }
 
@@ -178,6 +180,7 @@ depositQuenchTraj (const Dofs & oldx,
   H00 += tmp1 * dt / sigma;
 
   countQuench ++;
+  // std::cout << "quench count is " << countQuench << std::endl;
   if (countQuench == quenchNumStep){
     quench_order0[countNoneqSeg].deposite (newx);
     if (countNoneqSeg != 0){

### 2012-07-27 one.test/template/parameters.sh
add sh scripts for response simulations

@@ -5,6 +5,7 @@ noneq_time=150.0
 corr_step=0.01
 corr_time=10.0
 warm_time=100.0
+refe_strength=0.0
 pert_strength=1.0
 project_name=splitSingleWell
 #project_name=tiltDoubleWell

### 2012-07-27 one.test/template/run.resp.sh

@@ -1,0 +1,24 @@
+#!/bin/bash
+
+source parameters.sh
+
+make -C tools/simulator makedir &> make.log
+make -C tools/simulator -j8 &>> make.log
+
+rm -f run.resp.log
+
+command="`pwd`/tools/simulator/$project_name.noneqResp \
+--dt $dt --nst $nst -p 100000 --gamma $gamma --temperature $temperature \
+$command_line_param_print \
+--branch-feq $branch_feq \
+--noneq-check-feq $noneq_check_feq --noneq-time $noneq_time \
+--quench-temperature $quench_temperature --quench-time $quench_time \
+--x-low $x0 --x-up $x1 --v-low $v0 --v-up $v1 --x-grid $nx --v-grid $nv \
+--warm-time $warm_time --seed $seed \
+--pert-strength0 $refe_strength --pert-strength1 $pert_strength"
+
+echo "# run on `hostname`" &>> run.resp.log
+echo "# command is $command" &>> run.resp.log
+$command > output.resp 2> error.resp &
+echo "# pid is $!" &>> run.resp.log
+

### 2012-07-27 one.test/template/tools/simulator/app/splitSingleWell.noneqResp.cpp

@@ -96,8 +96,8 @@ int main(int argc, char * argv[])
   std::cout << "# sigma: " << sigma << " [nm]" << std::endl;
   // std::cout << "# barrier: " << 0.5 * kk * aa*aa*aa*aa << " [kJ/mol]" << std::endl;
   std::cout << "# gamma: " << gamma << " [ps^-1]" << std::endl;
-  std::cout << "# pert st0: " << pertSt0 << " [kJ/(mol nm)]" << std::endl;
-  std::cout << "# pert st1: " << pertSt1 << " [kJ/(mol nm)]" << std::endl;
+  std::cout << "# pert st0 (reference)   : " << pertSt0 << " [kJ/(mol nm)]" << std::endl;
+  std::cout << "# pert st1 (perturbation): " << pertSt1 << " [kJ/(mol nm)]" << std::endl;
   std::cout << "# warm time: " << warmTime << " [ps]" << std::endl;
   std::cout << "# quench T: " << quenchTemperture << " [K]" << std::endl;
   std::cout << "# quench kT: " << quenchkT << " [kJ/mol]" << std::endl;

### 2012-07-28 one.test/template/tools/simulator/src/NoneqResponseInfo.cpp
fix bug

@@ -153,7 +153,7 @@ depositMainTraj (const Dofs & oldx,
     for (int jj = 0; jj < numMode; ++jj){
       order1[countNoneqSeg][jj].deposite (newx, Gj[jj]);
       for (int kk = 0; kk < numMode; ++kk){
-	order2[countNoneqSeg][jj][kk].deposite (newx, Hjk[jj][kk]);
+	order2[countNoneqSeg][jj][kk].deposite (newx, Gj[jj] * Gj[kk] - Hjk[jj][kk]);
       }
     }
   }
@@ -185,12 +185,12 @@ depositQuenchTraj (const Dofs & oldx,
     quench_order0[countNoneqSeg].deposite (newx);
     if (countNoneqSeg != 0){
       quench_order1_term1[countNoneqSeg].deposite (newx, G0);
-      quench_order2_term1[countNoneqSeg].deposite (newx, H00);
+      quench_order2_term1[countNoneqSeg].deposite (newx, G0*G0 - H00);
       for (int jj = 0; jj < numMode; ++jj){
 	quench_order1_term2[countNoneqSeg][jj].deposite (newx, Gj[jj]);
 	quench_order2_term3[countNoneqSeg][jj].deposite (newx, G0 * Gj[jj]);
 	for (int kk = 0; kk < numMode; ++kk){
-	quench_order2_term2[countNoneqSeg][jj][kk].deposite (newx, Gj[jj] * Gj[kk]);
+	quench_order2_term2[countNoneqSeg][jj][kk].deposite (newx, Gj[jj] * Gj[kk] - Hjk[jj][kk]);
 	}
       }
     }
@@ -260,7 +260,7 @@ calculate (const double & time,
   // order 2
   for (int jj = 0; jj < numMode; ++jj){
     for (int kk = 0; kk < numMode; ++kk){
-      dist.add (-0.5 * pref1[jj] * pref1[kk], order2[index][jj][kk]);
+      dist.add (0.5 * pref1[jj] * pref1[kk], order2[index][jj][kk]);
     }
   }
 
@@ -273,11 +273,11 @@ calculate (const double & time,
     quench_dist.add (pref1[jj], quench_order1_term2[index][jj]);
   }
   // order 2
-  quench_dist.add (-0.5 * Fe1 * Fe1, quench_order2_term1[index]);
+  quench_dist.add (0.5 * Fe1 * Fe1, quench_order2_term1[index]);
[…]

### 2012-07-30 one.test/template/tools/simulator/app/splitSingleWell.noneqResp.cpp
save and load response from file

@@ -46,7 +46,9 @@ int main(int argc, char * argv[])
   unsigned branchNumFeq;
   double x0, x1, v0, v1;
   unsigned nx, nv;
-      
+
+  string sfile, lfile;
+  
   unsigned long seed;
   
   po::options_description desc ("Allow options");
@@ -67,6 +69,8 @@ int main(int argc, char * argv[])
       ("pert-strength0",po::value<double > (&pertSt0)->default_value(0.0), "perturbation strength 0 [kJ/(mol nm)]")
       ("pert-strength1",po::value<double > (&pertSt1)->default_value(1.0), "perturbation strength 1 [kJ/(mol nm)]")
       ("warm-time", po::value<double > (&warmTime)->default_value(100.), "warm up time of perturbation [ps]")
+      ("save-corr", po::value<string > (&sfile), "save correlation")
+      ("load-corr", po::value<string > (&lfile), "load saved correlation")
       ("x-low", po::value<double > (&x0)->default_value (-2.0), "the lower bound of x range considered")
       ("x-up",  po::value<double > (&x1)->default_value ( 2.0), "the upper bound of x range considered")
       ("v-low", po::value<double > (&v0)->default_value (-8.0), "the lower bound of v range considered")
@@ -137,8 +141,7 @@ int main(int argc, char * argv[])
   xx.xx[0] = 0.;
   xx.vv[0] = 0.;
 
-  unsigned numNoneqCheck = noneqTime / noneqCheckFeq;
-  numNoneqCheck ++;
+  unsigned numNoneqCheck = int((noneqTime + 0.5 * noneqCheckFeq) / noneqCheckFeq) + 1;
   Distribution_1d dist;
   Distribution_1d distQuench;
   vector<double > checkTimes(numNoneqCheck);
@@ -175,87 +178,72 @@ int main(int argc, char * argv[])
 
   NoneqResponseInfo resInfo;
   resInfo.reinit (x0, x1, nx, v0, v1, nv, dt, noneqTime, noneqCheckFeq, quenchTime, pert);
-  // resInfo.newTraj ();
 
-  for (double ii = 0.; ii < nst+0.1; ii += 1.){
-    inte.step (xx, 0.);
-    count ++;
-    countBranch ++;
-    time += dt;
[…]

### 2012-07-30 one.test/template/tools/simulator/include/Distribution.h

@@ -46,6 +46,9 @@ public:
   void add (const double & scalor,
 	    const Distribution_1d & d);
   double getNframe () const {return nframe;}
+public:
+  void save (FILE * fp) const;
+  void load (FILE * fp);
 }
     ;
 

### 2012-07-30 one.test/template/tools/simulator/include/NoneqResponseInfo.h

@@ -75,6 +75,9 @@ public:
 		  const Perturbation & pert1,
 		  Distribution_1d & dist,
 		  Distribution_1d & quench_dist);
+public:
+  void save (const string & filename) const;
+  void load (const string & filename);
 }
     ;
 

### 2012-07-30 one.test/template/tools/simulator/src/Distribution.cpp

@@ -200,5 +200,107 @@ add (const double & scalor,
     }
   }
 }
-  
+
+
+void Distribution_1d::
+save (FILE * fp) const
+{
+  size_t rv;
+  rv = fwrite (&x0, sizeof(double), 1, fp);
+  if (rv != 1){
+    cerr << "error writing corr file " << endl;
+    exit(1);
+  }
+  rv = fwrite (&x1, sizeof(double), 1, fp);
+  if (rv != 1){
+    cerr << "error writing corr file " << endl;
+    exit(1);
+  }
+  rv = fwrite (&v0, sizeof(double), 1, fp);
+  if (rv != 1){
+    cerr << "error writing corr file " << endl;
+    exit(1);
+  }
+  rv = fwrite (&v1, sizeof(double), 1, fp);
+  if (rv != 1){
+    cerr << "error writing corr file " << endl;
+    exit(1);
+  }
+  rv = fwrite (&nx, sizeof(unsigned), 1, fp);
+  if (rv != 1){
+    cerr << "error writing corr file " << endl;
+    exit(1);
+  }
+  rv = fwrite (&nv, sizeof(unsigned), 1, fp);
+  if (rv != 1){
+    cerr << "error writing corr file " << endl;
+    exit(1);
+  }
+
+  double * buff = (double *) malloc (sizeof(double) * nx * nv);
+  for (unsigned ii = 0; ii < nx; ++ii) {
+    for (unsigned jj = 0; jj < nv; ++jj) {
+      buff[ii*nv +jj] = values[ii][jj];
+    }
+  }
+  rv = fwrite (buff, sizeof(double), nx*nv, fp);
+  if (rv != nx*nv){
+    cerr << "error writing corr file " << endl;
+    exit(1);
+  }
+
+  free (buff);
+}
+
+
+void Distribution_1d::
+load (FILE * fp)
+{
+  size_t rv;
+  rv = fread (&x0, sizeof(double), 1, fp);
+  if (rv != 1){
+    cerr << "error reading corr file " << endl;
+    exit(1);
+  }
+  rv = fread (&x1, sizeof(double), 1, fp);
+  if (rv != 1){
+    cerr << "error reading corr file " << endl;
+    exit(1);
+  }
+  rv = fread (&v0, sizeof(double), 1, fp);
+  if (rv != 1){
+    cerr << "error reading corr file " << endl;
+    exit(1);
+  }
[…]

### 2012-07-30 one.test/template/tools/simulator/src/NoneqResponseInfo.cpp

@@ -82,6 +82,200 @@ reinit (const double & x0_,
 
 
 void NoneqResponseInfo::
+save (const string & filename) const
+{
+  FILE * fp = fopen (filename.c_str(), "w");
+
+  size_t rv;
+  rv = fwrite (&x0, sizeof(double), 1, fp);
+  if (rv != 1){
+    cerr << "error writing corr file " << endl;
+    exit(1);
+  }
+  rv = fwrite (&x1, sizeof(double), 1, fp);
+  if (rv != 1){
+    cerr << "error writing corr file " << endl;
+    exit(1);
+  }
+  rv = fwrite (&v0, sizeof(double), 1, fp);
+  if (rv != 1){
+    cerr << "error writing corr file " << endl;
+    exit(1);
+  }
+  rv = fwrite (&v1, sizeof(double), 1, fp);
+  if (rv != 1){
+    cerr << "error writing corr file " << endl;
+    exit(1);
+  }
+  rv = fwrite (&nx, sizeof(unsigned), 1, fp);
+  if (rv != 1){
+    cerr << "error writing corr file " << endl;
+    exit(1);
+  }
+  rv = fwrite (&nv, sizeof(unsigned), 1, fp);
+  if (rv != 1){
+    cerr << "error writing corr file " << endl;
+    exit(1);
+  }
+
+  rv = fwrite (&noneqTime, sizeof(double), 1, fp);
+  if (rv != 1){
+    cerr << "error writing corr file " << endl;
+    exit(1);
+  }
+  rv = fwrite (&noneqCheckFeq, sizeof(double), 1, fp);
+  if (rv != 1){
+    cerr << "error writing corr file " << endl;
+    exit(1);
+  }
+  rv = fwrite (&quenchTime, sizeof(double), 1, fp);
+  if (rv != 1){
+    cerr << "error writing corr file " << endl;
+    exit(1);
+  }
+
+  rv = fwrite (&numMode, sizeof(int), 1, fp);
+  if (rv != 1){
+    cerr << "error writing corr file " << endl;
+    exit(1);
+  }
+
+  
+  for (int tt = 0; tt < numCheck; ++tt){
+    order0[tt]				.save (fp);
+    quench_order0[tt]			.save (fp);
+    quench_order1_term1[tt]		.save (fp);
+    quench_order2_term1[tt]		.save (fp);
+    for (int jj = 0; jj < numMode; ++jj){
+      order1[tt][jj]			.save (fp);
[…]

### 2012-07-30 one.test/template/tools/simulator/app/splitSingleWell.noneqResp.cpp
fix bug

@@ -234,8 +234,8 @@ int main(int argc, char * argv[])
 	  }
 	}
       }
-      resInfo.average ();
     }
+    resInfo.average ();
     if (vm.count("save-corr")){
       resInfo.save (sfile);
     }

### 2012-08-05 writedown/linear.response/langevin.tex
fix bug
fix bug of the text
add tilt result of the gneral response theory

@@
   -
-  \frac 1{2\sigma}\int_0^t
+  \frac 1{2\sigma^2}\int_0^t
   [\Delta F_e^{i-1}(s)]^2\,\vect D^2(\vect q_s,\vect p_s)\,\dd s
@@
   -
-  \frac 1{2\sigma}\int_0^\tau
+  \frac 1{2\sigma^2}\int_0^\tau
   [\Delta F_e^{i-1}(t)]^2\,\vect D^2(\vect q_\eta,\vect p_\eta)\,\dd \eta
@@
   =&\,
-  \frac 1{\sigma}
+  \frac 1{\sigma^2}
   \int_0^t
@@
   =&\,
-  \frac 1{\sigma}
+  \frac 1{\sigma^2}
   \int_0^\tau
@@
   \Delta s_j^{i-1}G_j[\vect X_s]
-  -
-  \frac12
-  \sum_{j,k=1}^\infty
-  \Delta s_j^{i-1} \Delta s_k^{i-1}H_{jk}[\vect X_s]
+  % -
+  % \frac12
+  % \sum_{j,k=1}^\infty
+  % \Delta s_j^{i-1} \Delta s_k^{i-1}H_{jk}[\vect X_s]
   +
@@
   \sum_{j,k=1}^\infty
-  \Delta s_j^{i-1}\Delta s_k^{i-1}G_j[\vect X_s]G_k[\vect X_s]
+  \Delta s_j^{i-1}\Delta s_k^{i-1}
+  \big\{
+  G_j[\vect X_s]G_k[\vect X_s] - H_{jk}[\vect X_s]
+  \big\}
   +
@@
   G_0[\vect X_\eta]
-  -
-  \frac12[\Delta F_e^{i-1}(t)]^2
-  H_{00}[\vect X_\eta]
+  % -
+  % \frac12[\Delta F_e^{i-1}(t)]^2
+  % H_{00}[\vect X_\eta]
   +
@@
   \big\{
-  G_0[\vect X_\eta]
-  \big\}^2
+  G^2_0[\vect X_\eta] - H_{00}[\vect X_\eta]
+  \big\}
   +
@@
   &\, 
-  -
+  +
   \frac12[\Delta F_e^{i-1}(t)]^2
@@
   K_{\sigma,\vect y}(\vect X_\tau)\,
+  \big\{
+  G^2_0[\vect X_s] - 
   H_{00}[\vect X_\eta]
+  \big\}
   \big\rangle_{0,t,\tau}^{i-1}
-  -
+  % +
+  % \frac12[\Delta F_e^{i-1}(t)]^2
+  % \big\langle
+  % K_{\sigma,\vect y}(\vect X_\tau)\,
+  % \big\{G_{0}[\vect X_\eta]\big\}^2
+  % \big\rangle_{0,t,\tau}^{i-1}
+  \\\nonumber
+  &\,
+  +
   \frac 12
@@
   K_{\sigma,\vect y}(\vect X_\tau)\,
+  \big\{
+  G_j[\vect X_s] G_k[\vect X_s] - 
   H_{jk}[\vect X_s]
+  \big\}
   \big\rangle_{0,t,\tau}^{i-1}
+  % +
+  % \frac 12
+  % \sum_{j,k=1}^\infty
+  % \Delta s_j^{i-1}\Delta s_k^{i-1}
+  % \big\langle
+  % K_{\sigma,\vect y}(\vect X_\tau)\,
+  % G_{j}[\vect X_s] G_k[\vect X_s]
+  % \big\rangle_{0,t,\tau}^{i-1}
   \\\label{eqn:pert-approx-1}
@@
 is the zeroth order approximation,  the second line is the
-first order approximation, and the last two lines are the
+first order approximation, and the last three lines are the
 second order approximation.
@@
   &\,
-  -
+  +
   \frac 12
@@
   K_{\sigma,\vect y}(\vect X_t)\,
+  \big\{
+  G_j[\vect X_s] G_k[\vect X_s] -
   H_{jk}[\vect X_s]
+  \big\}
   \big\rangle_{0,t}^{i-1}  
@@
   \label{fig:tmp1}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.95\textwidth]
+  {figs/fig-tilt-str2-2d-more.eps}
+  \caption{The plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
+    $F_e^{\textrm{max}} = 2$.
+    The 1st row: the direct non-equilibrium
+    simulation. The 2nd row: the results of linear response
+    formula~\eqref{eqn:core-identify-approx-2}.
+    The 3rd row: the results of response
+    formula~\eqref{eqn:pert-approx-1} and \eqref{eqn:pert-approx-2}.
+    The order is 2, the reference state is $F_e^{\textrm{max}} = 0$, i.e.
+    the equilibrium state.
+  }
+  \label{fig:tmp2}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.95\textwidth]
+  {figs/fig-tilt-str4-2d-more.eps}
+  \caption{The plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
+    $F_e^{\textrm{max}} = 4$.  The 1st row: the direct non-equilibrium
+    simulation. The 2nd row: the results of linear response
+    formula~\eqref{eqn:core-identify-approx-2}.
+    The 3rd row: the results of response
+    formula~\eqref{eqn:pert-approx-1} and \eqref{eqn:pert-approx-2}.
+    The order is 2, the reference state is $F_e^{\textrm{max}} = 0$, i.e.
+    the equilibrium state.
+    The 4th row: the results of response
+    formula~\eqref{eqn:pert-approx-1} and \eqref{eqn:pert-approx-2}.
+    The order is 1, the reference state is $F_e^{\textrm{max}} = 2$.
+  }
+  \label{fig:tmp3}
 \end{figure}
@@
 See Fig. \ref{fig:tmp2} and \ref{fig:tmp3} for numerical result.  Both
-the direct non-equilibrium simulation and the linear response
-approximation are presented. Time slices $t = 0$, 10, 15 and 20
+the direct non-equilibrium simulation and the response
+approximations are presented. Time slices $t = 0$, 10, 15 and 20
 \textsf{ps} are shown.  When the perturbation is small,
 i.e. $F_e^{\textrm{max}} = 2$, the direct non-equilibrium simulation
-and the linear response results are accurately consistent. As the
+and the linear response results are accurately consistent.
+The 2nd order response formula~\eqref{eqn:pert-approx-1} and
+\eqref{eqn:pert-approx-2} are even better, but the numerical error
+is bigger.
+As the
 perturbation grows stronger, the left well becomes less stable, while
 the right well becomes more stable.  When the perturbation is big,
-i.e. $F_e^{\textrm{max}} = 4$, the linear response results are no
-longer precise. When $t\geq 10$ \textsf{ps}, the left well basically
+i.e. $F_e^{\textrm{max}} = 4$, the linear response result is no
+longer precise.
+When $t\geq 20$ \textsf{ps}, the left well basically
 disappears, however, the linear response calculation presents
 artifical occilation at the right well.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.95\textwidth]
-  {figs/tilt.warm020.str2.0.nst1e09.smallgrid/fig-2d.eps}
-  \caption{The plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
-    $F_e^{\textrm{max}} = 2$.
-    Upper row: the direct non-equilibrium simulation. Lower row: the
-    results of linear response formula~\eqref{eqn:core-identify-approx-2}.
-  }
-  \label{fig:tmp2}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.95\textwidth]
-  {figs/tilt.warm020.str4.0.nst1e09.smallgrid/fig-2d.eps}
-  \caption{The plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
-    $F_e^{\textrm{max}} = 4$.  Upper row: the direct non-equilibrium
-    simulation. Lower row: the results of linear response
-    formula~\eqref{eqn:core-identify-approx-2}.  }
-  \label{fig:tmp3}
-\end{figure}
+The 2nd order response also has artifical effect, and the numerical
+uncertainty is high.
+The 4th rwo of Fig.~\ref{fig:tmp3} presents the 1st order result of the
+response formula~\eqref{eqn:pert-approx-1} and
+\eqref{eqn:pert-approx-2}, with
+respect to the reference state $F_e^{\textrm{max}} = 2$. The result
+is impressively improved compared with the linear response with respect to the
+equilibrium state.
 
@@
 
-In this splitting single-well potential example, the unperturbed
-potential is simply the harmonic potential given by:
-\begin{align}
-  U(\vect q) = \frac12\,k\,\vect q^2 
-\end{align}
-Here $k = 8$~$\textsf{kJ} / (\textsf{mol nm}^2)$.
-Notice at room temperature $300\ \textsf{K}$, $k_BT = 2.48$~\textsf{kJ/mol}.
-See the red line in Fig.~\ref{fig:tmp4}.
-%$\textsf{kJ} / (\textsf{mol nm}^4)}$
-The perturbation is again given by
-\begin{align}
-  \vect C(\vect p, \vect q) = 0; \qquad
-  \vect D(\vect p, \vect q) = -\nabla_{\vect q} V(\vect q) 
-\end{align}
-where $V(\vect q)$ is a Gaussian type potential:
-\begin{align}
-  V(\vect q) = \frac{1}{\sqrt{2\pi \sigma^2}}
-  \exp\Big\{-\frac{\vect q^2}{2\sigma^2}\Big\}
-\end{align}
-we use $\sigma = 0.16$~\textsf{nm}. Please see Fig~\ref{fig:tmp4}.
-For the results, please see Fig.~\ref{fig:tmp5} -- Fig.~\ref{fig:tmp6}.
 \begin{figure}
@@
 \end{figure}
+
+In this splitting single-well potential example, the unperturbed
+potential is simply the harmonic potential given by:
+\begin{align}
+  U(\vect q) = \frac12\,k\,\vect q^2 
+\end{align}
+Here $k = 8$~$\textsf{kJ} / (\textsf{mol nm}^2)$.
+Notice at room temperature $300\ \textsf{K}$, $k_BT = 2.48$~\textsf{kJ/mol}.
+See the red line in Fig.~\ref{fig:tmp4}.
+%$\textsf{kJ} / (\textsf{mol nm}^4)}$
+The perturbation is again given by
+\begin{align}
+  \vect C(\vect p, \vect q) = 0; \qquad
+  \vect D(\vect p, \vect q) = -\nabla_{\vect q} V(\vect q) 
+\end{align}
+where $V(\vect q)$ is a Gaussian type potential:
+\begin{align}
+  V(\vect q) = \frac{1}{\sqrt{2\pi \sigma^2}}
+  \exp\Big\{-\frac{\vect q^2}{2\sigma^2}\Big\}
+\end{align}
+we use $\sigma = 0.16$~\textsf{nm}. Please see Fig~\ref{fig:tmp4}.
+For the results, please see Fig.~\ref{fig:tmp5} -- Fig.~\ref{fig:tmp6}.
 

### 2012-08-07 writedown/linear.response/langevin.tex
add new results for splitting single well potential

@@
   \caption{The plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
-    $F_e^{\textrm{max}} = 4$.  The 1st row: the direct non-equilibrium
+    $F_e^{\textrm{max}} = 4$.
+    The 1st row: the direct non-equilibrium
     simulation. The 2nd row: the results of linear response
@@
   \includegraphics[width=0.95\textwidth]
-  {figs/split.warm020.str1.0.nst1e09.smallgrid.more/fig-2d.eps}
+  {figs/fig-split-str1-2d-more.eps}
   \caption{The plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
     $F_e^{\textrm{max}} = 1.0$.
-    Upper row: the direct non-equilibrium simulation. Lower row: the
-    results of linear response formula~\eqref{eqn:core-identify-approx-2}.
+    The 1st row: the direct non-equilibrium
+    simulation. The 2nd row: the results of linear response
+    formula~\eqref{eqn:core-identify-approx-2}.
+    The 3rd row: the results of response
+    formula~\eqref{eqn:pert-approx-1} and \eqref{eqn:pert-approx-2}.
+    The order is 2, the reference state is $F_e^{\textrm{max}} = 0$, i.e.
+    the equilibrium state.
   }
@@
   \includegraphics[width=0.95\textwidth]
-  {figs/split.warm020.str2.0.nst1e09.smallgrid.more/fig-2d.eps}
+  {figs/fig-split-str2-2d-more.eps}
   \caption{The plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
-    $F_e^{\textrm{max}} = 2.0$.  Upper row: the direct non-equilibrium
-    simulation. Lower row: the results of linear response
-    formula~\eqref{eqn:core-identify-approx-2}.  }
+    $F_e^{\textrm{max}} = 2.0$.
+    The 1st row: the direct non-equilibrium
+    simulation. The 2nd row: the results of linear response
+    formula~\eqref{eqn:core-identify-approx-2}.
+    The 3rd row: the results of response
+    formula~\eqref{eqn:pert-approx-1} and \eqref{eqn:pert-approx-2}.
+    The order is 2, the reference state is $F_e^{\textrm{max}} = 0.0$, i.e.
+    the equilibrium state.
+    The 4th row: the results of response
+    formula~\eqref{eqn:pert-approx-1} and \eqref{eqn:pert-approx-2}.
+    The order is 1, the reference state is $F_e^{\textrm{max}} = 1.0$.
+  }
   \label{fig:tmp6}

### 2012-08-21 writedown/linear.response/langevin.tex
add accuracy consideration of the response theory

@@
 
+\section{On the accuracy of the response fomula}
+
+An important issue is how far can we trust the response theory.
+To investigate this issue, we should firstly look at the
+most important step: the trajectory measure reweighting,
+i.e. Eqn.~\eqref{eqn:exact-ratio-1} and \eqref{eqn:exact-ratio-2}.
+These formula are strict, however, the statistical error
+may be increased do to the measure reweighting, according to
+similar measure reweighting methods widely used in equilibrium sampling.
+We find it more convenient to start from the nonequilibrium steady
+state (NESS), then we extend it to more general, but short-time
+auto-correlated dynamics.
+
+\subsection{Starting from NESS sampling}
+
+For the NESS system, it is possible to define a time-independent
+measure of the configuration space, namely $p_0(\vect x)$, where
+the subscript ``0'' denotes the reference state. We want to measure
+the propties in the perturbed state, ``1'', where the NESS
+probability distribtion is denoted by $p_1(\vect x)$. Assume we have
+a sampling sequence $\vect x_1, \vect x_2, \cdots, \vect x_M$, which
+are independent, and are subject to the reference probability distribtion
+$p_0(\vect x)$. Any  observation $O$ of the perturbed system can be
+estimated by
+\begin{align}
+  \bar O_M = \frac 1M\sum_{i=0}^{M-1} O(\vect x_i) R(\vect x_i)
+\end{align}
+where $R(\vect x)$ is the ration between the two probability distributions:
+\begin{align}
+  R(\vect x) = \frac{p_1(\vect x)}{p_0(\vect x)}
+\end{align}
+$\bar O_M$ is an unbiased estimator for $\langle O\rangle_1$
+($\langle \cdot\rangle_1$ denotes the expectation value in the perturbed
+system), because
+\begin{align}
+  \langle \bar O_M\rangle_0
+  = \frac 1M\sum_{i=0}^{M-1}
+  \int\dd \vect x\, p_0(\vect x_i)\,O(\vect x_i) R(\vect x_i)
+  = \int \dd \vect x\,p_0(\vect x)\,O(\vect x)
+  \,\frac{p_1(\vect x)}{p_0(\vect x)}
+  = \langle O\rangle_1
+\end{align}
+The statistical uncertainty is defined by the variance:
+\begin{align}
+  \textrm{Var}_0(\bar O_M)
+  = \big\langle
+  [\bar O_M^2 - \langle \bar O_M\rangle_0]^2
+  \big\rangle_0
+  = \langle\bar O_M^2\rangle_0 -
+  \langle\bar O_M\rangle_0^2 
+\end{align}
+It can be calculated that (only some basic algibra are needed):
+\begin{align}
+  \textrm{Var}_0(\bar O_M)
+  =
+  \langle O^2\cdot R\rangle_1 - \langle O\rangle_1^2
+\end{align}
+Comparing this variance with that of a direct sampling in the perturbed system,
+the difference would be
+\begin{align}
+  \Delta\textrm{Var}_0(\bar O_M)
+  =
+  \langle O^2\cdot (R-1)\,\rangle_1
+\end{align}
+This means that the reweighted sampling can be more precise if the
+observable $O$ is more significant in the region where $p_0 > p_1$, or
+less precise if $O$ is more significant in the region where $p_0 <
+p_1$.  This is obvious, and be observed in the equilibrium sampling
+methods (for example, the umbrella sampling).  In our case, the
+observable is nothing but the mollifier $K_{\sigma, \vect y}(\vect
+x)$, which is the smoothified version of the delta function,
+so we can somehow think
+$\langle K_{\sigma, \vect y}(\vect x)\rangle_0
+\approx p_0(\vect y)$ and
+$\langle K_{\sigma, \vect y}(\vect x)\rangle_1
+\approx p_1(\vect y)$.  This
+means the average of the mollifier is always calculated with a worse
+precision when $\langle K_{\sigma, \vect y}(\vect x)\rangle_1 >
+\langle K_{\sigma, \vect y}(\vect x)\rangle_0$. Please note that we
+are always more interested in the region where $\langle K_{\sigma,
+  \vect y}(\vect x)\rangle_1 > \langle K_{\sigma, \vect y}(\vect
+x)\rangle_0$, so the sad news is that we will always end up with
+a worse preciseion compared with the direct sampling of perturbed
+system.
+To have a resonable approximation, it is required that
+the unperturbed distribution $p_0(\vect x)$ should always be
+overlapping with the perturbed version in the region where $p_1(\vect
+x)$, or $\langle K_{\sigma, \vect y}(\vect x)\rangle_1$ is relavant.
+
+\subsection{The accuracy of the reweighted trajectory sampling}
+
+We want to understand what is the precision of the
+reweighted trajectory sampling, namely:
+\begin{align}\label{eqn:traj-ob}
+  \Delta \mo (t) =
+  \int \dd \vect v\, f(\vect v, 0)
+  \int_{\mc\{\vect v, 0; t\}}
+  \dd \mathcal P_{i-1}^\beta[\vect X_s]\,
+  O(\vect X_t)
+  \Bigg[
+  \frac{\dd \mathcal P^\beta_i[\vect X_s]}
+  {\dd \mathcal P^\beta_{i-1}[\vect X_s]}
+  -1
+  \Bigg]
+\end{align}
+The leading order of the trajectory order reweighting is
+\begin{align}\label{eqn:leading-rew}
+  \frac{\dd \mathcal P^\beta_i[\vect X_s]}
+  {\dd \mathcal P^\beta_{i-1}[\vect X_s]}
+  -1
+  =
+  \frac1\sigma\int_0^t M(s)\,
+  \vect D(\vect X_s)\cdot \dd\vect W_t
+\end{align}
+Notice the observable $O$ only depends on the end point
+of the trajectory. Therefore, we can use this properties to derive the
+estimate of the path integral.
+Let a series of descrete point along the trajectory:
+$(\vect X_{t_1}, t_1), (\vect X_{t_2}, t_2), \cdots, (\vect X_{t_N}, t_N)$,
+with $t_0 = 0,\ t_N = t$. For convenience, we denote
+$\vect x_k = \vect X_{t_k}$. The descretization of the path integral
+\eqref{eqn:traj-ob} is 
+\begin{align}
+  \Delta\mo (t) =
+  \int\dd \vect x_0\, f(\vect x_0, 0)
+  \int \dd \vect x_N
+  \int
+  \prod_{k=0}^{N-1}
+  \dd \vect x_k\,
+  p(\vect x_{k+1},t_{k+1} | \vect x_{k},t_{k})\,
+  O(\vect x_N)\,
+  \frac1\sigma\,
+  \Big[
+  \sum_{m=0}^{N-1}
+  M(t_m) \vect D(\vect x_m) (\vect W_{m+1} - \vect W_{m})
+  \Big]
+\end{align}
+By using the Chapman-Kolmogorov property, we have
+\begin{align}\nonumber
+  \Delta\mo (t) = &\,
+  \frac1\sigma
+  \sum_{m=0}^{N-1}
+  \int \dd \vect x_0\,
+  \dd \vect x_m\,
+  \dd \vect x_{m+1}\,
+  \dd \vect x_N\,
+  p(\vect x_N,t_N | \vect x_{m+1}, t_{m+1})\,
+  p(\vect x_{m+1}, t_{m+1} | \vect x_{m}, t_{m})\,
+  p(\vect x_{m}, t_{m} | \vect x_{0}, t_{0}) \,
+  f(\vect x_0, t_0)\\
+  &\,
+  M(t_m) \vect D(\vect x_m) (\vect W_{m+1} - \vect W_{m})  
+  O(\vect x_N)
+\end{align}
+Here for simplicity, we assume the dynamics is over-damped (the equations
+are the same for $O$ as under-damped dynamics).
+\begin{align}
+  \dd \vect X_t =
+  \vect b(\vect X_t, t)\,\dd t +
+  \sigma\, \dd \vect W_t
+\end{align}
+the change of variables gives
+\begin{align}
+  \vect W_{m+1} - \vect W_{m}
+  =
+  \vect x_{m+1} - \vect x_{m}
+  -
+  \vect b(\vect x_m, t_m)\,\Delta t_m
+\end{align}
+so
+\begin{align}\nonumber
+  \Delta\mo (t) = &\,
+  \frac1\sigma
+  \sum_{m=0}^{N-1}
+  \int \dd \vect x_0\,
+  \dd \vect x_m\,
+  \dd \vect x_{m+1}\,
+  \dd \vect x_N\,
+  p(\vect x_N,t_N | \vect x_{m+1}, t_{m+1})\,
+  p(\vect x_{m+1}, t_{m+1} | \vect x_{m}, t_{m})\,
+  p(\vect x_{m}, t_{m} | \vect x_{0}, t_{0})\,
+  f(\vect x_0, t_0)
+  \\ \nonumber
+  &\,
+  M(t_m) \vect D(\vect x_m) (\vect x_{m+1} - \vect x_{m})  
+  O(\vect x_N)\\
+  -
+  &\,
+  \frac1\sigma
+  \sum_{m=0}^{N-1}
+  \int \dd \vect x_0\,
+  \dd \vect x_m\,
+  \dd \vect x_N\,
+  p(\vect x_N,t_N | \vect x_{m}, t_{m})\,
+  p(\vect x_{m}, t_{m} | \vect x_{0}, t_{0}) \,
+  f(\vect x_0, t_0)\\
+  &\,
+  M(t_m) \vect D(\vect x_m) \vect b(\vect x_m, t_m)\,\Delta t_m
+  O(\vect x_N)
+\end{align}
+% The last term equals to
+% \begin{align}
+%   \frac1\sigma
+%   \int \dd \tau
+%   M(\tau)
+%   \Big\langle
+%   O(\vect X_t) \vect b(\vect X_\tau,\tau) \vect D(\vect X_\tau)
+%   \Big\rangle
+% \end{align}
+By using the forward Kolmogorov equation, we have
+\begin{align}\nonumber
+  p(\vect x_{m+1},t_{m+1} | \vect x_{m},t_{m} )
+  = &\,
+  p(\vect x_{m+1},t_{m} | \vect x_{m},t_{m} )
+  -
+  \Delta t_m
+  \nabla_{\vect x_{m+1}}
+  \cdot
+  [ \vect b(\vect x_{m+1}, t_m)\,
+  p(\vect x_{m+1},t_{m} | \vect x_{m},t_{m} )\,]\\
+  &\,
+  +
+  \Delta t_m \frac{1} 2\,\sigma^2\,
+  \nabla_{\vect x_{m+1}}^2 p(\vect x_{m+1},t_{m} | \vect x_{m},t_{m} )
+\end{align}
+with initial condition
+\begin{align}
+  p(\vect x_{m+1},t_{m} | \vect x_{m},t_{m} ) =
+  \delta (\vect x_{m+1} - \vect x_{m})
+\end{align}
+so
+\begin{align}\nonumber
+  \int\dd \vect x_{m+1}\,
+  &
+  p(\vect x_N,t_N | \vect x_{m+1}, t_{m+1})\,
+  p(\vect x_{m+1}, t_{m+1} | \vect x_{m}, t_{m})\,
+  (\vect x_{m+1} -\vect x_{m})\\\nonumber
+  =&
+  -
+  \Delta t_m
+  \int \dd \vect x_{m+1}\,
+  p(\vect x_N,t_N | \vect x_{m+1}, t_{m+1})\,
+  (\vect x_{m+1} - \vect x_{m})
+  \nabla_{\vect x_{m+1}}
+  \cdot
+  [ \vect b(\vect x_{m+1}, t_m)\,
+  p(\vect x_{m+1},t_{m} | \vect x_{m},t_{m} )\,]\\\nonumber
+  &
+  +
+  \Delta t_m
+  \int \dd \vect x_{m+1}\,
+  p(\vect x_N,t_N | \vect x_{m+1}, t_{m+1})\,
+  (\vect x_{m+1} - \vect x_{m})
+  \frac{\sigma^2} 2\,
+  \nabla_{\vect x_{m+1}}^2 p(\vect x_{m+1},t_{m} | \vect x_{m},t_{m} ) \\\nonumber
+  =&
+  \Delta t_m
+  \int \dd \vect x_{m+1}\,
+  \vect b(\vect x_{m+1}, t_m)\,
+  \delta (\vect x_{m+1} - \vect x_{m})
+  p(\vect x_N,t_N | \vect x_{m+1}, t_{m+1}) \\\nonumber
+  &+
+  \Delta t_m
+  \int \dd \vect x_{m+1}\,
+  \delta (\vect x_{m+1} - \vect x_{m})
+  \nabla_{\vect x_{m+1}}p(\vect x_N,t_N | \vect x_{m+1}, t_{m+1})\times
+  \frac{\sigma^2} 2 \\
+  =&
+  \Delta t_m
+  \vect b(\vect x_{m}, t_m)\,
+  p(\vect x_N,t_N | \vect x_{m}, t_{m+1}) 
+  +
+  \,\frac{\sigma^2} 2\,\Delta t_m
+  \nabla_{\vect x_{m}}p(\vect x_N,t_N | \vect x_{m}, t_{m+1})
+\end{align}
+By using the backward Kolmogorov equation
+\begin{align}
+  p(\vect x_N,t_N | \vect x_{m}, t_{m+1})
+  = 
+  p(\vect x_N,t_N | \vect x_{m}, t_{m})
+  + \mathcal O(\Delta t_m)
+\end{align}
+So by keeping the leading order of $\Delta t_m$
+\begin{align}\nonumber
+  \int\dd \vect x_{m+1}\,
+  &
+  p(\vect x_N,t_N | \vect x_{m+1}, t_{m+1})\,
+  p(\vect x_{m+1}, t_{m+1} | \vect x_{m}, t_{m})\,
+  (\vect x_{m+1} -\vect x_{m})\\\nonumber
+  =&
+  \Delta t_m
+  \vect b(\vect x_{m}, t_m)\,
+  p(\vect x_N,t_N | \vect x_{m}, t_{m}) 
+  +
+  \,\frac{\sigma^2} 2\,\Delta t_m
+  \nabla_{\vect x_{m}}p(\vect x_N,t_N | \vect x_{m}, t_{m})
+\end{align}
+Therefore, we have
+\begin{align}\label{eqn:error-o-corr}
+  \Delta\mo (t) =
+  \frac\sigma 2
+  % \int \dd\vect x_0 f(\vect x_0,0)
+  \int_0^t \dd\tau M(\tau)
+  \Bigg[
+  \int \dd\vect x_t\, O(\vect x_t)
+  \int \dd\vect x_\tau\,
+  \nabla_{\vect x_{\tau}}p(\vect x_t,t | \vect x_{\tau}, \tau)
+  \cdot
+  \vect D(\vect x_\tau)\,
+  % p(\vect x_\tau,\tau | \vect x_0, 0)
+  f(\vect x_\tau,\tau)
+  \bigg]
+\end{align}
+\redc{
+The term in the bracket is the time correlation between
+the observable and the backward flux of the error force $\vect D$.
+}
+
+\vskip .5cm
+Discussions on Eqn.~\eqref{eqn:error-o-corr}:
+\begin{enumerate}
+\item This equation relate the error force with the error of observable,
+  which would be useful in the error studies.
+\item When the system does not have long-time memory:
+  \begin{align}
+    p(\vect x_t,t | \vect x_{\tau}, \tau) = p(\vect x_t,t ), \qquad
+    \textrm{for}\ t  - \tau > t_c
+  \end{align}
+  Here $t_c$ denotes the length of memory of the system. Then
+  Eqn.~\eqref{eqn:error-o-corr} becomes 
+  \begin{align}
+    \Delta\mo (t) =
+    \frac\sigma 2
+    % \int \dd\vect x_0 f(\vect x_0,0)
+    \int_{t_c}^t \dd\tau M(\tau)
+    \Bigg[
+    \int \dd\vect x_t\, O(\vect x_t)
+    \int \dd\vect x_\tau\,
+    \nabla_{\vect x_{\tau}}p(\vect x_t,t | \vect x_{\tau}, \tau)
+    \cdot
+    \vect D(\vect x_\tau)\,
+    % p(\vect x_\tau,\tau | \vect x_0, 0)
+    f(\vect x_\tau,\tau)
+    \bigg]
+  \end{align}
+  This means we can integrate \eqref{eqn:leading-rew} only from
+  $t_c$ to $t$.
+\item In our case of core set identification,
+  $O(\vect x) = K_{\sigma, \vect y}(\vect x)$, which is nearly a delta function.
+  A good precision of the response theory means 
+  that the trajectory generated by the reference system, should
+  properly sample the correlation between the
+  error force flux at time $\tau$, and the probability
+  density at time $t$.  In other words, we generate trajectories
+  samples the conditional probability $p(\vect x_t,t | \vect x_{\tau}, \tau)$,
+  and a good precision means
+  the backward flux $\nabla_{\vect x_{\tau}}p(\vect x_t,t | \vect x_{\tau}, \tau)
+    \cdot
+    \vect D(\vect x_\tau)$ should be significant on this sampling.
+\end{enumerate}
+
 \section{Numerical tests}

### 2013-02-20 alanine/template/build.jobs.sh
add template

@@ -1,0 +1,23 @@
+#!/bin/bash
+
+num_of_proc=5
+seed_dir=set
+
+stop_num=$(($num_of_proc-1))
+
+for i in `seq 0 $stop_num`
+do
+    printi=`printf %02d $i`
+    if test -d $seed_dir.$printi; then
+	echo "existing dir $seed_dir.$printi, exit"
+	exit
+    fi
+    echo "# prepare $seed_dir.$printi"
+    cp -a $seed_dir $seed_dir.$printi
+    cd $seed_dir.$printi
+    sed -e "s/pert_parallel_num_pro=.*/pert_parallel_num_pro=$num_of_proc/g" parameters.sh |
+    sed -e "s/pert_parallel_my_id=.*/pert_parallel_my_id=$i/g" > tmp.tmp.tmp
+    mv -f tmp.tmp.tmp parameters.sh
+    cd ..
+done
+    

### 2013-02-20 alanine/template/collect.jobs.sh

@@ -1,0 +1,31 @@
+#!/bin/bash
+
+num_of_proc=5
+seed_dir=set
+stop_num=$(($num_of_proc-1))
+
+cwd=`pwd`
+
+rm -f $seed_dir/angle.name
+rm -f $seed_dir/gxs.name
+rm -f $seed_dir/energy.name
+
+for i in `seq 0 $stop_num`
+do
+    printi=`printf %02d $i`
+
+    echo "# collect $seed_dir.$printi"
+    cd $seed_dir
+    test ! -d result.perts && mkdir result.perts
+    cd result.perts
+    if test -d ../../$seed_dir.$printi/result.perts; then
+	ln -sf ../../$seed_dir.$printi/result.perts/pert.*   .
+    fi
+    cd ..
+    test -f ../$seed_dir.$printi/angle.name  && cat ../$seed_dir.$printi/angle.name  >> ./angle.name
+    test -f ../$seed_dir.$printi/gxs.name  && cat ../$seed_dir.$printi/gxs.name  >> ./gxs.name
+    test -f ../$seed_dir.$printi/energy.name && cat ../$seed_dir.$printi/energy.name >> ./energy.name
+    
+    cd ..
+done
+    

### 2013-02-20 alanine/template/set/auto.adagio.pert.sh

@@ -1,0 +1,24 @@
+#!/bin/bash
+
+cwd=`pwd`
+jobname=${cwd##*/}
+
+rm -f adagio.out adagio.err
+
+echo "#!/bin/bash"				>  submit.sh
+echo "#PBS -N $jobname"				>> submit.sh
+echo "#PBS -d $cwd"				>> submit.sh
+echo "#PBS -o adagio.out"			>> submit.sh
+echo "#PBS -e adagio.err"			>> submit.sh
+echo "#PBS -M han.wang@fu-berlin.de"		>> submit.sh
+echo "#PBS -l walltime=100:00:00"		>> submit.sh
+echo "#PBS -l nodes=1:ppn=8"			>> submit.sh
+echo "#PBS -l mem=1000mb"			>> submit.sh
+
+echo "hostname"					>> submit.sh
+echo "date"					>> submit.sh
+echo "pwd"					>> submit.sh
+echo "./run.pert.sh"				>> submit.sh
+
+qsub submit.sh &> jobid
+

### 2013-02-20 alanine/template/set/functions.sh

@@ -1,0 +1,84 @@
+#!/bin/bash
+
+function set_parameters_equi () {
+    file=$1
+    equi_nstep_warm=`echo "$equi_warm_time/$equi_dt" | bc -l | cut -d '.' -f 1`
+    equi_nstep_count=`echo "$equi_frame_feq * $equi_num_frame / $equi_dt" | bc -l | cut -d '.' -f 1`
+    equi_nstep=`echo "$equi_nstep_warm + $equi_nstep_count" | bc`
+    equi_xvout_feq=`echo "$equi_frame_feq / $equi_dt" | bc -l | cut -d '.' -f 1`
+    sed -e "/^integrator/s/=.*/= sd/g" $file |\
+    sed -e "/^dt/s/=.*/= $equi_dt/g" |\
+    sed -e "/^Tcoupl/s/=.*/= no/g" |\
+    sed -e "/^nstep/s/=.*/= $equi_nstep/g" |\
+    sed -e "/^nstxout/s/=.*/= $equi_xvout_feq/g" |\
+    sed -e "/^nstvout/s/=.*/= $equi_xvout_feq/g" |\
+    sed -e "/^ld-seed/s/=.*/= $equi_seed/g" |\
+    sed -e "/^gen_vel/s/=.*/= no/g" |\
+    sed -e "/^tau_t/s/=.*/= $equi_taut/g" |\
+    sed -e "/^nstxtcout/s/=.*/= $equi_xvout_feq/g" > tmptmptmp.mdp
+    mv -f tmptmptmp.mdp $file
+}
+
+# function set_parameters_long_equi () {
+#     file=$1
+#     long_equi_nstep_warm=`echo "$long_equi_warm_time/$long_equi_dt" | bc -l | cut -d '.' -f 1`
+#     long_equi_nstep_count=`echo "$long_equi_frame_feq * $long_equi_num_frame / $long_equi_dt" | bc -l | cut -d '.' -f 1`
+#     long_equi_nstep=`echo "$long_equi_nstep_warm + $long_equi_nstep_count" | bc`
+#     long_equi_xvout_feq=`echo "$long_equi_frame_feq / $long_equi_dt" | bc -l | cut -d '.' -f 1`
+#     sed -e "/^integrator/s/=.*/= sd/g" $file |\
+#     sed -e "/^dt/s/=.*/= $long_equi_dt/g" |\
[…]

### 2013-02-20 alanine/template/set/parameters.sh

@@ -1,0 +1,68 @@
+# method of run:
+run_method=atom.langevin
+gro_dir=gromacs.conf.box.002.70
+
+# # productive equilibrium run settings
+# long_equi_warm_time=200		# ps
+# long_equi_frame_feq=5		# ps
+# long_equi_num_frame=100		# n
+# long_equi_dt=0.002		# ps
+# long_equi_taut=0.5		# ps
+# long_equi_seed=`date +%s`	# 
+
+# equilibrium run settings
+equi_warm_time=200		# ps
+equi_frame_feq=2		# ps
+equi_num_frame=10000		# n
+equi_dt=0.002			# ps
+equi_taut=0.5			# ps
+equi_seed=`date +%s`		# 
+
+# non-equilibrium settings
+# pert_conf_dir="result.equi/equiConfs/"
+pert_equi_result="result.equi"
+pert_num_conf_use=20000		# 
+pert_mode=1			# 1: relax, 2: peaks
+pert_strength=1.0		# nm/ps velocity
+pert_warm_time=10		# ps
+pert_time=20			# ps
+pert_frame_feq=0.1		# ps
+pert_dt=0.002			# ps
+pert_taut=1.0			# ps
+pert_noSdRange=1.0		# nm
+pert_barostat=no		# Parrinello-Rahman or no
+pert_taup=2.0			# ps
+pert_parallel_num_pro=5
+pert_parallel_my_id=0
+
+grompp_command="grompp -n index.ndx"
+mdrun_command="mdrun -v"
+
+if	echo $run_method | grep adress &> /dev/null; then
+    pert_integrator=sd
+    gro_dir=$gro_dir.adress
+    grompp_command="grompp -n "
+    gromacs_install_dir=~/study/thermo.convection/local.inhomo.thermostat.w.rdfCorr
+    source $gromacs_install_dir/bin/GMXRC.bash
+else if echo $run_method | grep "atom.inhomo.sd" | grep -v sd2 &> /dev/null; then
+    pert_integrator=sd1
+    gromacs_install_dir=~/study/adress.noneq/methane.solv/local.inhomo.sd
+    source $gromacs_install_dir/bin/GMXRC.bash
+else if echo $run_method | grep "atom.inhomo.sd2" &> /dev/null; then
+    pert_integrator=sd
+    gromacs_install_dir=~/study/adress.noneq/methane.solv/local.inhomo.sd2
+    source $gromacs_install_dir/bin/GMXRC.bash
[…]

### 2013-02-20 alanine/template/set/run.equi.sh

@@ -1,0 +1,26 @@
+#!/bin/bash
+
+source parameters.sh
+source functions.sh
+rm -f run.log
+
+if test -d result.equi; then
+    echo "existing dir result.equi, mv and backup"
+    mv result.equi result.equi.`date +%s`
+fi
+if test ! -d $gro_dir; then
+    echo "no dir $gro_dir, exit"
+    exit
+fi
+cp -a $gro_dir result.equi
+
+cd result.equi
+set_parameters_equi grompp.mdp
+
+echo "# run with command `which grompp`"
+$grompp_command &>> run.log
+echo "# run with command `which mdrun`"
+$mdrun_command &>> run.log
+
+split_trr
+cd ..

### 2013-02-20 alanine/template/set/run.genbox.sh

@@ -1,0 +1,81 @@
+#!/bin/bash
+
+if test $# -lt 1; then
+    echo "tell me the box size by using:"
+    echo "./run.genbox.sh boxsize"
+    exit
+fi
+if test ! -d gromacs.conf; then
+    echo "no dir gromacs.conf, exit"
+    exit
+fi
+if test ! -d gromacs.top; then
+    echo "no dir gromacs.top, exit"
+    exit
+fi
+
+box=$1
+p_box=`printf %.2f $box`
+i_box=`echo $p_box | cut -d '.' -f 1`
+d_box=`echo $p_box | cut -d '.' -f 2`
+p_box=`printf %03d.%02d $i_box $d_box`
+box=`printf %d.%02d $i_box $d_box`
+
+target_dir=gromacs.conf.box.$p_box
+
+if test -d $target_dir; then
+    echo "find dir $target_dir, do nothing"
+    exit
+fi
+
+mkdir -p $target_dir
+
+cp gromacs.conf/protein.gro $target_dir
+cp gromacs.conf/spc216.gro $target_dir
+cp gromacs.top/topol.top $target_dir
+cp gromacs.top/grompp.mdp $target_dir
+
+echo "# generate box"
+cd $target_dir
+sed -e "s/2\.5 2\.5 2\.5/$box $box $box/g" protein.gro > tmp.tmp
+mv -f tmp.tmp protein.gro
+grompp -c protein.gro &> genbox.log
+echo 3 0 | trjconv -f protein.gro -o out.gro -center -pbc whole &> genbox.log
+genbox -cp out.gro -cs spc216.gro -o conf.gro &> genbox.log
+nsol=`grep "Number of SOL molecules:" genbox.log | tail -n 1 | cut -d ":" -f 2`
+sed -e "s/SOL.*/SOL $nsol/g" topol.top > tmp.tmp
+mv -f tmp.tmp topol.top
+
+echo "# do NPT run"
+sed -e "/^Pcoupl/s/= no/= Parrinello-Rahman/g" grompp.mdp |
+sed -e "/^Tcoupl/s/= no/= v-rescale/g" |
+sed -e "/^freezegrps/s/= .*/= FIX/g" |
+sed -e "/^freezedim/s/= .*/= Y Y Y/g" |
+sed -e "/^nsteps/s/= .*/= 100000/g" > tmp.tmp
+mv -f tmp.tmp grompp.mdp
+echo "a 9" > command.index
[…]

### 2013-02-20 alanine/template/set/run.pert.sh

@@ -1,0 +1,61 @@
+#!/bin/bash
+
+source parameters.sh
+source functions.sh
+
+make -C tools/angles/ makedir
+make -C tools/angles/ -j4
+
+cwd=`pwd`
+# prepare files
+if test ! -d $pert_equi_result ; then
+    echo "no dir $pert_equi_result, exit"
+    exit
+fi
+if test ! -f $pert_equi_result/equi.frame ; then
+    echo "no file $pert_equi_result/equi.frame, exit"
+    exit
+fi
+if test ! -d $gro_dir; then
+    echo "no dir $gro_dir, exit"
+    exit
+fi
+#rm -f angle.name
+#rm -f gxs.name
+
+targets=`awk '{print $1}' $pert_equi_result/equi.frame | head -n $pert_num_conf_use`
+
+pert_main_dir=result.perts
+for i in $targets;
+do
+    count=$i
+    runid=`echo "$count % $pert_parallel_num_pro" | bc`
+    test $runid -ne $pert_parallel_my_id && continue
+    test ! -d $pert_main_dir && mkdir -p $pert_main_dir
+    my_dir=$pert_main_dir/pert.$count
+    if test -d $my_dir; then
+	echo "existing dir $my_dir, continue"
+	continue
+	rm -fr $my_dir
+    fi
+    cp -a $gro_dir $my_dir
+    cd $my_dir
+    rm -f run.log
+    set_parameters_pert grompp.mdp
+    start_time=`grep $count $cwd/$pert_equi_result/equi.frame | awk '{print $2}'`
+    echo "# run with command `which grompp`" &>> run.log
+    $grompp_command -t $cwd/$pert_equi_result/traj.trr -time $start_time &>> run.log
+    echo "# run with command `which mdrun`" &>> run.log
+    $mdrun_command &>> run.log
+
+    echo 3 14 | trjconv -center -pbc whole
+    mv -f trajout.xtc alanine.xtc
+    $cwd/tools/angles/evolve -f alanine.xtc -s angle.dat &> angle.log
+    rm -f traj.xtc traj.trr state*.cpt topol.tpr conf.gro index.ndx 
+    
+    cd $cwd
+    
+    echo "$my_dir/angle.dat" >> angle.name    
+    echo "$my_dir/gxs.out" >> gxs.name    
+done
+

### 2013-02-20 alanine/template/set/tools/angles/app/average.traj.metastable.resp.cpp

@@ -1,0 +1,232 @@
+#include <algorithm>
+#include <fstream>
+#include <iomanip>
+#include <iostream>
+#include <iterator>
+#include <list>
+#include <map>
+#include <set>
+#include <sstream>
+#include <string>
+#include <queue>
+#include <vector>
+#include <cmath>
+
+#include "Defines.h"
+#include "Distribution.h"
+#include <boost/program_options.hpp>
+namespace po = boost::program_options;
+using namespace std;
+#define MaxLineLength 2048
+
+bool myread (FILE * fp,
+	     float & time,
+	     ValueType & phi,
+	     ValueType & psi)
+{
+  size_t rv;
+  rv = fread (&time, sizeof(float), 1, fp);
+  if (rv != 1){
+    // cerr << "error read time or reach EOF" << endl;
+    return false;
+  }
+  rv = fread (&phi, sizeof(double), 1, fp);
+  if (rv != 1){
+    cerr << "error read phi " << endl;
+    exit(1);
+  }
+  rv = fread (&psi, sizeof(double), 1, fp);
+  if (rv != 1){
+    cerr << "error read psi " << endl;
+    exit(1);
+  }
+  return true;
+}
+
+bool myreadGxs (FILE * fp,
+		float & time,
+		ValueType & gxs)
+{
+  size_t rv;
+  rv = fscanf (fp, "%f", &time);
+  if (rv != 1){
+    // cerr << "error read time or reach EOF" << endl;
+    return false;
+  }
+  rv = fscanf (fp, "%lf", &gxs);  
+  if (rv != 1){
+    cerr << "error read gxs " << endl;
+    exit(1);
+  }
+  return true;
+}
+
+void depositMetastable (const double & phi,
+			const double & psi,
+			const vector<MetastableSet> & sets,
+			vector<double > & counts)
+{
+  int cc = 0;
+  counts.resize (sets.size());
+  for (unsigned ii = 0; ii < sets.size(); ++ii){
+    counts[ii] = 0;
+    if (sets[ii].inSet(phi, psi)){
+      counts[ii] = 1;
+      cc ++;
+    }
+  }
+  
+  if (cc != 1) {
+    if (cc == 0){
+      cerr << "not in any set!!!" << endl;
+    }
+    else {
+      cerr << "in multi sets" << endl;
+    }
+  }
+}
+
+
+int main(int argc, char * argv[])
[…]

### 2013-02-20 alanine/template/set/tools/angles/app/main.cpp

@@ -1,0 +1,139 @@
+#include <algorithm>
+#include <fstream>
+#include <iomanip>
+#include <iostream>
+#include <iterator>
+#include <list>
+#include <map>
+#include <set>
+#include <sstream>
+#include <string>
+#include <queue>
+#include <vector>
+#include <cmath>
+#include <boost/program_options.hpp>
+namespace po = boost::program_options;
+
+#include "Analyzer.h"
+#include "Angle.h"
+#include "Distribution.h"
+
+void readTop (const std::string & file,
+              TopInfo & info)
+{
+  FILE * fp = fopen (file.c_str(), "r");
+  if (fp == NULL){
+    std::cout << "cannot open file " << file << std::endl;
+    std::cout << "do not log top" << std::endl;
+    return;
+  }
+  fscanf (fp, "%d %d", &(info.numAtomOnALA), &(info.comIndexALA));
+  fscanf (fp, "%d", &(info.numAtomOnH2o));
+  fscanf (fp, "%d %d %d", &(info.OIndexH2o), &(info.H1IndexH2o), &(info.H2IndexH2o));
+  fclose (fp);
+}
+
+
+int main(int argc, char * argv[])
+{
+  float begin, end;
+  std::string ifile, ofile, tfile;
+  float time_prec = .01;
+  TopInfo info;
+  double refh;
+  
+  po::options_description desc ("Allow options");
+  desc.add_options()
+      ("help,h", "print this message")
+      ("begin,b", po::value<float > (&begin)->default_value(0.f), "start time")
+      ("end,e",   po::value<float > (&end  )->default_value(0.f), "end   time")
+      ("refh,r",  po::value<double > (&refh)->default_value(10), "size of bin (deg.)")
+      ("top-file,t",po::value<std::string > (&tfile)->default_value ("mytop"), "topolgy of the system")
+      ("output,o",  po::value<std::string > (&ofile)->default_value ("angle.out"), "the output of angle distribution")
+      ("input,f",   po::value<std::string > (&ifile)->default_value ("traj.xtc"), "the input .xtc file");
+      
+  po::variables_map vm;
+  po::store(po::parse_command_line(argc, argv, desc), vm);
+  po::notify (vm);
+  if (vm.count("help")){
+    std::cout << desc<< "\n";
+    return 0;
+  }
+  if (vm.count("top-file")){
+    readTop (tfile, info);
+  }
+  
[…]

### 2013-02-20 alanine/template/set/tools/scripts/energy.avg.sh

@@ -1,0 +1,33 @@
+#!/bin/bash
+
+make -C tools/average -j8
+
+pert_main_dir=result.perts
+
+targets=`ls $pert_main_dir | grep ^pert`
+
+count=0
+for i in $targets;
+do
+    count=$(($count+1))
+done
+count=$(($count-1))
+
+rm -f energy.name
+count1=0
+for i in $targets;
+do
+    echo doing $pert_main_dir/$i
+    cd $pert_main_dir/$i
+    echo 14 15 17 | g_energy -xvg none &> /dev/null
+    cd - &>/dev/null
+    echo $pert_main_dir/$i/energy.xvg >> energy.name
+    
+    count1=$(($count1+1))
+    if test $count1 -eq $count; then
+	break;
+    fi
+done
+
+echo doing average
+tools/average/average.file -f energy.name -o avg.energy.out

### 2013-02-20 writedown/linear.response/main.tex

@@
 \newcommand{\mt}{\mathcal T}
-\newcommand{\EX}{{\textrm{EX}}}
-\newcommand{\CG}{{\textrm{CG}}}
-\newcommand{\HY}{{\Delta}}
+\newcommand{\mo}{\mathcal O}
+\newcommand{\mi}{\mathcal I}
+\newcommand{\mc}{\mathcal C}
+\newcommand{\proj}{\mathit\Pi}
+\newcommand{\fwg}{{\mathcal A}}
+\newcommand{\bwg}{{\mathcal B}}
+\newcommand{\bsigma}{\boldsymbol\sigma}
 
@@
 theory, for details see Ref.~\cite{tuckeman2010statistical} for
-example. Here we assume the system is governed by equation:
-\begin{align}\label{eqn:gov-1}
-  \dot{\vect q}_i & = \frac{\partial\mh}{\partial \vect p_i}
-  + \vect C_i(\vect q,\vect p) F_e(t)\\\label{eqn:gov-2}
-  \dot{\vect p}_i & =- \frac{\partial\mh}{\partial \vect q_i}
-  + \vect D_i(\vect q,\vect p) F_e(t)
+example. Here we assume the system is governed by the
+perturbed Langevin equation:
+\begin{align}\label{eqn:langevin-1}
+  \dot{\vect q} & = \nabla_{\vect p}\mh(\vect q,\vect p)
+  + \vect C(\vect q,\vect p) F_e(t)\\\label{eqn:langevin-2}
+  \dot{\vect p} & =- \nabla_{\vect q}\mh(\vect q,\vect p)
+  + \vect D(\vect q,\vect p) F_e(t)
+  - \gamma\vect p
+  + \sigma\dot{\vect W}
 \end{align}
 Where $\vect C$ and $\vect D$ are perturbations to the system, and
-$F_e(t)$ is the strength of the perturbation as a function of time.
-The phase space imcompressibility condition is assumed:
+$F_e(t)$ is the strength of the perturbation as a function of time,
+and is of order $\mo (\varepsilon)$, where $\varepsilon$ is assumed to be small.
+The phase space incompressibility condition is assumed:
 \begin{align}
-  \sum_{i=1}^{3N}
-  \bigg[
-  \frac{\partial \vect C_i}{\partial \vect q_i} +
-  \frac{\partial \vect D_i}{\partial \vect p_i}
-  \bigg] = 0
+  \nabla_{\vect q}\cdot\vect C + \nabla_{\vect p}\cdot\vect D = 0
 \end{align}
 Then the phase space distribution $f(\vect x, t)$ ($\vect x = \{\vect
-q_i, \vect p_i\}$) is subject to the Liouville equation:
+q, \vect p\}$) is subject to the Kolmogorov forward equation:
+\begin{align}\label{eqn:orig-forward}
+  \frac{\partial}{\partial t} f(\vect x, t) - \fwg^\beta(t) f(\vect x, t) = 0
+\end{align}
+where $\fwg^\beta(t)$ is the forward infiniesimal generator given by
 \begin{align}
-  \frac{\partial}{\partial t} f(\vect x, t) + i\ml f(\vect x, t) = 0
+  \fwg^\beta(t) =
+  \frac{\sigma^2}2\Delta_{\vect p}
+  - \Big(
+  \nabla_{\vect p}\mh + \vect C F_e(t)
+  \Big)\cdot\nabla_{\vect q}
+  - \Big(
+  -\nabla_{\vect q}\mh +
+  \vect D F_e(t) - \gamma\vect p
+  \Big)\cdot\nabla_{\vect p}
+  + 3N\gamma,
 \end{align}
-where $\ml$ is the Liouville operator given by $i\ml = \dot{\vect
-  x}\cdot\nabla_{\vect x}$.  We assume that the perturbation to the
+and $\beta$ is the inverse temperature.
+The fluctuation-dissipation relation is always true:
+$\beta = 2\gamma / \sigma^2$.
+The equilibrium and unperturbed system
+is governed by
+the standard Langevin equation, i.e. with
+vanishing input perturbation $F_e(t) = 0$ in
+Eqn.~\eqref{eqn:langevin-1} and \eqref{eqn:langevin-2}.
+We assume that the perturbation to the
 system is so small that both the phase space distribution and the
-Liouville operator can be viewed as perturbation from the equilibrium
-distribution (invariant measure) and the unperturbed operator,
-respectively. Namely
+infiniesimal generator can be viewed as perturbation from the initial
+distribution and the unperturbed operator
+($\vect C = \vect D = 0$),
+respectively. We the expansion with respect to small
+variable $\varepsilon$:
+\begin{align}\label{eqn:f-expan}
+  f(\vect x, t) &= f_0(\vect x, t) + \varepsilon f_1(\vect x, t)
+  +\varepsilon^2 f_2(\vect x, t) + \mo (\varepsilon^3),
+\end{align}
+and 
+\begin{align}\label{eqn:A-expan}
+  \fwg^\beta(t) = \fwg^\beta_0 + \varepsilon\fwg_1(t),
+\end{align}
+where
 \begin{align}
-  f(\vect x, t) &= f_0(\vect x) + \Delta  f(\vect x, t)
+  \fwg_0^\beta =&
+  \frac{\sigma^2}2\Delta_{\vect p}
+  -
+  \nabla_{\vect p}\mh\cdot\nabla_{\vect q}
+  - \big(
+  -\nabla_{\vect q}\mh - \gamma\vect p
+  \big)\cdot\nabla_{\vect p}
+  + 3N\gamma,\\
+  \fwg_1(t) =&
+  - \frac{F_e(t)}{\varepsilon} \big(
+  \vect C \cdot\nabla_{\vect q}
+  +
+  \vect D \cdot\nabla_{\vect p}
+  \big).
 \end{align}
-and since the Eqn.~\eqref{eqn:gov-1} and \eqref{eqn:gov-2} imply that
-$\dot{\vect x} = \dot{\vect x}_0 + \Delta \dot{\vect x}(t)$
+Notice $F_e(t) \sim \mo(\varepsilon)$, so $\Vert\fwg_1(t)\Vert\sim\mo(1)$.
+The invariant measure of $\fwg^\beta_0$ satisfies $\fwg^\beta_0f^\beta_{\inv}(\vect x) = 0$,
+which is nothing but the Boltzmann distribution
 \begin{align}
-  i\ml =
-  (\dot{\vect x}_0 + \Delta \dot{\vect x}(t))\cdot\nabla_{\vect x}
-  = i\ml_0 + i\Delta\ml(t)
+  f^\beta_{\inv}(\vect x) \propto e^{-\beta\mh(\vect x)}
 \end{align}
-Where the invariant measure satisfies $i\ml_0f_0(\vect x) = 0$. Also
-$f_0(\vect x)$ has the form $f_0(\mh(\vect x))$ (?). The Liouville equation
-for the perturbed system is therefore:
+Inserting \eqref{eqn:f-expan} and \eqref{eqn:A-expan} into
+\eqref{eqn:orig-forward}, 
+the Kolmogorov forward equation becomes:
 \begin{align}
   \frac{\partial}{\partial t}
-  [f_0(\mh(\vect x)) + \Delta  f(\vect x, t)]
-  +
-  [i\ml_0 + i\Delta\ml(t)]
-  [f_0(\mh(\vect x)) + \Delta  f(\vect x, t)] = 0
+  [f_0(\vect x, t) + \varepsilon  f_1(\vect x, t) + \varepsilon^2f_2(\vect x, t)]
+  -
+  [\,\fwg^\beta_0 + \varepsilon\fwg_1(t)\,]
+  [f_0(\vect x, t) + \varepsilon  f_1(\vect x, t) + \varepsilon^2f_2(\vect x, t)]
+  = 0
 \end{align}
-The basic assumption for the ``linear response'' is that the high order
-term $i\Delta\ml(t)  \Delta  f(\vect x, t)$ in the Liouville equation
-can be neglected. Noticing $i\ml_0 f_0 = 0$, we have the Liouville equation
-for the leading order of perturbation:
-\begin{align}
+Matching this equation at different orders of $\varepsilon$, the zeroth
+order gives:
+\begin{align}\label{eqn:e-o0}
   \bigg[
   \frac{\partial}{\partial t}
-  + i\ml_0
+  - \fwg_0^\beta
   \bigg]
-  \Delta  f(\vect x, t)
+  f_0(\vect x, t)
+  = 0
+\end{align}
+which is the unperturbed equation. Matching the first and the second order gives:
+\begin{align}\label{eqn:e-o1}
+  \bigg[
+  \frac{\partial}{\partial t}
+  - \fwg_0^\beta
+  \bigg]
+  f_1(\vect x, t)
+  =&
+  \fwg_1(t) f_0(\vect x, t)\\\label{eqn:e-o2}
+  \bigg[
+  \frac{\partial}{\partial t}
+  - \fwg_0^\beta
+  \bigg]
+  f_2(\vect x, t)
+  =&
+  \fwg_1(t) f_1(\vect x, t)
+\end{align}
+Formally solving \eqref{eqn:e-o0} -- \eqref{eqn:e-o2} with
+inital conditions $f_1(\vect x, 0) = 0$ and $f_2(\vect x, 0) = 0$
+gives
+\begin{align}\label{eqn:f0-0}
+  f_0(\vect x, t)
+  =&\,
+  e^{t\fwg^\beta_0}f_0(\vect x, 0) \\\label{eqn:f1-0}
+  f_1(\vect x, t)
+  =&\,
+  \int_0^t\dd s\,
+  e^{(t-s)\fwg^\beta_0}\circ
+  \fwg_1(s) f_0(\vect x,s)
   =
-  - i\Delta\ml(t) f_0(\mh(\vect x))
+  \int_0^t\dd s\,
+  e^{(t-s)\fwg^\beta_0}\circ
+  \fwg_1(s)\circ
+  e^{s\fwg^\beta_0}f_0(\vect x, 0) \\\nonumber
+  f_2(\vect x, t)
+  =&\,
+  \int_0^t\dd s\,
+  e^{(t-s)\fwg^\beta_0}\circ
+  \fwg_1(s) f_1(\vect x,s) \\\label{eqn:f2-0}
+  &=
+  \int_0^t\dd s
+  \int_0^s\dd u\,\,
+  e^{(t-s)\fwg^\beta_0}\circ
+  \fwg_1(s)\circ
+  e^{(s-u)\fwg^\beta_0}\circ
+  \fwg_1(u)\circ
+  e^{u\fwg^\beta_0}
+  f_0(\vect x, 0) 
 \end{align}
-The operator on l.h.s. is the same as the unperturbed Liouville equation.
-Formally solving this equation with initial condition
-$\Delta f(\vect x, 0) = 0$ gives
+If the initial distribution is the invariant distribution, i.e.
+$f_0(\vect x, 0) = f_{\inv}^\beta(\vect x)$,
+\eqref{eqn:f0-0} becomes:
 \begin{align}
-  \Delta  f(\vect x, t) =
-  - \int_0^t \dd s \:
-  e^{-i\ml_0(t-s)}
-  i\Delta\ml(s)f_0(\mh(\vect x))
+  f_0(\vect x, t) = f_{\inv}^\beta(\vect x)
 \end{align}
-By defining the \emph{dissipative flux},
+It can be shown that
+\begin{align}
+  \fwg_1(s) f_{\inv}^\beta(\vect x)
+  =
+  -\frac{F_e(s)}{\varepsilon}\beta
+  j(\vect x)
+  f_{\inv}^\beta(\vect x),
+\end{align}
+where the \emph{dissipative flux} $j(\vect x)$ is defined by,
 \begin{align}
   j(\vect x) =
-  -\sum_{i=1}^{3N}
-  \bigg[
-  \vect D_i(\vect x) \frac{\partial\mh}{\partial \vect p_i} +
-  \vect C_i(\vect x) \frac{\partial\mh}{\partial \vect q_i}   
-  \bigg]
+  -\vect C\cdot\nabla_{\vect q}\mh 
+  -\vect D\cdot\nabla_{\vect p}\mh
 \end{align}
-it has been shown that~\cite{tuckeman2010statistical}
+Then $f_1(\vect x, t)$ becomes:
 \begin{align}
-  i\Delta\ml(s)f_0(\mh(\vect x)) =
-  -\frac{\partial f_0(\mh)}{\partial\mh} j(\vect x)\,F_e(s)
+  f_1(\vect x, t)
+  =
+  -\frac\beta\varepsilon
+  \int_0^t\dd s\,
+  F_e(s)\,
+  e^{(t-s)\fwg^\beta_0}
+  \big[
+  j(\vect x)
+  f_{\inv}^\beta(\vect x)
+  \big]
 \end{align}
-For the canonical invariant measure, this is equivalent to
-\begin{align}
-  i\Delta\ml(s)f_0(\mh(\vect x)) =
-  \beta f_0(\mh(\vect x)) j(\vect x)F_e(s)
+Therefore, the linear order approximation to
+the solution of~\eqref{eqn:orig-forward} is 
+\begin{align}\label{eqn:solv-orig-linear}
+  f(\vect x, t) =
+  f_{\inv}^\beta(\vect x)
+  - \beta
+  \int_0^t\dd s\,
+  F_e(s)\,
+  e^{(t-s)\fwg^\beta_0}
+  \big[
+  j(\vect x)
+  f_{\inv}^\beta(\vect x)
+  \big]
+  + \mo(\varepsilon^2)
 \end{align}
-For any observation:
+For any time-dependent observation, the linear response approximation
+would be:
 \begin{align}\nonumber
-  \mathcal A(t)
+  \mathcal O(t)
   &=
-  \int\dd \vect x A(\vect x)f(\vect x, t)  \\\nonumber
+  \int\dd \vect x \:O(\vect x)f(\vect x, t)  \\\nonumber
   &=
-  \int\dd \vect x\, A(\vect x)\,
-  [f_0(\mh(\vect x)) + \Delta f(\vect x, t)] \\\nonumber
+  \int\dd \vect x\, O(\vect x)\,
+  [\,f_{\inv}^\beta(\vect x) + \varepsilon f_1(\vect x, t)] \\\nonumber
   &=
-  \langle A\rangle_0
+  \langle O\rangle
   -
   \beta
-  \int \dd \vect x
-  A(\vect x)
-  \int_0^t\dd s\,
-  e^{-i\ml_0(t-s)}
-  f_0(\mh(\vect x)) j(\vect x)F_e(s)  \\\nonumber
+  \int \dd \vect x\:
+  O(\vect x)
+  \int_0^t\dd s\,F_e(s)
+  e^{(t-s)\fwg_0^\beta}
+  \big[j(\vect x)f^\beta_{\inv}(\vect x)\big]
+  \\\nonumber
   &=
-  \langle A\rangle_0
+  \langle O\rangle
   -
@@
   \int \dd \vect x\,
-  A(\vect x)
-  e^{-i\ml_0(t-s)}
-  f_0(\mh(\vect x)) j(\vect x)  \\
+  O(\vect x)\,
+  e^{(t-s)\fwg_0^\beta}
+  \big[j(\vect x)f^\beta_{\inv}(\vect x)\big]
+  \\\nonumber
   &=
-  \langle A\rangle_0
+  \langle O\rangle
   -
@@
   \int \dd \vect x\,
-  f_0(\mh(\vect x))
-  A(\vect x)
-  e^{-i\ml_0(t-s)}
-  j(\vect x)  
-\end{align}
-Notice that $e^{i\ml_0(t-s)}A(\vect x) = A(\phi_{t-s}(\vect x))$,
-where $\phi_t(\vect x)$ is the flow mapping of the \emph{unperturbed}
-Hamiltonian system. We have
-\begin{align}
-  \mathcal A(t)
-  =
-  \langle A\rangle_0
+  f^\beta_{\inv}(\vect x) j(\vect x)\,
+  e^{(t-s)\bwg^\beta_0}
+  O(\vect x)\\ \label{eqn:eqi-pert-1}
+  &=
+  \langle O\rangle
   -
@@
   \int_0^t\dd s\,
-  F_e(s)
+  F_e(t-s)
   \int \dd \vect x\,
-  f_0(\mh(\vect x))
-  A(\phi_{t-s}(\vect x))
-  j(\vect x)  
+  f^\beta_{\inv}(\vect x) j(\vect x)\,
+  e^{s\bwg_0^\beta}
+  O(\vect x)
+\end{align}
+Where $\bwg^\beta_0$ is the infiniesimal generator of the \emph{unperturbed} 
+Kolmogorov backward equation, which is the adjoint operator of $\fwg^\beta_0$.
+We have
+\begin{align}
+  e^{s\bwg^\beta_0} O(\vect x) = \mathbb E_{\vect x} [O(\vect X^\beta_s)]
+\end{align}
+where $\vect X^\beta_t$ is the trajectory of the unperturbed Langevin equation
+starting at $\vect x$.
+Therefore, we have
+% Notice that $e^{i\ml_0(t-s)}O(\vect x) = O(\phi_{t-s}(\vect x))$,
+% where $\phi_t(\vect x)$ is the flow mapping of the \emph{unperturbed}
+% Hamiltonian system. We have
+\begin{align}
+  \mathcal O(t)
+  =
+  \langle O\rangle
+  -
+  \beta
+  \int_0^t\dd s\,
+  F_e(t-s)
+  \int \dd \vect x\,
+  f^\beta_{\inv}(\vect x)
+  j(\vect x)\,
+  \mathbb E_{\vect x} [O(\vect X^\beta_s)]
 \end{align}
@@
 \begin{align}\label{eqn:lr}
-  \mathcal A(t)
+  \mathcal O(t)
   =
-  \langle A\rangle_0
+  \langle O\rangle
   -
@@
   \int_0^t\dd s\,
-  F_e(s)
-  \langle
-  A(t-s)\,
+  F_e(t-s)
+  \big\langle
   j(0)\,
-  \rangle_0
+  \mathbb E_{\vect x} [O(\vect X^\beta_s)]
+  \big\rangle
 \end{align}
-where $\langle A(t-s)\,j(0)\,\rangle_0$ is the \emph{equilibrium time
-  correlation} between the observation $A$ at time $t-s$ and the
-dissipative flux $j$ at time 0.
+where $  \langle
+  j(0)\,
+  \mathbb E_{\vect x} [O(\vect X^\beta_s)]
+  \rangle$ is the \emph{equilibrium time
+  correlation} between the observation $O$ at time $s$ and the
+dissipative flux $j$.
 
-\subsection{The Core set as an observation of the system}
+\subsection{The linear response theory for the core set identification}
 
@@
 identify core set is to consider the ``stability'' of the invariant
-measure $f_0$ under some mapping $\mt_\alpha$ of time step $\alpha$:
+measure $f^\beta_{\inv}$ under a propagator $\mt_\tau$
+of time step $\tau$.
+For example, the propagator is chosen $\mt_\tau = e^{\tau\fwg_t^{2\beta}}$,
+i.e. integrating the Langevin equation~\eqref{eqn:langevin-1}
+and \eqref{eqn:langevin-2} at temperature $1/2\beta$ for time $\tau$.
+Notice $\tau$ is an artificial time that is ``orthogonal'' to the real
+time $t$.
+% , which is associated with a
+% stochastic process denoted by $\vect Y_\alpha$:
+The core set is identified by:
 \begin{align}
-  C_\alpha = \big\{
-  \vect x \,\vert\, \mt_\alpha f_0(\vect x) > f_0(\vect x)
+  C_\tau(t) = \big\{
+  \vect x \,\vert\, \mt_\tau f(\vect x, t) > f(\vect x, t)
   \big\}
@@
 \begin{align}
-  \mathcal P f_0 (\vect y)
+  \proj f (\vect y, t)
   =
@@
   \widetilde
-  C_\alpha = \big\{
+  C_\tau(t) = \big\{
   \vect y \,\vert\,
-  \mathcal P\mt_\alpha f_0(\vect y) > \mathcal P f_0(\vect y)
+  \proj\mt_\tau f(\vect y, t) > \proj f(\vect y, t)
   \big\}
@@
 % \end{align}
-For convenience, we define $\Delta \mt_\alpha =
-\mt_\alpha - \mathcal I$, where $\mathcal I$ is the
-identity. If $[\mathcal P\circ\Delta \mt_\alpha f_0](\vect y) > 0$, then
-$\vect y$ is in the core set $\widetilde C_\alpha $.
+For convenience, we define $\Delta \mt_\tau =
+\mt_\tau - \mathcal I$, where $\mathcal I$ is the
+identity. If $[\proj\circ\Delta \mt_\tau f](\vect y, t) > 0$, then
+$\vect y$ is in the core set $\widetilde C_\tau(t) $.
 \begin{align}\nonumber
-  [\mathcal P\circ\Delta \mt_\alpha f_0](\vect y)
+  [\proj\circ\Delta \mt_\tau f](\vect y, t)
   &=
@@
   \dd \vect x\,
-  \Delta \mt_\alpha f_0(\vect x) \,
-  K_{\sigma, \vect y} (\vect x)  \\\nonumber
-  &=
-  \lim_{\sigma \rightarrow 0}
-  \int
-  \dd \vect x\,
-  f_0(\vect x) \,
-  [\Delta \mt_\alpha]^\dagger K_{\sigma, \vect y} (\vect x)\\
-  &=
-  \lim_{\sigma \rightarrow 0}\,
-  \langle
-  [\Delta \mt_\alpha]^\dagger K_{\sigma, \vect y}
-  \rangle_0
+  \Delta \mt_\tau f(\vect x, t) \,
+  K_{\sigma, \vect y} (\vect x)
+  % \\\nonumber
+  % &=
+  % \lim_{\sigma \rightarrow 0}
+  % \int
+  % \dd \vect x\,
+  % f_0(\vect x) \,
+  % [\Delta \mt_\alpha]^\dagger K_{\sigma, \vect y} (\vect x)\\
+  % &=
+  % \lim_{\sigma \rightarrow 0}\,
+  % \langle
+  % [\Delta \mt_\alpha]^\dagger K_{\sigma, \vect y}
+  % \rangle_0
 \end{align}
-where $[\Delta \mt_\alpha]^\dagger$ is the adjoint operator of $\Delta
-\mt_\alpha$, which is actually $\Delta \mt_{-\alpha}$. The delta
-function is replaced by the limit of a mollifier $K_\sigma(\vect y)$,
-which is infinitely smooth, compact supported of size $\sigma$,
-bounded and $\int
-K_\sigma(\vect y)\dd \vect y = 1$.  The physical meaning is that we
-can consider the expectation value of the stability of the mollifier
-$K_{\sigma, \vect y}$ instead of the stability of the invariant measure.
-This provides the convenience of using the linear response theory showed
-in the previous section. By Eqn.~\eqref{eqn:lr},
+For convenience, we denote $f_t(\vect x) = f(\vect x, t)$.
+Notice that $\mt_\tau f_t(\vect x) = e^{\tau \fwg_t^{2\beta}} f_t(\vect x)$
+is the solution of
+\begin{align}
+  \frac{\partial}{\partial \tau} f_t(\vect x, \tau)
+  -
+  \fwg^{2\beta}(t)  f_t(\vect x, \tau)
+  =0
+\end{align}
+with inital condition $f_t(\vect x, \tau = 0) = f_t(\vect x)$. We do the same
+expansion as before, and then
+matching different orders of $\varepsilon$ gives:
+\begin{align}
+  &\bigg[
+  \frac{\partial}{\partial \tau}
+  - \fwg_0^{2\beta}
+  \bigg]
+  f_{t,0}(\vect x, \tau)
+  = 0 \\
+  &\bigg[
+  \frac{\partial}{\partial \tau}
+  - \fwg_0^{2\beta}
+  \bigg]
+  f_{t,1}(\vect x, \tau)
+  =
+  \fwg_1(t) f_{t,0}(\vect x, \tau)
+\end{align}
+Formally solving these equations with
+inital conditions $f_{t,0}(\vect x, 0) = f_t(\vect x) = f(\vect x, t)$
+and  $f_{t,1}(\vect x, 0) = 0$ 
+gives
+\begin{align}\label{eqn:f0-1}
+  f_{t,0}(\vect x, \tau)
+  =&\,
+  e^{\tau\fwg^{2\beta}_0}f_{t,0}(\vect x, 0) \\\label{eqn:f1-1}
+  f_{t,1}(\vect x, \tau)
+  =&\,
+  \int_0^\tau\dd \eta\,
+  e^{(\tau-\eta)\fwg^{2\beta}_0}\circ
+  \fwg_1(t) f_{t,0}(\vect x,\eta)
+  =
+  \int_0^\tau\dd \eta\,
+  e^{(\tau-\eta)\fwg^{2\beta}_0}\circ
+  \fwg_1(t)\circ
+  e^{\eta\fwg^{2\beta}_0}f_{t,0}(\vect x, 0) 
+\end{align}
+By inserting the linear order approximated solution~\eqref{eqn:solv-orig-linear}
+in to the above equations, we have:
 \begin{align}\nonumber
-  [\mathcal P\circ\Delta \mt_\alpha f_t](\vect y)
-  &=
-  \lim_{\sigma \rightarrow 0}\,
-  \langle
-  [\Delta \mt_\alpha]^\dagger K_{\sigma, \vect y}
-  \rangle_t \\
-  &=
-  \lim_{\sigma \rightarrow 0}\,
-  \Big\{
-  \langle
-  [\Delta \mt_\alpha]^\dagger K_{\sigma, \vect y}
-  \rangle_0  -
+  \mt_\tau f_t(\vect x)=
+  f_{t}(\vect x, \tau)
+  =&\,
+  e^{\tau\fwg^{2\beta}_0} f_{\inv}^\beta(\vect x) -
   \beta
   \int_0^t\dd s\,
+  F_e(s)\,
+  e^{\tau\fwg^{2\beta}_0} \circ
+  e^{(t-s)\fwg^\beta_0}
+  \big[
+  j(\vect x)
+  f_{\inv}^\beta(\vect x)
+  \big]
+  \\\label{eqn:solve-tau-linear}
+  &\,+ \varepsilon
+  \int_0^\tau\dd \eta\,
+  e^{(\tau-\eta)\fwg^{2\beta}_0}\circ
+  \fwg_1(t)\circ
+  e^{\eta\fwg^{2\beta}_0}f_{\inv}^\beta(\vect x)  + \mo(\varepsilon^2)
+\end{align}
+By inserting~\eqref{eqn:solv-orig-linear} and~\eqref{eqn:solve-tau-linear}
+into~\eqref{eqn:core-identify}, we have, up to the first order
+approximation,
+\begin{align}\nonumber
+  [\proj\circ\Delta\mt_\tau f](\vect y, t)
+  \approx
+  \lim_{\sigma \rightarrow 0}
+  \bigg\{
+    &\,
+    \int \dd \vect x\,
+    \big(e^{\tau\fwg_0^{2\beta}}-\mi\big)
+    f_{\inv}^\beta(\vect x)\,
+    K_{\sigma,\vect y}(\vect x)\\\nonumber
+    &
+    - \int_0^t\dd s\,
     F_e(s)
-  \langle
-  [\Delta \mt_\alpha]^\dagger K_{\sigma, \vect y}(t-s)\,
-  j(0)\,
-  \rangle_0
+    \int\dd \vect x\,
+    \big(e^{\tau\fwg^{2\beta}_0} - \mi\big)\circ
+    e^{(t-s)\fwg^\beta_0}
+    \big[
+    j(\vect x)
+    f_{\inv}^\beta(\vect x)
+    \big]\, K_{\sigma,\vect y}(\vect x) \\\nonumber
+    &
+    +
+    \varepsilon
+    \int_0^\tau\dd \eta
+    \int\dd \vect x\,
+    e^{(\tau-\eta)\fwg^{2\beta}_0}\circ
+    \fwg_1(t)\circ
+    e^{\eta\fwg^{2\beta}_0}f_{\inv}^\beta(\vect x)\,K_{\sigma,\vect y}(\vect x)
+    \bigg\}     \\\nonumber
+  \approx
+  \lim_{\sigma \rightarrow 0}
+  \bigg\{
+  &\,
+    \Big\langle
+    \mathbb E_{\vect x}[ K_{\sigma,\vect y}(\vect X^{2\beta}_\tau) ]
+    - K_{\sigma,\vect y}(\vect x)
+    \Big\rangle\\\nonumber
+    &
+    - \int_0^t\dd s\,
+    F_e(s)
+    \Big\langle
+    j(\vect x)\,
+    \mathbb E_{\vect x}\big\{
+    \mathbb E_{\vect X_{t-s}^\beta}[
+    K_{\sigma,\vect y}(\vect X_\tau^{2\beta})
+    ]
+    -
+    K_{\sigma,\vect y}(\vect X_{t-s}^\beta)
+    \big\}
+    \Big\rangle\\\label{eqn:core-identify-approx-1}
+    &
+    +
+    \varepsilon
+    \int_0^\tau\dd \eta
+    \int\dd \vect x\,
+    e^{(\tau-\eta)\fwg^{2\beta}_0}\circ
+    \fwg_1(t)\circ
+    e^{\eta\fwg^{2\beta}_0}f_{\inv}^\beta(\vect x)\,K_{\sigma,\vect y}(\vect x)
+    \bigg\}     
+\end{align}
+Unfortunately the last term on the r.h.s. of~\eqref{eqn:core-identify-approx-1}
+is not easy to calculate. We will deal with it with
+the path integral method introduced later.
+
+\subsection{Path integral}
+The path integral is very useful to develop the non-equilibrium version of
+linear response theory. 
+Comparing with the idea of ensemble, which is a collection of state points
+in the configuration space, the path integral consider the ``ensemble'' of
+trajectories.
+(In the following, the develop of formula are formal. The strict mathematical
+theory is still unclear for me.)
+First, we take the standard 1-dimensional Wiener process for example.
+Consider the time interval $[0,T]$, the probability density of one trajectory
+is given by
+\begin{align}
+  p[W_t] = \frac 1Z e^{-I[W_t]}
+\end{align}
+where $I$ is the action functional of Wiener process given by
+\begin{align}
+  I[W_t] = \int_0^T\frac12 \dot{W}_t^2\dd t
+\end{align}
+the partition function $Z$ is
+\begin{align}
+  Z = \int  e^{-I[W_t]} \prod_{0\leq t\leq T}\dd W_t
+\end{align}
+Consider the following two stochastic processes:
+\begin{align}\label{eqn:st-orig}
+  \dd \vect X_t &= \vect b(\vect X_t, t) \dd t + \bsigma(\vect X_t, t)\,\dd\vect W_t \\\label{eqn:st-pert}
+  \dd \vect Y_t &= \vect b(\vect Y_t, t) \dd t + \vect c(\vect Y_t, t) \dd t + \bsigma(\vect Y_t, t)\,\dd\vect W_t
+\end{align}
+Where $\vect W_t$ is the standard Wiener process. Obviously
+\eqref{eqn:st-pert} is the perturbed version of
+\eqref{eqn:st-orig}. It has been shown that, starting from the same
+measure saying $\mu$ (which is not necessarily equilibrium or
+stationary), the ratio of trajectory probability density between the perturbed
+and unperturbed process is~\cite{li2010lecture} (\recheck{I think here
+  the incompressibility of phase space should be assumed. Otherwise the
+  ratio should be more complicated.})
+\begin{align}\label{eqn:girsanov}
+  \frac{\dd\mathcal P_{\vect Y}}{\dd\mathcal P_{\vect X}}
+  =
+  \exp
+  \Big\{
+  -\frac12 \int_0^T\vert\bsigma^{-1}\cdot\vect c(\vect X_t,t)\vert^2\,\dd t
+  +\int_0^T[\dot{\vect X}_t - \vect b(\vect X_t,t)]^T
+  \cdot\bsigma^{-T}\cdot\bsigma^{-1}\cdot\vect c(\vect X_t,t)\,\dd t 
   \Big\}
 \end{align}
-Therefore, the core set at time $t$ is defined by
+where $\mathcal P_{\vect Y}$ and $\mathcal P_{\vect Y}$ are measures
+of the trajectories generated by \eqref{eqn:st-pert} and
+\eqref{eqn:st-orig}, respectively.  This result (Girsanov formula?) is
+(formally) precise without any approximation.
+
+To be more general, we study the two perturbed Langevin dynamics
+(i.e. Eqn.~\eqref{eqn:langevin-1} and \eqref{eqn:langevin-2}) with
+different perturbations that are given by $F_e^1(t)$ and $F_e^2(t)$,
+respectively. 
+By writing Eqn.~\eqref{eqn:langevin-1} and \eqref{eqn:langevin-2} as
 \begin{align}
-  \widetilde C_\alpha^t =
+  \left[
+    \begin{array}{l}
+      \dd\vect q\\
+      \dd\vect p
+    \end{array}
+  \right]
+  =
+  \left[
+    \begin{array}{l}
+      \nabla_{\vect p}\mh\\
+      -\nabla_{\vect q}\mh - \gamma\vect p + F_e\vect D
+    \end{array}
+  \right]
+  \dd t
+  +
+  \left[
+    \begin{array}{cc}
+      \epsilon \vect I & 0\\
+      0 & \sigma\vect I
+    \end{array}
+  \right]
+  \cdot
+  \left[
+    \begin{array}{l}
+      \dd\vect W_t\\
+      \dd\vect W_t
+    \end{array}
+  \right]
+\end{align}
+Using Girsanov's formula, to the leading order of $\Delta F_e = F_e^2 - F_e^1$,
+we have
+\begin{align}\label{eqn:girsanov-1}
+  \frac{\dd\mathcal P^2[\vect X_t]}{\dd\mathcal P^1[\vect X_t]}
+  &=
+  \exp
   \Big\{
-  \vect y
-  \,\Big\vert\,
-  [\mathcal P\circ\Delta \mt_\alpha f_t](\vect y) > 0
+  \frac 1 {\sigma^2}
+  \int_0^t
+  [\dot{\vect p}_s +
+  \gamma\vect p_s +
+  \nabla_{\vect q}\mh(\vect q_s,\vect p_s)
+  - F_e^1(s) \vect D(\vect q_s,\vect p_s)
+  ]\cdot
+  \vect D(\vect q_s,\vect p_s)
+  \Delta F_e(s)\,
+  \dd s
+  + \mathcal O(\Delta F_e^2)
   \Big\} 
 \end{align}
+
+\subsection{The linear response theory for the core set identification,
+  revisited by the path integral}
+For some observable $O[\vect X_\eta]$, which is a functional of the
+stochastic process $\vect X_\eta$,
+the trajectory ensemble average yields:
+\begin{align}
+  \mo(t)
+  =
+  \int \dd \mu_t(\vect x)
+  \int_{\mc\{\vect x,0;\tau\}}
+  O[\vect X_\eta]\,
+  \dd \mathcal P^{2\beta}_t[\vect X_\eta],
+\end{align}
+where $\mu_t(\vect x)$
+is the measure of the initial position, and obviously, we have
+$\dd \mu_t(\vect x) = f_t(\vect x)\,\dd \vect x$.
+The trajectory $\vect X_\eta \in {\mc\{\vect x,0;\tau\}}$, where
+${\mc\{\vect x,0;\tau\}}$ denotes the set of all continuous but not
+differentiable trajectories starting at $\vect x$ at artificial
+time 0, and ending
+at artificial time $\tau$.
+$\mathcal P^{2\beta}_t$ denotes the measure of trajectories
+developped under temperature $1/2\beta$
+at real time $t$.
+Since the perturbation is a function of time, this
+measure also depends on time. The reference artificial dynamics is chosen
+at $t=0$.
+In this case, \eqref{eqn:girsanov-1} rewrites:
+\begin{align}\nonumber
+  \frac{\dd \mathcal P^{2\beta}_t[\vect X_\eta]}
+  {\dd \mathcal P^{2\beta}_0[\vect X_\eta]}\,
+  &=
+  \exp
+  \Big\{
+  \frac 1 {\sigma^2}
+  \int_0^\tau
+  [\dot{\vect p}_\eta +
+  \gamma\vect p_\eta +
+  \nabla_{\vect q}\mh(\vect q_\eta,\vect p_\eta)
+  ]\cdot
+  \vect D(\vect q_\eta,\vect p_\eta)
+  F_e(t)\,
+  \dd \eta
+  + \mathcal O(\varepsilon^2)
+  \Big\}\\
+  &=
+  1+
+  F_e(t)\,
+  \frac 1 {\sigma^2}
+  \int_0^\tau
+  [\dot{\vect p}_\eta +
+  \gamma\vect p_\eta +
+  \nabla_{\vect q}\mh(\vect q_\eta,\vect p_\eta)
+  ]\cdot
+  \vect D(\vect q_\eta,\vect p_\eta)\,
+  \dd \eta
+  + \mathcal O(\varepsilon^2)  
+\end{align}
+Notice that $F_e(t)\sim \mo(\varepsilon)$.
+For convenience, we denote
+\begin{align}
+  G[\vect X_\eta] &=
+    \frac 1 {\sigma^2}
+  \int_0^\tau
+  [\dot{\vect p}_\eta +
+  \gamma\vect p_\eta +
+  \nabla_{\vect q}\mh(\vect q_\eta,\vect p_\eta)
+  ]\cdot
+  \vect D(\vect q_\eta,\vect p_\eta)\,
+  \dd \eta =
+  \frac1\sigma\int_0^\tau
+  \vect D(\vect q_\eta,\vect p_\eta)
+  \cdot\dd \vect W_\eta
+\end{align}
+where $\vect W_\eta$ is the Wiener process that generates
+$\vect X_\eta$.
+Therefore, we have
+\begin{align}\nonumber
+  \mo(t)
+  =&\,
+  \int \dd \vect x
+  f(\vect x, t)
+  \int_{\mc\{\vect x,0;\tau\}}
+  O[\vect X_\eta]\,
+  \frac{\dd \mathcal P^{2\beta}_t[\vect X_\eta]}
+  {\dd \mathcal P^{2\beta}_0[\vect X_\eta]}\,
+  \dd \mathcal P^{2\beta}_0[\vect X_\eta]  \\\nonumber
+  =&\,
+  \int \dd \vect x
+  \big[
+  f_{0}(\vect x, t)
+  + \varepsilon f_{1}(\vect x, t)
+  + \mo(\varepsilon^2)
+  \big]
+  \int_{\mc\{\vect x,0;\tau\}}
+  O[\vect X_\eta]\,
+  \big[
+  1 
+  + F_e(t) G[\vect X_\eta]
+  + \mo(\varepsilon^2)
+  \big]
+  \dd \mathcal P^{2\beta}_0[\vect X_\eta]
+  \\\nonumber
+  =&\,
+  \int \dd \vect x
+  f_{0}(\vect x, t)
+  \int_{\mc\{\vect x,0;\tau\}}
+  O[\vect X_\eta]\,
+  \dd \mathcal P^{2\beta}_0[\vect X_\eta]
+  +
+  \varepsilon
+  \int \dd \vect x\,
+  f_{1}(\vect x, t)
+  \int_{\mc\{\vect x,0;\tau\}}
+  O[\vect X_\eta]\,
+  \dd \mathcal P^{2\beta}_0[\vect X_\eta]\\\label{eqn:solve-tau-linear-2-0}
+  &+
+  F_e(t)
+  \int \dd \vect x\,
+  f_{0}(\vect x, t)
+  \int_{\mc\{\vect x,0;\tau\}}
+  O[\vect X_\eta]\,
+  G[\vect X_\eta]\,
+  \dd \mathcal P^{2\beta}_0[\vect X_\eta]
+  + \mo(\varepsilon^2) 
+\end{align}
+In the case of core set identification,
+\begin{align}\nonumber
+  [\proj\circ \mt_\tau] f(\vect y,t)
+  =&\,
+  \int\dd \vect x\,
+  \mt_\tau f(\vect x,t) K_{\sigma,\vect y}(\vect x)\\\nonumber
+  =&\,
+  \int\dd \vect x\,
+  e^{\tau\fwg_t^{2\beta}} f(\vect x,t) \,K_{\sigma,\vect y}(\vect x)\\\nonumber
+  =&\,
+  \int\dd \vect x\,
+  f(\vect x,t) \,e^{\tau\bwg_t^{2\beta}}K_{\sigma,\vect y}(\vect x)  \\\nonumber
+  =&\,
+  \int\dd \vect x\,
+  f(\vect x,t) \,\mathbb E_{\vect x}
+  \big[
+  K_{\sigma,\vect y}(\vect X^{2\beta}_{t,\tau})
+  \big]  
+  \\
+  =&\,
+  \int\dd \vect x\,
+  f(\vect x,t) 
+  \int_{\mc\{\vect x,0;\tau\}} K_{\sigma,\vect y}(\vect X_\tau)\,
+  \dd\mathcal P^{2\beta}_t[\vect X_\eta]
+\end{align}
+So the observable $O[\vect X_\eta]$
+is:
+\begin{align}
+  O[\vect X_\eta] = 
+  K_{\sigma,\vect y}(\vect X_\tau),
+  % -
+  % K_{\sigma,\vect y}(\vect x)
+\end{align}
+which is the end point of the artificial
+trajectory measured by $K_{\sigma,\vect y}$.
+Inserting everything we known in~\eqref{eqn:solve-tau-linear-2-0},
+and notice $G[\vect X_\eta] = 0$ when $\tau = 0$,
+\begin{align}\nonumber
+  [\proj\circ\Delta\mt_\tau f](\vect y, t)
+  =
+  \lim_{\sigma \rightarrow 0}
+  \big\{
+  &\,
+  [\proj\circ\mt_\tau f - \proj\circ\mt_0 f](\vect y, t)
+  \big\}\\\nonumber
+  \approx
+  \lim_{\sigma \rightarrow 0}
+  \bigg\{
+  &\,
+  \int \dd \vect x\,
+  f_\inv^\beta(\vect x)
+  \int_{\mc\{\vect x,0;\tau\}}
+  [K_{\sigma,\vect y}(\vect X_\tau)
+  -K_{\sigma,\vect y}(\vect x)]  \,
+  \dd \mathcal P^{2\beta}_0[\vect X_\eta]\\\nonumber
+  &\,-
+  \beta
+  \int_0^t\dd s\, F_e(s)
+  \int \dd \vect x\,
+  e^{(t-s)\fwg_0^\beta}
+  \big[
+  j(\vect x)
+  f_\inv^\beta(\vect x)
+  \big]
+  \int_{\mc\{\vect x,0;\tau\}}
+  [K_{\sigma,\vect y}(\vect X_\tau)
+  -K_{\sigma,\vect y}(\vect x)]  \,
+  \dd \mathcal P^{2\beta}_0[\vect X_\eta]\\\nonumber
+  &\,+
+  F_e(t)
+  \int \dd \vect x\,
+  f_\inv^\beta(\vect x)
+  \int_{\mc\{\vect x,0;\tau\}}
+  K_{\sigma,\vect y}(\vect X_\tau)
+  % \Big\{
+  % \int_0^\tau\dd \eta\,
+  % [\dot{\vect p}_\eta + \gamma\vect p_\eta + \nabla_{\vect q}\mh(\vect X_\eta)]
+  % \cdot\vect D(\vect X_\eta)
+  % \Big\}\,
+  \,G[\vect X_\eta]\,
+  \dd \mathcal P^{2\beta}_0[\vect X_\eta]
+  \bigg\}\\\nonumber
+  \approx
+  \lim_{\sigma \rightarrow 0}
+  \bigg\{
+  &\,
+  \Big\langle
+  \mathbb E_{\vect x}[K_{\sigma,\vect y}(\vect X_\tau^{2\beta})]
+  -
+  K_{\sigma,\vect y}(\vect x)
+  \Big\rangle\\\nonumber
+  &\,
+  - \beta
+  \int_0^t\dd s\, F_e(s)
+  \Big\langle
+  j(\vect x)
+  \mathbb E_{\vect x} \big\{
+  \mathbb E_{\vect X^\beta_{t-s}}\big[
+  K_{\sigma, \vect y}(\vect X^{2\beta}_\tau)
+  \big]
+  -
+  K_{\sigma, y}(\vect X^\beta_{t-s})
+  \big\}
+  \Big\rangle\\\label{eqn:core-identify-approx-2}
+  &\,+
+  F_e(t)
+  \int \dd \vect x\,
+  f_\inv^\beta(\vect x)
+  \int_{\mc\{\vect x,0;\tau\}}
+  K_{\sigma,\vect y}(\vect X_\tau)
+  \,G[\vect X_\eta]\,
+  \dd \mathcal P^{2\beta}_0[\vect X_\eta]
+  \bigg\}
+\end{align}
+Comparing~\eqref{eqn:core-identify-approx-2}
+with~\eqref{eqn:core-identify-approx-1}, the first
+two terms on the r.h.s. are identical. 
+The last term is now computable.
+
+\subsection{Higher orders response theory for core set identification}
+
+Here, we assume that the system undergoes a series of perturbations,
+which are denoted by $F^0_e(t),\ F^1_e(t),\ \cdots,\ F^i_e(t),\ \cdots$.
+Here, $F^0_e(t) = 0$.
+The measure for the trajectory space
+$\mc\{\vect x, 0; t\}$
+is given by
+$\mathcal P_0^\beta[\vect X_s],\ \mathcal P_1^\beta[\vect X_s],\ \cdots,\
+\mathcal P_i^\beta[\vect X_s], \cdots$, respectively.
+The measure for the quenching trajectory are denoted by
+$\mathcal P_{0,t}^{2\beta}[\vect X_\eta],\
+\mathcal P_{1,t}^{2\beta}[\vect X_\eta],\ \cdots,\
+\mathcal P_{i,t}^{2\beta}[\vect X_\eta], \cdots$, respectively.
+The notation $t$ on the subscript means the trajectory is branched from
+the main trajectory at time $t$.
+Now the phase space distribution $f_i(\vect x, t)$ for the
+$i$-th perturbation can be studied in terms
+of path integral:
+\begin{align}
+  f_i(\vect x, t) =
+  \int\dd\vect u\,
+  f_i(\vect u, 0)
+  \int_{\mc\{\vect u,0; \vect x,t\}}
+  \dd\mathcal P_i^\beta[\vect X_s]
+\end{align}
+Applied by operator $\mt_\tau$, we have
+\begin{align}\nonumber
+  \mt_\tau f_i(\vect x, t)
+  =&\,
+  \int\dd\vect u f_i(\vect u, t)
+  \int_{\mc\{\vect u,0; \vect x,\tau\}}
+  \dd\mathcal P_{i,t}^{2\beta}[\vect X_\eta]\\
+  =&\,
+  \int\dd\vect v\,
+  f_i(\vect v, 0)
+  \int\dd\vect u
+  \int_{\mc\{\vect v,0; \vect u,t\}}
+  \dd\mathcal P_i^\beta[\vect X_s]  
+  \int_{\mc\{\vect u,0; \vect x,\tau\}}
+  \dd\mathcal P_{i,t}^{2\beta}[\vect X_\eta]
+\end{align}
+For the core set identification,
+\begin{align}
+  [\proj\circ\mt_\tau f](\vect y, t)
+  &=
+  \int\dd\vect v\,
+  f_i(\vect v, 0)
+  \int\dd\vect u
+  \int_{\mc\{\vect v,0; \vect u,t\}}
+  \dd\mathcal P_i^\beta[\vect X_s]  
+  \int_{\mc\{\vect u,0; \tau\}}
+  K_{\sigma,\vect y}(\vect X_\tau)\,
+  \dd\mathcal P_{i,t}^{2\beta}[\vect X_\eta]  \\
+  [\proj\circ\mt_0 f](\vect y, t)
+  &=
+  \int\dd\vect v\,
+  f_i(\vect v, 0)
+  \int_{\mc\{\vect v,0; t\}}
+  K_{\sigma,\vect y}(\vect X_t)\,
+  \dd\mathcal P_i^\beta[\vect X_s]  
+  % \int_{\mc\{\vect u,0; \tau\}}
+  % K_{\sigma,\vect y}(\vect X_\tau)\,
+  % \dd\mathcal P_{i,t}^{2\beta}[\vect X_\eta]  
+\end{align}
+Here we assume that the system always starts from invariant measure,
+i.e. $f_i(\vect x, 0) = f_{\inv}^\beta(\vect x)$, and that we have a
+good sampling of the $(i-1)$-th perturbation, then
+\begin{align}\label{eqn:pert-exact-1}
+  [\proj\circ\mt_\tau f](\vect y, t)
+  &=
+  \int\dd\vect v\,
+  f_\inv^\beta(\vect v)
+  \int\dd\vect u
+  \int_{\mc\{\vect v,0; \vect u,t\}}
+  \dd\mathcal P_{i-1}^\beta[\vect X_s]  
+  \int_{\mc\{\vect u,0; \tau\}}
+  \dd\mathcal P_{i-1,t}^{2\beta}[\vect X_\eta]\cdot
+  K_{\sigma,\vect y}(\vect X_\tau)\,
+  \frac
+  {\dd\mathcal P_{i}^\beta[\vect X_s]}
+  {\dd\mathcal P_{i-1}^\beta[\vect X_s]}
+  \frac
+  {\dd\mathcal P_{i,t}^{2\beta}[\vect X_\eta]}
+  {\dd\mathcal P_{i-1,t}^{2\beta}[\vect X_\eta]}
+  \\\label{eqn:pert-exact-2}
+  [\proj\circ\mt_0 f](\vect y, t)
+  &=
+  \int\dd\vect v\,
+  f_{\inv}^\beta(\vect v)
+  \int_{\mc\{\vect v,0; t\}}
+  \dd\mathcal P_{i-1}^\beta[\vect X_s]
+  \cdot
+  K_{\sigma,\vect y}(\vect X_t)\,
+  \frac
+  {\dd\mathcal P_{i}^\beta[\vect X_s]}
+  {\dd\mathcal P_{i-1}^\beta[\vect X_s]}  
+  % \int_{\mc\{\vect u,0; \tau\}}
+  % K_{\sigma,\vect y}(\vect X_\tau)\,
+  % \dd\mathcal P_{i,t}^{2\beta}[\vect X_\eta]  
+\end{align}
+Here
+\begin{align}\label{eqn:exact-ratio-1}
+  \frac
+  {\dd\mathcal P_{i}^\beta[\vect X_s]}
+  {\dd\mathcal P_{i-1}^\beta[\vect X_s]}  
+  =&\,
+  \exp
+  \bigg\{
+  \frac{1}\sigma\int_0^t
+  \Delta F_e^{i-1}(s)\,\vect D(\vect q_s,\vect p_s)\cdot\dd\vect W_s
+  -
+  \frac 1{2\sigma^2}\int_0^t
+  [\Delta F_e^{i-1}(s)]^2\,\vect D^2(\vect q_s,\vect p_s)\,\dd s
+  \bigg\}\\\label{eqn:exact-ratio-2}
+  \frac
+  {\dd\mathcal P_{i,t}^{2\beta}[\vect X_\eta]}
+  {\dd\mathcal P_{i-1,t}^{2\beta}[\vect X_\eta]}
+  =&\,
+  \exp
+  \bigg\{
+  \frac{1}\sigma\int_0^\tau
+  \Delta F_e^{i-1}(t)\,\vect D(\vect q_\eta,\vect p_\eta)\cdot\dd\vect W_\eta
+  -
+  \frac 1{2\sigma^2}\int_0^\tau
+  [\Delta F_e^{i-1}(t)]^2\,\vect D^2(\vect q_\eta,\vect p_\eta)\,\dd \eta
+  \bigg\}
+\end{align}
+where $\Delta F_e^{i-1}(t) = F_e^{i}(t) - F_e^{i-1}(t) $.
+If the trajectory at perturbation $ F_e^{i-1}(t) $ is well sampled, then
+\eqref{eqn:exact-ratio-1} and \eqref{eqn:exact-ratio-2} can be measured
+without linear approximation.
+This is the exact result, from where we can develop
+the response (also non-linear) theroy.
+Before doing so, we should assume the perturbation can be expanded,
+and well approximated in a finite dimensional space:
+\begin{align}
+  F^i_e(t)
+  =
+  \sum_{j = 1}^\infty
+  s^i_j M_j(t)
+\end{align}
+with this expansion, we have
+\begin{align}\label{eqn:exact-ratio-11}
+  \frac
+  {\dd\mathcal P_{i}^\beta[\vect X_s]}
+  {\dd\mathcal P_{i-1}^\beta[\vect X_s]}  
+  =&\,
+  \exp
+  \bigg\{
+  \sum_{j=1}^\infty
+  \Delta s_j^{i-1}G_j[\vect X_s]
+  -
+  \frac12 \sum_{j,k=1}^\infty
+  \Delta s_j^{i-1} \Delta s_k^{i-1}H_{jk}[\vect X_s]
+  \bigg\}\\\label{eqn:exact-ratio-21}
+  \frac
+  {\dd\mathcal P_{i,t}^{2\beta}[\vect X_\eta]}
+  {\dd\mathcal P_{i-1,t}^{2\beta}[\vect X_\eta]}
+  =&\,
+  \exp
+  \bigg\{
+  \Delta F_e^{i-1}(t)\,
+  G_0[\vect X_\eta]
+  -
+  \frac12[\Delta F_e^{i-1}(t)]^2
+  H_{00}[\vect X_\eta]
+  \bigg\}
+\end{align}
+where
+\begin{align}\label{eqn:Gj}
+  G_j[\vect X_s]
+  =&\,
+  \frac 1\sigma
+  \int_0^t
+  M_j(s)\,\vect D(\vect X_s)\cdot\dd\vect W_s\\\label{eqn:Hjk}
+  H_{jk}[\vect X_s]
+  =&\,
+  \frac 1{\sigma^2}
+  \int_0^t
+  M_j(s)\,M_k(s)\,\vect D^2(\vect X_s)\cdot\dd s
+\end{align}
+and 
+\begin{align}\label{eqn:G0}
+  G_0[\vect X_\eta]
+  =&\,
+  \frac 1\sigma
+  \int_0^\tau
+  \vect D(\vect X_\eta)\cdot\dd\vect W_\eta\\\label{eqn:H00}
+  H_{00}[\vect X_\eta]
+  =&\,
+  \frac 1{\sigma^2}
+  \int_0^\tau
+  \vect D^2(\vect X_\eta)\cdot\dd \eta
+\end{align}
+Notice that Eqn.~\eqref{eqn:Gj}--\eqref{eqn:H00} are perturbation
+independent, therefore can be measured with the dynamics $i-1$.
+Expanding \eqref{eqn:exact-ratio-11} and \eqref{eqn:exact-ratio-21}
+up to $\mathcal O(\varepsilon^3)$ yields
+\begin{align}\label{eqn:expan-ratio-1}
+  \frac
+  {\dd\mathcal P_{i}^\beta[\vect X_s]}
+  {\dd\mathcal P_{i-1}^\beta[\vect X_s]}
+  =&\,
+  1
+  +
+  \sum_{j=1}^\infty
+  \Delta s_j^{i-1}G_j[\vect X_s]
+  % -
+  % \frac12
+  % \sum_{j,k=1}^\infty
+  % \Delta s_j^{i-1} \Delta s_k^{i-1}H_{jk}[\vect X_s]
+  +
+  \frac12  
+  \sum_{j,k=1}^\infty
+  \Delta s_j^{i-1}\Delta s_k^{i-1}
+  \big\{
+  G_j[\vect X_s]G_k[\vect X_s] - H_{jk}[\vect X_s]
+  \big\}
+  +
+  \mathcal O(\varepsilon^3)
+  \\\label{eqn:expan-ratio-2}
+  \frac
+  {\dd\mathcal P_{i,t}^{2\beta}[\vect X_\eta]}
+  {\dd\mathcal P_{i-1,t}^{2\beta}[\vect X_\eta]}
+  =&\,
+  1 +
+  \Delta F_e^{i-1}(t)\,
+  G_0[\vect X_\eta]
+  % -
+  % \frac12[\Delta F_e^{i-1}(t)]^2
+  % H_{00}[\vect X_\eta]
+  +
+  \frac12[\Delta F_e^{i-1}(t)]^2
+  \big\{
+  G^2_0[\vect X_\eta] - H_{00}[\vect X_\eta]
+  \big\}
+  +
+  \mathcal O(\varepsilon^3)
+\end{align}
+Therefore, by the expansion formula \eqref{eqn:expan-ratio-1} and
+\eqref{eqn:expan-ratio-2}, the response theory for the
+\eqref{eqn:pert-exact-1} and \eqref{eqn:pert-exact-2} can be
+established.
+For convenience, we denote
+\begin{align}
+  \big\langle
+  O[\vect X_s]\tilde O[\vect X_\eta]
+  \big\rangle_{0,t,\tau}^{i-1}
+  =&\,
+  \int\dd\vect v\,
+  f_\inv^\beta(\vect v)
+  \int\dd\vect u
+  \int_{\mc\{\vect v, 0; \vect u, t\}}
+  \dd\mathcal P^\beta_{i-1}[\vect X_s]
+  \,
+  O[\vect X_s]
+  \int_{\mc\{\vect u, 0; \tau\}}
+  \dd\mathcal P^{2\beta}_{i-1,t}[\vect X_\eta]
+  \,
+  \tilde O[\vect X_\eta]
+  \\
+  \big\langle
+  O[\vect X_s]
+  \big\rangle_{0,t,\tau}^{i-1}
+  = &\,
+  \int\dd\vect v\,
+  f_\inv^\beta(\vect v)
+  \int_{\mc\{\vect v, 0; t\}}
+  \dd\mathcal P^\beta_{i-1}[\vect X_s]
+  \,
+  O[\vect X_s]  
+\end{align}
+Then the approximated \eqref{eqn:pert-exact-1} is
+\begin{align}\nonumber
+  [\proj\circ\mt_\tau f](\vect y, t)
+  =&\,
+  \big\langle
+  K_{\sigma,\vect y}(\vect X_\tau)
+  \big\rangle_{0,t,\tau}^{i-1}
+  \\ \nonumber
+  &\, +
+  \Delta F_e^{i-1}(t) 
+  \big\langle
+  K_{\sigma,\vect y}(\vect X_\tau)\,
+  G_0[\vect X_\eta]
+  \big\rangle_{0,t,\tau}^{i-1}
+  +
+  \sum_{j=1}^\infty
+  \Delta s_j^{i-1}
+  \big\langle
+  K_{\sigma,\vect y}(\vect X_\tau)\,
+  G_j[\vect X_s]
+  \big\rangle_{0,t,\tau}^{i-1}
+  \\\nonumber
+  &\, 
+  +
+  \frac12[\Delta F_e^{i-1}(t)]^2
+  \big\langle
+  K_{\sigma,\vect y}(\vect X_\tau)\,
+  \big\{
+  G^2_0[\vect X_s] - 
+  H_{00}[\vect X_\eta]
+  \big\}
+  \big\rangle_{0,t,\tau}^{i-1}
+  % +
+  % \frac12[\Delta F_e^{i-1}(t)]^2
+  % \big\langle
+  % K_{\sigma,\vect y}(\vect X_\tau)\,
+  % \big\{G_{0}[\vect X_\eta]\big\}^2
+  % \big\rangle_{0,t,\tau}^{i-1}
+  \\\nonumber
+  &\,
+  +
+  \frac 12
+  \sum_{j,k=1}^\infty
+  \Delta s_j^{i-1}\Delta s_k^{i-1}
+  \big\langle
+  K_{\sigma,\vect y}(\vect X_\tau)\,
+  \big\{
+  G_j[\vect X_s] G_k[\vect X_s] - 
+  H_{jk}[\vect X_s]
+  \big\}
+  \big\rangle_{0,t,\tau}^{i-1}
+  % +
+  % \frac 12
+  % \sum_{j,k=1}^\infty
+  % \Delta s_j^{i-1}\Delta s_k^{i-1}
+  % \big\langle
+  % K_{\sigma,\vect y}(\vect X_\tau)\,
+  % G_{j}[\vect X_s] G_k[\vect X_s]
+  % \big\rangle_{0,t,\tau}^{i-1}
+  \\\label{eqn:pert-approx-1}
+  &\,+
+  \sum_{j=1}^\infty
+  \Delta F_e^{i-1}(t) \Delta s_j^{i-1}
+  \big\langle
+  K_{\sigma,\vect y}(\vect X_\tau)\,
+  G_0[\vect X_\eta]\,
+  G_j[\vect X_s]
+  \big\rangle_{0,t,\tau}^{i-1}
+\end{align}
+The first line on the r.h.s. of \eqref{eqn:pert-approx-1}
+is the zeroth order approximation,  the second line is the
+first order approximation, and the last three lines are the
+second order approximation.
+And the approximated \eqref{eqn:pert-exact-2} is
+\begin{align}
+  \nonumber
+  [\proj\circ\mt_0 f](\vect y, t)
+  =&\,
+  \big\langle
+  K_{\sigma,\vect y}(\vect X_t)
+  \big\rangle_{0,t}^{i-1}
+  \\ \nonumber
+  &\,+
+  \sum_{j=1}^\infty
+  \Delta s_j^{i-1}
+  \big\langle
+  K_{\sigma,\vect y}(\vect X_t)\,
+  G_j[\vect X_s]
+  \big\rangle_{0,t}^{i-1}
+  \\\label{eqn:pert-approx-2}
+  &\,
+  +
+  \frac 12
+  \sum_{j,k=1}^\infty
+  \Delta s_j^{i-1}\Delta s_k^{i-1}
+  \big\langle
+  K_{\sigma,\vect y}(\vect X_t)\,
+  \big\{
+  G_j[\vect X_s] G_k[\vect X_s] -
+  H_{jk}[\vect X_s]
+  \big\}
+  \big\rangle_{0,t}^{i-1}  
+\end{align}
+The first, second and third line on the r.h.s. of \eqref{eqn:pert-approx-2}
+are the zeroth order, first order and second order approximations, respectively.
+
+\section{On the accuracy of the response fomula}
+
+An important issue is how far can we trust the response theory.
+To investigate this issue, we should firstly look at the
+most important step: the trajectory measure reweighting,
+i.e. Eqn.~\eqref{eqn:exact-ratio-1} and \eqref{eqn:exact-ratio-2}.
+These formula are strict, however, the statistical error
+may be increased do to the measure reweighting, according to
+similar measure reweighting methods widely used in equilibrium sampling.
+We find it more convenient to start from the nonequilibrium steady
+state (NESS), then we extend it to more general, but short-time
+auto-correlated dynamics.
+
+\subsection{Starting from NESS sampling}
+
+For the NESS system, it is possible to define a time-independent
+measure of the configuration space, namely $p_0(\vect x)$, where
+the subscript ``0'' denotes the reference state. We want to measure
+the propties in the perturbed state, ``1'', where the NESS
+probability distribtion is denoted by $p_1(\vect x)$. Assume we have
+a sampling sequence $\vect x_1, \vect x_2, \cdots, \vect x_M$, which
+are independent, and are subject to the reference probability distribtion
+$p_0(\vect x)$. Any  observation $O$ of the perturbed system can be
+estimated by
+\begin{align}
+  \bar O_M = \frac 1M\sum_{i=0}^{M-1} O(\vect x_i) R(\vect x_i)
+\end{align}
+where $R(\vect x)$ is the ration between the two probability distributions:
+\begin{align}
+  R(\vect x) = \frac{p_1(\vect x)}{p_0(\vect x)}
+\end{align}
+$\bar O_M$ is an unbiased estimator for $\langle O\rangle_1$
+($\langle \cdot\rangle_1$ denotes the expectation value in the perturbed
+system), because
+\begin{align}
+  \langle \bar O_M\rangle_0
+  = \frac 1M\sum_{i=0}^{M-1}
+  \int\dd \vect x\, p_0(\vect x_i)\,O(\vect x_i) R(\vect x_i)
+  = \int \dd \vect x\,p_0(\vect x)\,O(\vect x)
+  \,\frac{p_1(\vect x)}{p_0(\vect x)}
+  = \langle O\rangle_1
+\end{align}
+The statistical uncertainty is defined by the variance:
+\begin{align}
+  \textrm{Var}_0(\bar O_M)
+  = \big\langle
+  [\bar O_M^2 - \langle \bar O_M\rangle_0]^2
+  \big\rangle_0
+  = \langle\bar O_M^2\rangle_0 -
+  \langle\bar O_M\rangle_0^2 
+\end{align}
+It can be calculated that (only some basic algibra are needed):
+\begin{align}
+  \textrm{Var}_0(\bar O_M)
+  =
+  \langle O^2\cdot R\rangle_1 - \langle O\rangle_1^2
+\end{align}
+Comparing this variance with that of a direct sampling in the perturbed system,
+the difference would be
+\begin{align}
+  \Delta\textrm{Var}_0(\bar O_M)
+  =
+  \langle O^2\cdot (R-1)\,\rangle_1
+\end{align}
+This means that the reweighted sampling can be more precise if the
+observable $O$ is more significant in the region where $p_0 > p_1$, or
+less precise if $O$ is more significant in the region where $p_0 <
+p_1$.  This is obvious, and be observed in the equilibrium sampling
+methods (for example, the umbrella sampling).  In our case, the
+observable is nothing but the mollifier $K_{\sigma, \vect y}(\vect
+x)$, which is the smoothified version of the delta function,
+so we can somehow think
+$\langle K_{\sigma, \vect y}(\vect x)\rangle_0
+\approx p_0(\vect y)$ and
+$\langle K_{\sigma, \vect y}(\vect x)\rangle_1
+\approx p_1(\vect y)$.  This
+means the average of the mollifier is always calculated with a worse
+precision when $\langle K_{\sigma, \vect y}(\vect x)\rangle_1 >
+\langle K_{\sigma, \vect y}(\vect x)\rangle_0$. Please note that we
+are always more interested in the region where $\langle K_{\sigma,
+  \vect y}(\vect x)\rangle_1 > \langle K_{\sigma, \vect y}(\vect
+x)\rangle_0$, so the sad news is that we will always end up with
+a worse preciseion compared with the direct sampling of perturbed
+system.
+To have a resonable approximation, it is required that
+the unperturbed distribution $p_0(\vect x)$ should always be
+overlapping with the perturbed version in the region where $p_1(\vect
+x)$, or $\langle K_{\sigma, \vect y}(\vect x)\rangle_1$ is relavant.
+
+\subsection{The accuracy of the reweighted trajectory sampling}
+
+We want to understand what is the precision of the
+reweighted trajectory sampling, namely:
+\begin{align}\label{eqn:traj-ob}
+  \Delta \mo (t) =
+  \int \dd \vect v\, f(\vect v, 0)
+  \int_{\mc\{\vect v, 0; t\}}
+  \dd \mathcal P_{i-1}^\beta[\vect X_s]\,
+  O(\vect X_t)
+  \Bigg[
+  \frac{\dd \mathcal P^\beta_i[\vect X_s]}
+  {\dd \mathcal P^\beta_{i-1}[\vect X_s]}
+  -1
+  \Bigg]
+\end{align}
+The leading order of the trajectory order reweighting is
+\begin{align}\label{eqn:leading-rew}
+  \frac{\dd \mathcal P^\beta_i[\vect X_s]}
+  {\dd \mathcal P^\beta_{i-1}[\vect X_s]}
+  -1
+  =
+  \frac1\sigma\int_0^t M(s)\,
+  \vect D(\vect X_s)\cdot \dd\vect W_t
+\end{align}
+Notice the observable $O$ only depends on the end point
+of the trajectory. Therefore, we can use this properties to derive the
+estimate of the path integral.
+Let a series of descrete point along the trajectory:
+$(\vect X_{t_1}, t_1), (\vect X_{t_2}, t_2), \cdots, (\vect X_{t_N}, t_N)$,
+with $t_0 = 0,\ t_N = t$. For convenience, we denote
+$\vect x_k = \vect X_{t_k}$. The descretization of the path integral
+\eqref{eqn:traj-ob} is 
+\begin{align}
+  \Delta\mo (t) =
+  \int\dd \vect x_0\, f(\vect x_0, 0)
+  \int \dd \vect x_N
+  \int
+  \prod_{k=0}^{N-1}
+  \dd \vect x_k\,
+  p(\vect x_{k+1},t_{k+1} | \vect x_{k},t_{k})\,
+  O(\vect x_N)\,
+  \frac1\sigma\,
+  \Big[
+  \sum_{m=0}^{N-1}
+  M(t_m) \vect D(\vect x_m) (\vect W_{m+1} - \vect W_{m})
+  \Big]
+\end{align}
+By using the Chapman-Kolmogorov property, we have
+\begin{align}\nonumber
+  \Delta\mo (t) = &\,
+  \frac1\sigma
+  \sum_{m=0}^{N-1}
+  \int \dd \vect x_0\,
+  \dd \vect x_m\,
+  \dd \vect x_{m+1}\,
+  \dd \vect x_N\,
+  p(\vect x_N,t_N | \vect x_{m+1}, t_{m+1})\,
+  p(\vect x_{m+1}, t_{m+1} | \vect x_{m}, t_{m})\,
+  p(\vect x_{m}, t_{m} | \vect x_{0}, t_{0}) \,
+  f(\vect x_0, t_0)\\
+  &\,
+  M(t_m) \vect D(\vect x_m) (\vect W_{m+1} - \vect W_{m})  
+  O(\vect x_N)
+\end{align}
+Here for simplicity, we assume the dynamics is over-damped (the equations
+are the same for $O$ as under-damped dynamics).
+\begin{align}
+  \dd \vect X_t =
+  \vect b(\vect X_t, t)\,\dd t +
+  \sigma\, \dd \vect W_t
+\end{align}
+changing the variables gives
+\begin{align}
+  \vect W_{m+1} - \vect W_{m}
+  =
+  \vect x_{m+1} - \vect x_{m}
+  -
+  \vect b(\vect x_m, t_m)\,\Delta t_m
+\end{align}
+so
+\begin{align}\nonumber
+  \Delta\mo (t) = &\,
+  \frac1\sigma
+  \sum_{m=0}^{N-1}
+  \int \dd \vect x_0\,
+  \dd \vect x_m\,
+  \dd \vect x_{m+1}\,
+  \dd \vect x_N\,
+  p(\vect x_N,t_N | \vect x_{m+1}, t_{m+1})\,
+  p(\vect x_{m+1}, t_{m+1} | \vect x_{m}, t_{m})\,
+  p(\vect x_{m}, t_{m} | \vect x_{0}, t_{0})\,
+  f(\vect x_0, t_0)
+  \\ \nonumber
+  &\,
+  M(t_m) \vect D(\vect x_m) (\vect x_{m+1} - \vect x_{m})  
+  O(\vect x_N)\\\nonumber
+  -
+  &\,
+  \frac1\sigma
+  \sum_{m=0}^{N-1}
+  \int \dd \vect x_0\,
+  \dd \vect x_m\,
+  \dd \vect x_N\,
+  p(\vect x_N,t_N | \vect x_{m}, t_{m})\,
+  p(\vect x_{m}, t_{m} | \vect x_{0}, t_{0}) \,
+  f(\vect x_0, t_0)\\
+  &\,
+  M(t_m) \vect D(\vect x_m) \vect b(\vect x_m, t_m)\,\Delta t_m
+  O(\vect x_N)
+\end{align}
+% The last term equals to
+% \begin{align}
+%   \frac1\sigma
+%   \int \dd \tau
+%   M(\tau)
+%   \Big\langle
+%   O(\vect X_t) \vect b(\vect X_\tau,\tau) \vect D(\vect X_\tau)
+%   \Big\rangle
+% \end{align}
+By using the forward Kolmogorov equation, we have
+\begin{align}\nonumber
+  p(\vect x_{m+1},t_{m+1} | \vect x_{m},t_{m} )
+  = &\,
+  p(\vect x_{m+1},t_{m} | \vect x_{m},t_{m} )
+  -
+  \Delta t_m
+  \nabla_{\vect x_{m+1}}
+  \cdot
+  [ \vect b(\vect x_{m+1}, t_m)\,
+  p(\vect x_{m+1},t_{m} | \vect x_{m},t_{m} )\,]\\
+  &\,
+  +
+  \Delta t_m \frac{1} 2\,\sigma^2\,
+  \nabla_{\vect x_{m+1}}^2 p(\vect x_{m+1},t_{m} | \vect x_{m},t_{m} )
+\end{align}
+with initial condition
+\begin{align}
+  p(\vect x_{m+1},t_{m} | \vect x_{m},t_{m} ) =
+  \delta (\vect x_{m+1} - \vect x_{m})
+\end{align}
+so
+\begin{align}\nonumber
+  \int\dd \vect x_{m+1}\,
+  &
+  p(\vect x_N,t_N | \vect x_{m+1}, t_{m+1})\,
+  p(\vect x_{m+1}, t_{m+1} | \vect x_{m}, t_{m})\,
+  (\vect x_{m+1} -\vect x_{m})\\\nonumber
+  =&
+  -
+  \Delta t_m
+  \int \dd \vect x_{m+1}\,
+  p(\vect x_N,t_N | \vect x_{m+1}, t_{m+1})\,
+  (\vect x_{m+1} - \vect x_{m})
+  \nabla_{\vect x_{m+1}}
+  \cdot
+  [ \vect b(\vect x_{m+1}, t_m)\,
+  p(\vect x_{m+1},t_{m} | \vect x_{m},t_{m} )\,]\\\nonumber
+  &
+  +
+  \Delta t_m
+  \int \dd \vect x_{m+1}\,
+  p(\vect x_N,t_N | \vect x_{m+1}, t_{m+1})\,
+  (\vect x_{m+1} - \vect x_{m})
+  \frac{\sigma^2} 2\,
+  \nabla_{\vect x_{m+1}}^2 p(\vect x_{m+1},t_{m} | \vect x_{m},t_{m} ) \\\nonumber
+  =&
+  \Delta t_m
+  \int \dd \vect x_{m+1}\,
+  \vect b(\vect x_{m+1}, t_m)\,
+  \delta (\vect x_{m+1} - \vect x_{m})
+  p(\vect x_N,t_N | \vect x_{m+1}, t_{m+1}) \\\nonumber
+  &+
+  \Delta t_m
+  \int \dd \vect x_{m+1}\,
+  \delta (\vect x_{m+1} - \vect x_{m})
+  \nabla_{\vect x_{m+1}}p(\vect x_N,t_N | \vect x_{m+1}, t_{m+1})\times
+  \frac{\sigma^2} 2 \\
+  =&
+  \Delta t_m
+  \vect b(\vect x_{m}, t_m)\,
+  p(\vect x_N,t_N | \vect x_{m}, t_{m+1}) 
+  +
+  \,\frac{\sigma^2} 2\,\Delta t_m
+  \nabla_{\vect x_{m}}p(\vect x_N,t_N | \vect x_{m}, t_{m+1})
+\end{align}
+By using the backward Kolmogorov equation
+\begin{align}
+  p(\vect x_N,t_N | \vect x_{m}, t_{m+1})
+  = 
+  p(\vect x_N,t_N | \vect x_{m}, t_{m})
+  + \mathcal O(\Delta t_m)
+\end{align}
+So by keeping the leading order of $\Delta t_m$
+\begin{align}\nonumber
+  \int\dd \vect x_{m+1}\,
+  &
+  p(\vect x_N,t_N | \vect x_{m+1}, t_{m+1})\,
+  p(\vect x_{m+1}, t_{m+1} | \vect x_{m}, t_{m})\,
+  (\vect x_{m+1} -\vect x_{m})\\\nonumber
+  =&
+  \Delta t_m
+  \vect b(\vect x_{m}, t_m)\,
+  p(\vect x_N,t_N | \vect x_{m}, t_{m}) 
+  +
+  \,\frac{\sigma^2} 2\,\Delta t_m
+  \nabla_{\vect x_{m}}p(\vect x_N,t_N | \vect x_{m}, t_{m})
+\end{align}
+Therefore, we have
+\begin{align}\label{eqn:error-o-corr}
+  \Delta\mo (t) =
+  \frac\sigma 2
+  % \int \dd\vect x_0 f(\vect x_0,0)
+  \int_0^t \dd\tau M(\tau)
+  \Bigg[
+  \int \dd\vect x_t\, O(\vect x_t)
+  \int \dd\vect x_\tau\,
+  \nabla_{\vect x_{\tau}}p(\vect x_t,t | \vect x_{\tau}, \tau)
+  \cdot
+  \vect D(\vect x_\tau)\,
+  % p(\vect x_\tau,\tau | \vect x_0, 0)
+  f(\vect x_\tau,\tau)
+  \bigg]
+\end{align}
+\redc{
+The term in the bracket is the time correlation between
+the observable and the backward flux of the error force $\vect D$.
+}
+
+\vskip .5cm
+Discussions on Eqn.~\eqref{eqn:error-o-corr}:
+\begin{enumerate}
+\item This equation relate the error force with the error of observable,
+  which would be useful in the error studies.
+\item When the system does not have long-time memory:
+  \begin{align}
+    p(\vect x_t,t | \vect x_{\tau}, \tau) = p(\vect x_t,t ), \qquad
+    \textrm{for}\ t  - \tau > t_c
+  \end{align}
+  Here $t_c$ denotes the length of memory of the system. Then
+  Eqn.~\eqref{eqn:error-o-corr} becomes 
+  \begin{align}
+    \Delta\mo (t) =
+    \frac\sigma 2
+    % \int \dd\vect x_0 f(\vect x_0,0)
+    \int_{t - t_c}^t \dd\tau M(\tau)
+    \Bigg[
+    \int \dd\vect x_t\, O(\vect x_t)
+    \int \dd\vect x_\tau\,
+    \nabla_{\vect x_{\tau}}p(\vect x_t,t | \vect x_{\tau}, \tau)
+    \cdot
+    \vect D(\vect x_\tau)\,
+    % p(\vect x_\tau,\tau | \vect x_0, 0)
+    f(\vect x_\tau,\tau)
+    \bigg]
+  \end{align}
+  This means we can integrate \eqref{eqn:leading-rew} only from
+  $t - t_c$ to $t$.
+\item In our case of core set identification,
+  $O(\vect x) = K_{\sigma, \vect y}(\vect x)$, which is nearly a delta function.
+  A good precision of the response theory means 
+  that the trajectory generated by the reference system, should
+  properly sample the correlation between the
+  error force flux at time $\tau$, and the probability
+  density at time $t$.  In other words, we generate trajectories
+  samples the conditional probability $p(\vect x_t,t | \vect x_{\tau}, \tau)$,
+  and a good precision means
+  the backward flux $\nabla_{\vect x_{\tau}}p(\vect x_t,t | \vect x_{\tau}, \tau)
+    \cdot
+    \vect D(\vect x_\tau)$ should be significant on this sampling.
+\end{enumerate}
+
+\section{Numerical tests}
+
+\subsection{The tilting double-well potential}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.4\textwidth]{figs/fig-tilt-pot.eps}
+  \caption{The double-well potential with tilting perturbation.}
+  \label{fig:tmp1}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.95\textwidth]
+  {figs/fig-tilt-str2-2d-more.eps}
+  \caption{The plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
+    $F_e^{\textrm{max}} = 2$.
+    The 1st row: the direct non-equilibrium
+    simulation. The 2nd row: the results of linear response
+    formula~\eqref{eqn:core-identify-approx-2}.
+    The 3rd row: the results of response
+    formula~\eqref{eqn:pert-approx-1} and \eqref{eqn:pert-approx-2}.
+    The order is 2, the reference state is $F_e^{\textrm{max}} = 0$, i.e.
+    the equilibrium state.
+  }
+  \label{fig:tmp2}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.95\textwidth]
+  {figs/fig-tilt-str4-2d-more.eps}
+  \caption{The plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
+    $F_e^{\textrm{max}} = 4$.
+    The 1st row: the direct non-equilibrium
+    simulation. The 2nd row: the results of linear response
+    formula~\eqref{eqn:core-identify-approx-2}.
+    The 3rd row: the results of response
+    formula~\eqref{eqn:pert-approx-1} and \eqref{eqn:pert-approx-2}.
+    The order is 2, the reference state is $F_e^{\textrm{max}} = 0$, i.e.
+    the equilibrium state.
+    The 4th row: the results of response
+    formula~\eqref{eqn:pert-approx-1} and \eqref{eqn:pert-approx-2}.
+    The order is 1, the reference state is $F_e^{\textrm{max}} = 2$.
+  }
+  \label{fig:tmp3}
+\end{figure}
+
+We test the idea firstly by a one-dimensional model system: one particle in a
+tilting double-well potential. For convenience, we let the mass of the
+particle to be 1 \textsf{amu}. The unperturbed
+Hamiltonian of the system is given by:
+\begin{align}
+  \mh (\vect p, \vect q) = \frac 12 \vect p^2 + U(\vect q) 
+\end{align}
+with potential
+\begin{align}
+  U(\vect q) = \frac12 k (\vect q^2 - a^2)^2
+\end{align}
+Here $k = 8$~$\textsf{kJ} / (\textsf{mol nm}^4)$, and $ a = 1\ \textsf{nm}$.
+Notice at room temperature $300\ \textsf{K}$, $k_BT = 2.48$~\textsf{kJ/mol}.
+See the red line in Fig.~\ref{fig:tmp1}.
+%$\textsf{kJ} / (\textsf{mol nm}^4)}$
+The perturbation is given by
+\begin{align}
+  \vect C(\vect p, \vect q) = 0; \qquad
+  \vect D(\vect p, \vect q) = -\nabla_{\vect q} V(\vect q) = 1
+\end{align}
+Here $V(\vect q) = -\vect q$ is  effectively tilt the original
+potential $U(\vect q)$. The strength of the perturbation $F_e(t)$ is given
+by the following function:
+\begin{align}
+  F_e(t) = 
+  \begin{cases}
+    F_e^{\textrm{max}}\times (t / t_c) & t < t_c \\
+    F_e^{\textrm{max}} & t \geq t_c
+  \end{cases}
+\end{align}
+$t_c = 20$~\textsf{ps} so that the perturbation increases slow
+enough: the typical decaying time scale of the correlation function
+in~\eqref{eqn:core-identify-approx-2} is roughly 3~\textsf{ps}.
+We choose $\tau = 1$~\textsf{ps}.
+% We choose $\mt_\alpha$
+% to be the propagator of the Langevin dynamics at temperature $150$
+% \textsf{K} with $\alpha = 1$ \textsf{ps}.
+Since the dynamics is in
+1-d, the projection $\mathcal P$ is identity.  We consider the stability of the
+distribution $f(\vect x, t)$ at time $t$: $\Delta\mt_\tau f(\vect x, t)$.
+The larger
+this value, the more stable the core sets are.
+
+See Fig. \ref{fig:tmp2} and \ref{fig:tmp3} for numerical result.  Both
+the direct non-equilibrium simulation and the response
+approximations are presented. Time slices $t = 0$, 10, 15 and 20
+\textsf{ps} are shown.  When the perturbation is small,
+i.e. $F_e^{\textrm{max}} = 2$, the direct non-equilibrium simulation
+and the linear response results are accurately consistent.
+The 2nd order response formula~\eqref{eqn:pert-approx-1} and
+\eqref{eqn:pert-approx-2} are even better, but the numerical error
+is bigger.
+As the
+perturbation grows stronger, the left well becomes less stable, while
+the right well becomes more stable.  When the perturbation is big,
+i.e. $F_e^{\textrm{max}} = 4$, the linear response result is no
+longer precise.
+When $t\geq 20$ \textsf{ps}, the left well basically
+disappears, however, the linear response calculation presents
+artifical occilation at the right well.
+The 2nd order response also has artifical effect, and the numerical
+uncertainty is high.
+The 4th rwo of Fig.~\ref{fig:tmp3} presents the 1st order result of the
+response formula~\eqref{eqn:pert-approx-1} and
+\eqref{eqn:pert-approx-2}, with
+respect to the reference state $F_e^{\textrm{max}} = 2$. The result
+is impressively improved compared with the linear response with respect to the
+equilibrium state.
+
+\subsection{Splitting single-well potential}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.4\textwidth]{figs/fig-split-pot.eps}
+  \caption{The single-well potential with splitting perturbation.}
+  \label{fig:tmp4}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.95\textwidth]
+  {figs/fig-split-str1-2d-more.eps}
+  \caption{The plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
+    $F_e^{\textrm{max}} = 1.0$.
+    The 1st row: the direct non-equilibrium
+    simulation. The 2nd row: the results of linear response
+    formula~\eqref{eqn:core-identify-approx-2}.
+    The 3rd row: the results of response
+    formula~\eqref{eqn:pert-approx-1} and \eqref{eqn:pert-approx-2}.
+    The order is 2, the reference state is $F_e^{\textrm{max}} = 0$, i.e.
+    the equilibrium state.
+  }
+  \label{fig:tmp5}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.95\textwidth]
+  {figs/fig-split-str2-2d-more.eps}
+  \caption{The plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
+    $F_e^{\textrm{max}} = 2.0$.
+    The 1st row: the direct non-equilibrium
+    simulation. The 2nd row: the results of linear response
+    formula~\eqref{eqn:core-identify-approx-2}.
+    The 3rd row: the results of response
+    formula~\eqref{eqn:pert-approx-1} and \eqref{eqn:pert-approx-2}.
+    The order is 2, the reference state is $F_e^{\textrm{max}} = 0.0$, i.e.
+    the equilibrium state.
+    The 4th row: the results of response
+    formula~\eqref{eqn:pert-approx-1} and \eqref{eqn:pert-approx-2}.
+    The order is 1, the reference state is $F_e^{\textrm{max}} = 1.0$.
+  }
+  \label{fig:tmp6}
+\end{figure}
+
+In this splitting single-well potential example, the unperturbed
+potential is simply the harmonic potential given by:
+\begin{align}
+  U(\vect q) = \frac12\,k\,\vect q^2 
+\end{align}
+Here $k = 8$~$\textsf{kJ} / (\textsf{mol nm}^2)$.
+Notice at room temperature $300\ \textsf{K}$, $k_BT = 2.48$~\textsf{kJ/mol}.
+See the red line in Fig.~\ref{fig:tmp4}.
+%$\textsf{kJ} / (\textsf{mol nm}^4)}$
+The perturbation is again given by
+\begin{align}
+  \vect C(\vect p, \vect q) = 0; \qquad
+  \vect D(\vect p, \vect q) = -\nabla_{\vect q} V(\vect q) 
+\end{align}
+where $V(\vect q)$ is a Gaussian type potential:
+\begin{align}
+  V(\vect q) = \frac{1}{\sqrt{2\pi \sigma^2}}
+  \exp\Big\{-\frac{\vect q^2}{2\sigma^2}\Big\}
+\end{align}
+we use $\sigma = 0.16$~\textsf{nm}. Please see Fig~\ref{fig:tmp4}.
+For the results, please see Fig.~\ref{fig:tmp5} -- Fig.~\ref{fig:tmp6}.
+
+\section{Non-equilibrium linear response}
+
+% \subsection{Path integral}
+% The path integral is very useful to develop the non-equilibrium version of
+% linear response theory. 
+% Comparing with the idea of ensemble, which is a collection of state points
+% in the configuration space, the path integral consider the ``ensemble'' of
+% trajectories.
+% (In the following, the develop of formula are formal. The strict mathematical
+% theory is still unclear for me.)
+% First, we take the standard 1-dimensional Wiener process for example.
+% Consider the time interval $[0,T]$, the probability density of one trajectory
+% is given by
+% \begin{align}
+%   p[W_t] = \frac 1Z e^{-I[W_t]}
+% \end{align}
+% where $I$ is the action functional of Wiener process given by
+% \begin{align}
+%   I[W_t] = \int_0^T\frac12 \dot{W}_t^2\dd t
+% \end{align}
+% the partition function $Z$ is
+% \begin{align}
+%   Z = \int  e^{-I[W_t]} \prod_{0\leq t\leq T}\dd W_t
+% \end{align}
+% Consider the following two stochastic processes:
+% \begin{align}\label{eqn:st-orig}
+%   \dd \vect X_t &= \vect b(\vect X_t, t) \dd t + \bsigma(\vect X_t, t)\,\dd\vect W_t \\\label{eqn:st-pert}
+%   \dd \vect Y_t &= \vect b(\vect Y_t, t) \dd t + \vect c(\vect Y_t, t) \dd t + \bsigma(\vect Y_t, t)\,\dd\vect W_t
+% \end{align}
+% Where $\vect W_t$ is the standard Wiener process. Obviously
+% \eqref{eqn:st-pert} is the perturbed version of
+% \eqref{eqn:st-orig}. It has been shown that, starting from the same
+% measure saying $\mu$ (which is not necessarily equilibrium or
+% stationary), the ratio of trajectory probability density between the perturbed
+% and unperturbed process is~\cite{li2010lecture} (\recheck{I think here
+%   the incompressibility of phase space should be assumed. Otherwise the
+%   ratio should be more complicated.})
+% \begin{align}\label{eqn:girsanov}
+%   \frac{\dd\mathcal P_{\vect Y}}{\dd\mathcal P_{\vect X}}
+%   =
+%   \exp
+%   \Big\{
+%   -\frac12 \int_0^T\vert\bsigma^{-1}\cdot\vect c(\vect X_t,t)\vert^2\,\dd t
+%   +\int_0^T[\dot{\vect X}_t - \vect b(\vect X_t,t)]^T
+%   \cdot\bsigma^{-T}\cdot\bsigma^{-1}\cdot\vect c(\vect X_t,t)\,\dd t 
+%   \Big\}
+% \end{align}
+% where $\mathcal P_{\vect Y}$ and $\mathcal P_{\vect Y}$ are measures
+% of the trajectories generated by \eqref{eqn:st-pert} and
+% \eqref{eqn:st-orig}, respectively.  This result (Girsanov formula?) is
+% (formally) precise without any approximation.
+
+\subsection{Non-equilibrium perturbation with linear approximation}
+
+To be more general, we study the two perturbed Langevin dynamics
+(i.e. Eqn.~\eqref{eqn:langevin-1} and \eqref{eqn:langevin-2}) with
+different perturbations that are given by $F_e^1(t)$ and $F_e^2(t)$,
+respectively. We consider the difference between the observations of
+these dynamics:
+\begin{align}\label{eqn:noneq-pert-1}
+  \langle Q(t)\rangle_{\mu}^{2}  -\langle Q(t)\rangle_{\mu}^{1}
+  &=
+  \int Q(\vect X_t) \,\dd\mathcal P^2[\vect X_t]
+  -\int Q(\vect X_t) \,\dd\mathcal P^1[\vect X_t]
+  =
+  \int Q(\vect X_t)
+  \Big\{
+  \frac{\dd\mathcal P^2[\vect X_t]}{\dd\mathcal P^1[\vect X_t]}
+  -1
+  \Big\}
+  \,\dd\mathcal P^1[\vect X_t]
+\end{align}
+\recheck{Here we further assume that $\vect C(\vect q, \vect p) = 0$.
+  (otherwise the Girsanov formula~\eqref{eqn:girsanov} may be unstable...
+  By letting the $\epsilon$ in the following formula to 0, if
+  $\vect C\neq 0$, the limit may not be exist. This
+  is intuitionally not likely... I do not know why.)}
+By writing Eqn.~\eqref{eqn:langevin-1} and \eqref{eqn:langevin-2} as
+\begin{align}
+  \frac{\dd }{\dd t}
+  \left[
+    \begin{array}{l}
+      \vect q\\
+      \vect p
+    \end{array}
+  \right]
+  =
+  \left[
+    \begin{array}{l}
+      \nabla_{\vect p}\mh\\
+      -\nabla_{\vect q}\mh - \gamma\vect p + F_e\vect D
+    \end{array}
+  \right]
+  \dd t
+  +
+  \left[
+    \begin{array}{cc}
+      \epsilon \vect I & 0\\
+      0 & \sigma\vect I
+    \end{array}
+  \right]
+  \cdot
+  \left[
+    \begin{array}{l}
+      \dd\vect W_t\\
+      \dd\vect W_t
+    \end{array}
+  \right]
+\end{align}
+Using Girsanov's formula, to the leading order of $\Delta F_e = F_e^2 - F_e^1$,
+we have
+\begin{align}\nonumber
+  \frac{\dd\mathcal P^2[\vect X_t]}{\dd\mathcal P^1[\vect X_t]}
+  -1
+  &=
+  \exp
+  \Big\{
+  \frac 1 {\sigma^2}
+  \int_0^t
+  [\dot{\vect p}_s +
+  \gamma\vect p_s +
+  \nabla_{\vect q}\mh(\vect q_s,\vect p_s)
+  - F_e^1(s) \vect D(\vect q_s,\vect p_s)
+  ]\cdot
+  \vect D(\vect q_s,\vect p_s)
+  \Delta F_e(s)\,
+  \dd s
+  + \mathcal O(\Delta F_e^2)
+  \Big\} - 1 \\\label{eqn:girsanov-langevin-1}
+  &\approx
+  \frac 1 {\sigma^2}
+  \int_0^t
+  [\dot{\vect p}_s +
+  \gamma\vect p_s +
+  \nabla_{\vect q}\mh(\vect q_s,\vect p_s)
+  - F_e^1(s) \vect D(\vect q_s,\vect p_s)
+  ]\cdot
+  \vect D(\vect q_s,\vect p_s)
+  \Delta F_e(s)\,
+  \dd s  
+\end{align}
+This is where the \emph{linear approximation} applies.
+By using the relation $\beta = 2\gamma / \sigma^2$ and split the
+Eqn.~\eqref{eqn:girsanov-langevin-1} into time symmetric and
+antisymmetric parts~\cite{baiesi2010nonequilibrium2}, we have:
+\begin{align}\nonumber
+  \frac{\dd\mathcal P^2[\vect X_t]}{\dd\mathcal P^1[\vect X_t]}
+  -1
+  \approx& \:
+  \frac\beta 2
+  \int_0^t
+  \Delta F_e(s)\,
+  \vect p_s\cdot \vect D(\vect q_s,\vect p_s)\,
+  \dd s\\  \label{eqn:girsanov-langevin-2}
+  &\,-
+  \frac 1{\sigma^2}
+  \int_0^t
+  \Delta F_e(s)\,
+  \vect F^1(\vect q_s,\vect p_s)\cdot \vect D(\vect q_s,\vect p_s)\,
+  \dd s  
+  +
+  \frac 1{\sigma^2}
+  \int_0^t
+  \Delta F_e(s) \vect D(\vect q_s,\vect p_s)\cdot \dd \vect p_s,
+\end{align}
+where $\vect F^1(\vect q_s,\vect p_s) = -\nabla_{\vect q}\mh(\vect
+q_s,\vect p_s) + F_e^1(t) \vect D(\vect q_s,\vect p_s)$.  The first
+line of the r.h.s. of Eqn.~\eqref{eqn:girsanov-langevin-2} is the time
+integrated excess entropy flux, which is time-antisymmetric, while the
+second line is the excess dynamic activity that is time-symmetric.
+Inserting Eqn.~\eqref{eqn:girsanov-langevin-2} into
+\eqref{eqn:noneq-pert-1}, we have
+\begin{align}\nonumber
+  \langle O(t)\rangle_{\mu}^{2}  -\langle O(t)\rangle_{\mu}^{1}
+  =&\,
+  \frac\beta 2
+  \int_0^t\dd s\,
+  \Delta F_e(s)\,
+  \big\langle
+  \vect p_s\cdot\vect D(\vect X_s)\,O(\vect X_t)
+  \big\rangle_\mu^1 \\  \label{eqn:noneq-pert-2}
+  &
+  -
+  \frac1{\sigma^2}
+  \int_0^t\dd s\,
+  \Delta F_e(s)\,
+  \big\langle
+  \vect F^1(\vect X_s)\cdot\vect D(\vect X_s)\,O(\vect X_t)
+  \big\rangle_\mu^1
+  +
+  \frac1{\sigma^2}
+  \int_0^t
+  \Delta F_e(s)\,
+  \big\langle
+  \dd \vect p_s\cdot\,
+  \vect D(\vect X_s)\,O(\vect X_t)
+  \big\rangle_\mu^1  
+\end{align}
+If $\vect D(\vect x)$ is independent of momentum and is conservative,
+saying $\vect D(\vect x) = \vect D(\vect q) = -\nabla_{\vect q}
+V(\vect q)$, then the last term of \eqref{eqn:noneq-pert-2} can be
+further simplified, because
+\begin{align}
+  -\dd\, [\nabla_{\vect q}V(\vect q_s)\cdot \vect p_s]
+  =
+  -\nabla_{\vect q}V(\vect q_s)\cdot\dd \vect p_s
+  - \vect p_s\cdot\nabla^2_{\vect q}V(\vect q_s)\cdot\vect p_s\,\dd s,
+\end{align}
+then \eqref{eqn:noneq-pert-2} changes to
+\begin{align}\nonumber
+  \langle O(t)\rangle_{\mu}^{2}  -\langle O(t)\rangle_{\mu}^{1}
+  =&\,
+  -\frac\beta 2
+  \int_0^t\dd s\,
+  \Delta F_e(s)\,
+  \big\langle
+  \vect p_s\cdot\nabla_{\vect q} V(\vect X_s)\,O(\vect X_t)
+  \big\rangle_\mu^1 \\\nonumber
+  & +
+  \frac1{\sigma^2}
+  \int_0^t\dd s\,
+  \Delta F_e(s)\,
+  \big\langle
+  \vect F^1(\vect X_s)
+  \cdot\nabla_{\vect q} V(\vect X_s)
+  \,O(\vect X_t)
+  \big\rangle_\mu^1 \\\nonumber
+  & -
+  \frac1{\sigma^2}
+  \int_0^t
+  \dd s\,
+  \Delta F_e(s)\,
+  \frac{\dd}{\dd s}
+  \big\langle
+  \vect p_s\cdot\,
+  \nabla_{\vect q} V(\vect X_s)
+  \,O(\vect X_t)
+  \big\rangle_\mu^1    \\\label{eqn:noneq-pert-3}
+  & +
+  \frac1{\sigma^2}
+  \int_0^t
+  \dd s\,
+  \Delta F_e(s)\,
+  \big\langle
+  \vect p_s\vect p_s:
+  \nabla^2_{\vect q} V(\vect X_s)
+  \,O(\vect X_t)
+  \big\rangle_\mu^1      
+\end{align}
+
+\subsection{Remarks and possibilities}
+\bluec{
+\begin{itemize}
+\item It can be shown that 
+  Eqn.~\eqref{eqn:noneq-pert-3} simply reduces to
+  \eqref{eqn:eqi-pert-1}
+  in the equilibrium case~\cite{baiesi2010nonequilibrium2}.
+\item  Eqn.~\eqref{eqn:noneq-pert-3} shows that the trajectory average of
+  $F_e^2$ is calculated from the trajectory average of $F_e^1$. We
+  assume: (1) they start from the same initial distribution $\mu$, and
+  (2) $F_e^2$ is not very far from $F_e^1$ so that the linear
+  approximation is reasonably valid.  The condition (1) can be made
+  weaker, for example $F_e^2$ starts from $\mu_2$ and $F_e^1$ starts
+  from $\mu_1$, however, $\mu_2$ should be similar to $\mu_1$,
+  otherwise the sampling is of poor accuracy.
+  (2) means all possible trajectories of perturbation $F_e^2$ should be
+  effciently sampled by perturbation $F_e^1$.
+\item Eqn.~\eqref{eqn:noneq-pert-3}, may be used in this way:
+  assume we want to study the dynamics of $F_e^2$, which is very slow
+  and difficult to handle directly.
+  We can instead study the dynamics of $F_e^1$, which is
+  similar to $F_e^2$, but is much faster, so that it can be handle by MD.
+  And then use \eqref{eqn:noneq-pert-3} to calculate the averages of
+  the target dynamics $F_e^2$.
+  A possible choice of  $F_e^1$ may be  $F_e^1 = \alpha F_e^2,\ \alpha > 1$.
+  On the other hand, perturbation $F_e^2$ should be not very far
+  from $F_e^1$, that means $F_e^2$ can not arbitrarily accelerate
+  $F_e^1$. Fortunately we have the freedom to design $F_e^2$, therefore,
+  an error estimate may help, and a path optimization of $F_e^2$
+  may be developed.
+\item The perturbation $F_e^1$ can be kept a constant so that the
+  system reaches a stationary state.
+  $F_e^2$ is perturbation from that stationary state.
+  The averages in
+  \eqref{eqn:noneq-pert-3} can be calculated easier, because
+  what we sample is the stationary state, rather the trajectories.
+  When the system is far from the stationary state, the
+  accuracy is low. 
+\item It may also be possible to study multiple reference dynamics,
+  and use interpolations to increase the accuracy of the linear
+  approximation.
+\item Extension to nonlinear response?
+\end{itemize}
+}
 

### 2013-02-20 writedown/linear.response/main.tex
also select Backbone

@@
 
+\newpage
+\section{Numerical results}
+\subsection{Tilting double-well potential}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.4\textwidth]{figs/fig-tilt-pot.eps}
+  \includegraphics[width=0.4\textwidth]{figs/fig-split-pot.eps}
+  \caption{Two testing cases presented by this paper.
+    Left: the double-well potential with tilting perturbation.
+    Right: the single-well potential with splitting perturbation.}
+  \label{fig:new-tmp1}
+\end{figure}
+
+We test the idea by a one-dimensional model system: one particle in a
+tilting double-well potential. For convenience, we let the mass of the
+particle to be 1 \textsf{amu}. The unperturbed
+Hamiltonian of the system is given by:
+\begin{align}
+  \mh (\vect p, \vect q) = \frac 12 \vect p^2 + U(\vect q) 
+\end{align}
+with potential
+\begin{align}
+  U(\vect q) = \frac12 k (\vect q^2 - a^2)^2
+\end{align}
+Here $k = 8$~$\textsf{kJ} / (\textsf{mol nm}^4)$, and $ a = 1\ \textsf{nm}$.
+Notice at room temperature $300\ \textsf{K}$, $k_BT = 2.48$~\textsf{kJ/mol}.
+See the red line in Fig.~\ref{fig:tmp1}.
+%$\textsf{kJ} / (\textsf{mol nm}^4)}$
+The perturbation is given by
+\begin{align}
+  \vect C(\vect p, \vect q) = 0; \qquad
+  \vect D(\vect p, \vect q) = -\nabla_{\vect q} V(\vect q) = 1
+\end{align}
+Here $V(\vect q) = -\vect q$ is  effectively tilting the original
+potential $U(\vect q)$. The strength of the perturbation $F_e(t)$ is given
+by the following function:
+\begin{align}
+  F_e(t) = 
+  \begin{cases}
+    F_e^{\textrm{max}}\times (t / t_c) & t < t_c \\
+    F_e^{\textrm{max}} & t \geq t_c
+  \end{cases}
+\end{align}
+$t_c = 20$~\textsf{ps} so that the perturbation increases slow
+enough: the typical decaying time scale of the correlation function
+in~\eqref{eqn:core-identify-approx-2} is roughly 3~\textsf{ps}.
+We choose $\tau = 1$~\textsf{ps}.
+% We choose $\mt_\alpha$
+% to be the propagator of the Langevin dynamics at temperature $150$
+% \textsf{K} with $\alpha = 1$ \textsf{ps}.
+Since the dynamics is in
+1-d, the projection $\mathcal P$ is identity.  We consider the stability of the
+distribution $f(\vect x, t)$ at time $t$: $\Delta\mt_\tau f(\vect x, t)$.
+The larger
+this value, the more stable the core sets are.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.95\textwidth]
+  {figs/fig-tilt-str2-simple.eps}
+  \caption{
+    Tilting double well potential testing case:
+    the plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
+    $F_e^{\textrm{max}} = 2.0$.
+    From left to right:
+    (a) $t=0$~\textsf{ps},
+    the brutal force non-equilibirum simulation,
+    (b) $t=20$~\textsf{ps},
+    the brutal force non-equilibirum simulation,
+    (c) $t=20$~\textsf{ps},
+    response formula truncated at order $n=1$,
+    and
+    (d) $t=20$~\textsf{ps},
+    response formula truncated at order $n=2$.
+    The equilibirum state serves as the reference of
+    the response simulations.
+  }
+  \label{fig:new-tmp2}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.95\textwidth]
+  {figs/fig-tilt-str4-simple.eps}
+  \caption{
+    Tilting double well potential:
+    the plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
+    $F_e^{\textrm{max}} = 4.0$.
+    From left to right:
+    (a) $t=0$~\textsf{ps},
+    the brutal force non-equilibirum simulation,
+    (b) $t=20$~\textsf{ps},
+    the brutal force non-equilibirum simulation,
+    (c) $t=20$~\textsf{ps},
+    response formula with reference process $F_e^{\max} = 0.0$
+    (i.e. equilibirum simulation),
+    and
+    (d) $t=20$~\textsf{ps},
+    response formula with reference process $F_e^{\max} = 2.0$.
+    The response formula are truncated at order $n=1$.
+  }
+  \label{fig:new-tmp3}
+\end{figure}
+
+The Fig.~\ref{fig:new-tmp2} and \ref{fig:new-tmp3} present
+the application of the response theory to
+calculate core sets of the tilting double-well potential.
+Fig.~\ref{fig:new-tmp2} presents
+the snapshot at $t=20$~\textsf{ps}
+of $\Delta\mt_\tau f(\vect x,t)$ under the maximum perturbation
+strength $F_e^{\max} = 2.0$. Response formula truncated
+at both order $n=1$ and $n=2$ are compared to 
+the brutal force non-equilibirum simulation.
+No obviously difference is shown. The second order response result
+is slightly more precise than the first order one, but the
+statistical error is also larger, which makes the profile noisy.
+This figure tells us that
+(1) the modeling error of the linear response (1st order) is
+negligibly small; (2) the statistical error of numerically measuring the
+linear response is also small.
+Fig.~\ref{fig:new-tmp3} presents the $t=20$~\textsf{ps}
+snapshots of $\Delta\mt_\tau f(\vect x,t)$
+under a larger perturbations $F_e^{\max} = 4.0$.
+The linear reponse using the equilibirum process as the reference
+shows poor accuracy: the left well vanishes in this case,
+while the linear response produces  artifical structures.
+Since the profile is smooth, which indicates
+the statistical error of calculating the linear term is small,
+the deviation comes from the modellng error, namely the contribution
+of higher order response terms.
+By including the second order response term reduces the left
+artifical core set, but also creates artifical finer grained
+structures. The statistical accuracy is also worse than the linear response.
+Plot (d) present the result of the linear response
+formula using reference process $F_e^{\max} = 2.0$, which
+is the process presented in Fig.~\ref{fig:new-tmp2}.
+Much better modelling accuracy is achieved:
+the left artifical core set disappear and the
+statistical accuracy is also satisfactory.
+
+% See Fig. \ref{fig:tmp2} and \ref{fig:tmp3} for numerical result.  Both
+% the direct non-equilibrium simulation and the response
+% approximations are presented. Time slices $t = 0$, 10, 15 and 20
+% \textsf{ps} are shown.  When the perturbation is small,
+% i.e. $F_e^{\textrm{max}} = 2$, the direct non-equilibrium simulation
+% and the linear response results are accurately consistent.
+% The 2nd order response formula~\eqref{eqn:pert-approx-1} and
+% \eqref{eqn:pert-approx-2} are even better, but the numerical error
+% is bigger.
+% As the
+% perturbation grows stronger, the left well becomes less stable, while
+% the right well becomes more stable.  When the perturbation is big,
+% i.e. $F_e^{\textrm{max}} = 4$, the linear response result is no
+% longer precise.
+% When $t\geq 20$ \textsf{ps}, the left well basically
+% disappears, however, the linear response calculation presents
+% artifical occilation at the right well.
+% The 2nd order response also has artifical effect, and the numerical
+% uncertainty is high.
+% The 4th rwo of Fig.~\ref{fig:tmp3} presents the 1st order result of the
+% response formula~\eqref{eqn:pert-approx-1} and
+% \eqref{eqn:pert-approx-2}, with
+% respect to the reference state $F_e^{\textrm{max}} = 2$. The result
+% is impressively improved compared with the linear response with respect to the
+% equilibrium state.
+
+\subsection{Splitting single-well potential}
+
+In this splitting single-well potential example, the unperturbed
+potential is simply the harmonic potential given by:
+\begin{align}
+  U(\vect q) = \frac12\,k\,\vect q^2 
+\end{align}
+Here $k = 8$~$\textsf{kJ} / (\textsf{mol nm}^2)$.
+Notice at room temperature $300\ \textsf{K}$, $k_BT = 2.48$~\textsf{kJ/mol}.
+See the red line in Fig.~\ref{fig:tmp4}.
+%$\textsf{kJ} / (\textsf{mol nm}^4)}$
+The perturbation is again given by
+\begin{align}
+  \vect C(\vect p, \vect q) = 0; \qquad
+  \vect D(\vect p, \vect q) = -\nabla_{\vect q} V(\vect q) 
+\end{align}
+where $V(\vect q)$ is a Gaussian type potential:
+\begin{align}
+  V(\vect q) = \frac{1}{\sqrt{2\pi \sigma^2}}
+  \exp\Big\{-\frac{\vect q^2}{2\sigma^2}\Big\}
+\end{align}
+we use $\sigma = 0.16$~\textsf{nm}. Please see Fig~\ref{fig:tmp4}.
+For the results, please see Fig.~\ref{fig:tmp5} -- Fig.~\ref{fig:tmp6}.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.95\textwidth]
+  {figs/fig-split-str1-simple.eps}
+  \caption{
+    Splitting single-well potential testing case:
+    the plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
+    $F_e^{\textrm{max}} = 1.0$.
+    From left to right:
+    (a) $t=0$~\textsf{ps},
+    the brutal force non-equilibirum simulation,
+    (b) $t=20$~\textsf{ps},
+    the brutal force non-equilibirum simulation,
+    (c) $t=20$~\textsf{ps},
+    response formula truncated at order $n=1$,
+    and
+    (d) $t=20$~\textsf{ps},
+    response formula truncated at order $n=2$.
+    The equilibirum state serves as the reference of
+    the response simulations.
+  }
+  \label{fig:new-tmp5}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.95\textwidth]
+  {figs/fig-split-str2-simple.eps}
+  \caption{
+    splitting single well potential:
+    the plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
+    $F_e^{\textrm{max}} = 2.0$.
+    From left to right:
+    (a) $t=0$~\textsf{ps},
+    the brutal force non-equilibirum simulation,
+    (b) $t=20$~\textsf{ps},
+    the brutal force non-equilibirum simulation,
+    (c) $t=20$~\textsf{ps},
+    response formula with reference process $F_e^{\max} = 0.0$
+    (i.e. equilibirum simulation),
+    and
+    (d) $t=20$~\textsf{ps},
+    response formula with reference process $F_e^{\max} = 1.0$.
+    The response formula are truncated at order $n=1$.
+  }
+  \label{fig:new-tmp6}
+\end{figure}
+
 \section*{References}

### 2013-06-02 writedown/linear.response/paper.tex
add tex file for paper

--- 2012-05-15 writedown/linear.response/main.tex
@@
 
-\section{Theoretical considerations}
-\subsection{The linear response theory}
-In this sub-section, we give a brief description of the linear response
+\section{Theoretical background}
+\subsection{Classical linear response theory}
+
+In this section, we want to firstly give a brief description of the classical linear response
 theory, for details see Ref.~\cite{tuckeman2010statistical} for
-example. Here we assume the system is governed by equation:
-\begin{align}\label{eqn:gov-1}
-  \dot{\vect q}_i & = \frac{\partial\mh}{\partial \vect p_i}
-  + \vect C_i(\vect q,\vect p) F_e(t)\\\label{eqn:gov-2}
-  \dot{\vect p}_i & =- \frac{\partial\mh}{\partial \vect q_i}
-  + \vect D_i(\vect q,\vect p) F_e(t)
+example.
+In this paper we want to study the non-equilibrium system driven by some external force.
+More strictly, we investigate the system that is governed by the
+following Langevin equation:
+\begin{align}\label{eqn:langevin-1}
+  \dot{\vect q} & = \nabla_{\vect p}\mh(\vect q,\vect p), \\\label{eqn:langevin-2}
+  \dot{\vect p} & =- \nabla_{\vect q}\mh(\vect q,\vect p)
+  + \vect D(\vect q,\vect p) F_e(t)
+  - \gamma\vect p
+  + \sigma\dot{\vect W}.
 \end{align}
-Where $\vect C$ and $\vect D$ are perturbations to the system, and
-$F_e(t)$ is the strength of the perturbation as a function of time.
-The phase space imcompressibility condition is assumed:
+Here $\vect D$ is the external driving force applied to the system, and
+$F_e(t)$ is the strength of the driving as a function of time.
+We always assume that the magnitude of $\vect D$ is controllable, and
+the magnitude of $F_e(t)$ is small and of order $\mo (\varepsilon)$, $\varepsilon \ll 1$.
+The phase space incompressibility condition is assumed:
 \begin{align}
-  \sum_{i=1}^{3N}
-  \bigg[
-  \frac{\partial \vect C_i}{\partial \vect q_i} +
-  \frac{\partial \vect D_i}{\partial \vect p_i}
-  \bigg] = 0
+  \nabla_{\vect p}\cdot\vect D = 0
 \end{align}
 Then the phase space distribution $f(\vect x, t)$ ($\vect x = \{\vect
-q_i, \vect p_i\}$) is subject to the Liouville equation:
+q, \vect p\}$) is subject to the Kolmogorov forward equation:
+\begin{align}\label{eqn:orig-forward}
+  \frac{\partial}{\partial t} f(\vect x, t) - \fwg^\beta(t) f(\vect x, t) = 0
+\end{align}
+where $\fwg^\beta(t)$ is the forward infiniesimal generator given by
 \begin{align}
-  \frac{\partial}{\partial t} f(\vect x, t) + i\ml f(\vect x, t) = 0
+  \fwg^\beta(t) =
+  \frac{\sigma^2}2\Delta_{\vect p}
+  - \Big(
+  \nabla_{\vect p}\mh + \vect C F_e(t)
+  \Big)\cdot\nabla_{\vect q}
+  - \Big(
+  -\nabla_{\vect q}\mh +
+  \vect D F_e(t) - \gamma\vect p
+  \Big)\cdot\nabla_{\vect p}
+  + 3N\gamma,
 \end{align}
-where $\ml$ is the Liouville operator given by $i\ml = \dot{\vect
-  x}\cdot\nabla_{\vect x}$.  We assume that the perturbation to the
+and $\beta$ is the inverse temperature.
+The fluctuation-dissipation relation is always true:
+$\beta = 2\gamma / \sigma^2$.
+The equilibrium and non-equilibrium system
+is governed by
+the standard Langevin equation, i.e. with
+vanishing driving force $F_e(t) = 0$ in
+Eqn.~\eqref{eqn:langevin-1} and \eqref{eqn:langevin-2}.
+We assume that the driving to the
 system is so small that both the phase space distribution and the
-Liouville operator can be viewed as perturbation from the equilibrium
-distribution (invariant measure) and the unperturbed operator,
-respectively. Namely
+infiniesimal generator can be viewed as perturbation from the initial
+distribution and the un-driven generator
+($\vect D = 0$),
+respectively. We the expansion with respect to small
+variable $\varepsilon$:
+\begin{align}\label{eqn:f-expan}
+  f(\vect x, t) &= f_0(\vect x, t) + \varepsilon f_1(\vect x, t)
+  +\varepsilon^2 f_2(\vect x, t) + \mo (\varepsilon^3),
+\end{align}
+and 
+\begin{align}\label{eqn:A-expan}
+  \fwg^\beta(t) = \fwg^\beta_0 + \varepsilon\fwg_1(t),
+\end{align}
+where
 \begin{align}
-  f(\vect x, t) &= f_0(\vect x) + \Delta  f(\vect x, t)
+  \fwg_0^\beta =&
+  \frac{\sigma^2}2\Delta_{\vect p}
+  -
+  \nabla_{\vect p}\mh\cdot\nabla_{\vect q}
+  - \big(
+  -\nabla_{\vect q}\mh - \gamma\vect p
+  \big)\cdot\nabla_{\vect p}
+  + 3N\gamma,\\
+  \fwg_1(t) =&
+  - \frac{F_e(t)}{\varepsilon} \big(
+  \vect C \cdot\nabla_{\vect q}
+  +
+  \vect D \cdot\nabla_{\vect p}
+  \big).
 \end{align}
-and since the Eqn.~\eqref{eqn:gov-1} and \eqref{eqn:gov-2} imply that
-$\dot{\vect x} = \dot{\vect x}_0 + \Delta \dot{\vect x}(t)$
+Notice $F_e(t) \sim \mo(\varepsilon)$, so $\Vert\fwg_1(t)\Vert\sim\mo(1)$.
+The invariant measure of $\fwg^\beta_0$ satisfies $\fwg^\beta_0f^\beta_{\inv}(\vect x) = 0$,
+which is nothing but the Boltzmann distribution
 \begin{align}
-  i\ml =
-  (\dot{\vect x}_0 + \Delta \dot{\vect x}(t))\cdot\nabla_{\vect x}
-  = i\ml_0 + i\Delta\ml(t)
+  f^\beta_{\inv}(\vect x) \propto e^{-\beta\mh(\vect x)}
 \end{align}
-Where the invariant measure satisfies $i\ml_0f_0(\vect x) = 0$. Also
-$f_0(\vect x)$ has the form $f_0(\mh(\vect x))$ (?). The Liouville equation
-for the perturbed system is therefore:
+Inserting \eqref{eqn:f-expan} and \eqref{eqn:A-expan} into
+\eqref{eqn:orig-forward}, 
+the Kolmogorov forward equation becomes:
 \begin{align}
   \frac{\partial}{\partial t}
-  [f_0(\mh(\vect x)) + \Delta  f(\vect x, t)]
-  +
-  [i\ml_0 + i\Delta\ml(t)]
-  [f_0(\mh(\vect x)) + \Delta  f(\vect x, t)] = 0
+  [f_0(\vect x, t) + \varepsilon  f_1(\vect x, t) + \varepsilon^2f_2(\vect x, t)]
+  -
+  [\,\fwg^\beta_0 + \varepsilon\fwg_1(t)\,]
+  [f_0(\vect x, t) + \varepsilon  f_1(\vect x, t) + \varepsilon^2f_2(\vect x, t)]
+  = 0
 \end{align}
-The basic assumption for the ``linear response'' is that the high order
-term $i\Delta\ml(t)  \Delta  f(\vect x, t)$ in the Liouville equation
-can be neglected. Noticing $i\ml_0 f_0 = 0$, we have the Liouville equation
-for the leading order of perturbation:
-\begin{align}
+Matching this equation at different orders of $\varepsilon$, the zeroth
+order gives:
+\begin{align}\label{eqn:e-o0}
   \bigg[
   \frac{\partial}{\partial t}
-  + i\ml_0
+  - \fwg_0^\beta
   \bigg]
-  \Delta  f(\vect x, t)
+  f_0(\vect x, t)
+  = 0
+\end{align}
+which is the un-driven equation. Matching the first and the second order gives:
+\begin{align}\label{eqn:e-o1}
+  \bigg[
+  \frac{\partial}{\partial t}
+  - \fwg_0^\beta
+  \bigg]
+  f_1(\vect x, t)
+  =&
+  \fwg_1(t) f_0(\vect x, t)\\\label{eqn:e-o2}
+  \bigg[
+  \frac{\partial}{\partial t}
+  - \fwg_0^\beta
+  \bigg]
+  f_2(\vect x, t)
+  =&
+  \fwg_1(t) f_1(\vect x, t)
+\end{align}
+Formally solving \eqref{eqn:e-o0} -- \eqref{eqn:e-o2} with
+initial conditions $f_1(\vect x, 0) = 0$ and $f_2(\vect x, 0) = 0$
+gives
+\begin{align}\label{eqn:f0-0}
+  f_0(\vect x, t)
+  =&\,
+  e^{t\fwg^\beta_0}f_0(\vect x, 0) \\\label{eqn:f1-0}
+  f_1(\vect x, t)
+  =&\,
+  \int_0^t\dd s\,
+  e^{(t-s)\fwg^\beta_0}\circ
+  \fwg_1(s) f_0(\vect x,s)
   =
-  - i\Delta\ml(t) f_0(\mh(\vect x))
+  \int_0^t\dd s\,
+  e^{(t-s)\fwg^\beta_0}\circ
+  \fwg_1(s)\circ
+  e^{s\fwg^\beta_0}f_0(\vect x, 0) \\\nonumber
+  f_2(\vect x, t)
+  =&\,
+  \int_0^t\dd s\,
+  e^{(t-s)\fwg^\beta_0}\circ
+  \fwg_1(s) f_1(\vect x,s) \\\label{eqn:f2-0}
+  &=
+  \int_0^t\dd s
+  \int_0^s\dd u\,\,
+  e^{(t-s)\fwg^\beta_0}\circ
+  \fwg_1(s)\circ
+  e^{(s-u)\fwg^\beta_0}\circ
+  \fwg_1(u)\circ
+  e^{u\fwg^\beta_0}
+  f_0(\vect x, 0) 
 \end{align}
-The operator on l.h.s. is the same as the unperturbed Liouville equation.
-Formally solving this equation with initial condition
-$\Delta f(\vect x, 0) = 0$ gives
+If the initial distribution is the invariant distribution, i.e.
+$f_0(\vect x, 0) = f_{\inv}^\beta(\vect x)$,
+\eqref{eqn:f0-0} becomes:
 \begin{align}
-  \Delta  f(\vect x, t) =
-  - \int_0^t \dd s \:
-  e^{-i\ml_0(t-s)}
-  i\Delta\ml(s)f_0(\mh(\vect x))
+  f_0(\vect x, t) = f_{\inv}^\beta(\vect x)
 \end{align}
-By defining the \emph{dissipative flux},
+It can be shown that
+\begin{align}
+  \fwg_1(s) f_{\inv}^\beta(\vect x)
+  =
+  -\frac{F_e(s)}{\varepsilon}\beta
+  j(\vect x)
+  f_{\inv}^\beta(\vect x),
+\end{align}
+where the \emph{dissipative flux} $j(\vect x)$ is defined by,
 \begin{align}
   j(\vect x) =
-  -\sum_{i=1}^{3N}
-  \bigg[
-  \vect D_i(\vect x) \frac{\partial\mh}{\partial \vect p_i} +
-  \vect C_i(\vect x) \frac{\partial\mh}{\partial \vect q_i}   
-  \bigg]
+  -\vect C\cdot\nabla_{\vect q}\mh 
+  -\vect D\cdot\nabla_{\vect p}\mh
 \end{align}
-it has been shown that~\cite{tuckeman2010statistical}
+Then $f_1(\vect x, t)$ becomes:
 \begin{align}
-  i\Delta\ml(s)f_0(\mh(\vect x)) =
-  -\frac{\partial f_0(\mh)}{\partial\mh} j(\vect x)\,F_e(s)
+  f_1(\vect x, t)
+  =
+  -\frac\beta\varepsilon
+  \int_0^t\dd s\,
+  F_e(s)\,
+  e^{(t-s)\fwg^\beta_0}
+  \big[
+  j(\vect x)
+  f_{\inv}^\beta(\vect x)
+  \big]
 \end{align}
-For the canonical invariant measure, this is equivalent to
-\begin{align}
-  i\Delta\ml(s)f_0(\mh(\vect x)) =
-  \beta f_0(\mh(\vect x)) j(\vect x)F_e(s)
+Therefore, the linear order approximation to
+the solution of~\eqref{eqn:orig-forward} is 
+\begin{align}\label{eqn:solv-orig-linear}
+  f(\vect x, t) =
+  f_{\inv}^\beta(\vect x)
+  - \beta
+  \int_0^t\dd s\,
+  F_e(s)\,
+  e^{(t-s)\fwg^\beta_0}
+  \big[
+  j(\vect x)
+  f_{\inv}^\beta(\vect x)
+  \big]
+  + \mo(\varepsilon^2)
 \end{align}
-For any observation:
+For any time-dependent observation, the linear response approximation
+would be:
 \begin{align}\nonumber
-  \mathcal A(t)
+  \mathcal O(t)
   &=
-  \int\dd \vect x A(\vect x)f(\vect x, t)  \\\nonumber
+  \int\dd \vect x \:O(\vect x)f(\vect x, t)  \\\nonumber
   &=
-  \int\dd \vect x\, A(\vect x)\,
-  [f_0(\mh(\vect x)) + \Delta f(\vect x, t)] \\\nonumber
+  \int\dd \vect x\, O(\vect x)\,
+  [\,f_{\inv}^\beta(\vect x) + \varepsilon f_1(\vect x, t)] \\\nonumber
   &=
-  \langle A\rangle_0
+  \langle O\rangle
   -
   \beta
-  \int \dd \vect x
-  A(\vect x)
-  \int_0^t\dd s\,
-  e^{-i\ml_0(t-s)}
-  f_0(\mh(\vect x)) j(\vect x)F_e(s)  \\\nonumber
+  \int \dd \vect x\:
+  O(\vect x)
+  \int_0^t\dd s\,F_e(s)
+  e^{(t-s)\fwg_0^\beta}
+  \big[j(\vect x)f^\beta_{\inv}(\vect x)\big]
+  \\\nonumber
   &=
-  \langle A\rangle_0
+  \langle O\rangle
   -
@@
   \int_0^t\dd s\,
-  F_e(s)
+  F_e(t-s)
   \int \dd \vect x\,
-  f_0(\mh(\vect x))
-  A(\phi_{t-s}(\vect x))
-  j(\vect x)  
+  f^\beta_{\inv}(\vect x) j(\vect x)\,
+  e^{s\bwg_0^\beta}
+  O(\vect x)
+\end{align}
+Where $\bwg^\beta_0$ is the infiniesimal generator of the \emph{un-driven} 
+Kolmogorov backward equation, which is the adjoint operator of $\fwg^\beta_0$.
+We have
+\begin{align}
+  e^{s\bwg^\beta_0} O(\vect x) = \mathbb E_{\vect x} [O(\vect X^\beta_s)]
+\end{align}
+where $\vect X^\beta_t$ is the trajectory of the un-driven Langevin equation
+starting at $\vect x$.
+Therefore, we have
+% Notice that $e^{i\ml_0(t-s)}O(\vect x) = O(\phi_{t-s}(\vect x))$,
+% where $\phi_t(\vect x)$ is the flow mapping of the \emph{unperturbed}
+% Hamiltonian system. We have
+\begin{align}
+  \mathcal O(t)
+  =
+  \langle O\rangle
+  -
+  \beta
+  \int_0^t\dd s\,
+  F_e(t-s)
+  \int \dd \vect x\,
+  f^\beta_{\inv}(\vect x)
+  j(\vect x)\,
+  \mathbb E_{\vect x} [O(\vect X^\beta_s)]
 \end{align}
@@
   \int_0^t\dd s\,
-  F_e(s)
-  \langle
-  A(t-s)\,
+  F_e(t-s)
+  \big\langle
   j(0)\,
-  \rangle_0
+  \mathbb E_{\vect x} [O(\vect X^\beta_s)]
+  \big\rangle
 \end{align}
-where $\langle A(t-s)\,j(0)\,\rangle_0$ is the \emph{equilibrium time
-  correlation} between the observation $A$ at time $t-s$ and the
-dissipative flux $j$ at time 0.
+where $  \langle
+  j(0)\,
+  \mathbb E_{\vect x} [O(\vect X^\beta_s)]
+  \rangle$ is the \emph{equilibrium time
+  correlation} between the observation $O$ at time $s$ and the
+dissipative flux $j$.
 
-\subsection{The Core set as an observation of the system}
+\subsection{The non-equilibrium response theory}
 
-The core set plays a very important role in the MSM theory. One way to
-identify core set is to consider the ``stability'' of the invariant
-measure $f_0$ under some mapping $\mt_\alpha$ of time step $\alpha$:
+The classical response theory mainly has the following two limitations:
+Firstly, one has to start from an equilibrium simulation. That means if the
+external driving is no longer small ($\varepsilon$ is relatively big), the
+first order (or linear) response~\eqref{eqn:lr} may not be a good approximation to
+the true non-equilibrium process any more.
+One possible way to improve the accuracy is to introduce the second
+order response term, i.e.~\eqref{eqn:f2-0}.
+Here one encounters the second limitation of the classical
+response theory: it does not provide a frame work, under which
+the second and even higher order responses can be easily derived.
+The limitations of the classical responses theory motivate us to
+develop a more general frame work that allows one starting from any
+non-equilibrium process (arbitrarily driven), and calculating from it as many orders of
+responses as necessary.
+
+In the language of path integral, the time-dependent average $O( t)$
+can be written as
 \begin{align}
-  C_\alpha = \big\{
-  \vect x \,\vert\, \mt_\alpha f_0(\vect x) > f_0(\vect x)
+  O( t) =
+  \langle O[\vect X_s]\,\rangle_{0,t} = 
+  \int\dd \vect u\, f(\vect u, 0) 
+  \int_{\mc\{\vect u,0; t\}}
+  O[\vect X_s]\, \dd\mathcal P[\vect X_s]  
+\end{align}
+where $O[\vect X_s]$ is the observable of the trajectories, which is a generalization of the
+observable that only depends on the configuration of the system.
+$\langle \cdot \,\rangle_{0,t}$ is the ensemble average over all possible trajectories in time interval $[0,t]$.
+$f(\vect x, 0)$ is the initial distribution that may not be the equilibrium distribution.
+$\mc\{\vect u,0; t\}$ is the set of all continuous trajectories starting from point
+$\vect u$ at time $0$, and ending at some position at time $t$.
+$\mathcal P[\vect X_s]$ is the
+probability measure of the trajectories generated by driven Langevin dynamics~\eqref{eqn:langevin-1}--\eqref{eqn:langevin-2},
+where the intensity $F_e$ need not to be 0.
+Now we want to consider a system that is a small perturbation of the previous system (we now call it ``reference system'')
+in the sense that it has a similar driven force: $F_e + \Delta F_e$, where $\Delta F_e\sim \mo(\varepsilon)$ is small.
+The probability measure of the trajectories of this perturbed system is denoted by
+$\tilde{\mathcal P}[\vect X_s]$.
+The time-dependent average of observable of the system $F_e + \Delta F_e$ can be written down in terms of
+the probability measure of the reference system:
+\begin{align}
+  \tilde O(t) =
+  \int \dd\vect u f(\vect x,0)
+  \int_{\mc\{\vect u,0; t\}} O[\vect X_s]\,
+  \frac{\dd\tilde{\mathcal P}[\vect X_s]}{\dd\mathcal P[\vect X_s]}
+  \, \dd\mathcal P[\vect X_s]    
+\end{align}
+by using the Girsanov formula, 
+\begin{align}\label{eqn:girsanov}
+  \frac
+  {\dd\tilde{\mathcal P}[\vect X_s]}
+  {\dd\mathcal P[\vect X_s]}  
+  =&\,
+  \exp
+  \big\{{G[\vect X_s] - \frac12 H[\vect X_s]}
   \big\}
 \end{align}
-Generally, it is impossible and not necessary to consider the core set
-in the $3N$ dimensional configuration space. For example, a
-macromolecule solving in water, the DOFs of water is not of interest
-at all. Therefore, we consider some phase space compressing function
-$\vect y = \vect y(\vect x)$, with $\vect y \in \Omega \subset \mathbb
-R^M,\ M\ll 3N$. We call $\Omega$ the observation space.  The projected
-invariant measure on the observation space is therefore
+where the short notations are defined by:
 \begin{align}
-  \mathcal P f_0 (\vect y)
+  G[\vect X_s]
+  &= \frac{1}\sigma\int_0^t
+  \Delta F_e(s)\,\vect D(\vect q_s,\vect p_s)\cdot\dd\vect W_s \\
+  H[\vect X_s]
+  &=
+  \frac 1{2\sigma^2}\int_0^t
+  \Delta F_e^2(s)\,\vect D^2(\vect q_s,\vect p_s)\,\dd s
+\end{align}
+If the difference $\Delta F_e(t)$ is of order $\mo (\varepsilon)$, then $G[\vect X_s] \sim \varepsilon$ is
+the first order term, and $H[\vect X_s] \sim \varepsilon^2$  is the second order term.
+Then with the Taylor expansion of the right-hand-side of Eq.~\eqref{eqn:girsanov}, we have
+\begin{align}\label{eqn:neq-response}
+  \tilde O(t) -
+  O(t)
   =
-  \int
-  \dd \vect x\,
-  f_0(\vect x) \,
-  \delta (\vect y - \vect y(\vect x)) 
+  \big\langle O[\vect X_s]\cdot G[\vect X_s]\big\rangle_{0,t}
+  +
+  \frac12 \big\langle O[\vect X_s] \cdot (G^2[\vect X_s] - H[\vect X_s] )\big\rangle_{0,t}
+  +
+  \mo (\varepsilon^3)
 \end{align}
-The core set on the observation space is
+Here are some remarks of the main result Eq.~\eqref{eqn:neq-response}:
+\begin{itemize}
+\item The first term on the right-hand-side of Eq.~\eqref{eqn:neq-response} is the linear
+  response of the reference non-equilibrium process (driven by $F_e$). The second term is the second order
+  response. One can easily write down higher orders of response by the Taylor expansion
+  of the right-hand-side of Eq.~\eqref{eqn:girsanov}.
+\item When the observable $O[\vect X_s]$ only depends on the end point of the trajectory, i.e.
+  $O[\vect X_s] = O(\vect X_t)$, then $\big\langle O(\vect X_t)\cdot G[\vect X_s]\big\rangle_{0,t}$
+  gives the well-know time correlation form of the linear response.
+\item If one wants to calculate a large amount of different perturbations, then apply one by one
+  Eq.~\eqref{eqn:neq-response} is not convenient, because one has to calculate the responses
+  again when the function form of $\Delta F_e$ is not a linear combination of previously calculated perturbations.
+  If it is possible to
+  express different $\Delta F_e$ by a group of base functions, then the responses are only need
+  to calculated on the bases. Then with a linear conbination of the responses on bases, one can derive
+  the responses of the whole group of perturbations. 
+\item In the prove, we assume that the reference and the perturbed non-equilibrium processes
+  have the same initial distribution (both of them may be different from the equilibrium distribution),
+  but this is not a necessary. One can easily
+  applies similar perturbation approach to the initial distribution as  to the trajectory ensemble.
+  To keep the story simple, we do not develop the theory from this aspect.
+\end{itemize}
+
+\section{Numerical tests}
+
+\subsection{The tilting double-well potential}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.4\textwidth]{figs/fig-tilt-pot.eps}
+  \caption{The double-well potential with tilting perturbation.}
+  \label{fig:tmp1}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.95\textwidth]
+  {figs/fig-tilt-str2-2d-more.eps}
+  \caption{The plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
+    $F_e^{\textrm{max}} = 2$.
+    The 1st row: the direct non-equilibrium
+    simulation. The 2nd row: the results of linear response
+    formula~\eqref{eqn:core-identify-approx-2}.
+    The 3rd row: the results of response
+    formula~\eqref{eqn:pert-approx-1} and \eqref{eqn:pert-approx-2}.
+    The order is 2, the reference state is $F_e^{\textrm{max}} = 0$, i.e.
+    the equilibrium state.
+  }
+  \label{fig:tmp2}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.95\textwidth]
+  {figs/fig-tilt-str4-2d-more.eps}
+  \caption{The plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
+    $F_e^{\textrm{max}} = 4$.
+    The 1st row: the direct non-equilibrium
+    simulation. The 2nd row: the results of linear response
+    formula~\eqref{eqn:core-identify-approx-2}.
+    The 3rd row: the results of response
+    formula~\eqref{eqn:pert-approx-1} and \eqref{eqn:pert-approx-2}.
+    The order is 2, the reference state is $F_e^{\textrm{max}} = 0$, i.e.
+    the equilibrium state.
+    The 4th row: the results of response
+    formula~\eqref{eqn:pert-approx-1} and \eqref{eqn:pert-approx-2}.
+    The order is 1, the reference state is $F_e^{\textrm{max}} = 2$.
+  }
+  \label{fig:tmp3}
+\end{figure}
+
+We test the idea firstly by a one-dimensional model system: one particle in a
+tilting double-well potential. For convenience, we let the mass of the
+particle to be 1 \textsf{amu}. The unperturbed
+Hamiltonian of the system is given by:
 \begin{align}
-  \widetilde
-  C_\alpha = \big\{
-  \vect y \,\vert\,
-  \mathcal P\mt_\alpha f_0(\vect y) > \mathcal P f_0(\vect y)
-  \big\}
+  \mh (\vect p, \vect q) = \frac 12 \vect p^2 + U(\vect q) 
 \end{align}
-% with $\widetilde \mt_\alpha$ defined by
-% $\widetilde\mt_\alpha = \mathcal P\circ\mt_\alpha$, namely
-% \begin{align}
-%   \widetilde\mt_\alpha\, p_0(\vect y)
-%   =
-%   \mathcal P\circ\mt_\alpha f_0 (\vect y)
-%   =
-%   \int
-%   \dd \vect x\,
-%   [\mt_\alpha f_0](\vect x) \,
-%   \delta (\vect y - \vect y(\vect x)) 
-% \end{align}
-For convenience, we define $\Delta \mt_\alpha =
-\mt_\alpha - \mathcal I$, where $\mathcal I$ is the
-identity. If $[\mathcal P\circ\Delta \mt_\alpha f_0](\vect y) > 0$, then
-$\vect y$ is in the core set $\widetilde C_\alpha $.
-\begin{align}\nonumber
-  [\mathcal P\circ\Delta \mt_\alpha f_0](\vect y)
-  &=
-  \int
-  \dd \vect x\,
-  \Delta \mt_\alpha f_0(\vect x) \,
-  \delta (\vect y - \vect y(\vect x))  \\\nonumber
-  &=
-  \lim_{\sigma \rightarrow 0}
-  \int
-  \dd \vect x\,
-  \Delta \mt_\alpha f_0(\vect x) \,
-  K_{\sigma} (\vect y - \vect y(\vect x))  \\\nonumber
-  &=
-  \lim_{\sigma \rightarrow 0}
-  \int
-  \dd \vect x\,
-  \Delta \mt_\alpha f_0(\vect x) \,
-  K_{\sigma, \vect y} (\vect x)  \\\nonumber
-  &=
-  \lim_{\sigma \rightarrow 0}
-  \int
-  \dd \vect x\,
-  f_0(\vect x) \,
-  [\Delta \mt_\alpha]^\dagger K_{\sigma, \vect y} (\vect x)\\
-  &=
-  \lim_{\sigma \rightarrow 0}\,
-  \langle
-  [\Delta \mt_\alpha]^\dagger K_{\sigma, \vect y}
-  \rangle_0
+with potential
+\begin{align}
+  U(\vect q) = \frac12 k (\vect q^2 - a^2)^2
 \end{align}
-where $[\Delta \mt_\alpha]^\dagger$ is the adjoint operator of $\Delta
-\mt_\alpha$, which is actually $\Delta \mt_{-\alpha}$. The delta
-function is replaced by the limit of a mollifier $K_\sigma(\vect y)$,
-which is infinitely smooth, compact supported of size $\sigma$,
-bounded and $\int
-K_\sigma(\vect y)\dd \vect y = 1$.  The physical meaning is that we
-can consider the expectation value of the stability of the mollifier
-$K_{\sigma, \vect y}$ instead of the stability of the invariant measure.
-This provides the convenience of using the linear response theory showed
-in the previous section. By Eqn.~\eqref{eqn:lr},
-\begin{align}\nonumber
-  [\mathcal P\circ\Delta \mt_\alpha f_t](\vect y)
-  &=
-  \lim_{\sigma \rightarrow 0}\,
-  \langle
-  [\Delta \mt_\alpha]^\dagger K_{\sigma, \vect y}
-  \rangle_t \\
-  &=
-  \lim_{\sigma \rightarrow 0}\,
-  \Big\{
-  \langle
-  [\Delta \mt_\alpha]^\dagger K_{\sigma, \vect y}
-  \rangle_0  -
-  \beta
-  \int_0^t\dd s\,
-  F_e(s)
-  \langle
-  [\Delta \mt_\alpha]^\dagger K_{\sigma, \vect y}(t-s)\,
-  j(0)\,
-  \rangle_0
-  \Big\}
+Here $k = 8$~$\textsf{kJ} / (\textsf{mol nm}^4)$, and $ a = 1\ \textsf{nm}$.
+Notice at room temperature $300\ \textsf{K}$, $k_BT = 2.48$~\textsf{kJ/mol}.
+See the red line in Fig.~\ref{fig:tmp1}.
+%$\textsf{kJ} / (\textsf{mol nm}^4)}$
+The perturbation is given by
+\begin{align}
+  \vect C(\vect p, \vect q) = 0; \qquad
+  \vect D(\vect p, \vect q) = -\nabla_{\vect q} V(\vect q) = 1
 \end{align}
-Therefore, the core set at time $t$ is defined by
+Here $V(\vect q) = -\vect q$ is  effectively tilt the original
+potential $U(\vect q)$. The strength of the perturbation $F_e(t)$ is given
+by the following function:
 \begin{align}
-  \widetilde C_\alpha^t =
-  \Big\{
-  \vect y
-  \,\Big\vert\,
-  [\mathcal P\circ\Delta \mt_\alpha f_t](\vect y) > 0
-  \Big\}
+  F_e(t) = 
+  \begin{cases}
+    F_e^{\textrm{max}}\times (t / t_c) & t < t_c \\
+    F_e^{\textrm{max}} & t \geq t_c
+  \end{cases}
 \end{align}
+$t_c = 20$~\textsf{ps} so that the perturbation increases slow
+enough: the typical decaying time scale of the correlation function
+in~\eqref{eqn:core-identify-approx-2} is roughly 3~\textsf{ps}.
+We choose $\tau = 1$~\textsf{ps}.
+% We choose $\mt_\alpha$
+% to be the propagator of the Langevin dynamics at temperature $150$
+% \textsf{K} with $\alpha = 1$ \textsf{ps}.
+Since the dynamics is in
+1-d, the projection $\mathcal P$ is identity.  We consider the stability of the
+distribution $f(\vect x, t)$ at time $t$: $\Delta\mt_\tau f(\vect x, t)$.
+The larger
+this value, the more stable the core sets are.
+
+See Fig. \ref{fig:tmp2} and \ref{fig:tmp3} for numerical result.  Both
+the direct non-equilibrium simulation and the response
+approximations are presented. Time slices $t = 0$, 10, 15 and 20
+\textsf{ps} are shown.  When the perturbation is small,
+i.e. $F_e^{\textrm{max}} = 2$, the direct non-equilibrium simulation
+and the linear response results are accurately consistent.
+The 2nd order response formula~\eqref{eqn:pert-approx-1} and
+\eqref{eqn:pert-approx-2} are even better, but the numerical error
+is bigger.
+As the
+perturbation grows stronger, the left well becomes less stable, while
+the right well becomes more stable.  When the perturbation is big,
+i.e. $F_e^{\textrm{max}} = 4$, the linear response result is no
+longer precise.
+When $t\geq 20$ \textsf{ps}, the left well basically
+disappears, however, the linear response calculation presents
+artifical occilation at the right well.
+The 2nd order response also has artifical effect, and the numerical
+uncertainty is high.
+The 4th rwo of Fig.~\ref{fig:tmp3} presents the 1st order result of the
+response formula~\eqref{eqn:pert-approx-1} and
+\eqref{eqn:pert-approx-2}, with
+respect to the reference state $F_e^{\textrm{max}} = 2$. The result
+is impressively improved compared with the linear response with respect to the
+equilibrium state.
+
+\subsection{Splitting single-well potential}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.4\textwidth]{figs/fig-split-pot.eps}
+  \caption{The single-well potential with splitting perturbation.}
+  \label{fig:tmp4}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.95\textwidth]
+  {figs/fig-split-str1-2d-more.eps}
+  \caption{The plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
+    $F_e^{\textrm{max}} = 1.0$.
+    The 1st row: the direct non-equilibrium
+    simulation. The 2nd row: the results of linear response
+    formula~\eqref{eqn:core-identify-approx-2}.
+    The 3rd row: the results of response
+    formula~\eqref{eqn:pert-approx-1} and \eqref{eqn:pert-approx-2}.
+    The order is 2, the reference state is $F_e^{\textrm{max}} = 0$, i.e.
+    the equilibrium state.
+  }
+  \label{fig:tmp5}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.95\textwidth]
+  {figs/fig-split-str2-2d-more.eps}
+  \caption{The plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
+    $F_e^{\textrm{max}} = 2.0$.
+    The 1st row: the direct non-equilibrium
+    simulation. The 2nd row: the results of linear response
+    formula~\eqref{eqn:core-identify-approx-2}.
+    The 3rd row: the results of response
+    formula~\eqref{eqn:pert-approx-1} and \eqref{eqn:pert-approx-2}.
+    The order is 2, the reference state is $F_e^{\textrm{max}} = 0.0$, i.e.
+    the equilibrium state.
+    The 4th row: the results of response
+    formula~\eqref{eqn:pert-approx-1} and \eqref{eqn:pert-approx-2}.
+    The order is 1, the reference state is $F_e^{\textrm{max}} = 1.0$.
+  }
+  \label{fig:tmp6}
+\end{figure}
+
+In this splitting single-well potential example, the unperturbed
+potential is simply the harmonic potential given by:
+\begin{align}
+  U(\vect q) = \frac12\,k\,\vect q^2 
+\end{align}
+Here $k = 8$~$\textsf{kJ} / (\textsf{mol nm}^2)$.
+Notice at room temperature $300\ \textsf{K}$, $k_BT = 2.48$~\textsf{kJ/mol}.
+See the red line in Fig.~\ref{fig:tmp4}.
+%$\textsf{kJ} / (\textsf{mol nm}^4)}$
+The perturbation is again given by
+\begin{align}
+  \vect C(\vect p, \vect q) = 0; \qquad
+  \vect D(\vect p, \vect q) = -\nabla_{\vect q} V(\vect q) 
+\end{align}
+where $V(\vect q)$ is a Gaussian type potential:
+\begin{align}
+  V(\vect q) = \frac{1}{\sqrt{2\pi \sigma^2}}
+  \exp\Big\{-\frac{\vect q^2}{2\sigma^2}\Big\}
+\end{align}
+we use $\sigma = 0.16$~\textsf{nm}. Please see Fig~\ref{fig:tmp4}.
+For the results, please see Fig.~\ref{fig:tmp5} -- Fig.~\ref{fig:tmp6}.
+
+\newpage
+\section{Numerical results}
+\subsection{Tilting double-well potential}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.4\textwidth]{figs/fig-tilt-pot.eps}
+  \includegraphics[width=0.4\textwidth]{figs/fig-split-pot.eps}
+  \caption{Two testing cases presented by this paper.
+    Left: the double-well potential with tilting perturbation.
+    Right: the single-well potential with splitting perturbation.}
+  \label{fig:new-tmp1}
+\end{figure}
+
+We test the idea by a one-dimensional model system: one particle in a
+tilting double-well potential. For convenience, we let the mass of the
+particle to be 1 \textsf{amu}. The unperturbed
+Hamiltonian of the system is given by:
+\begin{align}
+  \mh (\vect p, \vect q) = \frac 12 \vect p^2 + U(\vect q) 
+\end{align}
+with potential
+\begin{align}
+  U(\vect q) = \frac12 k (\vect q^2 - a^2)^2
+\end{align}
+Here $k = 8$~$\textsf{kJ} / (\textsf{mol nm}^4)$, and $ a = 1\ \textsf{nm}$.
+Notice at room temperature $300\ \textsf{K}$, $k_BT = 2.48$~\textsf{kJ/mol}.
+See the red line in Fig.~\ref{fig:tmp1}.
+%$\textsf{kJ} / (\textsf{mol nm}^4)}$
+The perturbation is given by
+\begin{align}
+  \vect C(\vect p, \vect q) = 0; \qquad
+  \vect D(\vect p, \vect q) = -\nabla_{\vect q} V(\vect q) = 1
+\end{align}
+Here $V(\vect q) = -\vect q$ is  effectively tilting the original
+potential $U(\vect q)$. The strength of the perturbation $F_e(t)$ is given
+by the following function:
+\begin{align}
+  F_e(t) = 
+  \begin{cases}
+    F_e^{\textrm{max}}\times (t / t_c) & t < t_c \\
+    F_e^{\textrm{max}} & t \geq t_c
+  \end{cases}
+\end{align}
+$t_c = 20$~\textsf{ps} so that the perturbation increases slow
+enough: the typical decaying time scale of the correlation function
+in~\eqref{eqn:core-identify-approx-2} is roughly 3~\textsf{ps}.
+We choose $\tau = 1$~\textsf{ps}.
+% We choose $\mt_\alpha$
+% to be the propagator of the Langevin dynamics at temperature $150$
+% \textsf{K} with $\alpha = 1$ \textsf{ps}.
+Since the dynamics is in
+1-d, the projection $\mathcal P$ is identity.  We consider the stability of the
+distribution $f(\vect x, t)$ at time $t$: $\Delta\mt_\tau f(\vect x, t)$.
+The larger
+this value, the more stable the core sets are.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.95\textwidth]
+  {figs/fig-tilt-str2-simple.eps}
+  \caption{
+    Tilting double well potential testing case:
+    the plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
+    $F_e^{\textrm{max}} = 2.0$.
+    From left to right:
+    (a) $t=0$~\textsf{ps},
+    the brutal force non-equilibirum simulation,
+    (b) $t=20$~\textsf{ps},
+    the brutal force non-equilibirum simulation,
+    (c) $t=20$~\textsf{ps},
+    response formula truncated at order $n=1$,
+    and
+    (d) $t=20$~\textsf{ps},
+    response formula truncated at order $n=2$.
+    The equilibirum state serves as the reference of
+    the response simulations.
+  }
+  \label{fig:new-tmp2}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.95\textwidth]
+  {figs/fig-tilt-str4-simple.eps}
+  \caption{
+    Tilting double well potential:
+    the plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
+    $F_e^{\textrm{max}} = 4.0$.
+    From left to right:
+    (a) $t=0$~\textsf{ps},
+    the brutal force non-equilibirum simulation,
+    (b) $t=20$~\textsf{ps},
+    the brutal force non-equilibirum simulation,
+    (c) $t=20$~\textsf{ps},
+    response formula with reference process $F_e^{\max} = 0.0$
+    (i.e. equilibirum simulation),
+    and
+    (d) $t=20$~\textsf{ps},
+    response formula with reference process $F_e^{\max} = 2.0$.
+    The response formula are truncated at order $n=1$.
+  }
+  \label{fig:new-tmp3}
+\end{figure}
+
+The Fig.~\ref{fig:new-tmp2} and \ref{fig:new-tmp3} present
+the application of the response theory to
+calculate core sets of the tilting double-well potential.
+Fig.~\ref{fig:new-tmp2} presents
+the snapshot at $t=20$~\textsf{ps}
+of $\Delta\mt_\tau f(\vect x,t)$ under the maximum perturbation
+strength $F_e^{\max} = 2.0$. Response formula truncated
+at both order $n=1$ and $n=2$ are compared to 
+the brutal force non-equilibirum simulation.
+No obviously difference is shown. The second order response result
+is slightly more precise than the first order one, but the
+statistical error is also larger, which makes the profile noisy.
+This figure tells us that
+(1) the modeling error of the linear response (1st order) is
+negligibly small; (2) the statistical error of numerically measuring the
+linear response is also small.
+Fig.~\ref{fig:new-tmp3} presents the $t=20$~\textsf{ps}
+snapshots of $\Delta\mt_\tau f(\vect x,t)$
+under a larger perturbations $F_e^{\max} = 4.0$.
+The linear reponse using the equilibirum process as the reference
+shows poor accuracy: the left well vanishes in this case,
+while the linear response produces  artifical structures.
+Since the profile is smooth, which indicates
+the statistical error of calculating the linear term is small,
+the deviation comes from the modellng error, namely the contribution
+of higher order response terms.
+By including the second order response term reduces the left
+artifical core set, but also creates artifical finer grained
+structures. The statistical accuracy is also worse than the linear response.
+Plot (d) present the result of the linear response
+formula using reference process $F_e^{\max} = 2.0$, which
+is the process presented in Fig.~\ref{fig:new-tmp2}.
+Much better modelling accuracy is achieved:
+the left artifical core set disappear and the
+statistical accuracy is also satisfactory.
+
+% See Fig. \ref{fig:tmp2} and \ref{fig:tmp3} for numerical result.  Both
+% the direct non-equilibrium simulation and the response
+% approximations are presented. Time slices $t = 0$, 10, 15 and 20
+% \textsf{ps} are shown.  When the perturbation is small,
+% i.e. $F_e^{\textrm{max}} = 2$, the direct non-equilibrium simulation
+% and the linear response results are accurately consistent.
+% The 2nd order response formula~\eqref{eqn:pert-approx-1} and
+% \eqref{eqn:pert-approx-2} are even better, but the numerical error
+% is bigger.
+% As the
+% perturbation grows stronger, the left well becomes less stable, while
+% the right well becomes more stable.  When the perturbation is big,
+% i.e. $F_e^{\textrm{max}} = 4$, the linear response result is no
+% longer precise.
+% When $t\geq 20$ \textsf{ps}, the left well basically
+% disappears, however, the linear response calculation presents
+% artifical occilation at the right well.
+% The 2nd order response also has artifical effect, and the numerical
+% uncertainty is high.
+% The 4th rwo of Fig.~\ref{fig:tmp3} presents the 1st order result of the
+% response formula~\eqref{eqn:pert-approx-1} and
+% \eqref{eqn:pert-approx-2}, with
+% respect to the reference state $F_e^{\textrm{max}} = 2$. The result
+% is impressively improved compared with the linear response with respect to the
+% equilibrium state.
+
+\subsection{Splitting single-well potential}
+
+In this splitting single-well potential example, the unperturbed
+potential is simply the harmonic potential given by:
+\begin{align}
+  U(\vect q) = \frac12\,k\,\vect q^2 
+\end{align}
+Here $k = 8$~$\textsf{kJ} / (\textsf{mol nm}^2)$.
+Notice at room temperature $300\ \textsf{K}$, $k_BT = 2.48$~\textsf{kJ/mol}.
+See the red line in Fig.~\ref{fig:tmp4}.
+%$\textsf{kJ} / (\textsf{mol nm}^4)}$
+The perturbation is again given by
+\begin{align}
+  \vect C(\vect p, \vect q) = 0; \qquad
+  \vect D(\vect p, \vect q) = -\nabla_{\vect q} V(\vect q) 
+\end{align}
+where $V(\vect q)$ is a Gaussian type potential:
+\begin{align}
+  V(\vect q) = \frac{1}{\sqrt{2\pi \sigma^2}}
+  \exp\Big\{-\frac{\vect q^2}{2\sigma^2}\Big\}
+\end{align}
+we use $\sigma = 0.16$~\textsf{nm}. Please see Fig~\ref{fig:tmp4}.
+For the results, please see Fig.~\ref{fig:tmp5} -- Fig.~\ref{fig:tmp6}.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.95\textwidth]
+  {figs/fig-split-str1-simple.eps}
+  \caption{
+    Splitting single-well potential testing case:
+    the plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
+    $F_e^{\textrm{max}} = 1.0$.
+    From left to right:
+    (a) $t=0$~\textsf{ps},
+    the brutal force non-equilibirum simulation,
+    (b) $t=20$~\textsf{ps},
+    the brutal force non-equilibirum simulation,
+    (c) $t=20$~\textsf{ps},
+    response formula truncated at order $n=1$,
+    and
+    (d) $t=20$~\textsf{ps},
+    response formula truncated at order $n=2$.
+    The equilibirum state serves as the reference of
+    the response simulations.
+  }
+  \label{fig:new-tmp5}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.95\textwidth]
+  {figs/fig-split-str2-simple.eps}
+  \caption{
+    splitting single well potential:
+    the plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
+    $F_e^{\textrm{max}} = 2.0$.
+    From left to right:
+    (a) $t=0$~\textsf{ps},
+    the brutal force non-equilibirum simulation,
+    (b) $t=20$~\textsf{ps},
+    the brutal force non-equilibirum simulation,
+    (c) $t=20$~\textsf{ps},
+    response formula with reference process $F_e^{\max} = 0.0$
+    (i.e. equilibirum simulation),
+    and
+    (d) $t=20$~\textsf{ps},
+    response formula with reference process $F_e^{\max} = 1.0$.
+    The response formula are truncated at order $n=1$.
+  }
+  \label{fig:new-tmp6}
+\end{figure}
 

### 2013-06-04 writedown/linear.response/paper.tex
rewrite the result part, so that it is consistent with the updated theoy language

@@
 \newcommand{\mh}{\mathcal H}
+\newcommand{\eps}{\varepsilon}
 \newcommand{\ml}{\mathcal L}
@@
 \author{Han Wang}
-% \affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}
+\author{Carsten Hartmann}
+\affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}
 \author{Christof Sch\"utte}
 \affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}
+\affiliation{Zuse Institute Berlin (ZIB), Germany}
 % \affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}
@@
 \section{Theoretical background}
-\subsection{Classical linear response theory}
+\subsection{Classical equilibrium linear response theory}
 
@@
 example.
-In this paper we want to study the non-equilibrium system driven by some external force.
+We study the non-equilibrium system driven by some external force.
 More strictly, we investigate the system that is governed by the
@@
   \dot{\vect p} & =- \nabla_{\vect q}\mh(\vect q,\vect p)
-  + \vect D(\vect q,\vect p) F_e(t)
+  + \eps F_e(t) \vect D(\vect q) 
   - \gamma\vect p
@@
 \end{align}
-Here $\vect D$ is the external driving force applied to the system, and
-$F_e(t)$ is the strength of the driving as a function of time.
-We always assume that the magnitude of $\vect D$ is controllable, and
-the magnitude of $F_e(t)$ is small and of order $\mo (\varepsilon)$, $\varepsilon \ll 1$.
-The phase space incompressibility condition is assumed:
-\begin{align}
-  \nabla_{\vect p}\cdot\vect D = 0
-\end{align}
+Here $\vect D$ is the external driving force applied to the system,
+the magnitude of which is always assumed to be controllable.
+$\eps F_e(t)$ is the strength of the driving as a function of time.
+The prefactor $\eps$ indicates that we regard the driving force to be considerably smaller than the internal forces. We will comment on what is meant by "considerably smaller" further below.
+
+% , and
+% the magnitude of $F_e(t)$ is small and of order $\mo (\eps)$, $\eps \ll 1$.
+
 Then the phase space distribution $f(\vect x, t)$ ($\vect x = \{\vect
@@
   \frac{\sigma^2}2\Delta_{\vect p}
-  - \Big(
-  \nabla_{\vect p}\mh + \vect C F_e(t)
-  \Big)\cdot\nabla_{\vect q}
+  - \nabla_{\vect p}\mh 
+  \cdot\nabla_{\vect q}
   - \Big(
   -\nabla_{\vect q}\mh +
-  \vect D F_e(t) - \gamma\vect p
+  \eps F_e(t)  \vect D  - \gamma\vect p
   \Big)\cdot\nabla_{\vect p}
@@
 \end{align}
-and $\beta$ is the inverse temperature.
-The fluctuation-dissipation relation is always true:
-$\beta = 2\gamma / \sigma^2$.
-The equilibrium and non-equilibrium system
+and $\beta2=\gamma / \sigma^2$ is the inverse temperature.
+The equilibrium system
 is governed by
 the standard Langevin equation, i.e. with
-vanishing driving force $F_e(t) = 0$ in
+vanishing driving force $\eps = 0$ in
 Eqn.~\eqref{eqn:langevin-1} and \eqref{eqn:langevin-2}.
-We assume that the driving to the
-system is so small that both the phase space distribution and the
-infiniesimal generator can be viewed as perturbation from the initial
-distribution and the un-driven generator
-($\vect D = 0$),
-respectively. We the expansion with respect to small
-variable $\varepsilon$:
+
+Under our smallness assumption on the driving force the full phase space distribution has the following asymptotic expansion in $\eps$:
 \begin{align}\label{eqn:f-expan}
-  f(\vect x, t) &= f_0(\vect x, t) + \varepsilon f_1(\vect x, t)
-  +\varepsilon^2 f_2(\vect x, t) + \mo (\varepsilon^3),
+  f(\vect x, t) &= f_0(\vect x, t) + \eps f_1(\vect x, t)
+  +\eps^2 f_2(\vect x, t) + \mo (\eps^3),
 \end{align}
-and 
+where $f_0=f_0(x,t)$ denotes the evolution of the phase space distribution for the undriven system (with $\eps=0$).
+In addition we have:
 \begin{align}\label{eqn:A-expan}
-  \fwg^\beta(t) = \fwg^\beta_0 + \varepsilon\fwg_1(t),
+  \fwg^\beta(t) = \fwg^\beta_0 + \eps\fwg_1(t),
 \end{align}
@@
   \fwg_1(t) =&
-  - \frac{F_e(t)}{\varepsilon} \big(
-  \vect C \cdot\nabla_{\vect q}
-  +
-  \vect D \cdot\nabla_{\vect p}
-  \big).
+  - F_e(t)  \vect D \cdot\nabla_{\vect p}.
 \end{align}
-Notice $F_e(t) \sim \mo(\varepsilon)$, so $\Vert\fwg_1(t)\Vert\sim\mo(1)$.
-The invariant measure of $\fwg^\beta_0$ satisfies $\fwg^\beta_0f^\beta_{\inv}(\vect x) = 0$,
-which is nothing but the Boltzmann distribution
+
+The invariant measure of $\fwg^\beta_0$ satisfies $\fwg^\beta_0f^\beta_{\inv}(\vect x) = 0$, and is given by the Boltzmann distribution
 \begin{align}
@@
   \frac{\partial}{\partial t}
-  [f_0(\vect x, t) + \varepsilon  f_1(\vect x, t) + \varepsilon^2f_2(\vect x, t)]
+  [f_0(\vect x, t) + \eps  f_1(\vect x, t) + \eps^2f_2(\vect x, t)]
   -
-  [\,\fwg^\beta_0 + \varepsilon\fwg_1(t)\,]
-  [f_0(\vect x, t) + \varepsilon  f_1(\vect x, t) + \varepsilon^2f_2(\vect x, t)]
+  [\,\fwg^\beta_0 + \eps\fwg_1(t)\,]
+  [f_0(\vect x, t) + \eps  f_1(\vect x, t) + \eps^2f_2(\vect x, t)]
   = 0
 \end{align}
-Matching this equation at different orders of $\varepsilon$, the zeroth
+Matching this equation at different orders of $\eps$, the zeroth
 order gives:
@@
 \end{align}
-which is the un-driven equation. Matching the first and the second order gives:
+which is the Kolmogorov forward equation of the undriven system. Matching the first and the second order gives:
 \begin{align}\label{eqn:e-o1}
@@
   =
-  -\frac{F_e(s)}{\varepsilon}\beta
+  -\beta {F_e(s)}
   j(\vect x)
@@
   j(\vect x) =
-  -\vect C\cdot\nabla_{\vect q}\mh 
   -\vect D\cdot\nabla_{\vect p}\mh
@@
   =
-  -\frac\beta\varepsilon
+  -\beta
   \int_0^t\dd s\,
@@
   f_{\inv}^\beta(\vect x)
-  - \beta
+  - \eps\beta
   \int_0^t\dd s\,
@@
   \big]
-  + \mo(\varepsilon^2)
+  + \mo(\eps^2)
 \end{align}
-For any time-dependent observation, the linear response approximation
-would be:
+Now let $\mathcal{O}$ be an observable of the system. Its expectation value satisfies the linear response formula:
 \begin{align}\nonumber
@@
   \int\dd \vect x\, O(\vect x)\,
-  [\,f_{\inv}^\beta(\vect x) + \varepsilon f_1(\vect x, t)] \\\nonumber
+  [\,f_{\inv}^\beta(\vect x) + \eps f_1(\vect x, t)] +\mathcal{O}(\eps^2)\\\nonumber
+   \label{eqn:eqi-pert-1}
   &=
@@
   -
-  \beta
-  \int \dd \vect x\:
-  O(\vect x)
-  \int_0^t\dd s\,F_e(s)
-  e^{(t-s)\fwg_0^\beta}
-  \big[j(\vect x)f^\beta_{\inv}(\vect x)\big]
-  \\\nonumber
-  &=
-  \langle O\rangle
-  -
-  \beta
-  \int_0^t\dd s\,
-  F_e(s)
-  \int \dd \vect x\,
-  O(\vect x)\,
-  e^{(t-s)\fwg_0^\beta}
-  \big[j(\vect x)f^\beta_{\inv}(\vect x)\big]
-  \\\nonumber
-  &=
-  \langle O\rangle
-  -
-  \beta
-  \int_0^t\dd s\,
-  F_e(s)
-  \int \dd \vect x\,
-  f^\beta_{\inv}(\vect x) j(\vect x)\,
-  e^{(t-s)\bwg^\beta_0}
-  O(\vect x)\\ \label{eqn:eqi-pert-1}
-  &=
-  \langle O\rangle
-  -
-  \beta
+ \eps \beta
   \int_0^t\dd s\,
@@
   e^{s\bwg_0^\beta}
-  O(\vect x)
+  O(\vect x)+\mathcal{O}(\eps^2)
 \end{align}
-Where $\bwg^\beta_0$ is the infiniesimal generator of the \emph{un-driven} 
-Kolmogorov backward equation, which is the adjoint operator of $\fwg^\beta_0$.
-We have
+where $ \langle O\rangle=\int \dd\vect x\,O(\vect x) f_{\inv}^\beta(\vect x)$ is the expectation value of the undriven equilibrium system, and $\bwg^\beta_0$ is the infiniesimal generator of the undriven
+Kolmogorov backward equation, which is the adjoint operator of $\fwg^\beta_0$. That is,
+we have
 \begin{align}
@@
 \end{align}
-where $\vect X^\beta_t$ is the trajectory of the un-driven Langevin equation
+where $\vect X^\beta_t$ is the trajectory of the undriven Langevin equation (with $\eps=0$)
 starting at $\vect x$.
-Therefore, we have
+Therefore, we have up to order 2 in $\eps$:
 % Notice that $e^{i\ml_0(t-s)}O(\vect x) = O(\phi_{t-s}(\vect x))$,
@@
 % Hamiltonian system. We have
-\begin{align}
+\begin{align}\nonumber
   \mathcal O(t)
-  =
+ & =
   \langle O\rangle
   -
-  \beta
+  \eps\beta
   \int_0^t\dd s\,
@@
   j(\vect x)\,
-  \mathbb E_{\vect x} [O(\vect X^\beta_s)]
-\end{align}
-or equivalently:
-\begin{align}\label{eqn:lr}
-  \mathcal O(t)
-  =
+  \mathbb E_{\vect x} [O(\vect X^\beta_s)]\\
+\label{eqn:eq-lr}  
+ & =
   \langle O\rangle
   -
-  \beta
+  \eps \beta
   \int_0^t\dd s\,
@@
   \big\langle
-  j(0)\,
-  \mathbb E_{\vect x} [O(\vect X^\beta_s)]
+  j\cdot
+  \mathbb E [O(\vect X^\beta_s)]
   \big\rangle
@@
 where $  \langle
-  j(0)\,
-  \mathbb E_{\vect x} [O(\vect X^\beta_s)]
+  j\cdot
+  \mathbb E [O(\vect X^\beta_s)]
   \rangle$ is the \emph{equilibrium time
-  correlation} between the observation $O$ at time $s$ and the
-dissipative flux $j$.
+  correlation} between the observation $O$ at time $s$ and the flux $j$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 Firstly, one has to start from an equilibrium simulation. That means if the
-external driving is no longer small ($\varepsilon$ is relatively big), the
-first order (or linear) response~\eqref{eqn:lr} may not be a good approximation to
+external driving is no longer small ($\eps$ is relatively big), the
+first order (or linear) response~\eqref{eqn:eq-lr} may not be a good approximation to
 the true non-equilibrium process any more.
 One possible way to improve the accuracy is to introduce the second
-order response term, i.e.~\eqref{eqn:f2-0}.
+order response term, e.g., based on equation~\eqref{eqn:f2-0}.
 Here one encounters the second limitation of the classical
@@
 The limitations of the classical responses theory motivate us to
-develop a more general frame work that allows one starting from any
-non-equilibrium process (arbitrarily driven), and calculating from it as many orders of
-responses as necessary.
+develop a more general framework that allows one starting from any
+non-equilibrium process (arbitrarily driven), and calculating linear responses from it.
 
-In the language of path integral, the time-dependent average $O( t)$
+This time, we consider the following Langevin equation:
+\begin{align}\label{eqn:langevin-1b}
+  \dot{\vect q} & = \nabla_{\vect p}\mh(\vect q,\vect p), \\\label{eqn:langevin-2b}
+  \dot{\vect p} & =- \nabla_{\vect q}\mh(\vect q,\vect p)
+  + (k_0 F_e(t)+\eps \Delta F_e(t)) \vect D(\vect q)
+  - \gamma\vect p
+  + \sigma\dot{\vect W},
+\end{align}
+such that for $\eps=0$ we get a driven system that is further
+perturbed by the addition time-dependent driving force $\eps \Delta
+F_e(t) \vect D(\vect q)$. Let $\vect X^\eps_t$ denote the associated solution process, and
+$\mathcal P_\eps$ the probability measure of the trajectories
+generated by driven Langevin
+dynamics~\eqref{eqn:langevin-1b}--\eqref{eqn:langevin-2b}.
+
+The time-dependent expectation value $O_0(t)$ of the path observable $O$ under the dynamics of the Langevin equation (\ref{eqn:langevin-1b})-(\ref{eqn:langevin-2b}) for $\eps=0$
 can be written as
 \begin{align}
-  O( t) =
-  \langle O[\vect X_s]\,\rangle_{0,t} = 
+  O_0( t) =
+  \langle O[\vect X^0_s]\,\rangle_{0,t} = 
   \int\dd \vect u\, f(\vect u, 0) 
   \int_{\mc\{\vect u,0; t\}}
-  O[\vect X_s]\, \dd\mathcal P[\vect X_s]  
+  O[\vect X_s]\, \dd\mathcal P_0[\vect X_s],
 \end{align}
-where $O[\vect X_s]$ is the observable of the trajectories, which is a generalization of the
-observable that only depends on the configuration of the system.
-$\langle \cdot \,\rangle_{0,t}$ is the ensemble average over all possible trajectories in time interval $[0,t]$.
-$f(\vect x, 0)$ is the initial distribution that may not be the equilibrium distribution.
-$\mc\{\vect u,0; t\}$ is the set of all continuous trajectories starting from point
+where 
+$\langle \cdot \,\rangle_{0,t}$ is the ensemble average over all possible trajectories for $\eps=0$ in time interval $[0,t]$,
+$f(\vect x, 0)$ is the initial distribution that may not be the equilibrium distribution, and
+$\mc\{\vect u,0; t\}$ is the set of all admissible continuous trajectories starting from point
 $\vect u$ at time $0$, and ending at some position at time $t$.
-$\mathcal P[\vect X_s]$ is the
-probability measure of the trajectories generated by driven Langevin dynamics~\eqref{eqn:langevin-1}--\eqref{eqn:langevin-2},
-where the intensity $F_e$ need not to be 0.
-Now we want to consider a system that is a small perturbation of the previous system (we now call it ``reference system'')
-in the sense that it has a similar driven force: $F_e + \Delta F_e$, where $\Delta F_e\sim \mo(\varepsilon)$ is small.
-The probability measure of the trajectories of this perturbed system is denoted by
-$\tilde{\mathcal P}[\vect X_s]$.
-The time-dependent average of observable of the system $F_e + \Delta F_e$ can be written down in terms of
-the probability measure of the reference system:
+
+For $\eps\ge 0$ the  expectation value $O_\eps(t)$ takes the form
 \begin{align}
-  \tilde O(t) =
+  O_\eps(t) & =
   \int \dd\vect u f(\vect x,0)
   \int_{\mc\{\vect u,0; t\}} O[\vect X_s]\,
-  \frac{\dd\tilde{\mathcal P}[\vect X_s]}{\dd\mathcal P[\vect X_s]}
-  \, \dd\mathcal P[\vect X_s]    
+  \, \dd\mathcal P_\eps[\vect X_s]   \\
+& = \int \dd\vect u f(\vect x,0)
+  \int_{\mc\{\vect u,0; t\}} O[\vect X_s]\,
+  \frac{\dd {\mathcal P}_\eps[\vect X_s]}{\dd\mathcal P_0[\vect X_s]}
+  \, \dd\mathcal P_0[\vect X_s] .   
 \end{align}
-by using the Girsanov formula, 
+By using the Girsanov formula, 
 \begin{align}\label{eqn:girsanov}
   \frac
-  {\dd\tilde{\mathcal P}[\vect X_s]}
-  {\dd\mathcal P[\vect X_s]}  
+  {\dd{\mathcal P_\eps}[\vect X_s]}
+  {\dd\mathcal P_0[\vect X_s]}  
   =&\,
   \exp
-  \big\{{G[\vect X_s] - \frac12 H[\vect X_s]}
+  \big\{{\eps G[\vect X_s] - \eps^2 H[\vect X_s]}
   \big\}
@@
   &= \frac{1}\sigma\int_0^t
-  \Delta F_e(s)\,\vect D(\vect q_s,\vect p_s)\cdot\dd\vect W_s \\
+  \Delta F_e(s)\,\vect D(\vect q_s)\cdot\dd\vect W_s \\
   H[\vect X_s]
@@
   \frac 1{2\sigma^2}\int_0^t
-  \Delta F_e^2(s)\,\vect D^2(\vect q_s,\vect p_s)\,\dd s
+  \Delta F_e^2(s)\,\vect D^2(\vect q_s)\,\dd s.
 \end{align}
-If the difference $\Delta F_e(t)$ is of order $\mo (\varepsilon)$, then $G[\vect X_s] \sim \varepsilon$ is
-the first order term, and $H[\vect X_s] \sim \varepsilon^2$  is the second order term.
-Then with the Taylor expansion of the right-hand-side of Eq.~\eqref{eqn:girsanov}, we have
+Expansion of the right-hand-side of Eq.~\eqref{eqn:girsanov} in $\eps$, we have
 \begin{align}\label{eqn:neq-response}
-  \tilde O(t) -
-  O(t)
+   O_\eps(t) -
+  O_0(t)
   =
-  \big\langle O[\vect X_s]\cdot G[\vect X_s]\big\rangle_{0,t}
+  \eps\, \big\langle O[\vect X_s]\cdot G[\vect X_s]\big\rangle_{0,t}
   +
-  \frac12 \big\langle O[\vect X_s] \cdot (G^2[\vect X_s] - H[\vect X_s] )\big\rangle_{0,t}
+  \frac{\eps^2}{2} \big\langle O[\vect X_s] \cdot (G^2[\vect X_s] - H[\vect X_s] )\big\rangle_{0,t}
   +
-  \mo (\varepsilon^3)
+  \mo (\eps^3)
 \end{align}
-Here are some remarks of the main result Eq.~\eqref{eqn:neq-response}:
+Some remarks concerning Eq.~\eqref{eqn:neq-response} seem in order:
 \begin{itemize}
 \item The first term on the right-hand-side of Eq.~\eqref{eqn:neq-response} is the linear
-  response of the reference non-equilibrium process (driven by $F_e$). The second term is the second order
-  response. One can easily write down higher orders of response by the Taylor expansion
-  of the right-hand-side of Eq.~\eqref{eqn:girsanov}.
+  response to the reference non-equilibrium process (driven by $F_e$, $\eps=0$).
+The next term is the second order response term.
+Obviously, one can easily write down higher orders of response by the Taylor expansion of the right-hand-side of Eq.~\eqref{eqn:girsanov}.
 \item When the observable $O[\vect X_s]$ only depends on the end point of the trajectory, i.e.
   $O[\vect X_s] = O(\vect X_t)$, then $\big\langle O(\vect X_t)\cdot G[\vect X_s]\big\rangle_{0,t}$
-  gives the well-know time correlation form of the linear response.
-\item If one wants to calculate a large amount of different perturbations, then apply one by one
-  Eq.~\eqref{eqn:neq-response} is not convenient, because one has to calculate the responses
-  again when the function form of $\Delta F_e$ is not a linear combination of previously calculated perturbations.
+  gives the well-know time correlation form of the linear response formula.
+\item If one wants to calculate the same expectation value for a family of non-equilibrium perturbations $\Delta F_e(t)$ then 
+one does not need to repeat the calculations of Eq.~\eqref{eqn:neq-response} for every member of the family. 
   If it is possible to
-  express different $\Delta F_e$ by a group of base functions, then the responses are only need
-  to calculated on the bases. Then with a linear conbination of the responses on bases, one can derive
-  the responses of the whole group of perturbations. 
-\item In the prove, we assume that the reference and the perturbed non-equilibrium processes
+  express different $\Delta F_e$ in the same basis, then the responses must only be
+  calculated for the single basis functions. Then with a linear combination of the responses on basis functions, one can derive
+  the responses of the whole family. 
+\item In the above derivation, we assume that the reference and the perturbed non-equilibrium processes
   have the same initial distribution (both of them may be different from the equilibrium distribution),
   but this is not a necessary. One can easily
-  applies similar perturbation approach to the initial distribution as  to the trajectory ensemble.
-  To keep the story simple, we do not develop the theory from this aspect.
+  apply a similar reweighing approach to the initial distribution as we used it for the trajectory ensemble.
 \end{itemize}
 
-\section{Numerical tests}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Numerical tests: the core set identification of  the Markov State Model }
+
+The core set plays a very important role in the MSM theory.  We
+therefore want to test our idea of non-equilibrium response theory by
+identifying the core sets in a non-equilibrium driven system.
+There are several ways to define a core set.
+One possible choice  is to calculate the stability of the phase space
+distribution: $\Delta \mt_\tau f(\vect x, t) = (\mt_\tau -\mi) f(\vect
+x, t)$. And the core sets are defined by
+\begin{align}
+  C_\tau(t) = \{\vect x\vert \Delta \mt_\tau f(\vect x, t) \geq 0\}
+\end{align}
+Here the propagator $\mt_\tau$ is defined by $\mt_\tau =
+e^{\tau\fwg_t^{2\beta}}$.  Notice that here the physical time $t$ is
+frozen, and the distribution $f$ evolves a long the artificial time
+$\tau$ at temperature $1/(2\beta)$. The artificial time $\tau$ should
+be chosen smaller than the investigated relaxation time scales of the
+original system. However, on the other hand, $\tau$ cannot be
+arbitrarily small, because otherwise the statistical uncertainty would
+not allow us to distinguish $\mt_\tau f$ and $f$.
 
@@
   \includegraphics[width=0.4\textwidth]{figs/fig-tilt-pot.eps}
-  \caption{The double-well potential with tilting perturbation.}
+  \caption{The double-well potential with tilting perturbation at
+    time $t\geq 20$ ps..}
   \label{fig:tmp1}
@@
   {figs/fig-tilt-str2-2d-more.eps}
-  \caption{The plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
-    $F_e^{\textrm{max}} = 2$.
-    The 1st row: the direct non-equilibrium
-    simulation. The 2nd row: the results of linear response
-    formula~\eqref{eqn:core-identify-approx-2}.
-    The 3rd row: the results of response
-    formula~\eqref{eqn:pert-approx-1} and \eqref{eqn:pert-approx-2}.
-    The order is 2, the reference state is $F_e^{\textrm{max}} = 0$, i.e.
-    the equilibrium state.
+  \caption{The plot of $\Delta\mt_\tau f(\vect x,t)$ under
+    perturbation with strength $\eps = 2$.  From left to right the
+    columns present results at time $t = 0$, 10, 15 and
+    20~\textsf{ps}.  The 1st row: the direct non-equilibrium
+    simulation. The 2nd row: the results of traditional equilibrium
+    linear response theory.  The 3rd row: the results of the response
+    theory developed by this work.  We calculate response terms up to
+    the second order from an equilibrium simulation (i.e. letting $k_0
+    = 0$ and $\eps = 0$ in Eq.~\eqref{eqn:langevin-2b}).
   }
@@
   \caption{The plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
-    $F_e^{\textrm{max}} = 4$.
-    The 1st row: the direct non-equilibrium
-    simulation. The 2nd row: the results of linear response
-    formula~\eqref{eqn:core-identify-approx-2}.
-    The 3rd row: the results of response
-    formula~\eqref{eqn:pert-approx-1} and \eqref{eqn:pert-approx-2}.
-    The order is 2, the reference state is $F_e^{\textrm{max}} = 0$, i.e.
-    the equilibrium state.
-    The 4th row: the results of response
-    formula~\eqref{eqn:pert-approx-1} and \eqref{eqn:pert-approx-2}.
-    The order is 1, the reference state is $F_e^{\textrm{max}} = 2$.
+    $\eps = 4$. From left to right the
+    columns present results at time $t = 0$, 10, 15 and
+    20~\textsf{ps}.
+    The first three rows use the same algorithm as Fig.~\ref{fig:tmp2}.    
+    The 4th row: the results of the response theory developed by this work.
+    Instead of calculating the responses from the equilibrium simulation,
+    here we calculate the linear response from a non-equilibrium simulation
+    of strength 2 (i.e. $k_0 = 2$ and $\eps = 4$
+    in Eq.~\eqref{eqn:langevin-2b}).
+    Therefore, the effective perturbation in this case is only
+    $\eps = 2$.
   }
@@
 tilting double-well potential. For convenience, we let the mass of the
-particle to be 1 \textsf{amu}. The unperturbed
+particle to be 1~\textsf{amu},
+and the friction coefficient to be $1\,\textsf{ps}^{-1}$.
+The unperturbed
 Hamiltonian of the system is given by:
@@
 Here $k = 8$~$\textsf{kJ} / (\textsf{mol nm}^4)$, and $ a = 1\ \textsf{nm}$.
-Notice at room temperature $300\ \textsf{K}$, $k_BT = 2.48$~\textsf{kJ/mol}.
-See the red line in Fig.~\ref{fig:tmp1}.
+At room temperature $300\ \textsf{K}$, $k_BT = 2.48$~\textsf{kJ/mol}.
+The red line in Fig.~\ref{fig:tmp1} presents the unperturbed potential.
 %$\textsf{kJ} / (\textsf{mol nm}^4)}$
@@
 \begin{align}
-  \vect C(\vect p, \vect q) = 0; \qquad
-  \vect D(\vect p, \vect q) = -\nabla_{\vect q} V(\vect q) = 1
+  \vect D(\vect q) = -\nabla_{\vect q} V(\vect q) = 1
 \end{align}
-Here $V(\vect q) = -\vect q$ is  effectively tilt the original
-potential $U(\vect q)$. The strength of the perturbation $F_e(t)$ is given
-by the following function:
-\begin{align}
-  F_e(t) = 
-  \begin{cases}
-    F_e^{\textrm{max}}\times (t / t_c) & t < t_c \\
-    F_e^{\textrm{max}} & t \geq t_c
-  \end{cases}
-\end{align}
-$t_c = 20$~\textsf{ps} so that the perturbation increases slow
+Here $V(\vect q) = -\vect q$   effectively tilts the original
+potential $U(\vect q)$ (see the gree and blue lines in Fig.~\ref{fig:tmp1}).
+The strength of $F_e(t)$ is
+set to be 1, when the time $t$ is larger than some warm-up time $t_c$.
+When $t \leq t_c$, the strength of $F_e(t)$ is linearly growing,
+i.e. $ F_e(t) = t/t_c$.
+We use the value $t_c = 20$~\textsf{ps}, so that the perturbation increases slow
 enough: the typical decaying time scale of the correlation function
-in~\eqref{eqn:core-identify-approx-2} is roughly 3~\textsf{ps}.
-We choose $\tau = 1$~\textsf{ps}.
+in~\eqref{eqn:eq-lr} is roughly 3~\textsf{ps}.
+We choose $\tau = 1$~\textsf{ps}, and
 % We choose $\mt_\alpha$
@@
 % \textsf{K} with $\alpha = 1$ \textsf{ps}.
-Since the dynamics is in
-1-d, the projection $\mathcal P$ is identity.  We consider the stability of the
+consider the stability of the
 distribution $f(\vect x, t)$ at time $t$: $\Delta\mt_\tau f(\vect x, t)$.
@@
 
-See Fig. \ref{fig:tmp2} and \ref{fig:tmp3} for numerical result.  Both
-the direct non-equilibrium simulation and the response
-approximations are presented. Time slices $t = 0$, 10, 15 and 20
-\textsf{ps} are shown.  When the perturbation is small,
-i.e. $F_e^{\textrm{max}} = 2$, the direct non-equilibrium simulation
-and the linear response results are accurately consistent.
-The 2nd order response formula~\eqref{eqn:pert-approx-1} and
-\eqref{eqn:pert-approx-2} are even better, but the numerical error
-is bigger.
-As the
+See Fig.~\ref{fig:tmp2} and \ref{fig:tmp3} for numerical result.  Both
+the results of the brute-force non-equilibrium simulation
+and that of the response theory are presented.
+In Fig.~\ref{fig:tmp2},
+the good agreement between the brute-force (1st row)
+and the linear response result (2nd row) indicates that $\eps = 2$ is
+indeed a relatively small perturbation to the equilibrium.
+In the third row of Fig.~\ref{fig:tmp2},
+the second order response is considered. We do not see any
+substantial improvement of the result by including this higher order term,
+because the
+driving force to the system is small,
+and the linear term is anyway a good approximation.
+
+In Fig.~\ref{fig:tmp3}, as the
 perturbation grows stronger, the left well becomes less stable, while
-the right well becomes more stable.  When the perturbation is big,
-i.e. $F_e^{\textrm{max}} = 4$, the linear response result is no
-longer precise.
-When $t\geq 20$ \textsf{ps}, the left well basically
+the right well becomes more stable.  When the perturbation is as strong as
+$\eps = 4$, the linear response result (2nd row) is no
+longer precise:
+At $t\geq 20$ \textsf{ps}, the left well basically
 disappears, however, the linear response calculation presents
-artifical occilation at the right well.
-The 2nd order response also has artifical effect, and the numerical
-uncertainty is high.
-The 4th rwo of Fig.~\ref{fig:tmp3} presents the 1st order result of the
-response formula~\eqref{eqn:pert-approx-1} and
-\eqref{eqn:pert-approx-2}, with
-respect to the reference state $F_e^{\textrm{max}} = 2$. The result
-is impressively improved compared with the linear response with respect to the
-equilibrium state.
+artificial structure (that is a small artificial
+stable region accompanied by a small unstable region)
+at the position of the left well.
+The second order response (3rd row) improves the accuracy a little, 
+however, it also includes artificial finer-grained oscillation in the profile.
+At the same time, the numerical uncertainty is higher, so the figure
+looks noisy.
+In the 4th row of Fig.~\ref{fig:tmp3}, we present the linear response result
+that calculate from 
+a intermediate
+non-equilibrium simulation with driving force of $k_0 = 2$.
+Notice that the reference simulation in this case is also a
+non-equilibrium simulation, so the effective strength of perturbation
+is reduced to $\eps = 2$. According to Fig.~\ref{fig:tmp1}, this can be
+considered as a small perturbation.
+A substantial improvement is achieved: the profile
+is correctly calculated, and the statistical error is small.
 
-\subsection{Splitting single-well potential}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.4\textwidth]{figs/fig-split-pot.eps}
-  \caption{The single-well potential with splitting perturbation.}
-  \label{fig:tmp4}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.95\textwidth]
-  {figs/fig-split-str1-2d-more.eps}
-  \caption{The plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
-    $F_e^{\textrm{max}} = 1.0$.
-    The 1st row: the direct non-equilibrium
-    simulation. The 2nd row: the results of linear response
-    formula~\eqref{eqn:core-identify-approx-2}.
-    The 3rd row: the results of response
-    formula~\eqref{eqn:pert-approx-1} and \eqref{eqn:pert-approx-2}.
-    The order is 2, the reference state is $F_e^{\textrm{max}} = 0$, i.e.
-    the equilibrium state.
-  }
-  \label{fig:tmp5}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.95\textwidth]
-  {figs/fig-split-str2-2d-more.eps}
-  \caption{The plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
-    $F_e^{\textrm{max}} = 2.0$.
-    The 1st row: the direct non-equilibrium
-    simulation. The 2nd row: the results of linear response
-    formula~\eqref{eqn:core-identify-approx-2}.
-    The 3rd row: the results of response
-    formula~\eqref{eqn:pert-approx-1} and \eqref{eqn:pert-approx-2}.
-    The order is 2, the reference state is $F_e^{\textrm{max}} = 0.0$, i.e.
-    the equilibrium state.
-    The 4th row: the results of response
-    formula~\eqref{eqn:pert-approx-1} and \eqref{eqn:pert-approx-2}.
-    The order is 1, the reference state is $F_e^{\textrm{max}} = 1.0$.
-  }
-  \label{fig:tmp6}
-\end{figure}
-
-In this splitting single-well potential example, the unperturbed
-potential is simply the harmonic potential given by:
-\begin{align}
-  U(\vect q) = \frac12\,k\,\vect q^2 
-\end{align}
-Here $k = 8$~$\textsf{kJ} / (\textsf{mol nm}^2)$.
-Notice at room temperature $300\ \textsf{K}$, $k_BT = 2.48$~\textsf{kJ/mol}.
-See the red line in Fig.~\ref{fig:tmp4}.
-%$\textsf{kJ} / (\textsf{mol nm}^4)}$
-The perturbation is again given by
-\begin{align}
-  \vect C(\vect p, \vect q) = 0; \qquad
-  \vect D(\vect p, \vect q) = -\nabla_{\vect q} V(\vect q) 
-\end{align}
-where $V(\vect q)$ is a Gaussian type potential:
-\begin{align}
-  V(\vect q) = \frac{1}{\sqrt{2\pi \sigma^2}}
-  \exp\Big\{-\frac{\vect q^2}{2\sigma^2}\Big\}
-\end{align}
-we use $\sigma = 0.16$~\textsf{nm}. Please see Fig~\ref{fig:tmp4}.
-For the results, please see Fig.~\ref{fig:tmp5} -- Fig.~\ref{fig:tmp6}.
-
-\newpage
-\section{Numerical results}
-\subsection{Tilting double-well potential}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.4\textwidth]{figs/fig-tilt-pot.eps}
-  \includegraphics[width=0.4\textwidth]{figs/fig-split-pot.eps}
-  \caption{Two testing cases presented by this paper.
-    Left: the double-well potential with tilting perturbation.
-    Right: the single-well potential with splitting perturbation.}
-  \label{fig:new-tmp1}
-\end{figure}
-
-We test the idea by a one-dimensional model system: one particle in a
-tilting double-well potential. For convenience, we let the mass of the
-particle to be 1 \textsf{amu}. The unperturbed
-Hamiltonian of the system is given by:
-\begin{align}
-  \mh (\vect p, \vect q) = \frac 12 \vect p^2 + U(\vect q) 
-\end{align}
-with potential
-\begin{align}
-  U(\vect q) = \frac12 k (\vect q^2 - a^2)^2
-\end{align}
-Here $k = 8$~$\textsf{kJ} / (\textsf{mol nm}^4)$, and $ a = 1\ \textsf{nm}$.
-Notice at room temperature $300\ \textsf{K}$, $k_BT = 2.48$~\textsf{kJ/mol}.
-See the red line in Fig.~\ref{fig:tmp1}.
-%$\textsf{kJ} / (\textsf{mol nm}^4)}$
-The perturbation is given by
-\begin{align}
-  \vect C(\vect p, \vect q) = 0; \qquad
-  \vect D(\vect p, \vect q) = -\nabla_{\vect q} V(\vect q) = 1
-\end{align}
-Here $V(\vect q) = -\vect q$ is  effectively tilting the original
-potential $U(\vect q)$. The strength of the perturbation $F_e(t)$ is given
-by the following function:
-\begin{align}
-  F_e(t) = 
-  \begin{cases}
-    F_e^{\textrm{max}}\times (t / t_c) & t < t_c \\
-    F_e^{\textrm{max}} & t \geq t_c
-  \end{cases}
-\end{align}
-$t_c = 20$~\textsf{ps} so that the perturbation increases slow
-enough: the typical decaying time scale of the correlation function
-in~\eqref{eqn:core-identify-approx-2} is roughly 3~\textsf{ps}.
-We choose $\tau = 1$~\textsf{ps}.
-% We choose $\mt_\alpha$
-% to be the propagator of the Langevin dynamics at temperature $150$
-% \textsf{K} with $\alpha = 1$ \textsf{ps}.
-Since the dynamics is in
-1-d, the projection $\mathcal P$ is identity.  We consider the stability of the
-distribution $f(\vect x, t)$ at time $t$: $\Delta\mt_\tau f(\vect x, t)$.
-The larger
-this value, the more stable the core sets are.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.95\textwidth]
-  {figs/fig-tilt-str2-simple.eps}
-  \caption{
-    Tilting double well potential testing case:
-    the plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
-    $F_e^{\textrm{max}} = 2.0$.
-    From left to right:
-    (a) $t=0$~\textsf{ps},
-    the brutal force non-equilibirum simulation,
-    (b) $t=20$~\textsf{ps},
-    the brutal force non-equilibirum simulation,
-    (c) $t=20$~\textsf{ps},
-    response formula truncated at order $n=1$,
-    and
-    (d) $t=20$~\textsf{ps},
-    response formula truncated at order $n=2$.
-    The equilibirum state serves as the reference of
-    the response simulations.
-  }
-  \label{fig:new-tmp2}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.95\textwidth]
-  {figs/fig-tilt-str4-simple.eps}
-  \caption{
-    Tilting double well potential:
-    the plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
-    $F_e^{\textrm{max}} = 4.0$.
-    From left to right:
-    (a) $t=0$~\textsf{ps},
-    the brutal force non-equilibirum simulation,
-    (b) $t=20$~\textsf{ps},
-    the brutal force non-equilibirum simulation,
-    (c) $t=20$~\textsf{ps},
-    response formula with reference process $F_e^{\max} = 0.0$
-    (i.e. equilibirum simulation),
-    and
-    (d) $t=20$~\textsf{ps},
-    response formula with reference process $F_e^{\max} = 2.0$.
-    The response formula are truncated at order $n=1$.
-  }
-  \label{fig:new-tmp3}
-\end{figure}
-
-The Fig.~\ref{fig:new-tmp2} and \ref{fig:new-tmp3} present
-the application of the response theory to
-calculate core sets of the tilting double-well potential.
-Fig.~\ref{fig:new-tmp2} presents
-the snapshot at $t=20$~\textsf{ps}
-of $\Delta\mt_\tau f(\vect x,t)$ under the maximum perturbation
-strength $F_e^{\max} = 2.0$. Response formula truncated
-at both order $n=1$ and $n=2$ are compared to 
-the brutal force non-equilibirum simulation.
-No obviously difference is shown. The second order response result
-is slightly more precise than the first order one, but the
-statistical error is also larger, which makes the profile noisy.
-This figure tells us that
-(1) the modeling error of the linear response (1st order) is
-negligibly small; (2) the statistical error of numerically measuring the
-linear response is also small.
-Fig.~\ref{fig:new-tmp3} presents the $t=20$~\textsf{ps}
-snapshots of $\Delta\mt_\tau f(\vect x,t)$
-under a larger perturbations $F_e^{\max} = 4.0$.
-The linear reponse using the equilibirum process as the reference
-shows poor accuracy: the left well vanishes in this case,
-while the linear response produces  artifical structures.
-Since the profile is smooth, which indicates
-the statistical error of calculating the linear term is small,
-the deviation comes from the modellng error, namely the contribution
-of higher order response terms.
-By including the second order response term reduces the left
-artifical core set, but also creates artifical finer grained
-structures. The statistical accuracy is also worse than the linear response.
-Plot (d) present the result of the linear response
-formula using reference process $F_e^{\max} = 2.0$, which
-is the process presented in Fig.~\ref{fig:new-tmp2}.
-Much better modelling accuracy is achieved:
-the left artifical core set disappear and the
-statistical accuracy is also satisfactory.
-
-% See Fig. \ref{fig:tmp2} and \ref{fig:tmp3} for numerical result.  Both
-% the direct non-equilibrium simulation and the response
-% approximations are presented. Time slices $t = 0$, 10, 15 and 20
-% \textsf{ps} are shown.  When the perturbation is small,
-% i.e. $F_e^{\textrm{max}} = 2$, the direct non-equilibrium simulation
-% and the linear response results are accurately consistent.
-% The 2nd order response formula~\eqref{eqn:pert-approx-1} and
-% \eqref{eqn:pert-approx-2} are even better, but the numerical error
-% is bigger.
-% As the
-% perturbation grows stronger, the left well becomes less stable, while
-% the right well becomes more stable.  When the perturbation is big,
-% i.e. $F_e^{\textrm{max}} = 4$, the linear response result is no
-% longer precise.
-% When $t\geq 20$ \textsf{ps}, the left well basically
-% disappears, however, the linear response calculation presents
-% artifical occilation at the right well.
-% The 2nd order response also has artifical effect, and the numerical
-% uncertainty is high.
-% The 4th rwo of Fig.~\ref{fig:tmp3} presents the 1st order result of the
+% the 1st order result of the
 % response formula~\eqref{eqn:pert-approx-1} and
@@
 
-In this splitting single-well potential example, the unperturbed
+\begin{figure}
+  \centering
+  \includegraphics[width=0.4\textwidth]{figs/fig-split-pot.eps}
+  \caption{The single-well potential with splitting perturbation at
+    time $t\geq 20$ ps.}
+  \label{fig:tmp4}
+\end{figure}
+
+% \begin{figure}
+%   \centering
+%   \includegraphics[width=0.95\textwidth]
+%   {figs/fig-split-str1-2d-more.eps}
+%   \caption{The plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
+%     $F_e^{\textrm{max}} = 1.0$.
+%     The 1st row: the direct non-equilibrium
+%     simulation. The 2nd row: the results of linear response
+%     formula~\eqref{eqn:core-identify-approx-2}.
+%     The 3rd row: the results of response
+%     formula~\eqref{eqn:pert-approx-1} and \eqref{eqn:pert-approx-2}.
+%     The order is 2, the reference state is $F_e^{\textrm{max}} = 0$, i.e.
+%     the equilibrium state.
+%   }
+%   \label{fig:tmp5}
+% \end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.95\textwidth]
+  {figs/fig-split-str2-2d-more.eps}
+  \caption{
+    The plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
+    $\eps = 2$. From left to right the
+    columns present results at time $t = 0$, 5, 10 and
+    20~\textsf{ps}.
+    The 1st row: the direct non-equilibrium
+    simulation. The 2nd row: the results of traditional equilibrium
+    linear response theory.  The 3rd row: the results of the response
+    theory developed by this work.  We calculate response terms up to
+    the second order from an equilibrium simulation (i.e. letting $k_0
+    = 0$ and $\eps = 2$ in Eq.~\eqref{eqn:langevin-2b}).
+    The 4th row: the results of the response theory developed by this work.
+    Instead of calculating the responses from the equilibrium simulation,
+    here we calculate the linear response from a non-equilibrium simulation
+    of strength 2 (i.e. $k_0 = 1$ and $\eps = 1$
+    in Eq.~\eqref{eqn:langevin-2b}).
+    Only the linear term of response is calculated.
+    Notice, the effective perturbation in this case is only
+    $\eps = 1$.
+  }
+  \label{fig:tmp6}
+\end{figure}
+
+In the previous subsection, we considered an example that the stability of the
+core sets is switched by an external driving to the system.
+In this section, we consider ``splitting single-well potential'' example
+that two new core sets grows out of 
+an old core sets by the external driving force.
+In this example, the unperturbed
 potential is simply the harmonic potential given by:
@@
 Here $k = 8$~$\textsf{kJ} / (\textsf{mol nm}^2)$.
-Notice at room temperature $300\ \textsf{K}$, $k_BT = 2.48$~\textsf{kJ/mol}.
-See the red line in Fig.~\ref{fig:tmp4}.
+See the red line in Fig.~\ref{fig:tmp4} for the unperturbed potential.
 %$\textsf{kJ} / (\textsf{mol nm}^4)}$
-The perturbation is again given by
+The perturbation has again given by
 \begin{align}
-  \vect C(\vect p, \vect q) = 0; \qquad
-  \vect D(\vect p, \vect q) = -\nabla_{\vect q} V(\vect q) 
+  \vect D(\vect q) = -\nabla_{\vect q} V(\vect q) 
 \end{align}
-where $V(\vect q)$ is a Gaussian type potential:
+where $V(\vect q)$ has a Gaussian profile:
 \begin{align}
@@
 \end{align}
-we use $\sigma = 0.16$~\textsf{nm}. Please see Fig~\ref{fig:tmp4}.
-For the results, please see Fig.~\ref{fig:tmp5} -- Fig.~\ref{fig:tmp6}.
+we use $\sigma = 0.16$~\textsf{nm}.
 
-\begin{figure}
-  \centering
-  \includegraphics[width=0.95\textwidth]
-  {figs/fig-split-str1-simple.eps}
-  \caption{
-    Splitting single-well potential testing case:
-    the plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
-    $F_e^{\textrm{max}} = 1.0$.
-    From left to right:
-    (a) $t=0$~\textsf{ps},
-    the brutal force non-equilibirum simulation,
-    (b) $t=20$~\textsf{ps},
-    the brutal force non-equilibirum simulation,
-    (c) $t=20$~\textsf{ps},
-    response formula truncated at order $n=1$,
-    and
-    (d) $t=20$~\textsf{ps},
-    response formula truncated at order $n=2$.
-    The equilibirum state serves as the reference of
-    the response simulations.
-  }
-  \label{fig:new-tmp5}
-\end{figure}
+In this example, we also consider two different strengths of
+perturbations. Please see the green and blue lines in Fig~\ref{fig:tmp4} for
+perturbations of $\eps = 1$ and $\eps = 2$, respectively.
+Numerical results show that the perturbation $\eps=1$ can be considered to be
+relatively small. Since the phenomenon is basically the same as
+Fig.~\ref{fig:tmp2}, we do not present them here.
+Fig.~\ref{fig:tmp6} presents the numerical results of perturbation $\eps =
+2$.  Only with a linear response calculated from the equilibrium
+simulation (2nd row), the stability of the two new core sets are too
+strong, and the unstable region between them is over estimated, comparing with
+the brute force simulation presented on the first row.
+We present the  second order response results on the 3rd row, which is
+also calculated from the equilibrium simulation. In this case, the statistical
+error is too large, so that
+the profile of $\mt_\tau f$ is indeed too noisy to draw any useful information.
+On the 4th row of Fig.~\ref{fig:tmp6} presents the linear response starting
+from a non-equilibrium simulation with external driving $k_0 = 1$.
+In this case the effective strength of the perturbation to the reference
+simulation is $\eps = 1$, which is considered to be relatively small.
+The numerical results is satisfactorily consistent with the brute force
+non-equilibrium simulation.
 
-\begin{figure}
-  \centering
-  \includegraphics[width=0.95\textwidth]
-  {figs/fig-split-str2-simple.eps}
-  \caption{
-    splitting single well potential:
-    the plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
-    $F_e^{\textrm{max}} = 2.0$.
-    From left to right:
-    (a) $t=0$~\textsf{ps},
-    the brutal force non-equilibirum simulation,
-    (b) $t=20$~\textsf{ps},
-    the brutal force non-equilibirum simulation,
-    (c) $t=20$~\textsf{ps},
-    response formula with reference process $F_e^{\max} = 0.0$
-    (i.e. equilibirum simulation),
-    and
-    (d) $t=20$~\textsf{ps},
-    response formula with reference process $F_e^{\max} = 1.0$.
-    The response formula are truncated at order $n=1$.
-  }
-  \label{fig:new-tmp6}
-\end{figure}
+\section{Conclusions and Remarks}
+
+When the strength of the perturbation $\eps$ is relatively small, then
+the traditional linear response theory gives good approximation.  When
+the perturbation reaches the strength, that is not small for the
+linear response, including the second order response term does not
+substantially improve the accuracy.
+In the splitting single-well example, the statistical
+noise of the second order response is so big that overwhelms the
+information. Spending more computational effort on reducing the noise
+is not economic. The non-equilibrium response theory developed by this
+work allows one to calculate the response from some intermediate
+non-equilibrium simulation, so the effective perturbation is much
+smaller than it would be calculated from an equilibrium
+simulation.
+In both examples considered by this paper, the
+numerical results of the proposed idea  satisfactory consistent with the
+brute force non-equilibrium simulation.
+From the face that, in a strongly driven system,
+one should do an intermediate non-equilibrium
+simulation and calculate the response from it,
+one may argue that this may not be able to
+save a lot computational cost comparing with a brute force
+non-equilibrium simulation.  However, in practice, one need the
+response theory for the following reasons: (1) One may want to study
+multiple non-equilibrium system that are not very far from each other,
+so only do one system and apply the linear response idea takes
+advantage over doing non-equilibrium simulation for each of them.
+(2) A stronger driven non-equilibrium system may be more difficult to
+simulation directly comparing with a weakly driven system. For
+example, the energy barrier between two rare events may be increased
+greatly due to the external driven. (3) By the traditional linear
+response theory, some information concerning the intermediate
+non-equilibrium system can be derived from an standard equilibrium
+simulation.
+With this extra information, may be able to design better simulation methods
+for the intermediate non-equilibrium system,
+so that it is not as expensive as a brute force non-equilibrium
+simulation starting with nothing know before hand.
+(4) A thermodynamic multistep integration method can be developed from the
+response theory plus idea (3): the non-equilibrium driving is increased
+step-by-step and the total simulation effort at all the steps is still
+cheaper than doing one brute force non-equilibrium simulation for the
+target driven system.
+Till now we have not investigated idea (3) and (4) mentioned above,
+however, we believe the non-equilibrium response theory is a very
+important step toward it and by itself deserves a separate careful
+investigation.
+
+% \newpage
+% \section{Numerical results}
+% \subsection{Tilting double-well potential}
+
+% \begin{figure}
+%   \centering
+%   \includegraphics[width=0.4\textwidth]{figs/fig-tilt-pot.eps}
+%   \includegraphics[width=0.4\textwidth]{figs/fig-split-pot.eps}
+%   \caption{Two testing cases presented by this paper.
+%     Left: the double-well potential with tilting perturbation.
+%     Right: the single-well potential with splitting perturbation.}
+%   \label{fig:new-tmp1}
+% \end{figure}
+
+% We test the idea by a one-dimensional model system: one particle in a
+% tilting double-well potential. For convenience, we let the mass of the
+% particle to be 1 \textsf{amu}. The unperturbed
+% Hamiltonian of the system is given by:
+% \begin{align}
+%   \mh (\vect p, \vect q) = \frac 12 \vect p^2 + U(\vect q) 
+% \end{align}
+% with potential
+% \begin{align}
+%   U(\vect q) = \frac12 k (\vect q^2 - a^2)^2
+% \end{align}
+% Here $k = 8$~$\textsf{kJ} / (\textsf{mol nm}^4)$, and $ a = 1\ \textsf{nm}$.
+% Notice at room temperature $300\ \textsf{K}$, $k_BT = 2.48$~\textsf{kJ/mol}.
+% See the red line in Fig.~\ref{fig:tmp1}.
+% %$\textsf{kJ} / (\textsf{mol nm}^4)}$
+% The perturbation is given by
+% \begin{align}
+%   \vect C(\vect p, \vect q) = 0; \qquad
+%   \vect D(\vect p, \vect q) = -\nabla_{\vect q} V(\vect q) = 1
+% \end{align}
+% Here $V(\vect q) = -\vect q$ is  effectively tilting the original
+% potential $U(\vect q)$. The strength of the perturbation $F_e(t)$ is given
+% by the following function:
+% \begin{align}
+%   F_e(t) = 
+%   \begin{cases}
+%     F_e^{\textrm{max}}\times (t / t_c) & t < t_c \\
+%     F_e^{\textrm{max}} & t \geq t_c
+%   \end{cases}
+% \end{align}
+% $t_c = 20$~\textsf{ps} so that the perturbation increases slow
+% enough: the typical decaying time scale of the correlation function
+% in~\eqref{eqn:core-identify-approx-2} is roughly 3~\textsf{ps}.
+% We choose $\tau = 1$~\textsf{ps}.
+% % We choose $\mt_\alpha$
+% % to be the propagator of the Langevin dynamics at temperature $150$
+% % \textsf{K} with $\alpha = 1$ \textsf{ps}.
+% Since the dynamics is in
+% 1-d, the projection $\mathcal P$ is identity.  We consider the stability of the
+% distribution $f(\vect x, t)$ at time $t$: $\Delta\mt_\tau f(\vect x, t)$.
+% The larger
+% this value, the more stable the core sets are.
+
+% \begin{figure}
+%   \centering
+%   \includegraphics[width=0.95\textwidth]
+%   {figs/fig-tilt-str2-simple.eps}
+%   \caption{
+%     Tilting double well potential testing case:
+%     the plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
+%     $F_e^{\textrm{max}} = 2.0$.
+%     From left to right:
+%     (a) $t=0$~\textsf{ps},
+%     the brutal force non-equilibirum simulation,
+%     (b) $t=20$~\textsf{ps},
+%     the brutal force non-equilibirum simulation,
+%     (c) $t=20$~\textsf{ps},
+%     response formula truncated at order $n=1$,
+%     and
+%     (d) $t=20$~\textsf{ps},
+%     response formula truncated at order $n=2$.
+%     The equilibirum state serves as the reference of
+%     the response simulations.
+%   }
+%   \label{fig:new-tmp2}
+% \end{figure}
+
+% \begin{figure}
+%   \centering
+%   \includegraphics[width=0.95\textwidth]
+%   {figs/fig-tilt-str4-simple.eps}
+%   \caption{
+%     Tilting double well potential:
+%     the plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
+%     $F_e^{\textrm{max}} = 4.0$.
+%     From left to right:
+%     (a) $t=0$~\textsf{ps},
+%     the brutal force non-equilibirum simulation,
+%     (b) $t=20$~\textsf{ps},
+%     the brutal force non-equilibirum simulation,
+%     (c) $t=20$~\textsf{ps},
+%     response formula with reference process $F_e^{\max} = 0.0$
+%     (i.e. equilibirum simulation),
+%     and
+%     (d) $t=20$~\textsf{ps},
+%     response formula with reference process $F_e^{\max} = 2.0$.
+%     The response formula are truncated at order $n=1$.
+%   }
+%   \label{fig:new-tmp3}
+% \end{figure}
+
+% The Fig.~\ref{fig:new-tmp2} and \ref{fig:new-tmp3} present
+% the application of the response theory to
+% calculate core sets of the tilting double-well potential.
+% Fig.~\ref{fig:new-tmp2} presents
+% the snapshot at $t=20$~\textsf{ps}
+% of $\Delta\mt_\tau f(\vect x,t)$ under the maximum perturbation
+% strength $F_e^{\max} = 2.0$. Response formula truncated
+% at both order $n=1$ and $n=2$ are compared to 
+% the brutal force non-equilibirum simulation.
+% No obviously difference is shown. The second order response result
+% is slightly more precise than the first order one, but the
+% statistical error is also larger, which makes the profile noisy.
+% This figure tells us that
+% (1) the modeling error of the linear response (1st order) is
+% negligibly small; (2) the statistical error of numerically measuring the
+% linear response is also small.
+% Fig.~\ref{fig:new-tmp3} presents the $t=20$~\textsf{ps}
+% snapshots of $\Delta\mt_\tau f(\vect x,t)$
+% under a larger perturbations $F_e^{\max} = 4.0$.
+% The linear reponse using the equilibirum process as the reference
+% shows poor accuracy: the left well vanishes in this case,
+% while the linear response produces  artifical structures.
+% Since the profile is smooth, which indicates
+% the statistical error of calculating the linear term is small,
+% the deviation comes from the modellng error, namely the contribution
+% of higher order response terms.
+% By including the second order response term reduces the left
+% artifical core set, but also creates artifical finer grained
+% structures. The statistical accuracy is also worse than the linear response.
+% Plot (d) present the result of the linear response
+% formula using reference process $F_e^{\max} = 2.0$, which
+% is the process presented in Fig.~\ref{fig:new-tmp2}.
+% Much better modelling accuracy is achieved:
+% the left artifical core set disappear and the
+% statistical accuracy is also satisfactory.
+
+% % See Fig. \ref{fig:tmp2} and \ref{fig:tmp3} for numerical result.  Both
+% % the direct non-equilibrium simulation and the response
+% % approximations are presented. Time slices $t = 0$, 10, 15 and 20
+% % \textsf{ps} are shown.  When the perturbation is small,
+% % i.e. $F_e^{\textrm{max}} = 2$, the direct non-equilibrium simulation
+% % and the linear response results are accurately consistent.
+% % The 2nd order response formula~\eqref{eqn:pert-approx-1} and
+% % \eqref{eqn:pert-approx-2} are even better, but the numerical error
+% % is bigger.
+% % As the
+% % perturbation grows stronger, the left well becomes less stable, while
+% % the right well becomes more stable.  When the perturbation is big,
+% % i.e. $F_e^{\textrm{max}} = 4$, the linear response result is no
+% % longer precise.
+% % When $t\geq 20$ \textsf{ps}, the left well basically
+% % disappears, however, the linear response calculation presents
+% % artifical occilation at the right well.
+% % The 2nd order response also has artifical effect, and the numerical
+% % uncertainty is high.
+% % The 4th rwo of Fig.~\ref{fig:tmp3} presents the 1st order result of the
+% % response formula~\eqref{eqn:pert-approx-1} and
+% % \eqref{eqn:pert-approx-2}, with
+% % respect to the reference state $F_e^{\textrm{max}} = 2$. The result
+% % is impressively improved compared with the linear response with respect to the
+% % equilibrium state.
+
+% \subsection{Splitting single-well potential}
+
+% In this splitting single-well potential example, the unperturbed
+% potential is simply the harmonic potential given by:
+% \begin{align}
+%   U(\vect q) = \frac12\,k\,\vect q^2 
+% \end{align}
+% Here $k = 8$~$\textsf{kJ} / (\textsf{mol nm}^2)$.
+% Notice at room temperature $300\ \textsf{K}$, $k_BT = 2.48$~\textsf{kJ/mol}.
+% See the red line in Fig.~\ref{fig:tmp4}.
+% %$\textsf{kJ} / (\textsf{mol nm}^4)}$
+% The perturbation is again given by
+% \begin{align}
+%   \vect C(\vect p, \vect q) = 0; \qquad
+%   \vect D(\vect p, \vect q) = -\nabla_{\vect q} V(\vect q) 
+% \end{align}
+% where $V(\vect q)$ is a Gaussian type potential:
+% \begin{align}
+%   V(\vect q) = \frac{1}{\sqrt{2\pi \sigma^2}}
+%   \exp\Big\{-\frac{\vect q^2}{2\sigma^2}\Big\}
+% \end{align}
+% we use $\sigma = 0.16$~\textsf{nm}. Please see Fig~\ref{fig:tmp4}.
+% For the results, please see Fig.~\ref{fig:tmp5} -- Fig.~\ref{fig:tmp6}.
+
+% \begin{figure}
+%   \centering
+%   \includegraphics[width=0.95\textwidth]
+%   {figs/fig-split-str1-simple.eps}
+%   \caption{
+%     Splitting single-well potential testing case:
+%     the plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
+%     $F_e^{\textrm{max}} = 1.0$.
+%     From left to right:
+%     (a) $t=0$~\textsf{ps},
+%     the brutal force non-equilibirum simulation,
+%     (b) $t=20$~\textsf{ps},
+%     the brutal force non-equilibirum simulation,
+%     (c) $t=20$~\textsf{ps},
+%     response formula truncated at order $n=1$,
+%     and
+%     (d) $t=20$~\textsf{ps},
+%     response formula truncated at order $n=2$.
+%     The equilibirum state serves as the reference of
+%     the response simulations.
+%   }
+%   \label{fig:new-tmp5}
+% \end{figure}
+
+% \begin{figure}
+%   \centering
+%   \includegraphics[width=0.95\textwidth]
+%   {figs/fig-split-str2-simple.eps}
+%   \caption{
+%     splitting single well potential:
+%     the plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
+%     $F_e^{\textrm{max}} = 2.0$.
+%     From left to right:
+%     (a) $t=0$~\textsf{ps},
+%     the brutal force non-equilibirum simulation,
+%     (b) $t=20$~\textsf{ps},
+%     the brutal force non-equilibirum simulation,
+%     (c) $t=20$~\textsf{ps},
+%     response formula with reference process $F_e^{\max} = 0.0$
+%     (i.e. equilibirum simulation),
+%     and
+%     (d) $t=20$~\textsf{ps},
+%     response formula with reference process $F_e^{\max} = 1.0$.
+%     The response formula are truncated at order $n=1$.
+%   }
+%   \label{fig:new-tmp6}
+% \end{figure}
 

### 2013-06-10 one.test/template.optimal.ctrl/parameters.sh
add template for the optimal control, copy from the old template

@@ -1,0 +1,51 @@
+dt=1e-4
+nst=1e10
+noneq_check_feq=1.0
+noneq_time=20.0
+corr_step=0.01
+corr_time=10.0
+warm_time=20.0
+refe_strength=2.0
+pert_strength=3.0
+resp_order=1
+#project_name=splitSingleWell
+project_name=tiltDoubleWell
+
+load_saved_corr=no
+saved_corr_dir=~/study/noneq.msm/one.test/saved.corr.result5
+saved_corr_nst=1e9
+load_saved_resp=yes
+saved_resp_dir=~/study/noneq.msm/one.test/saved.corr.result5
+
+gamma=1.0
+temperature=300.0
+branch_feq=1.0
+quench_temperature=150
+quench_time=1.0
+x0=-2.0
+x1=2.0
+v0=-8.0
+v1=8.0
+nx=30
+nv=30
+seed=`date +%s`
+
+# tilt double well parameters
+double_well_k=8.0
+double_well_a=1.0
+# split single well parameters
+single_well_k=8.0
+gaussian_sigma=0.16
+
+if echo $project_name | grep tiltDoubleWell &> /dev/null; then
+    command_line_param_print="--double-well-k $double_well_k --double-well-a $double_well_a"
+    print_k=`printf "%.1f" $double_well_k`
+    print_a=`printf "%.1f" $double_well_a`
+    save_corr_param_note="k$print_k.a$print_a"
+else if echo $project_name | grep splitSingleWell &> /dev/null; then 
+    command_line_param_print="--single-well-k $single_well_k --gaussian-sigma $gaussian_sigma"
+    print_k=`printf "%.1f" $single_well_k`
+    print_s=`printf "%.3f" $gaussian_sigma`
+    save_corr_param_note="k$print_k.s$print_s"
+fi
+fi

### 2013-06-10 one.test/template.optimal.ctrl/run.corr.sh

@@ -1,0 +1,65 @@
+#!/bin/bash
+
+source parameters.sh
+
+make -C tools/simulator makedir &> make.log
+make -C tools/simulator -j8 &>> make.log
+
+rm -f run.corr.log
+
+print_corr_step=`printf %.3f $corr_step`
+print_corr_time=`printf %.1f $corr_time`
+print_x0=`printf %.1f $x0`
+print_x1=`printf %.1f $x1`
+print_v0=`printf %.1f $v0`
+print_v1=`printf %.1f $v1`
+print_nx=`printf %d $nx`
+print_nv=`printf %d $nv`
+
+if test ! -d $saved_corr_dir; then
+    echo "# make dir $saved_corr_dir"
+    mkdir -p $saved_corr_dir
+fi
+
+if echo $load_saved_corr | grep yes &> /dev/null; then
+    equi_time=`echo "$dt * $saved_corr_nst" | octave | grep ans | cut -d '=' -f 2`
+    print_equi_time=`printf %.2f $equi_time`
+    saved_corr_name=$project_name.$save_corr_param_note.save.corr.step$print_corr_step.time$print_corr_time.equiT$print_equi_time.grid$print_x0.$print_x1.$print_nx.$print_v0.$print_v1.$print_nv
+    if test -f $saved_corr_dir/$saved_corr_name; then
+	echo "# existing saved corr $saved_corr_dir/$saved_corr_name, load it"
+	command="`pwd`/tools/simulator/$project_name.timeCorr \
+--dt $dt --nst $nst -p 1000000 --gamma $gamma --temperature $temperature \
+$command_line_param_print \
+--branch-feq $branch_feq \
+--noneq-check-feq $noneq_check_feq --noneq-time $noneq_time \
+--quench-temperature $quench_temperature --quench-time $quench_time \
+--corr-time $corr_time --corr-step $corr_step \
+--load-corr $saved_corr_dir/$saved_corr_name \
+--x-low $x0 --x-up $x1 --v-low $v0 --v-up $v1 --x-grid $nx --v-grid $nv \
+--warm-time $warm_time --seed $seed --pert-strength $pert_strength"
+    else
+	echo "# no saved corr $saved_corr_dir/$saved_corr_name, exit"
+	exit
+    fi
+else
+    equi_time=`echo "$dt * $nst" | octave | grep ans | cut -d '=' -f 2`
+    print_equi_time=`printf %.2f $equi_time`
[…]

### 2013-06-10 one.test/template.optimal.ctrl/run.noneq.sh

@@ -1,0 +1,23 @@
+#!/bin/bash
+
+source parameters.sh
+
+make -C tools/simulator makedir &> make.log
+make -C tools/simulator -j8 &>> make.log
+
+rm -f run.noneq.log
+
+command="`pwd`/tools/simulator/$project_name.noneq \
+--dt $dt --nst $nst -p 100000 --gamma $gamma --temperature $temperature \
+$command_line_param_print \
+--branch-feq $branch_feq \
+--noneq-check-feq $noneq_check_feq --noneq-time $noneq_time \
+--quench-temperature $quench_temperature --quench-time $quench_time \
+--x-low $x0 --x-up $x1 --v-low $v0 --v-up $v1 --x-grid $nx --v-grid $nv \
+--warm-time $warm_time --seed $seed --pert-strength $pert_strength"
+
+echo "# run on `hostname`" &>> run.noneq.log
+echo "# command is $command" &>> run.noneq.log
+$command > output.noneq 2> error.noneq &
+echo "# pid is $!" &>> run.noneq.log
+

### 2013-06-10 one.test/template.optimal.ctrl/run.resp.sh

@@ -1,0 +1,79 @@
+#!/bin/bash
+
+source parameters.sh
+
+make -C tools/simulator makedir &> make.log
+make -C tools/simulator -j8 &>> make.log
+
+rm -f run.resp.log
+
+print_noneq_step=`printf %.2f $noneq_check_feq`
+print_noneq_time=`printf %.1f $noneq_time`
+print_quench_time=`printf %.1f $quench_time`
+print_x0=`printf %.1f $x0`
+print_x1=`printf %.1f $x1`
+print_v0=`printf %.1f $v0`
+print_v1=`printf %.1f $v1`
+print_nx=`printf %d $nx`
+print_nv=`printf %d $nv`
+
+equi_time=`echo "$dt * $nst" | octave | grep ans | cut -d '=' -f 2`
+print_equi_time=`printf %.2f $equi_time`
+print_quench_time=`printf %.2f $quench_time`
+print_refStr=`printf %.2f $refe_strength`
+
+if test ! -d $saved_resp_dir; then
+    echo "# make dir $saved_resp_dir"
+    mkdir -p $saved_resp_dir
+fi
+
+if echo $load_saved_resp | grep yes &> /dev/null; then
+    saved_resp_name=$project_name.$save_corr_param_note.save.resp.refStr$print_refStr.step$print_noneq_step.time$print_noneq_time.quench$print_quench_time.equiT$print_equi_time.grid$print_x0.$print_x1.$print_nx.$print_v0.$print_v1.$print_nv
+    if test -f $saved_resp_dir/$saved_resp_name; then
+	echo "# existing saved resp $saved_resp_dir/$saved_resp_name, load it"
+	command="`pwd`/tools/simulator/$project_name.noneqResp \
+--dt $dt --nst $nst -p 1000000 --gamma $gamma --temperature $temperature \
+$command_line_param_print \
+--branch-feq $branch_feq \
+--noneq-check-feq $noneq_check_feq --noneq-time $noneq_time \
+--quench-temperature $quench_temperature --quench-time $quench_time \
+--load-corr $saved_resp_dir/$saved_resp_name \
+--x-low $x0 --x-up $x1 --v-low $v0 --v-up $v1 --x-grid $nx --v-grid $nv \
+--warm-time $warm_time --seed $seed \
+--order $resp_order \
+--pert-strength0 $refe_strength --pert-strength1 $pert_strength"
+    else
+	echo "# no saved resp $saved_resp_dir/$saved_resp_name, exit"
+	exit
+    fi
+else
[…]

### 2013-06-10 one.test/template.optimal.ctrl/tools/simulator/app/equi.cpp

@@ -1,0 +1,109 @@
+#include <algorithm>
+#include <fstream>
+#include <iomanip>
+#include <iostream>
+#include <iterator>
+#include <list>
+#include <map>
+#include <set>
+#include <sstream>
+#include <string>
+#include <queue>
+#include <vector>
+#include <cmath>
+
+#include "Force.h"
+#include "Integrator.h"
+#include "Distribution.h"
+#include <boost/program_options.hpp>
+
+namespace po = boost::program_options;
+
+int main(int argc, char * argv[])
+{
+  double gamma;
+  double kT;
+  double T;
+  double dt;
+  // double pertSt;
+  double aa;
+  double kk;
+  double nst;
+  unsigned nstprint = 1;
+  
+  po::options_description desc ("Allow options");
+  desc.add_options()
+      ("help,h", "print this message")
+      ("dt,d",po::value<double > (&dt)->default_value(0.01), "time step [ps]")
+      ("nst,n",po::value<double > (&nst)->default_value(1000), "number of time step")
+      ("print-feq,p",po::value<unsigned > (&nstprint)->default_value(10), "print frequency")
+      ("gamma,g", po::value<double > (&gamma)->default_value(1.), "gamma [ps^-1]")
+      ("temperature,t",po::value<double > (&T)->default_value(300.), "temperature [K]")
+      ("double-well-k,k",po::value<double > (&kk)->default_value(8.0), "k parameter of the double well potential [kJ/(mol nm^4)]")
+      ("double-well-a,a",po::value<double > (&aa)->default_value(1.0), "a parameter of the double well potential [nm]");
+      // ("pert-strength,s",po::value<double > (&pertSt)->default_value(1.0), "perturbation strength [kJ/(mol nm)]");
+      
+  po::variables_map vm;
+  po::store(po::parse_command_line(argc, argv, desc), vm);
+  po::notify (vm);
+  if (vm.count("help")){
+    std::cout << desc<< "\n";
+    return 0;
+  }
+  kT = T * 1.38 * 6.02 * 1e-3;
+  
+  std::cout << "###################################################" << std::endl;
+  std::cout << "# T: " << T << " [K]" << std::endl;
[…]

### 2013-06-10 one.test/template.optimal.ctrl/tools/simulator/app/splitSingleWell.timeCorr.cpp

@@ -1,0 +1,231 @@
+#include <algorithm>
+#include <fstream>
+#include <iomanip>
+#include <iostream>
+#include <iterator>
+#include <list>
+#include <map>
+#include <set>
+#include <sstream>
+#include <string>
+#include <queue>
+#include <vector>
+#include <cmath>
+
+#include "Force.h"
+#include "Perturbation.h"
+#include "Integrator.h"
+#include "Distribution.h"
+#include <boost/program_options.hpp>
+#include "RandomGenerator.h"
+#include "TimeCorrelation.h"
+#include "QuenchCorr.h"
+
+namespace po = boost::program_options;
+
+int main(int argc, char * argv[])
+{
+  double gamma;
+  double kT;
+  double T;
+  double dt;
+  double pertSt;
+  double sigma;
+  double kk;
+  double nst;
+  unsigned nstprint = 1;
+  double warmTime;
+  
+  double noneqTime;
+  double noneqCheckFeq;
+  double branchFeq;
+  double quenchTime;
+  double quenchTemperture;
+  double quenchkT;
+  double corrTime;
+  double corrStep;
+  // unsigned noneqCheckNumFeq;
+  unsigned branchNumFeq;
+  string sfile, lfile;
+  double x0, x1, v0, v1;
+  unsigned nx, nv;
+  
+  unsigned long seed;
+  
+  po::options_description desc ("Allow options");
+  desc.add_options()
+      ("help,h", "print this message")
+      ("dt,d", po::value<double > (&dt)->default_value(0.01), "time step [ps]")
+      ("nst,n", po::value<double > (&nst)->default_value(1000.), "number of time step")
+      ("print-feq,p", po::value<unsigned > (&nstprint)->default_value(10000), "print frequency")
+      ("gamma,g", po::value<double > (&gamma)->default_value(1.), "gamma [ps^-1]")
+      ("temperature,t", po::value<double > (&T)->default_value(300.), "temperature [K]")
+      ("single-well-k,k",po::value<double > (&kk)->default_value(8.0), "k parameter of the double single potential [kJ/(mol nm^2)]")
+      ("gaussian-sigma,a", po::value<double > (&sigma)->default_value(0.16), "sigma of Gaussian perturbation [nm]")
[…]

### 2013-06-10 one.test/template.optimal.ctrl/tools/simulator/app/tiltDoubleWell.noneq.cpp

@@ -1,0 +1,242 @@
+#include <algorithm>
+#include <fstream>
+#include <iomanip>
+#include <iostream>
+#include <iterator>
+#include <list>
+#include <map>
+#include <set>
+#include <sstream>
+#include <string>
+#include <queue>
+#include <vector>
+#include <cmath>
+
+#include "Force.h"
+#include "Perturbation.h"
+#include "Integrator.h"
+#include "Distribution.h"
+#include <boost/program_options.hpp>
+#include "RandomGenerator.h"
+
+namespace po = boost::program_options;
+
+int main(int argc, char * argv[])
+{
+  double gamma;
+  double kT;
+  double T;
+  double dt;
+  double pertSt;
+  double aa;
+  double kk;
+  double nst;
+  unsigned nstprint = 1;
+  double warmTime;
+  
+  double noneqTime;
+  double noneqCheckFeq;
+  double branchFeq;
+  double quenchTime;
+  double quenchTemperture;
+  double quenchkT;
+  unsigned noneqCheckNumFeq;
+  unsigned branchNumFeq;
+  double x0, x1, v0, v1;
+  unsigned nx, nv;
+      
+  unsigned long seed;
+  
+  po::options_description desc ("Allow options");
+  desc.add_options()
+      ("help,h", "print this message")
+      ("dt,d", po::value<double > (&dt)->default_value(0.01), "time step [ps]")
+      ("nst,n", po::value<double > (&nst)->default_value(10000), "number of time step")
+      ("print-feq,p", po::value<unsigned > (&nstprint)->default_value(1000), "print frequency")
+      ("gamma,g", po::value<double > (&gamma)->default_value(1.), "gamma [ps^-1]")
+      ("temperature,t", po::value<double > (&T)->default_value(300.), "temperature [K]")
+      ("double-well-k,k", po::value<double > (&kk)->default_value(8.0), "k parameter of the double well potential [kJ/(mol nm^4)]")
+      ("double-well-a,a", po::value<double > (&aa)->default_value(1.0), "a parameter of the double well potential [nm]")
+      ("branch-feq", po::value<double > (&branchFeq)->default_value(1.), "branch frequency [ps]")
[…]

### 2013-06-10 one.test/template.optimal.ctrl/tools/simulator/app/tiltDoubleWell.timeCorr.cpp

@@ -1,0 +1,231 @@
+#include <algorithm>
+#include <fstream>
+#include <iomanip>
+#include <iostream>
+#include <iterator>
+#include <list>
+#include <map>
+#include <set>
+#include <sstream>
+#include <string>
+#include <queue>
+#include <vector>
+#include <cmath>
+
+#include "Force.h"
+#include "Perturbation.h"
+#include "Integrator.h"
+#include "Distribution.h"
+#include <boost/program_options.hpp>
+#include "RandomGenerator.h"
+#include "TimeCorrelation.h"
+#include "QuenchCorr.h"
+
+namespace po = boost::program_options;
+
+int main(int argc, char * argv[])
+{
+  double gamma;
+  double kT;
+  double T;
+  double dt;
+  double pertSt;
+  double aa;
+  double kk;
+  double nst;
+  unsigned nstprint = 1;
+  double warmTime;
+  
+  double noneqTime;
+  double noneqCheckFeq;
+  double branchFeq;
+  double quenchTime;
+  double quenchTemperture;
+  double quenchkT;
+  double corrTime;
+  double corrStep;
+  // unsigned noneqCheckNumFeq;
+  unsigned branchNumFeq;
+  string sfile, lfile;
+  double x0, x1, v0, v1;
+  unsigned nx, nv;
+  
+  unsigned long seed;
+  
+  po::options_description desc ("Allow options");
+  desc.add_options()
+      ("help,h", "print this message")
+      ("dt,d", po::value<double > (&dt)->default_value(0.01), "time step [ps]")
+      ("nst,n", po::value<double > (&nst)->default_value(1000.), "number of time step")
+      ("print-feq,p", po::value<unsigned > (&nstprint)->default_value(10000), "print frequency")
+      ("gamma,g", po::value<double > (&gamma)->default_value(1.), "gamma [ps^-1]")
+      ("temperature,t", po::value<double > (&T)->default_value(300.), "temperature [K]")
+      ("double-well-k,k", po::value<double > (&kk)->default_value(8.0), "k parameter of the double well potential [kJ/(mol nm^4)]")
+      ("double-well-a,a", po::value<double > (&aa)->default_value(1.0), "a parameter of the double well potential [nm]")
[…]

### 2013-06-10 one.test/template.optimal.ctrl/tools/simulator/include/Force.h

@@ -1,0 +1,55 @@
+#ifndef __Force_h_NONEQMSM_wanghan__
+#define __Force_h_NONEQMSM_wanghan__
+
+#include "Defines.h"
+
+class Force 
+{
+public:
+  virtual double potential (const Dofs & dofs,
+			    const double & time) const = 0;
+  virtual void operator () (const Dofs & dofs,
+			    vector<double> & fvalue) const = 0;
+  virtual void operator () (const Dofs & dofs,
+			    const double & time,
+			    vector<double> & fvalue) const = 0;
+}
+    ;
+
+// U(r) = 0.5 * kk * (r^2 - aa^2)^2
+class DoubleWell : public Force
+{
+  double kk;
+  double aa;
+public:
+  DoubleWell (const double & k,
+	      const double & a);
+  virtual double potential (const Dofs & dofs,
+			    const double & time) const;
+  virtual void operator () (const Dofs & dofs,
+			    vector<double> & fvalue) const ;
+  virtual void operator () (const Dofs & dofs,
+			    const double & time,
+			    vector<double> & fvalue) const ;
+}
+    ;
+
+// U(r) = 0.5 * kk * r^2
+class SingleWell : public Force 
+{
+  double kk;
+public:
+  public:
+  SingleWell (const double k = 8.);
+  virtual double potential (const Dofs & dofs,
+			    const double & time) const;
+  virtual void operator () (const Dofs & dofs,
+			    vector<double> & fvalue) const ;
+  virtual void operator () (const Dofs & dofs,
+			    const double & time,
+			    vector<double> & fvalue) const ;
+}
+    ;
+
+
+#endif

### 2013-06-10 one.test/template.optimal.ctrl/tools/simulator/include/NoneqResponseInfo.h

@@ -1,0 +1,86 @@
+#ifndef __NoneqResponseInfo_h_NONEQMSM_wanghan__
+#define __NoneqResponseInfo_h_NONEQMSM_wanghan__
+
+#include <vector>
+#include "Defines.h"
+#include "Distribution.h"
+#include "Perturbation.h"
+
+using namespace std;
+
+class NoneqResponseInfo 
+{
+  double x0, x1, v0, v1;
+  unsigned nx, nv;
+  
+  double dt;
+  double noneqTime;
+  double noneqCheckFeq;
+  int numCheck;
+  int checkNumFeq;
+  double quenchTime;
+  int quenchNumStep;
+
+  int countNoneq;
+  int countNoneqSeg;
+  int countQuench;
+
+  const Perturbation * ppert;
+  int numMode;
+  
+  vector<Distribution_1d >			order0;		// time
+  vector<vector<Distribution_1d > >		order1;		// time x vec
+  vector<vector<vector<Distribution_1d > > >	order2;		// time x mat  
+  
+  vector<Distribution_1d >			quench_order0;		// time
+  vector<Distribution_1d >			quench_order1_term1;	// time
+  vector<vector<Distribution_1d > >		quench_order1_term2;	// time x vec
+  vector<Distribution_1d >			quench_order2_term1;	// time
+  vector<vector<vector<Distribution_1d > > >	quench_order2_term2;	// time x mat
+  vector<vector<Distribution_1d > >		quench_order2_term3;	// time x vec
+  
+  vector<double >				Gj;		// vec 
+  vector<vector<double > >			Hjk;		// mat 
+  double					G0;		// 1
+  double					H00;		// 1
+
+public:
+  NoneqResponseInfo ();
+  // ~NoneqResponseInfo ();
+public:
+  void reinit (const double & x0,
+	       const double & x1,
+	       const unsigned & nx,
+	       const double & v0,
+	       const double & v1,
+	       const unsigned & nv,
+	       const double & dt,
+	       const double & noneqTime,
+	       const double & noneqCheckFeq,
+	       const double & quenchTime,
+	       const Perturbation & pert);
+  void newTraj ();
+  void newQuenchTraj ();
+  void depositMainTraj (const Dofs & oldx,
+			const Dofs & newx,
+			const double & sigma,
+			const Dofs & dw);
+  void depositQuenchTraj (const Dofs & oldx,
+			  const Dofs & newx,
+			  const double & sigma,
+			  const Dofs & dw);
[…]

### 2013-06-10 one.test/template.optimal.ctrl/tools/simulator/include/Perturbation.h

@@ -1,0 +1,93 @@
+#ifndef __Perturbation_h_NONEQMSM_wanghan__
+#define __Perturbation_h_NONEQMSM_wanghan__
+
+#include "Defines.h"
+#include "Force.h"
+#include <vector>
+
+using namespace std;
+
+class Perturbation 
+{
+public:
+  virtual double Fe (const double & time) const = 0;
+  virtual int numMode () const;
+  virtual void FeMode     (const double & time,
+			   vector<double > & modes) const;
+  virtual void FeModePref (vector<double > & pref ) const;
+public:
+  virtual void operator () (const Dofs & dofs,
+  			    Dofs & pvalue) const = 0;
+  virtual void operator () (const Dofs & dofs,
+			    const double & time,
+			    Dofs & pvalue) const = 0;
+}
+    ;
+
+class DissipativeFlux 
+{
+  const Perturbation * pert;
+  const Force * force;
+  mutable Dofs pvalue;
+  mutable vector<double > fvalue;
+public:
+  DissipativeFlux (const Perturbation * p,
+		   const Force * f);
+  double operator () (const Dofs & dofs) const;
+}
+    ;
+
+// V(r) = -r
+class PertConstTilt : public Perturbation
+{
+  double strength;
+  double warmTime;
+public:
+  PertConstTilt (const double & s,
+		 const double warmTime = 0.);
+public:
+  virtual double Fe (const double & time) const;
+  virtual int numMode () const;
+  virtual void FeMode (const double & time,
+		       vector<double > & modes) const;
+  virtual void FeModePref (vector<double > & pref ) const;
+public:
+  virtual void operator () (const Dofs & dofs,
+  			    Dofs & pvalue) const;
+  virtual void operator () (const Dofs & dofs,
+			    const double & time,
+			    Dofs & pvalue) const;
+}
+    ;
+
+// V(r) = 1/(2 \pi \sigma^2) * \exp (-(r - \mu)^2 / (2 \sigma^2))
+class PertGaussian : public Perturbation 
+{
+  double strength;
+  double mu;
+  double sigma;
+  double warmTime;
+private:
+  double factor0;
+  double factor1;  
+public:
+  PertGaussian (const double s = 1.,
+		const double m = 0.,
+		const double sig = 1.,
+		const double warmTime = 0.);
+public:
+  virtual double Fe (const double & time) const;
+  virtual int numMode () const;
+  virtual void FeMode (const double & time,
[…]

### 2013-06-10 one.test/template.optimal.ctrl/tools/simulator/include/QuenchCorr.h

@@ -1,0 +1,38 @@
+#ifndef __QuenchCorr_h_NONEQMSM_wanghan__
+#define __QuenchCorr_h_NONEQMSM_wanghan__
+
+#include "Defines.h"
+#include "Distribution.h"
+#include "Perturbation.h"
+#include "Integrator.h"
+
+class QuenchCorr
+{
+  Distribution_1d corr;
+  double sigma;
+  double dt;
+  const Perturbation * pert;
+  double calG (const std::vector<Dofs > & dofs,
+	       const std::vector<Dofs > & dw);
+  FILE * fp;
+public:
+  QuenchCorr (const double & x0,
+	      const double & x1,
+	      const unsigned & nx,
+	      const double & v0,
+	      const double & v1,
+	      const unsigned & nv,
+	      const EulerMaruyama & quench_inte0,	// quench integrator at equilibrium
+	      const Perturbation & pert);		// perturbation
+  ~QuenchCorr ();
+  void deposit (const Dofs & endpoint,
+		const std::vector<Dofs > & dofs,
+		const std::vector<Dofs > & dw);
+  void average ();
+  const Distribution_1d & getDist () const {return corr;}
+}
+    ;
+
+
+
+#endif

### 2013-06-10 one.test/template.optimal.ctrl/tools/simulator/include/Stopwatch.h

@@ -1,0 +1,53 @@
+#ifndef __Stopwatch_h_wanghan__
+#define __Stopwatch_h_wanghan__
+
+#include <sys/param.h>
+#include <sys/times.h>
+#include <sys/types.h>
+
+class Stopwatch 
+{ 
+public: 
+  Stopwatch(); 
+ 
+  void              start(); 
+  void              stop(); 
+  double            system() const; 
+  double            user() const; 
+  double            real() const; 
+ 
+  static double     resolution() {return 1./HZ;}; 
+private:
+  struct tms tic, toc;
+  long r1, r0;
+  double HZi;
+};               
+
+inline double Stopwatch::user () const
+{
+  return (double)(toc.tms_utime - tic.tms_utime) * HZi;
+}
+
+inline double Stopwatch::system () const
+{
+  return (double)(toc.tms_stime - tic.tms_stime) * HZi;
+}
+
+inline double Stopwatch::real () const
+{
+  return (double)(r1 - r0) * HZi;
+}
+
+inline void Stopwatch::stop ()
+{
+  r1 = times (&toc);
+}
+
+inline void Stopwatch::start() 
+{
+  r0 = times (&tic);
+}
+
+#endif
+// end of file 
+

### 2013-06-10 one.test/template.optimal.ctrl/tools/simulator/include/TimeCorrelation.h

@@ -1,0 +1,53 @@
+#ifndef __TimeCorrelation_h_NONEQMSM_wanghan__
+#define __TimeCorrelation_h_NONEQMSM_wanghan__
+
+#include "Defines.h"
+#include "Force.h"
+#include "Integrator.h"
+#include "Distribution.h"
+
+class TimeCorrelation 
+{
+  double step;
+  double time;
+  double dt;
+  unsigned nFrame;
+  unsigned nStep;
+  unsigned calCorrFeq;
+  std::vector<Distribution_1d > dists0;
+  std::vector<Distribution_1d > dists1;
+  const Integrator * equiInte;
+  double quenchAtTime;
+  unsigned quenchNumStep;
+  const Integrator * quenchInte;
+  const DissipativeFlux * flux;
+public:
+  void reinit (const double & step,
+	       const double & time,
+	       const double & x0,
+	       const double & x1,
+	       const unsigned & nx,
+	       const double & v0,
+	       const double & v1,
+	       const unsigned & nv,
+	       const Integrator * equiInte,
+	       const Integrator * quenchInte,
+	       const unsigned & quenchNumStep,
+	       const double & quenchAtTime,
+	       const DissipativeFlux * flux);
+  void calCorr (const Dofs & initDof,
+		const double & totalTime);
+  void calIndicator (const vector<vector<double > > & old,
+		     const double & idTime,
+		     const double & idStep,
+		     const double & beta,
+		     const Perturbation & pert,
+		     vector<vector<vector<double > > > & timeNew);
+  void print () const ;
+  void save (const string & filename) const;
+  void load (const string & filename);
+}
+    ;
+
+
+#endif

### 2013-06-10 one.test/template.optimal.ctrl/tools/simulator/src/Distribution.cpp

@@ -1,0 +1,306 @@
+#include "Distribution.h"
+
+Distribution_1d::
+Distribution_1d (const double & x0_,
+		 const double & x1_,
+		 const unsigned & nx_,
+		 const double & v0_,
+		 const double & v1_,
+		 const unsigned & nv_)
+{
+  reinit (x0_, x1_, nx_, v0_, v1_, nv_);
+}
+
+
+void Distribution_1d::
+reinit (const double & x0_,
+	const double & x1_,
+	const unsigned & nx_,
+	const double & v0_,
+	const double & v1_,
+	const unsigned & nv_)
+{
+  x0 = x0_;
+  x1 = x1_;
+  v0 = v0_;
+  v1 = v1_;
+  nx = nx_;
+  nv = nv_;
+
+  hx = (x1 - x0) / double (nx);
+  hv = (v1 - v0) / double (nv);
+  valuepp = 1./(hx*hv);
+  
+  gridx.resize(nx);
+  gridv.resize(nv);
+
+  for (unsigned ii = 0; ii < gridx.size(); ++ii){
+    gridx[ii] = x0 + (0.5 + ii) * hx;
+  }
+  for (unsigned ii = 0; ii < gridv.size(); ++ii){
+    gridv[ii] = v0 + (0.5 + ii) * hv;
+  }
+
+  values.resize(nx);
+  for (unsigned ii = 0; ii < nx; ++ii){
+    values[ii].resize(nv);
+  }
+  backup_values.resize(nx);
+  for (unsigned ii = 0; ii < nx; ++ii){
+    backup_values[ii].resize(nv);
+  }
+
+  backup_number = 1e6;
+  backup_unbacked_count = 0.;
+  clear ();
+}
+
+
+void Distribution_1d::
+clear ()
+{
+  nframe = 0.;
+  for (unsigned ii = 0; ii < nx; ++ii){
+    for (unsigned jj = 0; jj < nv; ++jj){
+      values[ii][jj] = 0.;
+      backup_values[ii][jj] = 0.;
+    }
+  }
+}
+
+void Distribution_1d::
+deposite (const Dofs & dof)
+{
+  deposite (dof, 1.);
+}
+
+void Distribution_1d::
+deposite (const Dofs & dof,
+	  const double& scale)
+{
+  int ix = (dof.xx[0] - x0) / hx;
+  int iv = (dof.vv[0] - v0) / hv;
[…]

### 2013-06-10 one.test/template.optimal.ctrl/tools/simulator/src/Integrator.cpp

@@ -1,0 +1,105 @@
+#include "Integrator.h"
+#include "RandomGenerator.h"
+
+#include <gsl/gsl_randist.h>
+
+void EulerMaruyama::
+step (Dofs & dofs,
+      const double time) const
+{
+  Dofs pvalue;
+  if (pert != NULL){
+    (*pert) (dofs, time, pvalue);
+  }
+  else {
+    for (unsigned dd = 0; dd < NUMDOFS; ++dd){
+      pvalue.xx[dd] = 0.;
+      pvalue.vv[dd] = 0.;
+    }
+  }
+  vector<double > fvalue (NUMDOFS);
+  if (force != NULL){
+    (*force) (dofs, time, fvalue);
+  }
+  else {
+    for (unsigned dd = 0; dd < NUMDOFS; ++dd){
+      fvalue[dd] = 0.;
+    }
+  }
+
+  // printf ("%f\n", fvalue[0]);
+  Dofs oldDofs(dofs);
+  for (unsigned dd = 0; dd < NUMDOFS; ++dd){
+    // double rand = RandomGenerator_MT19937::gaussian();
+    double rand = gsl_ran_gaussian_ziggurat (rg, 1.);
+    dofs.xx[dd] = oldDofs.xx[dd] + dt * oldDofs.vv[dd] + dt * pvalue.xx[dd];
+    dofs.vv[dd] = oldDofs.vv[dd] + dt * fvalue[dd] - dt * gamma * oldDofs.vv[dd] + dt * pvalue.vv[dd]
+	+ sqrtdt * sigma * rand;
+    storedw.vv[dd] = rand * sqrtdt;
+    // if (! (dofs.xx[dd] >= -10 && dofs.xx[dd] <= 10. && dofs.vv[dd]>= -20. && dofs.vv[dd] <= 20.)  ){
+    //   fprintf (stderr, "dd: %d   old: %f %f    new: %f %f\n",
+    // 	       dd,
+    // 	       oldDofs.xx[dd], oldDofs.vv[dd],
+    // 	       dofs.xx[dd], dofs.vv[dd]);
+    //   exit (1);
+    // }
+  }
+}
+
+
+EulerMaruyama::
+EulerMaruyama ()
+    : dt(0.), sqrtdt(0.), gamma(0.), sigma(0.),
+      pert(NULL), force(NULL), rg (NULL)
+{
+}
+
+EulerMaruyama::
+EulerMaruyama (const double & gamma,
+	       const double & kT,
[…]

### 2013-06-10 one.test/template.optimal.ctrl/tools/simulator/src/NoneqResponseInfo.cpp

@@ -1,0 +1,494 @@
+#include "NoneqResponseInfo.h"
+
+NoneqResponseInfo::
+NoneqResponseInfo ()
+    : ppert (NULL)
+{
+}
+
+
+void NoneqResponseInfo::
+reinit (const double & x0_,
+	const double & x1_,
+	const unsigned & nx_,
+	const double & v0_,
+	const double & v1_,
+	const unsigned & nv_,
+	const double & dt_,
+	const double & noneqTime_,
+	const double & noneqCheckFeq_,
+	const double & quenchTime_,
+	const Perturbation & pert)
+{
+  x0 = x0_;
+  x1 = x1_;
+  nx = nx_;
+  v0 = v0_;
+  v1 = v1_;
+  nv = nv_;
+  dt = dt_;
+  noneqTime = noneqTime_;
+  noneqCheckFeq = noneqCheckFeq_;
+  quenchTime = quenchTime_;
+  ppert = &pert;
+
+  numMode = ppert->numMode();  
+  numCheck = int((noneqTime + 0.5 * noneqCheckFeq) / noneqCheckFeq) + 1;
+  checkNumFeq = int((noneqCheckFeq + 0.5 * dt) / dt);
+  quenchNumStep = int((quenchTime + 0.5 * dt) / dt);
+  
+  order0.resize(numCheck);
+  order1.resize(numCheck);
+  order2.resize(numCheck);
+  
+  quench_order0.resize(numCheck);
+  quench_order1_term1.resize(numCheck);
+  quench_order1_term2.resize(numCheck);
+  quench_order2_term1.resize(numCheck);
+  quench_order2_term2.resize(numCheck);
+  quench_order2_term3.resize(numCheck);
+
+  for (int jj = 0; jj < numCheck; ++jj){
+    order1[jj].resize (numMode);
+    order2[jj].resize (numMode);
+    quench_order1_term2[jj].resize (numMode);
+    quench_order2_term2[jj].resize (numMode);
+    quench_order2_term3[jj].resize (numMode);
+    for (int kk = 0; kk < numMode; ++kk){
+      order2[jj][kk].resize (numMode);
+      quench_order2_term2[jj][kk].resize (numMode);
+    }
+  }
+
+  for (int tt = 0; tt < numCheck; ++tt){
+    order0[tt]				.reinit (x0, x1, nx, v0, v1, nv);
[…]

### 2013-06-10 one.test/template.optimal.ctrl/tools/simulator/src/Perturbation.cpp

@@ -1,0 +1,176 @@
+#include "Perturbation.h"
+
+DissipativeFlux::
+DissipativeFlux (const Perturbation * p,
+		 const Force * f)
+    :pert (p), force(f), fvalue(NUMDOFS)
+{
+}
+
+double DissipativeFlux::
+operator () (const Dofs & dof) const
+{
+  (*pert)  (dof, pvalue);
+  (*force) (dof, fvalue);
+  double tmp = 0.;
+  for (unsigned dd = 0; dd < NUMDOFS; ++dd){
+    tmp += pvalue.xx[dd] * fvalue[dd] - pvalue.vv[dd] * dof.vv[dd];
+  }
+  return tmp;
+}
+
+int Perturbation::
+numMode () const
+{
+  return 1;
+}
+
+void Perturbation::
+FeMode (const double & time,
+	vector<double > & modes) const
+{
+  modes.resize(1);
+  modes[0] = Fe(time);
+}
+
+void Perturbation::
+FeModePref (vector<double > & perf) const
+{
+  perf.resize(1);
+  perf[0] = 1;
+}
+
+
+PertConstTilt::
+PertConstTilt (const double & s,
+	       const double w)
+    : strength(s), warmTime(w)
+{    
+}
+
+void PertConstTilt::
+operator () (const Dofs & dofs,
+	     const double & time,
+	     Dofs & pvalue) const
+{
+  (*this) (dofs, pvalue);
+  pvalue.vv[0] *= Fe(time);
+}
+
+void PertConstTilt::
+operator () (const Dofs & dofs,
+	     Dofs & pvalue) const
+{
+  pvalue.vv[0] = 1.0;
+}
+
+double PertConstTilt::
+Fe (const double & time) const
+{
+  double tmp = strength;
+  if (warmTime > 0.0 && time < warmTime){
+    tmp *= time / warmTime;
+  }
+  return tmp;
+}
+
+int PertConstTilt::
+numMode () const
+{
+  return 1;
+}
+
+void PertConstTilt::
+FeMode (const double & time,
+	vector<double > & modes) const
+{
+  modes.resize(1);
+  modes[0] = 1.;
+  if (warmTime > 0.0 && time < warmTime){
+    modes[0] *= time / warmTime;
+  }
+}
+
+void PertConstTilt::
+FeModePref (vector<double > & pref ) const
+{
+  pref.resize(1);
+  pref[0] = strength;
+}
+
[…]

### 2013-06-10 one.test/template.optimal.ctrl/tools/simulator/src/QuenchCorr.cpp

@@ -1,0 +1,58 @@
+#include "QuenchCorr.h"
+
+QuenchCorr::
+QuenchCorr (const double & x0,
+	    const double & x1,
+	    const unsigned & nx,
+	    const double & v0,
+	    const double & v1,
+	    const unsigned & nv,
+	    const EulerMaruyama & quench_inte0,
+	    const Perturbation & pert_)
+{
+  corr.reinit (x0, x1, nx, v0, v1, nv);
+  sigma = quench_inte0.getSigma();
+  dt = quench_inte0.getDt();
+  pert = &pert_;
+  fp = fopen ("gx.record.out", "w");
+}
+
+QuenchCorr::
+~QuenchCorr ()
+{
+  fclose (fp);
+}
+
+void QuenchCorr::
+deposit (const Dofs & endpoint,
+	 const std::vector<Dofs > & dofs,
+	 const std::vector<Dofs > & dw)
+{
+  double Gx = calG(dofs, dw);
+  corr.deposite (endpoint, Gx);
+  fprintf (fp, "%f\n", Gx);
+}
+
+double QuenchCorr::
+calG (const std::vector<Dofs > & dofs,
+      const std::vector<Dofs > & dw)
+{
+  double sum = 0.;
+  for (unsigned ii = 0; ii < dofs.size(); ++ii){
+    Dofs pvalue;
+    pert->operator()(dofs[ii], pvalue);
+    double dot = 0.;
+    for (unsigned dd = 0; dd < NUMDOFS; ++dd){
+      dot += pvalue.vv[dd] * dw[ii].vv[dd];
+    }
+    sum += 1./sigma * dot;
+  }
+  return sum;
+}
+
+void QuenchCorr::
+average ()
+{
+  std::cout << "quench: nframe is " << corr.getNframe() << std::endl;
+  corr.average();
+}

### 2013-06-10 one.test/template.optimal.ctrl/tools/simulator/src/Stopwatch.cpp

@@ -1,0 +1,29 @@
+#include "Stopwatch.h"
+#include <iostream>
+
+Stopwatch::Stopwatch ()
+    : HZi (1./HZ)
+{
+}
+
+// int main(int argc, char * argv[])
+// {
+//   Stopwatch sw;
+//   double a = 1000;
+//   double b = 19;
+//   double c;
+//   sw.start();
+//   for (double i = 0; i < 1e8; i = i + 1.) c = a * b;
+//   sw.stop();  
+//   std::cout << "user time " << sw.user() << std::endl;
+//   std::cout << "system time " << sw.system() << std::endl;
+//   std::cout << "real time " << sw.real() << std::endl;
+
+// //   sw.start();
+//   for (double i = 0; i < 1e8; i = i + 1.) c = a * b;
+//   sw.stop();  
+//   std::cout << "user time " << sw.user() << std::endl;
+//   std::cout << "system time " << sw.system() << std::endl;
+//   std::cout << "real time " << sw.real() << std::endl;
+  
+// }

### 2013-06-10 one.test/template.optimal.ctrl/tools/simulator/src/TimeCorrelation.cpp

@@ -1,0 +1,289 @@
+#include "TimeCorrelation.h"
+
+void TimeCorrelation::
+reinit (const double & step_,
+	const double & time_,
+	const double & x0,
+	const double & x1,
+	const unsigned & nx,
+	const double & v0,
+	const double & v1,
+	const unsigned & nv,
+	const Integrator * equiInte_,
+	const Integrator * quenchInte_,
+	const unsigned & quenchNumStep_,
+	const double & quenchAtTime_,
+	const DissipativeFlux * flux_)
+{
+  equiInte = equiInte_;
+  quenchInte = quenchInte_;
+  flux = flux_;
+  step = step_;
+  time = time_;
+  dt = equiInte->getDt();
+  if (step / dt < 1){
+    std::cerr << "step should be larger than dt" << std::endl;
+    exit (1);
+  }
+  nFrame = (time + dt) / step + 1;
+  nStep = nFrame - 1;
+  quenchNumStep = quenchNumStep_;
+  quenchAtTime = quenchAtTime_;
+  calCorrFeq = (step + .5 * dt) / dt;
+
+  dists0.resize (nFrame);
+  dists1.resize (nFrame);
+  
+  for (unsigned ii = 0; ii < nFrame; ++ii){
+    dists0[ii].reinit (x0, x1, nx, v0, v1, nv);
+    dists1[ii].reinit (x0, x1, nx, v0, v1, nv);
+  }
+}
+
+static inline unsigned
+decreasePosi (const vector<double > & ring,
+	      const unsigned &posi)
+{
+  int p (posi);
+  if ((--p) < 0){
+    p += int (ring.size());
+  }
+  return unsigned(p);
+}
+
+static inline unsigned
+increasePosi (const vector<double > & ring,
+	      const unsigned &posi)
+{
+  int p (posi);
+  if ((++p) >= int(ring.size())){
+    p -= int (ring.size());
+  }
+  return unsigned(p);
+}
+
+void TimeCorrelation::
+calCorr (const Dofs & initDof,
+	 const double & totalTime)
+{
+  unsigned countStep = 0;
+  Dofs dof (initDof);
+  vector<double > savedFlux (nFrame, 0.);
+  unsigned savePosi = 0;
+  unsigned countValidSaved = 0;
+  int printCount = 0;
+  
+  for (double now = 0; now <= totalTime; now += dt){
[…]

### 2013-06-10 alanine/template/set/parameters.sh
back up, del wrong files

@@ -33,6 +33,7 @@ pert_noSdRange=1.0		# nm
 pert_barostat=no		# Parrinello-Rahman or no
 pert_taup=2.0			# ps
 pert_resp_corr_range=0.25	# nm
+pert_num_branch_traj=5		# number of branching trajs at each frame.
 
 pert_parallel_num_pro=5
 pert_parallel_my_id=0

### 2013-06-10 alanine/template/set/run.pert.sh

@@ -32,30 +32,33 @@ do
     runid=`echo "$count % $pert_parallel_num_pro" | bc`
     test $runid -ne $pert_parallel_my_id && continue
     test ! -d $pert_main_dir && mkdir -p $pert_main_dir
-    my_dir=$pert_main_dir/pert.$count
-    if test -d $my_dir; then
-	echo "existing dir $my_dir, continue"
-	continue
-	rm -fr $my_dir
-    fi
-    cp -a $gro_dir $my_dir
-    cd $my_dir
-    rm -f run.log
-    set_parameters_pert grompp.mdp
-    start_time=`grep $count $cwd/$pert_equi_result/equi.frame | awk '{print $2}'`
-    echo "# run with command `which grompp`" &>> run.log
-    $grompp_command -t $cwd/$pert_equi_result/traj.trr -time $start_time &>> run.log
-    echo "# run with command `which mdrun`" &>> run.log
-    $mdrun_command &>> run.log
-
-    echo 3 14 | trjconv -center -pbc whole
-    mv -f trajout.xtc alanine.xtc
-    $cwd/tools/angles/evolve -f alanine.xtc -s angle.dat &> angle.log
-    rm -f traj.xtc traj.trr state*.cpt topol.tpr conf.gro index.ndx 
-    
-    cd $cwd
-    
-    echo "$my_dir/angle.dat" >> angle.name    
-    echo "$my_dir/gxs.out" >> gxs.name    
+    for jj in `seq 1 $pert_num_branch_traj`;
+    do
+	pjj=`printf %03d $jj`
+	my_dir=$pert_main_dir/pert.$count.$pjj
+	if test -d $my_dir; then
+	    echo "existing dir $my_dir, continue"
+	    continue
+	    rm -fr $my_dir
+	fi
+	cp -a $gro_dir $my_dir
+	cd $my_dir
+	rm -f run.log
+	set_parameters_pert grompp.mdp
+	start_time=`grep $count $cwd/$pert_equi_result/equi.frame | awk '{print $2}'`
+	echo "# run with command `which grompp`" &>> run.log
+	$grompp_command -t $cwd/$pert_equi_result/traj.trr -time $start_time &>> run.log
+	echo "# run with command `which mdrun`" &>> run.log
+	$mdrun_command &>> run.log
+	
+	echo 3 14 | trjconv -center -pbc whole
[…]

### 2013-06-10 writedown/linear.response/paper.tex

@@
 \end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Optimal Forcing}
+
+Let us remain in the setting of the last section and assume that we are interested in a finite dimensional family of forcings
+\[
+\mathcal{F}=\{F_e(t)=\sum_{k=1}^N a_k \Phi_k(t)\},
+\]
+with explicitly given basis functions $\Phi_k$, $k=1,\ldots,N$.
+We are interested in computing the gradient of the expectation value at time $T$ of the observable $\mathcal{O}$ wrt the coefficient vector $(a_k)_{k=1,\ldots,N}$ in the point $F_e(t)=\sum_{k=1}^N a_k \Phi_k(t)$.
+From Eq.
+(\ref{eqn:neq-response}) we get with $\Delta F_e(t)=\Phi_k(t)$:
+\[
+\frac{\partial O(T)}{\partial a_j} = \lim_{\eps\to 0}\frac{1}{\eps}( O_\eps(T) -
+  O_0(T))=\big\langle O[\vect X_s]\cdot G_j[\vect X_s]\big\rangle_{0,T},
+\]
+where $X_t$ and $O_0$ are paths and expectation value wrt. the dynamics \eqref{eqn:langevin-1b}--\eqref{eqn:langevin-2b} with $F_e(t)=\sum_{k=1}^N a_k \Phi_k(t)$ and $\eps=0$, and
+\[
+G_j[\vect X_t]
+  = \frac{1}\sigma\int_0^t
+  \Phi_j(s)\,\vect D(\vect q_s)\cdot\dd\vect W_s. 
+\]
+That is, we can compute all derivatives $\partial O(T)/\partial a_j$ from just one ensemble of paths of $X_t$.
+
+Now we consider the optimization problem
+\[
+O_{max}(T) = \max_{F_e\in \mathcal{F}} O[F_e](T),
+\]
+where $O[F](T)$ is the expectation value of the path observable $\mathcal{O}$ that results from \eqref{eqn:langevin-1b}--\eqref{eqn:langevin-2b} with $F_e=F$ and $\eps=0$.
+With the gradient being computed one can solve the optimization problem by means of gradient descent in coefficient space,
+\[
+a^{(i+1)}=a^{(i)} -\gamma \frac{\partial O(T)}{\partial a}.
+\]
 

### 2013-06-12 writedown/linear.response/paper.tex
new title and add abs

@@
 
-\title{Linear response theory for core set identification}
+% \title{Linear response theory for core set identification}
+\title{The response theory for the system under a non-equilibrium driving}
 \author{Han Wang}
@@
 \begin{abstract}
+  In this talk, we firstly review the traditional linear response
+  theory, and then generalize it to the system under a non-equilibrium
+  driving. We calculate not only the linear response term, but also
+  the higher order terms with respect to the perturbation to the
+  reference dynamics. The response theory is applied to investigate
+  the dynamics of core set that plays a central role in the Markov
+  State Model (MSM). The advantages and limitations are discussed from
+  the numerical tests.
   % In this draft, we are trying to study the core set based Markov

### 2013-06-12 one.test/template.optimal.ctrl/tools/simulator/app/splitSingleWell.timeCorr.cpp
working on the ctrl problem, not tested

@@ -1,231 +1,0 @@
-#include <algorithm>
-#include <fstream>
-#include <iomanip>
-#include <iostream>
-#include <iterator>
-#include <list>
-#include <map>
-#include <set>
-#include <sstream>
-#include <string>
-#include <queue>
-#include <vector>
-#include <cmath>
-
-#include "Force.h"
-#include "Perturbation.h"
-#include "Integrator.h"
-#include "Distribution.h"
-#include <boost/program_options.hpp>
-#include "RandomGenerator.h"
-#include "TimeCorrelation.h"
-#include "QuenchCorr.h"
-
-namespace po = boost::program_options;
-
-int main(int argc, char * argv[])
-{
-  double gamma;
-  double kT;
-  double T;
-  double dt;
-  double pertSt;
-  double sigma;
-  double kk;
-  double nst;
-  unsigned nstprint = 1;
-  double warmTime;
-  
-  double noneqTime;
-  double noneqCheckFeq;
-  double branchFeq;
-  double quenchTime;
-  double quenchTemperture;
-  double quenchkT;
-  double corrTime;
-  double corrStep;
-  // unsigned noneqCheckNumFeq;
-  unsigned branchNumFeq;
-  string sfile, lfile;
-  double x0, x1, v0, v1;
-  unsigned nx, nv;
-  
-  unsigned long seed;
-  
-  po::options_description desc ("Allow options");
-  desc.add_options()
-      ("help,h", "print this message")
-      ("dt,d", po::value<double > (&dt)->default_value(0.01), "time step [ps]")
-      ("nst,n", po::value<double > (&nst)->default_value(1000.), "number of time step")
-      ("print-feq,p", po::value<unsigned > (&nstprint)->default_value(10000), "print frequency")
-      ("gamma,g", po::value<double > (&gamma)->default_value(1.), "gamma [ps^-1]")
-      ("temperature,t", po::value<double > (&T)->default_value(300.), "temperature [K]")
-      ("single-well-k,k",po::value<double > (&kk)->default_value(8.0), "k parameter of the double single potential [kJ/(mol nm^2)]")
-      ("gaussian-sigma,a", po::value<double > (&sigma)->default_value(0.16), "sigma of Gaussian perturbation [nm]")
[…]

### 2013-06-12 one.test/template.optimal.ctrl/tools/simulator/app/tiltDoubleWell.ctr.cpp

@@ -30,24 +30,23 @@ int main(int argc, char * argv[])
   double kT;
   double T;
   double dt;
-  double pertSt0, pertSt1;
-  double sigma;
+  double pertSt0;
+  double aa;
   double kk;
   double nst;
   unsigned nstprint = 1;
-  double warmTime;
   
   double noneqTime;
   double noneqCheckFeq;
   double branchFeq;
-  double quenchTime;
-  double quenchTemperture;
-  double quenchkT;
   unsigned noneqCheckNumFeq;
   unsigned branchNumFeq;
   double x0, x1, v0, v1;
   unsigned nx, nv;
   int order;
+  double beta;
+  double timeResolution;
+  unsigned nTimeFrame;
 
   string sfile, lfile;
   
@@ -60,23 +59,21 @@ int main(int argc, char * argv[])
   desc.add_options()
       ("help,h", "print this message")
       ("dt,d", po::value<double > (&dt)->default_value(0.0001), "time step [ps]")
+      ("dt,d", po::value<double > (&beta)->default_value(1.0), "the punishment constant, unknown unit...")
       ("nst,n", po::value<double > (&nst)->default_value(10000), "number of time step")
       ("print-feq,p", po::value<unsigned > (&nstprint)->default_value(10000), "print frequency")
       ("gamma,g", po::value<double > (&gamma)->default_value(1.), "gamma [ps^-1]")
       ("temperature,t", po::value<double > (&T)->default_value(300.), "temperature [K]")
-      ("single-well-k,k",po::value<double > (&kk)->default_value(8.0), "k parameter of the double single potential [kJ/(mol nm^2)]")
-      ("gaussian-sigma,a", po::value<double > (&sigma)->default_value(0.16), "sigma of Gaussian perturbation [nm]")
+      ("double-well-k,k", po::value<double > (&kk)->default_value(8.0), "k parameter of the double well potential [kJ/(mol nm^2)]")
+      ("double-well-a,a", po::value<double > (&aa)->default_value(1.0), "a parameter of the double well potential [nm]")
       ("branch-feq", po::value<double > (&branchFeq)->default_value(1.), "branch frequency [ps]")
[…]

### 2013-06-12 one.test/template.optimal.ctrl/tools/simulator/app/tiltDoubleWell.noneqResp.cpp

@@ -1,304 +1,0 @@
-#include <algorithm>
-#include <fstream>
-#include <iomanip>
-#include <iostream>
-#include <iterator>
-#include <list>
-#include <map>
-#include <set>
-#include <sstream>
-#include <string>
-#include <queue>
-#include <vector>
-#include <cmath>
-
-#include "Force.h"
-#include "Perturbation.h"
-#include "Integrator.h"
-#include "Distribution.h"
-#include "NoneqResponseInfo.h"
-#include "Stopwatch.h"
-
-#include <boost/program_options.hpp>
-#include "RandomGenerator.h"
-
-namespace po = boost::program_options;
-
-int main(int argc, char * argv[])
-{
-  double gamma;
-  double kT;
-  double T;
-  double dt;
-  double pertSt0, pertSt1;
-  double aa;
-  double kk;
-  double nst;
-  unsigned nstprint = 1;
-  double warmTime;
-  
-  double noneqTime;
-  double noneqCheckFeq;
-  double branchFeq;
-  double quenchTime;
-  double quenchTemperture;
-  double quenchkT;
-  unsigned noneqCheckNumFeq;
-  unsigned branchNumFeq;
-  double x0, x1, v0, v1;
-  unsigned nx, nv;
-  int order;
-
-  string sfile, lfile;
-  
-  unsigned long seed;
-
-  // Stopwatch sw_eq, sw_noneq, sw_quench;
-  Stopwatch sw_total;
-  
-  po::options_description desc ("Allow options");
-  desc.add_options()
-      ("help,h", "print this message")
-      ("dt,d", po::value<double > (&dt)->default_value(0.0001), "time step [ps]")
-      ("nst,n", po::value<double > (&nst)->default_value(10000), "number of time step")
-      ("print-feq,p", po::value<unsigned > (&nstprint)->default_value(10000), "print frequency")
-      ("gamma,g", po::value<double > (&gamma)->default_value(1.), "gamma [ps^-1]")
-      ("temperature,t", po::value<double > (&T)->default_value(300.), "temperature [K]")
-      ("double-well-k,k", po::value<double > (&kk)->default_value(8.0), "k parameter of the double well potential [kJ/(mol nm^2)]")
[…]

### 2013-06-12 one.test/template.optimal.ctrl/tools/simulator/app/tiltDoubleWell.timeCorr.cpp

@@ -1,231 +1,0 @@
-#include <algorithm>
-#include <fstream>
-#include <iomanip>
-#include <iostream>
-#include <iterator>
-#include <list>
-#include <map>
-#include <set>
-#include <sstream>
-#include <string>
-#include <queue>
-#include <vector>
-#include <cmath>
-
-#include "Force.h"
-#include "Perturbation.h"
-#include "Integrator.h"
-#include "Distribution.h"
-#include <boost/program_options.hpp>
-#include "RandomGenerator.h"
-#include "TimeCorrelation.h"
-#include "QuenchCorr.h"
-
-namespace po = boost::program_options;
-
-int main(int argc, char * argv[])
-{
-  double gamma;
-  double kT;
-  double T;
-  double dt;
-  double pertSt;
-  double aa;
-  double kk;
-  double nst;
-  unsigned nstprint = 1;
-  double warmTime;
-  
-  double noneqTime;
-  double noneqCheckFeq;
-  double branchFeq;
-  double quenchTime;
-  double quenchTemperture;
-  double quenchkT;
-  double corrTime;
-  double corrStep;
-  // unsigned noneqCheckNumFeq;
-  unsigned branchNumFeq;
-  string sfile, lfile;
-  double x0, x1, v0, v1;
-  unsigned nx, nv;
-  
-  unsigned long seed;
-  
-  po::options_description desc ("Allow options");
-  desc.add_options()
-      ("help,h", "print this message")
-      ("dt,d", po::value<double > (&dt)->default_value(0.01), "time step [ps]")
-      ("nst,n", po::value<double > (&nst)->default_value(1000.), "number of time step")
-      ("print-feq,p", po::value<unsigned > (&nstprint)->default_value(10000), "print frequency")
-      ("gamma,g", po::value<double > (&gamma)->default_value(1.), "gamma [ps^-1]")
-      ("temperature,t", po::value<double > (&T)->default_value(300.), "temperature [K]")
-      ("double-well-k,k", po::value<double > (&kk)->default_value(8.0), "k parameter of the double well potential [kJ/(mol nm^4)]")
-      ("double-well-a,a", po::value<double > (&aa)->default_value(1.0), "a parameter of the double well potential [nm]")
[…]

### 2013-06-12 one.test/template.optimal.ctrl/tools/simulator/include/NoneqResponseInfo.h

@@ -12,43 +12,37 @@ class NoneqResponseInfo
 {
   double x0, x1, v0, v1;
   unsigned nx, nv;
+
+  double beta;
   
   double dt;
   double noneqTime;
   double noneqCheckFeq;
   int numCheck;
   int checkNumFeq;
-  double quenchTime;
-  int quenchNumStep;
 
   int countNoneq;
   int countNoneqSeg;
-  int countQuench;
 
   const Perturbation * ppert;
   int numMode;
   
-  vector<Distribution_1d >			order0;		// time
-  vector<vector<Distribution_1d > >		order1;		// time x vec
-  vector<vector<vector<Distribution_1d > > >	order2;		// time x mat  
-  
-  vector<Distribution_1d >			quench_order0;		// time
-  vector<Distribution_1d >			quench_order1_term1;	// time
-  vector<vector<Distribution_1d > >		quench_order1_term2;	// time x vec
-  vector<Distribution_1d >			quench_order2_term1;	// time
-  vector<vector<vector<Distribution_1d > > >	quench_order2_term2;	// time x mat
-  vector<vector<Distribution_1d > >		quench_order2_term3;	// time x vec
+  vector<double >				order0;		// time
+  vector<double >				order0punish;	// time
+  vector<vector<double > >			order1;		// time x vec
+  vector<vector<vector<double > > >		order2;		// time x mat  
   
   vector<double >				Gj;		// vec 
-  vector<vector<double > >			Hjk;		// mat 
-  double					G0;		// 1
-  double					H00;		// 1
+  vector<vector<double > >			Hjk;		// mat
 
+  double					punish;
+  int						ntraj;
 public:
   NoneqResponseInfo ();
   // ~NoneqResponseInfo ();
 public:
-  void reinit (const double & x0,
+  void reinit (const double & beta,
+	       const double & x0,
 	       const double & x1,
 	       const unsigned & nx,
 	       const double & v0,
@@ -57,28 +51,23 @@ public:
 	       const double & dt,
 	       const double & noneqTime,
 	       const double & noneqCheckFeq,
-	       const double & quenchTime,
 	       const Perturbation & pert);
+  int inSet (const Dofs & x) ;
   void newTraj ();
-  void newQuenchTraj ();
   void depositMainTraj (const Dofs & oldx,
 			const Dofs & newx,
 			const double & sigma,
 			const Dofs & dw);
-  void depositQuenchTraj (const Dofs & oldx,
[…]

### 2013-06-12 one.test/template.optimal.ctrl/tools/simulator/include/Perturbation.h

@@ -3,7 +3,11 @@
 
 #include "Defines.h"
 #include "Force.h"
+#include "Poly.h"
+#include "Interpolation.h"
 #include <vector>
+#include <fstream>
+#include <string>
 
 using namespace std;
 
@@ -46,6 +50,31 @@ public:
   PertConstTilt (const double & s,
 		 const double warmTime = 0.);
 public:
+  virtual double Fe (const double & time) const;
+  virtual int numMode () const;
+  virtual void FeMode (const double & time,
+		       vector<double > & modes) const;
+  virtual void FeModePref (vector<double > & pref ) const;
+public:
+  virtual void operator () (const Dofs & dofs,
+  			    Dofs & pvalue) const;
+  virtual void operator () (const Dofs & dofs,
+			    const double & time,
+			    Dofs & pvalue) const;
+}
+    ;
+
+
+class PertConstTiltTable : public Perturbation
+{
+  PiecewisePoly pwl;
+public:
+  PertConstTiltTable (const double & s,
+		      const vector<double > & xx,
+		      const vector<double > & vv);
+public:
+  void reinit (const vector<double > & xx,
+	       const vector<double > & vv);
   virtual double Fe (const double & time) const;
   virtual int numMode () const;
   virtual void FeMode (const double & time,

### 2013-06-12 one.test/template.optimal.ctrl/tools/simulator/include/Poly.h

@@ -1,0 +1,77 @@
+#ifndef __wanghan_Poly_h__
+#define __wanghan_Poly_h__
+
+#include <iostream>
+#include <string>
+#include <vector>
+
+class Poly
+{
+  std::vector<double > a;
+  unsigned order;
+public:
+  Poly ();
+  Poly (const std::vector<double > & out);
+  void reinit (const std::vector<double > & out);
+  void zero () {a.clear(); a.resize(1,0); order = 0;}
+  void one  () {a.clear(); a.resize(1,1); order = 0;}
+public:
+  Poly & operator = (const Poly & poly);
+  Poly & operator += (const Poly & poly);
+  Poly & operator += (const double & b);
+  Poly & operator *= (const Poly & poly);
+  Poly & operator *= (const double & scale);
+  Poly & derivative ();
+public:
+  unsigned & getOrder () {return order;}
+  const unsigned & getOrder () const {return order;}
+  std::vector<double > & getCoeffs () {return a;}
+  const std::vector<double > & getCoeffs () const {return a;}
+public:
+  void print ();
+  void print (const std::string & x);
+  void printCode (const std::string & x);
+public :
+  double value ( const double & x ) const;
+public:
+  // p = f(ax + b)
+  Poly & valueLinearPoly (const double & a, const double & b,
+			  Poly & p);
+}
+    ;
+
+class PiecewisePoly
+{
+  std::vector<double > x;
+  std::vector<Poly > p;
+private:
+  void value (const unsigned & xbegin,
+	      const unsigned & xend,
+	      const std::vector<double > & r,
+	      const unsigned & rbegin,
+	      const unsigned & rend,
+	      std::vector<double > & y) const;
+  double value (const double & xx,
+		unsigned & begin,
+		unsigned & end) const;  
+public:
+  std::vector<double > & get_x () {return x;}
+  std::vector<Poly   > & get_p () {return p;}
+  const std::vector<double > & get_x () const {return x;}
+  const std::vector<Poly   > & get_p () const {return p;}
+public:
+  void clear () {x.clear(); p.clear();}
+  bool valid () const;
+public:
+  double value (const double & r) const;
+  void   value (const std::vector<double > & r,
+		std::vector<double > & y) const;
+  double value_periodic (const double & r) const;
+  void   value_periodic (const std::vector<double > & r,
+			 std::vector<double > & y) const;
+}
+    ;
+
+
+
+#endif

### 2013-06-12 one.test/template.optimal.ctrl/tools/simulator/src/Perturbation.cpp

@@ -1,4 +1,5 @@
 #include "Perturbation.h"
+#include <algorithm>
 
 DissipativeFlux::
 DissipativeFlux (const Perturbation * p,
@@ -97,6 +98,75 @@ FeModePref (vector<double > & pref ) const
   pref.resize(1);
   pref[0] = strength;
 }
+
+
+
+
+PertConstTiltTable::
+PertConstTiltTable (const double & s,
+		    const vector<double > & xx,
+		    const vector<double > & vv)
+{
+  reinit (xx, vv);
+}
+
+void PertConstTiltTable::
+reinit (const vector<double > & xx,
+	const vector<double > & vv)
+{
+  Interpolation::piecewiseLinear (xx, vv, pwl);
+}
+
+
+void PertConstTiltTable::
+operator () (const Dofs & dofs,
+	     const double & time,
+	     Dofs & pvalue) const
+{
+  (*this) (dofs, pvalue);
+  pvalue.vv[0] *= Fe(time);
+}
+
+void PertConstTiltTable::
+operator () (const Dofs & dofs,
+	     Dofs & pvalue) const
+{
+  pvalue.vv[0] = 1.0;
+}
+
+double PertConstTiltTable::
+Fe (const double & time) const
+{
+  return pwl.value (time);
+}
+
+int PertConstTiltTable::
+numMode () const
+{
+  return pwl.get_x().size();
+}
+
+void PertConstTiltTable::
+FeMode (const double & time,
+	vector<double > & modes) const
+{
+  modes.resize(numMode());
+  std::fill (modes.begin(), modes.end(), 0.0);
+  int index = int(time / (pwl.get_x()[1] - pwl.get_x()[0]));
+  if (index == numMode() - 1) index --;
+  if (index >= 0 && index < numMode() - 1){
+    modes[index  ] = (pwl.get_x()[index+1] - time) / (pwl.get_x()[index+1] - pwl.get_x()[index]);
+    modes[index+1] = (time - pwl.get_x()[index  ]) / (pwl.get_x()[index+1] - pwl.get_x()[index]);
+  }
+}
+
+void PertConstTiltTable::
+FeModePref (vector<double > & pref ) const
+{
+  pref.resize(numMode());
+  pwl.value (pwl.get_x(), pref);
+}
+
 
 
 

### 2013-06-19 writedown/linear.response/paper.tex
add numerical result for optimal forcing. better notation for the magnitude of F_e

@@
   \dot{\vect p} & =- \nabla_{\vect q}\mh(\vect q,\vect p)
-  + (k_0 F_e(t)+\eps \Delta F_e(t)) \vect D(\vect q)
+  + (F_e(t)+\eps \Delta F_e(t)) \vect D(\vect q)
   - \gamma\vect p
@@
 such that for $\eps=0$ we get a driven system that is further
-perturbed by the addition time-dependent driving force $\eps \Delta
+perturbed by the addition time-dependent driving force $ \Delta
 F_e(t) \vect D(\vect q)$. Let $\vect X^\eps_t$ denote the associated solution process, and
@@
 
-\section{Optimal Forcing}
+\subsection{Optimal Forcing}
 
@@
 \[
-O_{max}(T) = \max_{F_e\in \mathcal{F}} O[F_e](T),
+O_{min}(T) = \min_{F_e\in \mathcal{F}} O[F_e](T),
 \]
@@
     theory developed by this work.  We calculate response terms up to
-    the second order from an equilibrium simulation (i.e. letting $k_0
+    the second order from an equilibrium simulation (i.e. letting $k_e
     = 0$ and $\eps = 0$ in Eq.~\eqref{eqn:langevin-2b}).
@@
     here we calculate the linear response from a non-equilibrium simulation
-    of strength 2 (i.e. $k_0 = 2$ and $\eps = 4$
+    of strength 2 (i.e. $k_e = 2$ and $\eps = 4$
     in Eq.~\eqref{eqn:langevin-2b}).
@@
 The strength of $F_e(t)$ is
-set to be 1, when the time $t$ is larger than some warm-up time $t_c$.
+set to be $k_e$, when the time $t$ is larger than some warm-up time $t_c$.
 When $t \leq t_c$, the strength of $F_e(t)$ is linearly growing,
-i.e. $ F_e(t) = t/t_c$.
+i.e. $ F_e(t) = k_e\cdot t/t_c$.
 We use the value $t_c = 20$~\textsf{ps}, so that the perturbation increases slow
@@
 a intermediate
-non-equilibrium simulation with driving force of $k_0 = 2$.
+non-equilibrium simulation with driving force of $k_e = 2$.
 Notice that the reference simulation in this case is also a
@@
     theory developed by this work.  We calculate response terms up to
-    the second order from an equilibrium simulation (i.e. letting $k_0
+    the second order from an equilibrium simulation (i.e. letting $k_e
     = 0$ and $\eps = 2$ in Eq.~\eqref{eqn:langevin-2b}).
@@
     here we calculate the linear response from a non-equilibrium simulation
-    of strength 2 (i.e. $k_0 = 1$ and $\eps = 1$
+    of strength 2 (i.e. $k_e = 1$ and $\eps = 1$
     in Eq.~\eqref{eqn:langevin-2b}).
@@
 On the 4th row of Fig.~\ref{fig:tmp6} presents the linear response starting
-from a non-equilibrium simulation with external driving $k_0 = 1$.
+from a non-equilibrium simulation with external driving $k_e = 1$.
 In this case the effective strength of the perturbation to the reference
@@
 non-equilibrium simulation.
+
+\subsection{Optimal forcing}
+\begin{figure}
+  \centering
+  \includegraphics[]{figs/fig-ctr-stat.eps}
+  \caption{The optimal forcing.
+    The red color indicates $\eta = 0.01$ and the
+    green color indicates $\eta = 0.05$.
+    In the lower plot, the dashed lines display the
+    value of $P(\vect X_t)$ as the time eloves.
+    The solid lines present $O[F_e](t)$.  In the upper
+    plot, the solid lines present the magnitude of the
+    control $F_e(t)$.
+  }\label{fig:tmp7}
+\end{figure}
+
+We investigate the example of the double well potential, and
+consider the problem of optimally forcing the population from the
+left well to the right well. That is equal to consider the 
+the following optimal forcing problem:
+\begin{align}
+  O_{\min}(T)& = \min_{F_e\in\mathcal F} O[F_e](T) 
+\end{align}
+with
+\begin{align}
+  O[F_e](T) &= P(\vect X_T) + \eta \int_0^T\vert F_e(t)\vect D(\vect q)\vert^2\, dt
+\end{align}
+where $P(\vect X_T)$ is equal to 1 when the end point of the
+trajectory $\vect X_T$ is in the right well, and equal to 0 otherwise.
+The integral is the ``cost'' of the control and $\eta$ is a constant
+that indicates the relative magnitude of the cost. We consider the
+ansatz function family $\mathcal F$ of piecewise linear functions on
+$[0, T]$. 
+Fig.~\ref{fig:tmp7} presents the numerical result. The magnitude of control
+is shown in the upper plot, and the value of $O[F_e](T)$ and $P(\vect X_T)$
+are given in the lower plot.
 

### 2013-06-20 writedown/linear.response/paper.tex
add observation from optimal forcing

@@
   \exp
-  \big\{{\eps G[\vect X_s] - \eps^2 H[\vect X_s]}
+  \big\{{\eps G[\vect X_s] - \frac{\eps^2}{2} H[\vect X_s]}
   \big\}
@@
   &=
-  \frac 1{2\sigma^2}\int_0^t
+  \frac 1{\sigma^2}\int_0^t
   \Delta F_e^2(s)\,\vect D^2(\vect q_s)\,\dd s.
@@
 \[
-O_{min}(T) = \min_{F_e\in \mathcal{F}} O[F_e](T),
+O_{max}(T) = \max_{F_e\in \mathcal{F}} O[F_e](T),
 \]
@@
   \includegraphics[]{figs/fig-ctr-stat.eps}
-  \caption{The optimal forcing.
-    The red color indicates $\eta = 0.01$ and the
-    green color indicates $\eta = 0.05$.
-    In the lower plot, the dashed lines display the
-    value of $P(\vect X_t)$ as the time eloves.
-    The solid lines present $O[F_e](t)$.  In the upper
-    plot, the solid lines present the magnitude of the
-    control $F_e(t)$.
-  }\label{fig:tmp7}
+  \caption{The optimal forcing.  In this plot, $T=20$ ps.  The red
+    color indicates $\eta = 0.01$ and the green color indicates $\eta
+    = 0.05$.  In the upper plot, the solid lines present the magnitude
+    of the control $F_e(t)$, which is calculated on a family of
+    piecewise linear ansatz functions. The time interval of a linear
+    piece is 1 ps.  In the lower plot, the dashed lines display the
+    value of $P(\vect X_t)$ as the time evolves.  The solid lines
+    present $O[F_e](t)$.  }\label{fig:tmp7}
 \end{figure}
@@
 \begin{align}
-  O_{\min}(T)& = \min_{F_e\in\mathcal F} O[F_e](T) 
+  O_{\max}(T)& = \max_{F_e\in\mathcal F} O[F_e](T) 
 \end{align}
@@
 \begin{align}
-  O[F_e](T) &= P(\vect X_T) + \eta \int_0^T\vert F_e(t)\vect D(\vect q)\vert^2\, dt
-\end{align}
-where $P(\vect X_T)$ is equal to 1 when the end point of the
+  O[F_e](T) &= P(\vect X_T) - \eta \int_0^T\vert F_e(t)\vect D(\vect q)\vert^2\, dt
+\end{align}
+where $P(\vect X_T)$ is equal to $1$ when the end point of the
 trajectory $\vect X_T$ is in the right well, and equal to 0 otherwise.
-The integral is the ``cost'' of the control and $\eta$ is a constant
-that indicates the relative magnitude of the cost. We consider the
-ansatz function family $\mathcal F$ of piecewise linear functions on
-$[0, T]$. 
-Fig.~\ref{fig:tmp7} presents the numerical result. The magnitude of control
-is shown in the upper plot, and the value of $O[F_e](T)$ and $P(\vect X_T)$
-are given in the lower plot.
+It is clear that for an unbiased double-well, the average of $P(\vect
+X_T)$ is 0.5.  The integral is the ``cost'' of the control and $\eta$
+is a positive constant that indicates the relative magnitude of the
+cost. We consider the ansatz function family $\mathcal F$ of piecewise
+linear functions on $[0, T]$.
+
+Fig.~\ref{fig:tmp7} presents the numerical result. The magnitude of
+control $F_e(t)$ is shown in the upper plot, and the value of
+optimization target $O[F_e](T)$ and the value of $P(\vect X_T)$ are
+given in the lower plot by the solid and dashed lines, respectively.
+The plot shows that with a smaller $\eta$ (the relative magnitude of
+the cost of the control is smaller), a stronger control is calculated
+from the optimization, and the optimized value of $O[F_e](T)$ and
+$P(\vect X_T)$ are bigger, which means the end position of $\vect X_t$
+is more biased towards the right well. This observation is consistent
+with intuition.  When $t<10$~ps, the magnitude of the control is
+almost 0.  Near the end time $T$, the magnitude of the control firstly
+quickly goes up, and then falls down a little bit.  This implies some
+interesting information.  If the system were able to immediately relax
+to its instantaneous equilibrium state (sometimes called
+quasi-equilibrium), the population in the right well (dashed lines in
+Fig.~\ref{fig:tmp7}) would immediately go down, as the control
+decreases.  The fact that this does not happen, indicates that the
+speed of changing the control is relatively faster that the internal
+time scale of the system, so the system does not have enough time to
+fully relax itself. Therefore, the observed phenomenon is truly
+non-equilibrium, and our non-equilibrium linear response theory is a
+tool that facilitates the investigation of this optimal forcing
+problem in the non-equilibrium setting.  Intuitively, one concludes
+that as the control goes up, the population in the right well actually
+does not have enough time to go as high as it could be, so when the
+control decreases, it continues going up rather than following the
+decreasing control.
 

### 2013-06-26 writedown/linear.response/paper.tex
add finite diff SD

@@
 
-\section{Numerical tests: the core set identification of  the Markov State Model }
-
-The core set plays a very important role in the MSM theory.  We
-therefore want to test our idea of non-equilibrium response theory by
-identifying the core sets in a non-equilibrium driven system.
-There are several ways to define a core set.
-One possible choice  is to calculate the stability of the phase space
-distribution: $\Delta \mt_\tau f(\vect x, t) = (\mt_\tau -\mi) f(\vect
-x, t)$. And the core sets are defined by
-\begin{align}
-  C_\tau(t) = \{\vect x\vert \Delta \mt_\tau f(\vect x, t) \geq 0\}
-\end{align}
-Here the propagator $\mt_\tau$ is defined by $\mt_\tau =
-e^{\tau\fwg_t^{2\beta}}$.  Notice that here the physical time $t$ is
+\section{Numerical tests}
+
+\subsection{The metastability of a single-well potential}
+The metastability of the system at time $t$ can be investigated by
+calculating the function $\Delta \mt_\tau f(\vect x, t) = (\mt_\tau
+-\mi) f(\vect x, t)$, where $f(\vect x, t)$ is the probability
+distribution at position $\vect x$ and time $t$, the propagator
+$\mt_\tau$ is defined by $\mt_\tau = e^{\tau\fwg_t^{2\beta}}$, and
+$\mi$ is the identity.  Notice that here the physical time $t$ is
 frozen, and the distribution $f$ evolves a long the artificial time
-$\tau$ at temperature $1/(2\beta)$. The artificial time $\tau$ should
-be chosen smaller than the investigated relaxation time scales of the
-original system. However, on the other hand, $\tau$ cannot be
-arbitrarily small, because otherwise the statistical uncertainty would
-not allow us to distinguish $\mt_\tau f$ and $f$.
-
-\subsection{The tilting double-well potential}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.4\textwidth]{figs/fig-tilt-pot.eps}
-  \caption{The double-well potential with tilting perturbation at
-    time $t\geq 20$ ps..}
-  \label{fig:tmp1}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.95\textwidth]
-  {figs/fig-tilt-str2-2d-more.eps}
-  \caption{The plot of $\Delta\mt_\tau f(\vect x,t)$ under
-    perturbation with strength $\eps = 2$.  From left to right the
-    columns present results at time $t = 0$, 10, 15 and
-    20~\textsf{ps}.  The 1st row: the direct non-equilibrium
-    simulation. The 2nd row: the results of traditional equilibrium
-    linear response theory.  The 3rd row: the results of the response
-    theory developed by this work.  We calculate response terms up to
-    the second order from an equilibrium simulation (i.e. letting $k_e
-    = 0$ and $\eps = 0$ in Eq.~\eqref{eqn:langevin-2b}).
-  }
-  \label{fig:tmp2}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.95\textwidth]
-  {figs/fig-tilt-str4-2d-more.eps}
-  \caption{The plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
-    $\eps = 4$. From left to right the
-    columns present results at time $t = 0$, 10, 15 and
-    20~\textsf{ps}.
-    The first three rows use the same algorithm as Fig.~\ref{fig:tmp2}.    
-    The 4th row: the results of the response theory developed by this work.
-    Instead of calculating the responses from the equilibrium simulation,
-    here we calculate the linear response from a non-equilibrium simulation
-    of strength 2 (i.e. $k_e = 2$ and $\eps = 4$
-    in Eq.~\eqref{eqn:langevin-2b}).
-    Therefore, the effective perturbation in this case is only
-    $\eps = 2$.
-  }
-  \label{fig:tmp3}
-\end{figure}
-
-We test the idea firstly by a one-dimensional model system: one particle in a
-tilting double-well potential. For convenience, we let the mass of the
-particle to be 1~\textsf{amu},
-and the friction coefficient to be $1\,\textsf{ps}^{-1}$.
-The unperturbed
-Hamiltonian of the system is given by:
-\begin{align}
-  \mh (\vect p, \vect q) = \frac 12 \vect p^2 + U(\vect q) 
-\end{align}
-with potential
-\begin{align}
-  U(\vect q) = \frac12 k (\vect q^2 - a^2)^2
-\end{align}
-Here $k = 8$~$\textsf{kJ} / (\textsf{mol nm}^4)$, and $ a = 1\ \textsf{nm}$.
-At room temperature $300\ \textsf{K}$, $k_BT = 2.48$~\textsf{kJ/mol}.
-The red line in Fig.~\ref{fig:tmp1} presents the unperturbed potential.
-%$\textsf{kJ} / (\textsf{mol nm}^4)}$
-The perturbation is given by
-\begin{align}
-  \vect D(\vect q) = -\nabla_{\vect q} V(\vect q) = 1
-\end{align}
-Here $V(\vect q) = -\vect q$   effectively tilts the original
-potential $U(\vect q)$ (see the gree and blue lines in Fig.~\ref{fig:tmp1}).
-The strength of $F_e(t)$ is
-set to be $k_e$, when the time $t$ is larger than some warm-up time $t_c$.
-When $t \leq t_c$, the strength of $F_e(t)$ is linearly growing,
-i.e. $ F_e(t) = k_e\cdot t/t_c$.
-We use the value $t_c = 20$~\textsf{ps}, so that the perturbation increases slow
-enough: the typical decaying time scale of the correlation function
-in~\eqref{eqn:eq-lr} is roughly 3~\textsf{ps}.
-We choose $\tau = 1$~\textsf{ps}, and
-% We choose $\mt_\alpha$
-% to be the propagator of the Langevin dynamics at temperature $150$
-% \textsf{K} with $\alpha = 1$ \textsf{ps}.
-consider the stability of the
-distribution $f(\vect x, t)$ at time $t$: $\Delta\mt_\tau f(\vect x, t)$.
-The larger
-this value, the more stable the core sets are.
-
-See Fig.~\ref{fig:tmp2} and \ref{fig:tmp3} for numerical result.  Both
-the results of the brute-force non-equilibrium simulation
-and that of the response theory are presented.
-In Fig.~\ref{fig:tmp2},
-the good agreement between the brute-force (1st row)
-and the linear response result (2nd row) indicates that $\eps = 2$ is
-indeed a relatively small perturbation to the equilibrium.
-In the third row of Fig.~\ref{fig:tmp2},
-the second order response is considered. We do not see any
-substantial improvement of the result by including this higher order term,
-because the
-driving force to the system is small,
-and the linear term is anyway a good approximation.
-
-In Fig.~\ref{fig:tmp3}, as the
-perturbation grows stronger, the left well becomes less stable, while
-the right well becomes more stable.  When the perturbation is as strong as
-$\eps = 4$, the linear response result (2nd row) is no
-longer precise:
-At $t\geq 20$ \textsf{ps}, the left well basically
-disappears, however, the linear response calculation presents
-artificial structure (that is a small artificial
-stable region accompanied by a small unstable region)
-at the position of the left well.
-The second order response (3rd row) improves the accuracy a little, 
-however, it also includes artificial finer-grained oscillation in the profile.
-At the same time, the numerical uncertainty is higher, so the figure
-looks noisy.
-In the 4th row of Fig.~\ref{fig:tmp3}, we present the linear response result
-that calculate from 
-a intermediate
-non-equilibrium simulation with driving force of $k_e = 2$.
-Notice that the reference simulation in this case is also a
-non-equilibrium simulation, so the effective strength of perturbation
-is reduced to $\eps = 2$. According to Fig.~\ref{fig:tmp1}, this can be
-considered as a small perturbation.
-A substantial improvement is achieved: the profile
-is correctly calculated, and the statistical error is small.
-
-% the 1st order result of the
-% response formula~\eqref{eqn:pert-approx-1} and
-% \eqref{eqn:pert-approx-2}, with
-% respect to the reference state $F_e^{\textrm{max}} = 2$. The result
-% is impressively improved compared with the linear response with respect to the
-% equilibrium state.
-
-\subsection{Splitting single-well potential}
+$\tau$ at temperature $1/(2\beta)$.  If the position $\vect x$ is
+``stable'', then intuitively the probability of the system being at
+position $\vect x$ will be higher when the temperature is lower, i.e.,
+$\Delta \mt_\tau f(\vect x, t) > 0$.  Otherwise, when position $\vect
+x$ is not stable, $\Delta \mt_\tau f(\vect x, t) < 0$.  The
+artificial time $\tau$ should be chosen smaller than the investigated
+relaxation time scales of the original system. However, on the other
+hand, $\tau$ cannot be arbitrarily small, because otherwise the
+statistical uncertainty would not allow us to distinguish $\mt_\tau f$
+and $f$.
 
@@
   \includegraphics[width=0.4\textwidth]{figs/fig-split-pot.eps}
-  \caption{The single-well potential with splitting perturbation at
-    time $t\geq 20$ ps.}
+  \caption{The single-well potential with splitting splitting drving
+    force at time $t_c = 20$ ps.  Since the drving force is
+    conservative ($\vect D(\vect q) = -\nabla_{\vect q}V(\vect q)$),
+    we plot the non-equilibrium drving energy of the system.  The red,
+    green and blue lines are the potential energy without
+    non-equilibrium driving, with non-equilibrium driving and
+    the non-equilibrium driving perturbed by $\eps\Delta F_e(t)$,
+    respectively.  }
   \label{fig:tmp4}
@@
 
-% \begin{figure}
-%   \centering
-%   \includegraphics[width=0.95\textwidth]
-%   {figs/fig-split-str1-2d-more.eps}
-%   \caption{The plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
-%     $F_e^{\textrm{max}} = 1.0$.
-%     The 1st row: the direct non-equilibrium
-%     simulation. The 2nd row: the results of linear response
-%     formula~\eqref{eqn:core-identify-approx-2}.
-%     The 3rd row: the results of response
-%     formula~\eqref{eqn:pert-approx-1} and \eqref{eqn:pert-approx-2}.
-%     The order is 2, the reference state is $F_e^{\textrm{max}} = 0$, i.e.
-%     the equilibrium state.
-%   }
-%   \label{fig:tmp5}
-% \end{figure}
+% The core set plays a very important role in the MSM theory.  We
+% therefore want to test our idea of non-equilibrium response theory by
+% identifying the core sets in a non-equilibrium driven system.
+% There are several ways to define a core set.
+% One possible choice  is to calculate the stability of the phase space
+% distribution: $\Delta \mt_\tau f(\vect x, t) = (\mt_\tau -\mi) f(\vect
+% x, t)$. And the core sets are defined by
+% \begin{align}
+%   C_\tau(t) = \{\vect x\vert \Delta \mt_\tau f(\vect x, t) \geq 0\}
+% \end{align}
+% Here the propagator $\mt_\tau$ is defined by $\mt_\tau =
+% e^{\tau\fwg_t^{2\beta}}$.  Notice that here the physical time $t$ is
+% frozen, and the distribution $f$ evolves a long the artificial time
+% $\tau$ at temperature $1/(2\beta)$. The artificial time $\tau$ should
+% be chosen smaller than the investigated relaxation time scales of the
+% original system. However, on the other hand, $\tau$ cannot be
+% arbitrarily small, because otherwise the statistical uncertainty would
+% not allow us to distinguish $\mt_\tau f$ and $f$.
+
+We use the idea of non-equilibrium linear response to investigate the
+metastability a one-dimensional model system: one particle in a
+splitting single-well potential.  For convenience, we let the mass of
+the particle to be 1~\textsf{amu}, and the friction coefficient to be
+$1\,\textsf{ps}^{-1}$.
+The temperature is the
+room temperature $300\ \textsf{K}$, $k_BT = 2.48$~\textsf{kJ/mol}.
+The unperturbed Hamiltonian of the system is
+given by:
+\begin{align}
+  \mh (\vect p, \vect q) = \frac 12 \vect p^2 + U(\vect q) 
+\end{align}
+with potential
+\begin{align}
+  U(\vect q) = \frac12\,k\,\vect q^2 
+\end{align}
+Here $k = 8$~$\textsf{kJ} / (\textsf{mol nm}^2)$.
+See the red line in Fig.~\ref{fig:tmp4} for the  potential $U$.
+%$\textsf{kJ} / (\textsf{mol nm}^4)}$
+The non-equilibrium driving $\vect D$ is given by a conservative force:
+\begin{align}
+  \vect D(\vect q) = -\nabla_{\vect q} V(\vect q) 
+\end{align}
+where the driving potential $V(\vect q)$ has a Gaussian profile:
+\begin{align}
+  V(\vect q) = \frac{1}{\sqrt{2\pi \sigma^2}}
+  \exp\Big\{-\frac{\vect q^2}{2\sigma^2}\Big\}
+\end{align}
+we use $\sigma = 0.16$~\textsf{nm}.  The strength of non-equilibrium
+driving $F_e(t)$ is set to be a unitless constant $k_e$, when the time $t$ is
+larger than some warm-up time $t_c$.  When $t \leq t_c$, the strength
+of $F_e(t)$ is linearly growing, i.e. $ F_e(t) = k_e\cdot t/t_c$.  We
+consider the perturbation to the system given by $\eps\Delta F_e(t) =
+\eps$ when $t> t_c$, and $\eps\Delta F_e(t) = \eps \cdot t/t_c$ when
+$t\leq t_c$.
+We consider the following parameters: $t_c = 20$~\textsf{ps}, 
+$k_e = 1$ and $\eps = 1$, see Fig.~\ref{fig:tmp4}
+for the non-equilibrium driving potential and perturbed
+potential at time $t = t_c$.
+$\tau$ is set to be $1~\textsf{ps}$.
 
@@
   {figs/fig-split-str2-2d-more.eps}
-  \caption{
-    The plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
-    $\eps = 2$. From left to right the
-    columns present results at time $t = 0$, 5, 10 and
-    20~\textsf{ps}.
-    The 1st row: the direct non-equilibrium
-    simulation. The 2nd row: the results of traditional equilibrium
-    linear response theory.  The 3rd row: the results of the response
-    theory developed by this work.  We calculate response terms up to
-    the second order from an equilibrium simulation (i.e. letting $k_e
-    = 0$ and $\eps = 2$ in Eq.~\eqref{eqn:langevin-2b}).
-    The 4th row: the results of the response theory developed by this work.
-    Instead of calculating the responses from the equilibrium simulation,
-    here we calculate the linear response from a non-equilibrium simulation
-    of strength 2 (i.e. $k_e = 1$ and $\eps = 1$
-    in Eq.~\eqref{eqn:langevin-2b}).
-    Only the linear term of response is calculated.
-    Notice, the effective perturbation in this case is only
-    $\eps = 1$.
-  }
+  \caption{ The plot of $\Delta\mt_\tau f(\vect x,t)$ in the phase
+    space under perturbed non-equilibrium driving ($k_e = 1$ and $\eps
+    = 1$). From left to right the columns present results at time $t =
+    0$, 5, 10 and 20~\textsf{ps}.  The 1st row: the brute force
+    non-equilibrium simulation. The 2nd row: the results of
+    traditional equilibrium linear response theory. In this case we
+    effectivly consider an equilibrium system free of driving ($k_e =
+    0$), and the perturbation $\eps$ is set to 2.  The 3rd row: the
+    second order equilibrium response result.  The 4th row: the
+    non-equilibrium linear response result.  }
   \label{fig:tmp6}
@@
 
-In the previous subsection, we considered an example that the stability of the
-core sets is switched by an external driving to the system.
-In this section, we consider ``splitting single-well potential'' example
-that two new core sets grows out of 
-an old core sets by the external driving force.
-In this example, the unperturbed
-potential is simply the harmonic potential given by:
-\begin{align}
-  U(\vect q) = \frac12\,k\,\vect q^2 
-\end{align}
-Here $k = 8$~$\textsf{kJ} / (\textsf{mol nm}^2)$.
-See the red line in Fig.~\ref{fig:tmp4} for the unperturbed potential.
-%$\textsf{kJ} / (\textsf{mol nm}^4)}$
-The perturbation has again given by
-\begin{align}
-  \vect D(\vect q) = -\nabla_{\vect q} V(\vect q) 
-\end{align}
-where $V(\vect q)$ has a Gaussian profile:
-\begin{align}
-  V(\vect q) = \frac{1}{\sqrt{2\pi \sigma^2}}
-  \exp\Big\{-\frac{\vect q^2}{2\sigma^2}\Big\}
-\end{align}
-we use $\sigma = 0.16$~\textsf{nm}.
-
-In this example, we also consider two different strengths of
-perturbations. Please see the green and blue lines in Fig~\ref{fig:tmp4} for
-perturbations of $\eps = 1$ and $\eps = 2$, respectively.
-Numerical results show that the perturbation $\eps=1$ can be considered to be
-relatively small. Since the phenomenon is basically the same as
-Fig.~\ref{fig:tmp2}, we do not present them here.
-Fig.~\ref{fig:tmp6} presents the numerical results of perturbation $\eps =
-2$.  Only with a linear response calculated from the equilibrium
-simulation (2nd row), the stability of the two new core sets are too
-strong, and the unstable region between them is over estimated, comparing with
-the brute force simulation presented on the first row.
-We present the  second order response results on the 3rd row, which is
-also calculated from the equilibrium simulation. In this case, the statistical
-error is too large, so that
-the profile of $\mt_\tau f$ is indeed too noisy to draw any useful information.
-On the 4th row of Fig.~\ref{fig:tmp6} presents the linear response starting
-from a non-equilibrium simulation with external driving $k_e = 1$.
-In this case the effective strength of the perturbation to the reference
-simulation is $\eps = 1$, which is considered to be relatively small.
-The numerical results is satisfactorily consistent with the brute force
-non-equilibrium simulation.
-
-\subsection{Optimal forcing}
+Fig.~\ref{fig:tmp6} presents the numerical results of the splitting
+single-well potential. From left to right the four rows presents the
+stability of the system at time $t = 0$, 5, 10 and 20~\textsf{ps}. The first
+row presents the brute force non-equilibrium simulation.  It is clear
+that at the beginning, there is only one stable region in the phase,
+i.e. arount $q = 0$. As the time evolves, an energy barrier develops
+in the center of the simulation region, therefore, the single stable
+region splits into two equally sized stable regions.  In the end, the two
+stable regions are entirly seperated.  This brute non-equilibrium
+simulation serves as the precise result that the response theory
+should compare with. The second row shows the result of the
+traditional equilibrium linear response theory.  Please notice that in
+this case, since the reference simulation is in equilibrium (i.e. $k_e = 0$),
+the effective perturbation is of strength $\eps = 2$.  At $t =
+5$~\textsf{ps}, the accuracy of the equilibrium linear response is
+perfect. When $t = 10$~\textsf{ps}, the left stable region is a little
+over estimate. This may be due to the statistical uncertainty. At $t =
+20$~\textsf{ps}, stability of the two stable regions are too strong,
+and the unstable region in between is less stable than the brute
+force result. The accuracy of the equilibrium linear response is poor
+at $t \geq 10$~\textsf{ps}, because the strength of perturbation is no
+longer small so that the preliminary assumption of the reponse theory
+is not well fulfilled. On the third row, we present second order
+equilibrium response theory. The using of a higher order response term
+should improve the numerical accuracy, however, this is not the case,
+because the second order response term cannot be accuratly calculated
+due to the huge statistical error. By doing longer reference
+equilibrium simulations may improve the accuracy of statistics, but it
+is not ecnomic to do that.  On the fourth row of Fig.~\ref{fig:tmp6},
+we present the linear response starting from a non-equilibrium driving
+of $k_e = 1$.  The numerical results is satisfactorily consistent with
+the brute force non-equilibrium simulation, because the perturbation
+($\eps = 1$) is so small that the linear perturbation theory achieves
+good accuracy.
+
+% Fig.~\ref{fig:tmp6} presents the numerical results of perturbation $\eps =
+% 2$.  Only with a linear response calculated from the equilibrium
+% simulation (2nd row), the stability of the two new core sets are too
+% strong, and the unstable region between them is over estimated, comparing with
+% the brute force simulation presented on the first row.
+% We present the  second order response results on the 3rd row, which is
+% also calculated from the equilibrium simulation. In this case, the statistical
+% error is too large, so that
+% the profile of $\mt_\tau f$ is indeed too noisy to draw any useful information.
+
+% \subsection{The tilting double-well potential}
+
+% We test the idea firstly by a one-dimensional model system: one particle in a
+% tilting double-well potential. For convenience, we let the mass of the
+% particle to be 1~\textsf{amu},
+% and the friction coefficient to be $1\,\textsf{ps}^{-1}$.
+% The unperturbed
+% Hamiltonian of the system is given by:
+% \begin{align}
+%   \mh (\vect p, \vect q) = \frac 12 \vect p^2 + U(\vect q) 
+% \end{align}
+% with potential
+% \begin{align}
+%   U(\vect q) = \frac12 k (\vect q^2 - a^2)^2
+% \end{align}
+% Here $k = 8$~$\textsf{kJ} / (\textsf{mol nm}^4)$, and $ a = 1\ \textsf{nm}$.
+% At room temperature $300\ \textsf{K}$, $k_BT = 2.48$~\textsf{kJ/mol}.
+% The red line in Fig.~\ref{fig:tmp1} presents the unperturbed potential.
+% %$\textsf{kJ} / (\textsf{mol nm}^4)}$
+% The perturbation is given by
+% \begin{align}
+%   \vect D(\vect q) = -\nabla_{\vect q} V(\vect q) = 1
+% \end{align}
+% Here $V(\vect q) = -\vect q$   effectively tilts the original
+% potential $U(\vect q)$ (see the gree and blue lines in Fig.~\ref{fig:tmp1}).
+% The strength of $F_e(t)$ is
+% set to be $k_e$, when the time $t$ is larger than some warm-up time $t_c$.
+% When $t \leq t_c$, the strength of $F_e(t)$ is linearly growing,
+% i.e. $ F_e(t) = k_e\cdot t/t_c$.
+% We use the value $t_c = 20$~\textsf{ps}, so that the perturbation increases slow
+% enough: the typical decaying time scale of the correlation function
+% in~\eqref{eqn:eq-lr} is roughly 3~\textsf{ps}.
+% We choose $\tau = 1$~\textsf{ps}, and
+% % We choose $\mt_\alpha$
+% % to be the propagator of the Langevin dynamics at temperature $150$
+% % \textsf{K} with $\alpha = 1$ \textsf{ps}.
+% consider the stability of the
+% distribution $f(\vect x, t)$ at time $t$: $\Delta\mt_\tau f(\vect x, t)$.
+% The larger
+% this value, the more stable the core sets are.
+
+% See Fig.~\ref{fig:tmp2} and \ref{fig:tmp3} for numerical result.  Both
+% the results of the brute-force non-equilibrium simulation
+% and that of the response theory are presented.
+% In Fig.~\ref{fig:tmp2},
+% the good agreement between the brute-force (1st row)
+% and the linear response result (2nd row) indicates that $\eps = 2$ is
+% indeed a relatively small perturbation to the equilibrium.
+% In the third row of Fig.~\ref{fig:tmp2},
+% the second order response is considered. We do not see any
+% substantial improvement of the result by including this higher order term,
+% because the
+% driving force to the system is small,
+% and the linear term is anyway a good approximation.
+
+% In Fig.~\ref{fig:tmp3}, as the
+% perturbation grows stronger, the left well becomes less stable, while
+% the right well becomes more stable.  When the perturbation is as strong as
+% $\eps = 4$, the linear response result (2nd row) is no
+% longer precise:
+% At $t\geq 20$ \textsf{ps}, the left well basically
+% disappears, however, the linear response calculation presents
+% artificial structure (that is a small artificial
+% stable region accompanied by a small unstable region)
+% at the position of the left well.
+% The second order response (3rd row) improves the accuracy a little, 
+% however, it also includes artificial finer-grained oscillation in the profile.
+% At the same time, the numerical uncertainty is higher, so the figure
+% looks noisy.
+% In the 4th row of Fig.~\ref{fig:tmp3}, we present the linear response result
+% that calculate from 
+% a intermediate
+% non-equilibrium simulation with driving force of $k_e = 2$.
+% Notice that the reference simulation in this case is also a
+% non-equilibrium simulation, so the effective strength of perturbation
+% is reduced to $\eps = 2$. According to Fig.~\ref{fig:tmp1}, this can be
+% considered as a small perturbation.
+% A substantial improvement is achieved: the profile
+% is correctly calculated, and the statistical error is small.
+
+% % the 1st order result of the
+% % response formula~\eqref{eqn:pert-approx-1} and
+% % \eqref{eqn:pert-approx-2}, with
+% % respect to the reference state $F_e^{\textrm{max}} = 2$. The result
+% % is impressively improved compared with the linear response with respect to the
+% % equilibrium state.
+
+\subsection{Optimal tilting of a double-well potential}
 \begin{figure}
@@
     piecewise linear ansatz functions. The time interval of a linear
-    piece is 1 ps.  In the lower plot, the dashed lines display the
-    value of $P(\vect X_t)$ as the time evolves.  The solid lines
-    present $O[F_e](t)$.  }\label{fig:tmp7}
+    piece is 1 ps.  The blue insertions show the shape of the
+    non-equilibrium driving potential $U(\vect q) + F_e(t)V(\vect
+    q)$ at time indicated by the black arrow.  In the lower plot, the
+    dashed lines display the value of $P(\vect X_t)$ as the time
+    evolves.  The solid lines present $O[F_e](t)$.  }\label{fig:tmp7}
 \end{figure}
 
-We investigate the example of the double well potential, and
-consider the problem of optimally forcing the population from the
-left well to the right well. That is equal to consider the 
+In this section, we consider the following double well potential:
+\begin{align}
+  U(\vect q) = \frac12 k (\vect q^2 - a^2)^2
+\end{align}
+Here $k = 8$~$\textsf{kJ} / (\textsf{mol nm}^4)$, and $ a = 1\
+\textsf{nm}$.  See the leftmost blue insertion of Fig.~\ref{fig:tmp7} for
+the shape of the potential.
+%$\textsf{kJ} / (\textsf{mol nm}^4)}$
+The perturbation is given by a conservative tilting to $U(\vect q)$:
+\begin{align}
+  \vect D(\vect q) = -\nabla_{\vect q} V(\vect q) = 1
+\end{align}
+Here $V(\vect q) = -\vect q$. We want to
+optimally design the tilting such that the probability
+of being in the right well is as high as possible 
+at the end time of the process. That is equal to consider the 
 the following optimal forcing problem:
@@
 \begin{align}
-  O[F_e](T) &= P(\vect X_T) - \eta \int_0^T\vert F_e(t)\vect D(\vect q)\vert^2\, dt
-\end{align}
-where $P(\vect X_T)$ is equal to $1$ when the end point of the
-trajectory $\vect X_T$ is in the right well, and equal to 0 otherwise.
-It is clear that for an unbiased double-well, the average of $P(\vect
-X_T)$ is 0.5.  The integral is the ``cost'' of the control and $\eta$
-is a positive constant that indicates the relative magnitude of the
+  O[F_e](T) &= P(\vect X_T) -
+  \eta\,
+  \big\langle \int_0^T\vert F_e(t)\vect D(\vect q)\vert^2\, dt\,
+  \big\rangle
+\end{align}
+where $P(\vect X_T)$ is probability of the end point of the trajectory
+$\vect X_T$ being in the right well.  It is
+clear that for an unbiased double-well, $P(\vect X_T)$
+is 0.5.  The integral is the ``cost'' of the control and $\eta$ is a
+positive constant that indicates the relative magnitude of the
 cost. We consider the ansatz function family $\mathcal F$ of piecewise

### 2013-06-26 one.test/template.optimal.ctrl/tools/simulator/app/tiltDoubleWell.ctr.finitDiff.cpp

@@ -1,0 +1,413 @@
+#include <algorithm>
+#include <fstream>
+#include <iomanip>
+#include <iostream>
+#include <iterator>
+#include <list>
+#include <map>
+#include <set>
+#include <sstream>
+#include <string>
+#include <queue>
+#include <vector>
+#include <cmath>
+
+#include "Force.h"
+#include "Perturbation.h"
+#include "Integrator.h"
+#include "Distribution.h"
+#include "NoneqInfo.h"
+#include "Stopwatch.h"
+
+#include <boost/program_options.hpp>
+#include "RandomGenerator.h"
+#include <mpi.h>
+
+namespace po = boost::program_options;
+using namespace MPI;
+
+void readFile (const string & filename,
+	       std::vector<double > & tt,
+	       std::vector<double > & value)
+{
+  tt.clear ();
+  value.clear ();
+  FILE * fp = fopen (filename.c_str(), "r");
+  if (fp == NULL){
+    cerr << "cannot open file " << filename << endl;
+    exit (1);
+  }
+
+  double tmp1, tmp2;
+  while (fscanf (fp, "%lf%lf", &tmp1, &tmp2) == 2){
+    tt.push_back (tmp1);
+    value.push_back (tmp2);
+  }
+  
+  fclose (fp);
+}
+
+
+void cal_order1punish (const double & beta,
+		       const vector<double > & tt,		// size num mode
+		       const vector<double > & ttvalue,
+		       vector<double > & order1punish)
+{
+  const std::vector<double > & vv (ttvalue);
+  order1punish.resize (tt.size());
+  order1punish[0] = 0.5 * beta * (2./3. * vv[0] + 1./3. * vv[1]) * (tt[1] - tt[0]);
+  int nn = tt.size() - 1;
+  order1punish[nn] = 0.5 * beta * (1./3. * vv[nn-1] + 2./3. * vv[nn]) * (tt[nn] - tt[nn-1]);
+  
+  for (unsigned ii = 1; ii < tt.size() - 1; ++ii){
+    order1punish[ii] = 0.5 * beta * ((1./3. * vv[ii-1] + 2./3. * vv[ii]) * (tt[ii] - tt[ii-1]) +
+				     (1./3. * vv[ii+1] + 2./3. * vv[ii]) * (tt[ii+1] - tt[ii]) );
+  }
+}
+
+double runTraj (const vector<double > & tt,		// size num mode
+		const vector<double > & ttvalue,
[…]

### 2013-06-26 one.test/template.optimal.ctrl/tools/simulator/include/NoneqInfo.h

@@ -1,0 +1,60 @@
+#ifndef __NoneqInfo_h_NONEQMSM_wanghan__
+#define __NoneqInfo_h_NONEQMSM_wanghan__
+
+#include <vector>
+#include "Defines.h"
+#include "Distribution.h"
+#include "Perturbation.h"
+
+using namespace std;
+
+class NoneqInfo 
+{
+  double x0, x1;
+
+  double beta;
+  
+  double dt;
+  double noneqTime;
+  double noneqCheckFeq;
+  int numCheck;
+  int checkNumFeq;
+
+  int countNoneq;
+  int countNoneqSeg;
+
+  const Perturbation * ppert;
+  int numMode;
+  
+  vector<double >				order0;		// time
+  vector<double >				order0punish;	// time
+  
+  double					punish;
+  int						ntraj;
+public:
+  NoneqInfo ();
+  // ~NoneqInfo ();
+public:
+  void reinit (const double & beta,
+	       const double & x0,
+	       const double & x1,
+	       const double & dt,
+	       const double & noneqTime,
+	       const double & noneqCheckFeq,
+	       const Perturbation & pert);
+  int inSet (const Dofs & x) ;
+  void newTraj ();
+  void depositMainTraj (const Dofs & oldx,
+			const Dofs & newx,
+			const double & sigma,
+			const Dofs & dw);
+  void average ();
+  void collectLast () ;
+public:
+  const vector<double > &	   get_order0 () const {return order0;}
+  const vector<double > &	   get_order0punish () const {return order0punish;}
+}
+    ;
+
+
+#endif

### 2013-06-26 one.test/template.optimal.ctrl/tools/simulator/src/NoneqInfo.cpp

@@ -1,0 +1,122 @@
+#include "NoneqInfo.h"
+
+NoneqInfo::
+NoneqInfo ()
+    : ppert (NULL)
+{
+}
+
+
+void NoneqInfo::
+reinit (const double & beta_,
+	const double & x0_,
+	const double & x1_,
+	const double & dt_,
+	const double & noneqTime_,
+	const double & noneqCheckFeq_,
+	const Perturbation & pert)
+{
+  beta = beta_;
+  x0 = x0_;
+  x1 = x1_;
+  dt = dt_;
+  noneqTime = noneqTime_;
+  noneqCheckFeq = noneqCheckFeq_;
+  ppert = &pert;
+
+  numMode = ppert->numMode();  
+  numCheck = int((noneqTime + 0.5 * noneqCheckFeq) / noneqCheckFeq) + 1;
+  checkNumFeq = int((noneqCheckFeq + 0.5 * dt) / dt);
+  
+  order0.resize(numCheck);
+  order0punish.resize(numCheck);
+
+  for (int tt = 0; tt < numCheck; ++tt){
+    order0[tt] = 0.;
+    order0punish[tt] = 0.;
+  }
+
+  ntraj = 0;
+  newTraj ();
+  ntraj = 0;
+}
+
+
+void NoneqInfo::
+newTraj ()
+{
+  ntraj ++;
+
+  punish = 0;
+  countNoneq = 0;
+  countNoneqSeg = 0;
+}
+
+int NoneqInfo::
+inSet (const Dofs & x) 
+{
+  if (x.xx[0] > (0.5 * (x0 + x1))){
+    return 1;
+  }
+  else {
+    return 0;
+  }
+}
+
+
+void NoneqInfo::
+depositMainTraj (const Dofs & oldx,
+		 const Dofs & newx,
+		 const double & sigma,
+		 const Dofs & dw)
+{
+  if (countNoneq == 0){
+    order0[0] += (-inSet(oldx));
+    // countNoneq ++;
+    // return;
+  }
+    
+  double nowTime = countNoneq * dt;
+  Dofs pvalueTime;
+  ppert->operator () (oldx, nowTime, pvalueTime);
+  
+  double tmp2 = 0.;
+  for (unsigned dd = 0; dd < NUMDOFS; ++dd){
+    tmp2 += pvalueTime.vv[dd] * pvalueTime.vv[dd];
+  }
+  punish += beta * 0.5 * tmp2 * dt;
+
+  countNoneq ++;
+
[…]

### 2013-06-28 writedown/linear.response/paper.tex
update changes

@@
 \usepackage[fleqn]{amsmath}
-\usepackage{amssymb}
+\usepackage{amssymb,amsthm}
 \usepackage[dvips]{graphicx}
@@
 
+\newtheorem{thm}{Theorem}
+
 \newcommand{\recheck}[1]{{\color{red} #1}}
@@
 
+\newcommand{\R}{{\mathbf R}}
+\newcommand{\Q}{{\mathbf Q}}
+\newcommand{\Z}{{\mathbf Z}}
+\newcommand{\N}{{\mathbf N}}
+\newcommand{\T}{{\mathbf T}}
 \newcommand{\mh}{\mathcal H}
@@
 \newcommand{\fwg}{{\mathcal A}}
+\newcommand{\cA}{\mathcal A}
 \newcommand{\bwg}{{\mathcal B}}
 \newcommand{\bsigma}{\boldsymbol\sigma}
+\newcommand{\bE}{{\mathbf E}}
+\newcommand{\bP}{{\mathbf P}}
+\newcommand{\one}{{\mathbf 1}}
+\newcommand{\zero}{{\mathbf 0}}
+
+\newcommand{\wrt}{with respect to }
 
@@
 % \title{Linear response theory for core set identification}
-\title{The response theory for the system under a non-equilibrium driving}
+\title{Linear response theory for a system under non-equilibrium conditions}
 \author{Han Wang}
@@
 \begin{abstract}
-  In this talk, we firstly review the traditional linear response
-  theory, and then generalize it to the system under a non-equilibrium
+  In this paper, we firstly review standard equilibrium linear response
+  theory in the context of molecular dynamics, and then generalize it to systems in nonequilibrium and under a nonequilibrium
   driving. We calculate not only the linear response term, but also
-  the higher order terms with respect to the perturbation to the
-  reference dynamics. The response theory is applied to investigate
-  the dynamics of core set that plays a central role in the Markov
-  State Model (MSM). The advantages and limitations are discussed from
-  the numerical tests.
-  % In this draft, we are trying to study the core set based Markov
-  % State Model (MSM) under non-equilirbium conditions.
+  second order terms for small perturbations to the reference dynamics.
+Furthermore, we outline how the novel nonequilibirum linear response formula can be used to compute optimal controls of molecular systems for cases in which one wants to steer the system to maximize a certain target expectation value.
+
 \end{abstract}
@@
 
-\section{Theoretical background}
-\subsection{Classical equilibrium linear response theory}
+\section{Introduction}
 
-In this section, we want to firstly give a brief description of the classical linear response
-theory, for details see Ref.~\cite{tuckeman2010statistical} for
-example.
-We study the non-equilibrium system driven by some external force.
-More strictly, we investigate the system that is governed by the
-following Langevin equation:
-\begin{align}\label{eqn:langevin-1}
-  \dot{\vect q} & = \nabla_{\vect p}\mh(\vect q,\vect p), \\\label{eqn:langevin-2}
-  \dot{\vect p} & =- \nabla_{\vect q}\mh(\vect q,\vect p)
-  + \eps F_e(t) \vect D(\vect q) 
-  - \gamma\vect p
-  + \sigma\dot{\vect W}.
-\end{align}
-Here $\vect D$ is the external driving force applied to the system,
-the magnitude of which is always assumed to be controllable.
-$\eps F_e(t)$ is the strength of the driving as a function of time.
-The prefactor $\eps$ indicates that we regard the driving force to be considerably smaller than the internal forces. We will comment on what is meant by "considerably smaller" further below.
+Standard molecular dynamics simulations are dealing with systems in thermal equilibrium; in this case they are tuned to the canonial or Boltzmann distribution in the sense that either (1) if one starts from this distribution it remains invariant under the dynamics or (2) if one generates a very long trajectory it samples state space with respect this distribution, that is, every possible state of the molecular system under consideration is visited according to the probability given by it.
+Obviously, this allows to compute equilibrium expectation values with respect to the  canonical distribution simply be computing long trajectories.
 
-% , and
-% the magnitude of $F_e(t)$ is small and of order $\mo (\eps)$, $\eps \ll 1$.
+Often, however, one is interested in knowing about the response of the molecular system to perturbation out of equilibrium.
+The standard linear response formula allows to answer this question, at least partially.
+In the standard setting it gives us the first order of the change to an equilibrium expectation value as resulting from the non-equilibrium perturbation.
+Here, first order means first order in the size of the perturbation.
+This linear response formula has a long history of extensions and generalizations.
+In some sense it has become one of the cornerstones of modern statistical physics since it can be related to the fluctuation dissipation theorem (FDT) which roughly states that for appropriate systems in statistical equilibrium, the average response to small external perturbations can be calculated through the knowledge of suitable correlation functions of the unperturbed statistical system.
+In this sense there the standard linear response theory holds in a  very general sense \cite{Majda} as long as one assumes that the unperturbed system is in equilibrium.
+%=Martin Hairer and Andrew J Majda, A simple framework to justify linear response theory, Nonlinearity 23 (2010) 909–922 doi:10.1088/0951-7715/23/4/008
 
-Then the phase space distribution $f(\vect x, t)$ ($\vect x = \{\vect
-q, \vect p\}$) is subject to the Kolmogorov forward equation:
-\begin{align}\label{eqn:orig-forward}
-  \frac{\partial}{\partial t} f(\vect x, t) - \fwg^\beta(t) f(\vect x, t) = 0
-\end{align}
-where $\fwg^\beta(t)$ is the forward infiniesimal generator given by
-\begin{align}
-  \fwg^\beta(t) =
-  \frac{\sigma^2}2\Delta_{\vect p}
-  - \nabla_{\vect p}\mh 
-  \cdot\nabla_{\vect q}
-  - \Big(
-  -\nabla_{\vect q}\mh +
-  \eps F_e(t)  \vect D  - \gamma\vect p
-  \Big)\cdot\nabla_{\vect p}
-  + 3N\gamma,
-\end{align}
-and $\beta2=\gamma / \sigma^2$ is the inverse temperature.
-The equilibrium system
-is governed by
-the standard Langevin equation, i.e. with
-vanishing driving force $\eps = 0$ in
-Eqn.~\eqref{eqn:langevin-1} and \eqref{eqn:langevin-2}.
+There is an increasing number of articles in the literature that report on applications of molecular dynamics to nonequilibrium settings.
+There are many generalization to so-called nonequilibrium steady states based on the generality of the FDT, but despite its wide use, the present authors do not know of a linear response formula that applies to fully general nonequilibirium cases.
+We will provide such a formula for the case that the underlying dynamics can be described by Langevin dynamics.
+Furthermore, we will even provide a formula for the second order response of a Langevin system in non-equilibrium to a small perturbation.
 
-Under our smallness assumption on the driving force the full phase space distribution has the following asymptotic expansion in $\eps$:
-\begin{align}\label{eqn:f-expan}
-  f(\vect x, t) &= f_0(\vect x, t) + \eps f_1(\vect x, t)
-  +\eps^2 f_2(\vect x, t) + \mo (\eps^3),
-\end{align}
-where $f_0=f_0(x,t)$ denotes the evolution of the phase space distribution for the undriven system (with $\eps=0$).
-In addition we have:
-\begin{align}\label{eqn:A-expan}
-  \fwg^\beta(t) = \fwg^\beta_0 + \eps\fwg_1(t),
-\end{align}
-where
-\begin{align}
-  \fwg_0^\beta =&
-  \frac{\sigma^2}2\Delta_{\vect p}
-  -
-  \nabla_{\vect p}\mh\cdot\nabla_{\vect q}
-  - \big(
-  -\nabla_{\vect q}\mh - \gamma\vect p
-  \big)\cdot\nabla_{\vect p}
-  + 3N\gamma,\\
-  \fwg_1(t) =&
-  - F_e(t)  \vect D \cdot\nabla_{\vect p}.
-\end{align}
+Instead of applying this theory to a molecular system we will go one step further.
+We will outline how the novel nonequilibrium linear response formula can be used to compute the optimal control of molecular systems.
+In optimal control one seeks the optimal way to perturbed a molecular system such that a certain target expectation value (e.g.
+population of certain states) is maximized under constraints on the energy of the control.
+In general the control drives the molecular system under control out of equilibrium.
+Thus, the nonequilibrium linear response formula can be used to find the optimal correction of the present control regarding the expectation value of interest.
 
-The invariant measure of $\fwg^\beta_0$ satisfies $\fwg^\beta_0f^\beta_{\inv}(\vect x) = 0$, and is given by the Boltzmann distribution
-\begin{align}
-  f^\beta_{\inv}(\vect x) \propto e^{-\beta\mh(\vect x)}
-\end{align}
-Inserting \eqref{eqn:f-expan} and \eqref{eqn:A-expan} into
-\eqref{eqn:orig-forward}, 
-the Kolmogorov forward equation becomes:
-\begin{align}
-  \frac{\partial}{\partial t}
-  [f_0(\vect x, t) + \eps  f_1(\vect x, t) + \eps^2f_2(\vect x, t)]
-  -
-  [\,\fwg^\beta_0 + \eps\fwg_1(t)\,]
-  [f_0(\vect x, t) + \eps  f_1(\vect x, t) + \eps^2f_2(\vect x, t)]
-  = 0
-\end{align}
-Matching this equation at different orders of $\eps$, the zeroth
-order gives:
-\begin{align}\label{eqn:e-o0}
-  \bigg[
-  \frac{\partial}{\partial t}
-  - \fwg_0^\beta
-  \bigg]
-  f_0(\vect x, t)
-  = 0
-\end{align}
-which is the Kolmogorov forward equation of the undriven system. Matching the first and the second order gives:
-\begin{align}\label{eqn:e-o1}
-  \bigg[
-  \frac{\partial}{\partial t}
-  - \fwg_0^\beta
-  \bigg]
-  f_1(\vect x, t)
-  =&
-  \fwg_1(t) f_0(\vect x, t)\\\label{eqn:e-o2}
-  \bigg[
-  \frac{\partial}{\partial t}
-  - \fwg_0^\beta
-  \bigg]
-  f_2(\vect x, t)
-  =&
-  \fwg_1(t) f_1(\vect x, t)
-\end{align}
-Formally solving \eqref{eqn:e-o0} -- \eqref{eqn:e-o2} with
-initial conditions $f_1(\vect x, 0) = 0$ and $f_2(\vect x, 0) = 0$
-gives
-\begin{align}\label{eqn:f0-0}
-  f_0(\vect x, t)
-  =&\,
-  e^{t\fwg^\beta_0}f_0(\vect x, 0) \\\label{eqn:f1-0}
-  f_1(\vect x, t)
-  =&\,
-  \int_0^t\dd s\,
-  e^{(t-s)\fwg^\beta_0}\circ
-  \fwg_1(s) f_0(\vect x,s)
-  =
-  \int_0^t\dd s\,
-  e^{(t-s)\fwg^\beta_0}\circ
-  \fwg_1(s)\circ
-  e^{s\fwg^\beta_0}f_0(\vect x, 0) \\\nonumber
-  f_2(\vect x, t)
-  =&\,
-  \int_0^t\dd s\,
-  e^{(t-s)\fwg^\beta_0}\circ
-  \fwg_1(s) f_1(\vect x,s) \\\label{eqn:f2-0}
-  &=
-  \int_0^t\dd s
-  \int_0^s\dd u\,\,
-  e^{(t-s)\fwg^\beta_0}\circ
-  \fwg_1(s)\circ
-  e^{(s-u)\fwg^\beta_0}\circ
-  \fwg_1(u)\circ
-  e^{u\fwg^\beta_0}
-  f_0(\vect x, 0) 
-\end{align}
-If the initial distribution is the invariant distribution, i.e.
-$f_0(\vect x, 0) = f_{\inv}^\beta(\vect x)$,
-\eqref{eqn:f0-0} becomes:
-\begin{align}
-  f_0(\vect x, t) = f_{\inv}^\beta(\vect x)
-\end{align}
-It can be shown that
-\begin{align}
-  \fwg_1(s) f_{\inv}^\beta(\vect x)
-  =
-  -\beta {F_e(s)}
-  j(\vect x)
-  f_{\inv}^\beta(\vect x),
-\end{align}
-where the \emph{dissipative flux} $j(\vect x)$ is defined by,
-\begin{align}
-  j(\vect x) =
-  -\vect D\cdot\nabla_{\vect p}\mh
-\end{align}
-Then $f_1(\vect x, t)$ becomes:
-\begin{align}
-  f_1(\vect x, t)
-  =
-  -\beta
-  \int_0^t\dd s\,
-  F_e(s)\,
-  e^{(t-s)\fwg^\beta_0}
-  \big[
-  j(\vect x)
-  f_{\inv}^\beta(\vect x)
-  \big]
-\end{align}
-Therefore, the linear order approximation to
-the solution of~\eqref{eqn:orig-forward} is 
-\begin{align}\label{eqn:solv-orig-linear}
-  f(\vect x, t) =
-  f_{\inv}^\beta(\vect x)
-  - \eps\beta
-  \int_0^t\dd s\,
-  F_e(s)\,
-  e^{(t-s)\fwg^\beta_0}
-  \big[
-  j(\vect x)
-  f_{\inv}^\beta(\vect x)
-  \big]
-  + \mo(\eps^2)
-\end{align}
-Now let $\mathcal{O}$ be an observable of the system. Its expectation value satisfies the linear response formula:
-\begin{align}\nonumber
-  \mathcal O(t)
-  &=
-  \int\dd \vect x \:O(\vect x)f(\vect x, t)  \\\nonumber
-  &=
-  \int\dd \vect x\, O(\vect x)\,
-  [\,f_{\inv}^\beta(\vect x) + \eps f_1(\vect x, t)] +\mathcal{O}(\eps^2)\\\nonumber
-   \label{eqn:eqi-pert-1}
-  &=
-  \langle O\rangle
-  -
- \eps \beta
-  \int_0^t\dd s\,
-  F_e(t-s)
-  \int \dd \vect x\,
-  f^\beta_{\inv}(\vect x) j(\vect x)\,
-  e^{s\bwg_0^\beta}
-  O(\vect x)+\mathcal{O}(\eps^2)
-\end{align}
-where $ \langle O\rangle=\int \dd\vect x\,O(\vect x) f_{\inv}^\beta(\vect x)$ is the expectation value of the undriven equilibrium system, and $\bwg^\beta_0$ is the infiniesimal generator of the undriven
-Kolmogorov backward equation, which is the adjoint operator of $\fwg^\beta_0$. That is,
-we have
-\begin{align}
-  e^{s\bwg^\beta_0} O(\vect x) = \mathbb E_{\vect x} [O(\vect X^\beta_s)]
-\end{align}
-where $\vect X^\beta_t$ is the trajectory of the undriven Langevin equation (with $\eps=0$)
-starting at $\vect x$.
-Therefore, we have up to order 2 in $\eps$:
-% Notice that $e^{i\ml_0(t-s)}O(\vect x) = O(\phi_{t-s}(\vect x))$,
-% where $\phi_t(\vect x)$ is the flow mapping of the \emph{unperturbed}
-% Hamiltonian system. We have
-\begin{align}\nonumber
-  \mathcal O(t)
- & =
-  \langle O\rangle
-  -
-  \eps\beta
-  \int_0^t\dd s\,
-  F_e(t-s)
-  \int \dd \vect x\,
-  f^\beta_{\inv}(\vect x)
-  j(\vect x)\,
-  \mathbb E_{\vect x} [O(\vect X^\beta_s)]\\
-\label{eqn:eq-lr}  
- & =
-  \langle O\rangle
-  -
-  \eps \beta
-  \int_0^t\dd s\,
-  F_e(t-s)
-  \big\langle
-  j\cdot
-  \mathbb E [O(\vect X^\beta_s)]
-  \big\rangle
-\end{align}
-where $  \langle
-  j\cdot
-  \mathbb E [O(\vect X^\beta_s)]
-  \rangle$ is the \emph{equilibrium time
-  correlation} between the observation $O$ at time $s$ and the flux $j$.
+The outline of the article is as follows: First, we will review the derivation of the standard linear response formula for  rather general diffusion processes which include which Langevin dynamics.
+Next, we will show how to derive first and second order response formulas for the nonequilibrium case and how to apply these formula for the computation of optimal controls.
+Finally, we will validate the nonequilibrium linear response formula and its use for optimal control for simple test cases.
+This numerical experiments will also outline that the use of the linear response formula is imperative for numerical efficiency and allows to extend the applicability of linear response theory to stronger perturbations.
+
+\section{Linear response}
+
+In this section, we want to firstly give a simple and formal derivation of the linear response for equilibrium and nonequilibrium systems.
+To this end we consider a general It\^o stochastic differential equation of the form
+\begin{equation}\label{sde}
+dX_{t} = (b(X_{t},t) + \eps v_{t})dt +a(X_{t})dB_{t}\,,\quad t\ge 0\,,
+\end{equation}
+where $X_{t}\in\R^{d}$, $b(\cdot,\cdot)$ is a smooth time-dependent vector field, $a(\cdot)$ a smooth field of $d\times n$ matrices and $B_{t}$ is standard Brownian motion in $\R^{n}$.
+Here $v_{t}\in\R^{d}$ is any given driving force applied to the system, which in general may depend on $X_{t}$ and $t$; $\eps>0$ is a small parameter.
+We consider the equilibrium and nonequilibrium case separately.
+
+\subsection{Small perturbations from equilibrium: Langevin dynamics}
+
+For the equilibrium case (see, e.g., \cite{XXX,YYY,ZZZ}), we assume that the infinitesimal generator
+\[
+\fwg^{\eps} = \fwg_{0} + \eps\fwg_{1}\,, 
+\]
+with 
+\[
+\fwg_{0} = \frac{1}{2}aa^{T}\colon\nabla^{2} + b\cdot\nabla\,,\quad \fwg_{1}=v\cdot\nabla 
+\]
+of (\ref{sde}) has an isolated eigenvalue 0 corresponding to the unique invariant measure of the dynamics for $\eps=0$ and satisfies a spectral gap condition (see, e.g., \cite{stoltz2012}), so that the operator $(\fwg_{0}^{*})^{-1}\fwg_{1}^{*}$ is bounded (on an appropriate domain), where $\cA_{0}^{*}$ and $\cA_{1}^{*}$ denote the formal adjoints in $L^{2}$, e.g., $\cA_{1}^{*}\phi=-\nabla(v \phi)$.
+
+Specifically, we are interested in the case that (\ref{sde}) has the form of a Langevin equation, in which case  
+\begin{equation}\label{eqn:langevin-1}
+b = \left(\nabla_{p}H, -\nabla_{q} H - \gamma\nabla_{p}H\right )^{T}\,,\quad a= (\zero,\sigma)^{T}\,,
+\end{equation}
+where $H\colon\R^{d}\to\R$, $d=2n$ is the Hamiltonian of the system, $\gamma=\gamma^{T}>0$ the positive semi-definite $n\times n$ friction matrix, $\sigma$ constant, satisfying $2\gamma=\beta\sigma\sigma^{T}$ for some $\beta>0$, and we have used the convention $x=(q,p)\in\R^{n}\times\R^{n}$.
+Under these conditions on friction and noise and some mild growth conditions on the Hamiltonian for large arguments, the unperturbed dynamics has a unique invariant measure with density
+\[
+\rho_{0}(q,p) = \frac{1}{Z}e^{-\beta H(q,p)}\,,\quad  Z=\int  e^{-\beta H(q,p)}\,dqdp\,.
+\] 
+and the above spectral gap condition is met.
+Now let $f$ be any integrable phase space function and let $\rho^{\eps}=\rho^{\eps}(q,p,t)$ denote the probability density of $X_t=(Q_{t},P_{t})$, the solution to the Langevin equation (\ref{sde})--(\ref{eqn:langevin-1}), assuming that $(Q_{0},P_{0})=(q,p)$ was distributed according to the invariant density $\rho_{0}$.
+We define the expectation \wrt $\rho^{\eps}$ as
+\[
+\bE_{\rho^{\eps}}[f] = \int f \rho^{\eps} \,dqdp\,.
+\] 
+A classical result (see, e.g., \cite{HairerMaijda08} and the references therein), that is usually derived using a formal expansion of the solution to the Kolmogorov forward or backward equations in powers of $\eps$, now states that
+\begin{equation}\label{lr}
+\lim_{\eps\to 0}\frac{\bE_{\rho^{\eps}}[f] - \bE_{\rho^{0}}[f]}{\eps} = \bE_{\rho^{0}}[\cA_{1}\cA_{0}^{-1}(f-\bE_{\rho^{0}}[f])]\,.
+\end{equation}
+
+\subsubsection*{Green-Kubo relations} 
+For the specific case of the Langevin equation (\ref{sde})--(\ref{eqn:langevin-1}), the general linear response (\ref{lr}) can be recast in form of the better known Green-Kubo formula \cite{MarconiPhysRep,Risken}.
+Using the formal operator identity
+\[
+-\cA_{0}^{-1} g = \int_{0}^{\infty}e^{t\cA_{0}}g \,dt\,,
+\]
+that can shown to hold for the generator of the Langevin equation under mild conditions (formally for functions $g$ that are orthogonal to the nullspace of $\cA_{0}^{*}$, i.e.
+functions satisfying $\bE_{\rho_{0}}[g]=0$), we find that
+\begin{equation}\label{GreenKubo}
+\lim_{\eps\to 0}\frac{\bE_{\rho^{\eps}}[f] - \bE_{\rho^{0}}[f]}{\eps} = -\beta\int_{0}^{\infty}\bE_{\rho^{0}}[J(Q_{0},P_{0}) (f(Q_{t},P_{t})-\bE_{\rho^{0}}[f])]\,,dt
+\end{equation}
+in terms of the dissipative flux $J=-v\cdot\nabla H$, where the expectation under the integral is taken over all realizations/path of the unperturbed Langevin equation with initial distribution $\rho^{0}(q,p,0)=\rho_{0}$.
+In other words:
+\[
+\bE_{\rho^{\eps}}[f] \approx \bE_{\rho_{0}}[f] - \eps\beta\int_{0}^{\infty}\bE_{\rho_{0}}[J(Q_{0},P_{0}) (f(Q_{t},P_{t})-\bE_{\rho_{0}}[f])]\,dt.
+\]
 
@@
 
-\subsection{The non-equilibrium response theory}
+\subsection{Nonequilibrium response theory: controlled Langevin dynamics}
 
-The classical response theory mainly has the following two limitations:
-Firstly, one has to start from an equilibrium simulation. That means if the
-external driving is no longer small ($\eps$ is relatively big), the
-first order (or linear) response~\eqref{eqn:eq-lr} may not be a good approximation to
-the true non-equilibrium process any more.
-One possible way to improve the accuracy is to introduce the second
-order response term, e.g., based on equation~\eqref{eqn:f2-0}.
-Here one encounters the second limitation of the classical
-response theory: it does not provide a frame work, under which
-the second and even higher order responses can be easily derived.
-The limitations of the classical responses theory motivate us to
-develop a more general framework that allows one starting from any
-non-equilibrium process (arbitrarily driven), and calculating linear responses from it.
+The classical response theory has the limitation that the reference (equilibrium) distribution $\rho_{0}$ must be stationary under the dynamics and is unique.
+Moreover the validity of the linear approximation relies on a spectral gap condition that is difficult to verify in practice.
+In particular the perturbation argument does not provide a framework, under which the second and even higher order responses can be easily derived.
 
-This time, we consider the following Langevin equation:
-\begin{align}\label{eqn:langevin-1b}
-  \dot{\vect q} & = \nabla_{\vect p}\mh(\vect q,\vect p), \\\label{eqn:langevin-2b}
-  \dot{\vect p} & =- \nabla_{\vect q}\mh(\vect q,\vect p)
-  + (F_e(t)+\eps \Delta F_e(t)) \vect D(\vect q)
-  - \gamma\vect p
-  + \sigma\dot{\vect W},
+\subsubsection*{Girsanov transformation}
+
+Here we propose an alternative (purely formal) derivation of the linear response result, that is based on a change of drift in the corresponding SDE and which allows for an easy generalization of the above linear response result to nonequilibrium systems.
+We will briefly review the idea of the change of drift via Girsanov transformations; for details we refer to the textbook \cite{oksendal}.
+Now let $x_{t},X_{t}$ be the solutions to the following stochastic differential equations:
+\begin{subequations}\label{sde2}
+\begin{align}
+dx_{t} & = b(x_{t},t)dt + a(x_{t})dB_{t} \label{sde2-1}\\
+dX_{t} & = (b(X_{t},t) + \eps v_{t})dt +a(X_{t})dB_{t} \label{sde2-2}
 \end{align}
-such that for $\eps=0$ we get a driven system that is further
-perturbed by the addition time-dependent driving force $ \Delta
-F_e(t) \vect D(\vect q)$. Let $\vect X^\eps_t$ denote the associated solution process, and
-$\mathcal P_\eps$ the probability measure of the trajectories
-generated by driven Langevin
-dynamics~\eqref{eqn:langevin-1b}--\eqref{eqn:langevin-2b}.
+\end{subequations}
+for $0\le t\le T$ and with fixed initial conditions
+\[
+x_{0} = X_{0} = x\,,
+\]
+Suppose that there exists a auxiliary stochastic process $\xi_{t}\in\R^{m}$ such that 
+\begin{equation}\label{xi}
+a(X_{t})\xi_{t} = v_{t}\,,
+\end{equation}
+The auxiliary variable $\xi$ will be called \emph{control variable}. We define  
+\[
+W_{t} = \eps\int_{0}^{t}\xi_{s}\,ds + B_{t}\,,\quad 0\le t\le T\,,
+\]
+which allows us to express (\ref{sde2-2}) by
+\begin{equation}\label{sde2-3}
+dX_{t} = b(X_{t},t)dt +a(X_{t})dW_{t} 
+\end{equation}
+It follows from the Girsanov theorem \cite[Thm.~8.6.8]{oksendal2003}, sometimes also called \emph{Cameron-Martin-Girsanov theorem} \cite{}, that $W_{t}$ is again a Brownian motion under a new probability measure that has a density \wrt the Gaussian probability measure that is generated by the Brownian motion $B_{t}$.
+Specifically, let $\nu$ denote the law of the Brownian motion $B_{t}$ and define a new probability measure $\mu$ on the space of continuous trajectories by
+\[
+d\mu=M_{T}d\nu
+\]
+with 
+\begin{equation}\label{likelihood}
+M_{t} = \exp\left(-\eps\int_{0}^{t}\xi_{s}\cdot dB_{s} - \frac{\eps^{2}}{2}\int_{0}^{t}|\xi_{s}|^{2}ds\right)\,,\quad 0\le t\le T\,.
+\end{equation}
+Technical details aside, the Girsanov theorem implies that $W_{t}$  for any function $f=f(\{X_{t}\}_{0\le t\le T})$ that is integrable \wrt $\nu$, we have the identity
+\[
+\bE_{\nu}[f] := \int f(\{X_{t}\}_{0\le t\le T})\, d\nu = \int f(\{X_{t}\}_{0\le t\le T}) \frac{d\nu}{d\mu}\,d\mu =: \bE_{\mu}[M^{-1}_{T}f]\,.
+\]
+where 
+\[
+M_{T}^{-1} = \exp\left(\eps\int_{0}^{T}\xi_{s}\cdot dW_{s} - \frac{\eps^{2}}{2}\int_{0}^{T}|\xi_{s}|^{2}ds\right)\,.
+\]
+is the density of $\nu$ relative to $\mu$.
 
-The time-dependent expectation value $O_0(t)$ of the path observable $O$ under the dynamics of the Langevin equation (\ref{eqn:langevin-1b})-(\ref{eqn:langevin-2b}) for $\eps=0$
-can be written as
-\begin{align}
-  O_0( t) =
-  \langle O[\vect X^0_s]\,\rangle_{0,t} = 
-  \int\dd \vect u\, f(\vect u, 0) 
-  \int_{\mc\{\vect u,0; t\}}
-  O[\vect X_s]\, \dd\mathcal P_0[\vect X_s],
-\end{align}
-where 
-$\langle \cdot \,\rangle_{0,t}$ is the ensemble average over all possible trajectories for $\eps=0$ in time interval $[0,t]$,
-$f(\vect x, 0)$ is the initial distribution that may not be the equilibrium distribution, and
-$\mc\{\vect u,0; t\}$ is the set of all admissible continuous trajectories starting from point
-$\vect u$ at time $0$, and ending at some position at time $t$.
+Linearization of $M_{T}^{-1}$ about $\eps=0$, assuming that the control has bounded variance, yields the alternative linear response formula 
+\begin{equation}\label{lr-alt}
+\lim_{\eps\to 0}\frac{\bE_{\nu}[f] + \bE_{\mu}[f]}{\eps} = \bE_{\mu}\left[f(\{X_{t}\}_{0\le t\le T})\int_{0}^{T}\xi_{s}\cdot dW_{s} \right]\,,
+\end{equation}
+where $X_{t}$ is the solution (\ref{sde2-3}).
+Note that this is the unperturbed nonequilibrium dynamics, because $W_{t}$ is a standard Brownian motion under $\mu$, and the expectation is over all realizations of (\ref{sde2-3}) starting from either any given initial condition $X_{0}=x$ or an arbitrary initial distribution.
+On the other hand, $X_{t}$ under $\nu$ corresponds to the perturbed dynamics, which should become clear upon comparing equations (\ref{sde2-2}) and (\ref{sde2-3}).
+Further note that $f$ and $W_{t}$ are not independent, hence the expectation on the right hand side of the equation is not zero in general.
 
-For $\eps\ge 0$ the  expectation value $O_\eps(t)$ takes the form
-\begin{align}
-  O_\eps(t) & =
-  \int \dd\vect u f(\vect x,0)
-  \int_{\mc\{\vect u,0; t\}} O[\vect X_s]\,
-  \, \dd\mathcal P_\eps[\vect X_s]   \\
-& = \int \dd\vect u f(\vect x,0)
-  \int_{\mc\{\vect u,0; t\}} O[\vect X_s]\,
-  \frac{\dd {\mathcal P}_\eps[\vect X_s]}{\dd\mathcal P_0[\vect X_s]}
-  \, \dd\mathcal P_0[\vect X_s] .   
-\end{align}
-By using the Girsanov formula, 
-\begin{align}\label{eqn:girsanov}
-  \frac
-  {\dd{\mathcal P_\eps}[\vect X_s]}
-  {\dd\mathcal P_0[\vect X_s]}  
-  =&\,
-  \exp
-  \big\{{\eps G[\vect X_s] - \frac{\eps^2}{2} H[\vect X_s]}
-  \big\}
-\end{align}
-where the short notations are defined by:
-\begin{align}
-  G[\vect X_s]
-  &= \frac{1}\sigma\int_0^t
-  \Delta F_e(s)\,\vect D(\vect q_s)\cdot\dd\vect W_s \\
-  H[\vect X_s]
-  &=
-  \frac 1{\sigma^2}\int_0^t
-  \Delta F_e^2(s)\,\vect D^2(\vect q_s)\,\dd s.
-\end{align}
-Expansion of the right-hand-side of Eq.~\eqref{eqn:girsanov} in $\eps$, we have
-\begin{align}\label{eqn:neq-response}
-   O_\eps(t) -
-  O_0(t)
-  =
-  \eps\, \big\langle O[\vect X_s]\cdot G[\vect X_s]\big\rangle_{0,t}
-  +
-  \frac{\eps^2}{2} \big\langle O[\vect X_s] \cdot (G^2[\vect X_s] - H[\vect X_s] )\big\rangle_{0,t}
-  +
-  \mo (\eps^3)
-\end{align}
-Some remarks concerning Eq.~\eqref{eqn:neq-response} seem in order:
+\paragraph*{Remark.} A quick-and-dirty derivation of the above change-of-measure formula can be easily obtained, if the noise covariance $a(\cdot)a(\cdot)^{T}$ has full rank almost everywhere with bounded inverse.
+Then, using Euler's method for (\ref{sde2}), it follows that (\ref{likelihood}) is basically the likelihood ratio between the time-discrete path densities of (\ref{sde2-2}) and (\ref{sde2-1}).
+
+\paragraph*{Remark.} By formally expanding $M^{-1}_T$ up to second order we get an analogous second order response formula: 
+\begin{equation}\label{2nd}
+\bE_{\nu}[f] \approx \bE_{\mu}[f] + \eps \bE_{\mu}\left[f\int_{0}^{T}\xi_{s}\cdot dW_{s} \right] +\frac{\eps^{2}}{2} \bE_{\mu}\left[f\left\{\left(\int_{0}^{T}\xi_{s}\cdot dW_{s}\right)^2 -\int_0^T |\xi_s|^2\,ds\right\} \right]
+\end{equation}
+
+\subsubsection*{Nonequilibrium Langevin equation}
+
+We now link our previous considerations with the previous case and consider a non-equilibrium Langevin equation. Specifically, we add a non-gradient perturbation to the Langevin equation 
+\begin{equation}\label{eqn:langevin-1b}
+b = \left(\nabla_{p}H, -\nabla_{q} H - \gamma\nabla_{p}H  + D(q) u \right)^{T}\,,\quad a= (\zero,\sigma)^{T}\,,
+\end{equation}
+where $u\in\R^{n}$ is some control variable and $D(\cdot)\in\R^{n\times n}$ satisfies the Fredholm alternative $\mathrm{range}(D(\cdot))\perp \mathrm{ker}(\sigma^{T})$, where $\mathrm{ker}(\sigma^T)$ denotes the kernel of $\sigma^T$.
+As we are interested in small perturbation $\delta u$ in the controls $u$, it is convenient to further specify what $v$ in (\ref{sde2}) is.
+Specifically, we assume that $v$ is of the form
+\begin{equation}\label{eqn:langevin-1c}
+v = (\zero, D(q)\,\delta u)
+\end{equation}
+so that equation (\ref{xi}) that determines the change of measure in terms of the auxiliary control variable $\xi$ (and thus the linear response) reads 
+\[
+\sigma \xi_{t} = D(Q_{t})\,\delta u_{t}\,.
+\]
+To be more explicit, the equation we are considering has the form
+\begin{equation}\label{eqn:langevin-1d}
+\begin{aligned}
+dQ_t & =  \nabla_p H(Q_t,P_t)\\
+dP_t & =  -\nabla_q H(Q_t,P_t) - \gamma \nabla_p H(Q_t,P_t) + D(Q_t)\,(u_t+\eps \delta u_t)+\sigma dB_t,
+\end{aligned}
+\end{equation}
+so that $X_t$ from above now is $(Q_t,P_t)$ and $x_t=(q_t,p_t)$ from above is the solution of our last equation for $\eps=0$.
+
+The equation is solvable by the requirement  $\mathrm{range}(D(\cdot))\perp \mathrm{ker}(\sigma^{T})$, which, if $\sigma$ has full rank, means that $D(\cdot)$ must be invertible almost everywhere.
+Hence the linear response formula (\ref{lr-alt})  becomes
+\begin{equation}\label{lr-alt2}
+\lim_{\eps\to 0}\frac{\bE_{\nu}[f] - \bE_{\mu}[f]}{\eps} =  \bE_{\mu}\left[f(\{(Q_{t},P_{t})_{0\le t\le T}\})\int_{0}^{T}(\sigma^{-1}D(Q_{s})\,\delta u_{s})\cdot dW_{s} \right]\,
+\end{equation}
+or, in other words: 
+\begin{equation}\label{eqn:neq-response}
+\bE_{\nu}[f] \approx \bE_{\mu}[f] + \eps\bE_{\mu}\left[f(\{Q_{t},P_{t}\}_{0\le t\le T})\int_{0}^{T}(\sigma^{-1}D(Q_{s})\,\delta u_{s})\cdot dW_{s} \right]\,.
+\end{equation}
+Here, as before, the expectation on the right is over all realizations of (\ref{eqn:langevin-1d}) for $\eps=0$ with arbitrary initial conditions (fixed or distributed); see also the remark below on the choice of initial conditions.
+
+\paragraph*{Remark.} Some comments on the above result are in order:
 \begin{itemize}
-\item The first term on the right-hand-side of Eq.~\eqref{eqn:neq-response} is the linear
-  response to the reference non-equilibrium process (driven by $F_e$, $\eps=0$).
-The next term is the second order response term.
-Obviously, one can easily write down higher orders of response by the Taylor expansion of the right-hand-side of Eq.~\eqref{eqn:girsanov}.
-\item When the observable $O[\vect X_s]$ only depends on the end point of the trajectory, i.e.
-  $O[\vect X_s] = O(\vect X_t)$, then $\big\langle O(\vect X_t)\cdot G[\vect X_s]\big\rangle_{0,t}$
-  gives the well-know time correlation form of the linear response formula.
-\item If one wants to calculate the same expectation value for a family of non-equilibrium perturbations $\Delta F_e(t)$ then 
-one does not need to repeat the calculations of Eq.~\eqref{eqn:neq-response} for every member of the family. 
+
+\item[(i)] The rightmost term in \eqref{eqn:neq-response} is the linear 
+  response to the reference non-equilibrium process (driven by $u_t$ with $\eps=0$).
+From (\ref{2nd}) we can also get the second order response term.
+The latter is bounded by the assumption that the controls have bounded second moment.
+
+\item[(ii)] In the above derivation, we have tacitly assumes that the reference and the perturbed non-equilibrium processes start from the same initial value or
+  have the same initial distribution.
+For fixed initial values (i.e.~points) this assumption cannot be relaxed (because otherwise $dP/dQ$ does not exist).
+For distributed initial values, however, there is no problem for the unperturbed and perturbed dynamics to have different initial distribution as long as both distributions are strictly positive almost everywhere.
+In this case one can apply a similar reweighing approach between the initial distribution as we used it for the trajectory ensemble.
+
+\item[(iii)] If one wants to calculate the same expectation value for a family of non-equilibrium perturbations $\delta u_t$ then 
+one does not need to repeat the calculations of \eqref{eqn:neq-response} for every member of the family. 
   If it is possible to
-  express different $\Delta F_e$ in the same basis, then the responses must only be
+  express different $\delta u_t$ in the same basis, then the responses must only be
   calculated for the single basis functions. Then with a linear combination of the responses on basis functions, one can derive
-  the responses of the whole family. 
-\item In the above derivation, we assume that the reference and the perturbed non-equilibrium processes
-  have the same initial distribution (both of them may be different from the equilibrium distribution),
-  but this is not a necessary. One can easily
-  apply a similar reweighing approach to the initial distribution as we used it for the trajectory ensemble.
+  the responses for the whole family. This feature will be used below when discussion optimal control as an application of the linear response formula.
+
 \end{itemize}
@@
 
-\subsection{Optimal Forcing}
+\section{Application of the nonequilibrium response formula: optimal control}
 
@@
 \subsection{The metastability of a single-well potential}
-The metastability of the system at time $t$ can be investigated by
-calculating the function $\Delta \mt_\tau f(\vect x, t) = (\mt_\tau
--\mi) f(\vect x, t)$, where $f(\vect x, t)$ is the probability
-distribution at position $\vect x$ and time $t$, the propagator
-$\mt_\tau$ is defined by $\mt_\tau = e^{\tau\fwg_t^{2\beta}}$, and
-$\mi$ is the identity.  Notice that here the physical time $t$ is
-frozen, and the distribution $f$ evolves a long the artificial time
-$\tau$ at temperature $1/(2\beta)$.  If the position $\vect x$ is
-``stable'', then intuitively the probability of the system being at
-position $\vect x$ will be higher when the temperature is lower, i.e.,
-$\Delta \mt_\tau f(\vect x, t) > 0$.  Otherwise, when position $\vect
-x$ is not stable, $\Delta \mt_\tau f(\vect x, t) < 0$.  The
-artificial time $\tau$ should be chosen smaller than the investigated
-relaxation time scales of the original system. However, on the other
-hand, $\tau$ cannot be arbitrarily small, because otherwise the
-statistical uncertainty would not allow us to distinguish $\mt_\tau f$
-and $f$.
 
@@
   \caption{The single-well potential with splitting splitting drving
-    force at time $t_c = 20$ ps.  Since the drving force is
-    conservative ($\vect D(\vect q) = -\nabla_{\vect q}V(\vect q)$),
+    force at time $T = 20$ ps.  Since the drving force is
+    conservative ($ D( q) = -\nabla_{q}V( q)$),
     we plot the non-equilibrium drving energy of the system.  The red,
@@
     non-equilibrium driving, with non-equilibrium driving and
-    the non-equilibrium driving perturbed by $\eps\Delta F_e(t)$,
+    the non-equilibrium driving perturbed by $\eps\delta u(t)$,
     respectively.  }
@@
 
-% The core set plays a very important role in the MSM theory.  We
-% therefore want to test our idea of non-equilibrium response theory by
-% identifying the core sets in a non-equilibrium driven system.
-% There are several ways to define a core set.
-% One possible choice  is to calculate the stability of the phase space
-% distribution: $\Delta \mt_\tau f(\vect x, t) = (\mt_\tau -\mi) f(\vect
-% x, t)$. And the core sets are defined by
-% \begin{align}
-%   C_\tau(t) = \{\vect x\vert \Delta \mt_\tau f(\vect x, t) \geq 0\}
-% \end{align}
-% Here the propagator $\mt_\tau$ is defined by $\mt_\tau =
-% e^{\tau\fwg_t^{2\beta}}$.  Notice that here the physical time $t$ is
-% frozen, and the distribution $f$ evolves a long the artificial time
-% $\tau$ at temperature $1/(2\beta)$. The artificial time $\tau$ should
-% be chosen smaller than the investigated relaxation time scales of the
-% original system. However, on the other hand, $\tau$ cannot be
-% arbitrarily small, because otherwise the statistical uncertainty would
-% not allow us to distinguish $\mt_\tau f$ and $f$.
-
 We use the idea of non-equilibrium linear response to investigate the
-metastability a one-dimensional model system: one particle in a
+nonequilibrium phase space probability density distribution, denoted by $f(q, p, t)$, of a one-dimensional model system: one particle in a
 splitting single-well potential.  For convenience, we let the mass of
@@
 \begin{align}
-  \mh (\vect p, \vect q) = \frac 12 \vect p^2 + U(\vect q) 
+  \mh ( p,  q) = \frac 12  p^2 + U( q) 
 \end{align}
@@
 \begin{align}
-  U(\vect q) = \frac12\,k\,\vect q^2 
+  U( q) = \frac12\,k\, q^2 
 \end{align}
@@
 %$\textsf{kJ} / (\textsf{mol nm}^4)}$
-The non-equilibrium driving $\vect D$ is given by a conservative force:
+The non-equilibrium driving $ D$ is given by a conservative force:
 \begin{align}
-  \vect D(\vect q) = -\nabla_{\vect q} V(\vect q) 
+   D( q) = -\nabla_{ q} V( q) 
 \end{align}
-where the driving potential $V(\vect q)$ has a Gaussian profile:
+where the driving potential $V( q)$ has a Gaussian profile:
 \begin{align}
-  V(\vect q) = \frac{1}{\sqrt{2\pi \sigma^2}}
-  \exp\Big\{-\frac{\vect q^2}{2\sigma^2}\Big\}
+  V( q) = \frac{1}{\sqrt{2\pi \sigma^2}}
+  \exp\Big\{-\frac{ q^2}{2\sigma^2}\Big\}
 \end{align}
 we use $\sigma = 0.16$~\textsf{nm}.  The strength of non-equilibrium
-driving $F_e(t)$ is set to be a unitless constant $k_e$, when the time $t$ is
-larger than some warm-up time $t_c$.  When $t \leq t_c$, the strength
-of $F_e(t)$ is linearly growing, i.e. $ F_e(t) = k_e\cdot t/t_c$.  We
-consider the perturbation to the system given by $\eps\Delta F_e(t) =
-\eps$ when $t> t_c$, and $\eps\Delta F_e(t) = \eps \cdot t/t_c$ when
-$t\leq t_c$.
-We consider the following parameters: $t_c = 20$~\textsf{ps}, 
+driving $u(t)$ is set to be linearly growing, i.e. $ u(t) = k_e\cdot t/T$,
+where $k_e$ is a unitless constant.
+We
+consider the perturbation to the system given by $\eps\delta u(t) = \eps \cdot t/T$.
+We consider the following parameters: end time $T = 20$~\textsf{ps}, 
 $k_e = 1$ and $\eps = 1$, see Fig.~\ref{fig:tmp4}
 for the non-equilibrium driving potential and perturbed
-potential at time $t = t_c$.
-$\tau$ is set to be $1~\textsf{ps}$.
+potential at time $t = T$. The initial distribution $f(q, p, 0)$ is set to
+be equilibrium distribution of the unperturbed system.
 
@@
   \includegraphics[width=0.95\textwidth]
-  {figs/fig-split-str2-2d-more.eps}
-  \caption{ The plot of $\Delta\mt_\tau f(\vect x,t)$ in the phase
-    space under perturbed non-equilibrium driving ($k_e = 1$ and $\eps
-    = 1$). From left to right the columns present results at time $t =
+  {figs/fig-split-str2-2d-distrib-no2nd.eps}
+  \caption{ The plot of $ f(q,p,t)$ in the phase
+    space under perturbed non-equilibrium driving. From left to right the columns present results at time $t =
     0$, 5, 10 and 20~\textsf{ps}.  The 1st row: the brute force
@@
     traditional equilibrium linear response theory. In this case we
-    effectivly consider an equilibrium system free of driving ($k_e =
-    0$), and the perturbation $\eps$ is set to 2.  The 3rd row: the
-    second order equilibrium response result.  The 4th row: the
+    effectivly consider an equilibrium system free of driving, and the perturbation $\eps$ is set to 2.  The 3rd row: the
     non-equilibrium linear response result.  }
@@
 
-Fig.~\ref{fig:tmp6} presents the numerical results of the splitting
+Fig.~\ref{fig:tmp6} presents the numerical results of the phase space probability distribution for the splitting
 single-well potential. From left to right the four rows presents the
-stability of the system at time $t = 0$, 5, 10 and 20~\textsf{ps}. The first
+distribution of the system at time $t = 0$, 5, 10 and 20~\textsf{ps}. The first
 row presents the brute force non-equilibrium simulation.  It is clear
-that at the beginning, there is only one stable region in the phase,
-i.e. arount $q = 0$. As the time evolves, an energy barrier develops
-in the center of the simulation region, therefore, the single stable
-region splits into two equally sized stable regions.  In the end, the two
-stable regions are entirly seperated.  This brute non-equilibrium
+that at the beginning, the distribution has only one peak,
+i.e. arount $q = 0$ and $p = 0$. As the time evolves, an energy barrier develops
+in the center of the simulation region, therefore, the single peak
+splits into two equally sized peaks.  In the end, the two
+peaks are entirly seperated.  This brute force non-equilibrium
 simulation serves as the precise result that the response theory
@@
 traditional equilibrium linear response theory.  Please notice that in
-this case, since the reference simulation is in equilibrium (i.e. $k_e = 0$),
-the effective perturbation is of strength $\eps = 2$.  At $t =
-5$~\textsf{ps}, the accuracy of the equilibrium linear response is
-perfect. When $t = 10$~\textsf{ps}, the left stable region is a little
-over estimate. This may be due to the statistical uncertainty. At $t =
-20$~\textsf{ps}, stability of the two stable regions are too strong,
-and the unstable region in between is less stable than the brute
-force result. The accuracy of the equilibrium linear response is poor
-at $t \geq 10$~\textsf{ps}, because the strength of perturbation is no
+this case, since the reference simulation is in equilibrium, so 
+the effective perturbation is of strength $\eps = 2$.  At $t \leq
+15$~\textsf{ps}, the accuracy of the equilibrium linear response is
+perfect. At $t =
+20$~\textsf{ps}, magnitude of the peaks are relatively too strong,
+and the gap between them is actually negative.
+Since the probability distribution is always positive, the
+numerical solution of the equilibrium linear response is qualitatively wrong.
+The poor accuracy is due to the fact that the strength of perturbation is no
 longer small so that the preliminary assumption of the reponse theory
-is not well fulfilled. On the third row, we present second order
-equilibrium response theory. The using of a higher order response term
-should improve the numerical accuracy, however, this is not the case,
-because the second order response term cannot be accuratly calculated
-due to the huge statistical error. By doing longer reference
-equilibrium simulations may improve the accuracy of statistics, but it
-is not ecnomic to do that.  On the fourth row of Fig.~\ref{fig:tmp6},
+is not well fulfilled.  On the third row of Fig.~\ref{fig:tmp6},
 we present the linear response starting from a non-equilibrium driving
-of $k_e = 1$.  The numerical results is satisfactorily consistent with
+of $u(t)$.  The numerical results is satisfactorily consistent with
 the brute force non-equilibrium simulation, because the perturbation
@@
 % \begin{align}
-%   \mh (\vect p, \vect q) = \frac 12 \vect p^2 + U(\vect q) 
+%   \mh ( p,  q) = \frac 12  p^2 + U( q) 
 % \end{align}
@@
 % \begin{align}
-%   U(\vect q) = \frac12 k (\vect q^2 - a^2)^2
+%   U( q) = \frac12 k ( q^2 - a^2)^2
 % \end{align}
@@
 % \begin{align}
-%   \vect D(\vect q) = -\nabla_{\vect q} V(\vect q) = 1
+%    D( q) = -\nabla_{ q} V( q) = 1
 % \end{align}
-% Here $V(\vect q) = -\vect q$   effectively tilts the original
-% potential $U(\vect q)$ (see the gree and blue lines in Fig.~\ref{fig:tmp1}).
-% The strength of $F_e(t)$ is
+% Here $V( q) = - q$   effectively tilts the original
+% potential $U( q)$ (see the gree and blue lines in Fig.~\ref{fig:tmp1}).
+% The strength of $u(t)$ is
 % set to be $k_e$, when the time $t$ is larger than some warm-up time $t_c$.
-% When $t \leq t_c$, the strength of $F_e(t)$ is linearly growing,
-% i.e. $ F_e(t) = k_e\cdot t/t_c$.
+% When $t \leq t_c$, the strength of $u(t)$ is linearly growing,
+% i.e. $ u(t) = k_e\cdot t/t_c$.
 % We use the value $t_c = 20$~\textsf{ps}, so that the perturbation increases slow
@@
 % consider the stability of the
-% distribution $f(\vect x, t)$ at time $t$: $\Delta\mt_\tau f(\vect x, t)$.
+% distribution $f( x, t)$ at time $t$: $\Delta\mt_\tau f( x, t)$.
 % The larger
@@
   \centering
-  \includegraphics[]{figs/fig-ctr-stat.eps}
+  \includegraphics[]{figs/fig-ctr-stat-2.eps}
   \caption{The optimal forcing.  In this plot, $T=20$ ps.  The red
@@
     piece is 1 ps.  The blue insertions show the shape of the
-    non-equilibrium driving potential $U(\vect q) + F_e(t)V(\vect
+    non-equilibrium driving potential $U( q) + F_e(t)V(
     q)$ at time indicated by the black arrow.  In the lower plot, the
-    dashed lines display the value of $P(\vect X_t)$ as the time
+    dashed lines display the value of $P( X_t)$ as the time
     evolves.  The solid lines present $O[F_e](t)$.  }\label{fig:tmp7}
@@
 \begin{align}
-  U(\vect q) = \frac12 k (\vect q^2 - a^2)^2
+  U( q) = \frac12 k ( q^2 - a^2)^2
 \end{align}
@@
 %$\textsf{kJ} / (\textsf{mol nm}^4)}$
-The perturbation is given by a conservative tilting to $U(\vect q)$:
+The perturbation is given by a conservative tilting to $U( q)$:
 \begin{align}
-  \vect D(\vect q) = -\nabla_{\vect q} V(\vect q) = 1
+   D( q) = -\nabla_{ q} V( q) = 1
 \end{align}
-Here $V(\vect q) = -\vect q$. We want to
+Here $V( q) = - q$. We want to
 optimally design the tilting such that the probability
@@
 \begin{align}
-  O[F_e](T) &= P(\vect X_T) -
+  O[F_e](T) &= P( X_T) -
   \eta\,
-  \big\langle \int_0^T\vert F_e(t)\vect D(\vect q)\vert^2\, dt\,
+  \big\langle \int_0^T\vert F_e(t) D( q)\vert^2\, dt\,
   \big\rangle
 \end{align}
-where $P(\vect X_T)$ is probability of the end point of the trajectory
-$\vect X_T$ being in the right well.  It is
-clear that for an unbiased double-well, $P(\vect X_T)$
+where $P( X_T)$ is probability of the end point of the trajectory
+$ X_T$ being in the right well.  It is
+clear that for an unbiased double-well, $P( X_T)$
 is 0.5.  The integral is the ``cost'' of the control and $\eta$ is a
@@
 control $F_e(t)$ is shown in the upper plot, and the value of
-optimization target $O[F_e](T)$ and the value of $P(\vect X_T)$ are
+optimization target $O[F_e](T)$ and the value of $P( X_T)$ are
 given in the lower plot by the solid and dashed lines, respectively.
@@
 from the optimization, and the optimized value of $O[F_e](T)$ and
-$P(\vect X_T)$ are bigger, which means the end position of $\vect X_t$
+$P( X_T)$ are bigger, which means the end position of $ X_t$
 is more biased towards the right well. This observation is consistent
@@
 
-When the strength of the perturbation $\eps$ is relatively small, then
-the traditional linear response theory gives good approximation.  When
-the perturbation reaches the strength, that is not small for the
-linear response, including the second order response term does not
-substantially improve the accuracy.
-In the splitting single-well example, the statistical
-noise of the second order response is so big that overwhelms the
-information. Spending more computational effort on reducing the noise
-is not economic. The non-equilibrium response theory developed by this
-work allows one to calculate the response from some intermediate
-non-equilibrium simulation, so the effective perturbation is much
-smaller than it would be calculated from an equilibrium
-simulation.
-In both examples considered by this paper, the
-numerical results of the proposed idea  satisfactory consistent with the
-brute force non-equilibrium simulation.
-From the face that, in a strongly driven system,
-one should do an intermediate non-equilibrium
-simulation and calculate the response from it,
-one may argue that this may not be able to
-save a lot computational cost comparing with a brute force
-non-equilibrium simulation.  However, in practice, one need the
-response theory for the following reasons: (1) One may want to study
-multiple non-equilibrium system that are not very far from each other,
-so only do one system and apply the linear response idea takes
-advantage over doing non-equilibrium simulation for each of them.
-(2) A stronger driven non-equilibrium system may be more difficult to
-simulation directly comparing with a weakly driven system. For
-example, the energy barrier between two rare events may be increased
-greatly due to the external driven. (3) By the traditional linear
-response theory, some information concerning the intermediate
-non-equilibrium system can be derived from an standard equilibrium
-simulation.
-With this extra information, may be able to design better simulation methods
-for the intermediate non-equilibrium system,
-so that it is not as expensive as a brute force non-equilibrium
-simulation starting with nothing know before hand.
-(4) A thermodynamic multistep integration method can be developed from the
-response theory plus idea (3): the non-equilibrium driving is increased
-step-by-step and the total simulation effort at all the steps is still
-cheaper than doing one brute force non-equilibrium simulation for the
-target driven system.
-Till now we have not investigated idea (3) and (4) mentioned above,
-however, we believe the non-equilibrium response theory is a very
-important step toward it and by itself deserves a separate careful
-investigation.
+We derived  first and second order response formulas for molecular dynamics (driven Langevin dynamics) starting from general nonequilibrium distributions.
+For the special case of the initial distribution being the equilibrium distribution of the unperturbed dynamics, the novel linear response formula siimplifies to the well-known standard formula.
+We validated the novel formula in numerical experiments in comparison to brute-force nonequilibrium simulations.
+There, we demonstrated that the novel linear response formula allows to extend the algorithmic use of linear response theory to significantly stronger perturbutions of the system since it permits intermediate steps based on partially propagated nonequilibrium distributions.
+
+By means of this theory we outlined how to use the novel linear response formula for the computation of optimal controls in molecular dynamics where one desires to find the optimal perturbation/control that maximizes a target functional, that is, a certain expectation value (like the population of a certain region of state space) under a constraint on the energy used in the perturbation.
+Application of the novel nonequilibrium theory allows to compute the gradient of the target functional by computing expectation values \emph{only} for the dynamics at hand which permits efficient application of standard optimization techniques.
+We illustrated this technique in application to a simple test case and validated it in comparison to brute force optimization.
 
@@
 % \begin{align}
-%   \vect C(\vect p, \vect q) = 0; \qquad
-%   \vect D(\vect p, \vect q) = -\nabla_{\vect q} V(\vect q) = 1
+%    C( p,  q) = 0; \qquad
+%    D( p,  q) = -\nabla_{ q} V( q) = 1
 % \end{align}
-% Here $V(\vect q) = -\vect q$ is  effectively tilting the original
-% potential $U(\vect q)$. The strength of the perturbation $F_e(t)$ is given
+% Here $V( q) = - q$ is  effectively tilting the original
+% potential $U( q)$. The strength of the perturbation $F_e(t)$ is given
 % by the following function:
@@
 % 1-d, the projection $\mathcal P$ is identity.  We consider the stability of the
-% distribution $f(\vect x, t)$ at time $t$: $\Delta\mt_\tau f(\vect x, t)$.
+% distribution $f( x, t)$ at time $t$: $\Delta\mt_\tau f( x, t)$.
 % The larger
@@
 %     Tilting double well potential testing case:
-%     the plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
+%     the plot of $\Delta\mt_\tau f( x,t)$  under perturbation of
 %     $F_e^{\textrm{max}} = 2.0$.
@@
 %     Tilting double well potential:
-%     the plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
+%     the plot of $\Delta\mt_\tau f( x,t)$  under perturbation of
 %     $F_e^{\textrm{max}} = 4.0$.
@@
 % the snapshot at $t=20$~\textsf{ps}
-% of $\Delta\mt_\tau f(\vect x,t)$ under the maximum perturbation
+% of $\Delta\mt_\tau f( x,t)$ under the maximum perturbation
 % strength $F_e^{\max} = 2.0$. Response formula truncated
@@
 % Fig.~\ref{fig:new-tmp3} presents the $t=20$~\textsf{ps}
-% snapshots of $\Delta\mt_\tau f(\vect x,t)$
+% snapshots of $\Delta\mt_\tau f( x,t)$
 % under a larger perturbations $F_e^{\max} = 4.0$.
@@
 % \begin{align}
-%   U(\vect q) = \frac12\,k\,\vect q^2 
+%   U( q) = \frac12\,k\, q^2 
 % \end{align}
@@
 % \begin{align}
-%   \vect C(\vect p, \vect q) = 0; \qquad
-%   \vect D(\vect p, \vect q) = -\nabla_{\vect q} V(\vect q) 
+%    C( p,  q) = 0; \qquad
+%    D( p,  q) = -\nabla_{ q} V( q) 
 % \end{align}
-% where $V(\vect q)$ is a Gaussian type potential:
+% where $V( q)$ is a Gaussian type potential:
 % \begin{align}
-%   V(\vect q) = \frac{1}{\sqrt{2\pi \sigma^2}}
-%   \exp\Big\{-\frac{\vect q^2}{2\sigma^2}\Big\}
+%   V( q) = \frac{1}{\sqrt{2\pi \sigma^2}}
+%   \exp\Big\{-\frac{ q^2}{2\sigma^2}\Big\}
 % \end{align}
@@
 %     Splitting single-well potential testing case:
-%     the plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
+%     the plot of $\Delta\mt_\tau f( x,t)$  under perturbation of
 %     $F_e^{\textrm{max}} = 1.0$.
@@
 %     splitting single well potential:
-%     the plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
+%     the plot of $\Delta\mt_\tau f( x,t)$  under perturbation of
 %     $F_e^{\textrm{max}} = 2.0$.

### 2013-06-28 writedown/linear.response/paper.tex
changed to Mol. Phys. template

@@
 % \documentclass[aip,jcp,preprint,unsortedaddress,a4paper,onecolum]{revtex4-1}
-\documentclass[aip,jcp,a4paper,reprint,onecolumn]{revtex4-1}
+\documentclass[]{tMPH2e}
 % \documentclass[aps,pre,twocolumn]{revtex4-1}
@@
 
-\usepackage[fleqn]{amsmath}
-\usepackage{amssymb,amsthm}
-\usepackage[dvips]{graphicx}
-\usepackage{color}
-\usepackage{tabularx}
-\usepackage{algorithm}
-\usepackage{algorithmic}
+% \usepackage[fleqn]{amsmath}
+% \usepackage{amssymb,amsthm}
+% \usepackage[dvips]{graphicx}
+% \usepackage{color}
+% \usepackage{tabularx}
+% \usepackage{algorithm}
+% \usepackage{algorithmic}
 
@@
 \title{Linear response theory for a system under non-equilibrium conditions}
-\author{Han Wang}
-\author{Carsten Hartmann}
-\affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}
-\author{Christof Sch\"utte}
-\affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}
-\affiliation{Zuse Institute Berlin (ZIB), Germany}
+
+\author{Han Wang$^{a}$ and Carsten Hartmann$^{a}$ and Christof Sch\"utte$^{a,b}$$^{\ast}$\thanks{$^\ast$Corresponding author.
+Email: Christof.Schuette@fu-berlin.de}\\\vspace{6pt}$^{a}${\em Institute for Mathematics, Freie Universit\"at Berlin, Germany}\\$^{b}${\em Zuse Institute Berlin (ZIB), Germany}}
+
+% \author{Han Wang}
+% \author{Carsten Hartmann}
+% \affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}
+% \author{Christof Sch\"utte}
+% \affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}
+% \affiliation{Zuse Institute Berlin (ZIB), Germany}
 % \affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}
 
-\date{\today}
+\maketitle
 
@@
 \end{abstract}
-
-\maketitle
 
@@
 Here we propose an alternative (purely formal) derivation of the linear response result, that is based on a change of drift in the corresponding SDE and which allows for an easy generalization of the above linear response result to nonequilibrium systems.
-We will briefly review the idea of the change of drift via Girsanov transformations; for details we refer to the textbook \cite{oksendal}.
+We will briefly review the idea of the change of drift via Girsanov transformations; for details we refer to the textbook~\cite{oksendal2003stochastic}.
 Now let $x_{t},X_{t}$ be the solutions to the following stochastic differential equations:
@@
 \end{equation}
-It follows from the Girsanov theorem \cite[Thm.~8.6.8]{oksendal2003}, sometimes also called \emph{Cameron-Martin-Girsanov theorem} \cite{}, that $W_{t}$ is again a Brownian motion under a new probability measure that has a density \wrt the Gaussian probability measure that is generated by the Brownian motion $B_{t}$.
+It follows from the Girsanov theorem \cite[Thm.~8.6.8]{oksendal2003}, sometimes also called \emph{Cameron-Martin-Girsanov theorem} \cite{a}, that $W_{t}$ is again a Brownian motion under a new probability measure that has a density \wrt the Gaussian probability measure that is generated by the Brownian motion $B_{t}$.
 Specifically, let $\nu$ denote the law of the Brownian motion $B_{t}$ and define a new probability measure $\mu$ on the space of continuous trajectories by
@@
 
-% \newpage
-% \section{Numerical results}
-% \subsection{Tilting double-well potential}
-
-% \begin{figure}
-%   \centering
-%   \includegraphics[width=0.4\textwidth]{figs/fig-tilt-pot.eps}
-%   \includegraphics[width=0.4\textwidth]{figs/fig-split-pot.eps}
-%   \caption{Two testing cases presented by this paper.
-%     Left: the double-well potential with tilting perturbation.
-%     Right: the single-well potential with splitting perturbation.}
-%   \label{fig:new-tmp1}
-% \end{figure}
-
-% We test the idea by a one-dimensional model system: one particle in a
-% tilting double-well potential. For convenience, we let the mass of the
-% particle to be 1 \textsf{amu}. The unperturbed
-% Hamiltonian of the system is given by:
-% \begin{align}
-%   \mh ( p,  q) = \frac 12  p^2 + U( q) 
-% \end{align}
-% with potential
-% \begin{align}
-%   U( q) = \frac12 k ( q^2 - a^2)^2
-% \end{align}
-% Here $k = 8$~$\textsf{kJ} / (\textsf{mol nm}^4)$, and $ a = 1\ \textsf{nm}$.
-% Notice at room temperature $300\ \textsf{K}$, $k_BT = 2.48$~\textsf{kJ/mol}.
-% See the red line in Fig.~\ref{fig:tmp1}.
-% %$\textsf{kJ} / (\textsf{mol nm}^4)}$
-% The perturbation is given by
-% \begin{align}
-%    C( p,  q) = 0; \qquad
-%    D( p,  q) = -\nabla_{ q} V( q) = 1
-% \end{align}
-% Here $V( q) = - q$ is  effectively tilting the original
-% potential $U( q)$. The strength of the perturbation $F_e(t)$ is given
-% by the following function:
-% \begin{align}
-%   F_e(t) = 
-%   \begin{cases}
-%     F_e^{\textrm{max}}\times (t / t_c) & t < t_c \\
-%     F_e^{\textrm{max}} & t \geq t_c
-%   \end{cases}
-% \end{align}
-% $t_c = 20$~\textsf{ps} so that the perturbation increases slow
-% enough: the typical decaying time scale of the correlation function
-% in~\eqref{eqn:core-identify-approx-2} is roughly 3~\textsf{ps}.
-% We choose $\tau = 1$~\textsf{ps}.
-% % We choose $\mt_\alpha$
-% % to be the propagator of the Langevin dynamics at temperature $150$
-% % \textsf{K} with $\alpha = 1$ \textsf{ps}.
-% Since the dynamics is in
-% 1-d, the projection $\mathcal P$ is identity.  We consider the stability of the
-% distribution $f( x, t)$ at time $t$: $\Delta\mt_\tau f( x, t)$.
-% The larger
-% this value, the more stable the core sets are.
-
-% \begin{figure}
-%   \centering
-%   \includegraphics[width=0.95\textwidth]
-%   {figs/fig-tilt-str2-simple.eps}
-%   \caption{
-%     Tilting double well potential testing case:
-%     the plot of $\Delta\mt_\tau f( x,t)$  under perturbation of
-%     $F_e^{\textrm{max}} = 2.0$.
-%     From left to right:
-%     (a) $t=0$~\textsf{ps},
-%     the brutal force non-equilibirum simulation,
-%     (b) $t=20$~\textsf{ps},
-%     the brutal force non-equilibirum simulation,
-%     (c) $t=20$~\textsf{ps},
-%     response formula truncated at order $n=1$,
-%     and
-%     (d) $t=20$~\textsf{ps},
-%     response formula truncated at order $n=2$.
-%     The equilibirum state serves as the reference of
-%     the response simulations.
-%   }
-%   \label{fig:new-tmp2}
-% \end{figure}
-
-% \begin{figure}
-%   \centering
-%   \includegraphics[width=0.95\textwidth]
-%   {figs/fig-tilt-str4-simple.eps}
-%   \caption{
-%     Tilting double well potential:
-%     the plot of $\Delta\mt_\tau f( x,t)$  under perturbation of
-%     $F_e^{\textrm{max}} = 4.0$.
-%     From left to right:
-%     (a) $t=0$~\textsf{ps},
-%     the brutal force non-equilibirum simulation,
-%     (b) $t=20$~\textsf{ps},
-%     the brutal force non-equilibirum simulation,
-%     (c) $t=20$~\textsf{ps},
-%     response formula with reference process $F_e^{\max} = 0.0$
-%     (i.e. equilibirum simulation),
-%     and
-%     (d) $t=20$~\textsf{ps},
-%     response formula with reference process $F_e^{\max} = 2.0$.
-%     The response formula are truncated at order $n=1$.
-%   }
-%   \label{fig:new-tmp3}
-% \end{figure}
-
-% The Fig.~\ref{fig:new-tmp2} and \ref{fig:new-tmp3} present
-% the application of the response theory to
-% calculate core sets of the tilting double-well potential.
-% Fig.~\ref{fig:new-tmp2} presents
-% the snapshot at $t=20$~\textsf{ps}
-% of $\Delta\mt_\tau f( x,t)$ under the maximum perturbation
-% strength $F_e^{\max} = 2.0$. Response formula truncated
-% at both order $n=1$ and $n=2$ are compared to 
-% the brutal force non-equilibirum simulation.
-% No obviously difference is shown. The second order response result
-% is slightly more precise than the first order one, but the
-% statistical error is also larger, which makes the profile noisy.
-% This figure tells us that
-% (1) the modeling error of the linear response (1st order) is
-% negligibly small; (2) the statistical error of numerically measuring the
-% linear response is also small.
-% Fig.~\ref{fig:new-tmp3} presents the $t=20$~\textsf{ps}
-% snapshots of $\Delta\mt_\tau f( x,t)$
-% under a larger perturbations $F_e^{\max} = 4.0$.
-% The linear reponse using the equilibirum process as the reference
-% shows poor accuracy: the left well vanishes in this case,
-% while the linear response produces  artifical structures.
-% Since the profile is smooth, which indicates
-% the statistical error of calculating the linear term is small,
-% the deviation comes from the modellng error, namely the contribution
-% of higher order response terms.
-% By including the second order response term reduces the left
-% artifical core set, but also creates artifical finer grained
-% structures. The statistical accuracy is also worse than the linear response.
-% Plot (d) present the result of the linear response
-% formula using reference process $F_e^{\max} = 2.0$, which
-% is the process presented in Fig.~\ref{fig:new-tmp2}.
-% Much better modelling accuracy is achieved:
-% the left artifical core set disappear and the
-% statistical accuracy is also satisfactory.
-
-% % See Fig. \ref{fig:tmp2} and \ref{fig:tmp3} for numerical result.  Both
-% % the direct non-equilibrium simulation and the response
-% % approximations are presented. Time slices $t = 0$, 10, 15 and 20
-% % \textsf{ps} are shown.  When the perturbation is small,
-% % i.e. $F_e^{\textrm{max}} = 2$, the direct non-equilibrium simulation
-% % and the linear response results are accurately consistent.
-% % The 2nd order response formula~\eqref{eqn:pert-approx-1} and
-% % \eqref{eqn:pert-approx-2} are even better, but the numerical error
-% % is bigger.
-% % As the
-% % perturbation grows stronger, the left well becomes less stable, while
-% % the right well becomes more stable.  When the perturbation is big,
-% % i.e. $F_e^{\textrm{max}} = 4$, the linear response result is no
-% % longer precise.
-% % When $t\geq 20$ \textsf{ps}, the left well basically
-% % disappears, however, the linear response calculation presents
-% % artifical occilation at the right well.
-% % The 2nd order response also has artifical effect, and the numerical
-% % uncertainty is high.
-% % The 4th rwo of Fig.~\ref{fig:tmp3} presents the 1st order result of the
-% % response formula~\eqref{eqn:pert-approx-1} and
-% % \eqref{eqn:pert-approx-2}, with
-% % respect to the reference state $F_e^{\textrm{max}} = 2$. The result
-% % is impressively improved compared with the linear response with respect to the
-% % equilibrium state.
-
-% \subsection{Splitting single-well potential}
-
-% In this splitting single-well potential example, the unperturbed
-% potential is simply the harmonic potential given by:
-% \begin{align}
-%   U( q) = \frac12\,k\, q^2 
-% \end{align}
-% Here $k = 8$~$\textsf{kJ} / (\textsf{mol nm}^2)$.
-% Notice at room temperature $300\ \textsf{K}$, $k_BT = 2.48$~\textsf{kJ/mol}.
-% See the red line in Fig.~\ref{fig:tmp4}.
-% %$\textsf{kJ} / (\textsf{mol nm}^4)}$
-% The perturbation is again given by
-% \begin{align}
-%    C( p,  q) = 0; \qquad
-%    D( p,  q) = -\nabla_{ q} V( q) 
-% \end{align}
-% where $V( q)$ is a Gaussian type potential:
-% \begin{align}
-%   V( q) = \frac{1}{\sqrt{2\pi \sigma^2}}
-%   \exp\Big\{-\frac{ q^2}{2\sigma^2}\Big\}
-% \end{align}
-% we use $\sigma = 0.16$~\textsf{nm}. Please see Fig~\ref{fig:tmp4}.
-% For the results, please see Fig.~\ref{fig:tmp5} -- Fig.~\ref{fig:tmp6}.
-
-% \begin{figure}
-%   \centering
-%   \includegraphics[width=0.95\textwidth]
-%   {figs/fig-split-str1-simple.eps}
-%   \caption{
-%     Splitting single-well potential testing case:
-%     the plot of $\Delta\mt_\tau f( x,t)$  under perturbation of
-%     $F_e^{\textrm{max}} = 1.0$.
-%     From left to right:
-%     (a) $t=0$~\textsf{ps},
-%     the brutal force non-equilibirum simulation,
-%     (b) $t=20$~\textsf{ps},
-%     the brutal force non-equilibirum simulation,
-%     (c) $t=20$~\textsf{ps},
-%     response formula truncated at order $n=1$,
-%     and
-%     (d) $t=20$~\textsf{ps},
-%     response formula truncated at order $n=2$.
-%     The equilibirum state serves as the reference of
-%     the response simulations.
-%   }
-%   \label{fig:new-tmp5}
-% \end{figure}
-
-% \begin{figure}
-%   \centering
-%   \includegraphics[width=0.95\textwidth]
-%   {figs/fig-split-str2-simple.eps}
-%   \caption{
-%     splitting single well potential:
-%     the plot of $\Delta\mt_\tau f( x,t)$  under perturbation of
-%     $F_e^{\textrm{max}} = 2.0$.
-%     From left to right:
-%     (a) $t=0$~\textsf{ps},
-%     the brutal force non-equilibirum simulation,
-%     (b) $t=20$~\textsf{ps},
-%     the brutal force non-equilibirum simulation,
-%     (c) $t=20$~\textsf{ps},
-%     response formula with reference process $F_e^{\max} = 0.0$
-%     (i.e. equilibirum simulation),
-%     and
-%     (d) $t=20$~\textsf{ps},
-%     response formula with reference process $F_e^{\max} = 1.0$.
-%     The response formula are truncated at order $n=1$.
-%   }
-%   \label{fig:new-tmp6}
-% \end{figure}
-
-\section*{References}
 \bibliography{ref}{}
-\bibliographystyle{unsrt}
+\bibliographystyle{tMPH}
 

### 2013-06-29 writedown/linear.response/paper.tex
bk

@@
 \newcommand{\cA}{\mathcal A}
+\newcommand{\cU}{\mathcal U}
 \newcommand{\bwg}{{\mathcal B}}
@@
 \begin{abstract}
-  In this paper, we firstly review standard equilibrium linear response
-  theory in the context of molecular dynamics, and then generalize it to systems in nonequilibrium and under a nonequilibrium
-  driving. We calculate not only the linear response term, but also
-  second order terms for small perturbations to the reference dynamics.
+  In this paper, we propose a straightforward generalization of linear response theory to systems in nonequilibrium that are subject to nonequilibrium driving. 
+  We briefly revisit the standard linear response result for equilibrium systems, where we consider Langevin dynamics as a special case, and then give an alternative derivation using a change-of-measure argument that does not rely on any stationarity or reversibility assumption.
+This procedure moreover easily enables us to calculate the second order correction to the linear response formula (which may or may not be useful in practice).
 Furthermore, we outline how the novel nonequilibirum linear response formula can be used to compute optimal controls of molecular systems for cases in which one wants to steer the system to maximize a certain target expectation value.
-
+We illustrate our approach with simple numerical examples.
 \end{abstract}
@@
 In some sense it has become one of the cornerstones of modern statistical physics since it can be related to the fluctuation dissipation theorem (FDT) which roughly states that for appropriate systems in statistical equilibrium, the average response to small external perturbations can be calculated through the knowledge of suitable correlation functions of the unperturbed statistical system.
-In this sense there the standard linear response theory holds in a  very general sense \cite{Majda} as long as one assumes that the unperturbed system is in equilibrium.
+The standard linear response theory holds in a  very general sense \cite{hairer2008} as long as one assumes that the unperturbed system is in equilibrium.
 %=Martin Hairer and Andrew J Majda, A simple framework to justify linear response theory, Nonlinearity 23 (2010) 909–922 doi:10.1088/0951-7715/23/4/008
@@
 Here $v_{t}\in\R^{d}$ is any given driving force applied to the system, which in general may depend on $X_{t}$ and $t$; $\eps>0$ is a small parameter.
-We consider the equilibrium and nonequilibrium case separately.
+Whenever (\ref{sde}) without the perturbation $\eps v$ is considered, we write
+\begin{equation}\label{sdewo}
+dx_{t} = b(x_{t},t)dt +a(x_{t})dB_{t}\,,\quad t\ge 0\,. 
+\end{equation}
 
@@
 
-For the equilibrium case (see, e.g., \cite{XXX,YYY,ZZZ}), we assume that the infinitesimal generator
+We consider the equilibrium and nonequilibrium case separately and start with the equilibrium case (see, e.g., \cite{kubo1957,tuckeman2010statistical}).
+To begin with, we assume that the infinitesimal generator
 \[
@@
 where $H\colon\R^{d}\to\R$, $d=2n$ is the Hamiltonian of the system, $\gamma=\gamma^{T}>0$ the positive semi-definite $n\times n$ friction matrix, $\sigma$ constant, satisfying $2\gamma=\beta\sigma\sigma^{T}$ for some $\beta>0$, and we have used the convention $x=(q,p)\in\R^{n}\times\R^{n}$.
-Under these conditions on friction and noise and some mild growth conditions on the Hamiltonian for large arguments, the unperturbed dynamics has a unique invariant measure with density
+Then, under some mild growth conditions on the Hamiltonian for large arguments, the unperturbed dynamics has a unique invariant measure with density
 \[
@@
 and the above spectral gap condition is met.
-Now let $f$ be any integrable phase space function and let $\rho^{\eps}=\rho^{\eps}(q,p,t)$ denote the probability density of $X_t=(Q_{t},P_{t})$, the solution to the Langevin equation (\ref{sde})--(\ref{eqn:langevin-1}), assuming that $(Q_{0},P_{0})=(q,p)$ was distributed according to the invariant density $\rho_{0}$.
+Now let $f$ be any integrable phase space function and let $\rho^{\eps}=\rho^{\eps}(q,p,t)$ denote the probability density of $X_t=(Q_{t},P_{t})$, the solution to the Langevin equation (\ref{sde})\&(\ref{eqn:langevin-1}), assuming that $(Q_{0},P_{0})=(q,p)$ was distributed according to the invariant density $\rho_{0}$.
 We define the expectation \wrt $\rho^{\eps}$ as
@@
 \] 
-A classical result (see, e.g., \cite{HairerMaijda08} and the references therein), that is usually derived using a formal expansion of the solution to the Kolmogorov forward or backward equations in powers of $\eps$, now states that
+A classical result, that is usually derived using a formal expansion of the solution to the Kolmogorov forward or backward equations in powers of $\eps$, now states that (see, e.g., \cite{hairer2008} and the references therein)
 \begin{equation}\label{lr}
@@
 \subsubsection*{Green-Kubo relations} 
-For the specific case of the Langevin equation (\ref{sde})--(\ref{eqn:langevin-1}), the general linear response (\ref{lr}) can be recast in form of the better known Green-Kubo formula \cite{MarconiPhysRep,Risken}.
+For the specific case of the Langevin equation (\ref{sde})\&(\ref{eqn:langevin-1}), the general linear response (\ref{lr}) can be recast in form of the better known Green-Kubo formula \cite{marconi2008,risken1996}.
 Using the formal operator identity
@@
 \begin{equation}\label{GreenKubo}
-\lim_{\eps\to 0}\frac{\bE_{\rho^{\eps}}[f] - \bE_{\rho^{0}}[f]}{\eps} = -\beta\int_{0}^{\infty}\bE_{\rho^{0}}[J(Q_{0},P_{0}) (f(Q_{t},P_{t})-\bE_{\rho^{0}}[f])]\,,dt
+\lim_{\eps\to 0}\frac{\bE_{\rho^{\eps}}[f] - \bE_{\rho^{0}}[f]}{\eps} = -\beta\int_{0}^{\infty}\bE_{\rho^{0}}[J(q_{0},p_{0}) (f(q_{t},p_{t})-\bE_{\rho^{0}}[f])]\,,dt
 \end{equation}
-in terms of the dissipative flux $J=-v\cdot\nabla H$, where the expectation under the integral is taken over all realizations/path of the unperturbed Langevin equation with initial distribution $\rho^{0}(q,p,0)=\rho_{0}$.
-In other words:
+in terms of the dissipative flux $J=-v\cdot\nabla H$. In other words: 
 \[
-\bE_{\rho^{\eps}}[f] \approx \bE_{\rho_{0}}[f] - \eps\beta\int_{0}^{\infty}\bE_{\rho_{0}}[J(Q_{0},P_{0}) (f(Q_{t},P_{t})-\bE_{\rho_{0}}[f])]\,dt.
+\bE_{\rho^{\eps}}[f] \approx \bE[f] - \eps\beta\int_{0}^{\infty}\bE[J(q_{0},p_{0}) (f(q_{t},p_{t})-\bE[f])]\,dt\,,
+\]
+ where the expectation under the integral is taken over all realizations of the unperturbed equilibrium Langevin equation (\ref{sdewo})\&(\ref{eqn:langevin-1}) with initial distribution 
+ \[
+ \rho^{0}(q,p,0)=\rho_{0}\,.
  \]
@@
 
-The classical response theory has the limitation that the reference (equilibrium) distribution $\rho_{0}$ must be stationary under the dynamics and is unique.
+The classical response theory has the limitation that the reference (equilibrium) distribution $\rho_{0}$ must be the unique stationary probability measure of the unperturbed dynamics.
 Moreover the validity of the linear approximation relies on a spectral gap condition that is difficult to verify in practice.
-In particular the perturbation argument does not provide a framework, under which the second and even higher order responses can be easily derived.
+In particular the perturbation argument does not provide a framework, under which the second and even higher order responses can easily be derived.
 
@@
 Here we propose an alternative (purely formal) derivation of the linear response result, that is based on a change of drift in the corresponding SDE and which allows for an easy generalization of the above linear response result to nonequilibrium systems.
-We will briefly review the idea of the change of drift via Girsanov transformations; for details we refer to the textbook~\cite{oksendal2003stochastic}.
-Now let $x_{t},X_{t}$ be the solutions to the following stochastic differential equations:
+We will briefly review the idea of the change of drift via Girsanov transformations; for details we refer to the textbook \cite{oksendal2003stochastic}.
+Let $x_{t},X_{t}$ be the solutions to the following stochastic differential equations:
 \begin{subequations}\label{sde2}
@@
 \[
-x_{0} = X_{0} = x\,,
+x_{0} = X_{0} = x\,.
 \]
@@
 \begin{equation}\label{xi}
-a(X_{t})\xi_{t} = v_{t}\,,
+a(X_{t})\xi_{t} = v_{t}\,.
 \end{equation}
@@
 \end{equation}
-It follows from the Girsanov theorem \cite[Thm.~8.6.8]{oksendal2003}, sometimes also called \emph{Cameron-Martin-Girsanov theorem} \cite{a}, that $W_{t}$ is again a Brownian motion under a new probability measure that has a density \wrt the Gaussian probability measure that is generated by the Brownian motion $B_{t}$.
-Specifically, let $\nu$ denote the law of the Brownian motion $B_{t}$ and define a new probability measure $\mu$ on the space of continuous trajectories by
+It follows from the Girsanov theorem \cite[Thm.~8.6.8]{oksendal2003stochastic}, sometimes also called \emph{Cameron-Martin-Girsanov theorem} \cite{stroock2006}, that $W_{t}$ is again a Brownian motion under a new probability measure that has a density \wrt the Gaussian probability measure that is generated by the Brownian motion $B_{t}$.\footnote{A quick-and-dirty derivation of the above change-of-measure formula can be easily obtained, if the noise covariance $a(\cdot)a(\cdot)^{T}$ has full rank with bounded inverse.
+Then, using Euler's method for (\ref{sde2}), it follows that (\ref{likelihood}) is basically the likelihood ratio between the time-discrete path densities of (\ref{sde2-2}) and (\ref{sde2-1}).
+} Specifically, let $\nu$ denote the law of the Brownian motion $B_{t}$ and define a new probability measure $\mu$ on the space of continuous trajectories by
 \[
@@
 \[
-M_{T}^{-1} = \exp\left(\eps\int_{0}^{T}\xi_{s}\cdot dW_{s} - \frac{\eps^{2}}{2}\int_{0}^{T}|\xi_{s}|^{2}ds\right)\,.
+M_{T}^{-1} = \exp\left(\eps\int_{0}^{T}\xi_{s}\cdot dW_{s} - \frac{\eps^{2}}{2}\int_{0}^{T}|\xi_{s}|^{2}ds\right)
 \]
 is the density of $\nu$ relative to $\mu$.
+Note that the expectation on the right hand side corresponds to the unperturbed dynamics, because $W_{t}$ is a standard Brownian motion under $\mu$, and the expectation is over all realizations of (\ref{sde2-3}) starting from either any given initial condition $X_{0}=x$ or an arbitrary initial distribution.
+On the other hand, $X_{t}$ under $\nu$ corresponds to the perturbed dynamics, which should become clear upon comparing equations (\ref{sde2-2}) and (\ref{sde2-3})
+
+\subsubsection*{An alternative linear response formula}
 
@@
 \begin{equation}\label{lr-alt}
-\lim_{\eps\to 0}\frac{\bE_{\nu}[f] + \bE_{\mu}[f]}{\eps} = \bE_{\mu}\left[f(\{X_{t}\}_{0\le t\le T})\int_{0}^{T}\xi_{s}\cdot dW_{s} \right]\,,
+\lim_{\eps\to 0}\frac{\bE_{\rho^{\eps}}[f] + \bE_{\rho^{0}}[f]}{\eps} = \bE_{\rho^{0}}\left[f(\{x_{t}\}_{0\le t\le T})\int_{0}^{T}\xi_{s}\cdot dB_{s} \right]\,,
 \end{equation}
-where $X_{t}$ is the solution (\ref{sde2-3}).
-Note that this is the unperturbed nonequilibrium dynamics, because $W_{t}$ is a standard Brownian motion under $\mu$, and the expectation is over all realizations of (\ref{sde2-3}) starting from either any given initial condition $X_{0}=x$ or an arbitrary initial distribution.
-On the other hand, $X_{t}$ under $\nu$ corresponds to the perturbed dynamics, which should become clear upon comparing equations (\ref{sde2-2}) and (\ref{sde2-3}).
-Further note that $f$ and $W_{t}$ are not independent, hence the expectation on the right hand side of the equation is not zero in general.
+where $x_{t}$ is the solution (\ref{sde2-1}) and $\rho^{\eps}$, $\rho^{0}$ denote the distributions of the perturbed and unperturbed dynamics, in accordance with the notation used in the previous section.
+Note that $\rho^{0}\neq\rho_{0}$ does not need to be an equilibrium distribution.
+Further note that $f$ and $B_{t}$ are not independent, hence $f$ times the integral over the Brownian motion does not average to zero in general.
 
-\paragraph*{Remark.} A quick-and-dirty derivation of the above change-of-measure formula can be easily obtained, if the noise covariance $a(\cdot)a(\cdot)^{T}$ has full rank almost everywhere with bounded inverse.
-Then, using Euler's method for (\ref{sde2}), it follows that (\ref{likelihood}) is basically the likelihood ratio between the time-discrete path densities of (\ref{sde2-2}) and (\ref{sde2-1}).
-
-\paragraph*{Remark.} By formally expanding $M^{-1}_T$ up to second order we get an analogous second order response formula: 
+\paragraph*{Remark.} 
+ By formally expanding $M^{-1}_T$ up to second order we get an analogous second order response formula: 
 \begin{equation}\label{2nd}
-\bE_{\nu}[f] \approx \bE_{\mu}[f] + \eps \bE_{\mu}\left[f\int_{0}^{T}\xi_{s}\cdot dW_{s} \right] +\frac{\eps^{2}}{2} \bE_{\mu}\left[f\left\{\left(\int_{0}^{T}\xi_{s}\cdot dW_{s}\right)^2 -\int_0^T |\xi_s|^2\,ds\right\} \right]
+\begin{aligned}
+\bE_{\rho^{\eps}}[f] & \approx \bE_{\rho^{0}}[f]  + \eps \bE_{\rho^{0}}\left[f\int_{0}^{T}\xi_{s}\cdot dB_{s} \right]\\  
+& + \frac{\eps^{2}}{2} \bE_{\rho^{0}}\left[f\left\{\left(\int_{0}^{T}\xi_{s}\cdot dW_{s}\right)^2 -\int_0^T |\xi_s|^2\,ds\right\} \right].
+\end{aligned}
 \end{equation}
@@
 \begin{aligned}
-dQ_t & =  \nabla_p H(Q_t,P_t)\\
-dP_t & =  -\nabla_q H(Q_t,P_t) - \gamma \nabla_p H(Q_t,P_t) + D(Q_t)\,(u_t+\eps \delta u_t)+\sigma dB_t,
+dQ_t & =  \nabla_p H(Q_t,P_t)dt\\
+dP_t & =  -\nabla_q H(Q_t,P_t)dt - \gamma \nabla_p H(Q_t,P_t)dt + D(Q_t)\,(u_t+\eps \delta u_t)dt+\sigma dB_t,
 \end{aligned}
@@
 \begin{equation}\label{lr-alt2}
-\lim_{\eps\to 0}\frac{\bE_{\nu}[f] - \bE_{\mu}[f]}{\eps} =  \bE_{\mu}\left[f(\{(Q_{t},P_{t})_{0\le t\le T}\})\int_{0}^{T}(\sigma^{-1}D(Q_{s})\,\delta u_{s})\cdot dW_{s} \right]\,
+\lim_{\eps\to 0}\frac{\bE_{\rho^{\eps}}[f] - \bE_{\rho^{0}}[f]}{\eps} =  \bE_{\rho^{\eps}}\left[f(\{(q_{t},p_{t})_{0\le t\le T}\})\int_{0}^{T}(\sigma^{-1}D(q_{s})\,\delta u_{s})\cdot dB_{s} \right]\,
 \end{equation}
@@
 \begin{equation}\label{eqn:neq-response}
-\bE_{\nu}[f] \approx \bE_{\mu}[f] + \eps\bE_{\mu}\left[f(\{Q_{t},P_{t}\}_{0\le t\le T})\int_{0}^{T}(\sigma^{-1}D(Q_{s})\,\delta u_{s})\cdot dW_{s} \right]\,.
+\bE_{\rho^{\eps}}[f] \approx \bE[f] + \eps\bE\left[f(\{q_{t},p_{t}\}_{0\le t\le T})\int_{0}^{T}(\sigma^{-1}D(q_{s})\,\delta u_{s})\cdot dB_{s} \right]\,.
 \end{equation}
@@
 
-\paragraph*{Remark.} Some comments on the above result are in order:
+\subsubsection*{Numerical realization}
+
+As one can easily get lost in the various integral transformations, measures $\nu$ and $\mu$, distributions $\rho^{\eps}$ and $\rho^{\eps}$ etc.
+it may be helpful to understand how the linear response formulas (\ref{lr-alt}) or (\ref{lr-alt2}) can be realized algorithmically.
+To this end, let $(x_0,\,x_{1},\,x_{2},\,x_{3},\ldots)$ with $x_{k}$ be the numerical realization of (\ref{sde2-1}).
+Let us further suppose that the initial value $x_{0}=x$ is fixed.
+The simplest possible numerical discretization would be the Euler scheme
+\[
+x_{n+1} = x_{n} + \Delta t\, b(x_{n},t_{n}) + \sqrt{\Delta t}\,a(x_{n})\eta_{n+1}\,,\quad x_{0}=x\,, 
+\]
+with time step $\Delta t=t_{k+1}-t_{k}$ and $\eta_{k}$ i.i.d.~Gaussian random variables with mean 0 and unit covariance.
+(For a Langevin equation such as (\ref{eqn:langevin-1d}) the Euler scheme is not recommended, but the basic idea stays the same.) Now a simple unbiased estimator of the linear response---i.e.
+the right hand side in (\ref{lr-alt})---would be
+\begin{equation}\label{lr-est}
+\hat{R} = \frac{1}{M}\sum_{i=1}^{M}\left\{f(\{x_{k}(\omega_{i})\}_{0\le k\le N})\sum_{j=0}^{N-1}\hat{\xi}_{i}(\omega_{i})\cdot \eta_{j+1}(\omega_{i})\right\}
+\end{equation}
+with $N=\lfloor T/\Delta t\rfloor$ and $x_{k}(\omega_{i})$ denoting the $i$-th realization of $x_{k}$, that is generated by the $i$-th realization $(\eta_{1}(\omega_{i}),\ldots,\eta_{k}(\omega_{i}))$ of the Gaussian noise sequence $(\eta_{k})_{k\in\N}$.
+The time-discrete control variable is given by
+\[
+a(x_{k})\hat{\xi}_{k} = v_{t_{k}}
+\]
+for any given perturbation $v\colon[0,T]\to\R^{m}$.
+
+\paragraph*{Remarks.} Some comments on the above result are in order:
 \begin{itemize}
@@
   have the same initial distribution.
-For fixed initial values (i.e.~points) this assumption cannot be relaxed (because otherwise $dP/dQ$ does not exist).
+For fixed initial values (i.e.~points) this assumption cannot be relaxed (because otherwise $d\nu/d\mu$ does not exist).
 For distributed initial values, however, there is no problem for the unperturbed and perturbed dynamics to have different initial distribution as long as both distributions are strictly positive almost everywhere.
@@
 
-Let us remain in the setting of the last section and assume that we are interested in a finite dimensional family of forcings
+The nonequilibrium linear response formula can be used for solving certain optimal control problems.
+To this end, let us remain in the setting of equations (\ref{eqn:langevin-1d})--(\ref{eqn:neq-response}) and assume that we are interested choosing the nonequilibrium forcing $u$, such that the expected value $\bE[f]$ of some utility function
 \[
-\mathcal{F}=\{F_e(t)=\sum_{k=1}^N a_k \Phi_k(t)\},
+f = \int_{0}^{T}\left\{\ell (q_{s}) - \frac{1}{2}|u_{s}|^{2}\right\}\,dt + L(q_{T}) 
 \]
-with explicitly given basis functions $\Phi_k$, $k=1,\ldots,N$.
-We are interested in computing the gradient of the expectation value at time $T$ of the observable $\mathcal{O}$ wrt the coefficient vector $(a_k)_{k=1,\ldots,N}$ in the point $F_e(t)=\sum_{k=1}^N a_k \Phi_k(t)$.
-From Eq.
-(\ref{eqn:neq-response}) we get with $\Delta F_e(t)=\Phi_k(t)$:
+is maximized where $q_{t}$ is the solution to the controlled Langevin equation (\ref{eqn:langevin-1d}) for $\eps=0$; the functions $\ell$ and $L$ are the running cost and the terminal cost, which are assumed to be continuous and bounded from above; without loss of generality, $\ell$ and $L$ are assumed to depend only on the positions.
+The quadratic term is a penalization that makes sure that the controls do not go through the roof \cite{schuette2012,hartmann2012}  Let us moreover assume that the controls are open-loop (i.e.
+without feedback) and can be represented by
 \[
-\frac{\partial O(T)}{\partial a_j} = \lim_{\eps\to 0}\frac{1}{\eps}( O_\eps(T) -
-  O_0(T))=\big\langle O[\vect X_s]\cdot G_j[\vect X_s]\big\rangle_{0,T},
+u_{t} = \sum_{k=1}^K a_k \Phi_k(t)\,,\quad a_{k}\in \R\,,
 \]
-where $X_t$ and $O_0$ are paths and expectation value wrt. the dynamics \eqref{eqn:langevin-1b}--\eqref{eqn:langevin-2b} with $F_e(t)=\sum_{k=1}^N a_k \Phi_k(t)$ and $\eps=0$, and
+with suitably chosen time dependent, bounded and Lipschitz continuous vector fields $\Phi_k\colon[0,T]\to\R^{n}$, $k=1,\ldots,K$.
+%In principle we could even go for closed-loop (i.e.~feedback) controls by making $\Phi_{k}$ state dependent.
+We want to solve the optimal control problem
+\begin{equation}\label{ocp1}
+\begin{aligned}
+ \max_{u\in\cU} \,& I(u)\quad \textrm{s.t.}\\
+dq_t & =  \nabla_p H(q_t,p_t)dt\\
+dp_t & =  -\nabla_q H(q_t,p_t)dt - \gamma \nabla_p H(q_t,p_t)dt + D(q_t)u_{t} dt+\sigma dB_t\,.
+\end{aligned}
+\end{equation}
+where 
+\begin{equation}\label{ocp2}
+I(u) = \bE\left[\int_{0}^{T}\left\{\ell (q_{s}) - \frac{1}{2}|u_{s}|^{2}\right\}\,dt + L(q_{T}) \right]
+\end{equation}
+and $\cU$, the space of admissible controls, consists of all bounded controls 
+\begin{equation}\label{admissible}
+u_{t} = \sum_{k=1}^K a_k \Phi_k(t)\,,\quad a_{k}\in \R\,,
+\end{equation}
+
+\subsubsection*{Gradient method from linear response}
+
+In principle optimal control problems such as (\ref{ocp1})--(\ref{ocp2}) can be solved by dynamic programming, i,e., by solving the corresponding Hamilton-Jacobi-Bellman PDE \cite{fleming2006}.
+Expect for very simple, essentially one-dimensional systems, solving Hamilton-Jacobi-Bellman equations is not an easy task, so we pursue a different strategy here.
+The idea is to use that we have restricted our space of admissible controls to functions of the form (\ref{admissible}) with given basis vector fields and that we can a gradient search in the unknown coefficients $a_{k}$, using the iteration
 \[
-G_j[\vect X_t]
-  = \frac{1}\sigma\int_0^t
-  \Phi_j(s)\,\vect D(\vect q_s)\cdot\dd\vect W_s. 
+u^{(n+1)} = u^{(n)} + \tau_{n}\nabla I(u^{(n)})\,,
 \]
-That is, we can compute all derivatives $\partial O(T)/\partial a_j$ from just one ensemble of paths of $X_t$.
-
-Now we consider the optimization problem
+with $\tau_{n}>0$ being an adjustable parameter that determines the length of each gradient step.
+The gradient of $I$ can be easily evaluated using the linear response formula.
+To see this recall the notion of functional (G\^ateaux) derivatives as directional derivatives along a function $v$ (from a suitable function space):
+\begin{equation*}
+\frac{\delta I}{\delta u} = \left.\frac{d}{d\eps}\right|_{\eps=0} I(u+\eps v) = \langle\nabla I(u),v\rangle\,,
+\end{equation*} 
+Now the idea is to use that we have restricted our space of admissible controls to functions of the form (\ref{admissible}) with given basis vector fields and do a gradient search in the unknown coefficients $a_{k}$.
+This requires to compute the gradient \wrt the coefficients.
+Let the vector
 \[
-O_{max}(T) = \max_{F_e\in \mathcal{F}} O[F_e](T),
+\delta u = \sum_{k=1}^K \delta a_k \Phi_k(t)
 \]   
-where $O[F](T)$ is the expectation value of the path observable $\mathcal{O}$ that results from \eqref{eqn:langevin-1b}--\eqref{eqn:langevin-2b} with $F_e=F$ and $\eps=0$.
-With the gradient being computed one can solve the optimization problem by means of gradient descent in coefficient space,
+denote the direction along which we want to differentiate, where $\delta a=(\delta_{1},\ldots,\delta a_{K})$ are coefficients of the vector $\delta u$ in the basis of the $\Phi_{k}$, and note that 
+\begin{equation*}
+\begin{aligned}
+\frac{\delta I}{\delta u} & = \left.\frac{d}{d\eps}\right|_{\eps=0} I(u+\eps \delta u)\\
+& = \left(\lim_{\eps\to 0}\frac{I(u+\eps\delta u) - I(u)}{\eps}\right)\cdot \delta u\\
+& = \left(\lim_{\eps\to 0}\frac{\bE_{\rho^{\eps}}[f] - \bE_{\rho^{0}}[f]}{\eps}\right)\cdot\delta u\\
+\end{aligned}
+\end{equation*} 
+provided that the limit exists. The above iteration therefore is equivalent to
 \[
-a^{(i+1)}=a^{(i)} -\gamma \frac{\partial O(T)}{\partial a}.
+a_{k}^{(n+1)} = a_{k}^{(n)} + \tau_{n}\left.\frac{\partial I}{\partial a_{k}}\right|_{a_{k}=a_{k}^{(n)}}\,,
 \]
+with
+\[
+\frac{\partial I}{\partial a_{k}} = \bE\left[f\int_{0}^{T}(\sigma^{-1}D(q_{s})\Phi_{k}(s))\cdot dB_{s} \right]\,.
+\]
+and $\bE[\cdot]$ being the expectation over all realization of (\ref{eqn:langevin-1d}) with $\eps=0$ and $u=u^{(n)}$ being the current iterate; note that we can compute all partial derivatives $\partial I/\partial a_k$, $k=1,\ldots,K$ from just one ensemble of nonequilibrium paths of $(q_{t},p_{t})$.
+Further note that, in general, the updated coefficient $a^{(n+1)}$ and hence the updated control $u^{(n+1)}$ will depend on the distribution of the initial conditions at the $n$-th iteration stage; in particular, if the controls are computed on-the-fly, the controls will actually be feedback controls depending on the current state.
 
@@
 
-\bibliography{ref}{}
+\bibliography{ref}
 \bibliographystyle{tMPH}

### 2013-06-29 writedown/linear.response/paper.tex
update

@@
 
-% \title{Linear response theory for core set identification}
-\title{Linear response theory for a system under non-equilibrium conditions}
+\title{Linear response theory and optimal control for a molecular system under non-equilibrium conditions}
 
@@
 In some sense it has become one of the cornerstones of modern statistical physics since it can be related to the fluctuation dissipation theorem (FDT) which roughly states that for appropriate systems in statistical equilibrium, the average response to small external perturbations can be calculated through the knowledge of suitable correlation functions of the unperturbed statistical system.
-The standard linear response theory holds in a  very general sense \cite{hairer2008} as long as one assumes that the unperturbed system is in equilibrium.
+The standard linear response theory holds in a  very general sense \cite{HairerMajda} as long as one assumes that the unperturbed system is in equilibrium.
 %=Martin Hairer and Andrew J Majda, A simple framework to justify linear response theory, Nonlinearity 23 (2010) 909–922 doi:10.1088/0951-7715/23/4/008
@@
 There is an increasing number of articles in the literature that report on applications of molecular dynamics to nonequilibrium settings.
-There are many generalization to so-called nonequilibrium steady states based on the generality of the FDT, but despite its wide use, the present authors do not know of a linear response formula that applies to fully general nonequilibirium cases.
+There are many generalization to so-called nonequilibrium steady states based on the generality of the FDT \cite{Seifert}, but despite its wide use, the present authors do not know of a linear response formula that applies to fully general nonequilibirium cases.
 We will provide such a formula for the case that the underlying dynamics can be described by Langevin dynamics.
@@
 \begin{equation}\label{lr-alt2}
-\lim_{\eps\to 0}\frac{\bE_{\rho^{\eps}}[f] - \bE_{\rho^{0}}[f]}{\eps} =  \bE_{\rho^{\eps}}\left[f(\{(q_{t},p_{t})_{0\le t\le T}\})\int_{0}^{T}(\sigma^{-1}D(q_{s})\,\delta u_{s})\cdot dB_{s} \right]\,
+\lim_{\eps\to 0}\frac{\bE_{\rho^{\eps}}[f] - \bE_{\rho^{0}}[f]}{\eps} =  \bE_{\rho^{0}}\left[f(\{(q_{t},p_{t})_{0\le t\le T}\})\int_{0}^{T}(\sigma^{-1}D(q_{s})\,\delta u_{s})\cdot dB_{s} \right]\,
 \end{equation}
@@
 \begin{equation}\label{eqn:neq-response}
-\bE_{\rho^{\eps}}[f] \approx \bE[f] + \eps\bE\left[f(\{q_{t},p_{t}\}_{0\le t\le T})\int_{0}^{T}(\sigma^{-1}D(q_{s})\,\delta u_{s})\cdot dB_{s} \right]\,.
+\bE_{\rho^{\eps}}[f] \approx \bE_{\rho^0}[f] + \eps\bE_{\rho^0}\left[f(\{q_{t},p_{t}\}_{0\le t\le T})\int_{0}^{T}(\sigma^{-1}D(q_{s})\,\delta u_{s})\cdot dB_{s} \right]\,.
 \end{equation}
@@
 \begin{equation}\label{ocp2}
-I(u) = \bE\left[\int_{0}^{T}\left\{\ell (q_{s}) - \frac{1}{2}|u_{s}|^{2}\right\}\,dt + L(q_{T}) \right]
+I(u) = \bE_{\rho^0}\left[\int_{0}^{T}\left\{\ell (q_{s}) - \frac{1}{2}|u_{s}|^{2}\right\}\,dt + L(q_{T}) \right]
 \end{equation}
@@
 provided that the limit exists. The above iteration therefore is equivalent to
-\[
+\begin{equation}\label{gd}
 a_{k}^{(n+1)} = a_{k}^{(n)} + \tau_{n}\left.\frac{\partial I}{\partial a_{k}}\right|_{a_{k}=a_{k}^{(n)}}\,,
-\]
+\end{equation}
 with
-\[
-\frac{\partial I}{\partial a_{k}} = \bE\left[f\int_{0}^{T}(\sigma^{-1}D(q_{s})\Phi_{k}(s))\cdot dB_{s} \right]\,.
-\]
-and $\bE[\cdot]$ being the expectation over all realization of (\ref{eqn:langevin-1d}) with $\eps=0$ and $u=u^{(n)}$ being the current iterate; note that we can compute all partial derivatives $\partial I/\partial a_k$, $k=1,\ldots,K$ from just one ensemble of nonequilibrium paths of $(q_{t},p_{t})$.
+\begin{equation}\label{dida}
+\frac{\partial I}{\partial a_{k}} = \bE_{\rho^0}\left[f\int_{0}^{T}(\sigma^{-1}D(q_{s})\Phi_{k}(s))\cdot dB_{s} \right]\,.
+\end{equation}
+and $\bE_{\rho^0}[\cdot]$ being the expectation over all realization of (\ref{eqn:langevin-1d}) with $\eps=0$ and $u=u^{(n)}$ being the current iterate; note that we can compute all partial derivatives $\partial I/\partial a_k$, $k=1,\ldots,K$ from just one ensemble of nonequilibrium paths of $(q_{t},p_{t})$.
 Further note that, in general, the updated coefficient $a^{(n+1)}$ and hence the updated control $u^{(n+1)}$ will depend on the distribution of the initial conditions at the $n$-th iteration stage; in particular, if the controls are computed on-the-fly, the controls will actually be feedback controls depending on the current state.
@@
 
-\section{Numerical tests}
+\section{Numerical Experiments}
 
-\subsection{The metastability of a single-well potential}
+\subsection{Splitting a single-well potential}
 
@@
   \includegraphics[width=0.4\textwidth]{figs/fig-split-pot.eps}
-  \caption{The single-well potential with splitting splitting drving
-    force at time $T = 20$ ps.  Since the drving force is
-    conservative ($ D( q) = -\nabla_{q}V( q)$),
-    we plot the non-equilibrium drving energy of the system.  The red,
+  \caption{The single-well potential with  splitting driving
+    force. Time $T = 20$ ps.  Since the driving force is
+    of gradient form, $ D( q) = -\nabla_{q}V( q)$,
+    we plot the non-equilibrium driving energy of the system.  The red,
     green and blue lines are the potential energy without
@@
 We use the idea of non-equilibrium linear response to investigate the
-nonequilibrium phase space probability density distribution, denoted by $f(q, p, t)$, of a one-dimensional model system: one particle in a
-splitting single-well potential.  For convenience, we let the mass of
+nonequilibrium phase space probability density distribution, denoted by $\rho^\eps(q, p, t)$, of a one-dimensional model system: one particle in a
+splitting single-well potential as shown in Fig.~\ref{fig:tmp4}.  For convenience, we let the mass of
 the particle to be 1~\textsf{amu}, and the friction coefficient to be
@@
 %$\textsf{kJ} / (\textsf{mol nm}^4)}$
-The non-equilibrium driving $ D$ is given by a conservative force:
+The non-equilibrium driving $ D$ is given by a force of gradient form:
 \begin{align}
-   D( q) = -\nabla_{ q} V( q) 
+   D( q) = -\nabla_{ q} V( q) ,
 \end{align}
@@
 We
-consider the perturbation to the system given by $\eps\delta u(t) = \eps \cdot t/T$.
+consider the perturbation to the system given by $\eps\delta u(t) = \eps k_e \cdot t/T$.
 We consider the following parameters: end time $T = 20$~\textsf{ps}, 
@@
 for the non-equilibrium driving potential and perturbed
-potential at time $t = T$. The initial distribution $f(q, p, 0)$ is set to
+potential at time $t = T$. The initial distribution $\rho^0(q, p, 0)$ is set to
 be equilibrium distribution of the unperturbed system.
@@
   {figs/fig-split-str2-2d-distrib-no2nd.eps}
-  \caption{ The plot of $ f(q,p,t)$ in the phase
-    space under perturbed non-equilibrium driving. From left to right the columns present results at time $t =
-    0$, 5, 10 and 20~\textsf{ps}.  The 1st row: the brute force
-    non-equilibrium simulation. The 2nd row: the results of
-    traditional equilibrium linear response theory. In this case we
-    effectivly consider an equilibrium system free of driving, and the perturbation $\eps$ is set to 2.  The 3rd row: the
-    non-equilibrium linear response result.  }
+  \caption{ The plot of $ \rho^\eps(q,p,t)$ in  phase
+    space under the perturbed non-equilibrium driving described in the text. From left to right the columns present results at times $t =
+    0$, 5, 10 and 20~\textsf{ps}.  First row: Results of a brute force
+    non-equilibrium simulation. Second row: Results of
+    classical equilibrium linear response theory. In this case we
+    effectively consider an equilibrium system free of driving, and the perturbation $\eps$ is set to 2.  Third row: Results using the
+    nonequilibrium linear response result.  }
   \label{fig:tmp6}
@@
 
-Fig.~\ref{fig:tmp6} presents the numerical results of the phase space probability distribution for the splitting
-single-well potential. From left to right the four rows presents the
+Fig.~\ref{fig:tmp6} presents the numerical results for  the phase space probability distribution for the splitting
+single-well potential. From left to right the four columns present the
 distribution of the system at time $t = 0$, 5, 10 and 20~\textsf{ps}. The first
-row presents the brute force non-equilibrium simulation.  It is clear
-that at the beginning, the distribution has only one peak,
-i.e. arount $q = 0$ and $p = 0$. As the time evolves, an energy barrier develops
-in the center of the simulation region, therefore, the single peak
+row presents the result of a brute force non-equilibrium simulation.  It is clear
+that at the beginning the distribution has only one peak around $q = 0$ and $p = 0$. As  time evolves, an energy barrier develops
+in the center of the simulation region and, therefore, the single peak
 splits into two equally sized peaks.  In the end, the two
-peaks are entirly seperated.  This brute force non-equilibrium
-simulation serves as the precise result that the response theory
-should compare with. The second row shows the result of the
+peaks are entirely seperated.  The brute force non-equilibrium
+simulation serves as the precise result to which the response theory
+should be compared. The second row shows the result of the
 traditional equilibrium linear response theory.  Please notice that in
-this case, since the reference simulation is in equilibrium, so 
-the effective perturbation is of strength $\eps = 2$.  At $t \leq
+this case, since the reference simulation is in equilibrium, we set the perturbation to
+\[
+\eps v_t = D(q) (u_t+\eps\delta u_t)=2\eps D(q)k_e\cdot t/T=2\eps D(q)\delta u_t,
+\] 
+so that the effective perturbation is of strength $2\eps = 2$.  At $t \leq
 15$~\textsf{ps}, the accuracy of the equilibrium linear response is
@@
 20$~\textsf{ps}, magnitude of the peaks are relatively too strong,
-and the gap between them is actually negative.
+and in the gap between them the distribution is actually negative.
 Since the probability distribution is always positive, the
@@
 The poor accuracy is due to the fact that the strength of perturbation is no
-longer small so that the preliminary assumption of the reponse theory
-is not well fulfilled.  On the third row of Fig.~\ref{fig:tmp6},
-we present the linear response starting from a non-equilibrium driving
-of $u(t)$.  The numerical results is satisfactorily consistent with
-the brute force non-equilibrium simulation, because the perturbation
-($\eps = 1$) is so small that the linear perturbation theory achieves
+longer small so that the preliminary assumption of the classical reponse theory ("small perturbation")
+is not satisfied.
+
+The third row of Fig.~\ref{fig:tmp6} presents the results computed using the novel nonequilibrium  linear response formula: we first start form the equilibrium distribution, apply  $u(t)$, arrive at a nonequilibrium distribution and then, in a second step, compute the effect of the non-equilibrium driving $\delta u(t)$.
+The numerical results are satisfactorily consistent with
+the brute force non-equilibrium simulation, because the perturbation is still  small enough and the novel linear response theory achieves
 good accuracy.
@@
   \includegraphics[]{figs/fig-ctr-stat-2.eps}
-  \caption{The optimal forcing.  In this plot, $T=20$ ps.  The red
-    color indicates $\eta = 0.01$ and the green color indicates $\eta
-    = 0.05$.  In the upper plot, the solid lines present the magnitude
-    of the control $F_e(t)$, which is calculated on a family of
-    piecewise linear ansatz functions. The time interval of a linear
-    piece is 1 ps.  The blue insertions show the shape of the
-    non-equilibrium driving potential $U( q) + F_e(t)V(
-    q)$ at time indicated by the black arrow.  In the lower plot, the
-    dashed lines display the value of $P( X_t)$ as the time
-    evolves.  The solid lines present $O[F_e](t)$.  }\label{fig:tmp7}
+  \caption{Illustration of the optimal control for tilting the potential.  
+    In this plot, $T=9$ ps.  In the gradient descent search, the
+    red lines  calculate
+    the gradient by the non-equilibrium linear respect theory, while
+    the green lines calculates by the central finite difference scheme.
+ In the upper plot, the solid lines present the magnitude
+    of the optimal control $u_t$ as calculated based on a family of
+    piecewise linear ansatz functions with time interval 1~ps.
+    The blue insertions show the shape of the
+    nonequilibrium driving potential $U( q) + u_t V(
+    q)$ at the time indicated by the black arrow.  In the lower plot, the
+    dashed lines display the value of $P(q_t)$ as the time
+    evolves,  and the solid lines present $\eta I$.  }\label{fig:tmp7}
 \end{figure}
@@
 %$\textsf{kJ} / (\textsf{mol nm}^4)}$
-The perturbation is given by a conservative tilting to $U( q)$:
+The perturbation is given by a gradient form tilting of $U( q)$ by means of
 \begin{align}
-   D( q) = -\nabla_{ q} V( q) = 1
+   D( q) = -\nabla_{ q} V( q) = 1,
 \end{align}
-Here $V( q) = - q$. We want to
+with $V( q) = - q$. We want to
 optimally design the tilting such that the probability
 of being in the right well is as high as possible 
-at the end time of the process. That is equal to consider the 
+at the end time of the process under a constraint on the energy used for the control in the sense of
 the following optimal forcing problem:
 \begin{align}
-  O_{\max}(T)& = \max_{F_e\in\mathcal F} O[F_e](T) 
+  I & = \max_{u_t \in\mathcal F} \bE_{\rho^0}\left[ -\int_{0}^{T} \frac{1}{2}|u_{s}|^{2}\,dt + L(q_T) \right],
 \end{align}
-with
-\begin{align}
-  O[F_e](T) &= P( X_T) -
-  \eta\,
-  \big\langle \int_0^T\vert F_e(t) D( q)\vert^2\, dt\,
-  \big\rangle
-\end{align}
-where $P( X_T)$ is probability of the end point of the trajectory
-$ X_T$ being in the right well.  It is
-clear that for an unbiased double-well, $P( X_T)$
-is 0.5.  The integral is the ``cost'' of the control and $\eta$ is a
-positive constant that indicates the relative magnitude of the
-cost. We consider the ansatz function family $\mathcal F$ of piecewise
-linear functions on $[0, T]$.
+with $L(q_T)=\chi_{[a-\delta,a+\delta]}(q_T)/\eta$ representing the probability of the end point of the trajectory, $ q_T$, being in the right well ($\chi_I$ denotes the indicator function of the interval $I$) with $\eta$ being a weighting constant, ${\mathcal F}$ denoting the space of function that are piecewise linear  on $[0, T]$ in uniform intervals of length 1 ps, and $\rho^0$ being the non-equilibrium distribution for $u_t$.
+With $P(T) =  \bE_{\rho^0}(L(q_T))$, we thus have
+\[
+  \eta I = \max_{u_t \in\mathcal F} \Big[ P(T) -
+  \frac{\eta}{2}\,
+   \int_0^T |u_t|^2\, dt\Big],
+\]
+where $P(T)$ is probability of the end point of the trajectory
+$ q_T$ being in the right well.  It is
+clear that for an unbiased double-well, $P(T)$
+is 0.5.  The integral is the ``cost'' of the control and $\eta$  indicates the relative magnitude of the
+cost.
 
-Fig.~\ref{fig:tmp7} presents the numerical result. The magnitude of
-control $F_e(t)$ is shown in the upper plot, and the value of
-optimization target $O[F_e](T)$ and the value of $P( X_T)$ are
+Fig.~\ref{fig:tmp7} presents the numerical results of $\eta = 0.01$.
+Starting from an initial guess of linear-incremental control from $u_0 = 0$ to
+$u_T = 1$,
+the gradient search~\eqref{gd} converges at the 22nd step, when the
+maximum increment of the control coefficient $\max_k\vert \delta a_k\vert$
+is smaller than $0.02$.
+The magnitude of the optimal
+control $u_t$ is presented in the upper plot,
+with blue insertions showing the shape of the controlled potential 
+$U(q) + u_tV(q)$ (tilted double-well potential). 
+The maximized $\eta I$ and the corresponding probability in the right well
+$P( X_T)$ are
 given in the lower plot by the solid and dashed lines, respectively.
-The plot shows that with a smaller $\eta$ (the relative magnitude of
-the cost of the control is smaller), a stronger control is calculated
-from the optimization, and the optimized value of $O[F_e](T)$ and
-$P( X_T)$ are bigger, which means the end position of $ X_t$
-is more biased towards the right well. This observation is consistent
-with intuition.  When $t<10$~ps, the magnitude of the control is
+The red lines in the figure are produced by the non-equilibrium
+linear response theory developed in this work, i.e. using~\eqref{dida},
+and the green lines are reference simulations with the left-hand-side
+partial derivatives in~\eqref{dida} calculated by the central
+finite difference scheme.
+The good agreement between the red and green line demonstrates that 
+the linear response theory computes the gradient correctly,
+Please note that the calculate gradient, the finite difference scheme
+needs to do $2K$ non-equilibrium simulations
+($K$ is the number of degrees of freedom of the control,
+in this example $K = 10$). By contrast, the non-equilibrium response theory
+only needs one ensemble non-equilibrium simulation, which is much cheaper than
+the finite difference scheme.
+
+When $t<8$~ps, the magnitude of the control is
 almost 0.  Near the end time $T$, the magnitude of the control firstly

### 2013-07-05 writedown/linear.response/paper.tex
upate

@@
 
-\title{Linear response theory and optimal control for a molecular system under non-equilibrium conditions}
+\title{Linear response theory and optimal control for a molecular system under nonequilibrium conditions}
 
@@
 The standard linear response formula allows to answer this question, at least partially.
-In the standard setting it gives us the first order of the change to an equilibrium expectation value as resulting from the non-equilibrium perturbation.
+In the standard setting it gives us the first order of the change to an equilibrium expectation value as resulting from the nonequilibrium perturbation.
 Here, first order means first order in the size of the perturbation.
@@
 We will provide such a formula for the case that the underlying dynamics can be described by Langevin dynamics.
-Furthermore, we will even provide a formula for the second order response of a Langevin system in non-equilibrium to a small perturbation.
+Furthermore, we will even provide a formula for the second order response of a Langevin system in nonequilibrium to a small perturbation.
 
@@
 \[
-\bE_{\rho^{\eps}}[f] \approx \bE[f] - \eps\beta\int_{0}^{\infty}\bE[J(q_{0},p_{0}) (f(q_{t},p_{t})-\bE[f])]\,dt\,,
+\bE_{\rho^{\eps}}[f] \approx \bE_{\rho^{0}}[f] - \eps\beta\int_{0}^{\infty}\bE_{\rho^{0}}[J(q_{0},p_{0}) (f(q_{t},p_{t})-\bE_{\rho^{0}}[f])]\,dt\,,
 \]
@@
 Note that the expectation on the right hand side corresponds to the unperturbed dynamics, because $W_{t}$ is a standard Brownian motion under $\mu$, and the expectation is over all realizations of (\ref{sde2-3}) starting from either any given initial condition $X_{0}=x$ or an arbitrary initial distribution.
-On the other hand, $X_{t}$ under $\nu$ corresponds to the perturbed dynamics, which should become clear upon comparing equations (\ref{sde2-2}) and (\ref{sde2-3})
+On the other hand, $X_{t}$ under $\nu$ corresponds to the perturbed dynamics, which should become clear upon comparing equations (\ref{sde2-2}) and (\ref{sde2-3}).
+Another route to the same result is as follows: One writes down the Onsager-Machlup functional-based path space distribution \cite{Pinski1} for $(X_t)$ and $(x_t)$, then computes the reweighting factor from one to the other and gets $M_T$ again.
 
@@
 
-\subsubsection*{Nonequilibrium Langevin equation}
+\subsubsection*{Nonequilibrium Langevin Dynamics}
 
-We now link our previous considerations with the previous case and consider a non-equilibrium Langevin equation. Specifically, we add a non-gradient perturbation to the Langevin equation 
+We now link our previous considerations with the previous case and consider a nonequilibrium Langevin equation. Specifically, we add a non-gradient perturbation to the Langevin equation 
 \begin{equation}\label{eqn:langevin-1b}
@@
 \item[(i)] The rightmost term in \eqref{eqn:neq-response} is the linear 
-  response to the reference non-equilibrium process (driven by $u_t$ with $\eps=0$).
+  response to the reference nonequilibrium process (driven by $u_t$ with $\eps=0$).
 From (\ref{2nd}) we can also get the second order response term.
@@
 
-\item[(ii)] In the above derivation, we have tacitly assumes that the reference and the perturbed non-equilibrium processes start from the same initial value or
+\item[(ii)] In the above derivation, we have tacitly assumes that the reference and the perturbed nonequilibrium processes start from the same initial value or
   have the same initial distribution.
@@
 
-\item[(iii)] If one wants to calculate the same expectation value for a family of non-equilibrium perturbations $\delta u_t$ then 
+\item[(iii)] If one wants to calculate the same expectation value for a family of nonequilibrium perturbations $\delta u_t$ then 
 one does not need to repeat the calculations of \eqref{eqn:neq-response} for every member of the family. 
@@
 \[
-f = \int_{0}^{T}\left\{\ell (q_{s}) - \frac{1}{2}|u_{s}|^{2}\right\}\,dt + L(q_{T}) 
+f(u) = \int_{0}^{T}\left\{\ell (q_{s}) - \frac{1}{2}|u_{s}|^{2}\right\}\,dt + L(q_{T}) 
 \]
 is maximized where $q_{t}$ is the solution to the controlled Langevin equation (\ref{eqn:langevin-1d}) for $\eps=0$; the functions $\ell$ and $L$ are the running cost and the terminal cost, which are assumed to be continuous and bounded from above; without loss of generality, $\ell$ and $L$ are assumed to depend only on the positions.
-The quadratic term is a penalization that makes sure that the controls do not go through the roof \cite{schuette2012,hartmann2012}  Let us moreover assume that the controls are open-loop (i.e.
+The quadratic term is a penalization that makes sure that the controls do not go through the roof \cite{schuette2012,hartmann2012}.
+Let us moreover assume that the controls are open-loop (i.e.
 without feedback) and can be represented by
@@
 \]   
-denote the direction along which we want to differentiate, where $\delta a=(\delta_{1},\ldots,\delta a_{K})$ are coefficients of the vector $\delta u$ in the basis of the $\Phi_{k}$, and note that 
+denote the direction along which we want to differentiate where $\delta a_{1},\ldots,\delta a_{K}$ are the coefficients of the vector $\delta u$ in the basis of the $\Phi_{k}$, and note that 
 \begin{equation*}
@@
 & = \left(\lim_{\eps\to 0}\frac{I(u+\eps\delta u) - I(u)}{\eps}\right)\cdot \delta u\\
-& = \left(\lim_{\eps\to 0}\frac{\bE_{\rho^{\eps}}[f] - \bE_{\rho^{0}}[f]}{\eps}\right)\cdot\delta u\\
+& = \left(\lim_{\eps\to 0}\frac{\bE_{\rho^{\eps}}[f(u+\eps \delta u)] - \bE_{\rho^{0}}[f(u)]}{\eps}\right)\cdot\delta u\\
 \end{aligned}
@@
 \begin{equation}\label{dida}
-\frac{\partial I}{\partial a_{k}} = \bE_{\rho^0}\left[f\int_{0}^{T}(\sigma^{-1}D(q_{s})\Phi_{k}(s))\cdot dB_{s} \right]\,.
+\frac{\partial I}{\partial a_{k}} = \bE_{\rho^0}\left[f\int_{0}^{T}(\sigma^{-1}D(q_{s})\Phi_{k}(s))\cdot dB_{s} + \int_{0}^{T}u_{s}\cdot\Phi_{k}(s)\,ds \right] \,
 \end{equation}
@@
     of gradient form, $ D( q) = -\nabla_{q}V( q)$,
-    we plot the non-equilibrium driving energy of the system.  The red,
+    we plot the nonequilibrium driving energy of the system.  The red,
     green and blue lines are the potential energy without
-    non-equilibrium driving, with non-equilibrium driving and
-    the non-equilibrium driving perturbed by $\eps\delta u(t)$,
+    nonequilibrium driving, with nonequilibrium driving and
+    the nonequilibrium driving perturbed by $\eps\delta u(t)$,
     respectively.  }
@@
 
-We use the idea of non-equilibrium linear response to investigate the
+We use the idea of nonequilibrium linear response to investigate the
 nonequilibrium phase space probability density distribution, denoted by $\rho^\eps(q, p, t)$, of a one-dimensional model system: one particle in a
@@
 \begin{align}
-  \mh ( p,  q) = \frac 12  p^2 + U( q) 
+  H ( p,  q) = \frac 12  p^2 + U( q) 
 \end{align}
@@
 %$\textsf{kJ} / (\textsf{mol nm}^4)}$
-The non-equilibrium driving $ D$ is given by a force of gradient form:
+The nonequilibrium driving $ D$ is given by a force of gradient form:
 \begin{align}
@@
 \end{align}
-we use $\sigma = 0.16$~\textsf{nm}.  The strength of non-equilibrium
+we use $\sigma = 0.16$~\textsf{nm}.  The strength of nonequilibrium
 driving $u(t)$ is set to be linearly growing, i.e. $ u(t) = k_e\cdot t/T$,
@@
 $k_e = 1$ and $\eps = 1$, see Fig.~\ref{fig:tmp4}
-for the non-equilibrium driving potential and perturbed
+for the nonequilibrium driving potential and perturbed
 potential at time $t = T$. The initial distribution $\rho^0(q, p, 0)$ is set to
@@
   \caption{ The plot of $ \rho^\eps(q,p,t)$ in  phase
-    space under the perturbed non-equilibrium driving described in the text. From left to right the columns present results at times $t =
+    space under the perturbed nonequilibrium driving described in the text. From left to right the columns present results at times $t =
     0$, 5, 10 and 20~\textsf{ps}.  First row: Results of a brute force
-    non-equilibrium simulation. Second row: Results of
-    classical equilibrium linear response theory. In this case we
-    effectively consider an equilibrium system free of driving, and the perturbation $\eps$ is set to 2.  Third row: Results using the
+    nonequilibrium simulation. Second row: Results of
+    classical equilibrium linear response theory, see the text for details. Third row: Results using the
     nonequilibrium linear response result.  }
@@
 distribution of the system at time $t = 0$, 5, 10 and 20~\textsf{ps}. The first
-row presents the result of a brute force non-equilibrium simulation.  It is clear
+row presents the result of a brute force nonequilibrium simulation.  It is clear
 that at the beginning the distribution has only one peak around $q = 0$ and $p = 0$. As  time evolves, an energy barrier develops
@@
 splits into two equally sized peaks.  In the end, the two
-peaks are entirely seperated.  The brute force non-equilibrium
+peaks are entirely seperated.  The brute force nonequilibrium
 simulation serves as the precise result to which the response theory
@@
 
-The third row of Fig.~\ref{fig:tmp6} presents the results computed using the novel nonequilibrium  linear response formula: we first start form the equilibrium distribution, apply  $u(t)$, arrive at a nonequilibrium distribution and then, in a second step, compute the effect of the non-equilibrium driving $\delta u(t)$.
+The third row of Fig.~\ref{fig:tmp6} presents the results computed using the novel nonequilibrium  linear response formula: we first start form the equilibrium distribution, apply  $u(t)$, arrive at a nonequilibrium distribution and then, in a second step, compute the effect of the nonequilibrium driving $\delta u(t)$.
 The numerical results are satisfactorily consistent with
-the brute force non-equilibrium simulation, because the perturbation is still  small enough and the novel linear response theory achieves
+the brute force nonequilibrium simulation, because the perturbation is still  small enough and the novel linear response theory achieves
 good accuracy.
@@
 % See Fig.~\ref{fig:tmp2} and \ref{fig:tmp3} for numerical result.  Both
-% the results of the brute-force non-equilibrium simulation
+% the results of the brute-force nonequilibrium simulation
 % and that of the response theory are presented.
@@
 % a intermediate
-% non-equilibrium simulation with driving force of $k_e = 2$.
+% nonequilibrium simulation with driving force of $k_e = 2$.
 % Notice that the reference simulation in this case is also a
-% non-equilibrium simulation, so the effective strength of perturbation
+% nonequilibrium simulation, so the effective strength of perturbation
 % is reduced to $\eps = 2$. According to Fig.~\ref{fig:tmp1}, this can be
@@
   \caption{Illustration of the optimal control for tilting the potential.  
-    In this plot, $T=9$ ps.  In the gradient descent search, the
-    red lines  calculate
-    the gradient by the non-equilibrium linear respect theory, while
-    the green lines calculates by the central finite difference scheme.
- In the upper plot, the solid lines present the magnitude
-    of the optimal control $u_t$ as calculated based on a family of
-    piecewise linear ansatz functions with time interval 1~ps.
-    The blue insertions show the shape of the
-    nonequilibrium driving potential $U( q) + u_t V(
-    q)$ at the time indicated by the black arrow.  In the lower plot, the
-    dashed lines display the value of $P(q_t)$ as the time
-    evolves,  and the solid lines present $\eta I$.  }\label{fig:tmp7}
+    In this plot, $T=9$ ps.
+Top panel: Optimal control $o_t$ as calculated based on a family of piecewise linear ansatz functions with time interval 1~ps; the blue insertions show the shape of the nonequilibrium driving potential $U + o_t V$ at the times indicated by the black arrows.
+Bottom panel: Optimal gain $I=F(o_t)$ (solid lines) and the probability $P$ of being in the right well associated with $o_t$ (dashed lines) as functions of time along the optimal control.
+Red lines: computation using   the nonequilibrium linear respect theory.
+Green lines: brute force computation as described in the text.
+}\label{fig:tmp7}
 \end{figure}
@@
 \begin{align}
-  I & = \max_{u_t \in\mathcal F} \bE_{\rho^0}\left[ -\int_{0}^{T} \frac{1}{2}|u_{s}|^{2}\,dt + L(q_T) \right],
+  \tilde{I} & = \max_{u_t \in\mathcal F} \bE_{\rho^0}\left[ -\int_{0}^{T} \frac{1}{2}|u_{s}|^{2}\,dt + L(q_T) \right],
 \end{align}
-with $L(q_T)=\chi_{[a-\delta,a+\delta]}(q_T)/\eta$ representing the probability of the end point of the trajectory, $ q_T$, being in the right well ($\chi_I$ denotes the indicator function of the interval $I$) with $\eta$ being a weighting constant, ${\mathcal F}$ denoting the space of function that are piecewise linear  on $[0, T]$ in uniform intervals of length 1 ps, and $\rho^0$ being the non-equilibrium distribution for $u_t$.
-With $P(T) =  \bE_{\rho^0}(L(q_T))$, we thus have
+with $L(q_T)=\chi_{[a-\delta,a+\delta]}(q_T)/\eta$ representing the probability of the end point of the trajectory, $ q_T$, being in the right well ($\chi_I$ denotes the indicator function of the interval $I$) with $\eta$ being a weighting constant, ${\mathcal F}$ denoting the space of function that are piecewise linear  on $[0, T]$ in uniform intervals of length 1 ps, and $\rho^0$ being the initial nonequilibrium distribution.
+With $P(T) = \eta \bE_{\rho^0}(L(q_T))$, the probability of ending up in the right well at time $T$, and $I=\eta \tilde{I}$ we thus have
 \[
-  \eta I = \max_{u_t \in\mathcal F} \Big[ P(T) -
+  I = \max_{u \in\mathcal F} F(u),\quad F(u)= P(T) -
   \frac{\eta}{2}\,
-   \int_0^T |u_t|^2\, dt\Big],
+   \int_0^T |u_t|^2\, dt\,.
 \]
-where $P(T)$ is probability of the end point of the trajectory
-$ q_T$ being in the right well.  It is
+ It is
 clear that for an unbiased double-well, $P(T)$
-is 0.5.  The integral is the ``cost'' of the control and $\eta$  indicates the relative magnitude of the
+is 0.5 if we choose $\rho^0=\rho_0$ as initial distribution.  The integral is the ``cost'' of the control and $\eta$  indicates the relative magnitude of the
 cost.
@@
 Fig.~\ref{fig:tmp7} presents the numerical results of $\eta = 0.01$.
-Starting from an initial guess of linear-incremental control from $u_0 = 0$ to
+Starting from an initial guess of linear control from $u_0 = 0$ to
 $u_T = 1$,
@@
 maximum increment of the control coefficient $\max_k\vert \delta a_k\vert$
-is smaller than $0.02$.
+is smaller than $0.02$, the termination threshold.
 The magnitude of the optimal
-control $u_t$ is presented in the upper plot,
-with blue insertions showing the shape of the controlled potential 
-$U(q) + u_tV(q)$ (tilted double-well potential). 
-The maximized $\eta I$ and the corresponding probability in the right well
-$P( X_T)$ are
-given in the lower plot by the solid and dashed lines, respectively.
-The red lines in the figure are produced by the non-equilibrium
-linear response theory developed in this work, i.e. using~\eqref{dida},
-and the green lines are reference simulations with the left-hand-side
-partial derivatives in~\eqref{dida} calculated by the central
-finite difference scheme.
-The good agreement between the red and green line demonstrates that 
-the linear response theory computes the gradient correctly,
-Please note that the calculate gradient, the finite difference scheme
-needs to do $2K$ non-equilibrium simulations
-($K$ is the number of degrees of freedom of the control,
-in this example $K = 10$). By contrast, the non-equilibrium response theory
-only needs one ensemble non-equilibrium simulation, which is much cheaper than
-the finite difference scheme.
+control 
+\[
+o_t = \textrm{argmax}_{u_t \in\mathcal F} F(u_t)
+\] 
+is presented in the upper panel of Fig.~\ref{fig:tmp7},
+with blue insertions showing the shape of the time-dependent optimal control potential 
+$U(q) + o_t V(q)$ (optimally tilted double-well potential). 
+The maximum $I$ and the corresponding optimal probability to end up in the right well are
+given as functions of time in the lower plot by the solid and dashed lines, respectively.
+The red lines in the figure are produced by the nonequilibrium
+linear response theory developed in this work, i.e., using~\eqref{dida},
+and the green lines represent the brute force reference simulations that has been performed as follows: The optimal control from ${\mathcal F}$ is calculated by a gradient descent based optimization method in which the gradient of the functional with respect to the control is computed by numerical differentiation (central finite differences) in each step.
+The good agreement between the red and green lines demonstrates that 
+the linear response theory computes the gradient correctly.
+Please note that in order to calculate the gradient by the finite difference scheme, one
+needs to do $2K$ nonequilibrium simulations
+(where $K$ is the dimension of ${\mathcal F}$; here $K = 10$). In contrast, the nonequilibrium response theory only needs one nonequilibrium simulation.
 
-When $t<8$~ps, the magnitude of the control is
-almost 0.  Near the end time $T$, the magnitude of the control firstly
+When $t<8$~ps, the magnitude of the control is still small.  Near the end time $T$, the magnitude of the control firstly
 quickly goes up, and then falls down a little bit.  This implies some
 interesting information.  If the system were able to immediately relax
-to its instantaneous equilibrium state (sometimes called
+to its equilibrium state (sometimes called
 quasi-equilibrium), the population in the right well (dashed lines in
@@
 decreases.  The fact that this does not happen, indicates that the
-speed of changing the control is relatively faster that the internal
-time scale of the system, so the system does not have enough time to
-fully relax itself. Therefore, the observed phenomenon is truly
-non-equilibrium, and our non-equilibrium linear response theory is a
+speed of changing the control is relatively fast compared to the 
+time scale of quasi-equilibration of the system, so the system does not have enough time to
+fully relax. Therefore, the observed phenomenon is truly
+nonequilibrium, and our nonequilibrium linear response theory is a
 tool that facilitates the investigation of this optimal forcing
-problem in the non-equilibrium setting.  Intuitively, one concludes
-that as the control goes up, the population in the right well actually
-does not have enough time to go as high as it could be, so when the
-control decreases, it continues going up rather than following the
-decreasing control.
+problem in the nonequilibrium setting.
+The fact that the optimal control is decreasing at the end of the interval is understandable since the population in the right well needs time to be build and increasing the control till the very end would be a waste of energy without corresponding gain in population.
+
+% Intuitively, one concludes
+%that as the control goes up, the population in the right well actually
+%does not have enough time to go as high as it could be, so when the
+%control decreases, it continues going up rather than following the
+%decreasing control.
 
@@
 We derived  first and second order response formulas for molecular dynamics (driven Langevin dynamics) starting from general nonequilibrium distributions.
-For the special case of the initial distribution being the equilibrium distribution of the unperturbed dynamics, the novel linear response formula siimplifies to the well-known standard formula.
-We validated the novel formula in numerical experiments in comparison to brute-force nonequilibrium simulations.
-There, we demonstrated that the novel linear response formula allows to extend the algorithmic use of linear response theory to significantly stronger perturbutions of the system since it permits intermediate steps based on partially propagated nonequilibrium distributions.
+For the special case of the initial distribution being the equilibrium distribution of the unperturbed dynamics, the novel linear response formula simplifies to the well-known standard formula.
+We validated the  formula in numerical experiments in comparison to brute-force nonequilibrium simulations.
+There, we demonstrated that the nonequilibrium linear response formula allows to extend the algorithmic use of linear response theory to significantly stronger perturbutions of the system since it permits intermediate steps based on partially propagated nonequilibrium distributions.
 
-By means of this theory we outlined how to use the novel linear response formula for the computation of optimal controls in molecular dynamics where one desires to find the optimal perturbation/control that maximizes a target functional, that is, a certain expectation value (like the population of a certain region of state space) under a constraint on the energy used in the perturbation.
-Application of the novel nonequilibrium theory allows to compute the gradient of the target functional by computing expectation values \emph{only} for the dynamics at hand which permits efficient application of standard optimization techniques.
+By means of this theory we outlined how to use linear response theory for the computation of optimal controls in molecular dynamics where one desires to find the optimal perturbation/control that maximizes a target functional, that is, a certain expectation value (like the population of a certain region of state space) under a constraint on the energy used in the perturbation.
+Application of our nonequilibrium theory allows to compute the gradient of the target functional by computing expectation values \emph{only} for the dynamics at hand which permits efficient application of standard optimization techniques.
 We illustrated this technique in application to a simple test case and validated it in comparison to brute force optimization.

### 2013-08-21 writedown/linear.response/paper.tex
first revise according to the referee 1

@@
 % \usepackage[dvips]{graphicx}
-% \usepackage{color}
+\usepackage{color}
 % \usepackage{tabularx}
@@
 \begin{abstract}
-  In this paper, we propose a straightforward generalization of linear response theory to systems in nonequilibrium that are subject to nonequilibrium driving. 
+  In this paper, we propose a straightforward generalization of linear response theory on a finite time-horizon to systems in nonequilibrium that are subject to external forcing driving.
 We briefly revisit the standard linear response result for equilibrium systems, where we consider Langevin dynamics as a special case, and then give an alternative derivation using a change-of-measure argument that does not rely on any stationarity or reversibility assumption.
-This procedure moreover easily enables us to calculate the second order correction to the linear response formula (which may or may not be useful in practice).
-Furthermore, we outline how the novel nonequilibirum linear response formula can be used to compute optimal controls of molecular systems for cases in which one wants to steer the system to maximize a certain target expectation value.
+This procedure easily enables us to calculate the second order correction to the linear response formula (which may or may not be useful in practice).
+Furthermore, we outline how the novel nonequilibrium linear response formula can be used to compute optimal controls of molecular systems for cases in which one wants to steer the system to maximize a certain target expectation value.
 We illustrate our approach with simple numerical examples.
@@
 Here, first order means first order in the size of the perturbation.
-This linear response formula has a long history of extensions and generalizations.
+This linear response formula has a long history of extensions and generalizations; see, e.g., \cite{marconi2008,ruelle2009} and the references therein.
 In some sense it has become one of the cornerstones of modern statistical physics since it can be related to the fluctuation dissipation theorem (FDT) which roughly states that for appropriate systems in statistical equilibrium, the average response to small external perturbations can be calculated through the knowledge of suitable correlation functions of the unperturbed statistical system.
-The standard linear response theory holds in a  very general sense \cite{HairerMajda} as long as one assumes that the unperturbed system is in equilibrium.
-%=Martin Hairer and Andrew J Majda, A simple framework to justify linear response theory, Nonlinearity 23 (2010) 909–922 doi:10.1088/0951-7715/23/4/008
+Linear response formulae on the infinite time-horizon have been shown to hold under rather mild assumptions and for nonequilibrium systems with a unique invariant measure \cite{hairer2008}.
 
 There is an increasing number of articles in the literature that report on applications of molecular dynamics to nonequilibrium settings.
-There are many generalization to so-called nonequilibrium steady states based on the generality of the FDT \cite{Seifert}, but despite its wide use, the present authors do not know of a linear response formula that applies to fully general nonequilibirium cases.
-We will provide such a formula for the case that the underlying dynamics can be described by Langevin dynamics.
-Furthermore, we will even provide a formula for the second order response of a Langevin system in nonequilibrium to a small perturbation.
+There are many generalization to so-called nonequilibrium steady states based on the generality of the FDT \cite{Seifert}, but despite its wide use, we are not aware of a linear response formula, even for nonequilibrium (i.e.~irreversible) systems, that does not rely on any kind of stationarity assumption.
+We will provide such a formula for the case that the underlying dynamics can be described by Langevin dynamics that may not have an invariant measure.
+Furthermore, we will provide a formula for the second order response of a nonequilibrium Langevin system subject to a small perturbation.
 
@@
 We will outline how the novel nonequilibrium linear response formula can be used to compute the optimal control of molecular systems.
-In optimal control one seeks the optimal way to perturbed a molecular system such that a certain target expectation value (e.g.
+In optimal control one seeks the optimal way to perturb a molecular system such that a certain target expectation value (e.g.
 population of certain states) is maximized under constraints on the energy of the control.
@@
 
-The outline of the article is as follows: First, we will review the derivation of the standard linear response formula for  rather general diffusion processes which include which Langevin dynamics.
-Next, we will show how to derive first and second order response formulas for the nonequilibrium case and how to apply these formula for the computation of optimal controls.
+The outline of the article is as follows: First, we will review the derivation of finite time-horizon linear response formulae for general diffusion processes and diffusions of Langevin type.
+Next, we will show how to derive first and second order response formulas for genuine nonequilibrium systems and how to apply these formula to the computation of optimal controls.
 Finally, we will validate the nonequilibrium linear response formula and its use for optimal control for simple test cases.
@@
 
-In this section, we want to firstly give a simple and formal derivation of the linear response for equilibrium and nonequilibrium systems.
+In this section, we first want to give a simple and formal derivation of the linear response for equilibrium and nonequilibrium systems.
 To this end we consider a general It\^o stochastic differential equation of the form
@@
 where $X_{t}\in\R^{d}$, $b(\cdot,\cdot)$ is a smooth time-dependent vector field, $a(\cdot)$ a smooth field of $d\times n$ matrices and $B_{t}$ is standard Brownian motion in $\R^{n}$.
-Here $v_{t}\in\R^{d}$ is any given driving force applied to the system, which in general may depend on $X_{t}$ and $t$; $\eps>0$ is a small parameter.
-Whenever (\ref{sde}) without the perturbation $\eps v$ is considered, we write
+Here $v_{t}\in\R^{d}$ is any given driving force applied to the system, typically an affine function of the form $v_{t}=B(X_{t})u(t)$, with $B(\cdot)\in\R^{d\times k}$ and a bounded measurable function $u\colon[0,\infty)\to\R^{k}$ such that (\ref{sde}) has a strong solution for all $t> 0$ whenever $\eps>0$ is sufficiently small.
+When (\ref{sde}) without the perturbation $\eps v$ is considered, we write
 \begin{equation}\label{sdewo}
@@
 
-We consider the equilibrium and nonequilibrium case separately and start with the equilibrium case (see, e.g., \cite{kubo1957,tuckeman2010statistical}).
-To begin with, we assume that the infinitesimal generator
+We consider the equilibrium and nonequilibrium case separately and start with the equilibrium case (see, e.g., \cite{kubo1957,tuckerman2010statistical}).
+To begin with, we assume that the perturbation-free part of the infinitesimal generator
 \[
@@
 \]
-with 
+of (\ref{sde}) with 
 \[
@@
 \]
-of (\ref{sde}) has an isolated eigenvalue 0 corresponding to the unique invariant measure of the dynamics for $\eps=0$ and satisfies a spectral gap condition (see, e.g., \cite{stoltz2012}), so that the operator $(\fwg_{0}^{*})^{-1}\fwg_{1}^{*}$ is bounded (on an appropriate domain), where $\cA_{0}^{*}$ and $\cA_{1}^{*}$ denote the formal adjoints in $L^{2}$, e.g., $\cA_{1}^{*}\phi=-\nabla(v \phi)$.
+has an isolated eigenvalue 0 corresponding to the unique invariant measure $d\mu_{\infty}=\rho_{\infty}dx$.
+We denote by $\cA_{0}^{*}$ and $\cA_{1}^{*}$ the formal adjoints in $L^{2}$, e.g., $\cA_{1}^{*}\phi=-\nabla(v \phi)$.
 
-Specifically, we are interested in the case that (\ref{sde}) has the form of a Langevin equation, in which case  
-\begin{equation}\label{eqn:langevin-1}
-b = \left(\nabla_{p}H, -\nabla_{q} H - \gamma\nabla_{p}H\right )^{T}\,,\quad a= (\zero,\sigma)^{T}\,,
-\end{equation}
-where $H\colon\R^{d}\to\R$, $d=2n$ is the Hamiltonian of the system, $\gamma=\gamma^{T}>0$ the positive semi-definite $n\times n$ friction matrix, $\sigma$ constant, satisfying $2\gamma=\beta\sigma\sigma^{T}$ for some $\beta>0$, and we have used the convention $x=(q,p)\in\R^{n}\times\R^{n}$.
-Then, under some mild growth conditions on the Hamiltonian for large arguments, the unperturbed dynamics has a unique invariant measure with density
-\[
-\rho_{0}(q,p) = \frac{1}{Z}e^{-\beta H(q,p)}\,,\quad  Z=\int  e^{-\beta H(q,p)}\,dqdp\,.
-\] 
-and the above spectral gap condition is met.
-Now let $f$ be any integrable phase space function and let $\rho^{\eps}=\rho^{\eps}(q,p,t)$ denote the probability density of $X_t=(Q_{t},P_{t})$, the solution to the Langevin equation (\ref{sde})\&(\ref{eqn:langevin-1}), assuming that $(Q_{0},P_{0})=(q,p)$ was distributed according to the invariant density $\rho_{0}$.
+Now let $f\colon\R^{2n}\to\R$ be any integrable function and let $\rho^{\eps}=\rho^{\eps}(x,t)$ denote the probability density of $X_t$, at time $t>0$, assuming that $X_{0}=x$ was distributed according to $\rho^{\eps}(x,0)=\rho_{\infty}(x)$.
 We define the expectation \wrt $\rho^{\eps}$ as
 \[
-\bE_{\rho^{\eps}}[f] = \int f \rho^{\eps} \,dqdp\,.
+\bE_{\rho^{\eps}}[f] = \int_{\R^{2n}} f(x) \rho^{\eps}(x,t) \,dx\,.
 \] 
-A classical result, that is usually derived using a formal expansion of the solution to the Kolmogorov forward or backward equations in powers of $\eps$, now states that (see, e.g., \cite{hairer2008} and the references therein)
+A classical result, that is usually derived using a formal expansion of the solution to the Fokker-Planck equation in powers of $\eps$ using the ansatz 
+\[
+\rho^{\eps}(x,t)=\rho_{0}(x,t) + \eps\rho_{1}(x,t) + \ldots\] 
+with 
+\[\rho^{\eps}(x,0)=\rho_{\infty}\,,\textrm{ i.e. }\; \rho_{0}(x,0)=\rho_{\infty}\textrm{  and  } \rho_{i}(x,0)=0\textrm{ for $i>0$},
+\] 
+then states that (see, e.g., \cite{tuckerman2010statistical,hartmann2010})
 \begin{equation}\label{lr}
-\lim_{\eps\to 0}\frac{\bE_{\rho^{\eps}}[f] - \bE_{\rho^{0}}[f]}{\eps} = \bE_{\rho^{0}}[\cA_{1}\cA_{0}^{-1}(f-\bE_{\rho^{0}}[f])]\,.
+\lim_{\eps\to 0}\frac{\bE_{\rho^{\eps}}[f] - \bE_{\rho^{0}}[f]}{\eps} =\int_{\R^{2n}}f(x)\left(\int_{0}^{t}e^{\cA_{0}^{*}(t-s)}\cA_{1}^{*}(s)\rho_{\infty}(x)\,ds\right)dx\,.
 \end{equation}
@@
 \subsubsection*{Green-Kubo relations}
-For the specific case of the Langevin equation (\ref{sde})\&(\ref{eqn:langevin-1}), the general linear response (\ref{lr}) can be recast in form of the better known Green-Kubo formula \cite{marconi2008,risken1996}.
-Using the formal operator identity
+
+Specifically, we are interested in the case that (\ref{sde}) has the form of a second-order Langevin equation, in which case $d=2n$ and 
+\begin{equation}\label{eqn:langevin-1}
+b(x,t) = (J-R)\nabla H(x)\,,\quad a = \sqrt{2\beta^{-1}}R^{1/2}\,,
+\end{equation}
+with $J=-J^{T}$ the canonical $2n\times 2n$ symplectic matrix,
 \[
--\cA_{0}^{-1} g = \int_{0}^{\infty}e^{t\cA_{0}}g \,dt\,,
+J = \left ( \begin{array}{cc} 
+0 & I_{n\times n} \\ -I_{n\times n} &0
+\end{array}\right),
 \]
-that can shown to hold for the generator of the Langevin equation under mild conditions (formally for functions $g$ that are orthogonal to the nullspace of $\cA_{0}^{*}$, i.e.
-functions satisfying $\bE_{\rho_{0}}[g]=0$), we find that
+$R=R^{T}\ge 0$ the positive semidefinite $2n\times 2n$ friction matrix 
+\[
+R = \left ( \begin{array}{cc}
+0 & 0 \\ 0 & \gamma
+\end{array}\right),
+\]
+for $\gamma\in\R^{n\times n}$ symmetric positive definite, and 
+\[
+H\colon\R^{2n}\to\R\,,\quad H(x) = \frac{1}{2}|x_{2}|^{2} + V(x_{1})
+\] 
+the Hamiltonian of the system.
+We assume that the potential $V$ is bounded from below.
+Then, under some mild growth conditions on $V$ for $|x_{1}|\to\infty$ (see, e.g., \cite{mattingly2002}), the unperturbed dynamics has a unique invariant measure with density
+\[
+\rho_{\infty}(x) = \frac{1}{Z}\exp(-\beta H(x))\,,\quad  Z=\int_{\R^{2n}}  \exp(-\beta H(x))\,dx\,.
+\] 
+In this case, the linear response result (\ref{lr}) can be recast in form of the better known Green-Kubo formula \cite{risken1996}.
+Assuming that the forcing is of the form $v_{t}=Bu(t)$ with $u\colon[0,\infty)\to\R^{n}$ and $B\in\R^{2n\times n}$ being any given time-dependent function, the above expression can be recast as
 \begin{equation}\label{GreenKubo}
-\lim_{\eps\to 0}\frac{\bE_{\rho^{\eps}}[f] - \bE_{\rho^{0}}[f]}{\eps} = -\beta\int_{0}^{\infty}\bE_{\rho^{0}}[J(q_{0},p_{0}) (f(q_{t},p_{t})-\bE_{\rho^{0}}[f])]\,,dt
+\lim_{\eps\to 0}\frac{\bE_{\rho^{\eps}}[f] - \bE_{\rho^{0}}[f]}{\eps} = \beta\int_{0}^{t}F(t-s)u(s)\,ds
 \end{equation}
-in terms of the dissipative flux $J=-v\cdot\nabla H$. In other words: 
+where 
 \[
-\bE_{\rho^{\eps}}[f] \approx \bE_{\rho^{0}}[f] - \eps\beta\int_{0}^{\infty}\bE_{\rho^{0}}[J(q_{0},p_{0}) (f(q_{t},p_{t})-\bE_{\rho^{0}}[f])]\,dt\,,
+F(t) = - \int_{\R^{2n}}\bE_{x}[f(x_{t})] B^{T}\nabla H(x)\rho_{\infty}(x)\,dx
 \]
- where the expectation under the integral is taken over all realizations of the unperturbed equilibrium Langevin equation (\ref{sdewo})\&(\ref{eqn:langevin-1}) with initial distribution 
+is the response function and the expectation under the integral is taken over all realization of the equilibrium process (\ref{sdewo}) starting from $x_{0}=x$. In other words: 
 \[
- \rho^{0}(q,p,0)=\rho_{0}\,.
+\bE_{\rho^{\eps}}[f] \approx \bE_{\rho_{\infty}}[f] + \eps\beta\int_{0}^{t}F(t-s)u(s)\,ds\,,
 \]
+where we have used the fact that $\rho^{0}(x,t)\to \rho_{\infty}(x)$ uniformly on $\R^{2n}\times[0,\infty)$.
+
+\paragraph*{Remark.}
+It is possible to extend the above framework to the case of observables that are functionals of the path, i.e.~to functions the form 
+\[
+\varphi^{\eps}(x,t) = \bE_{x}\left[\int_{0}^{t}f(X_{s})ds\right]\,,
+\]
+where the expectation is over all realizations of the \emph{nonequilibrium} process $X_{t}$ with initial condition $X_{0}=x$.
+In this case a linear response result can be obtained by expanding the solution of the Kolmogorov backward equation
+\[
+\frac{\partial\varphi^{\eps}}{\partial t} = \cA^{\eps}\varphi^{\eps} - f\,,\quad \varphi^{\eps}(x,0) = 0\,, 
+\]
+rather than the Fokker-Planck equation, where the equivalence between the solution of the backward equation and the conditional expectation of the path functional follows from the Feynman-Kac formula \cite[Thm.~1.3.17]{pham2009}.
 
@@
 
-The classical response theory has the limitation that the reference (equilibrium) distribution $\rho_{0}$ must be the unique stationary probability measure of the unperturbed dynamics.
-Moreover the validity of the linear approximation relies on a spectral gap condition that is difficult to verify in practice.
-In particular the perturbation argument does not provide a framework, under which the second and even higher order responses can easily be derived.
+The derivation of the classical response result heavily relies on the fact that the unperturbed system has a unique equilibrium distribution, which requires that the generator of the process has certain spectral properties; cf.~\cite{hairer2008,stoltz2012}.
+Moreover the usual perturbation argument does not provide a framework, under which the second and even higher order responses can easily be derived.
+Here we propose an alternative derivation of the linear response result, that is based on a change of drift in the corresponding SDE and which allows for an easy generalization of the above linear response result to nonequilibrium systems.
 
 \subsubsection*{Girsanov transformation}
-
-Here we propose an alternative (purely formal) derivation of the linear response result, that is based on a change of drift in the corresponding SDE and which allows for an easy generalization of the above linear response result to nonequilibrium systems.
 We will briefly review the idea of the change of drift via Girsanov transformations; for details we refer to the textbook \cite{oksendal2003stochastic}.
-Let $x_{t},X_{t}$ be the solutions to the following stochastic differential equations:
+To this end, let $x_{t}=x_{t}(\omega)$ and $X_{t}=X_{t}(\omega)$ be the solutions of our generic stochastic differential equations
 \begin{subequations}\label{sde2}
 \begin{align}
-dx_{t} & = b(x_{t},t)dt + a(x_{t})dB_{t} \label{sde2-1}\\
-dX_{t} & = (b(X_{t},t) + \eps v_{t})dt +a(X_{t})dB_{t} \label{sde2-2}
+dx_{t} & = b(x_{t},t)dt + a(x_{t})dB_{t}\,,\quad 0\le t\le T  \label{sde2-1}\\
+dX_{t} & = (b(X_{t},t) + \eps v_{t})dt +a(X_{t})dB_{t}\,,\quad 0\le t\le T \label{sde2-2}
 \end{align}
 \end{subequations}
-for $0\le t\le T$ and with fixed initial conditions
+for $T<\infty$ and deterministic initial conditions
 \[
-x_{0} = X_{0} = x\,.
+x_{0}(\omega) = X_{0}(\omega) = x\quad\textrm{(almost surely)}\,.
 \]
-Suppose that there exists a auxiliary stochastic process $\xi_{t}\in\R^{m}$ such that 
+Suppose that there exists an auxiliary stochastic process $\xi_{t}\in\R^{m}$ such that 
 \begin{equation}\label{xi}
@@
 \]
-which allows us to express (\ref{sde2-2}) by
+which allows us to rewrite (\ref{sde2-2}) as
 \begin{equation}\label{sde2-3}
@@
 \end{equation}
-It follows from the Girsanov theorem \cite[Thm.~8.6.8]{oksendal2003stochastic}, sometimes also called \emph{Cameron-Martin-Girsanov theorem} \cite{stroock2006}, that $W_{t}$ is again a Brownian motion under a new probability measure that has a density \wrt the Gaussian probability measure that is generated by the Brownian motion $B_{t}$.\footnote{A quick-and-dirty derivation of the above change-of-measure formula can be easily obtained, if the noise covariance $a(\cdot)a(\cdot)^{T}$ has full rank with bounded inverse.
-Then, using Euler's method for (\ref{sde2}), it follows that (\ref{likelihood}) is basically the likelihood ratio between the time-discrete path densities of (\ref{sde2-2}) and (\ref{sde2-1}).
-} Specifically, let $\nu$ denote the law of the Brownian motion $B_{t}$ and define a new probability measure $\mu$ on the space of continuous trajectories by
+It follows from the Girsanov theorem \cite[Thm.~8.6.8]{oksendal2003stochastic}, sometimes also called \emph{Cameron-Martin-Girsanov theorem} \cite{stroock2006}, that $W_{t}$ is again a Brownian motion under a new probability measure that has a density \wrt the Gaussian probability measure that is generated by the Brownian motion $B_{t}$.
+Specifically, let $\nu$ denote the law of the Brownian motion $B_{t}$ and define a new probability measure $\mu$ on the space of continuous trajectories by
 \[
@@
 \end{equation}
-Technical details aside, the Girsanov theorem implies that $W_{t}$  for any function $f=f(\{X_{t}\}_{0\le t\le T})$ that is integrable \wrt $\nu$, we have the identity
+Technical details aside, the Girsanov theorem implies that $W_{t}$  for any functional 
 \[
-\bE_{\nu}[f] := \int f(\{X_{t}\}_{0\le t\le T})\, d\nu = \int f(\{X_{t}\}_{0\le t\le T}) \frac{d\nu}{d\mu}\,d\mu =: \bE_{\mu}[M^{-1}_{T}f]\,.
+\varphi(\omega)=\int_{0}^{T}f((X_{s}(\omega))\,ds + g(X_{T}(\omega))
+\] 
+that is integrable \wrt $\nu$, we have the identity\footnote{One of the omitted technical details is Novikov's condition \cite[pp.~162]{oksendal2003stochastic} that guarantees that $(M_{t})_{0\le t\le T}$ is a Martingale, which by $\bE_{\mu}[1]=\bE_{\nu}[M_{T}]=1$ implies normalizability of the new probability measure $\mu$.}
+\[
+\bE_{\nu}[\varphi] := \int_{\Omega} \varphi(\omega)\, d\nu(\omega) = \int_{\Omega} \varphi(\omega) M_{T}^{-1}(\omega)\,d\mu(\omega) =: \bE_{\mu}[M^{-1}_{T}\varphi]\,.
 \]
@@
 On the other hand, $X_{t}$ under $\nu$ corresponds to the perturbed dynamics, which should become clear upon comparing equations (\ref{sde2-2}) and (\ref{sde2-3}).
-Another route to the same result is as follows: One writes down the Onsager-Machlup functional-based path space distribution \cite{Pinski1} for $(X_t)$ and $(x_t)$, then computes the reweighting factor from one to the other and gets $M_T$ again.
+
+\paragraph*{Remark.}
+
+A quick-and-dirty derivation of the above change-of-measure formula can be easily obtained, if the noise covariance $a(\cdot)a(\cdot)^{T}$ has full rank with bounded inverse.
+Then, using Euler's method for (\ref{sde2}), it follows that (\ref{likelihood}) is basically the likelihood ratio between the time-discrete path densities of (\ref{sde2-2}) and (\ref{sde2-1}).
+Another route to the same result is by using the Onsager-Machlup functional \cite{Pinski1} for $(X_t)_{0\le t\le T}$ and $(x_t)_{0\le t\le T}$; then $M_{T}$ is found as the likelihood ratio of the two densities.
 
@@
 \begin{equation}\label{lr-alt}
-\lim_{\eps\to 0}\frac{\bE_{\rho^{\eps}}[f] + \bE_{\rho^{0}}[f]}{\eps} = \bE_{\rho^{0}}\left[f(\{x_{t}\}_{0\le t\le T})\int_{0}^{T}\xi_{s}\cdot dB_{s} \right]\,,
+\lim_{\eps\to 0}\frac{\bE_{\rho_{x}^{\eps}}[\varphi] + \bE_{\rho_{x}^{0}}[\varphi]}{\eps} = \bE_{\rho_{x}^{0}}\left[\varphi \int_{0}^{T}\xi_{s}\cdot dB_{s} \right]\,,
 \end{equation}
-where $x_{t}$ is the solution (\ref{sde2-1}) and $\rho^{\eps}$, $\rho^{0}$ denote the distributions of the perturbed and unperturbed dynamics, in accordance with the notation used in the previous section.
-Note that $\rho^{0}\neq\rho_{0}$ does not need to be an equilibrium distribution.
-Further note that $f$ and $B_{t}$ are not independent, hence $f$ times the integral over the Brownian motion does not average to zero in general.
+where, in accordance with the previous linear response formula, $\rho_{x}^{\eps}$ and $\rho_{x}^{0}$ denote the path probability distributions of perturbed and unperturbed processes (\ref{sde2-2}) and (\ref{sde2-1}) with deterministic initial condition.
+On the right hand side of the last equation $\varphi$ is understood as a functional of $(x_{t})_{0\le t\le T}$, the solution to (\ref{sde2-1}) with initial condition $x_{0}=x$, and $\xi_{t}$ is the solution to $a(x_{t})\xi_{t}=v_{t}$.
+By averaging the the initial values $x$ on both sides of (\ref{lr-alt}) over any given initial distribution, one obtains an analogous formula for distributed initial conditions.
+
+Note that $\varphi$ and $B_{t}$ are not independent, hence the product of $\varphi$ and the integral over the Brownian motion does not average to zero in general.
 
@@
 \begin{aligned}
-\bE_{\rho^{\eps}}[f] & \approx \bE_{\rho^{0}}[f]  + \eps \bE_{\rho^{0}}\left[f\int_{0}^{T}\xi_{s}\cdot dB_{s} \right]\\  
-& + \frac{\eps^{2}}{2} \bE_{\rho^{0}}\left[f\left\{\left(\int_{0}^{T}\xi_{s}\cdot dW_{s}\right)^2 -\int_0^T |\xi_s|^2\,ds\right\} \right].
+\bE_{\rho_{x}^{\eps}}[f] & \approx \bE_{\rho_{x}^{0}}[\varphi]  + \eps \bE_{\rho_{x}^{0}}\left[\varphi\int_{0}^{T}\xi_{s}\cdot dB_{s} \right]\\  
+& + \frac{\eps^{2}}{2} \bE_{\rho_{x}^{0}}\left[\varphi\left\{\left(\int_{0}^{T}\xi_{s}\cdot dW_{s}\right)^2 -\int_0^T |\xi_s|^2\,ds\right\} \right].
 \end{aligned}
@@
 
-We now link our previous considerations with the previous case and consider a nonequilibrium Langevin equation. Specifically, we add a non-gradient perturbation to the Langevin equation 
+We now link our previous considerations with the previous case and consider a nonequilibrium Langevin equation.
+Specifically, we add a non-gradient perturbation to the drift (\ref{eqn:langevin-1}) of the previous Langevin equation, i.e., we set
 \begin{equation}\label{eqn:langevin-1b}
-b = \left(\nabla_{p}H, -\nabla_{q} H - \gamma\nabla_{p}H  + D(q) u \right)^{T}\,,\quad a= (\zero,\sigma)^{T}\,,
+b(x,t) = (J-R)\nabla H(x) +  B(x) u(t)\,,\quad a(x) = \sqrt{2\beta^{-1}}R^{1/2}
 \end{equation}
-where $u\in\R^{n}$ is some control variable and $D(\cdot)\in\R^{n\times n}$ satisfies the Fredholm alternative $\mathrm{range}(D(\cdot))\perp \mathrm{ker}(\sigma^{T})$, where $\mathrm{ker}(\sigma^T)$ denotes the kernel of $\sigma^T$.
-As we are interested in small perturbation $\delta u$ in the controls $u$, it is convenient to further specify what $v$ in (\ref{sde2}) is.
-Specifically, we assume that $v$ is of the form
+with the definitions of $J,R$ and $H$ unchanged, $u\in\R^{n}$ being some control variable, and $B(\cdot)\in\R^{2n\times n}$ given by 
+\[
+B(x) = \left( \begin{array}{c}
+0 \\ D(x) \end{array}\right)
+\]
+The matrix $D(\cdot)\in\R^{n\times n}$ must be chosen such that $B$ satisfies the Fredholm alternative $\mathrm{range}(B(\cdot))\perp \mathrm{ker}((a(\cdot))^{T})$, where $\mathrm{ker}(a^T)$ denotes the kernel of $a^T$.
+
+We are interested in the dynamics under a small perturbation $\delta u$ of the control $u$, Specifically, we assume that $v$ in (\ref{sde2}) is of the form 
 \begin{equation}\label{eqn:langevin-1c}
-v = (\zero, D(q)\,\delta u)
+v_{t} = B(X_{t})\delta u(t)
 \end{equation}
@@
 \[
-\sigma \xi_{t} = D(Q_{t})\,\delta u_{t}\,.
+\sqrt{2\beta^{-1}}R^{1/2} \xi_{t} = D(X_{t})\,\delta u_{t}\,.
 \]
+Hence the equation for $\xi_{t}$ is solvable if and only if $\mathrm{range}(D(\cdot))\perp \mathrm{ker}(R^{1/2})$, which, since $R$ has full rank, means that $D(\cdot)$ must be invertible almost everywhere.
+
 To be more explicit, the equation we are considering has the form
 \begin{equation}\label{eqn:langevin-1d}
-\begin{aligned}
-dQ_t & =  \nabla_p H(Q_t,P_t)dt\\
-dP_t & =  -\nabla_q H(Q_t,P_t)dt - \gamma \nabla_p H(Q_t,P_t)dt + D(Q_t)\,(u_t+\eps \delta u_t)dt+\sigma dB_t,
-\end{aligned}
+dX_t =  (J-R)\nabla H(X_{t})dt + B(X_t)(u(t)+\eps \delta u(t))dt+a\, dB_t\,.
 \end{equation}
-so that $X_t$ from above now is $(Q_t,P_t)$ and $x_t=(q_t,p_t)$ from above is the solution of our last equation for $\eps=0$.
-
-The equation is solvable by the requirement  $\mathrm{range}(D(\cdot))\perp \mathrm{ker}(\sigma^{T})$, which, if $\sigma$ has full rank, means that $D(\cdot)$ must be invertible almost everywhere.
-Hence the linear response formula (\ref{lr-alt})  becomes
+In this case the linear response formula (\ref{lr-alt})  becomes
 \begin{equation}\label{lr-alt2}
-\lim_{\eps\to 0}\frac{\bE_{\rho^{\eps}}[f] - \bE_{\rho^{0}}[f]}{\eps} =  \bE_{\rho^{0}}\left[f(\{(q_{t},p_{t})_{0\le t\le T}\})\int_{0}^{T}(\sigma^{-1}D(q_{s})\,\delta u_{s})\cdot dB_{s} \right]\,
+\lim_{\eps\to 0}\frac{\bE_{\rho_{x}^{\eps}}[\varphi] - \bE_{\rho_{x}^{0}}[\varphi]}{\eps} =  \bE_{\rho_{x}^{0}}\left[\varphi\int_{0}^{T}(\sigma^{-1}D(x_{s})\,\delta u_{s})\cdot dB_{s} \right]\,
 \end{equation}
-or, in other words: 
+with the shorthands $\sigma=\sqrt{2\beta^{-1}}R^{1/2}$ and 
+\[
+\varphi = \int_{0}^{T}f(x_{s})ds+g(x_{T})\,.
+\]
+In other words: 
 \begin{equation}\label{eqn:neq-response}
-\bE_{\rho^{\eps}}[f] \approx \bE_{\rho^0}[f] + \eps\bE_{\rho^0}\left[f(\{q_{t},p_{t}\}_{0\le t\le T})\int_{0}^{T}(\sigma^{-1}D(q_{s})\,\delta u_{s})\cdot dB_{s} \right]\,.
+\bE_{\rho_{x}^{\eps}}[\varphi] \approx \bE_{\rho_{x}^0}[\varphi] + \eps\bE_{\rho_{x}^0}\left[\varphi\int_{0}^{T}(\sigma^{-1}D(x_{s})\,\delta u_{s})\cdot dB_{s} \right]\,.
 \end{equation}
-Here, as before, the expectation on the right is over all realizations of (\ref{eqn:langevin-1d}) for $\eps=0$ with arbitrary initial conditions (fixed or distributed); see also the remark below on the choice of initial conditions.
+Here, as before, the expectation on the right hand side of the linear response formula is over all realizations of (\ref{eqn:langevin-1d}) for $\eps=0$ starting at $x$; see also the remark below on the choice of initial conditions.
 
@@
 
-As one can easily get lost in the various integral transformations, measures $\nu$ and $\mu$, distributions $\rho^{\eps}$ and $\rho^{\eps}$ etc.
+As one can easily get lost in the various transformations and changes of measure, involving $\nu$, $\mu$, $\rho_{x}^{\eps}$, $\rho^{0}_{x}$ etc.
 it may be helpful to understand how the linear response formulas (\ref{lr-alt}) or (\ref{lr-alt2}) can be realized algorithmically.
@@
 From (\ref{2nd}) we can also get the second order response term.
-The latter is bounded by the assumption that the controls have bounded second moment.
+The latter is bounded if the controls have bounded second moment, which is a typical requirement in optimal control applications.
+Higher order moments need not exist.
 
-\item[(ii)] In the above derivation, we have tacitly assumes that the reference and the perturbed nonequilibrium processes start from the same initial value or
+\item[(ii)] In the above derivation, we have tacitly assumed that the reference and the perturbed nonequilibrium processes start from the same initial value or
   have the same initial distribution.
@@
   calculated for the single basis functions. Then with a linear combination of the responses on basis functions, one can derive
-  the responses for the whole family. This feature will be used below when discussion optimal control as an application of the linear response formula.
+  the responses for the whole family. This feature will be used below when discussing optimal control as an application of the linear response formula.
 
@@
 be equilibrium distribution of the unperturbed system.
+\recheck{
+  We use Euler scheme with time step of $10^{-4}$ to discretize the
+  Langevin equation. The inital configurations for nonequilibrium
+  simulation are generated by taking configurations
+  from an equilibrium simulation of $10^6$~\textsf{ps} at an 1~\textsf{ps} time interval.
+  Then starting with these inital configurations,  the nonequilibrium
+  Langevin equation is integrated until 20~\textsf{ps} with the same numerical
+  scheme and time step.
+}
 
@@
 Bottom panel: Optimal gain $I=F(o_t)$ (solid lines) and the probability $P$ of being in the right well associated with $o_t$ (dashed lines) as functions of time along the optimal control.
-Red lines: computation using   the nonequilibrium linear respect theory.
+Red lines: computation using   the nonequilibrium linear response theory.
 Green lines: brute force computation as described in the text.
@@
 \end{align}
-with $L(q_T)=\chi_{[a-\delta,a+\delta]}(q_T)/\eta$ representing the probability of the end point of the trajectory, $ q_T$, being in the right well ($\chi_I$ denotes the indicator function of the interval $I$) with $\eta$ being a weighting constant, ${\mathcal F}$ denoting the space of function that are piecewise linear  on $[0, T]$ in uniform intervals of length 1 ps, and $\rho^0$ being the initial nonequilibrium distribution.
+with $L(q_T)=\chi_{[a-\delta,a+\delta]}(q_T)/\eta$ representing the probability of the end point of the trajectory, $ q_T$, being in the right well ($\chi_I$ denotes the indicator function of the interval $I$) with $\eta$ being a weighting constant, ${\mathcal F}$ denoting the space of function that are piecewise linear  on $[0, T]$ in uniform intervals of length 1 ps, and $\rho^0$ being the
+\recheck{initial nonequilibrium distribution $\rightarrow$ nonequilibrium distribution under control $u_t$}.
 With $P(T) = \eta \bE_{\rho^0}(L(q_T))$, the probability of ending up in the right well at time $T$, and $I=\eta \tilde{I}$ we thus have
@@
    \]
+   \recheck{shall we use $I(u)$ instead of $F(u)$?, so that the
+     notation is consistent with  Sec. 3.}
  It is
@@
 
+\recheck{ In all our numerical tests for the optimal control problem, the
+  Langevin dyanmics Eq.~\eqref{ocp1} is discretized by the Euler
+  scheme with a time step of $10^{-3}$~\textsf{ps}.
+  The initial distribution $\rho^0(q,p,0)$ is also set to be the
+  equilibrium distribution $\rho_0$ of the unperturbed system.
+  Therefore, an equilibrium simulation of $10^6$~\textsf{ps} is firstly performed (with
+  the same numerical scheme and time step), and
+  configurations are saved every 10~\textsf{ps} (in total $10^5$
+  configurations) along the trajectory.
+  Then by using these configurations as inital configurations,
+  $10^5$ nonequilibrium
+  trajectories are integrated until 9~\textsf{ps}, while at the same time
+  the nonequilibrium responses are calculated and then averaged to
+  estimate the ensemble average on the r.h.s of Eq.~\eqref{dida}.
+  The statistical
+  uncertainty of this estimate varies by each step of iteration, but
+  is roughly $7\times10^{-4}$.  The statistical uncertainty for the 
+  optimization target $I$ is roughly $1\times10^{-3}$.
+  % The gradient search~\eqref{gd} is performed with a constant increment
+  % $\tau_n = 30$.
+}
+
 Fig.~\ref{fig:tmp7} presents the numerical results of $\eta = 0.01$.
@@
 $u_T = 1$,
-the gradient search~\eqref{gd} converges at the 22nd step, when the
+the gradient search~\eqref{gd} \recheck{uses a constant $\tau_n = 30$ and} converges at the 22nd step, when the
 maximum increment of the control coefficient $\max_k\vert \delta a_k\vert$
@@
 and the green lines represent the brute force reference simulations that has been performed as follows: The optimal control from ${\mathcal F}$ is calculated by a gradient descent based optimization method in which the gradient of the functional with respect to the control is computed by numerical differentiation (central finite differences) in each step.
+% : \redc{$\partial I / \partial a_k \approx (I^+ - I^-) / (a_k^+ - a_k^-)$}
 The good agreement between the red and green lines demonstrates that 
@@
 
-When $t<8$~ps, the magnitude of the control is still small.  Near the end time $T$, the magnitude of the control firstly
+When $t<8$~\textsf{ps}, the magnitude of the control is still small.  Near the end time $T$, the magnitude of the control firstly
 quickly goes up, and then falls down a little bit.  This implies some

### 2013-08-21 writedown/linear.response/paper.tex
Carsten's revise, 2nd round

@@
 \begin{abstract}
-  In this paper, we propose a straightforward generalization of linear response theory on a finite time-horizon to systems in nonequilibrium that are subject to external forcing driving.
+  In this paper, we propose a straightforward generalization of linear response theory on a finite time-horizon to systems in nonequilibrium that are subject to external forcing.
 We briefly revisit the standard linear response result for equilibrium systems, where we consider Langevin dynamics as a special case, and then give an alternative derivation using a change-of-measure argument that does not rely on any stationarity or reversibility assumption.
@@
 In some sense it has become one of the cornerstones of modern statistical physics since it can be related to the fluctuation dissipation theorem (FDT) which roughly states that for appropriate systems in statistical equilibrium, the average response to small external perturbations can be calculated through the knowledge of suitable correlation functions of the unperturbed statistical system.
-Linear response formulae on the infinite time-horizon have been shown to hold under rather mild assumptions and for nonequilibrium systems with a unique invariant measure \cite{hairer2008}.
+Linear response formulae on the infinite time-horizon have been shown to hold under rather mild assumptions even for nonequilibrium systems that possess a unique invariant measure \cite{hairer2008}.
 
 There is an increasing number of articles in the literature that report on applications of molecular dynamics to nonequilibrium settings.
-There are many generalization to so-called nonequilibrium steady states based on the generality of the FDT \cite{Seifert}, but despite its wide use, we are not aware of a linear response formula, even for nonequilibrium (i.e.~irreversible) systems, that does not rely on any kind of stationarity assumption.
+There are many generalization to so-called nonequilibrium steady states based on the generality of the FDT \cite{Seifert}, but despite its wide use, we are not aware of a linear response formula for equilibrium or nonequilibrium (i.e.~irreversible) systems that does not rely on any kind of stationarity assumption.
 We will provide such a formula for the case that the underlying dynamics can be described by Langevin dynamics that may not have an invariant measure.
@@
 We will outline how the novel nonequilibrium linear response formula can be used to compute the optimal control of molecular systems.
-In optimal control one seeks the optimal way to perturb a molecular system such that a certain target expectation value (e.g.
+In optimal control one seeks the optimal way to perturb a molecular system, such that a certain target expectation value (e.g.
 population of certain states) is maximized under constraints on the energy of the control.
-In general the control drives the molecular system under control out of equilibrium.
+In general, the control drives the molecular system under control out of equilibrium.
 Thus, the nonequilibrium linear response formula can be used to find the optimal correction of the present control regarding the expectation value of interest.
@@
 where $X_{t}\in\R^{d}$, $b(\cdot,\cdot)$ is a smooth time-dependent vector field, $a(\cdot)$ a smooth field of $d\times n$ matrices and $B_{t}$ is standard Brownian motion in $\R^{n}$.
-Here $v_{t}\in\R^{d}$ is any given driving force applied to the system, typically an affine function of the form $v_{t}=B(X_{t})u(t)$, with $B(\cdot)\in\R^{d\times k}$ and a bounded measurable function $u\colon[0,\infty)\to\R^{k}$ such that (\ref{sde}) has a strong solution for all $t> 0$ whenever $\eps>0$ is sufficiently small.
+Here $v_{t}\in\R^{d}$ is any given driving force applied to the system, typically an affine function of the form $v_{t}=c(X_{t})u(t)$, with $c(\cdot)\in\R^{d\times k}$ and a bounded measurable function $u\colon[0,\infty)\to\R^{k}$ such that (\ref{sde}) has a strong solution for all $t> 0$ whenever $\eps>0$ is sufficiently small.
 When (\ref{sde}) without the perturbation $\eps v$ is considered, we write
@@
 
-\subsection{Small perturbations from equilibrium: Langevin dynamics}
+\subsection{Small perturbations from equilibrium: Langevin dynamics}\label{ssec:equiLRT}
 
@@
 \[
-H\colon\R^{2n}\to\R\,,\quad H(x) = \frac{1}{2}|x_{2}|^{2} + V(x_{1})
+H\colon\R^{2n}\to\R\,,\quad H(x) = \frac{1}{2}|p|^{2} + U(q) \quad (x=(q,p))
 \] 
 the Hamiltonian of the system.
-We assume that the potential $V$ is bounded from below.
-Then, under some mild growth conditions on $V$ for $|x_{1}|\to\infty$ (see, e.g., \cite{mattingly2002}), the unperturbed dynamics has a unique invariant measure with density
+We assume that the potential $U$ is bounded from below.
+Then, under some mild growth conditions on $U$ for $|q|\to\infty$ (see, e.g., \cite{mattingly2002}), the unperturbed dynamics has a unique invariant measure with density
 \[
@@
 In this case, the linear response result (\ref{lr}) can be recast in form of the better known Green-Kubo formula \cite{risken1996}.
-Assuming that the forcing is of the form $v_{t}=Bu(t)$ with $u\colon[0,\infty)\to\R^{n}$ and $B\in\R^{2n\times n}$ being any given time-dependent function, the above expression can be recast as
+Assuming that the forcing is of the form $v_{t}=Cu(t)$ with $u\colon[0,\infty)\to\R^{n}$ and $C\in\R^{2n\times n}$ being any suitable given time-dependent function, the above expression can be recast as
 \begin{equation}\label{GreenKubo}
@@
 \[
-F(t) = - \int_{\R^{2n}}\bE_{x}[f(x_{t})] B^{T}\nabla H(x)\rho_{\infty}(x)\,dx
+F(t) = - \int_{\R^{2n}}\bE_{x}[f(x_{t})] C^{T}\nabla H(x)\rho_{\infty}(x)\,dx
 \]
@@
 \]
-where we have used the fact that $\rho^{0}(x,t)\to \rho_{\infty}(x)$ uniformly on $\R^{2n}\times[0,\infty)$.
+where we have used the fact that $\rho^{0}(x,t)= \rho_{\infty}(x)$ for all $t\ge 0$.
 
@@
 It follows from the Girsanov theorem \cite[Thm.~8.6.8]{oksendal2003stochastic}, sometimes also called \emph{Cameron-Martin-Girsanov theorem} \cite{stroock2006}, that $W_{t}$ is again a Brownian motion under a new probability measure that has a density \wrt the Gaussian probability measure that is generated by the Brownian motion $B_{t}$.
-Specifically, let $\nu$ denote the law of the Brownian motion $B_{t}$ and define a new probability measure $\mu$ on the space of continuous trajectories by
+Specifically, let $P$ denote the law of the Brownian motion $B_{t}$ and define a new probability measure $Q$ on the space of continuous trajectories by
 \[
-d\mu=M_{T}d\nu
+dQ=M_{T}dP
 \]
@@
 \] 
-that is integrable \wrt $\nu$, we have the identity\footnote{One of the omitted technical details is Novikov's condition \cite[pp.~162]{oksendal2003stochastic} that guarantees that $(M_{t})_{0\le t\le T}$ is a Martingale, which by $\bE_{\mu}[1]=\bE_{\nu}[M_{T}]=1$ implies normalizability of the new probability measure $\mu$.}
+that is integrable \wrt $\nu$, we have the identity\footnote{One of the omitted technical details is Novikov's condition \cite[pp.~162]{oksendal2003stochastic} that guarantees that $(M_{t})_{0\le t\le T}$ is a Martingale, which by $\bE_{Q}[1]=\bE_{P}[M_{T}]=1$ implies normalizability of the new probability measure $Q$.}
 \[
-\bE_{\nu}[\varphi] := \int_{\Omega} \varphi(\omega)\, d\nu(\omega) = \int_{\Omega} \varphi(\omega) M_{T}^{-1}(\omega)\,d\mu(\omega) =: \bE_{\mu}[M^{-1}_{T}\varphi]\,.
+\bE_{P}[\varphi] := \int_{\Omega} \varphi(\omega)\, dP(\omega) = \int_{\Omega} \varphi(\omega) M_{T}^{-1}(\omega)\,dQ(\omega) =: \bE_{Q}[M^{-1}_{T}\varphi]\,.
 \]
@@
 \]
-is the density of $\nu$ relative to $\mu$.
-Note that the expectation on the right hand side corresponds to the unperturbed dynamics, because $W_{t}$ is a standard Brownian motion under $\mu$, and the expectation is over all realizations of (\ref{sde2-3}) starting from either any given initial condition $X_{0}=x$ or an arbitrary initial distribution.
-On the other hand, $X_{t}$ under $\nu$ corresponds to the perturbed dynamics, which should become clear upon comparing equations (\ref{sde2-2}) and (\ref{sde2-3}).
+is the density of $P$ relative to $Q$.
+Note that the expectation on the right hand side corresponds to the unperturbed dynamics, because $W_{t}$ is a standard Brownian motion under $Q$, and the expectation is over all realizations of (\ref{sde2-3}) starting from $X_{0}=x$.
+On the other hand, $X_{t}$ under $P$ corresponds to the perturbed dynamics, which should become clear upon comparing equations (\ref{sde2-2}) and (\ref{sde2-3}).
 
@@
 \begin{equation}\label{lr-alt}
-\lim_{\eps\to 0}\frac{\bE_{\rho_{x}^{\eps}}[\varphi] + \bE_{\rho_{x}^{0}}[\varphi]}{\eps} = \bE_{\rho_{x}^{0}}\left[\varphi \int_{0}^{T}\xi_{s}\cdot dB_{s} \right]\,,
+\lim_{\eps\to 0}\frac{\bE_{P_{x}^{\eps}}[\varphi] + \bE_{P_{x}^{0}}[\varphi]}{\eps} = \bE_{P_{x}^{0}}\left[\varphi \int_{0}^{T}\xi_{s}\cdot dB_{s} \right]\,,
 \end{equation}
-where, in accordance with the previous linear response formula, $\rho_{x}^{\eps}$ and $\rho_{x}^{0}$ denote the path probability distributions of perturbed and unperturbed processes (\ref{sde2-2}) and (\ref{sde2-1}) with deterministic initial condition.
-On the right hand side of the last equation $\varphi$ is understood as a functional of $(x_{t})_{0\le t\le T}$, the solution to (\ref{sde2-1}) with initial condition $x_{0}=x$, and $\xi_{t}$ is the solution to $a(x_{t})\xi_{t}=v_{t}$.
+where, in analogy with the previous linear response formula, $P_{x}^{\eps}$ and $P_{x}^{0}$ denote the path probability measures of perturbed and unperturbed paths---i.e.~the trajectories obtained from (\ref{sde2-2}) and (\ref{sde2-1})---with deterministic initial condition.
+The path measure $P^{\eps}_{x}$ should not be confused with the phase space density $\rho^{\eps}(x,t)$ in Section \ref{ssec:equiLRT} that solves the Fokker Planck equation.
+Note moreover that $\varphi$ on the right hand side of (\ref{lr-alt}) is understood as a functional of $(x_{t})_{0\le t\le T}$, i.e.~the solution to (\ref{sde2-1}) with initial condition $x_{0}=x$, with $\xi_{t}$ being the solution to $a(x_{t})\xi_{t}=v_{t}$.
+
 By averaging the the initial values $x$ on both sides of (\ref{lr-alt}) over any given initial distribution, one obtains an analogous formula for distributed initial conditions.
-
 Note that $\varphi$ and $B_{t}$ are not independent, hence the product of $\varphi$ and the integral over the Brownian motion does not average to zero in general.
@@
 \begin{aligned}
-\bE_{\rho_{x}^{\eps}}[f] & \approx \bE_{\rho_{x}^{0}}[\varphi]  + \eps \bE_{\rho_{x}^{0}}\left[\varphi\int_{0}^{T}\xi_{s}\cdot dB_{s} \right]\\  
-& + \frac{\eps^{2}}{2} \bE_{\rho_{x}^{0}}\left[\varphi\left\{\left(\int_{0}^{T}\xi_{s}\cdot dW_{s}\right)^2 -\int_0^T |\xi_s|^2\,ds\right\} \right].
+\bE_{P_{x}^{\eps}}[f] & \approx \bE_{P_{x}^{0}}[\varphi]  + \eps \bE_{P_{x}^{0}}\left[\varphi\int_{0}^{T}\xi_{s}\cdot dB_{s} \right]\\  
+& + \frac{\eps^{2}}{2} \bE_{P_{x}^{0}}\left[\varphi\left\{\left(\int_{0}^{T}\xi_{s}\cdot dW_{s}\right)^2 -\int_0^T |\xi_s|^2\,ds\right\} \right].
 \end{aligned}
@@
 \begin{equation}\label{eqn:langevin-1b}
-b(x,t) = (J-R)\nabla H(x) +  B(x) u(t)\,,\quad a(x) = \sqrt{2\beta^{-1}}R^{1/2}
+b(x,t) = (J-R)\nabla H(x) +  c(x) u(t)\,,\quad a(x) = \sqrt{2\beta^{-1}}R^{1/2}
 \end{equation}
-with the definitions of $J,R$ and $H$ unchanged, $u\in\R^{n}$ being some control variable, and $B(\cdot)\in\R^{2n\times n}$ given by 
+with the definitions of $J,R$ and $H$ unchanged, $u\in\R^{n}$ being some bounded measurable control, and $c(\cdot)\in\R^{2n\times n}$ given by 
 \[
-B(x) = \left( \begin{array}{c}
+c(x) = \left( \begin{array}{c}
 0 \\ D(x) \end{array}\right)
 \]
-The matrix $D(\cdot)\in\R^{n\times n}$ must be chosen such that $B$ satisfies the Fredholm alternative $\mathrm{range}(B(\cdot))\perp \mathrm{ker}((a(\cdot))^{T})$, where $\mathrm{ker}(a^T)$ denotes the kernel of $a^T$.
+The matrix $D(\cdot)\in\R^{n\times n}$ must be chosen such that $c$ satisfies the Fredholm alternative $\mathrm{range}(c(\cdot))\perp \mathrm{ker}((a(\cdot))^{T})$, where $\mathrm{ker}(a^T)$ denotes the kernel of $a^T$.
 
@@
 \begin{equation}\label{eqn:langevin-1c}
-v_{t} = B(X_{t})\delta u(t)
+v_{t} = c(X_{t})\delta u(t)
 \end{equation}
@@
 \[
-\sqrt{2\beta^{-1}}R^{1/2} \xi_{t} = D(X_{t})\,\delta u_{t}\,.
+\sqrt{2\beta^{-1}}R^{1/2} \xi_{t} = D(X_{t})\,\delta u(t)\,.
 \]
@@
 
-To be more explicit, the equation we are considering has the form
+To be more specific, the equation we are considering has the form
 \begin{equation}\label{eqn:langevin-1d}
-dX_t =  (J-R)\nabla H(X_{t})dt + B(X_t)(u(t)+\eps \delta u(t))dt+a\, dB_t\,.
+dX_t =  (J-R)\nabla H(X_{t})dt + c(X_t)(u(t)+\eps \delta u(t))dt+a\, dB_t\,.
 \end{equation}
@@
 \begin{equation}\label{lr-alt2}
-\lim_{\eps\to 0}\frac{\bE_{\rho_{x}^{\eps}}[\varphi] - \bE_{\rho_{x}^{0}}[\varphi]}{\eps} =  \bE_{\rho_{x}^{0}}\left[\varphi\int_{0}^{T}(\sigma^{-1}D(x_{s})\,\delta u_{s})\cdot dB_{s} \right]\,
+\lim_{\eps\to 0}\frac{\bE_{P_{x}^{\eps}}[\varphi] - \bE_{P_{x}^{0}}[\varphi]}{\eps} =  \bE_{P_{x}^{0}}\left[\varphi\int_{0}^{T}(\sigma^{-1}D(x_{s})\,\delta u(s))\cdot dB_{s} \right]\,
 \end{equation}
@@
 \begin{equation}\label{eqn:neq-response}
-\bE_{\rho_{x}^{\eps}}[\varphi] \approx \bE_{\rho_{x}^0}[\varphi] + \eps\bE_{\rho_{x}^0}\left[\varphi\int_{0}^{T}(\sigma^{-1}D(x_{s})\,\delta u_{s})\cdot dB_{s} \right]\,.
+\bE_{P_{x}^{\eps}}[\varphi] \approx \bE_{P_{x}^0}[\varphi] + \eps\bE_{P_{x}^0}\left[\varphi\int_{0}^{T}(\sigma^{-1}D(x_{s})\,\delta u(s))\cdot dB_{s} \right]\,.
 \end{equation}
@@
 
-As one can easily get lost in the various transformations and changes of measure, involving $\nu$, $\mu$, $\rho_{x}^{\eps}$, $\rho^{0}_{x}$ etc.
-it may be helpful to understand how the linear response formulas (\ref{lr-alt}) or (\ref{lr-alt2}) can be realized algorithmically.
+As one can easily get lost in the various transformations, involving path measures $P$, $Q$, $P_{x}^{\eps}$, $P^{0}_{x}$ or phase space densities $\rho^{\eps}$, $\rho^{0}$, $\rho_{\infty}$, it may be helpful to understand how the linear response formulas (\ref{lr-alt}) or (\ref{lr-alt2}) can be realized algorithmically.
 To this end, let $(x_0,\,x_{1},\,x_{2},\,x_{3},\ldots)$ with $x_{k}$ be the numerical realization of (\ref{sde2-1}).
@@
 \begin{equation}\label{lr-est}
-\hat{R} = \frac{1}{M}\sum_{i=1}^{M}\left\{f(\{x_{k}(\omega_{i})\}_{0\le k\le N})\sum_{j=0}^{N-1}\hat{\xi}_{i}(\omega_{i})\cdot \eta_{j+1}(\omega_{i})\right\}
+\hat{R} = \frac{1}{M}\sum_{i=1}^{M}\left\{\varphi(\{x_{k}(\omega_{i})\}_{0\le k\le N})\sum_{j=0}^{N-1}\hat{\xi}_{i}(\omega_{i})\cdot \eta_{j+1}(\omega_{i})\right\}
 \end{equation}
@@
   have the same initial distribution.
-For fixed initial values (i.e.~points) this assumption cannot be relaxed (because otherwise $d\nu/d\mu$ does not exist).
+For fixed initial values (i.e.~points) this assumption cannot be relaxed (because otherwise $dP/dQ$ and $dQ/dP$ do not exist).
 For distributed initial values, however, there is no problem for the unperturbed and perturbed dynamics to have different initial distribution as long as both distributions are strictly positive almost everywhere.
@@
 
-\item[(iii)] If one wants to calculate the same expectation value for a family of nonequilibrium perturbations $\delta u_t$ then 
+\item[(iii)] If one wants to calculate the same expectation value for a family of nonequilibrium perturbations $\delta u$ then 
 one does not need to repeat the calculations of \eqref{eqn:neq-response} for every member of the family. 
   If it is possible to
-  express different $\delta u_t$ in the same basis, then the responses must only be
+  express different $\delta u$ in the same basis, then the responses must only be
   calculated for the single basis functions. Then with a linear combination of the responses on basis functions, one can derive
@@
 The nonequilibrium linear response formula can be used for solving certain optimal control problems.
-To this end, let us remain in the setting of equations (\ref{eqn:langevin-1d})--(\ref{eqn:neq-response}) and assume that we are interested choosing the nonequilibrium forcing $u$, such that the expected value $\bE[f]$ of some utility function
+To this end, let us remain in the setting of equations (\ref{eqn:langevin-1d})--(\ref{eqn:neq-response}) and assume that we are interested choosing the nonequilibrium forcing $u$, such that the expected value $I=\bE[\psi]$ of some utility function
 \[
-f(u) = \int_{0}^{T}\left\{\ell (q_{s}) - \frac{1}{2}|u_{s}|^{2}\right\}\,dt + L(q_{T}) 
+\psi(u) = \int_{0}^{T}\left\{f(x_{s}) - \frac{1}{2}|u(s)|^{2}\right\}\,dt + g(x_{T}) 
 \]
-is maximized where $q_{t}$ is the solution to the controlled Langevin equation (\ref{eqn:langevin-1d}) for $\eps=0$; the functions $\ell$ and $L$ are the running cost and the terminal cost, which are assumed to be continuous and bounded from above; without loss of generality, $\ell$ and $L$ are assumed to depend only on the positions.
-The quadratic term is a penalization that makes sure that the controls do not go through the roof \cite{schuette2012,hartmann2012}.
+is maximized where $x_{t}$ is the solution to the controlled Langevin equation (\ref{eqn:langevin-1d}) for $\eps=0$; the functions $f$ and $g$ are the running cost and the terminal cost, which are assumed to be continuous and bounded from above; without loss of generality, $f$ and $g$ are assumed to depend only on the positions $q:=x_{1}$ and not on the momenta $p:=x_{2}$.
+The quadratic term is a penalization that makes sure that the optimal control remains bounded \cite{schuette2012,hartmann2012}.
 Let us moreover assume that the controls are open-loop (i.e.
@@
 \[
-u_{t} = \sum_{k=1}^K a_k \Phi_k(t)\,,\quad a_{k}\in \R\,,
+u(t) = \sum_{k=1}^K a_k \Phi_k(t)\,,\quad a_{k}\in \R\,,
 \]
@@
 %In principle we could even go for closed-loop (i.e.~feedback) controls by making $\Phi_{k}$ state dependent.
-We want to solve the optimal control problem
+Specifically, we want to solve the optimal control problem of the following standard form: 
 \begin{equation}\label{ocp1}
@@
 dq_t & =  \nabla_p H(q_t,p_t)dt\\
-dp_t & =  -\nabla_q H(q_t,p_t)dt - \gamma \nabla_p H(q_t,p_t)dt + D(q_t)u_{t} dt+\sigma dB_t\,.
+dp_t & =  -\nabla_q H(q_t,p_t)dt - \gamma \nabla_p H(q_t,p_t)dt + D(q_t)u(t) dt+\sigma dB_t\,.
 \end{aligned}
@@
 \begin{equation}\label{ocp2}
-I(u) = \bE_{\rho^0}\left[\int_{0}^{T}\left\{\ell (q_{s}) - \frac{1}{2}|u_{s}|^{2}\right\}\,dt + L(q_{T}) \right]
+I(u) = \bE\left[\int_{0}^{T}\left\{f(q_{s}) - \frac{1}{2}|u(s)|^{2}\right\}\,dt + g(q_{T}) \right]\,,
 \end{equation}
-and $\cU$, the space of admissible controls, consists of all bounded controls 
+with the expectation being over all realizations of the controlled Langevin dynamics in (\ref{ocp1}) with any prescribed initial distribution.
+Here, the space of admissible controls $\cU$ consists of all bounded controls of the form
 \begin{equation}\label{admissible}
-u_{t} = \sum_{k=1}^K a_k \Phi_k(t)\,,\quad a_{k}\in \R\,,
+u(t) = \sum_{k=1}^K a_k \Phi_k(t)\,,\quad a_{k}\in \R\,,
 \end{equation}
@@
 
-In principle optimal control problems such as (\ref{ocp1})--(\ref{ocp2}) can be solved by dynamic programming, i,e., by solving the corresponding Hamilton-Jacobi-Bellman PDE \cite{fleming2006}.
+In principle, optimal control problems such as (\ref{ocp1})--(\ref{ocp2}) can be solved by dynamic programming, i.e.~by solving the corresponding Hamilton-Jacobi-Bellman PDE \cite{fleming2006}.
 Expect for very simple, essentially one-dimensional systems, solving Hamilton-Jacobi-Bellman equations is not an easy task, so we pursue a different strategy here.
@@
 \end{equation*} 
+For existence of the functional derivative \wrt general It\^o stochastic processes, we refer to \cite{bell2006}.  
 Now the idea is to use that we have restricted our space of admissible controls to functions of the form (\ref{admissible}) with given basis vector fields and do a gradient search in the unknown coefficients $a_{k}$.
@@
 & = \left(\lim_{\eps\to 0}\frac{I(u+\eps\delta u) - I(u)}{\eps}\right)\cdot \delta u\\
-& = \left(\lim_{\eps\to 0}\frac{\bE_{\rho^{\eps}}[f(u+\eps \delta u)] - \bE_{\rho^{0}}[f(u)]}{\eps}\right)\cdot\delta u\\
+& = \left(\lim_{\eps\to 0}\frac{\bE_{P^{\eps}}[\psi(u+\eps \delta u)] - \bE_{P^{0}}[\psi(u)]}{\eps}\right)\cdot\delta u\\
 \end{aligned}
@@
 \begin{equation}\label{dida}
-\frac{\partial I}{\partial a_{k}} = \bE_{\rho^0}\left[f\int_{0}^{T}(\sigma^{-1}D(q_{s})\Phi_{k}(s))\cdot dB_{s} + \int_{0}^{T}u_{s}\cdot\Phi_{k}(s)\,ds \right] \,
+\frac{\partial I}{\partial a_{k}} = \bE_{P^0}\left[\psi\int_{0}^{T}(\sigma^{-1}D(q_{s})\Phi_{k}(s))\cdot dB_{s} + \int_{0}^{T}u(s)\cdot\Phi_{k}(s)\,ds \right] \,
 \end{equation}
-and $\bE_{\rho^0}[\cdot]$ being the expectation over all realization of (\ref{eqn:langevin-1d}) with $\eps=0$ and $u=u^{(n)}$ being the current iterate; note that we can compute all partial derivatives $\partial I/\partial a_k$, $k=1,\ldots,K$ from just one ensemble of nonequilibrium paths of $(q_{t},p_{t})$.
-Further note that, in general, the updated coefficient $a^{(n+1)}$ and hence the updated control $u^{(n+1)}$ will depend on the distribution of the initial conditions at the $n$-th iteration stage; in particular, if the controls are computed on-the-fly, the controls will actually be feedback controls depending on the current state.
+and $\bE_{P^0}[\cdot]$ being the expectation over all realizations of (\ref{eqn:langevin-1d}) with $\eps=0$ and $u=u^{(n)}$ being the current iterate and arbitrary initial conditions $x=(q,p)$, e.g., 
+\[
+\bE_{P^{0}}[\psi] = \int_{\R^{2n}}\bE_{P^{0}_{x}}[\psi]\,d\mu_{\infty}(x) \,,
+\]
+with $P_{x}^{0}$ denoting the corresponding path measure with deterministic initial condition.
+Note that we can compute all partial derivatives $\partial I/\partial a_k$, $k=1,\ldots,K$ from just one ensemble of nonequilibrium paths of $(q_{t},p_{t})$.
+Further note that, in general, the updated coefficient $a^{(n+1)}$ and hence the updated control $u^{(n+1)}$ will depend on the distribution of the initial conditions at the $n$-th iteration stage; in particular, if the controls are computed on-the-fly, the controls will actually be feedback controls depending on the current state and the time elapsed since $t=0$.
 
@@
 We use the idea of nonequilibrium linear response to investigate the
-nonequilibrium phase space probability density distribution, denoted by $\rho^\eps(q, p, t)$, of a one-dimensional model system: one particle in a
+nonequilibrium phase space probability density, denoted by $\rho^\eps(q, p, t)$, of a one-dimensional model system: one particle in a
 splitting single-well potential as shown in Fig.~\ref{fig:tmp4}.  For convenience, we let the mass of
@@
   We use Euler scheme with time step of $10^{-4}$ to discretize the
-  Langevin equation. The inital configurations for nonequilibrium
+  Langevin equation. The initial configurations for nonequilibrium
   simulation are generated by taking configurations
   from an equilibrium simulation of $10^6$~\textsf{ps} at an 1~\textsf{ps} time interval.
-  Then starting with these inital configurations,  the nonequilibrium
+  Then starting with these initial configurations according to $\rho_{\infty}\propto\exp(-\beta H)$,  the nonequilibrium
   Langevin equation is integrated until 20~\textsf{ps} with the same numerical
@@
 should be compared. The second row shows the result of the
-traditional equilibrium linear response theory.  Please notice that in
-this case, since the reference simulation is in equilibrium, we set the perturbation to
+traditional equilibrium linear response theory. In
+this case, the reference simulation is in equilibrium and we have set the perturbation to
 \[
-\eps v_t = D(q) (u_t+\eps\delta u_t)=2\eps D(q)k_e\cdot t/T=2\eps D(q)\delta u_t,
+\eps v_t = D(q_{t}) (u(t)+\eps\delta u(t))=2\eps D(q_{t})k_e\cdot t/T=:2\eps D(q_{t})\delta u(t),
 \] 
@@
 \begin{align}
-  U( q) = \frac12 k ( q^2 - a^2)^2
+  U(q) = \frac12 k ( q^2 - a^2)^2
 \end{align}
@@
 the following optimal forcing problem:
-\begin{align}
-  \tilde{I} & = \max_{u_t \in\mathcal F} \bE_{\rho^0}\left[ -\int_{0}^{T} \frac{1}{2}|u_{s}|^{2}\,dt + L(q_T) \right],
-\end{align}
-with $L(q_T)=\chi_{[a-\delta,a+\delta]}(q_T)/\eta$ representing the probability of the end point of the trajectory, $ q_T$, being in the right well ($\chi_I$ denotes the indicator function of the interval $I$) with $\eta$ being a weighting constant, ${\mathcal F}$ denoting the space of function that are piecewise linear  on $[0, T]$ in uniform intervals of length 1 ps, and $\rho^0$ being the
-\recheck{initial nonequilibrium distribution $\rightarrow$ nonequilibrium distribution under control $u_t$}.
-With $P(T) = \eta \bE_{\rho^0}(L(q_T))$, the probability of ending up in the right well at time $T$, and $I=\eta \tilde{I}$ we thus have
 \[
-  I = \max_{u \in\mathcal F} F(u),\quad F(u)= P(T) -
-  \frac{\eta}{2}\,
-   \int_0^T |u_t|^2\, dt\,.
+ \max_{u\in{\mathcal U}_{L}}I(u)\,,\quad I(u) = \bE\left[ -\int_{0}^{T} \frac{1}{2}|u(s)|^{2}\,dt + g(q_T) \right],
 \]
-   \recheck{shall we use $I(u)$ instead of $F(u)$?, so that the
-     notation is consistent with  Sec. 3.}
- It is
-clear that for an unbiased double-well, $P(T)$
-is 0.5 if we choose $\rho^0=\rho_0$ as initial distribution.  The integral is the ``cost'' of the control and $\eta$  indicates the relative magnitude of the
+with $g(q_T)=\chi_{[a-\delta,a+\delta]}(q_T)/\eta$ representing the probability of the end point of the trajectory, $ q_T$, being in the right well ($\chi_I$ denotes the indicator function of the interval $I$) with $\eta$ being a weighting constant, ${\mathcal U}_{L}$ denoting the space of function that are piecewise linear  on $[0, T]$ in uniform intervals of length 1 \textsf{ps}.
+Then
+\[
+\bP[q_{T}\in[a-\delta,a+\delta] ] = \eta \bE[g(q_T)]
+\]
+is the probability of being in the right well at time $T$. It is therefore convenient to scale the cost functional according to $I\mapsto\eta I$ and redefine $I$ as follows: 
+\[
+ I(u) = \bP[q_{T}\in[a-\delta,a+\delta] ] - \frac{\eta}{2}\,
+   \int_0^T |u(s)|^2\, ds\,.
+   \]
+It is clear that for an unbiased double-well, the probability to be in the right well at any time $T$ is 0.5 if we choose $\rho_{\infty}$ to be the initial distribution.
+The remaining integral term in $I$ denotes the ``cost'' of the control and $\eta$  indicates the relative magnitude of this
 cost.
@@
   scheme with a time step of $10^{-3}$~\textsf{ps}.
-  The initial distribution $\rho^0(q,p,0)$ is also set to be the
-  equilibrium distribution $\rho_0$ of the unperturbed system.
+  The initial distribution is set to be the
+  equilibrium distribution $\rho_\infty\propto\exp(-\beta H)$ of the unperturbed system.
   Therefore, an equilibrium simulation of $10^6$~\textsf{ps} is firstly performed (with
@@
 Fig.~\ref{fig:tmp7} presents the numerical results of $\eta = 0.01$.
-Starting from an initial guess of linear control from $u_0 = 0$ to
-$u_T = 1$,
-the gradient search~\eqref{gd} \recheck{uses a constant $\tau_n = 30$ and} converges at the 22nd step, when the
+Starting from an initial guess of a linearly interpolated control between $u_0 = 0$ and
+$u_T = 1$, the gradient search~\eqref{gd} \recheck{uses a constant $\tau_n = 30$ and} converges at the 22nd step, when the
 maximum increment of the control coefficient $\max_k\vert \delta a_k\vert$
 is smaller than $0.02$, the termination threshold.
-The magnitude of the optimal
-control 
+The magnitude of the optimal control 
 \[
-o_t = \textrm{argmax}_{u_t \in\mathcal F} F(u_t)
+o(\cdot) = \mathop{\rm argmax}_{u \in{\mathcal U}_{L}} I(u)
 \] 
@@
 with blue insertions showing the shape of the time-dependent optimal control potential 
-$U(q) + o_t V(q)$ (optimally tilted double-well potential). 
+$U(q) + o(t) V(q)$ (optimally tilted double-well potential). 
 The maximum $I$ and the corresponding optimal probability to end up in the right well are
@@
 linear response theory developed in this work, i.e., using~\eqref{dida},
-and the green lines represent the brute force reference simulations that has been performed as follows: The optimal control from ${\mathcal F}$ is calculated by a gradient descent based optimization method in which the gradient of the functional with respect to the control is computed by numerical differentiation (central finite differences) in each step.
+and the green lines represent the brute force reference simulations that has been performed as follows: The optimal control from ${\mathcal U}_{L}$ is calculated by a gradient descent based optimization method in which the gradient of the functional with respect to the control is computed by numerical differentiation (central finite differences) in each step.
 % : \redc{$\partial I / \partial a_k \approx (I^+ - I^-) / (a_k^+ - a_k^-)$}
@@
 the linear response theory computes the gradient correctly.
-Please note that in order to calculate the gradient by the finite difference scheme, one
+Note that in order to calculate the gradient by the finite difference scheme, one
 needs to do $2K$ nonequilibrium simulations
-(where $K$ is the dimension of ${\mathcal F}$; here $K = 10$). In contrast, the nonequilibrium response theory only needs one nonequilibrium simulation.
+(where $K$ is the dimension of ${\mathcal U}_{L}$; here $K = 10$). In contrast, the nonequilibrium response theory only needs one nonequilibrium simulation.
 
@@
 problem in the nonequilibrium setting.
-The fact that the optimal control is decreasing at the end of the interval is understandable since the population in the right well needs time to be build and increasing the control till the very end would be a waste of energy without corresponding gain in population.
+The fact that the optimal control is decreasing at the end of the interval is understandable since the population in the right well needs time to be build and increasing the control until $t=T$ would be a waste of control without leading to a significant population gain.
 

### 2013-08-22 writedown/linear.response/paper.tex
new revise by han, send to C and C.

@@
 
-Standard molecular dynamics simulations are dealing with systems in thermal equilibrium; in this case they are tuned to the canonial or Boltzmann distribution in the sense that either (1) if one starts from this distribution it remains invariant under the dynamics or (2) if one generates a very long trajectory it samples state space with respect this distribution, that is, every possible state of the molecular system under consideration is visited according to the probability given by it.
+Standard molecular dynamics simulations are dealing with systems in thermal equilibrium; in this case they are tuned to the canonical or Boltzmann distribution in the sense that either (1) if one starts from this distribution it remains invariant under the dynamics or (2) if one generates a very long trajectory it samples state space with respect this distribution, that is, every possible state of the molecular system under consideration is visited according to the probability given by it.
 Obviously, this allows to compute equilibrium expectation values with respect to the  canonical distribution simply be computing long trajectories.
@@
 In some sense it has become one of the cornerstones of modern statistical physics since it can be related to the fluctuation dissipation theorem (FDT) which roughly states that for appropriate systems in statistical equilibrium, the average response to small external perturbations can be calculated through the knowledge of suitable correlation functions of the unperturbed statistical system.
-Linear response formulae on the infinite time-horizon have been shown to hold under rather mild assumptions even for nonequilibrium systems that possess a unique invariant measure \cite{hairer2008}.
+Linear response formulas on the infinite time-horizon have been shown to hold under rather mild assumptions even for nonequilibrium systems that possess a unique invariant measure \cite{hairer2008}.
 
@@
 
-The outline of the article is as follows: First, we will review the derivation of finite time-horizon linear response formulae for general diffusion processes and diffusions of Langevin type.
+The outline of the article is as follows: First, we will review the derivation of finite time-horizon linear response formulas for general diffusion processes and diffusions of Langevin type.
 Next, we will show how to derive first and second order response formulas for genuine nonequilibrium systems and how to apply these formula to the computation of optimal controls.
@@
 \]
-Suppose that there exists an auxiliary stochastic process $\xi_{t}\in\R^{m}$ such that 
+Suppose that there exists an auxiliary stochastic process \redc{ $\xi_{t}\in\R^{n}$} such that 
 \begin{equation}\label{xi}
@@
 \begin{equation}\label{lr-alt}
-\lim_{\eps\to 0}\frac{\bE_{P_{x}^{\eps}}[\varphi] + \bE_{P_{x}^{0}}[\varphi]}{\eps} = \bE_{P_{x}^{0}}\left[\varphi \int_{0}^{T}\xi_{s}\cdot dB_{s} \right]\,,
+\lim_{\eps\to 0}\frac{\bE_{P_{x}^{\eps}}[\varphi] - \bE_{P_{x}^{0}}[\varphi]}{\eps} = \bE_{P_{x}^{0}}\left[\varphi \int_{0}^{T}\xi_{s}\cdot dB_{s} \right]\,,
 \end{equation}
@@
 \[
-\sqrt{2\beta^{-1}}R^{1/2} \xi_{t} = D(X_{t})\,\delta u(t)\,.
+\sqrt{2\beta^{-1}}\recheck{\gamma}^{1/2} \xi_{t} = D(X_{t})\,\delta u(t)\,.
 \]
-Hence the equation for $\xi_{t}$ is solvable if and only if $\mathrm{range}(D(\cdot))\perp \mathrm{ker}(R^{1/2})$, which, since $R$ has full rank, means that $D(\cdot)$ must be invertible almost everywhere.
+\recheck{Hence the equation for $\xi_{t}$ is solvable if and only if $\mathrm{range}(D(\cdot))\perp \mathrm{ker}(\gamma^{1/2})$, which, since $\gamma$ has full rank, means no restriction on $D(\cdot)$.}
+  % means that $D(\cdot)$ must be invertible almost everywhere.}
 
@@
 The nonequilibrium linear response formula can be used for solving certain optimal control problems.
-To this end, let us remain in the setting of equations (\ref{eqn:langevin-1d})--(\ref{eqn:neq-response}) and assume that we are interested choosing the nonequilibrium forcing $u$, such that the expected value $I=\bE[\psi]$ of some utility function
+To this end, let us remain in the setting of equations (\ref{eqn:langevin-1d})--(\ref{eqn:neq-response}) and assume that we are interested in choosing the nonequilibrium forcing $u$, such that the expected value $I=\bE[\psi]$ of some utility function
 \[
@@
 dq_t & =  \nabla_p H(q_t,p_t)dt\\
-dp_t & =  -\nabla_q H(q_t,p_t)dt - \gamma \nabla_p H(q_t,p_t)dt + D(q_t)u(t) dt+\sigma dB_t\,.
+dp_t & =  -\nabla_q H(q_t,p_t)dt - \gamma \nabla_p H(q_t,p_t)dt + D(q_t)u(t) dt+\sigma dB_t\,,
 \end{aligned}
 \end{equation}
-where 
+\redc{with $(q_0, p_0)$ being subject to a given initial distribution $\mu$, and}
 \begin{equation}\label{ocp2}
@@
 \end{equation}
-with the expectation being over all realizations of the controlled Langevin dynamics in (\ref{ocp1}) with any prescribed initial distribution.
+with the expectation being over all realizations of the controlled Langevin dynamics in (\ref{ocp1}) with \recheck{the} prescribed initial distribution.
 Here, the space of admissible controls $\cU$ consists of all bounded controls of the form
@@
 In principle, optimal control problems such as (\ref{ocp1})--(\ref{ocp2}) can be solved by dynamic programming, i.e.~by solving the corresponding Hamilton-Jacobi-Bellman PDE \cite{fleming2006}.
-Expect for very simple, essentially one-dimensional systems, solving Hamilton-Jacobi-Bellman equations is not an easy task, so we pursue a different strategy here.
-The idea is to use that we have restricted our space of admissible controls to functions of the form (\ref{admissible}) with given basis vector fields and that we can a gradient search in the unknown coefficients $a_{k}$, using the iteration
+Except for very simple, essentially one-dimensional systems, solving Hamilton-Jacobi-Bellman equations is not an easy task, so we pursue a different strategy here.
+The idea is to use that we have restricted our space of admissible controls to functions of the form (\ref{admissible}) with given basis vector fields and that we can do a gradient search in the unknown coefficients $a_{k}$, using the iteration
 \[
@@
 \begin{equation*}
-\frac{\delta I}{\delta u} = \left.\frac{d}{d\eps}\right|_{\eps=0} I(u+\eps v) = \langle\nabla I(u),v\rangle\,,
+\frac{\delta I}{\delta u} = \left.\frac{d}{d\eps}\right|_{\eps=0} I(u+\eps v) = \langle\nabla I(u),v\rangle\,.
 \end{equation*} 
@@
 \begin{equation}\label{dida}
-\frac{\partial I}{\partial a_{k}} = \bE_{P^0}\left[\psi\int_{0}^{T}(\sigma^{-1}D(q_{s})\Phi_{k}(s))\cdot dB_{s} + \int_{0}^{T}u(s)\cdot\Phi_{k}(s)\,ds \right] \,
+\frac{\partial I}{\partial a_{k}} = \bE_{P^0}\left[\psi\int_{0}^{T}(\sigma^{-1}D(q_{s})\Phi_{k}(s))\cdot dB_{s} + \int_{0}^{T}u(s)\cdot\Phi_{k}(s)\,ds \right] \,,
 \end{equation}
@@
 \[
-\bE_{P^{0}}[\psi] = \int_{\R^{2n}}\bE_{P^{0}_{x}}[\psi]\,d\mu_{\infty}(x) \,,
+\bE_{P^{0}}[\psi] = \int_{\R^{2n}}\bE_{P^{0}_{x}}[\psi]\,d\recheck{\mu}(x) \,,
 \]
-with $P_{x}^{0}$ denoting the corresponding path measure with deterministic initial condition.
+with $P_{x}^{0}$ denoting the corresponding path measure with deterministic initial condition,
+\redc{and $\mu$ is the prescribed initial distribution.}
 Note that we can compute all partial derivatives $\partial I/\partial a_k$, $k=1,\ldots,K$ from just one ensemble of nonequilibrium paths of $(q_{t},p_{t})$.
@@
   \centering
-  \includegraphics[width=0.4\textwidth]{figs/fig-split-pot.eps}
+  % \includegraphics[width=0.4\textwidth]{figs/fig-split-pot.eps}
+  \includegraphics[width=0.5\textwidth]{figs/fig-split-pot.eps}
   \caption{The single-well potential with  splitting driving
@@
 \begin{align}
-  V( q) = \frac{1}{\sqrt{2\pi \sigma^2}}
-  \exp\Big\{-\frac{ q^2}{2\sigma^2}\Big\}
+  V( q) = \frac{1}{\sqrt{2\pi \Sigma^2}}
+  \exp\Big\{-\frac{ q^2}{2\Sigma^2}\Big\}
 \end{align}
-we use $\sigma = 0.16$~\textsf{nm}.  The strength of nonequilibrium
+we use $\Sigma = 0.16$~\textsf{nm}.  The strength of nonequilibrium
 driving $u(t)$ is set to be linearly growing, i.e. $ u(t) = k_e\cdot t/T$,
@@
 consider the perturbation to the system given by $\eps\delta u(t) = \eps k_e \cdot t/T$.
-We consider the following parameters: end time $T = 20$~\textsf{ps}, 
+We use the following parameters: end time $T = 20$~\textsf{ps}, 
 $k_e = 1$ and $\eps = 1$, see Fig.~\ref{fig:tmp4}
@@
 be equilibrium distribution of the unperturbed system.
-\recheck{
   We use Euler scheme with time step of $10^{-4}$ to discretize the
@@
   scheme and time step.
-}
 
@@
 splits into two equally sized peaks.  In the end, the two
-peaks are entirely seperated.  The brute force nonequilibrium
+peaks are entirely separated.  The brute force nonequilibrium
 simulation serves as the precise result to which the response theory
@@
 so that the effective perturbation is of strength $2\eps = 2$.  At $t \leq
-15$~\textsf{ps}, the accuracy of the equilibrium linear response is
+10$~\textsf{ps}, the accuracy of the equilibrium linear response is
 perfect. At $t =
@@
 The poor accuracy is due to the fact that the strength of perturbation is no
-longer small so that the preliminary assumption of the classical reponse theory ("small perturbation")
+longer small so that the preliminary assumption of the classical response theory ("small perturbation")
 is not satisfied.
 
-The third row of Fig.~\ref{fig:tmp6} presents the results computed using the novel nonequilibrium  linear response formula: we first start form the equilibrium distribution, apply  $u(t)$, arrive at a nonequilibrium distribution and then, in a second step, compute the effect of the nonequilibrium driving $\delta u(t)$.
+The third row of Fig.~\ref{fig:tmp6} presents the results computed using the novel nonequilibrium  linear response formula: we first start from the equilibrium distribution, apply  $u(t)$, arrive at a nonequilibrium distribution and then, in a second step, compute the effect of the nonequilibrium driving $\delta u(t)$.
 The numerical results are satisfactorily consistent with
@@
 \end{align}
-with $V( q) = - q$. We want to
+with $V( q) = - q$. {Starting from a fully equilibriated system},
+we want to
 optimally design the tilting such that the probability
@@
 \[
- \max_{u\in{\mathcal U}_{L}}I(u)\,,\quad I(u) = \bE\left[ -\int_{0}^{T} \frac{1}{2}|u(s)|^{2}\,dt + g(q_T) \right],
+ I = \max_{u\in{\mathcal U}_{L}}I(u)\,,\quad I(u) = \bE\left[ -\int_{0}^{T} \frac{1}{2}|u(s)|^{2}\,dt + g(q_T) \right],
 \]
@@
 
-\recheck{ In all our numerical tests for the optimal control problem, the
-  Langevin dyanmics Eq.~\eqref{ocp1} is discretized by the Euler
+In our numerical tests for the optimal control problem, the
+  Langevin dynamics Eq.~\eqref{ocp1} is discretized by the Euler
   scheme with a time step of $10^{-3}$~\textsf{ps}.
@@
   configurations) along the trajectory.
-  Then by using these configurations as inital configurations,
+  Then by using these configurations as initial configurations,
   $10^5$ nonequilibrium
@@
   % $\tau_n = 30$.
-}
 
-Fig.~\ref{fig:tmp7} presents the numerical results of $\eta = 0.01$.
+Fig.~\ref{fig:tmp7} presents the numerical results for $\eta = 0.01$.
 Starting from an initial guess of a linearly interpolated control between $u_0 = 0$ and
-$u_T = 1$, the gradient search~\eqref{gd} \recheck{uses a constant $\tau_n = 30$ and} converges at the 22nd step, when the
-maximum increment of the control coefficient $\max_k\vert \delta a_k\vert$
+$u_T = 1$, the gradient search~\eqref{gd} {uses a constant $\tau_n = 30$ and} converges at the 22nd step, when the
+maximum increment of the control coefficient $\max_k\vert \delta a^{(n)}_k\vert$
 is smaller than $0.02$, the termination threshold.
@@
 linear response theory developed in this work, i.e., using~\eqref{dida},
-and the green lines represent the brute force reference simulations that has been performed as follows: The optimal control from ${\mathcal U}_{L}$ is calculated by a gradient descent based optimization method in which the gradient of the functional with respect to the control is computed by numerical differentiation (central finite differences) in each step.
+and the green lines represent the brute force reference simulations that has been performed as follows: The optimal control from ${\mathcal U}_{L}$ is calculated by a gradient descent based optimization method in which the gradient of the functional with respect to the control is computed by numerical differentiation (central finite differences) at each step.
 % : \redc{$\partial I / \partial a_k \approx (I^+ - I^-) / (a_k^+ - a_k^-)$}
@@
 
-We derived  first and second order response formulas for molecular dynamics (driven Langevin dynamics) starting from general nonequilibrium distributions.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% the old one
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% We derived  first and second order response formulas for molecular dynamics (driven Langevin dynamics) starting from general nonequilibrium distributions.
 For the special case of the initial distribution being the equilibrium distribution of the unperturbed dynamics, the novel linear response formula simplifies to the well-known standard formula.
@@
 There, we demonstrated that the nonequilibrium linear response formula allows to extend the algorithmic use of linear response theory to significantly stronger perturbutions of the system since it permits intermediate steps based on partially propagated nonequilibrium distributions.
+
+\recheck{
+We derived linear response formulas for nonequilibrium molecular
+dynamics (driven Langevin dynamics), which is a generalization of the
+standard finite-time equilibrium response theory.  The novel response theory
+dose not rely on any assumption on the existence of the invariant
+measure nor stationary state. We validated the formula in numerical
+experiments in comparison to brute-force nonequilibrium
+simulations. There, we demonstrated that the nonequilibrium linear
+response formula allows to extend the algorithmic use of linear
+response theory to significantly stronger perturbations of the system
+since it permits intermediate drivings, with which the effective perturbations
+still lie in the linear range.
+}
 

### 2013-08-22 writedown/linear.response/reply/reply.tex
new revise by han, send to C and C.

\documentclass[]{revtex4-1}

\usepackage[fleqn]{amsmath}
\usepackage{amssymb,amsthm}

\newcommand{\fwg}{{\mathcal A}}

\begin{document}

\subsection*{The scientific issues from referee 1}

\emph{
The abstract and the general tones of the introduction and conclusion do not make justice to
the content of the paper. Whether the approach proposed by the authors generalizes standard
linear response theory as described in Section 2.1 is (at best) unclear to me. In fact, I think
that most of the material from Section 2.1 could be omitted. If the authors decide to keep the
material, they should more precisely compare (4) and (10), and show how (10) generalizes (4).
}

\emph{
One trouble is that I do not understand (4). As far as I understand, the equation before (4)
is (denoting by $\fwg$ the generator)
\begin{align*}
  \textbf{E}_{\rho^\varepsilon}[f] = \int f\rho^\varepsilon = \int e^{t\fwg} f\rho_0 = \mathbb{E}_{\rho_0}(f(X_t))
\end{align*}
where the latter expectation is over all initial conditions $X_0\sim\rho_0$ and over all realizations of
the dynamics (1). Is $f$ a functional of the trajectory or a given static observable? Is $t$ fixed
or is it taken to the limit $t \rightarrow+\infty$? Is the expectation $\textbf{E}_{\rho^\varepsilon}$
an expectation over all initial
conditions and all realizations of the dynamics? Assuming that the limit $t \rightarrow\infty$ is taken,
\begin{align*}
  \textbf{E}_{\rho^\varepsilon}[f] = \int f\mu_\varepsilon
\end{align*}
where $\mu_\varepsilon = \lim_{t\rightarrow+\infty}\rho^{\varepsilon}(t)$ is the (assumed) unique invariant measure of the dynamics. To
obtain (4), a linear response result on $\mu_{\varepsilon}$ is needed, as given by (3.19) in Ref. [5]. I would however rather write $\textbf{E}_{\rho_0}$ on the right hand side of (4), and emphasize that this is a static
average (no time dependence; compare with (10)). The authors should also make clear that
$\rho_0 (t) = e^{t\fwg^\ast} = \rho_0$ is in fact independent of $t$. Unless the authors mean something else with
$\rho_0$? (see the last equation of Section 2.1).
}

\emph{
I am also not very satisfied by the discussion at the beginning of Section 2.2. It seems
that the authors say that they can tackle cases uncovered by standard linear response theory
-- namely degeneracy of the invariant measure, absence of spectral gap, etc. But again, their
approach is a finite time perturbation theory, so I do not see why the discussion on the
limitation of standard perturbation theory (on invariant measures) is relevant.
}\\

We have rewritten Sec. 2.1, and restricted the discussion of the
equilibrium response theory (3) only in the finite-time
horizon, so that it is comparable to the the nonequilibrium response theory (10), which is also
developed in the finite-time horizon. Our work
can be treated as a
generalization of the equilibrium response theory in the sense that the
reference process is not assumed to be in equilibrium nor in stationary.
The introduction and conclusion are revised accordingly.

The notations we used for the probability measures (both for
trajectory ensemble and initial condition ensemble) were misleading.
In the updated manuscript, we adopt the following convention:
$\rho^\varepsilon = \rho^\varepsilon(x,t)$ is now used only in the equilibrium
response theory part (Sec.~2.1). It denotes the probability density of
perturbed stochastic process $X_t$ (governed by SDE (1))
being at position $x$ at time $t$,
and is also the solution to the corresponding perturbed Fokker-Planck
equation. 
$\rho^0 = \rho^0(x, t)$
is the solution to the unperturbed Fokker-Planck that corresponds
to SDE (2). $\rho_\infty$ denotes the unique invariant measure of the
unperturbed Langevin dynamics.
If the initial distribution is assumed to be
$\rho^0(x, 0) = \rho_\infty$, it is obvious
that $\rho^0(x, t) = \rho_\infty(x)$ for any $t\geq 0$.
In the section 2.2 of nonequilibrium response theory, we adopt the new notation
$P_x^\varepsilon$ that denotes the trajectory probability density for the
perturbed process that starts with initial condition $X_0 = x$.
Correspondingly, $P_x^0$ denotes the probability density of
unperturbed trajectories with initial condition of $x_0 = x$.
We further
comment on the fact that if one starts the nonequilibrium process 
from a certain initial distribution (rather than the above fixed point
initial condition),
the response formula (10) can
be straightforwardly extend by averaging both sides of the equation over
the initial probability measure.
% We remind the reader about the difference
% between $P_x^\varepsilon$ and $\rho^\varepsilon$ under Eq.(10).
Under this setting, the $\phi$ on the r.h.s.~of (10) is understood
as a functional of the trajectory $(x_t)_{0\leq t\leq T}$. 
\\

% With the improved manuscript, we believe the referee is now clear
% on the fact 
% that do NOT want to discuss the infinite-time scenario, 
% the degeneracy of the inital distribution, nor the existence and uniqueness
% of the staionary probability distribution in this paper.
% In our finite-time nonequilibrium response theory,
% we do NOT presume neither the equilibrium of the system,
% nor the exsistence of a invarient (nor staionary) measure.
% % In our finite time setting,
% % The nonequilibrium dynamics is fully discribed by the trajectory
% % probability density $P^\varepsilon_x$ and the initial distribution
% % (provided that these distributions are well-defined), and the
% % Girsanov transformation used to derive the response theory
% % is valid under mild restrictions.
% % We do not touch the issue of the inifite time convergence of the probability
% % distribution.
% Therefore, when compareing the finite-time  equilibrium response theory with our
% nonequilibrium response theory, we conclude that our research overcomes the
% limitations 
% is a generalization of the equilibrium version.

\emph{
  The auxiliary process $\xi_t$ is not unique. This should be emphasized. Is there an optimal
choice, allowing for instance to reduce the variance of the estimator for the linear response
(e.g. (10) or (17))?
}\\

As far as we concerned, we do not have a clear answer to this question
in the most general case.
However, in this paper, we deal with  the Langevin dynamics.
The friction
matrix is positive definite, so the auxiliary process $\xi_t$ is uniquely
solvable.
\\

\emph{
Why do the authors use a gradient method, rather than, say, a
conjugate gradient? How is the increment $\tau_n$ chosen in last equation of
page 8: fixed or optimized by a line search for instance? This is
never mentioned... Besides, I do not understand the final discussion
at the end of Section 3. I understand (22) (except for the above
remarks on who is $\rho^0$ , and whether this expectation means generating
initial conditions according to $\rho^0 (0, q, p)$ and evolving according to
some reference dynamics such as (18)). But I do not understand the
discussion on the feedback controls. In particular, how do the authors
make sure that the initial conditions are always distributed according
to the same measure? Or do the distributions of these initial
conditions convergence to some (a priori unknown) measure?
}\\

In the second numerical example (Sec.4.2), our aim is to verify the
application of nonequilibrium response theory (Eq.(22)) to the
optimization problem, but not to dig out the most efficient numerical
method for solving this optimization problem.  We prefer the
gradient descent method with a fixed increment, because firstly it
works and converges in only 22 steps of iterations. Secondly,
it is simple and easy to implement. Therefore, the gradient descent
method is eligible to illustrate the power of the response theory, and
at the same time keeps the story as simple as possible.

The confusing notations used in the old manuscript was clarified
in the updated manuscript, see the previous discussions.
The initial condition of the optimization problem is subject to
a prescribed distribution that is invariant with respect
to steps of iterations. The form of the initial distribution
depends on how the question is raised, for example, in Example 2,
we want to  calculate the optimal way of driving the
equilibrium distribution towards the right well of the potential.
Therefore, the initial distribution is naturally chosen as the
equilibrium distribution $\rho_\infty$.
At each step of iteration, we numerically solve the
equilibrium Langevin equation,
and take the configurations along the trajectory as initial conditions
for the nonequilibrium simulation.
These discussions have been added to the manuscript.
\\

\emph{ More details on the numerical simulations are needed to make
  them reproducible: which integrator is used to approximate the time
  evolution, what is the time step, what is the number of realizations
  M in (17), how are the initial conditions sampled (iid or
  subsampling of a long Langevin trajectory), etc. These elements may
  be trivial/standard, and I certainly do not want the authors to
  write more than a short paragraph about it; but I certainly want to
  read this paragraph. Moreover, on page 13, the optimization
  procedure may be documented more precisely (number of realizations,
  convergence of the estimator for the derivative and statistical
  error, etc). In the central finite difference, is some variance
  reduction technique used? Such as using the same realization of the
  random noise for the two dynamics with different forcings (a control
  variate method, as in J. B. Goodman and K. K. Lin, Coupling control
  variates for Markov chain Monte Carlo, J. Comput. Phys., 228,
  (2009), 7127–7136).  From what I understand, I would say ``yes''.  }\\

We have added all detailed information the referee asked to the updated
manuscript. Regarding the finite difference scheme, we did not used
the technique described in the literature by Goddman and Lin.
However, we have numerical checked the accuracy of the finite
difference scheme, and its convergence.  The finite-difference step
was 0.4, and $10^5$ trajectories were used to calculate one
partial derivative (one component of the gradient), and in total
10 partial derivative were calculated at each step of iterations.
The statistical uncertainty of the
partial derivative was roughly
$5\times 10^{-3}$.  This value is 5 times larger than the
response result, so we observe that the optimal control calculated
by the finite-difference displays larger oscillations
larger than that calculated
by the response formula. We have also compared the result with 
different finite-difference step sizes:  Step 0.2 with
statistical uncertainty $3\times 10^{-3}$ 
and step 1.0 with statistical uncertainty
$1\times 10^{-3}$.
All the results are in good consistency, which means
that the finite-difference method used  in the paper for comparison with
the response theory is reliable.\\

\emph{
On a final note, especially since this is a tribute to Giovanni Ciccotti: this work reminds
me of the subtraction technique (G. Ciccotti and G. Jacucci, Direct computation of dynamical
response by molecular dynamics - mobility of a charged Lennard-Jones particle, Phys. Rev.
Lett., 35, 789792, 1975), with a underlying nonequilibrium/driven dynamics rather than an
equilibrium one. Could the authors comment on the analogy?
}\\

In the paper by Giovanni and Jacucci, the authors use direct
nonequilibrium molecular dynamics simulation to 
calculate the \emph{equilibrium response}, i.e. the mobility of the charged
Lennard-Jones particle, and show that
the nonequilibrium approach has the advantage of smaller statistical error over
the Green-Kubo formula. Our  nonequilibrium simulations,
in comparison, calculate the
\emph{nonequilibrium response}.
As far as the authors concern, the method presented in the paper is new and
has never been studied.

\subsection*{The minor scientific comments from referee 1}

We have revised the paper according to the referee's comments. \\

\emph{
(29) page 11, line 28: Do I understand correctly that $\varepsilon$ for the “standard” linear response is
not small at all? The approach then seems doomed from the beginning, and would be
worse and worse as the forcing $u_t$ in the reference driven dynamics increases.
}\\

The answer is yes. We calculate the time-dependent probability density of
the perturbed driven system. If one wants to calculate it  under the
framework of the equilibrium response theory, then the effective perturbation
(to the equilibrium system) is in fact the driving plus the perturbation, as
indicated by the equation on page. 12. Therefore, when the driving 
becomes larger, the effective perturbation to the equilibrium also increases, so the
accuracy of the traditional response theory becomes worse.

\subsection*{The typo corrections from referee 2}

All the typos pointed out by the referee 2 have been corrected in the new manuscript.

\end{document}

### 2013-08-23 writedown/linear.response/paper.tex
improves according to Carsten

@@
 \]
-where we have used the fact that $\rho^{0}(x,t)= \rho_{\infty}(x)$ for all $t\ge 0$.
+where we have used the fact that $\rho^{0}(x,t)= \rho_{\infty}(x)$ for
+all $t\ge 0$.
+
+\paragraph*{Remark.}
+\recheck{Ciccotti and Jacucci have pointed out that
+  the equilibrium linear response
+can also be calculated by a direct nonequilibrium simulation of the
+perturbed process, and  the nonequilibrium
+approach has a smaller statistical uncertainty in comparison with the
+standard Green-Kubo fomula~\cite{ciccotti1975direct}.
+}
 
@@
 \]
-Suppose that there exists an auxiliary stochastic process \redc{ $\xi_{t}\in\R^{n}$} such that 
+Suppose that there exists an auxiliary stochastic process $\xi_{t}\in\R^{n}$ such that 
 \begin{equation}\label{xi}
@@
 \[
-\sqrt{2\beta^{-1}}\recheck{\gamma}^{1/2} \xi_{t} = D(X_{t})\,\delta u(t)\,.
+\sqrt{2\beta^{-1}}{\gamma}^{1/2} \xi_{t} = D(X_{t})\,\delta u(t)\,.
 \]
-\recheck{Hence the equation for $\xi_{t}$ is solvable if and only if $\mathrm{range}(D(\cdot))\perp \mathrm{ker}(\gamma^{1/2})$, which, since $\gamma$ has full rank, means no restriction on $D(\cdot)$.}
+Hence the equation for $\xi_{t}$ is solvable if and only if $\mathrm{range}(D(\cdot))\perp \mathrm{ker}(\gamma^{1/2})$, which, since $\gamma$ has full rank, means no restriction on $D(\cdot)$.
   % means that $D(\cdot)$ must be invertible almost everywhere.}
@@
 \end{equation}
-\redc{with $(q_0, p_0)$ being subject to a given initial distribution $\mu$, and}
+with $(q_0, p_0)$ being subject to a given initial distribution $\mu$, and
 \begin{equation}\label{ocp2}
@@
 \end{equation}
-with the expectation being over all realizations of the controlled Langevin dynamics in (\ref{ocp1}) with \recheck{the} prescribed initial distribution.
+with the expectation being over all realizations of the controlled Langevin dynamics in (\ref{ocp1}) with the prescribed initial distribution.
 Here, the space of admissible controls $\cU$ consists of all bounded controls of the form
@@
 \[
-\bE_{P^{0}}[\psi] = \int_{\R^{2n}}\bE_{P^{0}_{x}}[\psi]\,d\recheck{\mu}(x) \,,
+\bE_{P^{0}}[\psi] = \int_{\R^{2n}}\bE_{P^{0}_{x}}[\psi]\,d\mu(x) \,,
 \]
-with $P_{x}^{0}$ denoting the corresponding path measure with deterministic initial condition,
-\redc{and $\mu$ is the prescribed initial distribution.}
+with $P_{x}^{0}$ denoting the corresponding path measure with deterministic initial condition, and $\mu$ is the prescribed initial distribution.
+\recheck{A related result on computing the parameter sensitivity can be found
+  in Ref.~\cite{warren2012malliavin}.}
 Note that we can compute all partial derivatives $\partial I/\partial a_k$, $k=1,\ldots,K$ from just one ensemble of nonequilibrium paths of $(q_{t},p_{t})$.
@@
 \recheck{
-We derived linear response formulas for nonequilibrium molecular
+  We derived the first and second order
+  response formulas for nonequilibrium molecular
 dynamics (driven Langevin dynamics), which is a generalization of the
-standard finite-time equilibrium response theory.  The novel response theory
-dose not rely on any assumption on the existence of the invariant
-measure nor stationary state. We validated the formula in numerical
+standard finite-time equilibrium linear response theory.
+The novel nonequilibrium response theory
+does not rely on any reversibility or stationarity assumption on the dynamics.
+We validated the formula in numerical
 experiments in comparison to brute-force nonequilibrium

### 2013-08-23 writedown/linear.response/reply/reply.tex
improves according to Carsten

@@
 equilibrium response theory (3) only in the finite-time
-horizon, so that it is comparable to the the nonequilibrium response theory (10), which is also
+horizon, so that it is comparable to the nonequilibrium response theory (10), which is also
 developed in the finite-time horizon. Our work
@@
 generalization of the equilibrium response theory in the sense that the
-reference process is not assumed to be in equilibrium nor in stationary.
+reference process is not assumed to be in equilibrium.
+Being more general, our theory does not assume the existence of nonequilibrium steady state.
 The introduction and conclusion are revised accordingly.
@@
 
-As far as we concerned, we do not have a clear answer to this question
-in the most general case.
-However, in this paper, we deal with  the Langevin dynamics.
-The friction
+% As far as we concerned, we do not have a clear answer to this question
+% in the most general case.
+In this paper, when we deal with  the Langevin dynamics, 
+the friction
 matrix is positive definite, so the auxiliary process $\xi_t$ is uniquely
 solvable.
+If the matrix in front of the noise does not have full rank, then
+there are some components of the driving Brownian motion in the SDE
+that are uncorrelated from $X_t$, in which case the extra terms are
+independent of $\varphi$ that depends only on $(X_t)$. Hence the
+corresponding 1st and 2nd order linear response expressions vanish
+because of the Martingale property $\textbf{E}[\int \xi^{extra}\cdot dB^{extra}]=0$
+and the Ito isometry $\textbf{E}[(\int \xi^{extra}\cdot dB^{extra})^2]= \textbf{E} [\int
+|\xi^{extra}|^2dt]$.
 \\
@@
 As far as the authors concern, the method presented in the paper is new and
-has never been studied.
+has never been studied. The citation and discussion to Giovanni and Jacucci's
+work have been added to the manuscript.
 

### 2013-08-27 writedown/linear.response/paper.tex
remove all red things

@@
 \paragraph*{Remark.}
-\recheck{Ciccotti and Jacucci have pointed out that
+{Ciccotti and Jacucci have pointed out that
   the equilibrium linear response
 can also be calculated by a direct nonequilibrium simulation of the
-perturbed process, and  the nonequilibrium
+perturbed process, and  this nonequilibrium
 approach has a smaller statistical uncertainty in comparison with the
@@
 with $P_{x}^{0}$ denoting the corresponding path measure with deterministic initial condition, and $\mu$ is the prescribed initial distribution.
-\recheck{A related result on computing the parameter sensitivity can be found
+{A related result on computing the parameter sensitivity can be found
   in Ref.~\cite{warren2012malliavin}.}
@@
 
-\recheck{
+{
   We derived the first and second order

### 2013-09-09 writedown/accelerate/main.tex
add considerations for accerlate

\documentclass[aip,jcp,a4paper,reprint,onecolumn]{revtex4-1}

\usepackage[fleqn]{amsmath}
\usepackage{amssymb}
\usepackage[dvips]{graphicx}
\usepackage{color}
\usepackage{tabularx}
\usepackage{algorithm}
\usepackage{algorithmic}

\makeatletter
\makeatother

\newcommand{\recheck}[1]{{\color{red} #1}}
\newcommand{\redc}[1]{{\color{red} #1}}
\newcommand{\bluec}[1]{{\color{blue} #1}}
\newcommand{\vect}[1]{\textbf{\textit{#1}}}
\newcommand{\dd}{\textsf{d}}
\newcommand{\inv}{\textrm{inv}}

\newcommand{\mh}{\mathcal H}
\newcommand{\eps}{\varepsilon}
\newcommand{\ml}{\mathcal L}
\newcommand{\me}{\mathcal E}
\newcommand{\mt}{\mathcal T}
\newcommand{\mo}{\mathcal O}
\newcommand{\mi}{\mathcal I}
\newcommand{\mc}{\mathcal C}
\newcommand{\proj}{\mathit\Pi}
\newcommand{\fwg}{{\mathcal A}}
\newcommand{\bwg}{{\mathcal B}}
\newcommand{\bsigma}{\boldsymbol\sigma}
\newcommand{\pathmeas}{\mathcal P}

\begin{document}

\title{Calculating the nonequilibrium observable by a speed-boosing
technique}
\author{Han Wang}
\affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}
\author{Christof Sch\"utte}
\affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}
\affiliation{Zuse Institute Berlin (ZIB), Germany}
% \affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}

\date{\today}

\begin{abstract}
\end{abstract}

\maketitle

\section{Theoretical background}

We consider the following SDE:
\begin{align}\label{eqn:sde-1}
  \dd \vect x_t = \vect b(\vect x_t) \dd t + \sigma\dd \vect w_t
\end{align}
And we are intested in some nonequilibrium observable of the system,
which is expressed by
\begin{align*}
  O(T) = \int \dd\vect x_0 \rho(\vect x_0, 0)
  \int_{\mc\{T; \vect x_0, 0\}} 
  O(\vect x_t)\,\dd\pathmeas
\end{align*}
where $\mc\{t; \vect x_0, 0\}$ donotes the set of all continuous
trajectories that initially starts at $\vect x_0$, and ends at time $T$.
$\rho(\vect x_0, 0)$ denotes the probability density of the initial position
$\vect x_0$. $\dd \pathmeas$ denotes the probability measure of the trajectories
generated by SDE~\eqref{eqn:sde-1}.

We want to understand if the nonequilibrium observable can be calculated
by an accelerating technique. More precisely, we want to design a new
SDE
\begin{align}\label{eqn:sde-2}
  \dd \vect y_t = \tilde{\vect b}(\vect y_t)\dd t + \tilde\sigma\dd \vect w_t,
\end{align}
such that the solution $\vect y_{t}$ satisfies
\begin{align}\label{eqn:tmp3}
  \vect y_t = \vect x_{\alpha t}, \quad t \in [0, T/\alpha)
\end{align}
We assumes the same initial condition for both the dynamics.
Here $\alpha$ is the scaling
prefactor. $\alpha > 1$ means Eq.~\eqref{eqn:sde-2} is an accelerated
dynamics w.r.t. Eq.~\eqref{eqn:sde-1}, while if $\alpha < 1$ means a
a slowdown.
The
observable of the original dynamics can be replace by the observation
of the accelerated dynamics via the following relation:
\begin{align}\nonumber
  \tilde O(T/\alpha) &=
  \int \dd\vect y_0 {\rho}(\vect y_0, 0)
  \int_{\mc\{T/\alpha; \vect y_0, 0\}} 
  O(\vect y_t)\,\dd\tilde\pathmeas[\vect y_t]\\\nonumber
  & =
  \int\dd\vect x_0\rho(\vect x_0, 0)
  \int_{\mc\{T/\alpha; \vect x_0, 0\}}
  O(\vect x_{\alpha t}) \dd\pathmeas[\vect x_{\alpha t}]\\\nonumber
  & = 
  \int\dd\vect x_0\rho(\vect x_0, 0)
  \int_{\mc\{T; \vect x_0, 0\}}
  O(\vect x_{ t}) \dd\pathmeas[\vect x_{ t}]\\
  & = O(t)
\end{align}
That means for rescale dynamics~\eqref{eqn:sde-2} we only need to simulate
it until time $T/\alpha$, and the calculated observable that
is the same as the original dynamics. Now the problem is how to design
the rescaled dynamics. Actually, the solution is very simple, we simply
let
\begin{align}
  \tilde {\vect b}(\vect y) = \alpha\, {\vect b}(\vect y)\quad
  \textrm{and}\quad
  \tilde\sigma = \sqrt{\alpha}\,\sigma
\end{align}
Then the Eq.~\eqref{eqn:tmp3} is automatically satisfied.
However, stablity or the largest allowed timestep of the numerical
solution of the dynamics depends on the fastest modes of the
force $\vect b$. When the force $\vect b$ is rescaled by $\alpha$, the
largest timestep is also rescaled by $\alpha$, so the total computational
cost of is not reduced by this rescaling technique.

Now we want to eliminate the fast modes in~\eqref{eqn:sde-2}, so that
the largest timestep is not smaller than the original
dynamics~\eqref{eqn:sde-1}, and at the same time the numerical
accuracy for calculating the observable should be preserved. Please notice
that the numerical accuracy of the trajectories are not necessarily
preserved.
Therefore, we need some esitmate on the trajectory, which will
be developed under the framework of  Girsanov  transformation.
Now assume that the original forcing can be splitted by
\begin{align}
  \vect b(\vect x) = \vect b_s (\vect x) + \vect b_f(\vect x)
\end{align}
where $\vect b_f(\vect x)$ is the fast part of the force, while $\vect b_s
(\vect x)$ is the slow part.
We revise the force of the  accelerated dynamics~\eqref{eqn:sde-2} by
\begin{align}
  \hat{\vect b} = \alpha \vect b_s + \vect b_f,
\end{align}
and keeps the diffusion coefficient $\tilde\sigma$ unchanged.
Therefore, the difference between the observable calculated by the
accelerated dynamics and the original observable is given by 
\begin{align}\nonumber
  O(T) &= \tilde O(T/\alpha)\\\nonumber  
  &=\int \dd\vect y_0 {\rho}(\vect y_0, 0)
  \int_{\mc\{T/\alpha; \vect y_0, 0\}} 
  O(\vect y_t)\,\dd\tilde\pathmeas[\vect y_t]\\\nonumber
  &=
  \int \dd\vect y_0 {\rho}(\vect y_0, 0)
  \int_{\mc\{T/\alpha; \vect y_0, 0\}}
  O(\vect y_t) \frac{\dd\tilde\pathmeas[\vect y_t]}{\dd\hat\pathmeas[\vect y_t]}
  \dd\hat\pathmeas[\vect y_t]
\end{align}
Since we have $\tilde{\vect b} - \hat{\vect b} = (\alpha - 1)\vect b_f$,
by the Girsanov transformation, we have
\begin{align}
  \frac{\dd\tilde\pathmeas[\vect y_t]}{\dd\hat\pathmeas[\vect y_t]}
  = \exp
  \Big[
  \frac{\alpha-1}{\tilde{\sigma}}\int_0^{T/\alpha}\vect b_f(\vect x_s)\cdot\dd\vect w_s
  -
  \frac{(\alpha-1)^2}{2\tilde\sigma^2} \int_0^{T/\alpha}\vect b_f^2(\vect x_s)\dd t\,
  \Big]
\end{align}
We therefore have the error estimate
\begin{align}
  \hat O(T/\alpha) = O(T) +
  \int \dd\vect y_0 {\rho}(\vect y_0, 0)\cdot \me(\vect y_0, T/\alpha)
\end{align}
where
\begin{align}
  \me(\vect y_0, t) = 
  \int_{\mc\{T/\alpha; \vect y_0, 0\}}
  O(\vect y_t)
  \cdot
  \Big\{
  1 -  \exp
  \Big[
  \frac{\alpha-1}{\tilde{\sigma}}\int_0^{T/\alpha}\vect b_f(\vect x_s)\cdot\dd\vect w_s
  -
  \frac{(\alpha-1)^2}{2\tilde\sigma^2} \int_0^{T/\alpha}\vect b_f^2(\vect x_s)\dd t\,
  \Big]
  \Big\} \,\dd\hat\pathmeas
\end{align}

\end{document}

### 2013-09-09 writedown/accelerate/main.tex
save, first edition

@@
 
-\title{Calculating the nonequilibrium observable by a speed-boosing
+\title{Calculating the nonequilibrium observable by a speed-boosting
 technique}
@@
 
-We consider the following SDE:
+All formula will be developed formally. We consider the following SDE:
 \begin{align}\label{eqn:sde-1}
@@
 \end{align}
-And we are intested in some nonequilibrium observable of the system,
+And we are interested in some nonequilibrium observable of the system,
 which is expressed by
@@
   \int_{\mc\{T; \vect x_0, 0\}} 
-  O(\vect x_t)\,\dd\pathmeas
+  O[\vect x_t]\,\dd\pathmeas[\vect x_t]
 \end{align*}
-where $\mc\{t; \vect x_0, 0\}$ donotes the set of all continuous
+where $\mc\{t; \vect x_0, 0\}$ denotes the set of all continuous
 trajectories that initially starts at $\vect x_0$, and ends at time $T$.
@@
 We want to understand if the nonequilibrium observable can be calculated
-by an accelerating technique. More precisely, we want to design a new
+by an accelerating technique. 
+More precisely, we want to design a new
 SDE
@@
   \int_{\mc\{T/\alpha; \vect y_0, 0\}} 
-  O(\vect y_t)\,\dd\tilde\pathmeas[\vect y_t]\\\nonumber
+  O[\vect y_t]\,\dd\tilde\pathmeas[\vect y_t]\\\nonumber
   & =
@@
   \int_{\mc\{T/\alpha; \vect x_0, 0\}}
-  O(\vect x_{\alpha t}) \dd\pathmeas[\vect x_{\alpha t}]\\\nonumber
+  O[\vect x_{\alpha t}] \dd\pathmeas[\vect x_{\alpha t}]\\\nonumber
   & = 
@@
   \int_{\mc\{T; \vect x_0, 0\}}
-  O(\vect x_{ t}) \dd\pathmeas[\vect x_{ t}]\\
+  O[\vect x_{ t}] \dd\pathmeas[\vect x_{ t}]\\
   & = O(t)
@@
 Then the Eq.~\eqref{eqn:tmp3} is automatically satisfied.
-However, stablity or the largest allowed timestep of the numerical
+% Moreover, by the fluctuation-dissipation theorem,
+% the temperature of the system remains the same, which implies
+% a preserved invariant measure in the equilibrium case.
+However, stability or the largest allowed time-step of the numerical
 solution of the dynamics depends on the fastest modes of the
 force $\vect b$. When the force $\vect b$ is rescaled by $\alpha$, the
-largest timestep is also rescaled by $\alpha$, so the total computational
+largest time-step is also rescaled by $\alpha$, so the total computational
 cost of is not reduced by this rescaling technique.
@@
 Now we want to eliminate the fast modes in~\eqref{eqn:sde-2}, so that
-the largest timestep is not smaller than the original
+the largest time-step is not smaller than the original
 dynamics~\eqref{eqn:sde-1}, and at the same time the numerical
@@
 preserved.
-Therefore, we need some esitmate on the trajectory, which will
+Therefore, we need some estimate on the trajectory, which will
 be developed under the framework of  Girsanov  transformation.
-Now assume that the original forcing can be splitted by
+Now assume that the original forcing can be split by
 \begin{align}
@@
   \int_{\mc\{T/\alpha; \vect y_0, 0\}} 
-  O(\vect y_t)\,\dd\tilde\pathmeas[\vect y_t]\\\nonumber
+  O[\vect y_t]\,\dd\tilde\pathmeas[\vect y_t]\\\nonumber
   &=
@@
   \int_{\mc\{T/\alpha; \vect y_0, 0\}}
-  O(\vect y_t) \frac{\dd\tilde\pathmeas[\vect y_t]}{\dd\hat\pathmeas[\vect y_t]}
+  O[\vect y_t] \frac{\dd\tilde\pathmeas[\vect y_t]}{\dd\hat\pathmeas[\vect y_t]}
   \dd\hat\pathmeas[\vect y_t]
@@
 by the Girsanov transformation, we have
-\begin{align}
+\begin{align}\label{eqn:meas-reweight}
   \frac{\dd\tilde\pathmeas[\vect y_t]}{\dd\hat\pathmeas[\vect y_t]}
@@
   \Big[
-  \frac{\alpha-1}{\tilde{\sigma}}\int_0^{T/\alpha}\vect b_f(\vect x_s)\cdot\dd\vect w_s
+  \frac{\alpha-1}{\tilde{\sigma}}\int_0^{T/\alpha}\vect b_f(\vect y_s)\cdot\dd\vect w_s
   -
-  \frac{(\alpha-1)^2}{2\tilde\sigma^2} \int_0^{T/\alpha}\vect b_f^2(\vect x_s)\dd t\,
+  \frac{(\alpha-1)^2}{2\tilde\sigma^2} \int_0^{T/\alpha}\vect b_f^2(\vect y_s)\dd t\,
   \Big]
@@
 We therefore have the error estimate
-\begin{align}
-  \hat O(T/\alpha) = O(T) +
+\begin{align}\label{eqn:o-error}
+  O(T) = \hat O(T/\alpha)  +
   \int \dd\vect y_0 {\rho}(\vect y_0, 0)\cdot \me(\vect y_0, T/\alpha)
@@
   \int_{\mc\{T/\alpha; \vect y_0, 0\}}
-  O(\vect y_t)
+  O[\vect y_t]
   \cdot
-  \Big\{
-  1 -  \exp
+  W[\vect y_t]\,\dd\hat\pathmeas
+\end{align}
+and $W[\vect y_s]$ is defined by
+\begin{align}
+  W[\vect y_t] = 
+  \exp
   \Big[
-  \frac{\alpha-1}{\tilde{\sigma}}\int_0^{T/\alpha}\vect b_f(\vect x_s)\cdot\dd\vect w_s
+  \frac{\alpha-1}{\tilde{\sigma}}\int_0^{T/\alpha}\vect b_f(\vect y_s)\cdot\dd\vect w_s
   -
-  \frac{(\alpha-1)^2}{2\tilde\sigma^2} \int_0^{T/\alpha}\vect b_f^2(\vect x_s)\dd t\,
+  \frac{(\alpha-1)^2}{2\tilde\sigma^2} \int_0^{T/\alpha}\vect b_f^2(\vect y_s)\dd t\,
   \Big]
-  \Big\} \,\dd\hat\pathmeas
+  -1
 \end{align}
+Now the problem is that when the accelerate
+approximation~\eqref{eqn:o-error} has a good accuracy. One possibility
+is that the measure reweighting~\eqref{eqn:meas-reweight} is small.
+However, as a matter of fact, the reweighting is usually not small,
+because the fast part of the force is usually large. The other possibility
+is that the observable $O[\vect y]$ has vanished correlation with
+$W[\vect y]$.
+This is usually true if in the system the fast DOFs are decoupled with
+the slow variables, and the observation is only on the space spanned
+by the slow variables.
+
+Being more specifically, in a molecular system, it is reasonable to
+assume that the vibrational DOFs are decoupled with the slow variable.
+Therefore, as a good approximation, people use rigid bonds instead of
+flexible bonds to increase time-step of the simulation. Please notice
+if the numerical scheme only has first order accuracy, than our accelerating
+technique is equivalent to increasing the time-step of the simulation.
+If higher order numerical schemes are used, our accelerating technique
+is more accurate than simply using larger time-steps.
+
+The other fast variable is the repulsive part of the Lennard-Jones
+interaction, which can be removed by the following way: we define a
+smooth switching function $s(r)$ that has value 0 at repulsive part of
+the LBJ potential, while has value of 1 at the attractive part. Then
+the accelerated LJ potential (denoted by $\hat u(r)$) can be given by
+\begin{align}
+  \hat u(r) = \alpha s(r) u(r) + (1 - s(r)) u(r)
+\end{align}
+where $u(r)$ is the original LJ potential. It is clear that
+ $\hat u(0) = u(0)$, and $\hat u(\infty) = \alpha u(\infty)$.
 

### 2013-10-09 accelerate/template.alanine/tools/parse.top/app/rescale.param.cpp
add support for charmm force field, not extensively tested

@@ -8,14 +8,16 @@ using namespace std;
 int main (int argc, char **argv) {
 
   std::string ifile, ofile, opfile;
-  double sbond, sangle, sdihedral;
+  double sbond, sangle, sdihedral, sanglebond, scmap;
 
   po::options_description desc ("Allow options");
   desc.add_options()
       ("help,h", "print this message")      
       ("scal-bond,b", po::value<double > (&sbond)->default_value (1.0), "bond scale")
       ("scal-angle,a", po::value<double > (&sangle)->default_value (1.0), "angle scale")
+      ("scal-angle-bond", po::value<double > (&sanglebond)->default_value (1.0), "angle-bond scale")
       ("scal-dihedral,d", po::value<double > (&sdihedral)->default_value (1.0), "dihedral scale")
+      ("scal-cmap,c", po::value<double > (&scmap)->default_value (1.0), "cmap scale")
       ("output,o", po::value<std::string > (&ofile)->default_value ("out.top"), "the output top")
       ("input,f",  po::value<std::string > (&ifile)->default_value ("topol.top"), "the input top");
 
@@ -35,15 +37,15 @@ int main (int argc, char **argv) {
   if (systype.bondtypes.size() > 0){
     for (unsigned ii = 0; ii < systype.bondtypes.size(); ++ii){
       if (systype.bondtypes[ii].funct == 1 ||
-	  systype.bondtypes[ii].funct == 2 ||
-	  systype.bondtypes[ii].funct == 6 ||
-	  systype.bondtypes[ii].funct == 7 ||
-	  systype.bondtypes[ii].funct == 8 ||
-	  systype.bondtypes[ii].funct == 9 ){
-	systype.bondtypes[ii].params[1] *= sbond;
+  	  systype.bondtypes[ii].funct == 2 ||
+  	  systype.bondtypes[ii].funct == 6 ||
+  	  systype.bondtypes[ii].funct == 7 ||
+  	  systype.bondtypes[ii].funct == 8 ||
+  	  systype.bondtypes[ii].funct == 9 ){
+  	systype.bondtypes[ii].params[1] *= sbond;
       }
       else {
-	cerr << "bond type is not 1,2,6,7,8,9, do nothing!" << endl;
[…]

### 2013-10-09 accelerate/template.alanine/tools/parse.top/src/GmxTop.cpp

@@ -78,6 +78,29 @@ notTrivalLine (const string & line)
   return false;
 }
 
+
+static void
+unfoldLines (vector<vector<string > > & lines)
+{
+  vector<vector<string > > tmplines (lines);
+  lines.clear ();
+
+  for (unsigned ii = 0; ii < tmplines.size(); ++ii){
+    vector<string > blockline;
+    for (unsigned jj = 0; jj < tmplines[ii].size(); ++jj){
+      if (blockline.size() != 0 && blockline.back()[blockline.back().size()-1] == '\\'){
+	blockline.back().erase(blockline.back().size()-1);
+	blockline.back().push_back(' ');
+	blockline.back() += tmplines[ii][jj];
+      }
+      else {
+	blockline.push_back (tmplines[ii][jj]);
+      }
+    }
+    lines.push_back (blockline);
+  }
+}
+
 static void 
 readBlocks (ifstream & file,
 	    vector<string> & keys,
@@ -121,6 +144,8 @@ readBlocks (ifstream & file,
     }
   }
   lines.push_back (blockLines);
+
+  unfoldLines (lines);
 }
 
 
@@ -311,6 +336,13 @@ print (FILE * fp) const
     fprintf (fp, "[ dihedraltypes ]\n");
     for (unsigned ii = 0; ii < dihedraltypes.size(); ++ii){
       dihedraltypes[ii].print (fp);
+    }
+    fprintf (fp, "\n");
+  }
+  if (cmaptypes.size() > 0){
+    fprintf (fp, "[ cmaptypes ]\n");
+    for (unsigned ii = 0; ii < cmaptypes.size(); ++ii){
+      cmaptypes[ii].print (fp);
     }
     fprintf (fp, "\n");
   }
@@ -621,6 +653,28 @@ parseType (const string & fname,
 	  tmp.params.push_back (atof(words[ii].c_str()));
 	}
 	type.dihedraltypes.push_back (tmp);
+      }
+    }
+  }
+
+  for (unsigned ii = 0; ii < keys.size(); ++ii){
+    if (keys[ii] == "cmaptypes"){
+      for (unsigned jj = 0; jj < lines[ii].size(); ++jj){
+	StringOperation::split (lines[ii][jj], words);
+	if (words.size() < 8) die_wrong_format (__FILE__, __LINE__);
+	gmx_cmaptypes_item tmp;
[…]

### 2013-10-09 accelerate/template.alanine/tools/parse.top/src/GmxType.cpp

@@ -83,4 +83,20 @@ print (FILE * fp) const
   fprintf (fp, "\n");
 }
 
-  
+void GmxTop::gmx_cmaptypes_item::  
+print (FILE * fp) const
+{
+  fprintf (fp, "%s\t%s\t%s\t%s\t%s\t%d\t%d\t%d",
+	   name0.c_str(), name1.c_str(), name2.c_str(), name3.c_str(), name4.c_str(),
+	   funct,
+	   ngrid0, ngrid1);
+  fprintf (fp, "\\\n");
+  for (unsigned ii = 0; ii < params.size(); ++ii){
+    fprintf (fp, "%.10f ", params[ii]);
+    if ((ii+1) % 10 == 0){
+      fprintf (fp, "\\\n");
+    }
+  }
+  fprintf (fp, "\n");
+}
+

### 2013-10-09 accelerate/template.alanine/tools/parse.top/src/GmxTop.cpp
better printing, better position to normalize the lines

@@ -113,12 +113,11 @@ readBlocks (ifstream & file,
   
   while (! file.eof() ){
     file.getline (line, MAX_LINE_LENGTH);
-    
+    normalizeLine (line);
     // cout << line <<endl;
     
     if (!inBlock){
       if (ifKeyWord(line, tmpKey)){
-	normalizeLine (line);
 	inBlock = true;
 	// keys.push_back ("");
 	keys.push_back (tmpKey);
@@ -126,19 +125,16 @@ readBlocks (ifstream & file,
 	blockLines.clear ();
       }
       else if (notTrivalLine(line)) {
-	normalizeLine (line);
 	blockLines.push_back(line);	
       }
     }
     else {
       if (ifKeyWord(line, tmpKey)){
-	normalizeLine (line);
 	keys.push_back (tmpKey);
 	lines.push_back (blockLines);
 	blockLines.clear ();	
       }
       else if (notTrivalLine(line)){
-	normalizeLine (line);
 	blockLines.push_back(line);
       }
     }
@@ -198,7 +194,7 @@ print (FILE * fp) const
 {
   fprintf (fp, "%d\t%d\t%d", ii, jj, funct);
   for (unsigned ii = 0; ii < params.size(); ++ii){
-    fprintf (fp, "\t%f", params[ii]);
+    fprintf (fp, "\t%.10e", params[ii]);
   }
   fprintf (fp, "\n");
 }
@@ -208,7 +204,7 @@ print (FILE * fp) const
 {
   fprintf (fp, "%d\t%d\t%d", ii, jj, funct);
   for (unsigned ii = 0; ii < params.size(); ++ii){
-    fprintf (fp, "\t%f", params[ii]);
+    fprintf (fp, "\t%.10e", params[ii]);
   }
   fprintf (fp, "\n");
 }
@@ -218,7 +214,7 @@ print (FILE * fp) const
 {
   fprintf (fp, "%d\t%d\t%d\t%d", ii, jj, kk, funct);
   for (unsigned ii = 0; ii < params.size(); ++ii){
-    fprintf (fp, "\t%f", params[ii]);
+    fprintf (fp, "\t%.10e", params[ii]);
   }
   fprintf (fp, "\n");
 }
@@ -228,7 +224,7 @@ print (FILE * fp) const
 {
   fprintf (fp, "%d\t%d\t%d\t%d\t%d", ii, jj, kk, ll, funct);
   for (unsigned ii = 0; ii < params.size(); ++ii){
[…]

### 2013-10-09 accelerate/template.alanine/tools/parse.top/src/GmxType.cpp

@@ -16,7 +16,7 @@ print (FILE * fp) const
 void GmxTop::gmx_atomtypes_item::
 print (FILE * fp) const
 {
-  fprintf (fp, "%s\t%d\t%f\t%f\t%s\t%e\t%e\n",
+  fprintf (fp, "%s\t%d\t%f\t%f\t%s\t%.10e\t%.10e\n",
 	   name.c_str(), atom_num, mass, charge, ptype.c_str(), c6, c12);
 }
 
@@ -26,7 +26,7 @@ print (FILE * fp) const
   fprintf (fp, "%s\t%s\t%d",
 	   name0.c_str(), name1.c_str(), funct);
   for (unsigned ii = 0; ii < params.size(); ++ii){
-    fprintf (fp, "\t%e", params[ii]);
+    fprintf (fp, "\t%.10e", params[ii]);
   }
   fprintf (fp, "\n");
 }
@@ -37,7 +37,7 @@ print (FILE * fp) const
   fprintf (fp, "%s\t%s\t%d",
 	   name0.c_str(), name1.c_str(), funct);
   for (unsigned ii = 0; ii < params.size(); ++ii){
-    fprintf (fp, "\t%f", params[ii]);
+    fprintf (fp, "\t%.10e", params[ii]);
   }
   fprintf (fp, "\n");
 }
@@ -48,7 +48,7 @@ print (FILE * fp) const
   fprintf (fp, "%s\t%s\t%d",
 	   name0.c_str(), name1.c_str(), funct);
   for (unsigned ii = 0; ii < params.size(); ++ii){
-    fprintf (fp, "\t%f", params[ii]);
+    fprintf (fp, "\t%.10e", params[ii]);
   }
   fprintf (fp, "\n");
 }
@@ -59,7 +59,7 @@ print (FILE * fp) const
   fprintf (fp, "%s\t%s\t%s\t%d",
 	   name0.c_str(), name1.c_str(), name2.c_str(), funct);
   for (unsigned ii = 0; ii < params.size(); ++ii){
-    fprintf (fp, "\t%f", params[ii]);
+    fprintf (fp, "\t%.10e", params[ii]);
   }
   fprintf (fp, "\n");
 }
@@ -71,13 +71,13 @@ print (FILE * fp) const
 	   name0.c_str(), name1.c_str(), name2.c_str(), name3.c_str(), funct);
   if (funct == 1 || funct == 4 || funct == 9){
[…]

### 2013-10-09 accelerate/template.alanine.pme/tools/parse.top/include/GmxType.h
add support for charmm, not well tested

@@ -90,6 +90,21 @@ namespace GmxTop {
   }
       ;
 
+  struct gmx_cmaptypes_item
+  {
+    string		name0;
+    string		name1;
+    string		name2;
+    string		name3;
+    string		name4;
+    int			funct;
+    int			ngrid0;
+    int			ngrid1;    
+    vector<double >	params;
+    void		print (FILE * fp) const;
+  }
+      ;
+
   struct gmx_sys_types
   {
     gmx_defaults_item			defaults;
@@ -99,6 +114,7 @@ namespace GmxTop {
     vector<gmx_bondtypes_item>		bondtypes;
     vector<gmx_angletypes_item>		angletypes;
     vector<gmx_dihedraltypes_item>	dihedraltypes;
+    vector<gmx_cmaptypes_item>		cmaptypes;
     void		print (FILE * fp) const;
   }
       ;

### 2013-10-09 accelerate/template.alanine.pme/tools/parse.top/src/GmxTop.cpp

@@ -78,6 +78,29 @@ notTrivalLine (const string & line)
   return false;
 }
 
+
+static void
+unfoldLines (vector<vector<string > > & lines)
+{
+  vector<vector<string > > tmplines (lines);
+  lines.clear ();
+
+  for (unsigned ii = 0; ii < tmplines.size(); ++ii){
+    vector<string > blockline;
+    for (unsigned jj = 0; jj < tmplines[ii].size(); ++jj){
+      if (blockline.size() != 0 && blockline.back()[blockline.back().size()-1] == '\\'){
+	blockline.back().erase(blockline.back().size()-1);
+	blockline.back().push_back(' ');
+	blockline.back() += tmplines[ii][jj];
+      }
+      else {
+	blockline.push_back (tmplines[ii][jj]);
+      }
+    }
+    lines.push_back (blockline);
+  }
+}
+
 static void 
 readBlocks (ifstream & file,
 	    vector<string> & keys,
@@ -90,12 +113,11 @@ readBlocks (ifstream & file,
   
   while (! file.eof() ){
     file.getline (line, MAX_LINE_LENGTH);
-    
+    normalizeLine (line);
     // cout << line <<endl;
     
     if (!inBlock){
       if (ifKeyWord(line, tmpKey)){
-	normalizeLine (line);
 	inBlock = true;
 	// keys.push_back ("");
 	keys.push_back (tmpKey);
@@ -103,24 +125,23 @@ readBlocks (ifstream & file,
 	blockLines.clear ();
       }
       else if (notTrivalLine(line)) {
-	normalizeLine (line);
 	blockLines.push_back(line);	
       }
     }
     else {
       if (ifKeyWord(line, tmpKey)){
-	normalizeLine (line);
 	keys.push_back (tmpKey);
 	lines.push_back (blockLines);
 	blockLines.clear ();	
       }
       else if (notTrivalLine(line)){
-	normalizeLine (line);
 	blockLines.push_back(line);
       }
     }
   }
   lines.push_back (blockLines);
+
+  unfoldLines (lines);
 }
 
 
@@ -173,7 +194,7 @@ print (FILE * fp) const
 {
   fprintf (fp, "%d\t%d\t%d", ii, jj, funct);
[…]

### 2013-10-09 accelerate/template.alanine.pme/tools/parse.top/src/GmxType.cpp

@@ -16,7 +16,7 @@ print (FILE * fp) const
 void GmxTop::gmx_atomtypes_item::
 print (FILE * fp) const
 {
-  fprintf (fp, "%s\t%d\t%f\t%f\t%s\t%e\t%e\n",
+  fprintf (fp, "%s\t%d\t%f\t%f\t%s\t%.10e\t%.10e\n",
 	   name.c_str(), atom_num, mass, charge, ptype.c_str(), c6, c12);
 }
 
@@ -26,7 +26,7 @@ print (FILE * fp) const
   fprintf (fp, "%s\t%s\t%d",
 	   name0.c_str(), name1.c_str(), funct);
   for (unsigned ii = 0; ii < params.size(); ++ii){
-    fprintf (fp, "\t%e", params[ii]);
+    fprintf (fp, "\t%.10e", params[ii]);
   }
   fprintf (fp, "\n");
 }
@@ -37,7 +37,7 @@ print (FILE * fp) const
   fprintf (fp, "%s\t%s\t%d",
 	   name0.c_str(), name1.c_str(), funct);
   for (unsigned ii = 0; ii < params.size(); ++ii){
-    fprintf (fp, "\t%f", params[ii]);
+    fprintf (fp, "\t%.10e", params[ii]);
   }
   fprintf (fp, "\n");
 }
@@ -48,7 +48,7 @@ print (FILE * fp) const
   fprintf (fp, "%s\t%s\t%d",
 	   name0.c_str(), name1.c_str(), funct);
   for (unsigned ii = 0; ii < params.size(); ++ii){
-    fprintf (fp, "\t%f", params[ii]);
+    fprintf (fp, "\t%.10e", params[ii]);
   }
   fprintf (fp, "\n");
 }
@@ -59,7 +59,7 @@ print (FILE * fp) const
   fprintf (fp, "%s\t%s\t%s\t%d",
 	   name0.c_str(), name1.c_str(), name2.c_str(), funct);
   for (unsigned ii = 0; ii < params.size(); ++ii){
-    fprintf (fp, "\t%f", params[ii]);
+    fprintf (fp, "\t%.10e", params[ii]);
   }
   fprintf (fp, "\n");
 }
@@ -71,16 +71,32 @@ print (FILE * fp) const
 	   name0.c_str(), name1.c_str(), name2.c_str(), name3.c_str(), funct);
   if (funct == 1 || funct == 4 || funct == 9){
[…]

### 2013-10-09 writedown/accelerate/main.tex
backup

@@
 \begin{align}\label{eqn:sde-1}
-  \dd \vect x_t = \vect b(\vect x_t) \dd t + \sigma\dd \vect w_t
+  \dd \vect x_t = \vect b(\vect x_t, t)\, \dd t + \sigma\dd \vect w_t
 \end{align}
@@
 \begin{align}\label{eqn:sde-2}
-  \dd \vect y_t = \tilde{\vect b}(\vect y_t)\dd t + \tilde\sigma\dd \vect w_t,
+  \dd \vect y_t = \tilde{\vect b}(\vect y_t, t)\,\dd t + \tilde\sigma\dd \vect w_t,
 \end{align}
@@
 \begin{align}
-  \tilde {\vect b}(\vect y) = \alpha\, {\vect b}(\vect y)\quad
+  \tilde {\vect b}(\vect y, t) = \alpha\, {\vect b}(\vect y, \alpha t)\quad
   \textrm{and}\quad
@@
 force $\vect b$. When the force $\vect b$ is rescaled by $\alpha$, the
-largest time-step is also rescaled by $\alpha$, so the total computational
+largest time-step is also rescaled by $1/\alpha$, so the total computational
 cost of is not reduced by this rescaling technique.

### 2013-10-09 accelerate/template.alanine.pme/tools/parse.top/src/GmxTop.cpp
comment out printings

@@ -361,14 +361,14 @@ parseTop (const string & fname,
   readBlocks (file, keys, lines);
   vector<string > words;
 
-  cout << "n. keys " << keys.size() << endl;
-  cout << "n. lines " << lines.size() << endl;
-  for (unsigned ii = 0; ii < keys.size(); ++ii){
-    cout << "key " << ii << ": " << keys[ii] << endl;
-    for (unsigned jj = 0; jj < lines[ii].size(); ++jj){
-      cout << lines[ii][jj] << endl;
-    }
-  }
+  // cout << "n. keys " << keys.size() << endl;
+  // cout << "n. lines " << lines.size() << endl;
+  // for (unsigned ii = 0; ii < keys.size(); ++ii){
+  //   cout << "key " << ii << ": " << keys[ii] << endl;
+  //   for (unsigned jj = 0; jj < lines[ii].size(); ++jj){
+  //     cout << lines[ii][jj] << endl;
+  //   }
+  // }
   
   for (unsigned ii = 0; ii < keys.size(); ++ii){
     if (keys[ii] == "system"){

### 2013-10-09 accelerate/template.alanine.pme/tools/parse.top/src/GmxTop.cpp
fix bug

@@ -185,8 +185,14 @@ clear ()
 void GmxTop::gmx_atom::
 print (FILE * fp) const
 {
-  fprintf (fp, "%d\t%s\t%d\t%s\t%s\t%d\t%f\t%f\n",
-	   id, at_type.c_str(), res_nr, res_name.c_str(), at_name.c_str(), cgnr, charge, mass);
+  if (mass <= 0){
+    fprintf (fp, "%d\t%s\t%d\t%s\t%s\t%d\t%f\n",
+	     id, at_type.c_str(), res_nr, res_name.c_str(), at_name.c_str(), cgnr, charge);
+  }
+  else {
+    fprintf (fp, "%d\t%s\t%d\t%s\t%s\t%d\t%f\t%f\n",
+	     id, at_type.c_str(), res_nr, res_name.c_str(), at_name.c_str(), cgnr, charge, mass);
+  }
 }
 
 void GmxTop::gmx_pairs_item::
@@ -555,8 +561,8 @@ parseType (const string & fname,
 	gmx_atomtypes_item tmp;
 	tmp.name = words[0];
 	tmp.atom_num = atoi(words[1].c_str());
-	tmp.mass = atoi(words[2].c_str());
-	tmp.charge = atoi(words[3].c_str());
+	tmp.mass = atof(words[2].c_str());
+	tmp.charge = atof(words[3].c_str());
 	tmp.ptype = string(words[4].c_str());
 	tmp.c6 = atof(words[5].c_str());
 	tmp.c12 = atof(words[6].c_str());

### 2013-10-22 writedown/accelerate/main.tex
add latest result

@@
 \newcommand{\inv}{\textrm{inv}}
+\newcommand{\hard}{\textrm{h}}
+\newcommand{\soft}{\textrm{s}}
+\newcommand{\vdw}{\textrm{vdw}}
+\newcommand{\ele}{\textrm{ele}}
+\newcommand{\dir}{\textrm{dir}}
+\newcommand{\rec}{\textrm{rec}}
+\newcommand{\corr}{\textrm{corr}}
 
@@
 
+\section{The slow and fast part of the CHARMM force field}
+
+In this study, we are using the following CHARMM27 force field~\cite{mackerell2004extending}, which can be described:
+\begin{align}\nonumber
+  U(\vect R) = &
+  \sum_{\textrm{bonds}} K_b(b-b_0)^2\\\nonumber
+  &
+  +\sum_{\textrm{angles}} K_\theta (\theta - \theta_0)^2
+  +\sum_{\textrm{Urey-Bradley}} K_{\textrm{UB}} (S - S_0)^2\\\nonumber
+  &
+  +\sum_{\textrm{dihedrals}} K_\phi( 1+\cos(n\phi - \delta)) 
+  +\sum_{\textrm{impropers}} K_\omega (\omega - \omega_0)^2\\ \nonumber
+  &
+  +\sum_{\textrm{residues}} U_{\textrm{CMAP}}(\phi, \psi) \\\label{eqn:tmp13}
+  &
+  +\sum_{\textrm{non-bonded}}
+  \big[\,U_{\vdw}(r_{ij}) + U_{\ele}(r_{ij})\,\big]
+  % \Bigg\{
+  % \eps_{ij}
+  % \bigg[\bigg(\frac{R_{ij}}{r_{ij}}\bigg)^{12}
+  % - 2\bigg(\frac{R_{ij}}{r_{ij}}\bigg)^{6}
+  % \bigg]
+  % + \frac{q_iq_j}{4\pi\eps_0\eps r_{ij}}
+  % \Bigg\}
+\end{align}
+The terms include bond ($b$), angle ($\theta$), Urey-Bradley ($S$),
+dihedral angle ($\phi$), improper angle ($\omega$), backbone torsional
+correction (CMAP) and non-bonded van der Waals and electrostatic
+interactions.  The van der Waals interaction is written by
+\begin{align}
+  U_{\vdw}(r_{ij}) = 
+  \eps_{ij}
+  \bigg[\bigg(\frac{R_{ij}}{r_{ij}}\bigg)^{12}
+  - 2\bigg(\frac{R_{ij}}{r_{ij}}\bigg)^{6}
+  \bigg]
+\end{align}
+The electrostatic interaction is
+\begin{align}
+  U_{\ele}(r_{ij}) = \frac{q_iq_j}{4\pi\eps_0\eps r_{ij}}
+\end{align}
+We treat the electrostatic interaction by the Particle Mesh Ewald (PME)
+method~\cite{darden1993pme, essmann1995spm}, which split the electrostatic interaction into three parts,
+namely the direct, the reciprocal and the correction parts.
+\begin{align}
+  U_{\ele}(r_{ij}) =
+  U_{\dir}(r_{ij}) +
+  U_{\rec}(r_{ij}) + U_{\corr},
+\end{align}
+where
+\begin {align}
+  U_{\dir} (r_{ij})& = \frac12 
+  \frac{q_iq_j \textrm{erfc}(\beta {r}_{ij})}
+  {{r}_{ij} } \\
+  U_{\rec} (r_{ij})& = \frac1{2\pi V} \sum_{\vect m \neq 0}
+  \frac{\exp(-\pi^2 m^2 / \beta^2)}{m^2}
+   q_iq_j e^ {2 \pi i m \cdot r_{ij}}\\
+  U_{\corr}& = -\frac\beta{\sqrt \pi} q_iq_j \delta_{ij}
+\end {align}
+The correction part is a constant term with respect to the position of
+the atoms, so it is not calculated during
+the force calculation.\\
+
+\noindent
+Among all the components of the CHARMM force field, the bond
+interaction is treated as fast interaction. The angle, Urey-Bradley,
+dihedral angle, improper angle, backbone torsional correction and the
+reciprocal part of the electrostatic interaction are treated as slow
+interactions. The van der Waals interaction has a hard core,
+so it is split into soft and hard contributions in the following way:
+\begin{align}
+  U_{\vdw} (r) &= U^\hard_{\vdw}(r) + U^\soft_{\vdw}(r)
+\end{align}
+and
+\begin{align}
+  U^\hard_{\vdw}(r) & =  [\, 1-w_\vdw - (1 - w_\vdw) s_\vdw(r) \,] \,U_{\vdw}(r)\\
+  U^\soft_{\vdw}(r) & =  [\, w_\vdw + (1 - w_\vdw) s_\vdw(r) \,] \,U_{\vdw}(r)
+\end{align}
+where $w_\vdw$ is a positive parameter between 0 and 1, and
+$s_\vdw(r)$ is a function smoothly switches from 0 to 1:
+\begin{align}
+  s_\vdw(r) =
+  \left\{
+    \begin{aligned}
+      &\frac12 - \frac12 \cos \Big(\frac{r}{r^\hard_\vdw} \pi\Big) &\quad & r < r^\hard_\vdw\\
+      &1 & \quad & r \geq r^\hard_\vdw
+    \end{aligned}
+  \right.
+\end{align}
+The soft part of the van der Waals $U^\soft_{\vdw}(r)$ switches
+smoothly from $w\,U_{\vdw}(r)$ to $U_{\vdw}(r)$ in the range
+$[0, r^\hard_\vdw]$. Here $[0, r^\hard_\vdw]$ is the range where the the van der Waals is
+treated as hard interaction, and therefore has to been reduced. The
+direct part of the electrostatic interaction is split in the same way:
+\begin{align}
+  U_{\dir} (r) &= U^\hard_{\dir}(r) + U^\soft_{\dir}(r)
+\end{align}
+and
+\begin{align}
+  U^\hard_{\dir}(r) & =  [\, 1-w_\dir - (1 - w_\dir) s(r) \,] \,U_{\dir}(r)\\
+  U^\soft_{\dir}(r) & =  [\, w_\dir + (1 - w_\dir) s(r) \,] \,U_{\dir}(r)
+\end{align}
+where $w_\dir$ is a positive parameter between 0 and 1 that
+has the similar meaning as $w_\vdw$. \\
+
+\noindent
+In summary, the CHARMM force field is split in the following way:
+\begin{align}
+  U(\vect R) = U^\hard(\vect R) + U^\soft(\vect R)
+\end{align}
+where
+\begin{align}
+  U^\hard(\vect R) &=
+  \sum_{\textrm{bonds}} K_b(b-b_0)^2 +
+  \sum_{\textrm{non-bonded}} [\,U_{\vdw}^\hard (r_{ij}) + U_{\dir}^\hard (r_{ij})\,] \\\nonumber
+  U^\soft(\vect R) &=
+  +\sum_{\textrm{angles}} K_\theta (\theta - \theta_0)^2
+  +\sum_{\textrm{Urey-Bradley}} K_{\textrm{UB}} (S - S_0)^2\\\nonumber
+  &
+  +\sum_{\textrm{dihedrals}} K_\phi( 1+\cos(n\phi - \delta)) 
+  +\sum_{\textrm{impropers}} K_\omega (\omega - \omega_0)^2\\ \nonumber
+  &
+  +\sum_{\textrm{residues}} U_{\textrm{CMAP}}(\phi, \psi) \\
+  &
+  +\sum_{\textrm{non-bonded}}
+  \big[\,U^\soft_{\vdw}(r_{ij})
+  + U^\soft_{\dir}(r_{ij})
+  + U_{\rec}(r_{ij})\,\big]  
+\end{align}
+Therefore, the interaction for the accelerated system is
+\begin{align}
+  \hat U(\vect R) = U^\hard(\vect R) + \alpha\,U^\soft(\vect R)  
+\end{align}
+
+\section{Numerical results}
+\noindent
+We study the alanine dipeptide under an oscillatory EF for example.
+This example is exactly the same as the period $T=40$~ps case in the previous
+manuscript~\cite{wang2013exploring}. The only difference is that we use
+flexible bonds instead of rigid bonds in the simulations, and the corresponding
+time-step is now 0.5~fs. The split parameter $\beta$ in the PME method is optimzied by
+the method mentioned in Ref.~\cite{wang2010optimizing}\\
+
+\begin{table}
+  \centering
+  \begin{tabular*}{0.48\textwidth}{@{\extracolsep{\fill}}lcccc}
+    \hline\hline
+    $\alpha$ & $w_\vdw$ & $r^\hard_\vdw$ [nm] & $w_\dir$ & $r^\hard_\vdw$ [nm] \\\hline
+    2           & 1.00     & 0.0   & 1.00    & 0.0\\
+    5           & 1.00     & 0.0   & 0.80    & 0.2\\
+    6           & 1.00     & 0.0   & 0.67    & 0.2\\
+    \hline\hline
+  \end{tabular*}
+  \caption{The systems tested and the parameters to split the slow and fast parts of the force.}
+  \label{tab:tmp1}
+\end{table}
+
+\noindent
+Till now we have tested the following three systems listed in
+Tab.~\ref{tab:tmp1}. The probabilities of different conformations as a
+function of time are plotted in Fig.~\ref{fig:tmp1}--\ref{fig:tmp3}.
+We can see that when $\alpha = 2$, the result is consistent with the
+reference simulation (no speed boosting, i.e.~$\alpha = 1$). For
+bigger $\alpha = 5$ and 6, the small time scale oscillating (20~ps
+that is roughly the same as half of the period of the EF) of
+conformations $A_1$, $A_2$, $B_1$ and $B_2$ is not well captured by
+the speed boosting algorithm. Moreover, the probability of $B_1$ is
+overestimated, while the probability of $B_2$ is underestimated (see
+the blue and purple lines in Fig.~\ref{fig:tmp2} and \ref{fig:tmp3}).
+We plot the time dependent probability density of $\alpha$-helix
+($A_1+A_2$) and $\beta$-sheet ($B_1+B_2$) conformations, on the right
+plots of Fig.~\ref{fig:tmp2} and \ref{fig:tmp3}. We can see that for
+$\alpha=5$, the result consistent with the reference simulation very
+well, while for $\alpha=6$ the precision is slightly lower, but is
+still qualitatively correct.  That means that the fine-scaled
+oscillating of conformation and the probability exchange between the
+two sub-conformations of the $\beta$-sheet are statistically relevant
+to the hard forces that are modified during the speed boosting
+simulation, while the conformation changes between the $\alpha$-helix
+$\beta$-sheet and left-handed $\alpha$-helix ($C$)  are not correlated
+with the hard forces.
+
+\begin{figure}
+  \centering
+  \includegraphics[]{figs/fig-meta02.eps}
+  \caption{$\alpha = 2$, the probability of different conformations as
+    a function of time.  Red: $A_1$, green: $A_2$, blue: $B_1$,
+    purple: $B_2$, cyan: $C$.  The lighter colors denote the reference
+    exact dynamics of the conformations ($\alpha = 1$), while the
+    darker colors denote $\alpha = 2$.}
+  \label{fig:tmp1}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[]{figs/fig-meta04.eps}
+  \includegraphics[]{figs/fig-meta04-1.eps}
+  \caption{$\alpha = 4$, the probability of different conformations as a function of time.
+    In the left plot: Red: $A_1$, green: $A_2$, blue: $B_1$, purple: $B_2$, cyan: $C$.
+    In the right plot: Red: $A_1+A_2$, blue: $B_1+B_2$, cyan: $C$.
+    The lighter colors denote the reference exact dynamics of the conformations ($\alpha = 1$), while the darker colors denote $\alpha = 4$.}
+  \label{fig:tmp2}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[]{figs/fig-meta05.eps}
+  \includegraphics[]{figs/fig-meta05-1.eps}
+  \caption{$\alpha = 5$, the probability of different conformations as a function of time.
+    In the left plot: Red: $A_1$, green: $A_2$, blue: $B_1$, purple: $B_2$, cyan: $C$.
+    In the right plot: Red: $A_1+A_2$, blue: $B_1+B_2$, cyan: $C$.
+    The lighter colors denote the reference exact dynamics of the conformations ($\alpha = 1$), while the darker colors denote $\alpha = 5$.}
+  \label{fig:tmp2}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[]{figs/fig-meta08.eps}
+  \includegraphics[]{figs/fig-meta08-1.eps}
+  \caption{$\alpha = 6$, the probability of different conformations as a function of time.
+    In the left plot: Red: $A_1$, green: $A_2$, blue: $B_1$, purple: $B_2$, cyan: $C$.
+    In the right plot: Red: $A_1+A_2$, blue: $B_1+B_2$, cyan: $C$.
+    The lighter colors denote the reference exact dynamics of the conformations ($\alpha = 1$), while the darker colors denote $\alpha = 6$.}
+  \label{fig:tmp3}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-02-1.eps}
+  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-02-2.eps}
+  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-02-3.eps}
+  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-02-4.eps}
+  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-02-5.eps}\\
+  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-04-1.eps}
+  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-04-2.eps}
+  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-04-3.eps}
+  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-04-4.eps}
+  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-04-5.eps}\\
+  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-05-1.eps}
+  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-05-2.eps}
+  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-05-3.eps}
+  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-05-4.eps}
+  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-05-5.eps}\\
+  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-08-1.eps}
+  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-08-2.eps}
+  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-08-3.eps}
+  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-08-4.eps}
+  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-08-5.eps}\\
+  \caption{The integrated probability fluxes. $\alpha = 2,4,5,8$.}
+  \label{fig:tmp3}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-04-1-1.eps}
+  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-04-1-2.eps}
+  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-04-1-3.eps}\\
+  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-05-1-1.eps}
+  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-05-1-2.eps}
+  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-05-1-3.eps}\\
+  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-08-1-1.eps}
+  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-08-1-2.eps}
+  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-08-1-3.eps}
+  \caption{The integrated probability fluxes. $\alpha = 4,5,8$.}
+  \label{fig:tmp3}
+\end{figure}
+
+\section*{References}
+\bibliography{ref}{}
+\bibliographystyle{unsrt}
+
 \end{document}

### 2013-10-22 accelerate/template.butane.resp/functions.sh
better random seed

@@ -55,7 +55,7 @@ function set_parameters_pert () {
     pert_epsilon_r=`echo "1. / $pert_rescale" | bc -l`
     pert_taut_r=`echo "$pert_taut / $pert_rescale" | bc -l`
     pert_seed=`date +%s`
-    pert_seed=`echo "$pert_seed + $pert_num_conf_use * $pert_parallel_my_id" | bc`
+    pert_seed=`echo "$pert_seed + $pert_num_conf_use * 10 * $pert_parallel_my_id" | bc`
     sed -e "/^dt/s/=.*/= $pert_dt/g" $file |\
     sed -e "/^integrator/s/=.*/= $pert_integrator/g" |\
     sed -e "/^Tcoupl /s/=.*/= no/g" |\

### 2013-10-22 accelerate/template.butane.resp/tools/angles/app/trans.flux.cpp
add app/trans.prob.resp.cpp, not tested

@@ -1,0 +1,233 @@
+#include <algorithm>
+#include <fstream>
+#include <iomanip>
+#include <iostream>
+#include <iterator>
+#include <list>
+#include <map>
+#include <set>
+#include <sstream>
+#include <string>
+#include <queue>
+#include <vector>
+#include <cmath>
+
+#include "Defines.h"
+#include "StringSplit.h"
+#include "Distribution.h"
+#include "Traj.h"
+
+#include <boost/program_options.hpp>
+namespace po = boost::program_options;
+
+using namespace std;
+
+#define MaxLineLength 2048
+
+bool myread (FILE * fp,
+	     float & time,
+	     ValueType & angle)
+{
+  char line [MaxLineLength];
+  
+  while (1){
+    char* rv = fgets (line, MaxLineLength, fp);
+    if (rv == NULL) return false;
+    if (line[0] == '#' || line[0] == '@') {
+      continue;
+    }
+    else {
+      vector<string> words;
+      StringOperation::split (string(line), words);
+      if (words.size() < 2) {
+	cerr << "error of reading a line!! exit" << endl;
+	exit(1);
+      }
+      time = atof(words[0].c_str());
+      angle = atof(words[1].c_str());
+      break;
+    }
+  }
+  return true;
+}
+
+void depositMetastable (const double & angle,
+			const double & tol,
+			vector<double > & counts)
+{
+  counts.resize (2, 0.);
+  if (angle < -180 + tol || angle > 180 - tol) {
+    counts[0] = 1;
+    counts[1] = 0;
+  }
+  else {
+    counts[0] = 0;
+    counts[1] = 1;
+  }
+}
+
+void calCorr (const vector<double> & counts0,
+	      const vector<double> & counts1,
+	      vector<vector<double > > & countCorr)
+{
+  countCorr.resize (counts0.size());
+  for (unsigned ii = 0; ii < countCorr.size(); ++ii){
+    countCorr[ii].resize (counts1.size());
+    for (unsigned jj = 0; jj < countCorr[ii].size(); ++jj){
+      if (counts1[ii] != 0 && counts0[jj] != 0){
+	countCorr[ii][jj] = 1.;
+      }
+      else {
+	countCorr[ii][jj] = 0.;
+      }
+    }
+  }
+}
+
+int main(int argc, char * argv[])
+{
[…]

### 2013-10-22 accelerate/template.butane.resp/tools/angles/app/average.traj.corr.cpp
fix bugs

@@ -20,10 +20,10 @@ namespace po = boost::program_options;
 using namespace std;
 #define MaxLineLength 2048
 
-bool myread (FILE * fp,
-	     float & time,
-	     ValueType & phi,
-	     ValueType & psi)
+static bool myread (FILE * fp,
+		    float & time,
+		    ValueType & phi,
+		    ValueType & psi)
 {
   size_t rv;
   rv = fread (&time, sizeof(float), 1, fp);
@@ -44,10 +44,10 @@ bool myread (FILE * fp,
   return true;
 }
 
-void depositMetastable (const double & phi,
-			const double & psi,
-			const vector<MetastableSet> & sets,
-			vector<double > & counts)
+static void depositMetastable (const double & phi,
+			       const double & psi,
+			       const vector<MetastableSet> & sets,
+			       vector<double > & counts)
 {
   int cc = 0;
   counts.resize (sets.size());
@@ -69,9 +69,9 @@ void depositMetastable (const double & phi,
   }
 }
 
-void calCorr (const vector<double> & counts0,
-	      const vector<double> & counts1,
-	      vector<vector<double > > & countCorr)
+static void calCorr (const vector<double> & counts0,
+		     const vector<double> & counts1,
+		     vector<vector<double > > & countCorr)
 {
   countCorr.resize (counts0.size());
   for (unsigned ii = 0; ii < countCorr.size(); ++ii){

### 2013-10-22 accelerate/template.butane.resp/tools/angles/app/trans.flux.cpp

@@ -24,12 +24,11 @@ using namespace std;
 
 #define MaxLineLength 2048
 
-bool myread (FILE * fp,
-	     float & time,
-	     ValueType & angle)
+static bool myread (FILE * fp,
+		    float & time,
+		    ValueType & angle)
 {
   char line [MaxLineLength];
-  
   while (1){
     char* rv = fgets (line, MaxLineLength, fp);
     if (rv == NULL) return false;
@@ -51,9 +50,9 @@ bool myread (FILE * fp,
   return true;
 }
 
-void depositMetastable (const double & angle,
-			const double & tol,
-			vector<double > & counts)
+static void depositMetastable (const double & angle,
+			       const double & tol,
+			       vector<double > & counts)
 {
   counts.resize (2, 0.);
   if (angle < -180 + tol || angle > 180 - tol) {
@@ -66,9 +65,9 @@ void depositMetastable (const double & angle,
   }
 }
 
-void calCorr (const vector<double> & counts0,
-	      const vector<double> & counts1,
-	      vector<vector<double > > & countCorr)
+static void calCorr (const vector<double> & counts0,
+		     const vector<double> & counts1,
+		     vector<vector<double > > & countCorr)
 {
   countCorr.resize (counts0.size());
   for (unsigned ii = 0; ii < countCorr.size(); ++ii){
@@ -121,7 +120,6 @@ int main(int argc, char * argv[])
   vector<vector<double > > counts;
   vector<vector<vector<double > > > corrsBw;
   unsigned countFile = 0;
-  unsigned countData = 0;
 
   while (fpname.getline(nameline, MaxLineLength)){
     if (nameline[0] == '#') continue;
@@ -142,7 +140,7 @@ int main(int argc, char * argv[])
     if (nLagTime == 1) nLagTime ++;
     Traj traj (nLagTime);
     fclose (fp);
-    fp = fopen (nameline, "r");
+    fp = fopen (nameline1, "r");
     if (fp == NULL){
       std::cerr << "cannot open file online " << nameline << std::endl;
       std::cerr << std::endl;
@@ -161,7 +159,6 @@ int main(int argc, char * argv[])
 	traj.push_back (tmpcount);
 	if (traj.full ()){
 	  calCorr (traj.front(), traj.back(), tmpcorr);
-	  countData ++;
 	}	
 	corrsBw.push_back (tmpcorr);
[…]

### 2013-10-30 writedown/accelerate/main.tex
improve

@@
 where
-\begin{align}
-  U^\hard(\vect R) &=
-  \sum_{\textrm{bonds}} K_b(b-b_0)^2 +
-  \sum_{\textrm{non-bonded}} [\,U_{\vdw}^\hard (r_{ij}) + U_{\dir}^\hard (r_{ij})\,] \\\nonumber
-  U^\soft(\vect R) &=
-  +\sum_{\textrm{angles}} K_\theta (\theta - \theta_0)^2
-  +\sum_{\textrm{Urey-Bradley}} K_{\textrm{UB}} (S - S_0)^2\\\nonumber
+\begin{align}\nonumber
+  U^\hard(\vect R) =\,&
+  (1 - w_b) \sum_{\textrm{bonds}} K_b(b-b_0)^2 \\\nonumber
+  &
+  +(1 - w_\theta)\bigg[\sum_{\textrm{angles}} K_\theta (\theta - \theta_0)^2
+  +\sum_{\textrm{Urey-Bradley}} K_{\textrm{UB}} (S - S_0)^2\bigg]\\\nonumber
+  &
+  +\sum_{\textrm{non-bonded}} [\,U_{\vdw}^\hard (r_{ij}) + U_{\dir}^\hard (r_{ij})\,] \\\nonumber
+  U^\soft(\vect R) =\,&
+  w_b\sum_{\textrm{bonds}} K_b(b-b_0)^2 \\\nonumber
+  &
+  +w_\theta \bigg[\sum_{\textrm{angles}} K_\theta (\theta - \theta_0)^2
+  +\sum_{\textrm{Urey-Bradley}} K_{\textrm{UB}} (S - S_0)^2\bigg]\\\nonumber
   &
@@
 \end{align}
+The parameters $w_b$ and $w_\theta$ are chosen so that the vibrational
+frequency of the bond $ \big[1 + (\alpha - 1) w_b\big]\cdot
+K_b(b-b_0)^2 $ and angle $ \big[1 + (\alpha - 1) w_\theta\big]\cdot
+\big[K_\theta (\theta - \theta_0)^2 +K_{\textrm{UB}} (S - S_0)^2] $ is
+no more than the vibrational frequency of the bond and angle
+potentials of water molecule, respectively.
 
@@
     2           & 1.00     & 0.0   & 1.00    & 0.0\\
-    5           & 1.00     & 0.0   & 0.80    & 0.2\\
-    6           & 1.00     & 0.0   & 0.67    & 0.2\\
+    4           & 1.00     & 0.0   & 0.88    & 0.2\\
+    % 5           & 1.00     & 0.0   & 0.80    & 0.2\\
+    % 6           & 1.00     & 0.0   & 0.67    & 0.2\\
+    8           & 1.00     & 0.0   & 0.50    & 0.2\\
     \hline\hline
@@
 
-\begin{figure}
-  \centering
-  \includegraphics[]{figs/fig-meta05.eps}
-  \includegraphics[]{figs/fig-meta05-1.eps}
-  \caption{$\alpha = 5$, the probability of different conformations as a function of time.
-    In the left plot: Red: $A_1$, green: $A_2$, blue: $B_1$, purple: $B_2$, cyan: $C$.
-    In the right plot: Red: $A_1+A_2$, blue: $B_1+B_2$, cyan: $C$.
-    The lighter colors denote the reference exact dynamics of the conformations ($\alpha = 1$), while the darker colors denote $\alpha = 5$.}
-  \label{fig:tmp2}
-\end{figure}
+% \begin{figure}
+%   \centering
+%   \includegraphics[]{figs/fig-meta05.eps}
+%   \includegraphics[]{figs/fig-meta05-1.eps}
+%   \caption{$\alpha = 5$, the probability of different conformations as a function of time.
+%     In the left plot: Red: $A_1$, green: $A_2$, blue: $B_1$, purple: $B_2$, cyan: $C$.
+%     In the right plot: Red: $A_1+A_2$, blue: $B_1+B_2$, cyan: $C$.
+%     The lighter colors denote the reference exact dynamics of the conformations ($\alpha = 1$), while the darker colors denote $\alpha = 5$.}
+%   \label{fig:tmp2}
+% \end{figure}
 
@@
   \includegraphics[]{figs/fig-meta08-1.eps}
-  \caption{$\alpha = 6$, the probability of different conformations as a function of time.
+  \caption{$\alpha = 8$, the probability of different conformations as a function of time.
     In the left plot: Red: $A_1$, green: $A_2$, blue: $B_1$, purple: $B_2$, cyan: $C$.
@@
   \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-04-5.eps}\\
-  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-05-1.eps}
-  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-05-2.eps}
-  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-05-3.eps}
-  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-05-4.eps}
-  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-05-5.eps}\\
+  % \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-05-1.eps}
+  % \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-05-2.eps}
+  % \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-05-3.eps}
+  % \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-05-4.eps}
+  % \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-05-5.eps}\\
   \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-08-1.eps}
@@
   \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-08-5.eps}\\
-  \caption{The integrated probability fluxes. $\alpha = 2,4,5,8$.}
+  \caption{The integrated probability fluxes. $\alpha = 2,4,8$.}
   \label{fig:tmp3}
@@
   \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-04-1-3.eps}\\
-  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-05-1-1.eps}
-  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-05-1-2.eps}
-  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-05-1-3.eps}\\
+  % \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-05-1-1.eps}
+  % \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-05-1-2.eps}
+  % \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-05-1-3.eps}\\
   \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-08-1-1.eps}
@@
   \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-08-1-3.eps}
-  \caption{The integrated probability fluxes. $\alpha = 4,5,8$.}
+  \caption{The integrated probability fluxes. $\alpha = 4,8$.}
   \label{fig:tmp3}

### 2013-10-30 accelerate/template.alanine.pme/parameters.sh
better parameters

@@ -2,13 +2,13 @@
 
 # non-equilibrium settings
 pert_equi_result="$HOME/study/noneq.msm/accelerate/equi.systems/alanine.charmm.pme.tab/"
-pert_num_conf_use=100000	#
+pert_num_conf_use=2500	#
 pert_mode=2			# 1: relax, 2: cos wave
 pert_strength=1.0		# nm/ps velocity
 pert_warm_time=40		# ps 1: warm time, 2: periodicity
 pert_shift=0			# unitless
 pert_phi=270			# deg. shift of the phase
-pert_time=100			# ps
+pert_time=500			# ps
 pert_frame_feq=0.5		# ps
 pert_dt=0.0005			# ps
 pert_rlist=1.33			# nm
@@ -26,7 +26,7 @@ pert_taup=2.0			# ps
 
 pert_rescale=1.0		# ps
 pert_bond_T_cap=8.631997e-15	# s taken from water
-pert_angle_T_cap=6.955613e-15	# s taken from water
+pert_angle_T_cap=17.037705e-15	# s taken from water
 pert_ele_rescale_scale0=1.0	# 
 pert_vdw_rescale_scale0=1.0	# 
 

### 2013-10-30 accelerate/template.alanine.pme/tools/angles/app/average.traj.cpp
add block average for calculating the meta sets

@@ -47,7 +47,7 @@ bool myread (FILE * fp,
 int main(int argc, char * argv[])
 {
   std::string ifile, ofile;
-  unsigned numBlock = 20;
+  // unsigned numBlock = 20;
   double refh;
   float time_prec = .01;
   

### 2013-10-30 accelerate/template.alanine.pme/tools/angles/app/average.traj.metastable.cpp

@@ -14,9 +14,13 @@
 
 #include "Defines.h"
 #include "Distribution.h"
+#include "BlockAverage.h"
+
 #include <boost/program_options.hpp>
 namespace po = boost::program_options;
+
 using namespace std;
+
 #define MaxLineLength 2048
 
 bool myread (FILE * fp,
@@ -71,14 +75,14 @@ void depositMetastable (const double & phi,
 
 int main(int argc, char * argv[])
 {
-  std::string ifile, ofile;
+  std::string ifile, ofile, oefile;
   // unsigned numBlock = 20;
   // double refh;
   // float time_prec = .01;
-  double setA_psi_b = 128, setA_psi_e = 13;
-  double setA_phi_b =-125, setA_phi_e = 74;
-  double setB_psi_b = 128, setB_psi_e = 13;
-  double setB_phi_b = 74,  setB_phi_e =-125;
+  // double setA_psi_b = 128, setA_psi_e = 13;
+  // double setA_phi_b =-125, setA_phi_e = 74;
+  // double setB_psi_b = 128, setB_psi_e = 13;
+  // double setB_phi_b = 74,  setB_phi_e =-125;
   double setC_psi_b = 13,  setC_psi_e = 128;
   double setC_phi_b = -180,setC_phi_e = 180;
   
@@ -97,7 +101,8 @@ int main(int argc, char * argv[])
       ("help,h", "print this message")
       ("output,o", po::value<std::string > (&ofile)->default_value ("metastable.out"), "the output of metastable propulation")
       ("input,f",  po::value<std::string > (&ifile)->default_value ("angle.name"), "the file of file names");
-
+  oefile = ofile + string(".error");
+  
   po::variables_map vm;
   po::store(po::parse_command_line(argc, argv, desc), vm);
   po::notify (vm);
@@ -128,7 +133,9 @@ int main(int argc, char * argv[])
   ValueType phi, psi;
   vector<float > times;
   vector<vector<double > > counts;
+  vector<vector<BlockAverage_acc > > bas;
   unsigned countFile = 0;
+  unsigned nDataInBlock = 5;
 
   while (fpname.getline(nameline, MaxLineLength)){
     if (nameline[0] == '#') continue;
@@ -140,11 +147,17 @@ int main(int argc, char * argv[])
     }
     countFile ++;
[…]

### 2013-10-30 accelerate/template.alanine.pme/tools/angles/include/BlockAverage.h

@@ -1,0 +1,55 @@
+#ifndef __BlockAverage_h_wanghan__
+#define __BlockAverage_h_wanghan__
+
+#include <vector>
+
+class BlockAverage 
+{
+  double avg;
+  double var;
+  double error_avg;
+  double error_var;
+  unsigned nBlock;
+  unsigned nDataInBlock;
+  unsigned nDataUsed;
+private:
+  void calAvg (const std::vector<double > & vec,
+	       double & my_avg,
+	       double & my_error_avg);
+public:
+  void processData (const std::vector<double > & vec,
+		    const unsigned & nBlock);
+  double getAvg () const {return avg;}
+  double getVar () const {return var;}
+  double getAvgError () const {return error_avg;}
+  double getVarError () const {return error_var;}
+  unsigned getNumDataUsed () const {return nDataUsed;}
+};
+
+
+class BlockAverage_acc
+{
+  double avg;
+  double error_avg;
+  unsigned nDataInBlock;
+private:
+  double sum;
+  double blockSum2;
+  unsigned nBlock;
+  double currentBlockSum;
+  unsigned currentNDataInBlock;
+public:
+  BlockAverage_acc (const unsigned & nDataInBlock_);
+  void clear ();
+  void reinit (const unsigned & nDataInBlock_);
+public:
+  void deposite (const double & vv);
+  void calculate ();  
+  double getAvg () const {return avg;}
+  double getAvgError () const {return error_avg;}  
+  unsigned getNumDataUsed () const {return nBlock * nDataInBlock;}
+}
+    ;
+
+
+#endif

### 2013-10-30 accelerate/template.alanine.pme/tools/angles/app/average.traj.metastable.cpp
fix bugs

@@ -100,8 +100,8 @@ int main(int argc, char * argv[])
   desc.add_options()
       ("help,h", "print this message")
       ("output,o", po::value<std::string > (&ofile)->default_value ("metastable.out"), "the output of metastable propulation")
+      ("output-error", po::value<std::string > (&oefile)->default_value ("metastable.err.out"), "the output of metastable propulation with error")
       ("input,f",  po::value<std::string > (&ifile)->default_value ("angle.name"), "the file of file names");
-  oefile = ofile + string(".error");
   
   po::variables_map vm;
   po::store(po::parse_command_line(argc, argv, desc), vm);
@@ -147,17 +147,17 @@ int main(int argc, char * argv[])
     }
     countFile ++;
     vector<double > tmpcount;
-    vector<BlockAverage_acc > tmpba;
     if (countFile == 1){
       while (myread(fp, time, phi, psi)){
 	times.push_back (time);
 	depositMetastable (psi, phi, sets, tmpcount);
 	counts.push_back (tmpcount);
+	vector<BlockAverage_acc > tmpba (tmpcount.size());
 	for (unsigned ii = 0; ii < tmpcount.size(); ++ii){
-	  BlockAverage_acc tmptmpba (nDataInBlock);
-	  tmptmpba.deposite (tmpcount[ii]);
-	  tmpba.push_back (tmptmpba);
+	  tmpba[ii].reinit(nDataInBlock);
+	  tmpba[ii].deposite (tmpcount[ii]);
 	}
+	bas.push_back (tmpba);
       }
     }
     else {
@@ -180,6 +180,7 @@ int main(int argc, char * argv[])
 
   FILE * fp = fopen (ofile.c_str(), "w");
   FILE * fpe = fopen (oefile.c_str(), "w");
+  // cout << "print to " << oefile << endl;
   for (unsigned ii = 0; ii < times.size(); ++ii){
     fprintf (fp, "%f ", times[ii]);
     fprintf (fpe, "%f ", times[ii]);

### 2013-10-30 accelerate/template.alanine.pme/tools/angles/include/BlockAverage.h

@@ -39,7 +39,8 @@ private:
   double currentBlockSum;
   unsigned currentNDataInBlock;
 public:
-  BlockAverage_acc (const unsigned & nDataInBlock_);
+  BlockAverage_acc ();
+  BlockAverage_acc (const unsigned nDataInBlock_);
   void clear ();
   void reinit (const unsigned & nDataInBlock_);
 public:

### 2013-10-30 accelerate/template.alanine.pme/tools/angles/src/BlockAverage.cpp

@@ -69,7 +69,13 @@ reinit (const unsigned & nDataInBlock_)
 }
 
 BlockAverage_acc::
-BlockAverage_acc (const unsigned & nDataInBlock_)
+BlockAverage_acc ()
+{
+  reinit (1);
+}
+
+BlockAverage_acc::
+BlockAverage_acc (const unsigned nDataInBlock_)
 {
   reinit (nDataInBlock_);
 }

### 2013-12-20 writedown/first.hitting/cross_entropy/cross_entropy.tex
add draft, update result

\documentclass[final]{siamltex}
%\usepackage{CJK}
%\usepackage{CJKnumb}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{fancyhdr}
\usepackage{indentfirst}
%\usepackage{titlesec}
\usepackage{placeins}
\usepackage{listings}
\usepackage{caption}
\usepackage{subfigure}
\usepackage{url}
\usepackage{bm}
\usepackage{dsfont}
\usepackage{comment}
\usepackage{titlesec}
\usepackage{multirow}
\usepackage{xcolor}
\usepackage{algorithm}
\usepackage{algpseudocode}
\renewcommand{\baselinestretch}{1.2}
\topmargin 0cm \oddsidemargin 0.66cm \evensidemargin 0.66cm
\textwidth 14.66cm \textheight 22.23cm
\parindent 5ex

\def\longrightharpoonup{\relbar\joinrel\rightharpoonup}
\def\longleftharpoondown{\leftharpoondown\joinrel\relbar}

\def\longrightleftharpoons{
  \mathop{
    \vcenter{
       \hbox{
       \ooalign{
          \raise1pt\hbox{$\longrightharpoonup\joinrel$}\crcr
  	  \lower1pt\hbox{$\longleftharpoondown\joinrel$}
	}
      }
    }
  }
}

\newcommand{\rates}[2]{\displaystyle \mathrel{\longrightleftharpoons^{#1\mathstrut}_{#2}}}
\newcommand{\cL}{{\mathcal L}}
\newcommand{\bR}{{\mathbb R}}
\newcommand{\argmin}{\mathop{\rm argmin}}%

\newcommand{\bE}{{\mathbf E}}

\newtheorem{remark}{Remark}
\titleformat{\subsubsection}[runin]{\normalfont\bfseries}{\thesubsection.}{3pt}{}

\begin{document}
\title{Applications of the cross-entropy method in diffusion processes}
%\author{Wei Zhang}
\author{Wei Zhang\textsuperscript{2}\and Han Wang\textsuperscript{1} \and
  Carsten Hartmann\textsuperscript{1,*}\and 
Marcus Weber\textsuperscript{2}\and Christof Sch\"{u}tte\textsuperscript{1,2}}
\footnotetext[1]{Institute of Mathematics, Freie Universit\"{a}t Berlin, Arnimallee 6, 14195 Berlin, Germany}
\footnotetext[2]{Zuse Institute Berlin, Takustrasse 7, 14195 Berlin, Germany}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\footnotetext[1]{Corresponding author. Email: chartman@mi.fu-berlin.de}
\date{}
\maketitle
\begin{abstract}
  The cross-entropy method in diffusion process setting is studied. The
  general algorithm is proposed and is applied to design efficient importance sampling for rare events as well as to solve optimal
  control problem. Several examples are studied numerically.
\end{abstract}
\begin{keywords}
importance sampling, optimal control, cross-entropy method, Monte Carlo
method, rare event, change of measure. 
\end{keywords}
\begin{AMS}
\end{AMS}
\section{Introduction}
\label{intro}
Diffusion process has been extensively investigated because of its wide
applications in mathematics, physics, economics, and so on.
We are mainly interested in two topics in diffusion process, i.e. importance sampling and optimal control. 
For importance sampling, it is one kind of variance reduction technique 
in Monte Carlo simulations and becomes especially necessary in rare event simulation.
For optimal control, the purpose is to driven the dynamical system in certain
optimal way and has strong application backgrounds in finance, climate
science, mechanics \cite{fleming2006}. We emphasize that these two different areas are actually related and can
be unified as the same problem, i.e. how to change the measure in certain optimal way.

Here we take an applied point of view, i.e. we aim at designing efficient
algorithms to find the way to do the importance sampling, or to find the
optimal control strategy. One main reason that hinder efficient algorithms is
due to the high dimensionality of the dynamical systems, as in the case of
climate system, biological system and so on. However, when the dynamical
systems have multiscale feature, efficient algorithms may be designed by
taking this feature into account. See \cite{ip-dupuis-multiscale, ip-eric,
zhws13} on importance sampling and \cite{control_schuette, zlph2013} on
optimal control for more details.

In this work, we consider these two topics together as designing the change of
measure in an optimal way. If we further assume the existence of the optimal
change of measure $\mu^*$, the problem becomes to find approximation of
$\mu^*$ under some measure distance.

On the other hand, the Kullback-Leibler divergence, although not a metric, is
widely used to quantified the similarity of two probability measures. 
What is more, given certain target probability measure $\mu^*$, the so-called cross-entropy method has been
developed recently and provides a general algorithm to find the minimizer of the
Kullback-Leibler divergence w.r.t. $\mu^*$, among a family of parameterized
probability measures. It has wide applications in different areas, such as
Monte Carlo simulation, optimization, and machine learning \cite{ce_book}. The
main purpose of this paper is to formulate the cross-entropy method in the
context of diffusion process, i.e. for measures defined in path space, and
discuss how to apply this method in importance sampling and optimal control
problems.

This paper is organized as follows. In Section~\ref{sec-om-functional},
using Girsanov theorem, we derive the Onsager-Machlup functional for a given trajectory.
In Section~\ref{sec-ce-is}, the cross-entropy method in path space is
formulated for the importance sampling problem.
In Section~\ref{sec-ocp}, by using the dual relationship between importance
sampling and optimal control problem, we discuss the cross-entropy method for
optimal control problem. The cross-entropy algorithm is summarized in
Section~\ref{sec-ce-algo}. Several numerical examples are studied in
Section~\ref{sec-examples}. Some discussions are presented in
Section~\ref{sec-discuss}.

\section{Onsager-Machlup functional}
\label{sec-om-functional}
Let $O \subset \mathbb{R}^n$, $z_s \in \mathbb{R}^n$ satisfies the SDE 
\begin{align}
\begin{split}
  d z_s &= b(z_s) ds + \sqrt{2\beta^{-1}} dw_s, \quad 0 < s < \tau  \\
  z_0&=x
  \end{split}
\label{dynamics-1}
\end{align}
$\beta >0$ is constant, $\tau = \inf_{s>0} \{s;\, z_s \in O\}$ is some
stopping time, $b : \mathbb{R}^n \rightarrow \mathbb{R}^n$, $w$ is a
$n$-dimensional Brownian motion.

We derive the Onsager-Machlup functional for dynamics (\ref{dynamics-1}) in
this section.
Let $\mu$ and $\mu_0$ be the probability measures on path space generated by
(\ref{dynamics-1}) and the scaled Brownian motion
\begin{align}
  \begin{split}
  dz_s &= \sqrt{2\beta^{-1}}dw_s, \hspace{3mm} 0 < s < \tau \\
   z_0 &= x .
  \end{split}
\end{align}
From Girsanov theorem \cite{oksendalSDE}, we have 
\begin{align}
  \frac{d\mu}{d\mu_0} &= \exp\Big(\sqrt{\frac{\beta}{2}}\int_0^\tau
  \langle b(z_s), dw_s\rangle + \frac{\beta}{4}\int_0^\tau |b(z_s)|^2ds\Big) \notag\\
  &=\exp\Big(\frac{\beta}{2}\int_0^\tau \langle b(z_s),dz_s\rangle -
  \frac{\beta}{4}\int_0^\tau |b(z_s)|^2ds\Big) 
\end{align}
and
\begin{align}
  d\mu_0 = \exp\Big(-\frac{\beta}{4}\int_0^\tau
  \Big|\frac{dz}{ds}\Big|^2ds\Big) d\nu\,
\end{align}
where $\nu$ is the Wiener measure. Together we obtain
\begin{align}
  d\mu &=\exp\Big(-\frac{\beta}{2} I_{1}(z)\Big)d\nu
  \label{measure-1}
\end{align}
where
\begin{align}
  I_{1}(z) = \frac{1}{2}\int_0^\tau \Big|\frac{dz}{ds} - b(z_s)\Big|^2 ds
\end{align}
and the integral is interpreted as Ito integral.
We could also rewrite it as Stratonovich integral with the relationship
\begin{align}
  \int_0^\tau \langle b(z_s), dz_s\rangle = \int_0^\tau b(z_s) \circ dz_s - \beta^{-1} \int_0^\tau \operatorname{div}\,b(z_s) ds
\end{align}
and therefore obtain 
\begin{align}
  d\mu = \exp \Big(-\frac{\beta}{2} I_2(z) \Big) d\nu\,
\end{align}
where functional 
\begin{align}
  I_2(z) = \frac{1}{2}\int_0^\tau \Big|\frac{dz}{ds}- b(z_s)\Big|^2 ds +
  \beta^{-1} \int_0^\tau \operatorname{div}\,b(z_s) ds
  \label{om-fun}
\end{align}
is often called Onsager-Machlup functional \cite{om1978, om_functional_pinski}. 
\begin{remark}
  The introduction of Onsager-Machlup functional and Stratonovich integral is
  not necessary. However, it is helpful to allow us to follow the presentation of
  \cite{ce_tutorial}, \cite{ce_book} and to compare with the case of random
  variables.
\end{remark}
\begin{remark}
  \label{remark-general}
  The Onsager-Machlup functional can also be generalized to the following dynamics.
  \begin{enumerate}
    \item
      SDEs with non-constant diffusion coefficients. We consider 
      \begin{align}
	\vspace{-9mm}
    \begin{split}
      d z_s &= b(z_s) ds + \sqrt{2\beta^{-1}} \sigma(z_s)\,dw_s, \quad 0 < s < \tau \\
      z_0&=x
    \end{split}
    \label{dynamics-1-sigma}
    \end{align}
    where $a(z) = \sigma(z)\sigma(z)^T$ is $n \times n$ positive definite matrix. We have 
    $d\mu = \exp(-\frac{\beta}{2} I_1(z)) d\nu$, with 
    \begin{align*}
      I_1(z) = \frac{1}{2} \int_0^\tau \Big|\frac{dz}{ds} - b(z_s)\Big|_a^2 ds
    \end{align*}
     where Ito integral is used and $|v|_a = \langle v, a^{-1}v\rangle$,
     $\forall v \in \mathbb{R}^n$.
    \item
      System contains both stochastic and deterministic equations. For
      simplicity, here we consider dynamics 
      \begin{align}
      \begin{split}
	dx_s & = b_1(x_s, y_s)\, ds \\
	dy_s & = b_2(x_s, y_s)\, ds + \sqrt{2\beta^{-1}} dw_s
      \end{split}
      \label{ode-sde}
      \end{align}
      where $x_s \in \mathbb{R}^k, y_s, w_s \in \mathbb{R}^{n-k}$, $b_1 :
      \mathbb{R}^k \times \mathbb{R}^{n-k} \rightarrow \mathbb{R}^k$, $b_2 :
      \mathbb{R}^k \times \mathbb{R}^{n-k} \rightarrow
      \mathbb{R}^{n-k}$. We can still assign a probability measure $\mu$ on the
      trajectory space, i.e. $\forall z = (x,y) $ is a trajectory, we have 
      $d\mu = \exp(-\frac{\beta}{2} I_1(z)) d\nu$ (\textbf{or formally :}  $\mu(dz) =
      \exp(-\frac{\beta}{2} I_1(z)) \nu(dy)$, which is better ?), where $\nu$ is Wiener
      measure on $\mathbb{R}^{n-k}$ and 
      \begin{align*}
	I_1(z) = \frac{1}{2} \int_0^\tau \Big|\frac{dy}{ds} - b_2(x_s, y_s)
	\Big|^2 ds\,.
      \end{align*}
      Notice that only the stochastic equations are used.
  \end{enumerate}
\end{remark}
\begin{comment}
\begin{remark}
  Since $I_2$ is simply the Stratonovich version of $I_1$, we use $I_1$ for
  simplicity.
  Starting from dynamics
\begin{align}
\begin{split}
  d z_s &= \widetilde{b}(z_s) ds + \sqrt{2\beta^{-1}} d\widetilde{w}_s, \quad s >0  \\
  z_0&=z
  \end{split}
\label{dynamics-11}
\end{align}
the above argument can be repeated to get 
\begin{align}
  d\widetilde{\mu} &=\exp\Big(- \frac{\beta}{4}\int_0^\tau \Big|\frac{dz}{ds} -
  \widetilde{b}(z_s)\Big|^2 ds\Big) d\nu
  \label{measure-2}
\end{align}
Divided it by (\ref{measure-1}), we have
\begin{align}
  \frac{d\widetilde{\mu}}{d\mu} = \exp\Big(- \frac{\beta}{4}\Big[2\int_0^\tau (dz_s,
  b(z_s) - \widetilde{b}(z_s)) + \int_0^\tau (|\widetilde{b}(z_s)|^2 - |b(z_s)|^2) ds\Big]
  \Big)
\end{align}
Then we substitute the dynamics (\ref{dynamics-1}) in above (in principle we
can also use dynamics (\ref{dynamics-11}), but $\widetilde{w}$ is not Brownian motion
under $\mu$ ...), after
simplification, we have 
\begin{align}
  \frac{d\widetilde{\mu}}{d\mu} = \exp\Big(-
  \frac{\beta^{1/2}}{\sqrt{2}}\int_0^\tau (dw_s,
b(z_s) - \widetilde{b}(z_s)) - \frac{\beta}{4}\int_0^\tau (|\widetilde{b}(z_s)
- b(z_s)|^2) ds\Big)
\end{align}
Let $u_s = \frac{1}{\sqrt{2}} (b(z_s) - \widetilde{b}(z_s))$, it gives
\begin{align}
  \frac{d\widetilde{\mu}}{d\mu} = \exp\Big(- \beta^{1/2}\int_0^\tau (dw_s, u_s)
) - \frac{\beta}{2}\int_0^\tau |u_s|^2 ds\Big)
\end{align}
So we recover the more general situation in our IS paper for $\sigma =
\sqrt{2}$.
\end{remark}
\end{comment}
\section{Cross entropy method for importance sampling}
\label{sec-ce-is}
In this section, we discuss how to apply the cross entropy
method \cite{ce_tutorial, ce_book} to guide the
importance sampling in diffusion process.

Let $\mathcal{D}$ be certain subset in the path space and
$\mathds{1}_{\mathcal{D}}$ be the
characteristic function on $\mathcal{D}$. 
In the following, for simplicity, we will use $z$ to denote a path trajectory
$\{z_s\}$ and keep the notation $z_s \in \mathbb{R}^n$ for the state at time $s$. 
Let $F(\cdot)$ be certain path functional and will be written as $F(z)$ for
path $z$. 
We consider Monte Carlo method with importance sampling to compute the quantity
\begin{align}
  l = \mathrm{E}(F(z)~|~z_0 = x)
  \label{mean-l}
\end{align}
where we assume $F(\cdot)$ has the following type
\begin{align}
  F(z) = \mathds{1}_{\mathcal{D}}(z) \quad \mbox{or} \quad F(z) = \exp(-\beta\int_0^\tau G(z_s) \,ds).
  \label{path-fun-f}
\end{align}
\begin{remark}
  A special and interesting case is when $\mathcal{D} = \{z, \tau(z) \le \lambda\}$,
  $\lambda > 0$ is fixed. Then $l = \mathrm{E}(\mathds{1}_{\mathcal{D}}(z)) =
  \mathcal{P}(\tau \le \lambda)$, i.e. the probability of the trajectories
  stopping before time $\lambda$. This quantity may provide us some details of the
  distribution of the stopping time $\tau$ and help us understand the
  behaviors of the dynamics.
\end{remark}

We suppose $m$ basis functions on state space $\{\phi_i\}_{1\le i \le m}$ are
available, where $\phi_i : \mathbb{R}^n\rightarrow \mathbb{R}$. 
We will add a control force to the original dynamics (\ref{dynamics-1})
and the importance sampling will be realized by simulating trajectories from
this modified dynamics. Specifically, given $\bm{w} = (w_1, w_2,
\cdots, w_m)^T \in \mathbb{R}^m$,
the corresponding control force is state-dependent and is supposed to be 
\begin{align}
  u(x; \bm{w}) = \sum\limits_{i = 1}^{m} w_i \nabla \phi_i(x), \quad x \in
  \mathbb{R}^n\,.
  \label{control-force}
\end{align}
Therefore, for fixed parameter $\bm{w}$, the controlled dynamics is 
\begin{align}
\begin{split}
  d z_s &= b(z_s) ds + \sqrt{2}u(z_s;\bm{w}) ds + \sqrt{2\beta^{-1}} dw_s, \quad t \le s \le \tau \\
     z_0 &=z
  \end{split}
\label{dynamics-2}
\end{align}
Let $\mu_{\bm{w}}$ be the probability measure of dynamics (\ref{dynamics-2})
w.r.t. $\bm{w}$. 
From Section \ref{sec-om-functional},  we know $\mu_{\bm{w}}$ has density 
$f(z;\bm{w}) = \exp(-\frac{\beta}{2} I(z ; \bm{w}))$ w.r.t. Wiener
measure $\nu$, with
\begin{align}
  I(z ; \bm{w}) = \frac{1}{2}\int_0^\tau \Big|\frac{dz}{ds}- b(z_s) -
  \sqrt{2}\sum\limits_{i = 1}^{m} w_i \nabla \phi_i(z_s)\Big|^2 ds + \beta^{-1}
  \int_0^\tau \operatorname{div}\, \Big(b(z_s) + \sqrt{2}\sum\limits_{i =
  1}^{m} w_i \nabla \phi_i(z_s)\Big) ds
  \label{om-fun-w}
\end{align}
and the original dynamics (\ref{dynamics-1}) corresponds to $\bm{w} = \bm{0}$ (we use $\mu = \mu_{\bm{0}}$).
Given a probability measure $\mu_g$ on path space with $d\mu_g = g(z)d\nu$, 
we can design a unbiased Monte Carlo estimator with importance sampling method as 
\begin{align}
  l_N = \frac{1}{N} \sum_{i=1}^{N}  \frac{F(z_i)f(z_i;\bm{0})}{g(z_i)}
\end{align}
where trajectories $z_i, i = 1, \cdots, N$ are sampled from measure $\mu_g$.
The optimal measure $\mu_{g^*}$, which gives
zero-variance estimator, should satisfy 
\begin{align}
  g^*(z) = \frac{F(z)f(z;\bm{0})}{l}, \hspace{3mm} \forall z
  \label{optimal_pdf}
\end{align}
However, $g^*(z)$ depends on the path $z$ and is not easy to compute.

The idea of cross entropy method is to find the best approximation of $g^*(z)$ among
$f(z;\bm{w}), \bm{w} \in \mathbb{R}^m$. More specifically, define the
Kullback-Leibler divergence between
two measures $d\mu_1 = g_1d\nu$, $d\mu_2 = g_2d\nu$ as
\begin{align}
  D(\mu_1, \mu_2) = \mathrm{E}_{\mu_1} \log \frac{d\mu_1}{d\mu_2} = \int
  g_1\log g_1\, d\nu - \int g_1\log g_2\,d\nu,
  \label{kl-dist}
\end{align}
and we try to minimize
\begin{align}
  \min_{\bm{w} \in \mathbb{R}^m} D(\mu_{g^*}, \mu_{\bm{w}}).
  \label{kl-dist-d}
\end{align}
Using (\ref{optimal_pdf}) and (\ref{kl-dist}), it is equivalent to solve
\begin{align}
  \max_{\bm{w} \in \mathbb{R}^m} \mathrm{E}_{\mu} ( F(z) \log f(z;\bm{w}))
  = \max_{\bm{w} \in \mathbb{R}^m} \mathrm{E}_{\mu_{\bm{v}}} ( F(z)
  w(z;\bm{v}) \log f(z;\bm{w})), \hspace{3mm} \forall \bm{v} \in
  \mathbb{R}^m
  \label{ce-2}
\end{align}
where
\begin{align}
  w(z;\bm{v}) = \frac{f(z;\bm{0})}{f(z;\bm{v})}.
\end{align}
For numerical purpose, we consider the discrete formulation of (\ref{ce-2})
\begin{align}
\max_{\bm{w} \in \mathbb{R}^m} \frac{1}{N} \sum_{i=1}^{N} F(z_i) w(z_i, \bm{v}) \log f(z_i;\bm{w}) 
\label{ce-2-discrete}
\end{align}
where $z_i, i = 1, \cdots, N$ are trajectories generated from $\mu_{\bm{v}}$. 
Taking gradient, the maximizer of (\ref{ce-2-discrete}) could be found by solving
\begin{align}
  \frac{1}{N} \sum_{i=1}^{N} F(z_i) w(z_i, \bm{v}) \nabla_{\bm{w}} \log
  f(z_i;\bm{w}) = 0.
  \label{max-gradient}
\end{align}
With formula (\ref{om-fun-w}), equations (\ref{max-gradient}) could be written more explicitly
as linear equations $A\bm{w} = \bm{r}$, where $\forall 1\le i,j\le m$, 
\begin{align}
  \begin{split}
  A_{ij} &= \frac{1}{N}\sum_{k = 1}^{N} F(z_k)w(z_k, \bm{v}) \int_0^\tau
    \nabla\phi_i(z_k)\nabla\phi_j(z_k)\,ds, \\
     r_i &= \frac{1}{\sqrt{2}N}\sum_{k = 1}^{N} F(z_k)w(z_k, \bm{v})
    \Big[\int_0^\tau \Big(\frac{dz_k}{ds} - b(z_s)\Big)\nabla\phi_i(z_k)\,ds - \beta^{-1} \int_0^\tau \operatorname{div} \nabla \phi_i(z_k)\,ds\Big]
  \end{split}
  \label{linear-eqn}
\end{align}
Notice that (\ref{max-gradient}) holds $\forall \bm{v} \in \mathbb{R}^m$ and
$A$ is in general a positive definite matrix when $\phi_i$'s are linearly independent.
\begin{remark}
  The more general type conditional expectation 
  $l = \mathrm{E}(F(z)~|~ z_0 \sim \pi)$
  can be similarly considered,
where $\pi(dx) = \rho(x)dx$ is a probability measure on $\Omega
\subset \mathbb{R}^n$, i.e. $\int_\Omega \rho(x)\,dx = 1$. Let $E = \{z ~|~ z \in
  C([0,\infty), \mathbb{R}^n), z_0 \in \Omega\}$ be the set of trajectories
  starting from $\Omega$. Denote $\mu^x$ as the aforementioned measure for
  dynamics (\ref{dynamics-1}) starting from $x \in \mathbb{R}^n$. Further
  define $\mu^\pi(\mathcal{D}) = \int_{\Omega} \mu^x(\mathcal{D}) \pi(dx)$,
  where $\mathcal{D}$ is a measurable subset of $E$.
  We have $d\mu^\pi_{\bm{v}} = f(z;\bm{v})\,d\nu^\pi, \forall
  \bm{v} \in \mathbb{R}^n$, where $\mu^\pi_{\bm{v}}$ and $\nu^{\pi}$ are
  defined similarly. The aim is to find the optimal $\bm{w} \in
  \mathbb{R}^m$ and the corresponding measure $\mu^\pi_{\bm{w}}$ to do
  the importance sampling. All the derivations are the same with the above
  except that, while simulating the trajectories, 
  the starting point $z_0$ should be generated from $z_0 \sim \pi$.
\end{remark}
\begin{remark}
   The above algorithm can be further applied to dynamics (\ref{ode-sde}), discussed
   in Remark \ref{remark-general}. In this case, since only stochastic
   equations appear in the Radon-Nikodym derivative, the control should
   also be only added to these equations, in order to obtain linear equations
   with full rank. Specifically, we study the controlled dynamics 
         \begin{align}
      \begin{split}
	dx_s & = b_1(x_s, y_s)\, ds \\
	dy_s & = b_2(x_s, y_s)\, ds + \sqrt{2} u(x_s, y_s ; \bm{w})\, ds + \sqrt{2\beta^{-1}} dw_s
      \end{split}
      \label{ode-sde-control}
      \end{align}
      with $u(x_s, y_s ; \bm{w}) = \sum\limits_{i=1}^{m} w_i \nabla \phi_i(x_s,
      y_s)$ and optimize the coefficients $\bm{w}$. All the derivations are
      the same as above.
\end{remark}
\section{Cross entropy method for optimal control problem}
\label{sec-ocp}
In this section, we consider applying the cross entropy method to solve optimal control problem
for diffusion process.

Consider the optimal control problem with cost function
\begin{equation}
  J(u) = \mathrm{E}_{\mu_u}\left[\int_0^\tau \left(G(z_s) + \frac{1}{2} |u(z_s,s)|^2
\right) ds ~\bigg|~ z_0 = x\right].
\end{equation}
where $G : \mathbb{R}^n\rightarrow \mathbb{R}$ is nonnegative, $u$ is the
control force and $\mathrm{E}_{\mu_u}$ denotes the expectation under probability measure $\mu_u$,
which is generated by the controlled dynamics
of (\ref{dynamics-1-sigma}), namely,
\begin{align}
 \begin{split}
   d z_s &= b(z_s) ds + \sqrt{2}\sigma(z_s) u(z_s, s) ds + \sqrt{2\beta^{-1}} \sigma(z_s)\,dw_s, \quad 0 < s < \tau \\
    z_0&=x
  \end{split}
  \label{dynamics-1-sigmma-c}
\end{align}

Let $\hat{u} = \argmin J(u)$. For simplicity, we denote $\mu, \hat{\mu}$ as
the probability measure corresponding to $u \equiv 0$ and $\hat{u}$ respectively. Using the dual relation
    \begin{align}
      -\beta^{-1}\log \mathrm{E}_\mu\left[\exp\Big(-\beta\int_0^\tau G(z_s)\, ds\Big)\right] = \inf_u J(u) = J(\hat{u})
      \label{dual-relation}
    \end{align}
    and Jensen's inequality, we know that 
    \begin{align}
    \exp\Big(-\beta\int_0^\tau G(z_s)\,ds\Big)\frac{d\mu}{d\hat{\mu}} =
    \mathrm{E}_\mu\left[\exp\Big(-\beta\int_0^\tau G(z_s)\, ds\Big)\right],
    \hspace{3mm} \hat{\mu}-a.s.
    \label{jensen-as}
  \end{align}
From the above we have 
  \begin{align}
    J(u) &= \mathrm{E}_{\hat{\mu}} \left\{\Big[\int_0^\tau \Big(G(z_s) + \frac{1}{2} |u(z_s,
    s)|^2\Big)\,ds\Big]\frac{d\mu_u}{d\hat{\mu}}\right\}  \\
    & = J(\hat{u}) + \mathrm{E}_{\hat{\mu}}\Big[\Big(\beta^{-1}\log
    \frac{d\mu}{d\hat{\mu}} + \int_0^\tau \frac{1}{2} |u(z_s,
    s)|^2\,ds\Big)\frac{d\mu_u}{d\hat{\mu}}\Big]
\end{align}
After simplification, we get the following simple relationship 
\begin{align}
J(u) = J(\hat{u}) + \beta^{-1}D(\mu_u|\, \hat{\mu}).
\label{J-entropy}
\end{align}
where $D(\cdot, \cdot)$ is the Kullback-Leibler divergence in (\ref{kl-dist}).

We are interested in finding the optimal control $\hat{u}$. For high
dimensional dynamics ($n$ is large), this becomes difficult.  
Similarly as the in previous section, we suppose the control force can be written
as (\ref{control-force}) with $m$ predefined basis functions $\{\phi_i\}_{1\le
i \le m}$. And our problem is simplified to solve
\begin{align}
  \min_{\bm{w} \in \mathbb{R}^m} J(u(z;\bm{w})).
  \label{J-w}
\end{align}
under dynamics (\ref{dynamics-1-sigmma-c}).

From (\ref{J-entropy}), we know that solving (\ref{J-w})
is equivalent to minimize the Kullback-Leibler divergence $D$ between
$\mu_u$ and $\hat{\mu}$, which, however, is still not easy. On the other hand, inspired by 
the discussions in Section \ref{sec-ce-is}, cross entropy method can be
applied to minimize $D(\hat{\mu} \,|\, \mu_u)$ (notice this is different with
(\ref{J-w}), because the Kullback-Leibler divergence is not symmetric). Following
the notations in Section \ref{sec-ce-is} and from (\ref{jensen-as}), we know the optimal measure
satisfies
\begin{align}
  g^*(z) \propto \exp\Big(-\beta\int_0^\tau G(z_s) \,ds\Big) f(z ; 0),
  \quad \forall z\,,
\end{align}
therefore the problem is reduced to the case in Section \ref{sec-ce-is} with $F(z) = \exp(-\beta\int_0^\tau G(z_s) \,ds)$.
\section{Cross entropy algorithm}
\label{sec-ce-algo}
In this section, we consider how to implement the cross entropy algorithm in
order to maximize (\ref{ce-2-discrete}), which is equivalent to (\ref{kl-dist-d}) in the discrete case. 
We suppose the optimal density $g^*(z)$ is given by (\ref{optimal_pdf}) and
$F(z)$ takes the forms in (\ref{path-fun-f}).

Ideally, the solution of the maximization problem (\ref{ce-2-discrete}) can be
obtained by solving the linear equation (\ref{linear-eqn}) without iterations. But in real
applications, when the expectation value $l$ of $F(z)$ in (\ref{mean-l}) is very rare, it will become
difficult to estimate the coefficients (\ref{linear-eqn}) and therefore the solution
will be inaccurate. In diffusion processes, one major reason for this rarity
is the metastability in the dynamics when $\beta \gg 1$. In this case, the
trajectories are long and $F(z)$ is small, which means computing
(\ref{linear-eqn}) is both time-consuming and inaccurate. Inspired by the
original cross entropy method \cite{ce_tutorial},
we may overcome this problem by starting from a higher temperature (smaller $\beta$) and compute
(\ref{linear-eqn}) while decreasing the temperature (increasing $\beta$). The proposed iterative
method to solve the maximization problem (\ref{ce-2-discrete}) is illustrated in Algorithm \ref{ce-algo}.
\begin{algorithm}
  \caption{Algorithm to solve the maximization problem (\ref{ce-2-discrete}) \label{ce-algo}}
  \begin{algorithmic}[1]
    \State Define $0 < \beta_0 < \beta_1 < \cdots <\beta_k = \beta$, set
    $\bm{v}^{(0)} = 0$.
    \For {$j=0$ to $k$}
    \State generate $N_j$ trajectories $z_i, i=1,2,\cdots, N$ from dynamics (\ref{dynamics-2}), with
    $\bm{w} = \bm{v}^{(j)}$, $\beta = \beta_j$.
    \State compute the coefficients of $A^{(j)}, \bm{r}^{(j)}$ from (\ref{linear-eqn}) with $\bm{v} =
    \bm{v}^{(j)}$, and solve the linear equations $A^{(j)}\bm{v}^{(j+1)} =
    \bm{r}^{(j)}$.
    \EndFor 
  \end{algorithmic}
  \label{algo1}
\end{algorithm}
\FloatBarrier
\section{Numerical examples}
\label{sec-examples}
In this section, we will study the cross entropy method with some concrete dynamics and
illustrate some numerical results.
\subsection{One-dimensional double well potential}
Here we consider the simplest 1D diffusion process with double well potential.
Let $b(x) = - V'(x), x\in \mathbb{R}^1$, with $V(x) = \frac{1}{2}(x^2 - 1)^2$.
The dynamics is 
\begin{align}
  dz_s = - V'(z_s) ds + \sqrt{2\beta^{-1}} dw_s
\end{align}
where $w_s$ is 1D Brownian motion. The potential $V$ has two minima points
$x_0 = -1, x_2 = 1$, while $x_1 = 0$ is
saddle point. Define $\tau = \inf\{s >0, |z_s - x_2| \le 1 \}$, and choose
the starting point $z_0 = x_0 =
-1$. We define three basis functions of Gaussian type as $\phi_i(x) = \exp(-\frac{(x -
x_i)^2}{2r^2})$, $i = 0, 1, 2$, and choose $r = 0.5, \beta = 5.0$ (see Figure. \ref{fig-ex1-basis}).

We first study the optimal control problem  
\begin{align}
  \inf_{u_s} \mathrm{E} \Big(\tau + \frac{1}{2} \int_0^\tau u_s^2 \,ds~\Big|~
  z_0 = -1\Big)\,,
  \label{ex1-cost}
\end{align}
under dynamics
\begin{align}
  dz_s = - V'(z_s) ds + \sqrt{2}u_s ds + \sqrt{2\beta^{-1}} dw_s\,.
\end{align}
$\mbox{dt} = 1.0 \times 10^{-3}$, $N_j = 10^{4}, j=0,1,2$ is used to compute the
coefficients $\bm{w}$ using Algorithm.\ref{ce-algo}, with $\{\beta_0,\beta_1, \beta_2\}
= \{1.0, 3.0, 5.0\}$.  Therefore $k=2$ in Algorithm.\ref{ce-algo} and $3$
iteration steps are performed. Denote $\bm{v}^{(0)} = \bm{0}$ and suppose
coefficients $\bm{v}^{(j+1)}$ are obtained
in $j$th steps, $j=0,1,2$. Then applying control force $u(x;\bm{v}^{(j)})$ is
equivalent to modifying the potential as $V^{(j)}(x) = V(x) - \sqrt{2}\sum\limits_{i=0}^{2} v^{(j)}_i
\phi_i(x)$. The numerical results are presented in Figure.\ref{fig-ex1-1} and Figure.\ref{fig-ex1-2}.
In Figure.\ref{fig-ex1-1}, we show the modified potential $V^{(j)}(x), 0\le
j\le 3$ when different basis functions are used. We can observe that the potential is
gradually modified during the iterations. In Figure.\ref{fig-ex1-2}, it shows
the modified potentials $V^{(3)}(x)$ with optimized coefficients
$\bm{v}^{(3)}$ in Algorithm.\ref{ce-algo} for different basis functions. In
one-dimension, the optimal control problem (\ref{ex1-cost}) can also be solved
from a Feymann-Kac type boundary value PDE \cite{control_schuette}. The
potential (the cost value) obtained in this way can be considered as exact solution and is also
shown in Figure.\ref{fig-ex1-2} (The last row of Table.\ref{tab-ex1-1}).
Later on, $N = 10^{6}$ samples are generated to compute
the cost function (\ref{ex1-cost}) and the results are presented in
Table.\ref{tab-ex1-1}. We can observe that better results are obtained when basis function
$\phi_1$ is used, since in job 2 and 3, the cost value (1.31) is closer to the exact
solution (1.25). Notice that this fact is also shown in
Figure.\ref{fig-ex1-2}, as the potential profiles with Label ``basis 0, 1'' and
Label ``basis 1'' are more similar to the profile with Label ``exact'', the
optimal potential. (Notice that only the region $x<0$ is meaningful and the dynamics remains the same when
the potential is shifted vertically) 
\begin{figure}[tphb]
  \centering
\includegraphics[width=7cm]{./fig/basis_1d.eps}
\caption{One-dimensional basis functions $\phi_0, \phi_1, \phi_2$. \label{fig-ex1-basis}}
\end{figure}
\begin{figure}[tphb]
  \centering
  \begin{tabular}{ll}
    \subfigure[basis $\phi_0$]{\includegraphics[width=5.0cm]{./fig/pot_1d_0_converge_control.eps}}
&
    \subfigure[basis $\phi_1$]{\includegraphics[width=5.0cm]{./fig/pot_1d_1_converge_control.eps}} \\
    \subfigure[basis $\phi_0,\phi_1$]{\includegraphics[width=5.0cm]{./fig/pot_1d_2_converge_control.eps}}
    &
    \subfigure[basis $\phi_0,\phi_2$]{\includegraphics[width=5.0cm]{./fig/pot_1d_3_converge_control.eps}}
\end{tabular}
\caption{One-dimensional dynamics. Convergence results for the modified
potentials $V^{(j)}, j=0,1,2,3$, with different sets of basis functions\label{fig-ex1-1}}
\end{figure}
\begin{figure}[tphb]
  \centering
    \includegraphics[width=7.0cm]{./fig/pot_1d_basis_control.eps}
    \caption{One-dimensional dynamics. Optimized potential $V^{(3)}$ with
    different sets of basis functions. The curve with Label ``exact'' is
  computed directly by solving a Feymann-Kac boundary value PDE. \label{fig-ex1-2}}
\end{figure}
\begin{table}
  \centering
  \begin{tabular}{c|c|c|c|c}
    \hline
    job & coefficients & cost & relative variance & mean of $\tau$ \\
    \hline
       $1$ & $(-1.771, 0, 0)$ & $5.14$ & $1.35$ & $2.02$  \\
    \hline
       $2$ & $(0, 1.857, 0)$ & $1.31$ & $0.44$ & $0.57$  \\
    \hline
       $3$ & $(-0.110, 1.762, 0)$ & $1.31$ & $0.44$ & $0.57$  \\
    \hline
       $4$ & $(-1.611, 0, 1.379)$ & $3.88$ & $1.22$ & $1.73$  \\
    \hline
       $5$ & $$ & $1.25$ & $0.30$ & $0.52$  \\
    \hline
  \end{tabular}
  \caption{Results of the optimal control problem for the one-dimensional dynamics with different sets
  of basis functions. The coefficients are solved by Algorithm.\ref{ce-algo}.
Job $1$-$4$ are the results with different sets of functions $\{\phi_0\}$,
$\{\phi_1\}$, $\{\phi_0, \phi_1\}$, $\{\phi_0, \phi_2\}$ respectively. The
costs are the exectation (\ref{ex1-cost}) computed with $N=10^6$ samples. The
last column is the mean of the stopping time $\tau$ under the modified
dynamics. The last row is the computed by solving a Feymann-Kac type PDE and
is considered as the ``accurate'' solution.\label{tab-ex1-1}}
\end{table}

We continue to study how to compute $P(\tau \le \lambda)$ for $\lambda > 0$.
We choose $\lambda = 2.0, 1.0, 0.5, 0.3, 0.1$ and two sets of basis functions are used : $\{\phi_1\}$
and $\{\phi_2\}$. We compare the results using importance sampling with the
standard Monte Carlo method. These results are shown in
Table.\ref{tab-ex1-2}--\ref{tab-ex1-4} and Figure.\ref{fig-ex1-3}.
Table.\ref{tab-ex1-2} and Table.\ref{tab-ex1-3} shows the results of
importance sampling with basis function $\{\phi_0\}$ and $\{\phi_1\}$,
respectively, while Table.\ref{tab-ex1-4} is the result using standard Monte Carlo method.
In each case, we can observe that the estimators' relative variance will increase when $\lambda$ becomes
small. However, for fixed $\lambda$, the relative variances of importance
sampling are largely reduced comparing to the standard Monte Carlo method. And this becomes
more apparent when $\lambda$ decreases, while the standard Monte Carlo method
may fail to work (see the last row of Table.\ref{tab-ex1-4}).
In Figure.\ref{fig-ex1-3}, the optimized potentials used to perform importance sampling
are ploted for different $\lambda$ and different sets of basis functions. As
$\lambda$ decreases,
the optimized potentials become more different from the original
potential, which indicates larger forces are needed.

\begin{table}
\centering
  \begin{tabular}{c|c|c|c|c|c}
    \hline
   $\lambda$ & coefficients & $P(\tau \le \lambda)$ & relative variance & $R_e$ & time (sec.) \\
    \hline
  $2.0$ & $(-0.837, 0, 0)$ & $9.22 \times 10^{-2}$ & $1.45$ & $6.1 \times 10^{-1}$ & $70.4$ \\
    \hline
    $1.0$ & $(-1.391, 0, 0)$ & $3.23 \times 10^{-2}$ & $1.48$ & $4.8 \times 10^{-1}$ & $39.4$ \\
    \hline
    $0.5$ & $(-2.221, 0, 0)$ & $6.45 \times 10^{-3}$ & $1.75$ & $3.9 \times 10^{-1}$ & $21.5$ \\
    \hline
    $0.3$ & $(-3.282, 0, 0)$ & $9.51 \times 10^{-4}$ & $2.30$ & $3.3 \times 10^{-1}$ & $13.6$ \\
    \hline
    $0.2$ & $(-4.553, 0, 0)$ & $9.53 \times 10^{-5}$ & $3.10$ & $2.7 \times 10^{-1}$ & $9.4$ \\
    \hline
    $0.1$ & $(-8.245, 0, 0)$ & $1.22 \times 10^{-7}$ & $6.84$ & $1.6 \times
 10^{-1}$ & $4.9$ \\
    \hline
  \end{tabular}
  \caption{ $\mathcal{P}(\tau \le \lambda)$ for one-dimensional dynamics.
  $N=10^6$ samples and importance sampling are used with basis function $\{\phi_0\}$. $R_e$ is the
ratio of trajectories satisfying $\tau \le \lambda$ under the modified dynamics. \label{tab-ex1-2}}
\end{table}
\begin{table}
  \centering
  \begin{tabular}{c|c|c|c|c|c}
    \hline
    $\lambda$ & coefficients & $P(\tau \le b)$ & relative variance & $R_e$ & time (sec.) \\
    \hline
  $2.0$ & $(0, 0.962, 0)$ & $9.23 \times 10^{-2}$ & $1.04$ & $8.5 \times 10^{-1}$ & $56.1$\\
    \hline
  $1.0$ & $(0, 1.498, 0)$ & $3.23 \times 10^{-2}$ & $1.00$ & $8.1 \times 10^{-1}$ & $32.7$ \\
    \hline
  $0.5$ & $(0, 2.314, 0)$ & $6.46 \times 10^{-3}$ & $1.33$ & $6.8 \times 10^{-1}$ & $19.4$ \\
    \hline
  $0.3$ & $(0, 3.338, 0)$ & $9.49 \times 10^{-4}$ & $1.84$ & $5.6 \times 10^{-1}$ & $12.8$ \\
    \hline
  $0.2$ & $(0, 4.578, 0)$ & $9.56 \times 10^{-5}$ & $2.54$ & $4.6 \times 10^{-1}$ & $9.02$\\
    \hline
  $0.1$ & $(0, 8.232, 0)$ & $1.21 \times 10^{-7}$ & $5.28$ & $2.6 \times 10^{-1}$ & $4.8$\\
    \hline
  \end{tabular}
  \caption{ $\mathcal{P}(\tau \le \lambda)$ for one-dimensional dynamics.
  $N=10^6$ samples and importance sampling are used with basis function $\{\phi_1\}$. $R_e$ is the ratio of trajectories satisfying $\tau \le \lambda$ under the modified dynamics. \label{tab-ex1-3}}
\end{table}
\begin{table}
  \centering
  \begin{tabular}{c|c|c|c|c}
    \hline
    $\lambda$ & $P(\tau \le \lambda)$ & relative variance & $R_e$ & time (sec.) \\
    \hline
    $2.0$ & $9.23 \times 10^{-2}$ & $3.14$ & $9.23 \times 10^{-2}$ & $95.0$ \\
    \hline
    $1.0$ & $3.22 \times 10^{-2}$ & $5.48$ & $3.22 \times 10^{-2}$ & $48.8$ \\
    \hline
    $0.5$ & $6.28 \times 10^{-3}$ & $12.58$ & $6.28 \times 10^{-3}$ & $24.6$ \\
    \hline
    $0.3$ & $1.00 \times 10^{-3}$ & $31.61$ & $1.00 \times 10^{-3}$ & $14.9$ \\
    \hline
    $0.2$ & $9.30 \times 10^{-5}$ & $103.69$ & $9.30 \times 10^{-5}$ & $9.9$ \\
    \hline
    $0.1$ & $0.00$ & $-$ & $0.00$ & $5.0$ \\
    \hline
  \end{tabular}
  \centering
  \caption{$\mathcal{P}(\tau \le \lambda)$ for one-dimensional dynamics. $R_e$ is the
ratio of trajectories satisfying $\tau \le \lambda$ . $N=10^6$ samples are
used with standard Monte Carlo method. In the last row, the estimation fails because there is no
  trajectories satisfying $\tau \le \lambda$ among $N=10^6$ trajectory samples. \label{fig-ex1-3}}
\end{table}
\begin{figure}[tphb]
  \centering
  \begin{tabular}{lll}
    \subfigure[basis $\phi_0$]{\includegraphics[width=5.0cm]{./fig/pot_1d_1_b_ip.eps}}
    &
    \subfigure[basis $\phi_1$]{\includegraphics[width=5.0cm]{./fig/pot_1d_2_b_ip.eps}}
    &
    \subfigure[$P(\tau \le \lambda)$]{\includegraphics[width=5.0cm]{./fig/pdf_tau.eps}}
\end{tabular}
\caption{The modified potentials for computing $\mathcal{P}(\tau \le \lambda)$ of
one-dimensional dynamics with different sets of basis functions. The
potentials
become more different from the original one when $\lambda$ decreases. (c) shows the cumulative distribution function of
$\tau$ (only for small $\lambda$). \label{tab-ex1-4}} 
\end{figure}

%\subsection{Non-gradient dynamical system}
%Possibly more interesting examples are the noise-driven non-gradient
%systems studied in \cite{mechanism_coherence, stochastic_resonance}.
\subsection{Two dimensional example}
We choose the potential to be 
\begin{align}
V(x) = (1 - x_1^2 - x_2^2)^2 + \frac{x_1^2}{x_1^2 + x_2^2}, \quad x = (x_1,x_2) \in \mathbb{R}^2
\end{align}
and consider the two-dimensional dynamics
\begin{align}
  dz_s = - \nabla V(z_s) ds + \sqrt{2\beta^{-1}} dw_s
\end{align}
where $w_s$ is a two-dimensional Brownian motion.
The potential $V(x)$ (see Figure.\ref{fig-ex2-0}) has two minima at $a = (-1, 0)$ and $b = (1,0)$.
$c_1 = (0, 1)$ and $c_2 = (0,-1)$ are two saddle points and $d=(0,0)$ is a
singular point. The basis functions we choose are 
\begin{align}
  & \phi_0(x) = \exp(-\frac{|x - a|^2}{2r_0^2}), \quad \phi_1(x) = \exp(-\frac{|x - c_1|^2}{2r_1^2}), \\
  & \phi_2(x) = \exp(-\frac{|x - c_2|^2}{2r_2^2}) \quad \phi_3(x) = \exp(-\frac{|x|^2}{2r_0^2}) 
\end{align}
with $r_0 = 0.5, r_1 = r_2 = 1.0$. We study the dynamics starting from $z_0 =
a$ and define the stopping time as $\tau = \inf\{s > 0, |z_s - b| \le 1.0\}$.

For the optimal control problem, the cost function is 
\begin{align}
  \inf_{u_s} \mathrm{E}\Big(\tau + \frac{1}{2}\int_0^\tau |u_s|^2 ds ~\Big|~ z_0 = a \Big)
  \label{cost-ex2}
\end{align}
under dynamics
\begin{align}
  dz_s = - \nabla V(z_s) ds + \sqrt{2}u_s ds + \sqrt{2\beta^{-1}} dw_s
\end{align}
We choose $\beta = 8.0$, $\mbox{dt} = 0.001$. When solving the coefficients
using Algorithm.\ref{ce-algo}, we
choose $(\beta_0, \beta_1 , \beta_2, \beta_3) = (2.0, 4.0, 6.0, 8.0)$, and
$N_i = 10^4$ for each $\beta_i, i = 0, 1, 2, 3$. Four different sets of
basis functions $\{\phi_0\}$, $\{\phi_0, \phi_1\}$, $\{\phi_0, \phi_1,
\phi_2\}$, $\{\phi_0, \phi_1, \phi_2, \phi_3\}$ are studied. After finding the coefficients, the cost function (\ref{cost-ex2}) is computed with $N =
10^6$ samples. Results are tabulated in Table.\ref{tab-ex2-1}. From the dual relation
(\ref{dual-relation}), we also compute the optimal cost by Monte Carlo method
and the value is $1.27$. It can be observed that the cost ($1.39$) is close to the
optimal cost ($1.27$) if all basis functions are used. Figure.\ref{fig-ex2-1} shows the optimized potential with different basis functions.
\begin{table}
  \begin{tabular}{c|c|c|c|c}
    \hline
    job & coefficients & cost & relative variance & mean of $\tau$ \\
    \hline
    $0$ & ($-1.535$, $0.000$, $0.000$, $0.000$) & $3.69$ & $0.67$ & $2.97$ \\
    \hline
    $1$ & ($-1.379$, $0.744$, $0.000$, $0.000$) & $3.86$ & $0.82$ & $3.16$ \\
    \hline
    $2$ & ($-0.943$, $1.527$, $1.500$, $0.000$) & $2.48$ & $0.94$ & $1.71$ \\
    \hline
    $3$ & ($-0.909$, $0.479$, $0.484$, $0.936$) & $1.39$ & $0.46$ & $0.63$ \\
    \hline
  \end{tabular}
  \centering
  \caption{2D example, optimal control problem with different sets of basis
  functions $\{\phi_0\}$, $\{\phi_0, \phi_1\}$, $\{\phi_0, \phi_1, \phi_2\}$, $\{\phi_0, \phi_1, \phi_2, \phi_3\}$.
  The coefficients are solved using Algorithm.\ref{ce-algo}. The cost value is
  computed with $N=10^{6}$ trajectories.
The last row is the mean value of stopping time $\tau$ under the modified
dynamics. \label{tab-ex2-1}}
\end{table}
\begin{figure}[tphb]
  \centering
    \includegraphics[width=9.0cm]{./fig/pot_2d_1_0_0.eps}
    \caption{Potential $V(x)$ for the 2D example. \label{fig-ex2-0}}
\end{figure}
\begin{figure}[tphb]
  \centering
  \begin{tabular}{ll}
    \subfigure[basis $\{\phi_0\}$]{\includegraphics[width=6.0cm]{./fig/pot_2d_1_0_4.eps}} 
    &
    \subfigure[basis $\{\phi_0, \phi_1\}$]{\includegraphics[width=6.0cm]{./fig/pot_2d_1_1_4.eps}} \\
    \subfigure[basis $\{\phi_0, \phi_1, \phi_2\}$]{\includegraphics[width=6.0cm]{./fig/pot_2d_1_2_4.eps}} 
    &
    \subfigure[basis $\{\phi_0, \phi_1, \phi_2, \phi_3\}$]{\includegraphics[width=6.0cm]{./fig/pot_2d_1_3_4.eps}}
  \end{tabular}
    \caption{2D example. The optimized potentials for optimal control problem with different sets of basis functions.\label{fig-ex2-1}}
\end{figure}

We continue to compute the probability of the rare events $\mathcal{P}(\tau
\le \lambda)$. We choose a large $\beta = 20.0$, $\mbox{dt} = 0.001$. Three sets of
basis functions as above are used : $\{\phi_0\}$, $\{\phi_0, \phi_1,
\phi_2\}$, $\{\phi_0, \phi_1, \phi_2, \phi_3\}$. $(\beta_0, \beta_1 , \beta_2,
\beta_3, \beta_4, \beta_5, \beta_6) = (4.0, 8.0, 10.0, 12.0, 16.0, 18.0,
20.0)$ and $N_i = 10^4, 0\le i \le 5, N_6 = 10^5$ are used when solving the coefficients.
The numerical results are shown in Table.\ref{tab-ex2-2}--\ref{tab-ex2-5} and Figure.\ref{fig-ex2-2}, Figure.\ref{fig-ex2-3}.
Table.\ref{tab-ex2-2}--\ref{tab-ex2-4} shows the
results of importance sampling with different sets of basis functions and the results
of standard Monte Carlo method are given in Table.\ref{tab-ex2-5}. $N=5 \times
10^{6}$ samples are used to compute the probability. We clearly
see that, comparing to the standard Monte Carlo method, the (relative) variances are always largely reduced with importance sampling, especially when $\lambda$ is small. Comparing different basis
functions, we observe that the performance is better when all four basis functions are used.
The cumulative distribution function is computed using importance sampling
with all four basis functions and is shown in Figure.\ref{fig-ex2-2}. The optimized potentials for different $\lambda$ are presented in Figure.\ref{fig-ex2-3}. We see once again that large force is
needed when $\lambda$ is small.

\begin{table}
  \begin{tabular}{c|c|c|c|c|c}
    \hline
  $\lambda$ & coefficients & $P(\tau \le \lambda)$ & relative variance & $R_e$ & time (sec.) \\
    \hline
    $12.0$ & ($-0.271$, $0.000$, $0.000$, $0.000$) & $9.36\times 10^{-3}$ & $7.00$ & $2.08\times 10^{-1}$ & $3219.7$ \\ 
    \hline
    $10.0$ & ($-0.305$, $0.000$, $0.000$, $0.000$) & $6.47\times 10^{-3}$ & $8.25$ & $1.87\times 10^{-1}$ & $2688.7$ \\
    \hline
    $8.0$ & ($-0.337$, $0.000$, $0.000$, $0.000$) & $3.97\times 10^{-3}$ & $8.93$ & $1.50\times 10^{-1}$ & $2184.4$ \\
    \hline
    $6.0$ & ($-0.391$, $0.000$, $0.000$, $0.000$) & $2.03\times 10^{-3}$ & $9.13$ & $1.13\times 10^{-1}$ & $1660.6$ \\
    \hline
    $4.0$ & ($-0.471$, $0.000$, $0.000$, $0.000$) & $7.50\times 10^{-4}$ & $12.38$ & $6.86\times 10^{-2}$ & $1121.4$ \\
    \hline
    $2.0$ & ($-0.593$, $0.000$, $0.000$, $0.000$) & $1.19\times 10^{-4}$ & $16.30$ & $2.33\times 10^{-2}$ & $569.6$ \\
    \hline
    $1.0$ & ($-0.945$, $0.000$, $0.000$, $0.000$) & $7.26\times 10^{-6}$ & $31.37$ & $1.86\times 10^{-2}$ & $286.2$ \\
    \hline
    $0.5$ & ($-1.654$, $0.000$, $0.000$, $0.000$) & $3.60\times 10^{-8}$ & $52.00$ & $2.03\times 10^{-2}$ & $143.3$ \\
    \hline
  \end{tabular}
  \centering
  \caption{2D example, $\mathcal{P}(\tau \le \lambda)$ using importance
  sampling method. Coefficients are computed using Algorithm.\ref{ce-algo}
with basis functions $\{\phi_0\}$. $R_e$ is the ratio of trajectories
satisfying $\tau \le \lambda$ under the modified dynamics. \label{tab-ex2-2}}
\end{table}
\begin{table}
  \begin{tabular}{c|c|c|c|c|c}
    \hline
    $\lambda$ & coefficients & $P(\tau \le \lambda)$ & relative variance & $R_e$ & time (sec.) \\
    \hline
  $12.0$ & ($-0.255$,  $0.030$, $0.029$, $0.000$) & $9.40\times 10^{-3}$ & $6.76$ & $1.95\times 10^{-1}$ & $4724.0$ \\
    \hline
  $10.0$ & ($-0.291$,  $0.035$, $0.040$, $0.000$) & $6.40\times 10^{-3}$ & $6.85$ & $1.81\times 10^{-1}$ & $3967.5$ \\
    \hline
  $8.0$ & ($-0.341$,  $0.049$, $0.051$, $0.000$) & $3.98\times 10^{-3}$ & $17.59$ & $1.63\times 10^{-1}$ & $3190.7$ \\
    \hline
  $6.0$ & ($-0.400$,  $0.065$, $0.069$, $0.000$) & $2.03\times 10^{-3}$ & $10.99$ & $1.30\times 10^{-1}$ & $2418.9$ \\
    \hline
  $4.0$ & ($-0.446$,  $0.127$, $0.127$, $0.000$) & $7.47\times 10^{-4}$ & $11.95$ & $8.30\times 10^{-2}$ & $1638.6$ \\
    \hline
  $2.0$ & ($-0.519$,  $0.317$, $0.318$, $0.000$) & $1.21\times 10^{-4}$ & $13.95$ & $6.78\times 10^{-2}$ & $860.5$ \\
    \hline
  $1.0$ & ($-0.719$,  $0.712$, $0.729$, $0.000$) & $7.24\times 10^{-6}$ & $20.08$ & $1.43\times 10^{-1}$ & $457.4$ \\
    \hline
  $0.5$ & ($-1.035$,  $1.681$, $1.552$, $0.000$) & $3.71\times 10^{-8}$ & $12.97$ & $2.69\times 10^{-1}$ & $210.4$ \\
    \hline
  \end{tabular}
  \centering
  \caption{2D example, $\mathcal{P}(\tau \le \lambda)$ using importance
    sampling method. Basis functions are $\{\phi_0, \phi_1, \phi_2\}$. $R_e$ is the ratio of trajectories
satisfying $\tau \le \lambda$ under the modified dynamics. \label{tab-ex2-3}}
\end{table}
\begin{table}
  \begin{tabular}{c|c|c|c|c|c}
    \hline
    $\lambda$ & coefficients & $P(\tau \le \lambda)$ & relative variance & $R_e$ & time (sec.) \\
    \hline
    $12.0$ & ($-0.270$, $-0.029$, $-0.034$, $0.074$) & $9.39 \times 10^{-3}$ & $6.65$ & $2.26\times 10^{-1}$ & $5446.4$ \\
    \hline
    $10.0$ & ($-0.319$, $-0.062$, $-0.064$, $0.108$) & $6.43\times 10^{-3}$ & $7.87$ & $2.35\times 10^{-1}$ & $4475.0$ \\
    \hline
    $8.0$ & ($-0.373$,  $-0.107$, $-0.098$, $0.155$) & $3.95\times 10^{-3}$ & $7.52$ & $2.35\times 10^{-1}$ & $3563.8$ \\
    \hline
    $6.0$ & ($-0.438$,  $-0.158$, $-0.148$, $0.232$) & $2.03\times 10^{-3}$ & $8.17$ & $2.48\times 10^{-1}$ & $2643.7$ \\
    \hline
    $4.0$ & ($-0.505$,  $-0.223$, $-0.209$, $0.343$) & $7.43\times 10^{-4}$ & $8.12$ & $2.76\times 10^{-1}$ & $1745.5$ \\
    \hline
    $2.0$ & ($-0.602$,  $-0.269$, $-0.254$, $0.556$) & $1.19\times 10^{-4}$ & $11.28$ & $3.85\times 10^{-1}$ & $893.5$ \\
    \hline
    $1.0$ & ($-0.787$,  $-0.003$, $-0.003$, $0.630$) & $7.16\times 10^{-6}$ & $12.15$ & $4.18\times 10^{-1}$ & $457.4$ \\
    \hline
    $0.5$ & ($-1.064$, $0.865$, $0.896$, $0.566$) & $3.71\times 10^{-8}$ & $9.28$ & $4.24\times 10^{-1}$ & $235.9$ \\
    \hline
  \end{tabular}
  \centering
  \caption{2D example, $\mathcal{P}(\tau \le \lambda)$ using importance sampling method. Basis functions $\{\phi_0, \phi_1, \phi_2, \phi_3\}$. $R_e$ is the ratio of trajectories satisfying $\tau \le \lambda$ under the modified dynamics. \label{tab-ex2-4}}
\end{table}

\begin{table}
  \begin{tabular}{c|c|c|c|c}
    \hline
    $\lambda$ & $P(\tau \le \lambda)$ & relative variance & $R_e$ & time (sec.) \\
    \hline
    $12.0$  & $9.44\times 10^{-3}$ & $10.25$ & $9.44\times 10^{-3}$ & $3431.5$ \\
    \hline
    $10.0$  & $6.38\times 10^{-3}$ & $12.48$ & $6.38\times 10^{-3}$ & $2861.9$ \\
    \hline
    $8.0$  & $3.94\times 10^{-3}$ & $15.90$ & $3.94\times 10^{-3}$ & $2291.0$ \\
    \hline
  $6.0$  & $2.04\times 10^{-3}$ & $22.10$ & $2.04\times 10^{-3}$ & $1719.1$ \\
    \hline
  $4.0$  & $7.40\times 10^{-4}$ & $36.74$ & $7.40\times 10^{-4}$ & $1146.9$ \\
    \hline
  $2.0$  & $1.25\times 10^{-4}$ & $89.44$ & $1.25\times 10^{-4}$ & $573.6$ \\
    \hline
  $1.0$  & $6.80\times 10^{-6}$ & $383.48$ & $6.80\times 10^{-6}$ & $287.0$ \\
    \hline
  $0.5$  & $0.00$ & $-$ & $0.00$ & $143.7$ \\
    \hline
  \end{tabular}
  \centering
  \caption{2D example, $\mathcal{P}(\tau \le \lambda)$ using standard Monte
  Carlo. $R_e$ is the ratio of trajectories satisfying $\tau \le \lambda$
under the orininal dynamics. \label{tab-ex2-5}}
\end{table}

\begin{figure}[tphb]
  \centering
    \includegraphics[width=8.0cm]{./fig/pdf_tau_2d.eps} \caption{2D example,
    cumulative distribution function of $\tau$, computed using importance
  sampling with basis $\{\phi_0, \phi_1, \phi_2, \phi_3\}$.\label{fig-ex2-2}}
\end{figure}
\begin{figure}[tphb]
  \centering
  \begin{tabular}{ll}
    \subfigure[$\lambda=12.0$]{\includegraphics[width=6.0cm]{./fig/pot_2d_2_2_7.eps}} 
    &
    \subfigure[$\lambda=6.0$]{\includegraphics[width=6.0cm]{./fig/pot_2d_2_11_7.eps}} \\
    \subfigure[$\lambda=2.0$]{\includegraphics[width=6.0cm]{./fig/pot_2d_2_17_7.eps}}
    &
    \subfigure[$\lambda=0.5$]{\includegraphics[width=6.0cm]{./fig/pot_2d_2_23_7.eps}}
  \end{tabular}
  \caption{2D example, optimized potentials for computing $\mathcal{P}(\tau
  \le \lambda)$ with different $\lambda$. Basis functions $\{\phi_0,
\phi_1, \phi_2, \phi_3\}$ are used.\label{fig-ex2-3}}
\end{figure}

\subsection{Conformational transition of butane dissolved in water}

\begin{figure}
  \centering
  \includegraphics[width=6cm]{./fig.md/butane.eps}
  \caption{The butane molecule in water}
  \label{fig:tmp1}
\end{figure}

The butane-dissolved-in-water system is set up in a $3.0\times
3.0\times 3.0\, \textrm{nm}^3$ periodic simulation region, with one
butane molecule described by the GROMOS 45a3 force
field~\cite{schuler2001improved} dissolved in 900
SPC/E~\cite{berendsen1987missing} water molecules.
%The EF exserts force $\vect F = r \vect E(t)$ on
%the partial charge $r$ in the system.
All simulations are performed by a home-modified GROMACS
4.5~\cite{pronk2013gromacs}.  The molecular dynamics of the system is
subject to the Langevin equation:
\begin{align}\label{eqn:langevin-r}
  dr_i &= m_i^{-1}p_idt, \\\label{eqn:langevin-p}
  dp_i &= [ f_i(r)  + u_i (r, w) ] dt - \gamma p_i dt + \sigma_i dw_t,
\end{align}
where $r$ is the Cartisian coordinates of the $i$-th atom in the
system and $p$ is the corresponding momenta. The friction
constant is denoted by $\gamma$. Due to the
fluctuation-dissipation theorem, the realtion $\sigma_i^2 = 2\gamma
m_i k_B T$ holds, where $k_B$ is the Boltzmann constant, $T$ is the
temperature and $m_i$ is the mass of the atom.  $f_i(r)$ and $u_i(r, w)$
are the force that modelled by the GROMOS 45a3 force field and the
control force on the $i$-th atom, respectively. Being more specific,
the force term is given by
\begin{align}\label{eqn:gromos-inter}
  f(r) = -\nabla_{r_i} V_{\textrm{bond}}(r) - \nabla_{r_i} V_{\textrm{non-bond}}(r),
\end{align}
where the bond interaction is
\begin{align}\nonumber
  V_{\textrm{bond}}(r) =\,&
  V_b(r) + V_\theta(r) + V_\phi(r) \\\nonumber
  =\,&
  \sum_{k = 1}^{N_b} \frac14\, k_b(b_k^2(r) - b_{k,0}^2)^2 
  + 
  \sum_{k = 1}^{N_\theta}\frac12\, k_\theta(\cos(\theta_k(r)) - \cos\theta_{k,0})^2\\\label{eqn:gromos-bd}
  &+
  \sum_{k = 1}^{N_\phi} k_\phi [ 1 + \cos (\delta_k) \cos (m_k\,\phi_k(r))]
\end{align}
The $V_b(r)$ is the energy due to the covalent bonds in the syste that
are not constrained as a constant. $b_k(r)$ is the instantaneous
length of the $k$-th bond, which depends on the coordinates of the
bonding atoms (that is why it is $r$ dependent), while constant
$b_{k,0}$ is the equilibrium length. $k_b$ is the constant denoting
the strength of the interaction. The term $V_\theta(r)$ is the energy
due to the bond angle. $\theta_k(r)$, $\theta_{k,0}$ and $k_\theta$
denote the instantaneous angle, equilibrium angle and interaction
constant, respectively. The third term is the energy due the torsional
dihedral angle.  $\phi_k(r)$, $\delta_k$, $m_k$ and $k_\phi$ denote
the instantaneous dihedral angle, phase shift, multiplicity and the
interaction constant, respectively. For a butane molecule, presented
by Fig.~\ref{fig:tmp1}, the covalent bonds exist for atom pairs 1--2,
2--3 and 3--4. The bond angle interactions exist for atom triples
1--2--3 and 2--3--4. The dihedral angle interaction exists for atom
quadruple 1--2--3--4. The strength of $k_b$ ($=7.15\times
10^6~\textrm{kJ/mol/nm}^4)$ for all bonds) and $k_\theta$
($=530~\textrm{kJ/mol}$ for all bond angles) are relatively much large
than the dihedral angle constant $k_\phi$ ($=5.92~\textrm{kJ/mol}$),
therefore, the vibrations of the bond and bond angle are rigid, so can
be viewed as fast motions in the system. For butane dihedral angle,
the multiplicity is $m=3$ and phase shift $\delta = 0$, so the
dihedral angle is more populated around three angles: $-60^\circ$,
$60^\circ$ and $180^\circ$. The first two are usually called
\emph{cis} conformation, and the last one is usually called
\emph{trans} conformation. And the transition between the
conformations are slow motion in the system. The bonds and bond angles
in water molecules are constrained by the SETTLE
algorithm~\cite{miyamoto2004settle}.  The non-bonded interaction,
i.e.~the second term of R.H.S.~of Eq.~\eqref{eqn:gromos-inter}, is
given by
\begin{align}\label{gromos-nb}
  V_{\textrm{non-bond}}(r) =
  \sum_{i,j} \bigg[\frac{C_{12,ij}}{r_{ij}^{12}} - \frac{C_{6,ij}}{r_{ij}^{6}}\bigg]
  +
  \sum_{i,j}\frac{1}{4\pi\varepsilon_0} \frac{q_i q_j}{ r_{ij}},
\end{align}
where $r_{ij}$ is the relative distance between atom $i$ and $j$. The
first term of Eq.~\eqref{gromos-nb} is the van der Waals interaction
and the second term is the Coulomb interaction, where $q_i$ denotes the
partial charge due to the atom $i$. The van der Waals interaction is
treated by the cut-off method with cut-off radius $9$~nm, while the Coulomb
interaction is treated by the smooth particle mesh Ewald
method (SPME)~\cite{essmann1995spm}.

In this butane dissolved in water system, we are interested in the
first hitting time $\tau$ of transition from cis to trans
conformation, and want to precisely calculate the probability of rare
event that $\tau$ being smaller than some given threshold $\lambda$,
i.e.~$P(\tau \leq\lambda)$. We firstly run an equilibrium simulation
(i.e.~with $u = 0$ in Eq.~\eqref{eqn:langevin-p}) at 300~K, with
friction constant $1.0~\textrm{ps}^{-1}$.  The dihedral angle of the
butane molecule is calculated every 10~ps. If $\phi$ is in the range
$[40^\circ, 80^\circ]$ (corresponding to the cis conformation), then
the conformation (including all water coordinates) is recorded, and
used for the later study of cis-to-trans transition. In our
simulation, we only record the first observed 5000 conformations.
Then controlled simulation (Eq.~\eqref{eqn:langevin-r} and
\eqref{eqn:langevin-p}) are performed also at 300~K with friction
constant $10~\textrm{ps}^{-1}$ until $\lambda$, or until the butane
changes to the trans conformation. Here the butane is of trans
conformation, when the dihedral angle $\phi$ is equal or larger than
$150^\circ$. The control  $u_i$ is defined as a function of
the dihedral angle:
\begin{align}
  u_i(r, w) = -\nabla_{r_i}V_{\textrm{ctrl}}(\phi(r))=
  -\nabla_{r_i} \bigg[ \sum_{k=1}^{N_c} w_k\cos(k \phi (r)) \bigg].
\end{align}
Notice that the dihedral angle is a function of atomic coordinates
$\phi = \phi(r)$.  The control is defined as a function of the
dihedral angle $\phi$, because it fully describes the conformational
change of the butane moelcule, moreover, bond lengths and bond angles
are fast variables, so (as a good approximation) they decouple from
the dynamics of the dihedral angle.  The transitional, rotational
movement and the neighboring solvation structure may also play a role
on the conformation of butane, and it is also fully justified to
design control also based on these degrees of freedoms (DOFs), however, We
will show later that they do not contribute significantly in the
investigated butane system.  The control potential
$V_{\textrm{ctrl}}(\phi(r))$ is chosen as a expansion of cosine base
functions of $\phi$, because the system is symmetric between range $\phi \in
[-180^\circ,0^\circ]$ and $\phi \in [0^\circ, 180^\circ]$, so the sine
bases are eliminated. In all our simulations, the number of control
base functions $N_c$ is chosen to be 8. Numerical results show that
the coefficient of the 8th base $w_8$ is already very small comparing
to the dominant contributing bases, that means the number of base is
large enough that can capture the essential structure of the control.
In all our simulations, the Langevin equation is discretized by the
BAOAB scheme~\cite{leimkuhler2013rational} with a time step of
$10^{-4}$~ps for the case $\lambda = 0.1$~ps, and $5\times10^{-4}$~ps
otherwise.

\begin{figure}
  \centering
  \includegraphics[width=8cm]{./fig.md/fig-pot.eps}
  \caption{The original dihedral potential and the dihedral potential
    with control.}
  \label{fig:tmp2}
\end{figure}

We considered $\lambda = 1.0$, 0.5, 0.2 and 0.1~ps and shown the
result Fig.~\ref{fig:tmp2}, Tab.~\ref{tab:tmp1} and~\ref{tab:tmp2}.
From each of the 5000 equilibrium conformation, 4 trajectories with
different randoms seeds are simulated for each $\lambda$, so we
generate 20000 trajectories for each case.  The control $u(r, w)$ is
calculated by simulations of butane in vacumm, in which all other
parameters are the same as the aforementioned descriptions, except
that the waters are removed. We do it this way because the vacumm
simulation is much cheaper than the in-water simulation, and
practically, the control calculated in the corresponding vacumm
systems perform good enough in the in-water systems, because we find
that no further iteration is need to refine the control. Furthermore,
in the vacumm system, the probability $P(\tau\leq\lambda) =
2.19\times10^{-2}$, $9.32\times 10^{-3}$, $1.35\times 10^{-3}$ and
$5.76\times 10^{-5}$ for $\lambda = 1.0$, 0.5, 0.2 and 0.1~ps,
respectively. These values does not significantly differ from the
dissolved system (see the second column of Tab.~\ref{tab:tmp1}).
Noticing the butane is symmetric with respect to transitional and
rotational movement, the above observation therefore indicates that
the transitional, rotational DOFs and the water structure does not
play a dominant role in the conformational change of butane, and the
definition of control only as a function as the dihedral angle, and
the computation of control in the vacumm system are reasonable
choices.

\begin{table}[t]
  \centering
  \caption{
    Butane dissolved in water. The variance reduced estimate of the probability
    $P (\tau \leq \lambda)$ by applying control on the dihedral angle.
    See the text for more details of the definition of the control.
    Column ``Error'' denotes the standard deviation of
    estimating the probablity $P (\tau \leq \lambda)$.
    Column ``Var'' denotes the reduced variation of the esimate.
    ``Accel.''(acceleration) is the ratio between
    the reduced variance and the brute force variance (see Tab.~\ref{tab:tmp2}),
    which is the actual speed-up of the simulation.
    ``Traj. Usage'' dentoes percentage of trajectoies that changes
    to the trans conformation within time interval $[0,\lambda]$.
  }
  \label{tab:tmp1}
  \begin{tabular*}{0.9\textwidth}{@{\extracolsep{\fill}}lcccrr}
    \hline\hline
    $\lambda$ [ps] & $P (\tau \leq \lambda)$ & Error & Var & Accel. & Traj. Usage \\\hline
    0.1 & $4.34\times 10^{-5}$ & $1.16\times 10^{-5}$ & $2.31\times10^{-6}$ &18.8 & 0.7\%\\
    0.2 & $9.03\times 10^{-4}$ & $0.77\times 10^{-4}$ & $1.19\times10^{-4}$ & 7.6 &10.0\%\\
    0.5 & $6.96\times 10^{-3}$ & $0.34\times 10^{-3}$ & $2.21\times10^{-3}$ & 3.1 &11.9\%\\
    1.0 & $1.68\times 10^{-2}$ & $0.06\times 10^{-2}$ & $6.30\times10^{-3}$ & 2.7 &13.4\%\\
    \hline\hline
  \end{tabular*}
  \caption{
    Butane dissolved in water, brute force simulations
    without any variance reduction. The meaning of the columns are the same
    as Tab.~\ref{tab:tmp1}.}\label{tab:tmp2}
  \begin{tabular*}{0.9\textwidth}{@{\extracolsep{\fill}}lcccrr}
    \hline\hline
    $\lambda$ [ps] & $P (\tau \leq \lambda)$ & Error & Var & Accel. & Traj. Usage \\\hline
    0.2 & $7.46\times 10^{-4}$ & $2.36\times 10^{-4}$ & $7.46\times10^{-4}$ & 1.0 & 0.1\%\\
    0.5 & $6.94\times 10^{-3}$ & $0.72\times 10^{-3}$ & $6.89\times10^{-3}$ & 1.0 & 0.7\%\\
    1.0 & $1.61\times 10^{-2}$ & $0.11\times 10^{-2}$ & $1.59\times10^{-2}$ & 1.0 & 1.6\%\\
    \hline\hline
  \end{tabular*}
\end{table}

The Fig.~\ref{fig:tmp1} plots the effective dihedral angle energy,
which is the original dihedral energy $V_\phi(\phi)$ plus the control
$V_{\textrm{ctrl}}(\phi)$. We only show the effective energy in the
range $[40^\circ, 150^\circ]$, because this this is the initial
conformations start in range $[40^\circ, 80^\circ]$, and the
trajectoies are stopped when they reach $\phi = 150^\circ$. For a easy
comparison, all effective energies are switched by a constant, so that
they are of vanishing value at $\phi = 150^\circ$. It clear that for
smaller $\lambda$ value, the control applied is stronger. Different
from the one-dimensional case, where the barrier is lowered, in this
case, only the potential well around $\phi = 60^\circ$ is fill, and
the barrier around $\phi = 120^\circ$ is almost kept the same.  The
resultant probability $P (\tau \leq \lambda)$ calculted by the
variance reduced simulation is summarized in Tab.~\ref{tab:tmp1},
which is consistent with the Tab.~\ref{tab:tmp2} that presents the
brute force results. Since the variance is redueced, the error of the
statistical esitmate is more precise in the variance reduced case. The
reduced variance is shown in the fourth column of Tab.~\ref{tab:tmp1},
which is much smaller than the brute force case (the fourth column of
Tab.~\ref{tab:tmp2}). The fifth column presents the acceleration ratio
with respect to the brute force simulation or the benefit from the
variance reduction, which is exactly the ratio between the brute force
variance and the reduced variance. The column of trajectory usage
presents the percentage of the trajectories that changes to the trans
conformation within time interval $[0, \lambda]$. It is increased by
the control with respect to the brute force result.

\section{Discussions}
\label{sec-discuss}

As a continuation of the work \cite{zlph2013, zhws13}, we study the cross-entropy
method in the setting of diffusion process and their application in the
importance sampling and optimal control.
In those previous work, some assumptions have been made on the dynamical systems and
on the type of quantities to estimate \cite{zhws13}. For example, 
in order to perform the asymptotic analysis, we assumed there are explicit parameters in dynamical systems and the dynamics
is in certain limit regime. In the case of multiple time-scale systems, we
also assumed that explicit slow/fast variables are known. In sampling problem \cite{zhws13}, we assumed 
the optimal change of measure is related to certain Hamilton-Jacobi-Bellman equation,
and therefore the type of quantity to estimate is restricted. With the
cross-entropy method, all these assumptions could be removed. General
diffusion processes can be investigated and importance sampling for more
meaningful quantity such as $\mathcal{P}(\tau \le \lambda)$ can be considered.

On the other hand, some basis functions are needed in order to apply the
cross-entropy method. The choice of these basis functions are crucial in order
to obtain good approximation of the target measure and therefore effcient
importance sampling or ``good control force'' in the optimal control problem.
The successful choice is often based on our knowledge about the given dyanmical
system, such as its metastable states, or reactive cordinates. It is easy to
imagine that there is no way to obtain satisfactory results when the system
has a reactive coordinate but the basis functions are supported somewhere else.
The optimal control problem for dynamical systems having a reactive coordinate
will be further studied in the future work.

\section*{Acknowledgements}
The authors acknowledge financial support by the DFG Research Center MATHEON.
\FloatBarrier

\bibliographystyle{siam}
\bibliography{reference}
\end{document}

### 2013-12-24 writedown/first.hitting/cross_entropy/cross_entropy.tex
revised by Wei

@@
 \begin{abstract}
-  The cross-entropy method in diffusion process setting is studied. The
+  The cross-entropy method in the diffusion process setting is studied. The
   general algorithm is proposed and is applied to design efficient importance sampling for rare events as well as to solve optimal
-  control problem. Several examples are studied numerically.
+  control problem of diffusion process. Several examples are studied
+  numerically to demonstrate this method.
 \end{abstract}
@@
 applications in mathematics, physics, economics, and so on.
-We are mainly interested in two topics in diffusion process, i.e. importance sampling and optimal control. 
+Here we are mainly interested in two topics in diffusion process, i.e. importance sampling and optimal control. 
 For importance sampling, it is one kind of variance reduction technique 
@@
 For optimal control, the purpose is to driven the dynamical system in certain
-optimal way and has strong application backgrounds in finance, climate
+optimal way and has strong applicational backgrounds in finance, climate
 science, mechanics \cite{fleming2006}. We emphasize that these two different areas are actually related and can
@@
 algorithms to find the way to do the importance sampling, or to find the
-optimal control strategy. One main reason that hinder efficient algorithms is
-due to the high dimensionality of the dynamical systems, as in the case of
-climate system, biological system and so on. However, when the dynamical
+optimal control strategy. One main reason that designing efficient algorithms
+is urgent and challenging is due to the high dimensionality of the dynamical
+systems, as in the case of climate system, biological system and so on. However, when the dynamical
 systems have multiscale feature, efficient algorithms may be designed by
@@
 
-In this work, we consider these two topics together as designing the change of
+In this work, we consider these two topics together as how to design the change of
 measure in an optimal way. If we further assume the existence of the optimal
-change of measure $\mu^*$, the problem becomes to find approximation of
-$\mu^*$ under some measure distance.
-
-On the other hand, the Kullback-Leibler divergence, although not a metric, is
+change of measure $\mu^*$, the problem can be roughly described as how to find the approximation of
+$\mu^*$ under some measure distance. For the choice of measure distances, we will use the Kullback-Leibler
+divergence, which, although not a metric, is
 widely used to quantified the similarity of two probability measures. 
@@
 probability measures. It has wide applications in different areas, such as
-Monte Carlo simulation, optimization, and machine learning \cite{ce_book}. The
+Monte Carlo simulation, optimization, and machine learning \cite{ce_tutorial, ce_book}. The
 main purpose of this paper is to formulate the cross-entropy method in the
 context of diffusion process, i.e. for measures defined in path space, and
-discuss how to apply this method in importance sampling and optimal control
+discuss the applications of this method in importance sampling and optimal control
 problems.
@@
 Section~\ref{sec-ce-algo}. Several numerical examples are studied in
-Section~\ref{sec-examples}. Some discussions are presented in
+Section~\ref{sec-examples}. Discussions are presented in
 Section~\ref{sec-discuss}.
@@
 and the integral is interpreted as Ito integral.
-We could also rewrite it as Stratonovich integral with the relationship
+We could also rewrite it as Stratonovich integral using the relationship
 \begin{align}
@@
     \begin{align*}
-      I_1(z) = \frac{1}{2} \int_0^\tau \Big|\frac{dz}{ds} - b(z_s)\Big|_a^2 ds
+      I_1(z) = \frac{1}{2} \int_0^\tau \Big|\frac{dz}{ds} -
+      b(z_s)\Big|_{a(z)}^2 ds
     \end{align*}
@@
       \mathbb{R}^{n-k}$. We can still assign a probability measure $\mu$ on the
-      trajectory space, i.e. $\forall z = (x,y) $ is a trajectory, we have 
-      $d\mu = \exp(-\frac{\beta}{2} I_1(z)) d\nu$ (\textbf{or formally :}  $\mu(dz) =
-      \exp(-\frac{\beta}{2} I_1(z)) \nu(dy)$, which is better ?), where $\nu$ is Wiener
+      \textbf{trajectory space}, i.e. $\forall z = (x,y) $ is a trajectory, we have 
+      $d\mu = \exp(-\frac{\beta}{2} I_1(z)) d\nu$, where $\nu$ is Wiener
       measure on $\mathbb{R}^{n-k}$ and 
@@
 \end{remark}
-\begin{comment}
-\begin{remark}
-  Since $I_2$ is simply the Stratonovich version of $I_1$, we use $I_1$ for
-  simplicity.
-  Starting from dynamics
-\begin{align}
-\begin{split}
-  d z_s &= \widetilde{b}(z_s) ds + \sqrt{2\beta^{-1}} d\widetilde{w}_s, \quad s >0  \\
-  z_0&=z
-  \end{split}
-\label{dynamics-11}
-\end{align}
-the above argument can be repeated to get 
-\begin{align}
-  d\widetilde{\mu} &=\exp\Big(- \frac{\beta}{4}\int_0^\tau \Big|\frac{dz}{ds} -
-  \widetilde{b}(z_s)\Big|^2 ds\Big) d\nu
-  \label{measure-2}
-\end{align}
-Divided it by (\ref{measure-1}), we have
-\begin{align}
-  \frac{d\widetilde{\mu}}{d\mu} = \exp\Big(- \frac{\beta}{4}\Big[2\int_0^\tau (dz_s,
-  b(z_s) - \widetilde{b}(z_s)) + \int_0^\tau (|\widetilde{b}(z_s)|^2 - |b(z_s)|^2) ds\Big]
-  \Big)
-\end{align}
-Then we substitute the dynamics (\ref{dynamics-1}) in above (in principle we
-can also use dynamics (\ref{dynamics-11}), but $\widetilde{w}$ is not Brownian motion
-under $\mu$ ...), after
-simplification, we have 
-\begin{align}
-  \frac{d\widetilde{\mu}}{d\mu} = \exp\Big(-
-  \frac{\beta^{1/2}}{\sqrt{2}}\int_0^\tau (dw_s,
-b(z_s) - \widetilde{b}(z_s)) - \frac{\beta}{4}\int_0^\tau (|\widetilde{b}(z_s)
-- b(z_s)|^2) ds\Big)
-\end{align}
-Let $u_s = \frac{1}{\sqrt{2}} (b(z_s) - \widetilde{b}(z_s))$, it gives
-\begin{align}
-  \frac{d\widetilde{\mu}}{d\mu} = \exp\Big(- \beta^{1/2}\int_0^\tau (dw_s, u_s)
-) - \frac{\beta}{2}\int_0^\tau |u_s|^2 ds\Big)
-\end{align}
-So we recover the more general situation in our IS paper for $\sigma =
-\sqrt{2}$.
-\end{remark}
-\end{comment}
-\section{Cross entropy method for importance sampling}
+\section{Cross-entropy method for importance sampling}
 \label{sec-ce-is}
-In this section, we discuss how to apply the cross entropy
+In this section, we discuss how to apply the cross-entropy
 method \cite{ce_tutorial, ce_book} to guide the
@@
 Let $\mathcal{D}$ be certain subset in the path space and
-$\mathds{1}_{\mathcal{D}}$ be the
+$\textbf{1}_{\mathcal{D}}$ be the
 characteristic function on $\mathcal{D}$. 
@@
 \begin{align}
-  F(z) = \mathds{1}_{\mathcal{D}}(z) \quad \mbox{or} \quad F(z) = \exp(-\beta\int_0^\tau G(z_s) \,ds).
+  F(z) = \textbf{1}_{\mathcal{D}}(z) \quad \mbox{or} \quad F(z) = \exp(-\beta\int_0^\tau G(z_s) \,ds).
   \label{path-fun-f}
 \end{align}
-\begin{remark}
   A special and interesting case is when $\mathcal{D} = \{z, \tau(z) \le \lambda\}$,
-  $\lambda > 0$ is fixed. Then $l = \mathrm{E}(\mathds{1}_{\mathcal{D}}(z)) =
+  $\lambda > 0$ is fixed. Then $l = \mathrm{E}(\textbf{1}_{\mathcal{D}}(z)) =
   \mathcal{P}(\tau \le \lambda)$, i.e. the probability of the trajectories
-  stopping before time $\lambda$. This quantity may provide us some details of the
+  stop before time $\lambda$. This quantity may provide us some details of the
   distribution of the stopping time $\tau$ and help us understand the
   behaviors of the dynamics.
-\end{remark}
 
@@
 and the importance sampling will be realized by simulating trajectories from
-this modified dynamics. Specifically, given $\bm{w} = (w_1, w_2,
-\cdots, w_m)^T \in \mathbb{R}^m$,
+this modified dynamics. Specifically, given $\bm{\omega} = (\omega_1, \omega_2,
+\cdots, \omega_m)^T \in \mathbb{R}^m$,
 the corresponding control force is state-dependent and is supposed to be 
 \begin{align}
-  u(x; \bm{w}) = \sum\limits_{i = 1}^{m} w_i \nabla \phi_i(x), \quad x \in
+  u(x; \bm{\omega}) = \sum\limits_{i = 1}^{m} \omega_i \nabla \phi_i(x), \quad x \in
   \mathbb{R}^n\,.
@@
 \end{align}
-Therefore, for fixed parameter $\bm{w}$, the controlled dynamics is 
+Therefore, for fixed parameter $\bm{\omega}$, the controlled dynamics is 
 \begin{align}
 \begin{split}
-  d z_s &= b(z_s) ds + \sqrt{2}u(z_s;\bm{w}) ds + \sqrt{2\beta^{-1}} dw_s, \quad t \le s \le \tau \\
+  d z_s &= b(z_s) ds + \sqrt{2}u(z_s;\bm{\omega}) ds + \sqrt{2\beta^{-1}} dw_s, \quad t \le s \le \tau \\
      z_0 &=z
@@
 \end{align}
-Let $\mu_{\bm{w}}$ be the probability measure of dynamics (\ref{dynamics-2})
-w.r.t. $\bm{w}$. 
-From Section \ref{sec-om-functional},  we know $\mu_{\bm{w}}$ has density 
-$f(z;\bm{w}) = \exp(-\frac{\beta}{2} I(z ; \bm{w}))$ w.r.t. Wiener
+Let $\mu_{\bm{\omega}}$ be the probability measure of dynamics (\ref{dynamics-2}). 
+From Section \ref{sec-om-functional},  we know $\mu_{\bm{\omega}}$ has density 
+$f(z;\bm{\omega}) = \exp(-\frac{\beta}{2} I(z ; \bm{\omega}))$ w.r.t. Wiener
 measure $\nu$, with
 \begin{align}
-  I(z ; \bm{w}) = \frac{1}{2}\int_0^\tau \Big|\frac{dz}{ds}- b(z_s) -
-  \sqrt{2}\sum\limits_{i = 1}^{m} w_i \nabla \phi_i(z_s)\Big|^2 ds + \beta^{-1}
+  I(z ; \bm{\omega}) = \frac{1}{2}\int_0^\tau \Big|\frac{dz}{ds}- b(z_s) -
+  \sqrt{2}\sum\limits_{i = 1}^{m} \omega_i \nabla \phi_i(z_s)\Big|^2 ds + \beta^{-1}
   \int_0^\tau \operatorname{div}\, \Big(b(z_s) + \sqrt{2}\sum\limits_{i =
-  1}^{m} w_i \nabla \phi_i(z_s)\Big) ds
+  1}^{m} \omega_i \nabla \phi_i(z_s)\Big) ds
   \label{om-fun-w}
 \end{align}
-and the original dynamics (\ref{dynamics-1}) corresponds to $\bm{w} = \bm{0}$ (we use $\mu = \mu_{\bm{0}}$).
+and the original dynamics (\ref{dynamics-1}) corresponds to $\bm{\omega} = \bm{0}$ (we use $\mu = \mu_{\bm{0}}$).
 Given a probability measure $\mu_g$ on path space with $d\mu_g = g(z)d\nu$, 
-we can design a unbiased Monte Carlo estimator with importance sampling method as 
+we can design a unbiased Monte Carlo estimator for (\ref{mean-l}) as 
 \begin{align}
@@
 
-The idea of cross entropy method is to find the best approximation of $g^*(z)$ among
-$f(z;\bm{w}), \bm{w} \in \mathbb{R}^m$. More specifically, define the
+The idea of cross-entropy method is to find the best approximation of $g^*(z)$ among
+$f(z;\bm{\omega}), \bm{\omega} \in \mathbb{R}^m$. More specifically, define the
 Kullback-Leibler divergence between
@@
 \begin{align}
-  \min_{\bm{w} \in \mathbb{R}^m} D(\mu_{g^*}, \mu_{\bm{w}}).
+  \min_{\bm{\omega} \in \mathbb{R}^m} D(\mu_{g^*}, \mu_{\bm{\omega}}).
   \label{kl-dist-d}
@@
 \begin{align}
-  \max_{\bm{w} \in \mathbb{R}^m} \mathrm{E}_{\mu} ( F(z) \log f(z;\bm{w}))
-  = \max_{\bm{w} \in \mathbb{R}^m} \mathrm{E}_{\mu_{\bm{v}}} ( F(z)
-  w(z;\bm{v}) \log f(z;\bm{w})), \hspace{3mm} \forall \bm{v} \in
+  \max_{\bm{\omega} \in \mathbb{R}^m} \mathrm{E}_{\mu} ( F(z) \log
+  f(z;\bm{\omega}))
+  = \max_{\bm{\omega} \in \mathbb{R}^m} \mathrm{E}_{\mu_{\bm{v}}} ( F(z)
+  \eta(z;\bm{v}) \log f(z;\bm{\omega})), \hspace{3mm} \forall \bm{v} \in
   \mathbb{R}^m
@@
 \begin{align}
-  w(z;\bm{v}) = \frac{f(z;\bm{0})}{f(z;\bm{v})}.
+  \eta(z;\bm{v}) = \frac{f(z;\bm{0})}{f(z;\bm{v})}.
 \end{align}
@@
 \begin{align}
-\max_{\bm{w} \in \mathbb{R}^m} \frac{1}{N} \sum_{i=1}^{N} F(z_i) w(z_i, \bm{v}) \log f(z_i;\bm{w}) 
+\max_{\bm{\omega} \in \mathbb{R}^m} \frac{1}{N} \sum_{i=1}^{N} F(z_i)
+\eta(z_i; \bm{v}) \log f(z_i;\bm{\omega}) 
 \label{ce-2-discrete}
@@
 \begin{align}
-  \frac{1}{N} \sum_{i=1}^{N} F(z_i) w(z_i, \bm{v}) \nabla_{\bm{w}} \log
-  f(z_i;\bm{w}) = 0.
+  \frac{1}{N} \sum_{i=1}^{N} F(z_i) \eta(z_i; \bm{v}) \nabla_{\bm{\omega}} \log
+  f(z_i;\bm{\omega}) = 0.
   \label{max-gradient}
@@
 With formula (\ref{om-fun-w}), equations (\ref{max-gradient}) could be written more explicitly
-as linear equations $A\bm{w} = \bm{r}$, where $\forall 1\le i,j\le m$, 
+as linear equations $A\bm{\omega} = \bm{r}$, where $\forall 1\le i,j\le m$, 
 \begin{align}
   \begin{split}
-  A_{ij} &= \frac{1}{N}\sum_{k = 1}^{N} F(z_k)w(z_k, \bm{v}) \int_0^\tau
-    \nabla\phi_i(z_k)\nabla\phi_j(z_k)\,ds, \\
-     r_i &= \frac{1}{\sqrt{2}N}\sum_{k = 1}^{N} F(z_k)w(z_k, \bm{v})
-    \Big[\int_0^\tau \Big(\frac{dz_k}{ds} - b(z_s)\Big)\nabla\phi_i(z_k)\,ds - \beta^{-1} \int_0^\tau \operatorname{div} \nabla \phi_i(z_k)\,ds\Big]
+  A_{ij} &= \frac{1}{N}\sum_{k = 1}^{N} F(z_k)\eta(z_k; \bm{v}) \int_0^\tau
+  \nabla\phi_i(z_{k,s})\nabla\phi_j(z_{k,s})\,ds, \\
+     r_i &= \frac{1}{\sqrt{2}N}\sum_{k = 1}^{N} F(z_k)\eta(z_k; \bm{v})
+  \Big[\int_0^\tau \Big(\frac{dz_k}{ds} - b(z_s)\Big)\nabla\phi_i(z_{k,s})\,ds
+  - \beta^{-1} \int_0^\tau \operatorname{div} \nabla \phi_i(z_{k,s})\,ds\Big]
   \end{split}
@@
   \bm{v} \in \mathbb{R}^n$, where $\mu^\pi_{\bm{v}}$ and $\nu^{\pi}$ are
-  defined similarly. The aim is to find the optimal $\bm{w} \in
-  \mathbb{R}^m$ and the corresponding measure $\mu^\pi_{\bm{w}}$ to do
+  defined similarly. The aim is to find the optimal $\bm{\omega} \in
+  \mathbb{R}^m$ and the corresponding measure $\mu^\pi_{\bm{\omega}}$ to do
   the importance sampling. All the derivations are the same with the above
@@
 	dx_s & = b_1(x_s, y_s)\, ds \\
-	dy_s & = b_2(x_s, y_s)\, ds + \sqrt{2} u(x_s, y_s ; \bm{w})\, ds + \sqrt{2\beta^{-1}} dw_s
+	dy_s & = b_2(x_s, y_s)\, ds + \sqrt{2} u(x_s, y_s ; \bm{\omega})\, ds + \sqrt{2\beta^{-1}} dw_s
       \end{split}
@@
       \end{align}
-      with $u(x_s, y_s ; \bm{w}) = \sum\limits_{i=1}^{m} w_i \nabla \phi_i(x_s,
-      y_s)$ and optimize the coefficients $\bm{w}$. All the derivations are
+      with $u(x_s, y_s ; \bm{\omega}) = \sum\limits_{i=1}^{m} \omega_i \nabla \phi_i(x_s,
+      y_s)$ and optimize the coefficients $\bm{\omega}$. All the derivations are
       the same as above.
 \end{remark}
-\section{Cross entropy method for optimal control problem}
+\section{Cross-entropy method for optimal control problem}
 \label{sec-ocp}
-In this section, we consider applying the cross entropy method to solve optimal control problem
+In this section, we consider applying the cross-entropy method to solve optimal control problem
 for diffusion process.
 
-Consider the optimal control problem with cost function
+Consider the optimal control problem with cost
+function \cite{control_schuette,zlph2013}
 \begin{equation}
   J(u) = \mathrm{E}_{\mu_u}\left[\int_0^\tau \left(G(z_s) + \frac{1}{2} |u(z_s,s)|^2
-\right) ds ~\bigg|~ z_0 = x\right].
+\right) ds ~\bigg|~ z_0 = x\right],
 \end{equation}
@@
 
-Let $\hat{u} = \argmin J(u)$. For simplicity, we denote $\mu, \hat{\mu}$ as
-the probability measure corresponding to $u \equiv 0$ and $\hat{u}$ respectively. Using the dual relation
+Suppose the lower bound of $J(u)$ can be achieved and let $\hat{u} = \argmin J(u)$. For simplicity, we denote $\mu, \hat{\mu}$ as
+the probability measure corresponding to $u \equiv 0$ and $\hat{u}$
+respectively. Using the dual relation \cite{DaiPra1996, zlph2013}
     \begin{align}
@@
 \begin{align}
-  \min_{\bm{w} \in \mathbb{R}^m} J(u(z;\bm{w})).
+  \min_{\bm{\omega} \in \mathbb{R}^m} J(u(z;\bm{\omega})).
   \label{J-w}
@@
 $\mu_u$ and $\hat{\mu}$, which, however, is still not easy. On the other hand, inspired by 
-the discussions in Section \ref{sec-ce-is}, cross entropy method can be
-applied to minimize $D(\hat{\mu} \,|\, \mu_u)$ (notice this is different with
-(\ref{J-w}), because the Kullback-Leibler divergence is not symmetric). Following
+the discussions in Section \ref{sec-ce-is}, we can apply cross-entropy method 
+to minimize $D(\hat{\mu} \,|\, \mu_u)$ (notice this is different from
+(\ref{J-w}), because the Kullback-Leibler divergence is not symmetric), and
+hope that (\ref{J-w}) can be solved approximately. Following
 the notations in Section \ref{sec-ce-is} and from (\ref{jensen-as}), we know the optimal measure
@@
 therefore the problem is reduced to the case in Section \ref{sec-ce-is} with $F(z) = \exp(-\beta\int_0^\tau G(z_s) \,ds)$.
-\section{Cross entropy algorithm}
+\section{Cross-entropy algorithm}
 \label{sec-ce-algo}
-In this section, we consider how to implement the cross entropy algorithm in
+In this section, we consider how to implement the cross-entropy algorithm in
 order to maximize (\ref{ce-2-discrete}), which is equivalent to (\ref{kl-dist-d}) in the discrete case. 
@@
 (\ref{linear-eqn}) is both time-consuming and inaccurate. Inspired by the
-original cross entropy method \cite{ce_tutorial},
+original cross-entropy method \cite{ce_tutorial},
 we may overcome this problem by starting from a higher temperature (smaller $\beta$) and compute
@@
     \State generate $N_j$ trajectories $z_i, i=1,2,\cdots, N$ from dynamics (\ref{dynamics-2}), with
-    $\bm{w} = \bm{v}^{(j)}$, $\beta = \beta_j$.
+    $\bm{\omega} = \bm{v}^{(j)}$, $\beta = \beta_j$.
     \State compute the coefficients of $A^{(j)}, \bm{r}^{(j)}$ from (\ref{linear-eqn}) with $\bm{v} =
@@
 \label{sec-examples}
-In this section, we will study the cross entropy method with some concrete dynamics and
+In this section, we will study the cross-entropy method with some concrete dynamics and
 illustrate some numerical results.
@@
 $x_0 = -1, x_2 = 1$, while $x_1 = 0$ is
-saddle point. Define $\tau = \inf\{s >0, |z_s - x_2| \le 1 \}$, and choose
+a saddle point. Define $\tau = \inf\{s >0, |z_s - x_2| \le 1 \}$, and choose
 the starting point $z_0 = x_0 =
@@
 \end{align}
-$\mbox{dt} = 1.0 \times 10^{-3}$, $N_j = 10^{4}, j=0,1,2$ is used to compute the
-coefficients $\bm{w}$ using Algorithm.\ref{ce-algo}, with $\{\beta_0,\beta_1, \beta_2\}
+$\mbox{dt} = 1.0 \times 10^{-3}$, $N_j = 10^{4}, j=0,1,2$ are used to compute the
+coefficients $\bm{\omega}$ using Algorithm.\ref{ce-algo}, with $\{\beta_0,\beta_1, \beta_2\}
 = \{1.0, 3.0, 5.0\}$.  Therefore $k=2$ in Algorithm.\ref{ce-algo} and $3$
@@
 from a Feymann-Kac type boundary value PDE \cite{control_schuette}. The
-potential (the cost value) obtained in this way can be considered as exact solution and is also
-shown in Figure.\ref{fig-ex1-2} (The last row of Table.\ref{tab-ex1-1}).
+potential (the cost value) obtained in this way can be considered as the exact solution and is also
+shown in Figure.\ref{fig-ex1-2} (The last row of Tab.~\ref{tab-ex1-1}).
 Later on, $N = 10^{6}$ samples are generated to compute
 the cost function (\ref{ex1-cost}) and the results are presented in
-Table.\ref{tab-ex1-1}. We can observe that better results are obtained when basis function
+Tab.~\ref{tab-ex1-1}. We can observe that better results are obtained when basis function
 $\phi_1$ is used, since in job 2 and 3, the cost value (1.31) is closer to the exact
@@
 \end{tabular}
-\caption{One-dimensional dynamics. Convergence results for the modified
+\caption{Optimal control problem for one-dimensional dynamics. Convergence results for the modified
 potentials $V^{(j)}, j=0,1,2,3$, with different sets of basis functions\label{fig-ex1-1}}
@@
     \includegraphics[width=7.0cm]{./fig/pot_1d_basis_control.eps}
-    \caption{One-dimensional dynamics. Optimized potential $V^{(3)}$ with
+    \caption{Optimal control problem for one-dimensional dynamics. Optimized potential $V^{(3)}$ with
     different sets of basis functions. The curve with Label ``exact'' is
@@
   \centering
-  \begin{tabular}{c|c|c|c|c}
+%  \begin{tabular}{c|c|c|c|c}
+  \begin{tabular*}{0.8\textwidth}{@{\extracolsep{\fill}}ccccc}
+    \hline
     \hline
@@
        $1$ & $(-1.771, 0, 0)$ & $5.14$ & $1.35$ & $2.02$  \\
-    \hline
        $2$ & $(0, 1.857, 0)$ & $1.31$ & $0.44$ & $0.57$  \\
-    \hline
        $3$ & $(-0.110, 1.762, 0)$ & $1.31$ & $0.44$ & $0.57$  \\
-    \hline
        $4$ & $(-1.611, 0, 1.379)$ & $3.88$ & $1.22$ & $1.73$  \\
-    \hline
        $5$ & $$ & $1.25$ & $0.30$ & $0.52$  \\
     \hline
-  \end{tabular}
+    \hline
+  \end{tabular*}
   \caption{Results of the optimal control problem for the one-dimensional dynamics with different sets
@@
 
-We continue to study how to compute $P(\tau \le \lambda)$ for $\lambda > 0$.
+We continue to study the computation of $P(\tau \le \lambda)$ for $\lambda > 0$.
 We choose $\lambda = 2.0, 1.0, 0.5, 0.3, 0.1$ and two sets of basis functions are used : $\{\phi_1\}$
@@
 standard Monte Carlo method. These results are shown in
-Table.\ref{tab-ex1-2}--\ref{tab-ex1-4} and Figure.\ref{fig-ex1-3}.
-Table.\ref{tab-ex1-2} and Table.\ref{tab-ex1-3} shows the results of
+Tab.~\ref{tab-ex1-2}--\ref{tab-ex1-4} and Figure.\ref{fig-ex1-3}.
+Tab.~\ref{tab-ex1-2} and Tab.~\ref{tab-ex1-3} shows the results of
 importance sampling with basis function $\{\phi_0\}$ and $\{\phi_1\}$,
-respectively, while Table.\ref{tab-ex1-4} is the result using standard Monte Carlo method.
+respectively, while Tab.~\ref{tab-ex1-4} is the result using standard Monte Carlo method.
 In each case, we can observe that the estimators' relative variance will increase when $\lambda$ becomes
@@
 more apparent when $\lambda$ decreases, while the standard Monte Carlo method
-may fail to work (see the last row of Table.\ref{tab-ex1-4}).
+may fail to work (see the last row of Tab.~\ref{tab-ex1-4}).
 In Figure.\ref{fig-ex1-3}, the optimized potentials used to perform importance sampling
@@
 \centering
-  \begin{tabular}{c|c|c|c|c|c}
-    \hline
-   $\lambda$ & coefficients & $P(\tau \le \lambda)$ & relative variance & $R_e$ & time (sec.) \\
+  \begin{tabular*}{0.9\textwidth}{@{\extracolsep{\fill}}cccccc}
+    \hline
+    \hline
+   $\lambda$ & coefficients & $P(\tau \le \lambda)$ & relative variance & Traj. Usage & time (sec.) \\
     \hline
   $2.0$ & $(-0.837, 0, 0)$ & $9.22 \times 10^{-2}$ & $1.45$ & $6.1 \times 10^{-1}$ & $70.4$ \\
-    \hline
     $1.0$ & $(-1.391, 0, 0)$ & $3.23 \times 10^{-2}$ & $1.48$ & $4.8 \times 10^{-1}$ & $39.4$ \\
-    \hline
     $0.5$ & $(-2.221, 0, 0)$ & $6.45 \times 10^{-3}$ & $1.75$ & $3.9 \times 10^{-1}$ & $21.5$ \\
-    \hline
     $0.3$ & $(-3.282, 0, 0)$ & $9.51 \times 10^{-4}$ & $2.30$ & $3.3 \times 10^{-1}$ & $13.6$ \\
-    \hline
     $0.2$ & $(-4.553, 0, 0)$ & $9.53 \times 10^{-5}$ & $3.10$ & $2.7 \times 10^{-1}$ & $9.4$ \\
-    \hline
     $0.1$ & $(-8.245, 0, 0)$ & $1.22 \times 10^{-7}$ & $6.84$ & $1.6 \times
@@
     \hline
-  \end{tabular}
+    \hline
+  \end{tabular*}
   \caption{ $\mathcal{P}(\tau \le \lambda)$ for one-dimensional dynamics.
-  $N=10^6$ samples and importance sampling are used with basis function $\{\phi_0\}$. $R_e$ is the
+  $N=10^6$ samples and importance sampling are used with basis function
+  $\{\phi_0\}$. ``Traj. Usage'' is the
 ratio of trajectories satisfying $\tau \le \lambda$ under the modified dynamics. \label{tab-ex1-2}}
@@
   \centering
-  \begin{tabular}{c|c|c|c|c|c}
-    \hline
-    $\lambda$ & coefficients & $P(\tau \le b)$ & relative variance & $R_e$ & time (sec.) \\
+  \begin{tabular*}{0.9\textwidth}{@{\extracolsep{\fill}}cccccc}
+    \hline
+    \hline
+    $\lambda$ & coefficients & $P(\tau \le b)$ & relative variance & Traj. Usage & time (sec.) \\
     \hline
   $2.0$ & $(0, 0.962, 0)$ & $9.23 \times 10^{-2}$ & $1.04$ & $8.5 \times 10^{-1}$ & $56.1$\\
-    \hline
   $1.0$ & $(0, 1.498, 0)$ & $3.23 \times 10^{-2}$ & $1.00$ & $8.1 \times 10^{-1}$ & $32.7$ \\
-    \hline
   $0.5$ & $(0, 2.314, 0)$ & $6.46 \times 10^{-3}$ & $1.33$ & $6.8 \times 10^{-1}$ & $19.4$ \\
-    \hline
   $0.3$ & $(0, 3.338, 0)$ & $9.49 \times 10^{-4}$ & $1.84$ & $5.6 \times 10^{-1}$ & $12.8$ \\
-    \hline
   $0.2$ & $(0, 4.578, 0)$ & $9.56 \times 10^{-5}$ & $2.54$ & $4.6 \times 10^{-1}$ & $9.02$\\
-    \hline
   $0.1$ & $(0, 8.232, 0)$ & $1.21 \times 10^{-7}$ & $5.28$ & $2.6 \times 10^{-1}$ & $4.8$\\
     \hline
-  \end{tabular}
+    \hline
+  \end{tabular*}
   \caption{ $\mathcal{P}(\tau \le \lambda)$ for one-dimensional dynamics.
-  $N=10^6$ samples and importance sampling are used with basis function $\{\phi_1\}$. $R_e$ is the ratio of trajectories satisfying $\tau \le \lambda$ under the modified dynamics. \label{tab-ex1-3}}
+  $N=10^6$ samples and importance sampling are used with basis function
+$\{\phi_1\}$. ``Traj. Usage'' is the ratio of trajectories satisfying $\tau \le \lambda$ under the modified dynamics. \label{tab-ex1-3}}
 \end{table}
@@
   \centering
-  \begin{tabular}{c|c|c|c|c}
-    \hline
-    $\lambda$ & $P(\tau \le \lambda)$ & relative variance & $R_e$ & time (sec.) \\
+  \begin{tabular*}{0.9\textwidth}{@{\extracolsep{\fill}}ccccc}
+    \hline
+    \hline
+    $\lambda$ & $P(\tau \le \lambda)$ & relative variance & Traj. Usage & time (sec.) \\
     \hline
     $2.0$ & $9.23 \times 10^{-2}$ & $3.14$ & $9.23 \times 10^{-2}$ & $95.0$ \\
-    \hline
     $1.0$ & $3.22 \times 10^{-2}$ & $5.48$ & $3.22 \times 10^{-2}$ & $48.8$ \\
-    \hline
     $0.5$ & $6.28 \times 10^{-3}$ & $12.58$ & $6.28 \times 10^{-3}$ & $24.6$ \\
-    \hline
     $0.3$ & $1.00 \times 10^{-3}$ & $31.61$ & $1.00 \times 10^{-3}$ & $14.9$ \\
-    \hline
     $0.2$ & $9.30 \times 10^{-5}$ & $103.69$ & $9.30 \times 10^{-5}$ & $9.9$ \\
-    \hline
     $0.1$ & $0.00$ & $-$ & $0.00$ & $5.0$ \\
     \hline
-  \end{tabular}
-  \centering
-  \caption{$\mathcal{P}(\tau \le \lambda)$ for one-dimensional dynamics. $R_e$ is the
+    \hline
+  \end{tabular*}
+  \centering
+  \caption{$\mathcal{P}(\tau \le \lambda)$ for one-dimensional dynamics.
+    ``Traj. Usage'' is the
 ratio of trajectories satisfying $\tau \le \lambda$ . $N=10^6$ samples are
@@
 \end{figure}
-
+%\FloatBarrier
 %\subsection{Non-gradient dynamical system}
@@
 \subsection{Two dimensional example}
-We choose the potential to be 
+We choose the potential 
 \begin{align}
@@
 
-For the optimal control problem, the cost function is 
+For the optimal control problem, we consider the cost function 
 \begin{align}
@@
 \phi_2\}$, $\{\phi_0, \phi_1, \phi_2, \phi_3\}$ are studied. After finding the coefficients, the cost function (\ref{cost-ex2}) is computed with $N =
-10^6$ samples. Results are tabulated in Table.\ref{tab-ex2-1}. From the dual relation
+10^6$ samples. Results are tabulated in Tab.~\ref{tab-ex2-1}. From the dual relation
 (\ref{dual-relation}), we also compute the optimal cost by Monte Carlo method
@@
 optimal cost ($1.27$) if all basis functions are used. Figure.\ref{fig-ex2-1} shows the optimized potential with different basis functions.
-\begin{table}
-  \begin{tabular}{c|c|c|c|c}
+\begin{figure}[tphb]
+  \centering
+    \includegraphics[width=7.5cm]{./fig/pot_2d_1_0_0.eps}
+    \caption{Potential $V(x)$ for the 2D example. \label{fig-ex2-0}}
+\end{figure}
+\begin{table}[tphb]
+  \begin{tabular*}{0.9\textwidth}{@{\extracolsep{\fill}}ccccc}
+    \hline
     \hline
@@
     $0$ & ($-1.535$, $0.000$, $0.000$, $0.000$) & $3.69$ & $0.67$ & $2.97$ \\
-    \hline
     $1$ & ($-1.379$, $0.744$, $0.000$, $0.000$) & $3.86$ & $0.82$ & $3.16$ \\
-    \hline
     $2$ & ($-0.943$, $1.527$, $1.500$, $0.000$) & $2.48$ & $0.94$ & $1.71$ \\
-    \hline
     $3$ & ($-0.909$, $0.479$, $0.484$, $0.936$) & $1.39$ & $0.46$ & $0.63$ \\
     \hline
-  \end{tabular}
+    \hline
+  \end{tabular*}
   \centering
@@
   computed with $N=10^{6}$ trajectories.
-The last row is the mean value of stopping time $\tau$ under the modified
+The last column is the mean value of stopping time $\tau$ under the modified
 dynamics. \label{tab-ex2-1}}
 \end{table}
-\begin{figure}[tphb]
-  \centering
-    \includegraphics[width=9.0cm]{./fig/pot_2d_1_0_0.eps}
-    \caption{Potential $V(x)$ for the 2D example. \label{fig-ex2-0}}
-\end{figure}
 \begin{figure}[tphb]
@@
 We continue to compute the probability of the rare events $\mathcal{P}(\tau
-\le \lambda)$. We choose a large $\beta = 20.0$, $\mbox{dt} = 0.001$. Three sets of
-basis functions as above are used : $\{\phi_0\}$, $\{\phi_0, \phi_1,
-\phi_2\}$, $\{\phi_0, \phi_1, \phi_2, \phi_3\}$. $(\beta_0, \beta_1 , \beta_2,
+\le \lambda)$. We choose a large $\beta = 20.0$, $\mbox{dt} = 0.001$. 
+For simplicity we only consider basis functions : 
+%$\{\phi_0\}$, $\{\phi_0, \phi_1, \phi_2\}$, 
+$\{\phi_0, \phi_1, \phi_2, \phi_3\}$. $(\beta_0, \beta_1 , \beta_2,
 \beta_3, \beta_4, \beta_5, \beta_6) = (4.0, 8.0, 10.0, 12.0, 16.0, 18.0,
 20.0)$ and $N_i = 10^4, 0\le i \le 5, N_6 = 10^5$ are used when solving the coefficients.
-The numerical results are shown in Table.\ref{tab-ex2-2}--\ref{tab-ex2-5} and Figure.\ref{fig-ex2-2}, Figure.\ref{fig-ex2-3}.
-Table.\ref{tab-ex2-2}--\ref{tab-ex2-4} shows the
-results of importance sampling with different sets of basis functions and the results
-of standard Monte Carlo method are given in Table.\ref{tab-ex2-5}. $N=5 \times
+The numerical results are shown in Tab.~\ref{tab-ex2-4}--\ref{tab-ex2-5}.
+Tab.~\ref{tab-ex2-4} shows the
+results of importance sampling and the results
+of standard Monte Carlo method are given in Tab.~\ref{tab-ex2-5}. $N=5 \times
 10^{6}$ samples are used to compute the probability. We clearly
-see that, comparing to the standard Monte Carlo method, the (relative) variances are always largely reduced with importance sampling, especially when $\lambda$ is small. Comparing different basis
+see that, comparing to the standard Monte Carlo method, the (relative) variances are always largely reduced with importance sampling, especially when $\lambda$ is small. 
+\begin{comment}
+Comparing different basis
 functions, we observe that the performance is better when all four basis functions are used.
-The cumulative distribution function is computed using importance sampling
-with all four basis functions and is shown in Figure.\ref{fig-ex2-2}. The optimized potentials for different $\lambda$ are presented in Figure.\ref{fig-ex2-3}. We see once again that large force is
-needed when $\lambda$ is small.
-
+The cumulative distribution function is computed using importance sampling and is shown in Figure.\ref{fig-ex2-2}. 
+The optimized potentials for different $\lambda$ are presented in Figure.\ref{fig-ex2-3}. We see once again that large force is needed when $\lambda$ is small.
 \begin{table}
@@
     \hline
+    \hline
   $\lambda$ & coefficients & $P(\tau \le \lambda)$ & relative variance & $R_e$ & time (sec.) \\
@@
     $12.0$ & ($-0.271$, $0.000$, $0.000$, $0.000$) & $9.36\times 10^{-3}$ & $7.00$ & $2.08\times 10^{-1}$ & $3219.7$ \\ 
-    \hline
     $10.0$ & ($-0.305$, $0.000$, $0.000$, $0.000$) & $6.47\times 10^{-3}$ & $8.25$ & $1.87\times 10^{-1}$ & $2688.7$ \\
-    \hline
     $8.0$ & ($-0.337$, $0.000$, $0.000$, $0.000$) & $3.97\times 10^{-3}$ & $8.93$ & $1.50\times 10^{-1}$ & $2184.4$ \\
-    \hline
     $6.0$ & ($-0.391$, $0.000$, $0.000$, $0.000$) & $2.03\times 10^{-3}$ & $9.13$ & $1.13\times 10^{-1}$ & $1660.6$ \\
-    \hline
     $4.0$ & ($-0.471$, $0.000$, $0.000$, $0.000$) & $7.50\times 10^{-4}$ & $12.38$ & $6.86\times 10^{-2}$ & $1121.4$ \\
-    \hline
     $2.0$ & ($-0.593$, $0.000$, $0.000$, $0.000$) & $1.19\times 10^{-4}$ & $16.30$ & $2.33\times 10^{-2}$ & $569.6$ \\
-    \hline
     $1.0$ & ($-0.945$, $0.000$, $0.000$, $0.000$) & $7.26\times 10^{-6}$ & $31.37$ & $1.86\times 10^{-2}$ & $286.2$ \\
-    \hline
     $0.5$ & ($-1.654$, $0.000$, $0.000$, $0.000$) & $3.60\times 10^{-8}$ & $52.00$ & $2.03\times 10^{-2}$ & $143.3$ \\
+    \hline
     \hline
@@
     \hline
+    \hline
     $\lambda$ & coefficients & $P(\tau \le \lambda)$ & relative variance & $R_e$ & time (sec.) \\
@@
   $12.0$ & ($-0.255$,  $0.030$, $0.029$, $0.000$) & $9.40\times 10^{-3}$ & $6.76$ & $1.95\times 10^{-1}$ & $4724.0$ \\
-    \hline
   $10.0$ & ($-0.291$,  $0.035$, $0.040$, $0.000$) & $6.40\times 10^{-3}$ & $6.85$ & $1.81\times 10^{-1}$ & $3967.5$ \\
-    \hline
   $8.0$ & ($-0.341$,  $0.049$, $0.051$, $0.000$) & $3.98\times 10^{-3}$ & $17.59$ & $1.63\times 10^{-1}$ & $3190.7$ \\
-    \hline
   $6.0$ & ($-0.400$,  $0.065$, $0.069$, $0.000$) & $2.03\times 10^{-3}$ & $10.99$ & $1.30\times 10^{-1}$ & $2418.9$ \\
-    \hline
   $4.0$ & ($-0.446$,  $0.127$, $0.127$, $0.000$) & $7.47\times 10^{-4}$ & $11.95$ & $8.30\times 10^{-2}$ & $1638.6$ \\
-    \hline
   $2.0$ & ($-0.519$,  $0.317$, $0.318$, $0.000$) & $1.21\times 10^{-4}$ & $13.95$ & $6.78\times 10^{-2}$ & $860.5$ \\
-    \hline
   $1.0$ & ($-0.719$,  $0.712$, $0.729$, $0.000$) & $7.24\times 10^{-6}$ & $20.08$ & $1.43\times 10^{-1}$ & $457.4$ \\
-    \hline
   $0.5$ & ($-1.035$,  $1.681$, $1.552$, $0.000$) & $3.71\times 10^{-8}$ & $12.97$ & $2.69\times 10^{-1}$ & $210.4$ \\
+    \hline
     \hline
@@
 \end{table}
-\begin{table}
-  \begin{tabular}{c|c|c|c|c|c}
-    \hline
-    $\lambda$ & coefficients & $P(\tau \le \lambda)$ & relative variance & $R_e$ & time (sec.) \\
-    \hline
-    $12.0$ & ($-0.270$, $-0.029$, $-0.034$, $0.074$) & $9.39 \times 10^{-3}$ & $6.65$ & $2.26\times 10^{-1}$ & $5446.4$ \\
-    \hline
-    $10.0$ & ($-0.319$, $-0.062$, $-0.064$, $0.108$) & $6.43\times 10^{-3}$ & $7.87$ & $2.35\times 10^{-1}$ & $4475.0$ \\
-    \hline
-    $8.0$ & ($-0.373$,  $-0.107$, $-0.098$, $0.155$) & $3.95\times 10^{-3}$ & $7.52$ & $2.35\times 10^{-1}$ & $3563.8$ \\
-    \hline
-    $6.0$ & ($-0.438$,  $-0.158$, $-0.148$, $0.232$) & $2.03\times 10^{-3}$ & $8.17$ & $2.48\times 10^{-1}$ & $2643.7$ \\
-    \hline
-    $4.0$ & ($-0.505$,  $-0.223$, $-0.209$, $0.343$) & $7.43\times 10^{-4}$ & $8.12$ & $2.76\times 10^{-1}$ & $1745.5$ \\
-    \hline
-    $2.0$ & ($-0.602$,  $-0.269$, $-0.254$, $0.556$) & $1.19\times 10^{-4}$ & $11.28$ & $3.85\times 10^{-1}$ & $893.5$ \\
-    \hline
-    $1.0$ & ($-0.787$,  $-0.003$, $-0.003$, $0.630$) & $7.16\times 10^{-6}$ & $12.15$ & $4.18\times 10^{-1}$ & $457.4$ \\
-    \hline
-    $0.5$ & ($-1.064$, $0.865$, $0.896$, $0.566$) & $3.71\times 10^{-8}$ & $9.28$ & $4.24\times 10^{-1}$ & $235.9$ \\
-    \hline
-  \end{tabular}
-  \centering
-  \caption{2D example, $\mathcal{P}(\tau \le \lambda)$ using importance sampling method.
-Basis functions $\{\phi_0, \phi_1, \phi_2, \phi_3\}$.
-$R_e$ is the ratio of trajectories satisfying $\tau \le \lambda$ under the modified dynamics.
-\label{tab-ex2-4}}
+\end{comment}
+\begin{table}[htbp]
+  \begin{tabular*}{0.9\textwidth}{@{\extracolsep{\fill}}ccccc}
+    \hline
+    \hline
+    $\lambda$ & coefficients & $P(\tau \le \lambda)$ & relative variance &
+    Traj. Usage  \\
+    \hline
+    $12.0$ & ($-0.270$, $-0.029$, $-0.034$, $0.074$) & $9.39 \times 10^{-3}$ & $6.65$ & $2.26\times 10^{-1}$  \\
+    $10.0$ & ($-0.319$, $-0.062$, $-0.064$, $0.108$) & $6.43\times 10^{-3}$ & $7.87$ & $2.35\times 10^{-1}$ \\
+    $8.0$ & ($-0.373$,  $-0.107$, $-0.098$, $0.155$) & $3.95\times 10^{-3}$ & $7.52$ & $2.35\times 10^{-1}$ \\
+    $6.0$ & ($-0.438$,  $-0.158$, $-0.148$, $0.232$) & $2.03\times 10^{-3}$ & $8.17$ & $2.48\times 10^{-1}$ \\
+    $4.0$ & ($-0.505$,  $-0.223$, $-0.209$, $0.343$) & $7.43\times 10^{-4}$ & $8.12$ & $2.76\times 10^{-1}$  \\
+    $2.0$ & ($-0.602$,  $-0.269$, $-0.254$, $0.556$) & $1.19\times 10^{-4}$ & $11.28$ & $3.85\times 10^{-1}$  \\
+    $1.0$ & ($-0.787$,  $-0.003$, $-0.003$, $0.630$) & $7.16\times 10^{-6}$ & $12.15$ & $4.18\times 10^{-1}$  \\
+    $0.5$ & ($-1.064$, $0.865$, $0.896$, $0.566$) & $3.71\times 10^{-8}$ & $9.28$ & $4.24\times 10^{-1}$ \\
+    \hline
+    \hline
+  \end{tabular*}
+  \centering
+  \caption{2D example, $\mathcal{P}(\tau \le \lambda)$ using importance
+  sampling method. Basis functions $\{\phi_0, \phi_1, \phi_2, \phi_3\}$.
+``Traj. Usage'' is the ratio of trajectories satisfying $\tau \le \lambda$ under the modified dynamics. \label{tab-ex2-4}}
 \end{table}
-
-\begin{table}
-  \begin{tabular}{c|c|c|c|c}
-    \hline
-    $\lambda$ & $P(\tau \le \lambda)$ & relative variance & $R_e$ & time (sec.) \\
-    \hline
-    $12.0$  & $9.44\times 10^{-3}$ & $10.25$ & $9.44\times 10^{-3}$ & $3431.5$ \\
-    \hline
-    $10.0$  & $6.38\times 10^{-3}$ & $12.48$ & $6.38\times 10^{-3}$ & $2861.9$ \\
-    \hline
-    $8.0$  & $3.94\times 10^{-3}$ & $15.90$ & $3.94\times 10^{-3}$ & $2291.0$ \\
-    \hline
-  $6.0$  & $2.04\times 10^{-3}$ & $22.10$ & $2.04\times 10^{-3}$ & $1719.1$ \\
-    \hline
-  $4.0$  & $7.40\times 10^{-4}$ & $36.74$ & $7.40\times 10^{-4}$ & $1146.9$ \\
-    \hline
-  $2.0$  & $1.25\times 10^{-4}$ & $89.44$ & $1.25\times 10^{-4}$ & $573.6$ \\
-    \hline
-  $1.0$  & $6.80\times 10^{-6}$ & $383.48$ & $6.80\times 10^{-6}$ & $287.0$ \\
-    \hline
-  $0.5$  & $0.00$ & $-$ & $0.00$ & $143.7$ \\
-    \hline
-  \end{tabular}
+\begin{table}[htbp]
+  \begin{tabular*}{0.9\textwidth}{@{\extracolsep{\fill}}cccc}
+    \hline
+    \hline
+    $\lambda$ & $P(\tau \le \lambda)$ & relative variance & Traj. Usage  \\
+    \hline
+    $12.0$  & $9.44\times 10^{-3}$ & $10.25$ & $9.44\times 10^{-3}$  \\
+    $10.0$  & $6.38\times 10^{-3}$ & $12.48$ & $6.38\times 10^{-3}$  \\
+    $8.0$  & $3.94\times 10^{-3}$ & $15.90$ & $3.94\times 10^{-3}$  \\
+  $6.0$  & $2.04\times 10^{-3}$ & $22.10$ & $2.04\times 10^{-3}$  \\
+  $4.0$  & $7.40\times 10^{-4}$ & $36.74$ & $7.40\times 10^{-4}$  \\
+  $2.0$  & $1.25\times 10^{-4}$ & $89.44$ & $1.25\times 10^{-4}$  \\
+  $1.0$  & $6.80\times 10^{-6}$ & $383.48$ & $6.80\times 10^{-6}$  \\
+  $0.5$  & $0.00$ & $-$ & $0.00$ \\
+    \hline
+    \hline
+  \end{tabular*}
   \centering
   \caption{2D example, $\mathcal{P}(\tau \le \lambda)$ using standard Monte
-  Carlo. $R_e$ is the ratio of trajectories satisfying $\tau \le \lambda$
+  Carlo method. ``Traj. Usage'' is the ratio of trajectories satisfying $\tau \le \lambda$
 under the orininal dynamics. \label{tab-ex2-5}}
 \end{table}
-
+\begin{comment}
 \begin{figure}[tphb]
@@
 \end{figure}
-
+\end{comment}
+%\FloatBarrier
 \subsection{Conformational transition of butane dissolved in water}
@@
 \end{align}
-The $V_b(r)$ is the energy due to the covalent bonds in the syste that
+The $V_b(r)$ is the energy due to the covalent bonds in the system that
 are not constrained as a constant. $b_k(r)$ is the instantaneous
@@
 \emph{cis} conformation, and the last one is usually called
-\emph{trans} conformation. And the transition between the
+\emph{trans} conformation. And the transitions between the
 conformations are slow motion in the system. The bonds and bond angles
@@
 event that $\tau$ being smaller than some given threshold $\lambda$,
-i.e.~$P(\tau \leq\lambda)$. We firstly run an equilibrium simulation
+i.e.~$\mathcal{P}(\tau \leq\lambda)$. We firstly run an equilibrium simulation
 (i.e.~with $u = 0$ in Eq.~\eqref{eqn:langevin-p}) at 300~K, with
@@
   u_i(r, w) = -\nabla_{r_i}V_{\textrm{ctrl}}(\phi(r))=
-  -\nabla_{r_i} \bigg[ \sum_{k=1}^{N_c} w_k\cos(k \phi (r)) \bigg].
+  -\nabla_{r_i} \bigg[ \sum_{k=1}^{N_c} \omega_k\cos(k \phi (r)) \bigg].
 \end{align}
@@
 on the conformation of butane, and it is also fully justified to
-design control also based on these degrees of freedoms (DOFs), however, We
+design control also based on these degrees of freedoms (DOFs), however, we
 will show later that they do not contribute significantly in the
@@
 base functions $N_c$ is chosen to be 8. Numerical results show that
-the coefficient of the 8th base $w_8$ is already very small comparing
+the coefficient of the 8th base $\omega_8$ is already very small comparing
 to the dominant contributing bases, that means the number of base is
@@
 We considered $\lambda = 1.0$, 0.5, 0.2 and 0.1~ps and shown the
-result Fig.~\ref{fig:tmp2}, Tab.~\ref{tab:tmp1} and~\ref{tab:tmp2}.
-From each of the 5000 equilibrium conformation, 4 trajectories with
+results in Fig.~\ref{fig:tmp2}, Tab.~\ref{tab:tmp1} and~\ref{tab:tmp2}.
+From each of the 5000 equilibrium conformations, 4 trajectories with
 different randoms seeds are simulated for each $\lambda$, so we
-generate 20000 trajectories for each case.  The control $u(r, w)$ is
+generate 20000 trajectories for each case.  The control $u(r, \omega)$ is
 calculated by simulations of butane in vacumm, in which all other
@@
 practically, the control calculated in the corresponding vacumm
-systems perform good enough in the in-water systems, because we find
+systems perform well enough in the in-water systems, because we find
 that no further iteration is need to refine the control. Furthermore,
-in the vacumm system, the probability $P(\tau\leq\lambda) =
+in the vacumm system, the probability $\mathcal{P}(\tau\leq\lambda) =
 2.19\times10^{-2}$, $9.32\times 10^{-3}$, $1.35\times 10^{-3}$ and
@@
     Butane dissolved in water. The variance reduced estimate of the probability
-    $P (\tau \leq \lambda)$ by applying control on the dihedral angle.
+    $\mathcal{P} (\tau \leq \lambda)$ by applying control on the dihedral angle.
     See the text for more details of the definition of the control.
     Column ``Error'' denotes the standard deviation of
-    estimating the probablity $P (\tau \leq \lambda)$.
+    estimating the probablity $\mathcal{P} (\tau \leq \lambda)$.
     Column ``Var'' denotes the reduced variation of the esimate.
@@
     \hline\hline
-    $\lambda$ [ps] & $P (\tau \leq \lambda)$ & Error & Var & Accel. & Traj. Usage \\\hline
+    $\lambda$ [ps] & $\mathcal{P} (\tau \leq \lambda)$ & Error & Var & Accel. & Traj. Usage \\\hline
     0.1 & $4.34\times 10^{-5}$ & $1.16\times 10^{-5}$ & $2.31\times10^{-6}$ &18.8 & 0.7\%\\
@@
     \hline\hline
-    $\lambda$ [ps] & $P (\tau \leq \lambda)$ & Error & Var & Accel. & Traj. Usage \\\hline
+    $\lambda$ [ps] & $\mathcal{P} (\tau \leq \lambda)$ & Error & Var & Accel. & Traj. Usage \\\hline
     0.2 & $7.46\times 10^{-4}$ & $2.36\times 10^{-4}$ & $7.46\times10^{-4}$ & 1.0 & 0.1\%\\
@@
 $V_{\textrm{ctrl}}(\phi)$. We only show the effective energy in the
-range $[40^\circ, 150^\circ]$, because this this is the initial
+range $[40^\circ, 150^\circ]$, because the initial
 conformations start in range $[40^\circ, 80^\circ]$, and the
-trajectoies are stopped when they reach $\phi = 150^\circ$. For a easy
+trajectoies are stopped when they reach $\phi = 150^\circ$. For an easy
 comparison, all effective energies are switched by a constant, so that
@@
 the barrier around $\phi = 120^\circ$ is almost kept the same.  The
-resultant probability $P (\tau \leq \lambda)$ calculted by the
+resultant probability $\mathcal{P} (\tau \leq \lambda)$ calculated by the
 variance reduced simulation is summarized in Tab.~\ref{tab:tmp1},
-which is consistent with the Tab.~\ref{tab:tmp2} that presents the
+which is consistent with Tab.~\ref{tab:tmp2} that presents the
 brute force results. Since the variance is redueced, the error of the
@@
 diffusion processes can be investigated and importance sampling for more
-meaningful quantity such as $\mathcal{P}(\tau \le \lambda)$ can be considered.
+meaningful quantity such as the probability $\mathcal{P}(\tau \le \lambda)$ can be considered.
 

### 2013-12-24 writedown/first.hitting/cross_entropy/cross_entropy.tex
small revise

@@
   \caption{
-    Butane dissolved in water. The variance reduced estimate of the probability
-    $\mathcal{P} (\tau \leq \lambda)$ by applying control on the dihedral angle.
+    Butane dissolved in water. The probability
+    $\mathcal{P} (\tau \leq \lambda)$ calculated by the important sampling with
+    control on the dihedral angle.
     See the text for more details of the definition of the control.
-    Column ``Error'' denotes the standard deviation of
+    Column ``Error'' denotes the statistical uncertainty of
     estimating the probablity $\mathcal{P} (\tau \leq \lambda)$.
-    Column ``Var'' denotes the reduced variation of the esimate.
+    If the  trajectories are statistically independent, the expected error is $\sqrt{\textrm{Var}/M}$, where
+    $M$ is the number of trajectories. If the trajectories are not independent, the error
+    can be estimated by the block average method~\cite{frenkel2001understanding}.
+    Column ``Var'' denotes the reduced variance of the esimate.
     ``Accel.''(acceleration) is the ratio between
@@
     Butane dissolved in water, brute force simulations
-    without any variance reduction. The meaning of the columns are the same
+    without any important sampling, i.e.~standard Monte Carlo. The meaning of the columns are the same
     as Tab.~\ref{tab:tmp1}.}\label{tab:tmp2}
@@
 resultant probability $\mathcal{P} (\tau \leq \lambda)$ calculated by the
-variance reduced simulation is summarized in Tab.~\ref{tab:tmp1},
+important sampling simulation is summarized in Tab.~\ref{tab:tmp1},
 which is consistent with Tab.~\ref{tab:tmp2} that presents the
 brute force results. Since the variance is redueced, the error of the
-statistical esitmate is more precise in the variance reduced case. The
+statistical esitmate is more precise in the important sampling case. The
 reduced variance is shown in the fourth column of Tab.~\ref{tab:tmp1},
@@
 with respect to the brute force simulation or the benefit from the
-variance reduction, which is exactly the ratio between the brute force
+important sampling, which is exactly the ratio between the brute force
 variance and the reduced variance. The column of trajectory usage

### 2014-04-29 committor/template.alanine/solve.w.sh
add solve.w.sh

@@ -1,0 +1,29 @@
+#!/bin/bash
+
+source parameters.sh
+
+nbase=`wc -l base.info | awk '{print $1}'`
+nbaseUse=6
+calAb_command=/home/mi/wanghan/study/noneq.msm/committor/template.alanine//tools/analysis/cal.Ab.noAvg
+
+nnotUse=`echo "$nbase - $nbaseUse" | bc`
+rangeStart=`echo "$nbase - $nbaseUse + 1" | bc`
+rangeCode="$rangeStart:$nbase"
+
+$calAb_command -n 100 --target-indicate $fht_coreset_target --input-coreset $fht_coreset_data &> calAb.out
+
+grep "a = " calAb.out				> solve.w.m
+grep "b = " calAb.out				>> solve.w.m
+echo "[v d] = eig(a);"				>> solve.w.m
+echo "d"					>> solve.w.m
+echo "rd = d($rangeCode,$rangeCode);"		>> solve.w.m
+echo "vb = v'*b'"				>> solve.w.m
+echo "rvb = vb($rangeCode)"			>> solve.w.m
+echo "rrsh = rd^-1 * rvb"			>> solve.w.m
+echo "rsh = [zeros($nnotUse,1); rrsh]"		>> solve.w.m
+echo "w = v * rsh"				>> solve.w.m
+
+
+
+
+rm -f calAb.out

### 2014-04-29 committor/template.alanine/tools/analysis/app/cal.Ab.noAvg.cpp
new printing format

@@ -267,174 +267,67 @@ int main(int argc, char * argv[])
   
   printf ("# committor\n");
   printf ("# value   var    error\n");
-  printf ("%e   %e   %e\n", ba.getAvg(), ba.getVar(), ba.getAvgError());
+  printf ("%e   %e   %e\n\n", ba.getAvg(), ba.getVar(), ba.getAvgError());
 
-  printf ("# vect b\n");
+  printf ("b = [ ");
   for (int ii = 0; ii < nBase; ++ii){
     vecb[ii].calculate();
     printf ("%e   ", vecb[ii].getAvg());
   }
-  printf ("\n# vect b error\n");
+  printf ("];\n\nbe = [ ");
   for (int ii = 0; ii < nBase; ++ii){
-    printf ("%e   ", vecb[ii].getAvgError());
+    printf ("%e ", vecb[ii].getAvgError());
   }
-  printf ("\n#;\n");
-  printf ("# mat A\n");
-  for (int ii = 0; ii < nBase; ++ii){
-    for (int jj = 0; jj < nBase; ++jj){
-      matA[ii][jj].calculate();
-      printf ("%e   ", matA[ii][jj].getAvg());
-    }
-    printf (";\n");
-  }
-  printf ("# mat A error\n");
+  printf (" ];\n\n");
+  printf ("a = [ ");
   for (int ii = 0; ii < nBase; ++ii){
     for (int jj = 0; jj < nBase; ++jj){
       matA[ii][jj].calculate();
-      printf ("%e   ", matA[ii][jj].getAvgError());
+      printf ("%e ", matA[ii][jj].getAvg());
     }
-    printf (";\n");
+    if (ii != nBase - 1) printf ("; ");
   }
+  printf (" ];\n\n");
+  printf ("ae = [ \n");
+  for (int ii = 0; ii < nBase; ++ii){
+    for (int jj = 0; jj < nBase; ++jj){
+      matA[ii][jj].calculate();
+      printf ("%e ", matA[ii][jj].getAvgError());
+    }
+    if (ii != nBase - 1) printf ("; ");
+  }
+  printf (" ];\n\n");
 
-
-  //   double sum_j = 0.;
-  //   if (ms.inSet(anglev[0])) {
-  //     record[angleIdx(anglev[0], bin)].push_back (0.0);
-  //     record_j[angleIdx(anglev[0], bin)].push_back (0.0);
-  //     countFound ++;
-  //   }
-  //   else {
[…]

### 2014-04-29 committor/template.alanine/solve.w.sh
fix bug in calculate w, make it complete

@@ -5,25 +5,70 @@ source parameters.sh
 nbase=`wc -l base.info | awk '{print $1}'`
 nbaseUse=6
 calAb_command=/home/mi/wanghan/study/noneq.msm/committor/template.alanine//tools/analysis/cal.Ab.noAvg
+outfileName=base.new.info
 
 nnotUse=`echo "$nbase - $nbaseUse" | bc`
 rangeStart=`echo "$nbase - $nbaseUse + 1" | bc`
 rangeCode="$rangeStart:$nbase"
 
+echo "# calculating the matrix A and rsh b"
 $calAb_command -n 100 --target-indicate $fht_coreset_target --input-coreset $fht_coreset_data &> calAb.out
+echo "# done"
 
+echo "# solve w by octave"
 grep "a = " calAb.out				> solve.w.m
 grep "b = " calAb.out				>> solve.w.m
 echo "[v d] = eig(a);"				>> solve.w.m
 echo "d"					>> solve.w.m
 echo "rd = d($rangeCode,$rangeCode);"		>> solve.w.m
-echo "vb = v'*b'"				>> solve.w.m
-echo "rvb = vb($rangeCode)"			>> solve.w.m
-echo "rrsh = rd^-1 * rvb"			>> solve.w.m
-echo "rsh = [zeros($nnotUse,1); rrsh]"		>> solve.w.m
-echo "w = v * rsh"				>> solve.w.m
-
-
-
+echo "vb = v'*b';"				>> solve.w.m
+echo "rvb = vb($rangeCode);"			>> solve.w.m
+echo "rrsh = rd^-1 * rvb;"			>> solve.w.m
+echo "rsh = [zeros($nnotUse,1); rrsh];"		>> solve.w.m
+echo "w = v * rsh;"				>> solve.w.m
+echo "fid=fopen('$outfileName', 'w');"		>> solve.w.m
+echo "fprintf(fid,'%.10f\n',w);"		>> solve.w.m
+echo "fclose(fid);"				>> solve.w.m
+echo "# done"
 
 rm -f calAb.out
+
+octave solve.w.m
+
+newNbase=`wc -l $outfileName | awk '{print $1}'`
+if [ $newNbase -ne $nbase ]; then
+    echo "# the number of bases in the output is not equal to the number of lines in base.info, something must be wrong"
+    exit
+fi
+expectedNbase=`echo "$fht_base_phi_number + $fht_base_psi_number" | bc`
+if [ $newNbase -ne $expectedNbase ]; then
[…]

### 2014-04-29 committor/template.alanine/solve.w.sh
may not calculate  Ab every time.

@@ -11,9 +11,15 @@ nnotUse=`echo "$nbase - $nbaseUse" | bc`
 rangeStart=`echo "$nbase - $nbaseUse + 1" | bc`
 rangeCode="$rangeStart:$nbase"
 
-echo "# calculating the matrix A and rsh b"
-$calAb_command -n 100 --target-indicate $fht_coreset_target --input-coreset $fht_coreset_data &> calAb.out
-echo "# done"
+if [ ! -f success.dir.name ]; then
+    echo "# no simulation is done, exit"
+    exit
+fi
+if [ ! -f calAb.out ] || [ calAb.out -ot success.dir.name ]; then
+    echo "# calculating the matrix A and rsh b"
+    $calAb_command -n 100 --target-indicate $fht_coreset_target --input-coreset $fht_coreset_data &> calAb.out
+    echo "# done"
+fi
 
 echo "# solve w by octave"
 grep "a = " calAb.out				> solve.w.m
@@ -30,8 +36,6 @@ echo "fid=fopen('$outfileName', 'w');"		>> solve.w.m
 echo "fprintf(fid,'%.10f\n',w);"		>> solve.w.m
 echo "fclose(fid);"				>> solve.w.m
 echo "# done"
-
-rm -f calAb.out
 
 octave solve.w.m
 

### 2014-04-29 committor/template.alanine/tools/analysis/app/cal.Ab.noAvg.cpp
fix bug of printing

@@ -288,7 +288,7 @@ int main(int argc, char * argv[])
     if (ii != nBase - 1) printf ("; ");
   }
   printf (" ];\n\n");
-  printf ("ae = [ \n");
+  printf ("ae = [ ");
   for (int ii = 0; ii < nBase; ++ii){
     for (int jj = 0; jj < nBase; ++jj){
       matA[ii][jj].calculate();

### 2014-04-29 committor/template.alanine/solve.w.sh
print relartive error of matrixe

@@ -24,8 +24,12 @@ fi
 echo "# solve w by octave"
 grep "a = " calAb.out				> solve.w.m
 grep "b = " calAb.out				>> solve.w.m
+grep "ae = " calAb.out				>> solve.w.m
+grep "be = " calAb.out				>> solve.w.m
+echo "relative_a_error=ae./(abs(a))"		>> solve.w.m
+echo "relative_b_error=be./(abs(b))"		>> solve.w.m
 echo "[v d] = eig(a);"				>> solve.w.m
-echo "d"					>> solve.w.m
+echo "diag(d)"					>> solve.w.m
 echo "rd = d($rangeCode,$rangeCode);"		>> solve.w.m
 echo "vb = v'*b';"				>> solve.w.m
 echo "rvb = vb($rangeCode);"			>> solve.w.m

### 2014-04-29 accelerate/template.butane.resp/run.pert.sh
backup

@@ -54,7 +54,7 @@ do
     cp $pert_equi_result/grompp.mdp	$my_dir
     cp $pert_equi_result/angle.ndx	$my_dir
     cp $pert_equi_result/topol.top	$my_dir
-    cd $my_dir
+    cd $my_dir    
     rm -f run.log
     ../../tools/potential/tab.coulomb --scale $pert_rescale --scale-ele-start 1000 --scale-ele-end 1001 --scale-ele-scale-0 1.0 --scale-vdw-scale-0 $pert_vdw_rescale_scale0 --scale-vdw-start $pert_vdw_rescale_start --scale-vdw-end $pert_vdw_rescale_end --rc $pert_rcut --r-smooth $pert_rsmooth --erf $pert_erf -o table.xvg
     ../../tools/potential/tab.coulomb --scale $pert_rescale --scale-ele-start 1000 --scale-ele-end 1001 --scale-ele-scale-0 1.0 --scale-vdw-scale-0 $pert_vdw_rescale_scale0 --scale-vdw-start $pert_vdw_rescale_start --scale-vdw-end $pert_vdw_rescale_end --rc $pert_rcut --r-smooth $pert_rcut --erf $pert_erf -o tablep.xvg
@@ -64,7 +64,7 @@ do
     ../../tools/parse.top/rescale.param -f topol.top -b $pert_bond_rescale -a $pert_angle_rescale --scal-angle-bond $pert_angle_bond_rescale -d $pert_rescale --scal-cmap $pert_rescale -o out.top
     mv -f out.top topol.top
     set_parameters_pert grompp.mdp
-    start_time=`grep $count $pert_equi_result/equi.frame | awk '{print $2}'`
+    start_time=`grep ^$count $pert_equi_result/equi.frame | awk '{print $2}'`
     echo "# run with command `which grompp`"
     $grompp_command -t $pert_equi_result/traj.trr -time $start_time &> run.log
     if [ $? -ne 0 ]; then

### 2014-04-29 first.hitting/template.inwater/functions.sh
backup

@@ -99,6 +99,7 @@ function set_parameters_fht () {
     fht_nstep=`echo "$fht_time / $fht_dt" | bc -l | cut -d '.' -f 1`
     fht_xtcout_feq=`echo "$fht_frame_feq / $fht_dt" | bc -l | cut -d '.' -f 1`
     fht_xvout_feq=$fht_xtcout_feq
+    fht_xvout_feq=0
     fht_energy_feq_r=`echo "$fht_energy_feq / $fht_dt" | bc -l | cut -d '.' -f 1`
     fht_seed=`date +%s`
     fht_seed=`echo "$fht_seed + $fht_num_conf_use * 10 * $fht_parallel_my_id" | bc`
@@ -110,8 +111,8 @@ function set_parameters_fht () {
     sed -e "/^tau_t/s/=.*/= $fht_taut/g" |\
     sed -e "/^tau_p/s/=.*/= $fht_taup/g" |\
     sed -e "/^nstep/s/=.*/= $fht_nstep/g" |\
-    sed -e "/^nstxout/s/=.*/= 0/g" |\
-    sed -e "/^nstvout/s/=.*/= 0/g" |\
+    sed -e "/^nstxout/s/=.*/= $fht_xvout_feq/g" |\
+    sed -e "/^nstvout/s/=.*/= $fht_xvout_feq/g" |\
     sed -e "/^nstfout/s/=.*/= 0/g" |\
     sed -e "/^nstenergy/s/=.*/= $fht_energy_feq_r/g" |\
     sed -e "/^nstcalcenergy/s/=.*/= $fht_energy_feq_r/g" |\

### 2014-04-29 writedown/accelerate/main.tex
backup

@@
 where
-\begin{align}
+\begin{align}\label{eqn:error}
   \me(\vect y_0, t) = 
@@
 and $W[\vect y_s]$ is defined by
-\begin{align}
+\begin{align}\label{eqn:error-exp}
   W[\vect y_t] = 
@@
 $\hat u(0) = u(0)$, and $\hat u(\infty) = \alpha u(\infty)$.
+\\
+
+\recheck{ A possible idea is to estimate the error~\eqref{eqn:error}
+  by the Girsanov transformation~\eqref{eqn:meas-reweight}. However, we
+  suffer from the following difficulty: since the fast part of the
+  force are generally very stiff, the magnitude of $\vect b_f$ is
+  large comparing to the noise in the system,
+  i.e.~$\tilde\sigma$. In our alanine dipeptide simulation,
+  the typical magnitude of
+  $\frac{\alpha-1}{\tilde{\sigma}}\int_0^{T/\alpha}\vect b_f(\vect
+  y_s)\cdot\dd\vect w_s$ is of order $10^3$, while the magnitude of
+  $\frac{(\alpha-1)^2}{2\tilde\sigma^2} \int_0^{T/\alpha}\vect
+  b_f^2(\vect y_s)\dd t$ is of order $10^7$.  This means that for the
+  trajectories we sampled, the reweighting~\eqref{eqn:meas-reweight}
+  vanishes. There are also trajectories, along which the magnitude of
+  the fast forces is very small (for example the bonds and angles stay
+  around the equilibrium value), which means a significant contribution
+  to the reweighting~\eqref{eqn:meas-reweight}, however  these
+  trajectories are seldomly (actually not) sampled in real
+  simulations.
+
+  When we remove the fast forces from the system, actually the
+  microscopic dynamics of the system changes a lot, that is why the
+  direct application of Girsanov
+  transformation~\eqref{eqn:meas-reweight} does not work.  However, we
+  are only interested by the dynamics of the slow observables, which may NOT
+  be perturbed a lot by removing the fast forces.  The question is how
+  to derive an effective reweighting (or correction) for only the slow variables?
+
+  If the fast varibles only contribute entropy to the slow degrees of
+  freedom, and if during the nonequilibrium dynamics the fast varibles
+  relaxes much faster than the slow variables (short-term entropy
+  contribution is well defined), it may be possible to effectively
+  add back the removed fast force (defined on fast DOFs)  by
+  applying addtional free energy correction $\nabla_{\vect r}( \Delta A)$
+  (on slow DOFs), 
+  where $\Delta A$ denotes the free energy difference between
+  the original system and the perturbed system (fast force removed).
+  % then it may be possible to consider the Girsanov
+  % transformation based on the free energy (on the slow variables)
+  % difference:
+%   \begin{align}\label{eqn:meas-reweight-da}
+%     \frac{\dd\tilde\pathmeas[\vect y_t]}{\dd\hat\pathmeas[\vect y_t]}
+%     = \exp
+%     \Big[
+%     \frac{\alpha-1}{\tilde{\sigma}}\int_0^{T/\alpha} \nabla( \Delta A)\cdot\dd\vect w_s
+%     -
+%     \frac{(\alpha-1)^2}{2\tilde\sigma^2} \int_0^{T/\alpha}\vert \nabla( \Delta A) \vert^2\dd t\,
+%     \Big]
+% \end{align}  
+}
 
@@
   \centering
+  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-01-check-1.eps}
+  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-01-check-2.eps}
+  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-01-check-3.eps}
+  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-01-check-4.eps}
+  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-01-check-5.eps}\\
   \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-02-1.eps}
@@
   \centering
+  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-01-check-1-1.eps}
+  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-01-check-1-2.eps}
+  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-01-check-1-3.eps}\\
+  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-02-1-1.eps}
+  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-02-1-2.eps}
+  \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-02-1-3.eps}\\
   \includegraphics[width=0.195\textwidth]{figs/fig-meta-iflux-04-1-1.eps}

### 2014-04-29 writedown/first.hitting/cross_entropy/cross_entropy.tex
update the data
backup

--- 2013-12-24 writedown/first.hitting/cross_entropy/cross_entropy.tex
@@
 \newcommand{\cL}{{\mathcal L}}
+\newcommand{\cI}{{\mathcal I}}
+\newcommand{\cF}{{\mathcal F}}
+\newcommand{\cD}{{\mathcal D}}
 \newcommand{\bR}{{\mathbb R}}
+\newcommand{\R}{{\mathbb R}}
+\newcommand{\Q}{{\mathbb Q}}
+\newcommand{\N}{{\mathbb N}}
+\newcommand{\bE}{{\mathbf E}}
+\newcommand{\bP}{{\mathbf P}}
+
+\newcommand{\eps}{\epsilon}
+\newcommand{\blambda}{\bm{\lambda}}
 \newcommand{\argmin}{\mathop{\rm argmin}}%
-
-\newcommand{\bE}{{\mathbf E}}
+\newcommand{\bk}[2]{\left\langle #1,#2\right\rangle}
+\newcommand{\ip}{\langle\cdot,\cdot\rangle}
+\newcommand{\di}{{\rm div}\,}
+\newcommand{\vol}{{\rm vol}}
+\newcommand{\wrt}{with respect to }
 
@@
 \begin{document}
-\title{Applications of the cross-entropy method in diffusion processes}
-%\author{Wei Zhang}
+\title{Applications of the cross-entropy method to importance sampling and optimal control of diffusions}
 \author{Wei Zhang\textsuperscript{2}\and Han Wang\textsuperscript{1} \and
@@
 \begin{abstract}
-  The cross-entropy method in the diffusion process setting is studied. The
-  general algorithm is proposed and is applied to design efficient importance sampling for rare events as well as to solve optimal
-  control problem of diffusion process. Several examples are studied
-  numerically to demonstrate this method.
+ We study the cross-entropy method for diffusions. One of the results is a
+  versatile cross-entropy algorithm that can be applied to the design of efficient importance sampling strategies for rare events as well as optimal
+  control problems.
+The approach is based on the minimization of a suitable cross-entropy functional, based on a parametric family of exponentially tilted probability distributions.
+We illustrate the new algorithm with several numerical examples and discuss algorithmic issues and possible extensions of the method.
 \end{abstract}
 \begin{keywords}
-importance sampling, optimal control, cross-entropy method, Monte Carlo
-method, rare event, change of measure. 
+importance sampling, optimal control, cross-entropy method, rare events, change of measure. 
 \end{keywords}
@@
 \end{AMS}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
 \label{intro}
-Diffusion process has been extensively investigated because of its wide
-applications in mathematics, physics, economics, and so on.
-Here we are mainly interested in two topics in diffusion process, i.e. importance sampling and optimal control. 
-For importance sampling, it is one kind of variance reduction technique 
-in Monte Carlo simulations and becomes especially necessary in rare event simulation.
-For optimal control, the purpose is to driven the dynamical system in certain
-optimal way and has strong applicational backgrounds in finance, climate
-science, mechanics \cite{fleming2006}. We emphasize that these two different areas are actually related and can
-be unified as the same problem, i.e. how to change the measure in certain optimal way.
-
-Here we take an applied point of view, i.e. we aim at designing efficient
-algorithms to find the way to do the importance sampling, or to find the
-optimal control strategy. One main reason that designing efficient algorithms
-is urgent and challenging is due to the high dimensionality of the dynamical
-systems, as in the case of climate system, biological system and so on. However, when the dynamical
-systems have multiscale feature, efficient algorithms may be designed by
-taking this feature into account. See \cite{ip-dupuis-multiscale, ip-eric,
-zhws13} on importance sampling and \cite{control_schuette, zlph2013} on
-optimal control for more details.
-
-In this work, we consider these two topics together as how to design the change of
-measure in an optimal way. If we further assume the existence of the optimal
-change of measure $\mu^*$, the problem can be roughly described as how to find the approximation of
-$\mu^*$ under some measure distance. For the choice of measure distances, we will use the Kullback-Leibler
-divergence, which, although not a metric, is
-widely used to quantified the similarity of two probability measures. 
-What is more, given certain target probability measure $\mu^*$, the so-called cross-entropy method has been
-developed recently and provides a general algorithm to find the minimizer of the
-Kullback-Leibler divergence w.r.t. $\mu^*$, among a family of parameterized
-probability measures. It has wide applications in different areas, such as
-Monte Carlo simulation, optimization, and machine learning \cite{ce_tutorial, ce_book}. The
-main purpose of this paper is to formulate the cross-entropy method in the
-context of diffusion process, i.e. for measures defined in path space, and
-discuss the applications of this method in importance sampling and optimal control
-problems.
-
-This paper is organized as follows. In Section~\ref{sec-om-functional},
-using Girsanov theorem, we derive the Onsager-Machlup functional for a given trajectory.
-In Section~\ref{sec-ce-is}, the cross-entropy method in path space is
-formulated for the importance sampling problem.
-In Section~\ref{sec-ocp}, by using the dual relationship between importance
-sampling and optimal control problem, we discuss the cross-entropy method for
-optimal control problem. The cross-entropy algorithm is summarized in
-Section~\ref{sec-ce-algo}. Several numerical examples are studied in
-Section~\ref{sec-examples}. Discussions are presented in
-Section~\ref{sec-discuss}.
-
-\section{Onsager-Machlup functional}
-\label{sec-om-functional}
-Let $O \subset \mathbb{R}^n$, $z_s \in \mathbb{R}^n$ satisfies the SDE 
+This article deals with the application of the cross-entropy method to diffusion processes, specifically, with the application to  
+%Diffusion processes have been extensively investigated because of their wide range of applications in mathematics, physics, economics, and other fields. 
+importance sampling for rare events and optimal control.
+Generally, the cross-entropy method is a Monte-Carlo method that was originally developed for the efficient simulation of rare events in queuing models and that has, in the meantime, been extended to, e.g., combinatorial optimization or analysis of networks \cite{ce_book,ce_tutorial}.
+To our knowledge, however, the cross-entropy approach has not been analysed or used in connection with diffusion processes, even though there have been significant research activities in the direction of efficient algorithms for importance sampling and optimal control high dimensional multiscale diffusions; see, e.g., \cite{ip-dupuis-multiscale, ip-eric,
+zhws13} for some ideas related to importance sampling of rare events or \cite{control_schuette, zlph2013} for problems in 
+optimal control.
+
+%Importance sampling is a quite general variance reduction technique 
+%for Monte Carlo simulations that is essential in rare event simulation, especially if applied to Monte Carlo simulations in path space. 
+%Seemingly different, optimal control aims at solving the problem of how to driven the dynamical system in certain
+%optimal way and has strong applicational backgrounds in finance, climate
+%science, mechanics \cite{fleming2006}.
+
+We will exploit the fundamental duality between importance sampling and optimal control, which arises due to the fact that both problems admit a variational formulation that boils down to finding an optimal transformation of the underlying path space measure \cite{WangDupuis2004}.
+Algorithms for computing an optimal change of measure problems usually seek an approximation of the optimal measure \wrt some distance on the space of (probability) measures.
+Here, we will use the Kullback-Leibler
+divergence, which, although not a metric, is a numerically convenient and widely used similarity measure for probability measures. 
+The cross-entropy method then provides a general algorithm to find the minimizer of the
+Kullback-Leibler divergence among a family of parameterized
+probability measures, and the main purpose of this paper is to formulate the 
+method in the context of diffusions, i.e.~for probability measures on the space of continuous paths, and discuss 
+its application to importance sampling and optimal control.
+
+This paper is organized as follows. The cross-entropy method in path space is
+outlined in Section~\ref{sec-ce-is} and discussed in the context of importance
+sampling and the dual optimal control problem. Section~\ref{sec-ce-algo} is
+devoted to the formulation of the cross-entropy algorithm for diffusions. Several numerical examples are studied in Section~\ref{sec-examples}. We summarize our findings in Section~\ref{sec-discuss}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{The cross-entropy method in path space}
+\label{sec-ce-is}
+In this section, we discuss how to use the cross-entropy
+method for stochastic differential equations.
+The mathematical set-up will be largely based on the application of the method to importance sampling, which is the most standard application of the cross-entropy method in the literature.
+The associated (dual) optimal control problem will be briefly discussed at the end of this section.
+
+\subsection{Problem set-up}
+
+Consider  $z_s \in \mathbb{R}^n$ satisfying 
+\begin{equation}\label{dynamics-1}
+\begin{aligned}
+  d z_s & =  b(z_s) ds + \sqrt{2\eps }\, dw_s\,, \quad 0\le s\le T\\
+  z_{0} & = x
+\end{aligned}
+\end{equation}
+where $\eps>0$ is constant, $b\colon \mathbb{R}^n \to \mathbb{R}^n$ is a smooth, locally Lipschitz-continuous vector field, and $w$ is 
+$n$-dimensional Brownian motion. Further let $O \subset \mathbb{R}^n$ be open and bounded and call 
+\begin{equation}\label{tau}
+\tau = \inf\{s>0\colon \, z_s \notin O\}  
+\end{equation}
+the first exit time of the set $O\subset\mathbb{R}^n$.
+%
+%Let $\mathcal{D}$ be certain subset in the path space and
+%$\textbf{1}_{\mathcal{D}}$ be the
+%characteristic function on $\mathcal{D}$. 
+In the following we will use $Z$ to denote a path trajectory
+$\{z_s\}$ and use the notation $z_s \in \mathbb{R}^n$ for the state at time $s$. 
+Accordingly we denote by $F(Z)$ a path functional that, throughout this paper, is assumed to be of the form
+\begin{equation}
+F(Z) = \exp\left(-\frac{1}{\eps}\int_0^{\tau\wedge T} G(z_s) \,ds - \frac{1}{\eps} H(z_{\tau\wedge T})\right).
+  \label{path-fun-f}
+\end{equation}
+for some continous and bounded functions $G,H\colon\R^{n}\to\R$ and with $\tau\wedge T=\min\{\tau,T\}$. We consider a Monte Carlo method to compute the quantity
+\begin{align}
+  \ell(x) =\bE(F(Z)),
+  \label{mean-l}
+\end{align}
+with $\bE(\cdot) = \bE(\cdot|\,z_0 = x)$ denoting the conditional expectation over all realizations of (\ref{dynamics-1}) starting at $z_{0}=x$.  
+A special and interesting case is when $G = 0$ and $H=-\eps\log{\mathbf 1}_{O^{c}}$ with $O^{c}=\R^{n}\setminus O$ denoting the complement of the set $O$, in which case 
+\begin{align}
+  \bE({\mathbf 1}_{O^{c}}(z_{\tau\wedge T})) = \bP(\tau\le T),
+  \label{mean-chi}
+\end{align}
+is the probability of trajectories starting at $z_{0}=x$ to reach the boundary of $O$ before time $T$.
+In many cases, this quantity provides details of the distribution of the stopping time $\tau$ and can be used to understand the behaviour of the dynamics.
+
+%%%%%%%%%%%%%%%%%%
+\subsection{Cross-entropy method for importance sampling}
+
+We now formulate the cross-entropy method in path space following the relevant literature \cite{ce_tutorial,ce_book}.
+To this end we first introduce  the general concept of importance sampling in path space: Suppose we are able to generate path samples from a family of probability
+measures $\{\mu_{\blambda}\}_{\blambda \in
+\cF}$ on the space of continuous functions $C([0,T],\R^{n})$ that are
+parametrized by $\blambda \in \cF\subset \mathbb{R}^m$ where the  dynamics
+(\ref{dynamics-1}) corresponds to $\blambda = \bm{0}$; for the sake of
+simplicity,  we set $\cF=\mathbb{R}^m$.
+We use the shorthand $\mu = \mu_{\bm{0}}$ and refer to $\mu_{\blambda\neq\bm{0}}$ as the \emph{tilted probability measure}.
+We further assume that every $\mu_{\blambda}$ has a probability density $f(\cdot;\blambda)$ \wrt the scaled Wiener measure $\nu_{\eps}$.\footnote{The scaled measure $\nu_{\eps}$ is the probability measure induced by the scaled Brownian motion $\sqrt{2\eps} w_{s}$ on the space $C([0,T],\R^{n})$.
+It is related to the standard Wiener measure underlying the standard Brownian motion $w_{s}$ by rescaling of time, which follows from the fact that $w_{s}$ and $\sqrt{2\eps}\,w_{s/(2\eps)}$ have the same law.}
+
+\subsubsection*{Importance sampling.}
+%
+The idea of importance sampling is---instead of drawing samples from the measure $\mu$---to generate samples from an alternative probability measure $\eta = g(\cdot)\nu_{\eps}$ that is absolutely continuous \wrt $\mu$, but that yields better Monte-Carlo estimators, e.g., having smaller variance or smaller relative error \cite{Liu2008}.
+Using independent draws from $\eta$ an unbiased Monte Carlo estimator of (\ref{mean-l}) is then given by
+\begin{align}
+  \ell_N = \frac{1}{N} \sum_{i=1}^{N}  \frac{F(\tilde{Z}_i)f(\tilde{Z}_i;\bm{0})}{g(\tilde{Z}_i)}\,,
+\end{align}
+where the trajectories $\tilde{Z}_i, i = 1, \cdots, N$ are independent realizations from $\eta$.
+It is well known that the optimal measure $\eta^*$ that minimizes the variance of the estimator has the density
+\begin{align}
+  g^*(Z) = \frac{F(Z)f(Z;\bm{0})}{\ell}\,.
+  \label{optimal-pdf}
+\end{align}
+It is easy to see that the thus defined $\eta^{*}$ yields a zero variance estimator, note however that it depends on $\ell=\bE(F(Z))$, which is the quantity that we want to compute.
+
+The idea of the cross-entropy method is to find the best approximation of $\eta^*$ among the family
+$\mu_{\blambda}, \blambda \in \cF$ of tilted probability measures. The approach is based on minimizing the 
+Kullback-Leibler divergence, which in our case can be defined as follows:
+given two probability measures $\mu_1 = g_1\nu_{\eps}$, $\mu_2 =
+g_2\nu_{\eps}$ that are absolutely continuous \wrt the scaled Wiener measure, the \emph{Kullback-Leibler divergence} or \emph{relative entropy} between $\mu_{1}$ and $\mu_{2}$ is defined as
+\begin{align}
+  D(\mu_1, \mu_2) = \bE_{\mu_1}\!\left(\log \frac{d\mu_1}{d\mu_2}\right) 
+  \label{kl-dist}
+\end{align}
+where the expectation \wrt the measure $\mu_{1}$ is defined as
+\begin{align}
+  \bE_{\mu_1}\!\left(\log \frac{d\mu_1}{d\mu_2}\right) = \int \log \frac{d\mu_1}{d\mu_2}\,d\mu_{1} = \int
+  g_1\log g_1\, d\nu_{\eps} - \int g_1\log g_2\,d\nu_{\eps}\,.
+  \label{kl-dist2}
+\end{align}
+
+\subsubsection*{Cross-entropy method I.}
+%
+The cross entropy method now seeks an optimal change of measure by solving the minimization task
+\begin{align}
+  \min_{\blambda\in\R^{m}} D(\eta^*, \mu_{\blambda}).
+  \label{kl-dist-d}
+\end{align}
+for the tilt parameter $\blambda\in\R^{m}$.
+Not knowing what $\eta^{*}$ is, this still sounds like an infeasible minimization problem.
+It turns out, however, that we need to know $\eta^{*}$ only up to a constant prefactor, which in our case, since $x$ is fixed, eliminates the unknown quantity $\ell=\ell(x)$.
+Using (\ref{optimal-pdf}) and (\ref{kl-dist}), the minimization problem is equivalent to the following maximization problem
+\begin{align}\label{ce-1}
+  \max_{\blambda \in \bR^m} \bE_{\mu} ( F(Z) \log
+  f(Z;\blambda))\,.
+\end{align}
+
+For the efficient numerical solution of (\ref{ce-1}) it is often convenient to allow for drawing samples from a probability measure that  somehow ``in between'' $\mu$ and $\mu_{\blambda^{*}}$.
+This will give us some extra freedom to use, e.g., an iterative solver for the maximization problem (\ref{ce-1}).
+Letting $\bm{v}\in \bR^{m}$  denote an arbitrary family parameter, our maximization problem reads
+\begin{align}
+  \max_{\blambda \in \mathbb{R}^m} \bE_{\mu_{\bm{v}}} ( F(Z)
+  h(Z;\bm{v}) \log f(Z;\blambda)),
+  \label{ce-2}
+\end{align}
+where
+\begin{align}
+  h(Z;\bm{v}) = \frac{f(Z;\bm{0})}{f(Z;\bm{v})}.
+\end{align}
+An unbiased estimator of (\ref{ce-2}) is 
+\begin{align}
+\max_{\blambda \in \mathbb{R}^m} \frac{1}{N} \sum_{i=1}^{N} F(\tilde{Z}_i)
+h(\tilde{Z}_i; \bm{v}) \log f(\tilde{Z}_i;\blambda) 
+\label{ce-2-discrete}
+\end{align}
+where $\tilde{Z}_i, i = 1, \cdots, N$ are independent realizations generated from $\mu_{\bm{v}}$.
+The necessary condition for $\blambda^{*}$ being a maximizer of (\ref{ce-2-discrete}) is obtained by taking the gradient \wrt $\blambda$:
+\begin{align}
+\sum_{i=1}^{N} F(\tilde{Z}_i) h(\tilde{Z}_i; \bm{v}) \nabla_{\blambda} \log
+  f(\tilde{Z}_i;\blambda) = 0\,.
+  \label{max-gradient}
+\end{align}
+
+The degree of difficulty when solving (\ref{max-gradient}) numerically of course depends on the parameterization of the tilted family of distributions that will be introduced in Section \ref{sec-ce-algo} below.
+
+%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Cross-entropy method for optimal control}
+\label{subsec-ocp}
+In this section, we consider applying the cross-entropy method to solve optimal control problems
+for diffusion processes. 
+%
+To this end consider the optimal control problem with cost
+function \cite{control_schuette,zlph2013}
+\begin{equation}\label{cost}
+  J(u) = \bE_{\mu_u}\left(\int_0^{\tau\wedge T} G(z_{s}) + \frac{1}{4} |u_{s}|^2
+ds  + H(z_{\tau\wedge T})\right),
+\end{equation}
+with bounded continuous function $G,H\colon\R^{n}\to\R$ and $u_{s}\in\R^{n}$ being a measurable control that will specified below.
+The expectation $\bE_{\mu_u}$ \wrt the probability measure $\mu_u$ is understood as the expectation over all realizations of the controlled dynamics
+\begin{equation}
+ \begin{aligned}
+   d z_s &= \left(b(z_s) +  u_{s}\right)ds + \sqrt{2\eps }\,dw_s, \quad 0 \le s \le T \\
+    z_0&=x
+  \end{aligned}
+  \label{ctr-sde}
+\end{equation}
+We suppose that $G\ge 0$ and, for the ease of notation, we set $H=0$.
+We wish to minimize (\ref{cost}) under the constraint (\ref{ctr-sde}) and over all measurable controls $u$ that are adapted to the filtration generated by the Brownian motion driving the dynamics (\ref{ctr-sde}).
+Then, given suitable conditions on the vector field $b$, it is known that this optimal control problem has a unique solution in form of a Markovian feedback control.
+That is, for some continuous and bounded function $c\colon[0,T]\times \R^{n}\to\R^{n}$, it holds that
+\begin{equation}\label{feedback}
+ \hat{u}_{s} = c(s, z_{s})\,.
+\end{equation}
+Now call $\mu, \hat{\mu}$ the probability measures on the path space  $C([0,T],\R^{n})$ corresponding to (\ref{ctr-sde}) with $u = 0$ and $\hat{u}$.
+Then, using the Legendre-type dual relation, we have \cite{DaiPra1996,HaEtal14}
+    \begin{align}
+      J(\hat{u}) = -\eps \log \bE_\mu\left(\exp\left(-\frac{1}{\eps}\int_0^{\tau\wedge T} G(z_s) \,ds\right) \right) ,
+      \label{dual-relation}
+    \end{align}
+    where, by Jensen's inequality (see \cite[Sec.~VI.2]{fleming2006}), we know that $\hat{\mu}$-a.s.
+    \begin{align}
+    \exp\left(-\frac{1}{\eps}\int_0^{\tau\wedge T} G(z_s)\,ds\right)\frac{d\mu}{d\hat{\mu}} =
+    \bE_\mu\left(\exp\left(-\frac{1}{\eps}\int_0^{\tau\wedge T} G(z_s)\, ds\right)\right) 
+    \label{jensen-as}
+  \end{align}
+From the above we conclude that (see \cite[Sec.~VI.3]{fleming2006} for details)
+  \begin{equation}\label{cost-2}
+  \begin{aligned}
+    J(u) &= \bE_{\hat{\mu}}\!\left(\int_0^{\tau\wedge T} \left(G(z_s) + \frac{1}{4} |u_{s}|^2\right)\,ds\right)\frac{d\mu_u}{d\hat{\mu}}  \\
+    & = J(\hat{u}) + \bE_{\hat{\mu}}\!\left(\left(\eps \log
+    \frac{d\mu}{d\hat{\mu}} + \frac{1}{4} \int_0^{\tau\wedge T} |u_{s}|^2\,ds\right)\frac{d\mu_u}{d\hat{\mu}}\right)
+\end{aligned}
+\end{equation}
+After some rearrangement and simplification, we get the following simple relationship: 
+\begin{align}
+J(u) = J(\hat{u}) + \eps D(\mu_u, \hat{\mu}).
+\label{J-entropy}
+\end{align}
+where $D(\cdot, \cdot)$ is the Kullback-Leibler divergence as defined in (\ref{kl-dist}).
+
+\subsubsection*{Cross-entropy method II.}
+%
+Computing the optimal control $\hat{u}$ can be tedious, or is infeasible if the dynamics are 
+high dimensional. As a remedy we suggest again to find the best approximation of $\hat{\mu}=\hat{g}(\cdot)\nu_{\eps}$ among a suitably 
+defined family $\mu_{\blambda}, \blambda \in \cF$ of tilted probability measures that are absolutely continuous \wrt $\nu_{\eps}$. 
+%
+Sticking to the notation from Section \ref{sec-ce-is}, it readily follows that the minimizer of (\ref{J-entropy}) has the density   
+\begin{align}
+  \hat{g}(Z) \propto \exp\left(-\frac{1}{\eps}\int_0^{\tau\wedge T} G(z_s) \,ds\right) f(Z; 0)\,
+\end{align}
+\wrt $\nu_{\eps}$ (cf.~\pageref{optimal-pdf}). By Girsanov's theorem
+there is a one-to-one correspondence between the control force $u=u(\blambda)$ and a certain family of \emph{exponentially tilted} probability measures $\mu_{\blambda}$. 
+Instead of minimizing (\ref{cost}) subject to the dynamics (\ref{ctr-sde}), we solve the constrained optimization problem  
+\begin{align}
+  \min_{\blambda \in \mathbb{R}^m} J(u(\blambda)).
+  \label{J-w}
+\end{align}
+subject to the dynamics (\ref{ctr-sde}). From (\ref{J-entropy}), we know that solving (\ref{J-w})
+is equivalent to minimizing the Kullback-Leibler divergence $D$ between
+$\mu_u$ and $\hat{\mu}$, which, however, is still not easy. On the other hand, inspired by 
+the discussions in Section \ref{sec-ce-is}, we can apply cross-entropy method 
+to minimize the relaxed entropy functional $D(\hat{\mu}, \mu_u)$ rather than $D(\mu_{u},\hat{\mu})$. As a consequence, 
+the problem is reduced to the case in Section \ref{sec-ce-is}.
+
+Note that the relaxed problem solved by the cross-entropy method is different from (\ref{J-w}), as the Kullback-Leibler divergence is not symmetric.
+Yet both (\ref{J-w}) and its relaxed version agree at the minimum, therefore the hope is that the latter yields a reasonable approximation of the optimal control problem---given that the family of tilted distributions is cleverly chosen.
+
+\FloatBarrier
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Cross-entropy algorithm}
+\label{sec-ce-algo}
+
+In this section, we will specify the family  $\{\mu_{\blambda}\}_{\blambda \in
+\cF}$ of tilted probability measures that we are going to use for the
+procedure introduced above and formulate the cross-entropy algorithm. As a
+first step, let $\mu$  denote the probability measure on $C([0,T],\R^{n})$
+that is induced by the dynamics (\ref{dynamics-1}) and let $\nu_{\eps}$ be scaled Wiener measure associated with 
 \begin{align}
   \begin{split}
-  d z_s &= b(z_s) ds + \sqrt{2\beta^{-1}} dw_s, \quad 0 < s < \tau  \\
-  z_0&=x
-  \end{split}
-\label{dynamics-1}
-\end{align}
-$\beta >0$ is constant, $\tau = \inf_{s>0} \{s;\, z_s \in O\}$ is some
-stopping time, $b : \mathbb{R}^n \rightarrow \mathbb{R}^n$, $w$ is a
-$n$-dimensional Brownian motion.
-
-We derive the Onsager-Machlup functional for dynamics (\ref{dynamics-1}) in
-this section.
-Let $\mu$ and $\mu_0$ be the probability measures on path space generated by
-(\ref{dynamics-1}) and the scaled Brownian motion
-\begin{align}
-  \begin{split}
-  dz_s &= \sqrt{2\beta^{-1}}dw_s, \hspace{3mm} 0 < s < \tau \\
+  dz_s &= \sqrt{2\eps }\,dw_s, \hspace{3mm} 0 \le  s \le T \\
    z_0 &= x .
@@
 \end{align}
-From Girsanov theorem \cite{oksendalSDE}, we have 
-\begin{align}
-  \frac{d\mu}{d\mu_0} &= \exp\Big(\sqrt{\frac{\beta}{2}}\int_0^\tau
-  \langle b(z_s), dw_s\rangle + \frac{\beta}{4}\int_0^\tau |b(z_s)|^2ds\Big) \notag\\
-  &=\exp\Big(\frac{\beta}{2}\int_0^\tau \langle b(z_s),dz_s\rangle -
-  \frac{\beta}{4}\int_0^\tau |b(z_s)|^2ds\Big) 
-\end{align}
-and
-\begin{align}
-  d\mu_0 = \exp\Big(-\frac{\beta}{4}\int_0^\tau
-  \Big|\frac{dz}{ds}\Big|^2ds\Big) d\nu\,
-\end{align}
-where $\nu$ is the Wiener measure. Together we obtain
-\begin{align}
-  d\mu &=\exp\Big(-\frac{\beta}{2} I_{1}(z)\Big)d\nu
-  \label{measure-1}
-\end{align}
-where
-\begin{align}
-  I_{1}(z) = \frac{1}{2}\int_0^\tau \Big|\frac{dz}{ds} - b(z_s)\Big|^2 ds
-\end{align}
-and the integral is interpreted as Ito integral.
-We could also rewrite it as Stratonovich integral using the relationship
-\begin{align}
-  \int_0^\tau \langle b(z_s), dz_s\rangle = \int_0^\tau b(z_s) \circ dz_s - \beta^{-1} \int_0^\tau \operatorname{div}\,b(z_s) ds
-\end{align}
-and therefore obtain 
-\begin{align}
-  d\mu = \exp \Big(-\frac{\beta}{2} I_2(z) \Big) d\nu\,
-\end{align}
-where functional 
-\begin{align}
-  I_2(z) = \frac{1}{2}\int_0^\tau \Big|\frac{dz}{ds}- b(z_s)\Big|^2 ds +
-  \beta^{-1} \int_0^\tau \operatorname{div}\,b(z_s) ds
+By Girsanov's theorem \cite{oksendalSDE}, we have 
+\begin{equation}\label{girsanov1}
+  {d\mu}= \exp\left(-\frac{1}{\eps}S(Z)\right) d\nu_{\eps}\,, 
+\end{equation}
+with the action
+\begin{equation}\label{girsanov2}
+\begin{aligned}
+  S(Z) & = -\sqrt{\frac{\eps}{2}}\int_0^{\tau\wedge T}
+  b(z_s)\cdot dw_s - \frac{1}{4}\int_0^{\tau\wedge T} |b(z_s)|^2\,ds\\
+  &=-\frac{1}{2}\int_0^{\tau\wedge T} b(z_s)\cdot dz_s +
+  \frac{1}{4}\int_0^{\tau\wedge T} |b(z_s)|^2ds \,
+\end{aligned}
+\end{equation}
+where the stochastic integral is interpreted in the sense of It\^o \cite{oksendalSDE}.
+%
+A remark is in order. 
+%%
+\begin{remark}
+  We could rewrite (\ref{girsanov1})--(\ref{girsanov2}) as a Stratonovich integral using the relationship
+\begin{align}
+  \int_0^{\tau\wedge T} b(z_s)\cdot dz_s = \int_0^{\tau\wedge T} b(z_s) \circ dz_s - \eps  \int_0^{\tau\wedge T} \operatorname{div}\,b(z_s)\,ds\,,
+\end{align}
+by which we obtain 
+\begin{equation}\label{girsanov3}
+  {d\mu} =\exp\left(\frac{1}{2\eps}\int_0^{\tau\wedge T} b(z_s)\circ dz_s -
+  \frac{1}{4\eps}\int_0^{\tau\wedge T} (|b(z_s)|^2 + 2\eps \operatorname{div}\,b(z_s))\,ds\right)d\nu_{\eps} \,
+\end{equation}
+The associated action functional reads
+\begin{align}
+  S(Z)= -\frac{1}{2}\int_0^{\tau\wedge T} b(z_s)\cdot dz_s +
+  \frac{1}{4}\int_0^{\tau\wedge T} |b(z_s)|^2ds +
+  \frac{\eps}{2}  \int_0^{\tau\wedge T} \operatorname{div}\,b(z_{s}) ds
   \label{om-fun}
 \end{align}
-is often called Onsager-Machlup functional \cite{om1978, om_functional_pinski}. 
-\begin{remark}
-  The introduction of Onsager-Machlup functional and Stratonovich integral is
-  not necessary. However, it is helpful to allow us to follow the presentation of
-  \cite{ce_tutorial}, \cite{ce_book} and to compare with the case of random
-  variables.
+It is closely related to what is known as the \emph{Onsager-Machlup functional} in the physical literature. See \cite{om1978,
+om_functional_pinski} for details. We will stick to It\^o interpretation of (\ref{girsanov2}) in the
+following.
 \end{remark}
-\begin{remark}
-  \label{remark-general}
-  The Onsager-Machlup functional can also be generalized to the following dynamics.
-  \begin{enumerate}
-    \item
-      SDEs with non-constant diffusion coefficients. We consider 
-      \begin{align}
-	\vspace{-9mm}
+%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Choosing a family of path space measures}
+\label{subsec-om-functional}
+
+We will confine our attention to a special class of tilted probability distributions that is suggested by the 
+the optimal control problem from in Section \ref{subsec-ocp}. To this end, let  
+ $\{\phi_i\}_{1\le i \le m}$ denote a set of continuously differentiable basis functions $\phi_i\colon [0,T]\times \mathbb{R}^n\rightarrow \mathbb{R}$. 
+ The cross-entropy method will be based on realizations of 
+ \begin{equation}
+\begin{aligned}
+  d z_s &= \left(b(z_s) + c(s,z_{s};\blambda) \right)ds + \sqrt{2\eps } dw_s, \quad 0 \le s \le T \\
+     z_0 &=x\,,
+     \end{aligned}
+\label{dynamics-2}
+\end{equation}
+with 
+\begin{align}
+  c(s,x;\blambda) = 2\sum_{i = 1}^{m} \lambda_i \nabla \phi_i(s,x)\,.
+  \label{control-force}
+\end{align}
+(The scaling factor $2$ is merely conventional.) It follows from Girsanov's theorem that  
+the associated path probability measure $\mu_{\blambda}$ has a density $f(\cdot;\blambda)$ \wrt the scaled Wiener measure $\nu_{\eps}$. It 
+is given by the usual exponential expression
+\begin{equation}\label{girsanov4}
+  f= \exp\left(\frac{1}{2\eps}\int_0^{\tau\wedge T} (b(z_s)+c(s,z_{s};\blambda))\cdot dz_s 
+  - \frac{1}{4\eps}\int_0^{\tau\wedge T} |b(z_s) + c(s,z_{s};\blambda)|^2\,ds\right) \,.
+\end{equation}
+As a consequence, we can generate independent samples from $\mu_{\blambda}$ by 
+repeatedly simulating the controlled dynamics (\ref{dynamics-2}). Since the tilting parameter 
+$\blambda=(\lambda_{1},\ldots,\lambda_{m})$ enters linearly, the associated 
+cross-entropy maximization problem (\ref{ce-2-discrete}) is strictly convex and thus has a unique solution. 
+Note that, indeed, $\mu_{\bm{0}} = \mu$ is the probability measure corresponding (\ref{dynamics-1}).
+
+\subsubsection*{Cross-entropy method III.}
+Defining 
+\begin{equation}\label{girsanov5}
+   f(Z;\blambda) = \exp\left(-\frac{1}{\eps}S(Z;\blambda)\right) \,, 
+\end{equation}
+with the action
+\begin{equation}\label{girsanov6}
+  S 
+ =-\frac{1}{2}\int_0^{\tau\wedge T} (b(z_s)+c(s,z_{s};\blambda))\cdot dz_s +
+  \frac{1}{4}\int_0^{\tau\wedge T} |b(z_s)+c(s,z_{s};\blambda)|^2ds \,
+\end{equation}
+and noting that 
+\begin{align}
+\nabla_{\blambda} \log f(Z;\blambda) = -\frac{1}{\eps}  \nabla_{\blambda} S(Z;\blambda),
+\end{align}
+the necessary optimality condition (\ref{max-gradient}) can be recast as a linear system of equations: 
+\begin{align}
+A\blambda = \bm{r},
+\end{align}
+where $A=(A_{ij})_{1\le i,j\le m}$ and $\bm{r}=(r_{i})_{1\le i\le m}$ with  
+\begin{align}
   \begin{split}
-      d z_s &= b(z_s) ds + \sqrt{2\beta^{-1}} \sigma(z_s)\,dw_s, \quad 0 < s < \tau \\
-      z_0&=x
-    \end{split}
-    \label{dynamics-1-sigma}
-    \end{align}
-    where $a(z) = \sigma(z)\sigma(z)^T$ is $n \times n$ positive definite matrix. We have 
-    $d\mu = \exp(-\frac{\beta}{2} I_1(z)) d\nu$, with 
-    \begin{align*}
-      I_1(z) = \frac{1}{2} \int_0^\tau \Big|\frac{dz}{ds} -
-      b(z_s)\Big|_{a(z)}^2 ds
-    \end{align*}
-     where Ito integral is used and $|v|_a = \langle v, a^{-1}v\rangle$,
-     $\forall v \in \mathbb{R}^n$.
-    \item
-      System contains both stochastic and deterministic equations. For
-      simplicity, here we consider dynamics 
-      \begin{align}
-      \begin{split}
-	dx_s & = b_1(x_s, y_s)\, ds \\
-	dy_s & = b_2(x_s, y_s)\, ds + \sqrt{2\beta^{-1}} dw_s
-      \end{split}
-      \label{ode-sde}
-      \end{align}
-      where $x_s \in \mathbb{R}^k, y_s, w_s \in \mathbb{R}^{n-k}$, $b_1 :
-      \mathbb{R}^k \times \mathbb{R}^{n-k} \rightarrow \mathbb{R}^k$, $b_2 :
-      \mathbb{R}^k \times \mathbb{R}^{n-k} \rightarrow
-      \mathbb{R}^{n-k}$. We can still assign a probability measure $\mu$ on the
-      \textbf{trajectory space}, i.e. $\forall z = (x,y) $ is a trajectory, we have 
-      $d\mu = \exp(-\frac{\beta}{2} I_1(z)) d\nu$, where $\nu$ is Wiener
-      measure on $\mathbb{R}^{n-k}$ and 
-      \begin{align*}
-	I_1(z) = \frac{1}{2} \int_0^\tau \Big|\frac{dy}{ds} - b_2(x_s, y_s)
-	\Big|^2 ds\,.
-      \end{align*}
-      Notice that only the stochastic equations are used.
-  \end{enumerate}
-\end{remark}
-\section{Cross-entropy method for importance sampling}
-\label{sec-ce-is}
-In this section, we discuss how to apply the cross-entropy
-method \cite{ce_tutorial, ce_book} to guide the
-importance sampling in diffusion process.
-
-Let $\mathcal{D}$ be certain subset in the path space and
-$\textbf{1}_{\mathcal{D}}$ be the
-characteristic function on $\mathcal{D}$. 
-In the following, for simplicity, we will use $z$ to denote a path trajectory
-$\{z_s\}$ and keep the notation $z_s \in \mathbb{R}^n$ for the state at time $s$. 
-Let $F(\cdot)$ be certain path functional and will be written as $F(z)$ for
-path $z$. 
-We consider Monte Carlo method with importance sampling to compute the quantity
-\begin{align}
-  l = \mathrm{E}(F(z)~|~z_0 = x)
-  \label{mean-l}
-\end{align}
-where we assume $F(\cdot)$ has the following type
-\begin{align}
-  F(z) = \textbf{1}_{\mathcal{D}}(z) \quad \mbox{or} \quad F(z) = \exp(-\beta\int_0^\tau G(z_s) \,ds).
-  \label{path-fun-f}
-\end{align}
-  A special and interesting case is when $\mathcal{D} = \{z, \tau(z) \le \lambda\}$,
-  $\lambda > 0$ is fixed. Then $l = \mathrm{E}(\textbf{1}_{\mathcal{D}}(z)) =
-  \mathcal{P}(\tau \le \lambda)$, i.e. the probability of the trajectories
-  stop before time $\lambda$. This quantity may provide us some details of the
-  distribution of the stopping time $\tau$ and help us understand the
-  behaviors of the dynamics.
-
-We suppose $m$ basis functions on state space $\{\phi_i\}_{1\le i \le m}$ are
-available, where $\phi_i : \mathbb{R}^n\rightarrow \mathbb{R}$. 
-We will add a control force to the original dynamics (\ref{dynamics-1})
-and the importance sampling will be realized by simulating trajectories from
-this modified dynamics. Specifically, given $\bm{\omega} = (\omega_1, \omega_2,
-\cdots, \omega_m)^T \in \mathbb{R}^m$,
-the corresponding control force is state-dependent and is supposed to be 
-\begin{align}
-  u(x; \bm{\omega}) = \sum\limits_{i = 1}^{m} \omega_i \nabla \phi_i(x), \quad x \in
-  \mathbb{R}^n\,.
-  \label{control-force}
-\end{align}
-Therefore, for fixed parameter $\bm{\omega}$, the controlled dynamics is 
-\begin{align}
-\begin{split}
-  d z_s &= b(z_s) ds + \sqrt{2}u(z_s;\bm{\omega}) ds + \sqrt{2\beta^{-1}} dw_s, \quad t \le s \le \tau \\
-     z_0 &=z
-  \end{split}
-\label{dynamics-2}
-\end{align}
-Let $\mu_{\bm{\omega}}$ be the probability measure of dynamics (\ref{dynamics-2}). 
-From Section \ref{sec-om-functional},  we know $\mu_{\bm{\omega}}$ has density 
-$f(z;\bm{\omega}) = \exp(-\frac{\beta}{2} I(z ; \bm{\omega}))$ w.r.t. Wiener
-measure $\nu$, with
-\begin{align}
-  I(z ; \bm{\omega}) = \frac{1}{2}\int_0^\tau \Big|\frac{dz}{ds}- b(z_s) -
-  \sqrt{2}\sum\limits_{i = 1}^{m} \omega_i \nabla \phi_i(z_s)\Big|^2 ds + \beta^{-1}
-  \int_0^\tau \operatorname{div}\, \Big(b(z_s) + \sqrt{2}\sum\limits_{i =
-  1}^{m} \omega_i \nabla \phi_i(z_s)\Big) ds
-  \label{om-fun-w}
-\end{align}
-and the original dynamics (\ref{dynamics-1}) corresponds to $\bm{\omega} = \bm{0}$ (we use $\mu = \mu_{\bm{0}}$).
-Given a probability measure $\mu_g$ on path space with $d\mu_g = g(z)d\nu$, 
-we can design a unbiased Monte Carlo estimator for (\ref{mean-l}) as 
-\begin{align}
-  l_N = \frac{1}{N} \sum_{i=1}^{N}  \frac{F(z_i)f(z_i;\bm{0})}{g(z_i)}
-\end{align}
-where trajectories $z_i, i = 1, \cdots, N$ are sampled from measure $\mu_g$.
-The optimal measure $\mu_{g^*}$, which gives
-zero-variance estimator, should satisfy 
-\begin{align}
-  g^*(z) = \frac{F(z)f(z;\bm{0})}{l}, \hspace{3mm} \forall z
-  \label{optimal_pdf}
-\end{align}
-However, $g^*(z)$ depends on the path $z$ and is not easy to compute.
-
-The idea of cross-entropy method is to find the best approximation of $g^*(z)$ among
-$f(z;\bm{\omega}), \bm{\omega} \in \mathbb{R}^m$. More specifically, define the
-Kullback-Leibler divergence between
-two measures $d\mu_1 = g_1d\nu$, $d\mu_2 = g_2d\nu$ as
-\begin{align}
-  D(\mu_1, \mu_2) = \mathrm{E}_{\mu_1} \log \frac{d\mu_1}{d\mu_2} = \int
-  g_1\log g_1\, d\nu - \int g_1\log g_2\,d\nu,
-  \label{kl-dist}
-\end{align}
-and we try to minimize
-\begin{align}
-  \min_{\bm{\omega} \in \mathbb{R}^m} D(\mu_{g^*}, \mu_{\bm{\omega}}).
-  \label{kl-dist-d}
-\end{align}
-Using (\ref{optimal_pdf}) and (\ref{kl-dist}), it is equivalent to solve
-\begin{align}
-  \max_{\bm{\omega} \in \mathbb{R}^m} \mathrm{E}_{\mu} ( F(z) \log
-  f(z;\bm{\omega}))
-  = \max_{\bm{\omega} \in \mathbb{R}^m} \mathrm{E}_{\mu_{\bm{v}}} ( F(z)
-  \eta(z;\bm{v}) \log f(z;\bm{\omega})), \hspace{3mm} \forall \bm{v} \in
-  \mathbb{R}^m
-  \label{ce-2}
-\end{align}
-where
-\begin{align}
-  \eta(z;\bm{v}) = \frac{f(z;\bm{0})}{f(z;\bm{v})}.
-\end{align}
-For numerical purpose, we consider the discrete formulation of (\ref{ce-2})
-\begin{align}
-\max_{\bm{\omega} \in \mathbb{R}^m} \frac{1}{N} \sum_{i=1}^{N} F(z_i)
-\eta(z_i; \bm{v}) \log f(z_i;\bm{\omega}) 
-\label{ce-2-discrete}
-\end{align}
-where $z_i, i = 1, \cdots, N$ are trajectories generated from $\mu_{\bm{v}}$. 
-Taking gradient, the maximizer of (\ref{ce-2-discrete}) could be found by solving
-\begin{align}
-  \frac{1}{N} \sum_{i=1}^{N} F(z_i) \eta(z_i; \bm{v}) \nabla_{\bm{\omega}} \log
-  f(z_i;\bm{\omega}) = 0.
-  \label{max-gradient}
-\end{align}
-With formula (\ref{om-fun-w}), equations (\ref{max-gradient}) could be written more explicitly
-as linear equations $A\bm{\omega} = \bm{r}$, where $\forall 1\le i,j\le m$, 
-\begin{align}
-  \begin{split}
-  A_{ij} &= \frac{1}{N}\sum_{k = 1}^{N} F(z_k)\eta(z_k; \bm{v}) \int_0^\tau
-  \nabla\phi_i(z_{k,s})\nabla\phi_j(z_{k,s})\,ds, \\
-     r_i &= \frac{1}{\sqrt{2}N}\sum_{k = 1}^{N} F(z_k)\eta(z_k; \bm{v})
-  \Big[\int_0^\tau \Big(\frac{dz_k}{ds} - b(z_s)\Big)\nabla\phi_i(z_{k,s})\,ds
-  - \beta^{-1} \int_0^\tau \operatorname{div} \nabla \phi_i(z_{k,s})\,ds\Big]
+  A_{ij} &= 2\sum_{k = 1}^{N} F(Z_k)h(Z_k; \bm{v}) \int_0^{\tau\wedge T}
+  \nabla\phi_i(s,z^{k}_{s})\nabla\phi_j(s,z^{k}_{s})\,ds, \\
+     r_i &= \sum_{k = 1}^{N} F(Z_k)h(Z_k; \bm{v})
+  \left(\int_0^{\tau\wedge T} \nabla\phi_i(s,z^{k}_{s})\cdot dz^{k}_{s} - \int_0^{\tau\wedge T}  \nabla\phi_i(s,z^{k}_{s})\cdot b(z^{k}_{s})\,ds\right)\,. 
   \end{split}
@@
 \end{align}
-Notice that (\ref{max-gradient}) holds $\forall \bm{v} \in \mathbb{R}^m$ and
-$A$ is in general a positive definite matrix when $\phi_i$'s are linearly independent.
-\begin{remark}
-  The more general type conditional expectation 
-  $l = \mathrm{E}(F(z)~|~ z_0 \sim \pi)$
-  can be similarly considered,
-where $\pi(dx) = \rho(x)dx$ is a probability measure on $\Omega
-\subset \mathbb{R}^n$, i.e. $\int_\Omega \rho(x)\,dx = 1$. Let $E = \{z ~|~ z \in
-  C([0,\infty), \mathbb{R}^n), z_0 \in \Omega\}$ be the set of trajectories
-  starting from $\Omega$. Denote $\mu^x$ as the aforementioned measure for
-  dynamics (\ref{dynamics-1}) starting from $x \in \mathbb{R}^n$. Further
-  define $\mu^\pi(\mathcal{D}) = \int_{\Omega} \mu^x(\mathcal{D}) \pi(dx)$,
-  where $\mathcal{D}$ is a measurable subset of $E$.
-  We have $d\mu^\pi_{\bm{v}} = f(z;\bm{v})\,d\nu^\pi, \forall
-  \bm{v} \in \mathbb{R}^n$, where $\mu^\pi_{\bm{v}}$ and $\nu^{\pi}$ are
-  defined similarly. The aim is to find the optimal $\bm{\omega} \in
-  \mathbb{R}^m$ and the corresponding measure $\mu^\pi_{\bm{\omega}}$ to do
-  the importance sampling. All the derivations are the same with the above
-  except that, while simulating the trajectories, 
-  the starting point $z_0$ should be generated from $z_0 \sim \pi$.
-\end{remark}
-\begin{remark}
-   The above algorithm can be further applied to dynamics (\ref{ode-sde}), discussed
-   in Remark \ref{remark-general}. In this case, since only stochastic
-   equations appear in the Radon-Nikodym derivative, the control should
-   also be only added to these equations, in order to obtain linear equations
-   with full rank. Specifically, we study the controlled dynamics 
-         \begin{align}
-      \begin{split}
-	dx_s & = b_1(x_s, y_s)\, ds \\
-	dy_s & = b_2(x_s, y_s)\, ds + \sqrt{2} u(x_s, y_s ; \bm{\omega})\, ds + \sqrt{2\beta^{-1}} dw_s
-      \end{split}
-      \label{ode-sde-control}
-      \end{align}
-      with $u(x_s, y_s ; \bm{\omega}) = \sum\limits_{i=1}^{m} \omega_i \nabla \phi_i(x_s,
-      y_s)$ and optimize the coefficients $\bm{\omega}$. All the derivations are
-      the same as above.
-\end{remark}
-\section{Cross-entropy method for optimal control problem}
-\label{sec-ocp}
-In this section, we consider applying the cross-entropy method to solve optimal control problem
-for diffusion process.
-
-Consider the optimal control problem with cost
-function \cite{control_schuette,zlph2013}
-\begin{equation}
-  J(u) = \mathrm{E}_{\mu_u}\left[\int_0^\tau \left(G(z_s) + \frac{1}{2} |u(z_s,s)|^2
-\right) ds ~\bigg|~ z_0 = x\right],
-\end{equation}
-where $G : \mathbb{R}^n\rightarrow \mathbb{R}$ is nonnegative, $u$ is the
-control force and $\mathrm{E}_{\mu_u}$ denotes the expectation under probability measure $\mu_u$,
-which is generated by the controlled dynamics
-of (\ref{dynamics-1-sigma}), namely,
-\begin{align}
- \begin{split}
-   d z_s &= b(z_s) ds + \sqrt{2}\sigma(z_s) u(z_s, s) ds + \sqrt{2\beta^{-1}} \sigma(z_s)\,dw_s, \quad 0 < s < \tau \\
-    z_0&=x
-  \end{split}
-  \label{dynamics-1-sigmma-c}
-\end{align}
-
-Suppose the lower bound of $J(u)$ can be achieved and let $\hat{u} = \argmin J(u)$. For simplicity, we denote $\mu, \hat{\mu}$ as
-the probability measure corresponding to $u \equiv 0$ and $\hat{u}$
-respectively. Using the dual relation \cite{DaiPra1996, zlph2013}
-    \begin{align}
-      -\beta^{-1}\log \mathrm{E}_\mu\left[\exp\Big(-\beta\int_0^\tau G(z_s)\, ds\Big)\right] = \inf_u J(u) = J(\hat{u})
-      \label{dual-relation}
-    \end{align}
-    and Jensen's inequality, we know that 
-    \begin{align}
-    \exp\Big(-\beta\int_0^\tau G(z_s)\,ds\Big)\frac{d\mu}{d\hat{\mu}} =
-    \mathrm{E}_\mu\left[\exp\Big(-\beta\int_0^\tau G(z_s)\, ds\Big)\right],
-    \hspace{3mm} \hat{\mu}-a.s.
-    \label{jensen-as}
-  \end{align}
-From the above we have 
-  \begin{align}
-    J(u) &= \mathrm{E}_{\hat{\mu}} \left\{\Big[\int_0^\tau \Big(G(z_s) + \frac{1}{2} |u(z_s,
-    s)|^2\Big)\,ds\Big]\frac{d\mu_u}{d\hat{\mu}}\right\}  \\
-    & = J(\hat{u}) + \mathrm{E}_{\hat{\mu}}\Big[\Big(\beta^{-1}\log
-    \frac{d\mu}{d\hat{\mu}} + \int_0^\tau \frac{1}{2} |u(z_s,
-    s)|^2\,ds\Big)\frac{d\mu_u}{d\hat{\mu}}\Big]
-\end{align}
-After simplification, we get the following simple relationship 
-\begin{align}
-J(u) = J(\hat{u}) + \beta^{-1}D(\mu_u|\, \hat{\mu}).
-\label{J-entropy}
-\end{align}
-where $D(\cdot, \cdot)$ is the Kullback-Leibler divergence in (\ref{kl-dist}).
-
-We are interested in finding the optimal control $\hat{u}$. For high
-dimensional dynamics ($n$ is large), this becomes difficult.  
-Similarly as the in previous section, we suppose the control force can be written
-as (\ref{control-force}) with $m$ predefined basis functions $\{\phi_i\}_{1\le
-i \le m}$. And our problem is simplified to solve
-\begin{align}
-  \min_{\bm{\omega} \in \mathbb{R}^m} J(u(z;\bm{\omega})).
-  \label{J-w}
-\end{align}
-under dynamics (\ref{dynamics-1-sigmma-c}).
-
-From (\ref{J-entropy}), we know that solving (\ref{J-w})
-is equivalent to minimize the Kullback-Leibler divergence $D$ between
-$\mu_u$ and $\hat{\mu}$, which, however, is still not easy. On the other hand, inspired by 
-the discussions in Section \ref{sec-ce-is}, we can apply cross-entropy method 
-to minimize $D(\hat{\mu} \,|\, \mu_u)$ (notice this is different from
-(\ref{J-w}), because the Kullback-Leibler divergence is not symmetric), and
-hope that (\ref{J-w}) can be solved approximately. Following
-the notations in Section \ref{sec-ce-is} and from (\ref{jensen-as}), we know the optimal measure
-satisfies
-\begin{align}
-  g^*(z) \propto \exp\Big(-\beta\int_0^\tau G(z_s) \,ds\Big) f(z ; 0),
-  \quad \forall z\,,
-\end{align}
-therefore the problem is reduced to the case in Section \ref{sec-ce-is} with $F(z) = \exp(-\beta\int_0^\tau G(z_s) \,ds)$.
-\section{Cross-entropy algorithm}
-\label{sec-ce-algo}
-In this section, we consider how to implement the cross-entropy algorithm in
-order to maximize (\ref{ce-2-discrete}), which is equivalent to (\ref{kl-dist-d}) in the discrete case. 
-We suppose the optimal density $g^*(z)$ is given by (\ref{optimal_pdf}) and
-$F(z)$ takes the forms in (\ref{path-fun-f}).
-
-Ideally, the solution of the maximization problem (\ref{ce-2-discrete}) can be
-obtained by solving the linear equation (\ref{linear-eqn}) without iterations. But in real
-applications, when the expectation value $l$ of $F(z)$ in (\ref{mean-l}) is very rare, it will become
-difficult to estimate the coefficients (\ref{linear-eqn}) and therefore the solution
-will be inaccurate. In diffusion processes, one major reason for this rarity
-is the metastability in the dynamics when $\beta \gg 1$. In this case, the
-trajectories are long and $F(z)$ is small, which means computing
+Here $\bm{v}\in \mathbb{R}^m$ is an arbitrary vector, and $Z_k=(z^{k}_{s})_{0\le s\le T}$ denotes the sample paths of (\ref{dynamics-2}) with control $u_{s}=c(s,z^{k}_{s};\bm{v})$.
+Note that the realizations are generated from $\mu_{\bm{v}}$ and therefore do not depend on $\blambda$.
+Further notice
+$A$ is positive definite matrix if the basis functions $\phi_i$ are linearly
+independent,  which implies that (\ref{linear-eqn}) has a unique solution.
+
+It thus seems that the solution of the discrete maximization problem (\ref{ce-2-discrete}) can be
+obtained by just solving the linear equation (\ref{linear-eqn}). However in real
+applications, when the expectation value $\ell$ of $F(Z)$ in (\ref{mean-l}) is very small so that it is
+difficult to estimate the coefficients (\ref{linear-eqn}) accurately enough, the solution itself is in danger of being inaccurate.
+In many applications, the reason for this is metastability of the dynamics when $\eps\ll 1$.
+In this case, the
+trajectories are long and $F(Z)$ is small, which means computing
 (\ref{linear-eqn}) is both time-consuming and inaccurate. Inspired by the
-original cross-entropy method \cite{ce_tutorial},
-we may overcome this problem by starting from a higher temperature (smaller $\beta$) and compute
-(\ref{linear-eqn}) while decreasing the temperature (increasing $\beta$). The proposed iterative
-method to solve the maximization problem (\ref{ce-2-discrete}) is illustrated in Algorithm \ref{ce-algo}.
+original cross-entropy method \cite{ce_tutorial}, we may overcome this problem by starting from a higher temperature (here: $\eps$) and compute
+(\ref{linear-eqn}) while decreasing the temperature. The proposed iterative
+method to solve (\ref{ce-2-discrete}) is illustrated in Algorithm \ref{ce-algo}.
+
 \begin{algorithm}
-  \caption{Algorithm to solve the maximization problem (\ref{ce-2-discrete}) \label{ce-algo}}
+  \caption{Cross-entropy algorithm \label{ce-algo}}
   \begin{algorithmic}[1]
-    \State Define $0 < \beta_0 < \beta_1 < \cdots <\beta_k = \beta$, set
+    \State Define $\eps_0 > \eps_{1} > \ldots > \eps_k = \eps$, set
     $\bm{v}^{(0)} = 0$.
@@
     \State generate $N_j$ trajectories $z_i, i=1,2,\cdots, N$ from dynamics (\ref{dynamics-2}), with
-    $\bm{\omega} = \bm{v}^{(j)}$, $\beta = \beta_j$.
+    $\blambda = \bm{v}^{(j)}$, $\eps = \eps_j$.
     \State compute the coefficients of $A^{(j)}, \bm{r}^{(j)}$ from (\ref{linear-eqn}) with $\bm{v} =
@@
 \end{algorithm}
-\FloatBarrier
+
+We conclude this subsection with a few remarks on possible extensions of the method.
+\begin{remark}
+It is straightforward to relax restriction of the fixed initial condition and consider distributed initial conditions instead, i.e.~$z_{0}=x\in\R^{n}$ following some probability distribution $\pi$ on $\R^{n}$.
+All considerations and the cross-entropy method remain valid without alterations, if the sum over the $N$ realizations of the dynamics is replaced by the sum over all realization \emph{and} the sum over sufficiently many independent initial conditions $x\sim \pi$.
+\end{remark}
+
+\begin{remark}
+  \label{remark-general}
+  We briefly mention two more possible generalization of the above algorithm. 
+  %
+The first generalization concerns dynamics with multiplicative noise: 
+      \begin{equation}
+      d z_s = b(z_s) ds + \sqrt{2\eps } \sigma(z_s)\,dw_s\,, 
+    \label{dynamics-1-sigma}
+    \end{equation}
+    where the $n \times n$  matrices $a(\cdot) = \sigma(\cdot)\sigma(\cdot)^T$  are positive definite with bounded inverses.
+Defining the weighted scalar product $\bk{u}{v}=u^{T} (a(z))^{-1}v$, then all considerations remain valid, with the dot product $u\cdot v$ being replaced by $\bk{u}{v}$.
+In particular, (\ref{girsanov2}) must be replaced by
+       \begin{equation}\label{I-1-sigma}
+  S(Z) = -\frac{1}{2}\int_0^{\tau\wedge T} \bk{b(z_s)}{dz_s} +
+  \frac{1}{4}\int_0^{\tau\wedge T} \|b(z_s)\|^2ds \,,
+\end{equation}
+     where  $\|v\|=\sqrt{\bk{v}{v}}$ is the norm induced by the scalar product $\ip$. 
+%   
+Another important class of systems are Langevin-type diffusions with degenerate noise:  
+      \begin{equation}
+      \begin{aligned}
+	dx_s & = y_s\, ds \\
+	dy_s & = -\left(\nabla V(x_s) + y_{s}\right)  ds + \sqrt{2\eps } dw_s\,,
+      \end{aligned}
+      \label{ode-sde}
+      \end{equation}
+      Here $(x_s,y_{s}) \in \mathbb{R}^{2n}$ are the state variables and $V\colon\R^{n}\to\R$ is a smooth potential energy that is bounded below 
+      and sufficiently growing at infinity; more general variants of (\ref{ode-sde}) can be considered too, but we refrain from discussing the most general scenario here. 
+      %
+      Langevin diffusions have the 
+      property that, even though the noise is degenerate and hence the tilting of the distribution can be only in the direction of some variables, one has control over the full path space distribution.\footnote{The reason for this lies in the fact that the Langevin equation satisfies a condition known as \emph{complete controllability}, which ensures that noise drives all degrees of freedom in the system \cite{mattingly2002}.}
+      \end{remark}
+
+      \begin{remark}
+      If the terminal time $T$ is large, it is possible to suppress the time dependence of the basis functions
+      and consider only functions $\phi_{j}=\phi_{i}(x)$. In this case the optimal tilting is not explicitly time-dependent as is the case in 
+      optimal stopping problems (see, e.g., \cite{Shiryaev2006}).    
+      \end{remark}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Numerical examples}
@@
 \subsection{One-dimensional double well potential}
-Here we consider the simplest 1D diffusion process with double well potential.
-Let $b(x) = - V'(x), x\in \mathbb{R}^1$, with $V(x) = \frac{1}{2}(x^2 - 1)^2$.
-The dynamics is 
-\begin{align}
-  dz_s = - V'(z_s) ds + \sqrt{2\beta^{-1}} dw_s
-\end{align}
-where $w_s$ is 1D Brownian motion. The potential $V$ has two minima points
-$x_0 = -1, x_2 = 1$, while $x_1 = 0$ is
-a saddle point. Define $\tau = \inf\{s >0, |z_s - x_2| \le 1 \}$, and choose
-the starting point $z_0 = x_0 =
--1$. We define three basis functions of Gaussian type as $\phi_i(x) = \exp(-\frac{(x -
-x_i)^2}{2r^2})$, $i = 0, 1, 2$, and choose $r = 0.5, \beta = 5.0$ (see Figure. \ref{fig-ex1-basis}).
-
-We first study the optimal control problem  
-\begin{align}
-  \inf_{u_s} \mathrm{E} \Big(\tau + \frac{1}{2} \int_0^\tau u_s^2 \,ds~\Big|~
-  z_0 = -1\Big)\,,
+We consider the one-dimensional diffusion process 
+\begin{align}
+  dz_s = - V'(z_s) ds + \sqrt{2\eps } dw_s\,,
+\end{align}
+with $w_s$ standard one-dimensional Brownian motion, $\eps=0.2$  and the double well potential 
+\begin{equation}
+V(x) = \frac{1}{2}(x^2 - 1)^2\,.
+\end{equation}
+The bistable potential $V$ has two minima at $x_0 = -1$ and $x_2 = 1$ and a local maximum at $x_1 = 0$. Define $\tau = \inf\{s >0 \colon z_s - x_2| \le 1 \}$, and choose
+the starting point $z_0 = x_0$ to be the left minimum. As basis function we use three (unnormalized) Gaussians of the form
+\begin{equation}
+\phi_i(x) = \exp\left(-\frac{(x - x_i)^2}{2r^2}\right)\,,\quad i = 0, 1, 2\,,
+\end{equation}
+with $r = 0.5$ (see Fig.~\ref{fig-ex1-basis}).
+Note that the basis functions are independent of time, which is due to the fact that the time dependence of the optimal tilting is relatively weak in our case.
+ %
+
+ \subsubsection*{Optimal control.}
+
+We begin by studying the following optimal control problem: minimize   
+\begin{align}
+  J(u) = \bE \left(\tau + \frac{1}{4} \int_0^{\tau}  u_s^2 \,ds \right)\,,
   \label{ex1-cost}
 \end{align}
-under dynamics
-\begin{align}
-  dz_s = - V'(z_s) ds + \sqrt{2}u_s ds + \sqrt{2\beta^{-1}} dw_s\,.
-\end{align}
-$\mbox{dt} = 1.0 \times 10^{-3}$, $N_j = 10^{4}, j=0,1,2$ are used to compute the
-coefficients $\bm{\omega}$ using Algorithm.\ref{ce-algo}, with $\{\beta_0,\beta_1, \beta_2\}
-= \{1.0, 3.0, 5.0\}$.  Therefore $k=2$ in Algorithm.\ref{ce-algo} and $3$
-iteration steps are performed. Denote $\bm{v}^{(0)} = \bm{0}$ and suppose
-coefficients $\bm{v}^{(j+1)}$ are obtained
-in $j$th steps, $j=0,1,2$. Then applying control force $u(x;\bm{v}^{(j)})$ is
-equivalent to modifying the potential as $V^{(j)}(x) = V(x) - \sqrt{2}\sum\limits_{i=0}^{2} v^{(j)}_i
-\phi_i(x)$. The numerical results are presented in Figure.\ref{fig-ex1-1} and Figure.\ref{fig-ex1-2}.
-In Figure.\ref{fig-ex1-1}, we show the modified potential $V^{(j)}(x), 0\le
-j\le 3$ when different basis functions are used. We can observe that the potential is
-gradually modified during the iterations. In Figure.\ref{fig-ex1-2}, it shows
-the modified potentials $V^{(3)}(x)$ with optimized coefficients
-$\bm{v}^{(3)}$ in Algorithm.\ref{ce-algo} for different basis functions. In
-one-dimension, the optimal control problem (\ref{ex1-cost}) can also be solved
-from a Feymann-Kac type boundary value PDE \cite{control_schuette}. The
-potential (the cost value) obtained in this way can be considered as the exact solution and is also
-shown in Figure.\ref{fig-ex1-2} (The last row of Tab.~\ref{tab-ex1-1}).
-Later on, $N = 10^{6}$ samples are generated to compute
-the cost function (\ref{ex1-cost}) and the results are presented in
-Tab.~\ref{tab-ex1-1}. We can observe that better results are obtained when basis function
-$\phi_1$ is used, since in job 2 and 3, the cost value (1.31) is closer to the exact
-solution (1.25). Notice that this fact is also shown in
-Figure.\ref{fig-ex1-2}, as the potential profiles with Label ``basis 0, 1'' and
-Label ``basis 1'' are more similar to the profile with Label ``exact'', the
-optimal potential. (Notice that only the region $x<0$ is meaningful and the dynamics remains the same when
-the potential is shifted vertically) 
+under the tilted dynamics
+\begin{align}
+  dz_s = \left( u_{s} - V'(z_s) \right) ds + \sqrt{2\eps } dw_s\,.
+\end{align}
+The cost functional considered here is a variant of (\ref{cost}) for
+$T\to\infty$ with running cost $G=1$ and terminal cost $H=0$.
+In all numerical computations, however, $T=\infty$ is replaced by a large but finite terminal time $T<\infty$, so that $\tau\wedge T\approx \tau <\infty$; the latter is to make sure that we can apply Girsanov's theorem and thus use the cross-entropy algorithm.
+
+We generated trajectories using the Euler-Maruyama scheme with time step $\mbox{dt} = 1.0 \times 10^{-3}$.
+The number of realizations used in Algorithm \ref{ce-algo} was set to $N_j = 10^{4}$ for all temperature steps $\eps_{j}=(2j+1)^{-1}$, $j=0,1,2$.
+The algorithm was initialized with $\bm{v}^{(0)} = \bm{0}$, from which $\bm{v}^{(j+1)}$ were obtained in the $j$th step with $j=0,1,2$.
+Note that applying a control force $u^{j}_{s}=c(s,z_{s};\bm{v}^{(j)})$ in the $j$-th iteration is equivalent to modifying the potential by
+\begin{equation}
+V^{I,j}(x) = V(x) - 2\sum_{i\in I} v^{(j)}_i \phi_i(x)
+\end{equation} 
+where $I\subset \N$ is the index set of the basis functions. We denote the
+final optimized potential as $V^I(x)=V^{I,3}(x)$. 
+%
+The numerical results for up to three basis functions (i.e. $I\subset\{0,1,2\}$) are presented in Figure \ref{fig-ex1-1} and 
+Figure \ref{fig-ex1-2}. Figure \ref{fig-ex1-1} shows the modified potentials using four different index sets 
+\begin{equation}
+I\in\left\{\{0\},\{1\},\{0,1\},\{0,2\}\right\} 
+\end{equation}
+that were kept fixed throughout the iteration over the different temperature levels $\eps_{j}$.
+It can be seen that the solution is relatively sensitive to basis functions that either do not capture the relevant region of state space (here: the transition region around the maximum at $x_{1}=0$) or that are supported in regions that are not sampled (here: rightmost energy well).
+
+To analyse the accuracy of the cross-entropy-algorithm in more detail, we computed the solution of the optimal control problem (\ref{ex1-cost}) 
+by solving a Feymann-Kac type elliptic boundary value problem using a highly accurate finite element discretization (see
+\cite{control_schuette} for details); this is our reference solution. We then apply the cross-entropy method with $17$ Gaussian basis functions
+with centres  $a_k = -1.5 + 0.1k$, $k=0,1,\cdots, 16$ and variance $r=0.1$ and compute the modified potential via Algorithm \ref{ce-algo}.
+From Fig.~\ref{fig-ex1-2}, we see that cross-entropy
+solutions with $17$ basis functions can approximate the reference solution quite well.
+We also observe that a similarly good approximation can be obtained with a \emph{single} well-chosen
+basis function $\phi_1$. This indicates the possibility to solve high-dimensional problems with few basis functions.
+
+With these optimized potential, we then generated $N = 10^{6}$ samples from the
+controlled dynamics and computed the value of the associated cost function (\ref{ex1-cost}). The results are presented in
+Tab.~\ref{tab-ex1-1}: We clearly observe that the best results are obtained when the basis functions capture the relevant transition region (here: 
+$\phi_1$), since in job 2 and 3, the cost value 1.31 is closer to the exact
+solution 1.25 than elsewhere.
+
 \begin{figure}[tphb]
@@
   \begin{tabular}{ll}
-    \subfigure[basis $\phi_0$]{\includegraphics[width=5.0cm]{./fig/pot_1d_0_converge_control.eps}}
+    \subfigure[basis $\{\phi_0\}$]{\includegraphics[width=5.0cm]{./fig/pot_1d_0_converge_control.eps}}
 &
-    \subfigure[basis $\phi_1$]{\includegraphics[width=5.0cm]{./fig/pot_1d_1_converge_control.eps}} \\
-    \subfigure[basis $\phi_0,\phi_1$]{\includegraphics[width=5.0cm]{./fig/pot_1d_2_converge_control.eps}}
+    \subfigure[basis $\{\phi_1\}$]{\includegraphics[width=5.0cm]{./fig/pot_1d_1_converge_control.eps}} \\
+    \subfigure[basis $\{\phi_0,\phi_1\}$]{\includegraphics[width=5.0cm]{./fig/pot_1d_2_converge_control.eps}}
     &
-    \subfigure[basis $\phi_0,\phi_2$]{\includegraphics[width=5.0cm]{./fig/pot_1d_3_converge_control.eps}}
+    \subfigure[basis $\{\phi_0,\phi_2\}$]{\includegraphics[width=5.0cm]{./fig/pot_1d_3_converge_control.eps}}
 \end{tabular}
-\caption{Optimal control problem for one-dimensional dynamics. Convergence results for the modified
-potentials $V^{(j)}, j=0,1,2,3$, with different sets of basis functions\label{fig-ex1-1}}
+\caption{Effective potentials for one-dimensional dynamics. The modified
+potentials $V^{I,j}(x)$, $j=0,1,2,3$, obtained by performing Algorithm~\ref{ce-algo}, with different sets of basis functions as explained in the text. \label{fig-ex1-1}}
 \end{figure}
+
 \begin{figure}[tphb]
@@
     \includegraphics[width=7.0cm]{./fig/pot_1d_basis_control.eps}
-    \caption{Optimal control problem for one-dimensional dynamics. Optimized potential $V^{(3)}$ with
-    different sets of basis functions. The curve with Label ``exact'' is
-  computed directly by solving a Feymann-Kac boundary value PDE. \label{fig-ex1-2}}
+    \caption{Optimal control problem. Optimized
+      potential $V^{I}(x)$ with a single
+    basis function $\phi_1$ compared to the cross-entropy optimized potential with $17$ basis functions and the (''exact'') reference  solution.
+The potentials are vertically shifted for better presentation.
+\label{fig-ex1-2}}
 \end{figure}
+
 \begin{table}
@@
 %  \begin{tabular}{c|c|c|c|c}
-  \begin{tabular*}{0.8\textwidth}{@{\extracolsep{\fill}}ccccc}
-    \hline
-    \hline
-    job & coefficients & cost & relative variance & mean of $\tau$ \\
-    \hline
-       $1$ & $(-1.771, 0, 0)$ & $5.14$ & $1.35$ & $2.02$  \\
-       $2$ & $(0, 1.857, 0)$ & $1.31$ & $0.44$ & $0.57$  \\
-       $3$ & $(-0.110, 1.762, 0)$ & $1.31$ & $0.44$ & $0.57$  \\
-       $4$ & $(-1.611, 0, 1.379)$ & $3.88$ & $1.22$ & $1.73$  \\
-       $5$ & $$ & $1.25$ & $0.30$ & $0.52$  \\
+  \begin{tabular*}{0.7\textwidth}{@{\extracolsep{\fill}}cccc}
+    \hline
+    \hline
+    basis set & coefficients & cost & mean of $\tau$ \\
+    \hline
+       $\{\phi_0\}$ & $(-1.252, 0, 0)$ & $5.14$ &  $2.02$  \\
+       $\{\phi_1\}$ & $(0, 1.313, 0)$ & $1.31$ &   $0.57$  \\
+       $\{\phi_0,\phi_1\}$ & $(-0.078, 1.246, 0)$ & $1.31$ &  $0.57$  \\
+       $\{\phi_0,\phi_2\}$ & $(-1.139, 0, 0.975)$ & $3.88$ &   $1.73$  \\
+       17 Gaussians /reference &  & $1.27$/$1.25$ & $0.52$/$0.52$  \\
     \hline
@@
   \end{tabular*}
-  \caption{Results of the optimal control problem for the one-dimensional dynamics with different sets
-  of basis functions. The coefficients are solved by Algorithm.\ref{ce-algo}.
-Job $1$-$4$ are the results with different sets of functions $\{\phi_0\}$,
-$\{\phi_1\}$, $\{\phi_0, \phi_1\}$, $\{\phi_0, \phi_2\}$ respectively. The
-costs are the exectation (\ref{ex1-cost}) computed with $N=10^6$ samples. The
-last column is the mean of the stopping time $\tau$ under the modified
-dynamics. The last row is the computed by solving a Feymann-Kac type PDE and
-is considered as the ``accurate'' solution.\label{tab-ex1-1}}
+  \caption{One-dimensional optimal control problem with different sets
+  of basis functions. The
+cost is given by the expectation (\ref{ex1-cost}) that is computed from $N=10^6$ samples. The
+last column shows the mean of the stopping time $\tau$ under the modified
+dynamics. The last row displays the results for $17$ Gaussian basis functions and the reference solution.
+\label{tab-ex1-1}}
 \end{table}
 
-We continue to study the computation of $P(\tau \le \lambda)$ for $\lambda > 0$.
-We choose $\lambda = 2.0, 1.0, 0.5, 0.3, 0.1$ and two sets of basis functions are used : $\{\phi_1\}$
-and $\{\phi_2\}$. We compare the results using importance sampling with the
-standard Monte Carlo method. These results are shown in
-Tab.~\ref{tab-ex1-2}--\ref{tab-ex1-4} and Figure.\ref{fig-ex1-3}.
-Tab.~\ref{tab-ex1-2} and Tab.~\ref{tab-ex1-3} shows the results of
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsubsection*{Importance sampling.}
+
+We continue our study with the computation of the exit time distribution
+$\bP(\tau \le T)$ for different values of $T$. Here we choose $T \in\{2.0,
+1.0, 0.5, 0.3, 0.2, 0.1\}$ with two sets of basis functions, $\{\phi_0\}$
+and $\{\phi_1\}$. We compare the outcome of the cross-entropy algorithm with that obtained by 
+standard Monte Carlo; the results are shown in 
+Tables \ref{tab-ex1-2}--\ref{tab-ex1-4} and Figure \ref{fig-ex1-3}.
+Tables \ref{tab-ex1-2} and \ref{tab-ex1-3} show the results of
 importance sampling with basis function $\{\phi_0\}$ and $\{\phi_1\}$,
-respectively, while Tab.~\ref{tab-ex1-4} is the result using standard Monte Carlo method.
-In each case, we can observe that the estimators' relative variance will increase when $\lambda$ becomes
-small. However, for fixed $\lambda$, the relative variances of importance
-sampling are largely reduced comparing to the standard Monte Carlo method. And this becomes
-more apparent when $\lambda$ decreases, while the standard Monte Carlo method
-may fail to work (see the last row of Tab.~\ref{tab-ex1-4}).
-In Figure.\ref{fig-ex1-3}, the optimized potentials used to perform importance sampling
-are ploted for different $\lambda$ and different sets of basis functions. As
-$\lambda$ decreases,
-the optimized potentials become more different from the original
-potential, which indicates larger forces are needed.
+respectively, while Table \ref{tab-ex1-4} records the result of vanilla Monte Carlo. 
+We can observe that, for each value of $T$, the variances of the importance
+samplers are largely reduced compared to those of standard Monte Carlo. The
+difference increases when $T$ decreases as is to be expected, for the event
+$\{\tau\le T\}$ rarer the smaller $T$.
+Note that for $T=0.1$ standard Monte Carlo cannot be used at all because not a single realization is  generated that hits the set boundary, while importance sampling still gives a resonable estimate; see Table~\ref{tab-ex1-4}.
+
+In order to measure the speed-up gained by importance sampling (IS) compard to standard Monte Carlo (MC) we define the accelaration index 
+\begin{equation}
+\cI=\frac{\textrm{Var}_{MC}}{\textrm{Var}_{IS}}\,
+\end{equation}
+as the ratio of MC and IS sample variances.
+According to the central limit theorem the variances of the two methods will decrease as $N^{-1}$ with the number $N$ of trajectories, and hence the accelaration index $\cI$ has the following interpretation: If $N$ is sufficiently large and IS reaches a certain error with $N$ trajectories, MC requires about $\cI N$ trajectories to achieve the same error.
+Thus, $\cI$ is the speed-up factor of IS relative to MC, provided that we can ignore the computational overhead associated with importance sampling (which is the case here).
+
+In Figure \ref{fig-ex1-3}, the optimized potentials resulting from the IS 
+are plotted for different values of $T$ and different sets of basis functions. As expected 
+the optimized potentials become increasingly different from the original 
+potential the smaller the value of $T$, indicating that larger forces are needed when the event is rarer.
 
@@
 \centering
-  \begin{tabular*}{0.9\textwidth}{@{\extracolsep{\fill}}cccccc}
-    \hline
-    \hline
-   $\lambda$ & coefficients & $P(\tau \le \lambda)$ & relative variance & Traj. Usage & time (sec.) \\
-    \hline
-  $2.0$ & $(-0.837, 0, 0)$ & $9.22 \times 10^{-2}$ & $1.45$ & $6.1 \times 10^{-1}$ & $70.4$ \\
-    $1.0$ & $(-1.391, 0, 0)$ & $3.23 \times 10^{-2}$ & $1.48$ & $4.8 \times 10^{-1}$ & $39.4$ \\
-    $0.5$ & $(-2.221, 0, 0)$ & $6.45 \times 10^{-3}$ & $1.75$ & $3.9 \times 10^{-1}$ & $21.5$ \\
-    $0.3$ & $(-3.282, 0, 0)$ & $9.51 \times 10^{-4}$ & $2.30$ & $3.3 \times 10^{-1}$ & $13.6$ \\
-    $0.2$ & $(-4.553, 0, 0)$ & $9.53 \times 10^{-5}$ & $3.10$ & $2.7 \times 10^{-1}$ & $9.4$ \\
-    $0.1$ & $(-8.245, 0, 0)$ & $1.22 \times 10^{-7}$ & $6.84$ & $1.6 \times
- 10^{-1}$ & $4.9$ \\
+  \begin{tabular*}{0.9\textwidth}{@{\extracolsep{\fill}}ccccrr}
+    \hline
+    \hline
+    $T$ & coefficients & $\bP(\tau \le T)$ & Var & Accel. $\cI$ & Traj. Usage \\
+    \hline
+    $2.0$ & $(-0.592, 0, 0)$ & $9.22 \times 10^{-2}$ & $1.8 \times 10^{-2}$ & $4.7 $ & $61\%$  \\
+    $1.0$ & $(-0.984, 0, 0)$ & $3.23 \times 10^{-2}$ & $2.2 \times 10^{-3}$ & $13.7$ & $48\%$  \\
+    $0.5$ & $(-1.570, 0, 0)$ & $6.45 \times 10^{-3}$ & $1.3 \times 10^{-4}$ & $50.3 $ & $39\%$  \\
+    $0.3$ & $(-2.321, 0, 0)$ & $9.51 \times 10^{-4}$ & $4.8\times 10^{-6}$ & $198.6 $ & $33\%$  \\
+    $0.2$ & $(-3.219, 0, 0)$ & $9.53 \times 10^{-5}$ & $8.7\times 10^{-8}$ & $1091.8 $ & $27\%$  \\
+    $0.1$ & $(-5.830, 0, 0)$ & $1.22 \times 10^{-7}$ & $7.0\times 10^{-13}$ & very large & $16\%$ \\
     \hline
@@
   \end{tabular*}
-  \caption{ $\mathcal{P}(\tau \le \lambda)$ for one-dimensional dynamics.
-  $N=10^6$ samples and importance sampling are used with basis function
-  $\{\phi_0\}$. ``Traj. Usage'' is the
-ratio of trajectories satisfying $\tau \le \lambda$ under the modified dynamics. \label{tab-ex1-2}}
+  \caption{Computation of $\bP(\tau \le T)$ by importance sampling, based on 
+  $N=10^6$ independent  realizations a single basis function
+  $\phi_0$.
+``Accel.'' $\cI$ (acceleration) measures the computational speed-up of importance sampling relative to standard MC.
+``Traj.
+Usage'' denotes the portion of trajectories satisfying $\tau \le T$ under the modified dynamics.
+\label{tab-ex1-2}}
 \end{table}
+
 \begin{table}
   \centering
-  \begin{tabular*}{0.9\textwidth}{@{\extracolsep{\fill}}cccccc}
-    \hline
-    \hline
-    $\lambda$ & coefficients & $P(\tau \le b)$ & relative variance & Traj. Usage & time (sec.) \\
-    \hline
-  $2.0$ & $(0, 0.962, 0)$ & $9.23 \times 10^{-2}$ & $1.04$ & $8.5 \times 10^{-1}$ & $56.1$\\
-  $1.0$ & $(0, 1.498, 0)$ & $3.23 \times 10^{-2}$ & $1.00$ & $8.1 \times 10^{-1}$ & $32.7$ \\
-  $0.5$ & $(0, 2.314, 0)$ & $6.46 \times 10^{-3}$ & $1.33$ & $6.8 \times 10^{-1}$ & $19.4$ \\
-  $0.3$ & $(0, 3.338, 0)$ & $9.49 \times 10^{-4}$ & $1.84$ & $5.6 \times 10^{-1}$ & $12.8$ \\
-  $0.2$ & $(0, 4.578, 0)$ & $9.56 \times 10^{-5}$ & $2.54$ & $4.6 \times 10^{-1}$ & $9.02$\\
-  $0.1$ & $(0, 8.232, 0)$ & $1.21 \times 10^{-7}$ & $5.28$ & $2.6 \times 10^{-1}$ & $4.8$\\
+  \begin{tabular*}{0.9\textwidth}{@{\extracolsep{\fill}}ccccrr}
+    \hline
+    \hline
+    $T$ & coefficients & $\bP(\tau \le T)$ & Var & Accel. $\cI$ & Traj. Usage \\
+    \hline
+    $2.0$ & $(0, 0.680, 0)$ & $9.23 \times 10^{-2}$ & $9.2 \times 10^{-3}$ & $9.1$ & $85\% $ \\
+    $1.0$ & $(0, 1.059, 0)$ & $3.23 \times 10^{-2}$ & $1.0 \times 10^{-3}$ & $30.0$ & $81\% $ \\
+    $0.5$ & $(0, 1.636, 0)$ & $6.46 \times 10^{-3}$ & $7.4 \times 10^{-5}$ & $86.9$ & $68\% $ \\
+    $0.3$ & $(0, 2.360, 0)$ & $9.49 \times 10^{-4}$ & $3.0 \times 10^{-6}$ & $310.9$ & $56\% $ \\
+    $0.2$ & $(0, 3.237, 0)$ & $9.56 \times 10^{-5}$ & $5.9\times 10^{-8}$  & $1621.2$ & $46\% $ \\
+    $0.1$ & $(0, 5.821, 0)$ & $1.21 \times 10^{-7}$ & $4.1 \times 10^{-13}$ & very large & $26\%$ \\
     \hline
@@
   \end{tabular*}
-  \caption{ $\mathcal{P}(\tau \le \lambda)$ for one-dimensional dynamics.
-  $N=10^6$ samples and importance sampling are used with basis function
-$\{\phi_1\}$. ``Traj. Usage'' is the ratio of trajectories satisfying $\tau \le \lambda$ under the modified dynamics. \label{tab-ex1-3}}
+  \caption{Computation of $\bP(\tau \le T)$ for the one-dimensional dynamics, based on
+  $N=10^6$  independent  realizations a single basis function
+$\{\phi_1\}$; see Tab.~\ref{tab-ex1-2} for comparison. \label{tab-ex1-3}}
 \end{table}
+
 \begin{table}
   \centering
-  \begin{tabular*}{0.9\textwidth}{@{\extracolsep{\fill}}ccccc}
-    \hline
-    \hline
-    $\lambda$ & $P(\tau \le \lambda)$ & relative variance & Traj. Usage & time (sec.) \\
-    \hline
-    $2.0$ & $9.23 \times 10^{-2}$ & $3.14$ & $9.23 \times 10^{-2}$ & $95.0$ \\
-    $1.0$ & $3.22 \times 10^{-2}$ & $5.48$ & $3.22 \times 10^{-2}$ & $48.8$ \\
-    $0.5$ & $6.28 \times 10^{-3}$ & $12.58$ & $6.28 \times 10^{-3}$ & $24.6$ \\
-    $0.3$ & $1.00 \times 10^{-3}$ & $31.61$ & $1.00 \times 10^{-3}$ & $14.9$ \\
-    $0.2$ & $9.30 \times 10^{-5}$ & $103.69$ & $9.30 \times 10^{-5}$ & $9.9$ \\
-    $0.1$ & $0.00$ & $-$ & $0.00$ & $5.0$ \\
+  \begin{tabular*}{0.9\textwidth}{@{\extracolsep{\fill}}cccrr}
+    \hline
+    \hline
+    $T$ & $\bP(\tau \le T)$ & Var & Accel. & Traj. Usage  \\
+    \hline
+    $2.0$ & $9.23 \times 10^{-2}$ & $8.4 \times 10^{-2}$ & $1.0$ & $9.2\%$  \\
+    $1.0$ & $3.22 \times 10^{-2}$ & $3.1 \times 10^{-2}$ & $1.0$ &$3.2\% $  \\
+    $0.5$ & $6.28 \times 10^{-3}$ & $6.2 \times 10^{-3}$ & $1.0$ &$0.6\% $ \\
+    $0.3$ & $1.00 \times 10^{-3}$ & $1.0 \times 10^{-3}$ & $1.0$ &$0.1\% $ \\
+    $0.2$ & $9.30 \times 10^{-5}$ & $9.3 \times 10^{-5}$ & $1.0$ & $0.009\% $ \\
+    $0.1$ & $0.00$ & $-$ & $1.0$ & $0.0\%$  \\
     \hline
@@
   \centering
-  \caption{$\mathcal{P}(\tau \le \lambda)$ for one-dimensional dynamics.
-    ``Traj. Usage'' is the
-ratio of trajectories satisfying $\tau \le \lambda$ . $N=10^6$ samples are
-used with standard Monte Carlo method. In the last row, the estimation fails because there is no
-  trajectories satisfying $\tau \le \lambda$ among $N=10^6$ trajectory samples. \label{fig-ex1-3}}
+  \caption{Computation of $\bP(\tau \le T)$ for the one-dimensional dynamics by standard Monte-Carlo, based on
+  $N=10^6$  independent  realizations; compare Tabs. \ref{tab-ex1-2} and \ref{tab-ex1-3}. \label{fig-ex1-3}}
 \end{table}
+
 \begin{figure}[tphb]
@@
     &
-    \subfigure[$P(\tau \le \lambda)$]{\includegraphics[width=5.0cm]{./fig/pdf_tau.eps}}
+    \subfigure[$\bP(\tau \le T)$]{\includegraphics[width=5.0cm]{./fig/pdf_tau.eps}}
 \end{tabular}
-\caption{The modified potentials for computing $\mathcal{P}(\tau \le \lambda)$ of
-one-dimensional dynamics with different sets of basis functions. The
-potentials
-become more different from the original one when $\lambda$ decreases. (c) shows the cumulative distribution function of
-$\tau$ (only for small $\lambda$). \label{tab-ex1-4}} 
+\caption{The modified potentials for $\bP(\tau \le T)$ with different sets of basis functions. The
+potentials become increasingly different from the original one when $T$
+decreases. The rightmost panel shows the cumulative distribution function of
+$\tau$ for $T \in [0.1, 2.0]$. \label{tab-ex1-4}} 
 \end{figure}
-%\FloatBarrier
-%\subsection{Non-gradient dynamical system}
-%Possibly more interesting examples are the noise-driven non-gradient
-%systems studied in \cite{mechanism_coherence, stochastic_resonance}.
-\subsection{Two dimensional example}
-We choose the potential 
-\begin{align}
-V(x) = (1 - x_1^2 - x_2^2)^2 + \frac{x_1^2}{x_1^2 + x_2^2}, \quad x = (x_1,x_2) \in \mathbb{R}^2
-\end{align}
-and consider the two-dimensional dynamics
-\begin{align}
-  dz_s = - \nabla V(z_s) ds + \sqrt{2\beta^{-1}} dw_s
-\end{align}
-where $w_s$ is a two-dimensional Brownian motion.
-The potential $V(x)$ (see Figure.\ref{fig-ex2-0}) has two minima at $a = (-1, 0)$ and $b = (1,0)$.
-$c_1 = (0, 1)$ and $c_2 = (0,-1)$ are two saddle points and $d=(0,0)$ is a
-singular point. The basis functions we choose are 
-\begin{align}
-  & \phi_0(x) = \exp(-\frac{|x - a|^2}{2r_0^2}), \quad \phi_1(x) = \exp(-\frac{|x - c_1|^2}{2r_1^2}), \\
-  & \phi_2(x) = \exp(-\frac{|x - c_2|^2}{2r_2^2}) \quad \phi_3(x) = \exp(-\frac{|x|^2}{2r_0^2}) 
-\end{align}
-with $r_0 = 0.5, r_1 = r_2 = 1.0$. We study the dynamics starting from $z_0 =
-a$ and define the stopping time as $\tau = \inf\{s > 0, |z_s - b| \le 1.0\}$.
-
-For the optimal control problem, we consider the cost function 
-\begin{align}
-  \inf_{u_s} \mathrm{E}\Big(\tau + \frac{1}{2}\int_0^\tau |u_s|^2 ds ~\Big|~ z_0 = a \Big)
-  \label{cost-ex2}
-\end{align}
-under dynamics
-\begin{align}
-  dz_s = - \nabla V(z_s) ds + \sqrt{2}u_s ds + \sqrt{2\beta^{-1}} dw_s
-\end{align}
-We choose $\beta = 8.0$, $\mbox{dt} = 0.001$. When solving the coefficients
-using Algorithm.\ref{ce-algo}, we
-choose $(\beta_0, \beta_1 , \beta_2, \beta_3) = (2.0, 4.0, 6.0, 8.0)$, and
-$N_i = 10^4$ for each $\beta_i, i = 0, 1, 2, 3$. Four different sets of
-basis functions $\{\phi_0\}$, $\{\phi_0, \phi_1\}$, $\{\phi_0, \phi_1,
-\phi_2\}$, $\{\phi_0, \phi_1, \phi_2, \phi_3\}$ are studied. After finding the coefficients, the cost function (\ref{cost-ex2}) is computed with $N =
-10^6$ samples. Results are tabulated in Tab.~\ref{tab-ex2-1}. From the dual relation
-(\ref{dual-relation}), we also compute the optimal cost by Monte Carlo method
-and the value is $1.27$. It can be observed that the cost ($1.39$) is close to the
-optimal cost ($1.27$) if all basis functions are used. Figure.\ref{fig-ex2-1} shows the optimized potential with different basis functions.
-\begin{figure}[tphb]
-  \centering
-    \includegraphics[width=7.5cm]{./fig/pot_2d_1_0_0.eps}
-    \caption{Potential $V(x)$ for the 2D example. \label{fig-ex2-0}}
-\end{figure}
-\begin{table}[tphb]
-  \begin{tabular*}{0.9\textwidth}{@{\extracolsep{\fill}}ccccc}
-    \hline
-    \hline
-    job & coefficients & cost & relative variance & mean of $\tau$ \\
-    \hline
-    $0$ & ($-1.535$, $0.000$, $0.000$, $0.000$) & $3.69$ & $0.67$ & $2.97$ \\
-    $1$ & ($-1.379$, $0.744$, $0.000$, $0.000$) & $3.86$ & $0.82$ & $3.16$ \\
-    $2$ & ($-0.943$, $1.527$, $1.500$, $0.000$) & $2.48$ & $0.94$ & $1.71$ \\
-    $3$ & ($-0.909$, $0.479$, $0.484$, $0.936$) & $1.39$ & $0.46$ & $0.63$ \\
-    \hline
-    \hline
-  \end{tabular*}
-  \centering
-  \caption{2D example, optimal control problem with different sets of basis
-  functions $\{\phi_0\}$, $\{\phi_0, \phi_1\}$, $\{\phi_0, \phi_1, \phi_2\}$, $\{\phi_0, \phi_1, \phi_2, \phi_3\}$.
-  The coefficients are solved using Algorithm.\ref{ce-algo}. The cost value is
-  computed with $N=10^{6}$ trajectories.
-The last column is the mean value of stopping time $\tau$ under the modified
-dynamics. \label{tab-ex2-1}}
-\end{table}
-\begin{figure}[tphb]
-  \centering
-  \begin{tabular}{ll}
-    \subfigure[basis $\{\phi_0\}$]{\includegraphics[width=6.0cm]{./fig/pot_2d_1_0_4.eps}} 
-    &
-    \subfigure[basis $\{\phi_0, \phi_1\}$]{\includegraphics[width=6.0cm]{./fig/pot_2d_1_1_4.eps}} \\
-    \subfigure[basis $\{\phi_0, \phi_1, \phi_2\}$]{\includegraphics[width=6.0cm]{./fig/pot_2d_1_2_4.eps}} 
-    &
-    \subfigure[basis $\{\phi_0, \phi_1, \phi_2, \phi_3\}$]{\includegraphics[width=6.0cm]{./fig/pot_2d_1_3_4.eps}}
-  \end{tabular}
-    \caption{2D example. The optimized potentials for optimal control problem with different sets of basis functions.\label{fig-ex2-1}}
-\end{figure}
-
-We continue to compute the probability of the rare events $\mathcal{P}(\tau
-\le \lambda)$. We choose a large $\beta = 20.0$, $\mbox{dt} = 0.001$. 
-For simplicity we only consider basis functions : 
-%$\{\phi_0\}$, $\{\phi_0, \phi_1, \phi_2\}$, 
-$\{\phi_0, \phi_1, \phi_2, \phi_3\}$. $(\beta_0, \beta_1 , \beta_2,
-\beta_3, \beta_4, \beta_5, \beta_6) = (4.0, 8.0, 10.0, 12.0, 16.0, 18.0,
-20.0)$ and $N_i = 10^4, 0\le i \le 5, N_6 = 10^5$ are used when solving the coefficients.
-The numerical results are shown in Tab.~\ref{tab-ex2-4}--\ref{tab-ex2-5}.
-Tab.~\ref{tab-ex2-4} shows the
-results of importance sampling and the results
-of standard Monte Carlo method are given in Tab.~\ref{tab-ex2-5}. $N=5 \times
-10^{6}$ samples are used to compute the probability. We clearly
-see that, comparing to the standard Monte Carlo method, the (relative) variances are always largely reduced with importance sampling, especially when $\lambda$ is small. 
-\begin{comment}
-Comparing different basis
-functions, we observe that the performance is better when all four basis functions are used.
-The cumulative distribution function is computed using importance sampling and is shown in Figure.\ref{fig-ex2-2}. 
-The optimized potentials for different $\lambda$ are presented in Figure.\ref{fig-ex2-3}. We see once again that large force is needed when $\lambda$ is small.
-\begin{table}
-  \begin{tabular}{c|c|c|c|c|c}
-    \hline
-    \hline
-  $\lambda$ & coefficients & $P(\tau \le \lambda)$ & relative variance & $R_e$ & time (sec.) \\
-    \hline
-    $12.0$ & ($-0.271$, $0.000$, $0.000$, $0.000$) & $9.36\times 10^{-3}$ & $7.00$ & $2.08\times 10^{-1}$ & $3219.7$ \\ 
-    $10.0$ & ($-0.305$, $0.000$, $0.000$, $0.000$) & $6.47\times 10^{-3}$ & $8.25$ & $1.87\times 10^{-1}$ & $2688.7$ \\
-    $8.0$ & ($-0.337$, $0.000$, $0.000$, $0.000$) & $3.97\times 10^{-3}$ & $8.93$ & $1.50\times 10^{-1}$ & $2184.4$ \\
-    $6.0$ & ($-0.391$, $0.000$, $0.000$, $0.000$) & $2.03\times 10^{-3}$ & $9.13$ & $1.13\times 10^{-1}$ & $1660.6$ \\
-    $4.0$ & ($-0.471$, $0.000$, $0.000$, $0.000$) & $7.50\times 10^{-4}$ & $12.38$ & $6.86\times 10^{-2}$ & $1121.4$ \\
-    $2.0$ & ($-0.593$, $0.000$, $0.000$, $0.000$) & $1.19\times 10^{-4}$ & $16.30$ & $2.33\times 10^{-2}$ & $569.6$ \\
-    $1.0$ & ($-0.945$, $0.000$, $0.000$, $0.000$) & $7.26\times 10^{-6}$ & $31.37$ & $1.86\times 10^{-2}$ & $286.2$ \\
-    $0.5$ & ($-1.654$, $0.000$, $0.000$, $0.000$) & $3.60\times 10^{-8}$ & $52.00$ & $2.03\times 10^{-2}$ & $143.3$ \\
-    \hline
-    \hline
-  \end{tabular}
-  \centering
-  \caption{2D example, $\mathcal{P}(\tau \le \lambda)$ using importance
-  sampling method. Coefficients are computed using Algorithm.\ref{ce-algo}
-with basis functions $\{\phi_0\}$. $R_e$ is the ratio of trajectories
-satisfying $\tau \le \lambda$ under the modified dynamics. \label{tab-ex2-2}}
-\end{table}
-\begin{table}
-  \begin{tabular}{c|c|c|c|c|c}
-    \hline
-    \hline
-    $\lambda$ & coefficients & $P(\tau \le \lambda)$ & relative variance & $R_e$ & time (sec.) \\
-    \hline
-  $12.0$ & ($-0.255$,  $0.030$, $0.029$, $0.000$) & $9.40\times 10^{-3}$ & $6.76$ & $1.95\times 10^{-1}$ & $4724.0$ \\
-  $10.0$ & ($-0.291$,  $0.035$, $0.040$, $0.000$) & $6.40\times 10^{-3}$ & $6.85$ & $1.81\times 10^{-1}$ & $3967.5$ \\
-  $8.0$ & ($-0.341$,  $0.049$, $0.051$, $0.000$) & $3.98\times 10^{-3}$ & $17.59$ & $1.63\times 10^{-1}$ & $3190.7$ \\
-  $6.0$ & ($-0.400$,  $0.065$, $0.069$, $0.000$) & $2.03\times 10^{-3}$ & $10.99$ & $1.30\times 10^{-1}$ & $2418.9$ \\
-  $4.0$ & ($-0.446$,  $0.127$, $0.127$, $0.000$) & $7.47\times 10^{-4}$ & $11.95$ & $8.30\times 10^{-2}$ & $1638.6$ \\
-  $2.0$ & ($-0.519$,  $0.317$, $0.318$, $0.000$) & $1.21\times 10^{-4}$ & $13.95$ & $6.78\times 10^{-2}$ & $860.5$ \\
-  $1.0$ & ($-0.719$,  $0.712$, $0.729$, $0.000$) & $7.24\times 10^{-6}$ & $20.08$ & $1.43\times 10^{-1}$ & $457.4$ \\
-  $0.5$ & ($-1.035$,  $1.681$, $1.552$, $0.000$) & $3.71\times 10^{-8}$ & $12.97$ & $2.69\times 10^{-1}$ & $210.4$ \\
-    \hline
-    \hline
-  \end{tabular}
-  \centering
-  \caption{2D example, $\mathcal{P}(\tau \le \lambda)$ using importance
-    sampling method. Basis functions are $\{\phi_0, \phi_1, \phi_2\}$. $R_e$ is the ratio of trajectories
-satisfying $\tau \le \lambda$ under the modified dynamics. \label{tab-ex2-3}}
-\end{table}
-\end{comment}
-\begin{table}[htbp]
-  \begin{tabular*}{0.9\textwidth}{@{\extracolsep{\fill}}ccccc}
-    \hline
-    \hline
-    $\lambda$ & coefficients & $P(\tau \le \lambda)$ & relative variance &
-    Traj. Usage  \\
-    \hline
-    $12.0$ & ($-0.270$, $-0.029$, $-0.034$, $0.074$) & $9.39 \times 10^{-3}$ & $6.65$ & $2.26\times 10^{-1}$  \\
-    $10.0$ & ($-0.319$, $-0.062$, $-0.064$, $0.108$) & $6.43\times 10^{-3}$ & $7.87$ & $2.35\times 10^{-1}$ \\
-    $8.0$ & ($-0.373$,  $-0.107$, $-0.098$, $0.155$) & $3.95\times 10^{-3}$ & $7.52$ & $2.35\times 10^{-1}$ \\
-    $6.0$ & ($-0.438$,  $-0.158$, $-0.148$, $0.232$) & $2.03\times 10^{-3}$ & $8.17$ & $2.48\times 10^{-1}$ \\
-    $4.0$ & ($-0.505$,  $-0.223$, $-0.209$, $0.343$) & $7.43\times 10^{-4}$ & $8.12$ & $2.76\times 10^{-1}$  \\
-    $2.0$ & ($-0.602$,  $-0.269$, $-0.254$, $0.556$) & $1.19\times 10^{-4}$ & $11.28$ & $3.85\times 10^{-1}$  \\
-    $1.0$ & ($-0.787$,  $-0.003$, $-0.003$, $0.630$) & $7.16\times 10^{-6}$ & $12.15$ & $4.18\times 10^{-1}$  \\
-    $0.5$ & ($-1.064$, $0.865$, $0.896$, $0.566$) & $3.71\times 10^{-8}$ & $9.28$ & $4.24\times 10^{-1}$ \\
-    \hline
-    \hline
-  \end{tabular*}
-  \centering
-  \caption{2D example, $\mathcal{P}(\tau \le \lambda)$ using importance
-  sampling method. Basis functions $\{\phi_0, \phi_1, \phi_2, \phi_3\}$.
-``Traj. Usage'' is the ratio of trajectories satisfying $\tau \le \lambda$ under the modified dynamics. \label{tab-ex2-4}}
-\end{table}
-\begin{table}[htbp]
-  \begin{tabular*}{0.9\textwidth}{@{\extracolsep{\fill}}cccc}
-    \hline
-    \hline
-    $\lambda$ & $P(\tau \le \lambda)$ & relative variance & Traj. Usage  \\
-    \hline
-    $12.0$  & $9.44\times 10^{-3}$ & $10.25$ & $9.44\times 10^{-3}$  \\
-    $10.0$  & $6.38\times 10^{-3}$ & $12.48$ & $6.38\times 10^{-3}$  \\
-    $8.0$  & $3.94\times 10^{-3}$ & $15.90$ & $3.94\times 10^{-3}$  \\
-  $6.0$  & $2.04\times 10^{-3}$ & $22.10$ & $2.04\times 10^{-3}$  \\
-  $4.0$  & $7.40\times 10^{-4}$ & $36.74$ & $7.40\times 10^{-4}$  \\
-  $2.0$  & $1.25\times 10^{-4}$ & $89.44$ & $1.25\times 10^{-4}$  \\
-  $1.0$  & $6.80\times 10^{-6}$ & $383.48$ & $6.80\times 10^{-6}$  \\
-  $0.5$  & $0.00$ & $-$ & $0.00$ \\
-    \hline
-    \hline
-  \end{tabular*}
-  \centering
-  \caption{2D example, $\mathcal{P}(\tau \le \lambda)$ using standard Monte
-  Carlo method. ``Traj. Usage'' is the ratio of trajectories satisfying $\tau \le \lambda$
-under the orininal dynamics. \label{tab-ex2-5}}
-\end{table}
-\begin{comment}
-\begin{figure}[tphb]
-  \centering
-    \includegraphics[width=8.0cm]{./fig/pdf_tau_2d.eps} \caption{2D example,
-    cumulative distribution function of $\tau$, computed using importance
-  sampling with basis $\{\phi_0, \phi_1, \phi_2, \phi_3\}$.\label{fig-ex2-2}}
-\end{figure}
-\begin{figure}[tphb]
-  \centering
-  \begin{tabular}{ll}
-    \subfigure[$\lambda=12.0$]{\includegraphics[width=6.0cm]{./fig/pot_2d_2_2_7.eps}} 
-    &
-    \subfigure[$\lambda=6.0$]{\includegraphics[width=6.0cm]{./fig/pot_2d_2_11_7.eps}} \\
-    \subfigure[$\lambda=2.0$]{\includegraphics[width=6.0cm]{./fig/pot_2d_2_17_7.eps}}
-    &
-    \subfigure[$\lambda=0.5$]{\includegraphics[width=6.0cm]{./fig/pot_2d_2_23_7.eps}}
-  \end{tabular}
-  \caption{2D example, optimized potentials for computing $\mathcal{P}(\tau
-  \le \lambda)$ with different $\lambda$. Basis functions $\{\phi_0,
-\phi_1, \phi_2, \phi_3\}$ are used.\label{fig-ex2-3}}
-\end{figure}
-\end{comment}
-%\FloatBarrier
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Conformational transition of butane dissolved in water}
@@
   \includegraphics[width=6cm]{./fig.md/butane.eps}
-  \caption{The butane molecule in water}
+  \caption{The butane molecule.}
   \label{fig:tmp1}
@@
 
-The butane-dissolved-in-water system is set up in a $3.0\times
+The butane dissolved in water system is set up in a $3.0\times
 3.0\times 3.0\, \textrm{nm}^3$ periodic simulation region, with one
@@
 %the partial charge $r$ in the system.
-All simulations are performed by a home-modified GROMACS
+All simulations are performed by a modified GROMACS
 4.5~\cite{pronk2013gromacs}.  The molecular dynamics of the system is
@@
 \end{align}
-where $r$ is the Cartisian coordinates of the $i$-th atom in the
-system and $p$ is the corresponding momenta. The friction
+where $r\in\R^{3}$ denotes the Cartesian coordinate of the $i$-th atom in the
+system and $p\in\R^{3}$ is the corresponding momentum. The friction
 constant is denoted by $\gamma$. Due to the
-fluctuation-dissipation theorem, the realtion $\sigma_i^2 = 2\gamma
-m_i k_B T$ holds, where $k_B$ is the Boltzmann constant, $T$ is the
-temperature and $m_i$ is the mass of the atom.  $f_i(r)$ and $u_i(r, w)$
-are the force that modelled by the GROMOS 45a3 force field and the
+fluctuation-dissipation theorem, the relation $\sigma_i^2 = 2\gamma
+m_i k_B \mathcal T$ holds, where $k_B$ is the Boltzmann constant, $\mathcal T$ is the
+temperature and $m_i$ is the mass of the atom.  $f_i(r)$
+denotes the force resulting from the GROMOS 45a3 force field and $u_i(r, w)$ the
 control force on the $i$-th atom, respectively. Being more specific,
@@
 \begin{align}\label{eqn:gromos-inter}
-  f(r) = -\nabla_{r_i} V_{\textrm{bond}}(r) - \nabla_{r_i} V_{\textrm{non-bond}}(r),
+  f_i(r) = -\nabla_{r_i} V_{\textrm{bond}}(r) - \nabla_{r_i} V_{\textrm{non-bond}}(r),
 \end{align}
@@
 \end{align}
-The $V_b(r)$ is the energy due to the covalent bonds in the system that
-are not constrained as a constant. $b_k(r)$ is the instantaneous
+The bond potential $V_b(r)$ is the energy due to the covalent bonds in the system that
+are not constrained as fixed. $b_k(r)$ is the instantaneous
 length of the $k$-th bond, which depends on the coordinates of the
@@
 the instantaneous dihedral angle, phase shift, multiplicity and the
-interaction constant, respectively. For a butane molecule, presented
-by Fig.~\ref{fig:tmp1}, the covalent bonds exist for atom pairs 1--2,
-2--3 and 3--4. The bond angle interactions exist for atom triples
-1--2--3 and 2--3--4. The dihedral angle interaction exists for atom
+interaction constant, respectively. For a butane molecule as illustrated in Fig.~\ref{fig:tmp1}, the covalent bonds exist for atom pairs 1--2,
+2--3 and 3--4, while the bond angle interactions exist for atom triples
+1--2--3 and 2--3--4, and the dihedral angle interaction for atom
 quadruple 1--2--3--4. The strength of $k_b$ ($=7.15\times
-10^6~\textrm{kJ/mol/nm}^4)$ for all bonds) and $k_\theta$
-($=530~\textrm{kJ/mol}$ for all bond angles) are relatively much large
-than the dihedral angle constant $k_\phi$ ($=5.92~\textrm{kJ/mol}$),
-therefore, the vibrations of the bond and bond angle are rigid, so can
-be viewed as fast motions in the system. For butane dihedral angle,
+10^6$~\textrm{kJ/(mol$\times$nm$^4$)} for all bonds) and $k_\theta$
+($=530~\textrm{kJ/mol}$ for all bond angles) are much larger
+than the dihedral angle constant $k_\phi$ ($=5.92~\textrm{kJ/mol}$).
+Therefore, the vibrations of the bonds and bond angles  can
+be viewed as fast motions in the system. For the butane dihedral angle,
 the multiplicity is $m=3$ and phase shift $\delta = 0$, so the
 dihedral angle is more populated around three angles: $-60^\circ$,
-$60^\circ$ and $180^\circ$. The first two are usually called
-\emph{cis} conformation, and the last one is usually called
-\emph{trans} conformation. And the transitions between the
-conformations are slow motion in the system. The bonds and bond angles
+$60^\circ$ and $180^\circ$. The first two are called
+\emph{gauche conformations}, the other one is called
+\emph{trans conformation}. And the transitions between these
+conformations are the slow motions in the system. The bonds and bond angles
 in water molecules are constrained by the SETTLE
 algorithm~\cite{miyamoto2004settle}.  The non-bonded interaction,
-i.e.~the second term of R.H.S.~of Eq.~\eqref{eqn:gromos-inter}, is
+i.e.~the second term in equation \eqref{eqn:gromos-inter}, is
 given by
@@
 where $r_{ij}$ is the relative distance between atom $i$ and $j$. The
-first term of Eq.~\eqref{gromos-nb} is the van der Waals interaction
+first term of \eqref{gromos-nb} is the van der Waals interaction
 and the second term is the Coulomb interaction, where $q_i$ denotes the
@@
 In this butane dissolved in water system, we are interested in the
-first hitting time $\tau$ of transition from cis to trans
-conformation, and want to precisely calculate the probability of rare
-event that $\tau$ being smaller than some given threshold $\lambda$,
-i.e.~$\mathcal{P}(\tau \leq\lambda)$. We firstly run an equilibrium simulation
+first hitting time $\tau$ of transitions from the \emph{gauche} to the
+\emph{trans}
+conformations, and we want to precisely calculate the probability of the rare
+event that $\tau$ is smaller than some given threshold $T$,
+i.e.~$\bP(\tau \leq T)$. We firstly run an equilibrium simulation
 (i.e.~with $u = 0$ in Eq.~\eqref{eqn:langevin-p}) at 300~K, with
-friction constant $1.0~\textrm{ps}^{-1}$.  The dihedral angle of the
-butane molecule is calculated every 10~ps. If $\phi$ is in the range
-$[40^\circ, 80^\circ]$ (corresponding to the cis conformation), then
-the conformation (including all water coordinates) is recorded, and
-used for the later study of cis-to-trans transition. In our
-simulation, we only record the first observed 5000 conformations.
+friction constant $1.0~\textrm{ps}^{-1}$.  The dihedral angle $\phi$ of the
+butane molecule is monitored for every interval of 10~ps. If $\phi$ is in the range
+$[40^\circ, 80^\circ]$ (corresponding to the \emph{gauche} conformation), then
+the system state (including all water coordinates) is recorded, and
+used for the later study of \emph{gauche}-to-\emph{trans} transition. In our
+simulation, we only record the first 5000 of these system states as an
+equilibrium sample of the \emph{gauche} conformation.
+
 Then controlled simulation (Eq.~\eqref{eqn:langevin-r} and
 \eqref{eqn:langevin-p}) are performed also at 300~K with friction
-constant $10~\textrm{ps}^{-1}$ until $\lambda$, or until the butane
-changes to the trans conformation. Here the butane is of trans
-conformation, when the dihedral angle $\phi$ is equal or larger than
+constant $10~\textrm{ps}^{-1}$ until $T$, or until the transition to the
+\emph{trans} conformation happens. Here the butane molecule is in \emph{trans}
+conformation when the dihedral angle $\phi$ is equal or larger than
 $150^\circ$. The control  $u_i$ is defined as a function of
@@
 dihedral angle $\phi$, because it fully describes the conformational
-change of the butane moelcule, moreover, bond lengths and bond angles
+change of the butane moelcule, and, moreover, bond lengths and bond angles
 are fast variables, so (as a good approximation) they decouple from
-the dynamics of the dihedral angle.  The transitional, rotational
-movement and the neighboring solvation structure may also play a role
+the dynamics of the dihedral angle.  The transitional and rotational
+movements and the neighboring solvation structure may also have an effect
 on the conformation of butane, and it is also fully justified to
-design control also based on these degrees of freedoms (DOFs), however, we
+design the control as depending on these degrees of freedoms (DOFs); however, we
 will show later that they do not contribute significantly in the
@@
 $V_{\textrm{ctrl}}(\phi(r))$ is chosen as a expansion of cosine base
-functions of $\phi$, because the system is symmetric between range $\phi \in
-[-180^\circ,0^\circ]$ and $\phi \in [0^\circ, 180^\circ]$, so the sine
-bases are eliminated. In all our simulations, the number of control
-base functions $N_c$ is chosen to be 8. Numerical results show that
-the coefficient of the 8th base $\omega_8$ is already very small comparing
-to the dominant contributing bases, that means the number of base is
-large enough that can capture the essential structure of the control.
+functions of $\phi$ because the system is symmetric between range $\phi \in
+[-180^\circ,0^\circ]$ and $\phi \in [0^\circ, 180^\circ]$, so that the corresponding sine
+functions are eliminated. In all of our simulations, the number of control
+basis functions $N_c$ is chosen to be 8. Numerical results show that
+the coefficient of the 8th base $\omega_8$ is already very small compared
+to the dominant coefficients, which means that the number of basis functions is
+large enough for  capturing the essential structure of the control.
 In all our simulations, the Langevin equation is discretized by the
 BAOAB scheme~\cite{leimkuhler2013rational} with a time step of
-$10^{-4}$~ps for the case $\lambda = 0.1$~ps, and $5\times10^{-4}$~ps
-otherwise.
+ $5\times10^{-4}$~ps.
 
@@
   \caption{The original dihedral potential and the dihedral potential
-    with control.}
+    with optimal control potential.}
   \label{fig:tmp2}
@@
 
-We considered $\lambda = 1.0$, 0.5, 0.2 and 0.1~ps and shown the
-results in Fig.~\ref{fig:tmp2}, Tab.~\ref{tab:tmp1} and~\ref{tab:tmp2}.
-From each of the 5000 equilibrium conformations, 4 trajectories with
-different randoms seeds are simulated for each $\lambda$, so we
-generate 20000 trajectories for each case.  The control $u(r, \omega)$ is
-calculated by simulations of butane in vacumm, in which all other
+We considered $T = 1.0$, 0.5, 0.2 and 0.1~ps. The results of the importance sampling computations are shown in Fig.~\ref{fig:tmp2}, Tab.~\ref{tab:tmp1} and Tab.~\ref{tab:tmp2}.
+From each of the 5,000 equilibrium system states from the \emph{gauche} conformation, 4 trajectories with
+different random seeds are simulated for each $T$, so we
+generate $M_{IS}=20,000$ trajectories for $T = 1.0$, 0.5, 0.2.
+For $T = 0.1$~ps, 12 trajectories are simulated from each
+initial system state, so 60,000 trajectories in total.
+The optimal control $u(r, \omega)$ is
+calculated by simulations of butane in vacuum, in which all other
 parameters are the same as the aforementioned descriptions, except
-that the waters are removed. We do it this way because the vacumm
+that the water molecules are removed. This is done because the vacuum
 simulation is much cheaper than the in-water simulation, and
-practically, the control calculated in the corresponding vacumm
-systems perform well enough in the in-water systems, because we find
-that no further iteration is need to refine the control. Furthermore,
-in the vacumm system, the probability $\mathcal{P}(\tau\leq\lambda) =
-2.19\times10^{-2}$, $9.32\times 10^{-3}$, $1.35\times 10^{-3}$ and
-$5.76\times 10^{-5}$ for $\lambda = 1.0$, 0.5, 0.2 and 0.1~ps,
-respectively. These values does not significantly differ from the
+practically, the control calculated in the corresponding vacuum
+systems performs well enough in the in-water system, because we find, when tested,
+ no further iteration is needed to refine the control. In the vacuum system, we find  probabilities $\bP(\tau\leq T) =
+2.16\times10^{-2}$, $8.66\times 10^{-3}$, $1.48\times 10^{-3}$ and
+$6.13\times 10^{-5}$ for $T = 1.0$, 0.5, 0.2 and 0.1~ps,
+respectively. These values do not significantly differ from those of the
 dissolved system (see the second column of Tab.~\ref{tab:tmp1}).
-Noticing the butane is symmetric with respect to transitional and
-rotational movement, the above observation therefore indicates that
-the transitional, rotational DOFs and the water structure does not
+Noticing that butane is invariant with respect to transitional and
+rotational movement, the above observations  indicate that
+the transitional, rotational DOFs and the water structure do not
 play a dominant role in the conformational change of butane, and the
 definition of control only as a function as the dihedral angle, and
-the computation of control in the vacumm system are reasonable
+the computation of control in the vacuum system are reasonable
 choices.
@@
   \caption{
-    Butane dissolved in water. The variance reduced estimate of the probability
-    $\mathcal{P} (\tau \leq \lambda)$ by applying control on the dihedral angle.
-    See the text for more details of the definition of the control.
-    Column ``Error'' denotes the standard deviation of
-    estimating the probablity $\mathcal{P} (\tau \leq \lambda)$.
-    Column ``Var'' denotes the reduced variation of the esimate.
-    ``Accel.''(acceleration) is the ratio between
-    the reduced variance and the brute force variance (see Tab.~\ref{tab:tmp2}),
-    which is the actual speed-up of the simulation.
-    ``Traj. Usage'' dentoes percentage of trajectoies that changes
-    to the trans conformation within time interval $[0,\lambda]$.
+    Results for butane dissolved in water: The probability
+    $\bP (\tau \leq T)$ is calculated by the importance sampling procedure with
+    control acting on the dihedral angle only; see the text for more details.
+    The column ``Error'' denotes the statistical uncertainty of
+    estimating the probablity $\bP (\tau \leq T)$.
+    If the  trajectories are statistically independent, the expected error is $\sqrt{\textrm{Var}/M_{IS}}$, where
+    $M_{IS}$ is the number of trajectories used. If the trajectories are not independent, the error
+    can be estimated by the block average method~\cite{frenkel2001understanding}.
+    The meaning of the other columns are the same as Tab.~\ref{tab-ex1-2}. 
+    %here, the accelaration index has to be computed as  $\cI=\textrm{Var}_{MC} M_{MC}/(\textrm{Var}_{IS}M_{IS})$ since the numbers of trajectories used in the IS and MC procedures are different.
   }
@@
     \hline\hline
-    $\lambda$ [ps] & $\mathcal{P} (\tau \leq \lambda)$ & Error & Var & Accel. & Traj. Usage \\\hline
-    0.1 & $4.34\times 10^{-5}$ & $1.16\times 10^{-5}$ & $2.31\times10^{-6}$ &18.8 & 0.7\%\\
-    0.2 & $9.03\times 10^{-4}$ & $0.77\times 10^{-4}$ & $1.19\times10^{-4}$ & 7.6 &10.0\%\\
-    0.5 & $6.96\times 10^{-3}$ & $0.34\times 10^{-3}$ & $2.21\times10^{-3}$ & 3.1 &11.9\%\\
-    1.0 & $1.68\times 10^{-2}$ & $0.06\times 10^{-2}$ & $6.30\times10^{-3}$ & 2.7 &13.4\%\\
+    $T$ [ps] & $\bP (\tau \leq T)$ & Error & Var & Accel. $\cI$ & Traj. Usage \\\hline
+    0.1 & $4.30\times 10^{-5}$ & $0.77\times 10^{-5}$ & $3.53\times10^{-6}$ & 12.2 & 0.4\%\\
+    0.2 & $1.21\times 10^{-3}$ & $0.11\times 10^{-3}$ & $2.50\times10^{-4}$ & 4.8 & 5.4\%\\
+    0.5 & $6.85\times 10^{-3}$ & $0.38\times 10^{-3}$ & $2.88\times10^{-3}$ & 2.4 & 8.3\%\\
+    1.0 & $1.74\times 10^{-2}$ & $0.08\times 10^{-2}$ & $1.21\times10^{-2}$ & 1.4 &12.3\%\\
     \hline\hline
   \end{tabular*}
+
+%%HERE
   \caption{
-    Butane dissolved in water, brute force simulations
-    without any variance reduction. The meaning of the columns are the same
-    as Tab.~\ref{tab:tmp1}.}\label{tab:tmp2}
+    Results for butane dissolved in water: Brute force / standard Monte Carlo computations of $\bP (\tau \leq T)$ 
+    without any importance sampling.}\label{tab:tmp2}
   \begin{tabular*}{0.9\textwidth}{@{\extracolsep{\fill}}lcccrr}
     \hline\hline
-    $\lambda$ [ps] & $\mathcal{P} (\tau \leq \lambda)$ & Error & Var & Accel. & Traj. Usage \\\hline
-    0.2 & $7.46\times 10^{-4}$ & $2.36\times 10^{-4}$ & $7.46\times10^{-4}$ & 1.0 & 0.1\%\\
-    0.5 & $6.94\times 10^{-3}$ & $0.72\times 10^{-3}$ & $6.89\times10^{-3}$ & 1.0 & 0.7\%\\
-    1.0 & $1.61\times 10^{-2}$ & $0.11\times 10^{-2}$ & $1.59\times10^{-2}$ & 1.0 & 1.6\%\\
+    $T$ [ps] & $\bP (\tau \leq T)$ & Error & Var & Accel. & Traj. Usage \\\hline
+    0.1 & $9.00\times 10^{-5}$ & $3.00\times 10^{-5}$ & $9.00\times10^{-5}$ & 1.0 & 0.009\%\\
+    0.2 & $1.29\times 10^{-3}$ & $0.11\times 10^{-3}$ & $1.29\times10^{-3}$ & 1.0 & 0.1\%\\
+    0.5 & $7.41\times 10^{-3}$ & $0.27\times 10^{-3}$ & $7.36\times10^{-3}$ & 1.0 & 0.7\%\\
+    1.0 & $1.78\times 10^{-2}$ & $0.04\times 10^{-2}$ & $1.75\times10^{-2}$ & 1.0 & 1.8\%\\
+    % 0.2 & $7.46\times 10^{-4}$ & $2.36\times 10^{-4}$ & $7.46\times10^{-4}$ & 1.0 & 0.1\%\\
+    % 0.5 & $6.94\times 10^{-3}$ & $0.72\times 10^{-3}$ & $6.89\times10^{-3}$ & 1.0 & 0.7\%\\
+    % 1.0 & $1.61\times 10^{-2}$ & $0.11\times 10^{-2}$ & $1.59\times10^{-2}$ & 1.0 & 1.6\%\\
     \hline\hline
@@
 
-The Fig.~\ref{fig:tmp1} plots the effective dihedral angle energy,
-which is the original dihedral energy $V_\phi(\phi)$ plus the control
+In Fig.~\ref{fig:tmp1} the effective dihedral angle energy is plotted being defined as the original dihedral energy $V_\phi(\phi)$ plus the control
 $V_{\textrm{ctrl}}(\phi)$. We only show the effective energy in the
-range $[40^\circ, 150^\circ]$, because the initial
-conformations start in range $[40^\circ, 80^\circ]$, and the
-trajectoies are stopped when they reach $\phi = 150^\circ$. For an easy
-comparison, all effective energies are switched by a constant, so that
-they are of vanishing value at $\phi = 150^\circ$. It clear that for
-smaller $\lambda$ value, the control applied is stronger. Different
-from the one-dimensional case, where the barrier is lowered, in this
-case, only the potential well around $\phi = 60^\circ$ is fill, and
-the barrier around $\phi = 120^\circ$ is almost kept the same.  The
-resultant probability $\mathcal{P} (\tau \leq \lambda)$ calculated by the
-variance reduced simulation is summarized in Tab.~\ref{tab:tmp1},
+range $[40^\circ, 150^\circ]$, because the initial states of the trajectories are located in the range $[40^\circ, 80^\circ]$, and the
+trajectories are stopped when they reach $\phi = 150^\circ$. For an easy
+comparison, all effective energies are shifted by a constant, so that
+they are of  value zero at $\phi = 150^\circ$. It clear that for
+smaller $T$ values, the control applied is stronger.  The
+resulting probabilities $\bP (\tau \leq T)$ calculated by the
+importance sampling procedure are summarized in Tab.~\ref{tab:tmp1},
 which is consistent with Tab.~\ref{tab:tmp2} that presents the
-brute force results. Since the variance is redueced, the error of the
-statistical esitmate is more precise in the variance reduced case. The
+brute force results (calculated from $M_{MC}=100,000$ trajectoies).
+The consistency at $T = 0.1$~ps is worse, because
+among the 100,000 trajectoies, only 9 of them hit the \emph{trans}
+conformation within 0.1~ps, therefore both the estimate of the probability
+$\bP (\tau \leq T)$ and its error are not reliable due to the
+insufficient sampling.
+Since the variance is reduced, the error of the
+statistical estimate is more precise in the importance sampling case. The
 reduced variance is shown in the fourth column of Tab.~\ref{tab:tmp1},
 which is much smaller than the brute force case (the fourth column of
-Tab.~\ref{tab:tmp2}). The fifth column presents the acceleration ratio
-with respect to the brute force simulation or the benefit from the
-variance reduction, which is exactly the ratio between the brute force
-variance and the reduced variance. The column of trajectory usage
-presents the percentage of the trajectories that changes to the trans
-conformation within time interval $[0, \lambda]$. It is increased by
+Tab.~\ref{tab:tmp2}). The fifth column presents the acceleration index $\cI$
+of the importance sampling procedure relative to standard Monte Carlo. It measures the speed-up gained by the
+importance sampling. The column "trajectory usage"
+presents the percentage of the trajectories that changes to the \emph{trans}
+conformation within time interval $[0, T]$. It is increased by
 the control with respect to the brute force result.
@@
 
-As a continuation of the work \cite{zlph2013, zhws13}, we study the cross-entropy
-method in the setting of diffusion process and their application in the
+As a continuation of our works \cite{zlph2013, zhws13}, we propose a cross-entropy
+algorithm for diffusion processes and study its application to
 importance sampling and optimal control.
-In those previous work, some assumptions have been made on the dynamical systems and
-on the type of quantities to estimate \cite{zhws13}. For example, 
-in order to perform the asymptotic analysis, we assumed there are explicit parameters in dynamical systems and the dynamics
-is in certain limit regime. In the case of multiple time-scale systems, we
-also assumed that explicit slow/fast variables are known. In sampling problem \cite{zhws13}, we assumed 
-the optimal change of measure is related to certain Hamilton-Jacobi-Bellman equation,
-and therefore the type of quantity to estimate is restricted. With the
-cross-entropy method, all these assumptions could be removed. General
-diffusion processes can be investigated and importance sampling for more
-meaningful quantity such as the probability $\mathcal{P}(\tau \le \lambda)$ can be considered.
-
-On the other hand, some basis functions are needed in order to apply the
-cross-entropy method. The choice of these basis functions are crucial in order
-to obtain good approximation of the target measure and therefore effcient
-importance sampling or ``good control force'' in the optimal control problem.
-The successful choice is often based on our knowledge about the given dyanmical
-system, such as its metastable states, or reactive cordinates. It is easy to
-imagine that there is no way to obtain satisfactory results when the system
-has a reactive coordinate but the basis functions are supported somewhere else.
-The optimal control problem for dynamical systems having a reactive coordinate
-will be further studied in the future work.
-
-\section*{Acknowledgements}
+For instance, in our previous work  \cite{zhws13}, we have analysed the effect of the use of suboptimal controls in multiscale systems with explicit scale separation, e.g., slow-fast systems or diffusions in the small-noise limit.
+Here the situation is different, in that no such small parameter or detailed information regarding the relevant degrees of freedom is used.
+On the other hand, in the cross-entropy method, the approximation of the target measure and, consequently, effcient importance sampling or control strategies crucially depend on a sensible choice of a function basis.
+%
+A good choice can be often based on prior knowledge about the dynamical
+system, such as metastable states or reactive cordinates; it is easy to
+imagine that there is no way to obtain satisfactory results when the basis functions used are not supported along the relevant degrees of freedom.
+The relation between the cross-entropy method and methods for multiscale dynamical systems is an interesting and yet open question that will be addressed in future work.
+
+\section*{Acknowledgement}
 The authors acknowledge financial support by the DFG Research Center MATHEON.
